### 2022-08-03 Scripts/Map_study_locations.Rmd
Scripts to visualize locations of the networks

---
title: "Untitled"
output: html_document
---

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE}
library(gtools) #Library that allows to sort files by numerical order
library(tidyverse)
library(sf)
#Read all files with their paths
files <- dir("../Data/Clean_data", pattern="*.RData", full.names=T)
files <- mixedsort(files)

#Generate list of lists
all_data = files %>%
map(readRDS)

#Read all file names
file_names <- dir("../Data/Clean_data", pattern="*.RData", full.names=F)
file_names <- mixedsort(file_names)

#Delete extension
file_names = str_replace(file_names, ".RData", "")
#Rename all elements of the list
names(all_data) <- file_names

#Select coordinates from the list of lists----
data <- list()

for (i in 1:length(file_names)) {

  int_list <- all_data[[file_names[i]]]$InteractionData
  data[[i]] <- bind_rows(map(int_list, function(x) x %>% #map here is the same as a lapply
                select(Longitude, Latitude)),  .id = 'Internal_id')
}

#Rename list and 
names(data) <- file_names
data = bind_rows(data,  .id = 'Study_id')
data_coord <- st_as_sf(data, coords = c(3:4)) %>% 
st_set_crs(4326) %>% 
st_transform(3035)

data_coord$Study_id <- factor(data_coord$Study_id, levels = file_names)

# install.packages("sf")
library(sf)
# install.packages("giscoR")
library(giscoR)

# Year
year_ref <- 2016

# Get all countries and transform to the same CRS
cntries <- gisco_get_countries(year = year_ref,
        resolution = 03) %>%
        st_transform(3035)

set.seed(1)

colors <- c("#8AD69E", "#74C5CE",  "#FDE9DE", "#45A631", "#95F7F0", "#CB53E1",
            "#FE90BB", "#B187E1", "#F8A982", "#0075DC", 
            "#993F00",  "#5EF1F2", "#005C31", "#74C5CE", "black", "lightblue",  "darkorange3","#8F7C00",
            "#8F7C00","#8F7C00", "#9DCC00", "#C20088", "gold1","#FFA8BB","#426600","94FFB5","#003380",  
            "#FF0010","#FF0010", "salmon", "seashell3","greenyellow","mediumaquamarine")
#To generate more random colors 
#randomcoloR::distinctColorPalette(k = length(file_names), altCol = F, runTsne = T)

# Plot
g <- ggplot() +
geom_sf(data = cntries, fill = "grey80", color = NA) +
xlim(c(2200000, 7150000)) +
ylim(c(1380000, 5500000)) +
theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), panel.background = element_rect(fill = "aliceblue"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
geom_sf(data = data_coord, aes(color=Study_id), size = 0.8, stroke = 0, shape = 16) +
theme(legend.position="none") + 
scale_colour_manual(name = "Study \n locations", values = colors)

library(plotly)
ggplotly(g)

```

### 2022-09-01 Scripts/Map_study_locations.Rmd
karise dataset
manicor dataset

@@
 ---
-title: "Untitled"
+title: "Study locations"
 output: html_document
@@
 data = bind_rows(data,  .id = 'Study_id')
+
+#Add manual jitter to these three points
+data = data %>% 
+mutate(Latitude = case_when(Study_id == "16_Manicor" ~ Latitude+0.25, TRUE ~ Latitude)) %>% mutate(Latitude = case_when(Study_id == "17_Manicor" ~ Latitude+0.25, TRUE ~ Latitude)) %>%
+mutate(Latitude = case_when(Study_id == "18_Manicor" ~ Latitude+0.25, TRUE ~ Latitude))
+
 data_coord <- st_as_sf(data, coords = c(3:4)) %>% 

### 2022-11-02 Scripts/Map_study_locations.Rmd
Jauke dataset

@@
 #Read all files with their paths
-files <- dir("../Data/Clean_data", pattern="*.RData", full.names=T)
+files <- dir("../Data/Clean_data", pattern="*.rds", full.names=T)
 files <- mixedsort(files)
@@
 #Read all file names
-file_names <- dir("../Data/Clean_data", pattern="*.RData", full.names=F)
+file_names <- dir("../Data/Clean_data", pattern="*.rds", full.names=F)
 file_names <- mixedsort(file_names)
@@
 #Delete extension
-file_names = str_replace(file_names, ".RData", "")
+file_names = str_replace(file_names, ".rds", "")
 #Rename all elements of the list

### 2022-11-06 Scripts/Processing/10_Ockinger.R
file organization and renaming order

@@ -1,143 +1,0 @@
-#DATASET NUMBER 12; Ockinger
-
-#Load libraries
-library(tidyverse)
-
-
-#Prepare interaction data ----
-#Load interaction data
-data = read_csv2("Data/Raw_data/10_Ockinger/interaction_data_ockinger.csv")
-
-#Prepare interaction data with the same number of columns
-data = data %>% 
-remove_rownames() %>%
-column_to_rownames(var = '...1') %>% 
-rename(Plant_species = flw_visited_by_pollin) %>% 
-rename(Pollinator_species = pollin_species) %>% 
-drop_na(Plant_species) %>% 
-rename(Interaction = Visit) %>% 
-mutate(Sampling_method = "Transect") %>% 
-mutate(Year = "2021") %>% 
-mutate(Month = NA) %>% 
-mutate(Day = NA) %>% 
-mutate(Sampling_effort_minutes = NA) %>% 
-mutate(Sampling_area_square_meters = 200) %>% 
-mutate(Country = "Sweeden") %>% 
-mutate(Locality = NA) %>% 
-mutate(Coordinate_precision = NA) %>% 
-mutate(Elevation = NA) %>% 
-mutate(Comments = NA) %>% 
-mutate(Temperature = NA) %>% 
-mutate(Humidity = NA) 
-
-#Coordinates are in another dataset, load it
-coord = read_csv("Data/Raw_data/10_Ockinger/coordinate_data_ockinger.csv") %>% 
-rename(Latitude = Lat_WGS84) %>% 
-rename(Longitude = Long_WGS84) %>% 
-select(Name, site_ID, Latitude, Longitude)
-#glimpse(coord)
-
-data = left_join(data, coord, by = c("PlatsID" = "site_ID")) %>% 
-rename(Site_id = Site_name) 
-
-#Habitat type can be added from this other dataset
-habitat = read_csv2("Data/Raw_data/10_Ockinger/site_landscape_data.csv") %>% 
-select(Category, PlatsID)
-
-#Bind datasets  
-data = left_join(data, habitat, by = c("PlatsID" = "PlatsID")) %>% 
-rename(Habitat = Category)  %>% 
-select(Plant_species, Pollinator_species, Interaction, Sampling_method,
-         Sampling_effort_minutes, Sampling_area_square_meters,
-         Site_id, Habitat, Country, Locality, Latitude, Longitude,
-         Coordinate_precision, Elevation, Day, Month, Year, Comments,
[…]

### 2022-11-06 Scripts/Processing/10_Vanbergen.R

@@ -1,0 +1,83 @@
+#DATASET NUMBER 10; Vanbergen
+
+#Load libraries
+library(tidyverse)
+
+#Prepare interaction data ----
+data = read.csv("Data/Raw_data/10_Vanbergen/interaction_data_vanbergen.csv")
+#Create cols of Temperature and humidity
+data = data %>% 
+mutate(Temperature = NA) %>% 
+mutate(Humidity = NA) %>% 
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters))
+
+#Split by Site_id
+InteractionData <- split(data, data$Site_id)
+
+
+#Prepare flower count data ----
+flower_count = read.csv("Data/Raw_data/8_Vanbergen/flower_count_vanbergen.csv")
+
+#Filter out one column but saved the data
+flower_count = flower_count %>% 
+mutate(Units = paste0(Units, "; ", Inflorescence.type)) %>% 
+select(!Inflorescence.type)
+#Split data into different dataframes based on survey name
+FlowerCount <- split(flower_count, flower_count$Site_id)
+
+#Prepare metadata data ----
+
+#Store unique cases of plants and polls
+plant_single_cases = data %>% distinct(Plant_species)
+pollinator_single_cases = data %>%distinct(Pollinator_species)
+
+Metadata <- tibble(
+Doi = NA,
+Dataset_description = "This dataset is part of the VOODOO-project.
+It consists of plant-pollinator networks in 12 different landscapes
+(4 agricultural, 4 rural, 4 urban), sampled over 3 periods in 2021
+(mid april, end may and mid july). Coordinates are the centroids of each landscape.
+Transects were stratified in relation with the area of each type of pollinator habitat
+within a 500 m radius around the centroid. Habitat is given for each individual catch,
+not for the landscape as a whole. Transect length and duration are therefore pooled in
+this file, rather than giving the sampling time in each habitat
+(in total: 120 min per site visit, so 360 min in total, if the networks are pooled
+over the 3 sampling periods). If requested, we can also provide sampling time from
+each subtransect in different habitat types.",
+Taxa_recorded = "All flying insects, except really small stuff (< 1 mm)",
+Sampling_year = 2021,
+Country = "France",
+Habitat = "4 agricultural, 4 rural, 4 urban",
[…]

### 2022-11-06 Scripts/Processing/11_Clough.R

@@ -1,0 +1,107 @@
+#DATASET NUMBER 11; Clough
+
+#Load libraries
+library(tidyverse)
+
+#Prepare interaction data ----
+data = read_csv("Data/Raw_data/11_Clough/interaction_data_clough.csv") %>% 
+select(Plant_species, Pollinator_species, Interaction, Sampling_method,
+       Sampling_effort_minutes, Sampling_area_square_meters,
+       Site_id, Habitat, Country, Locality, Latitude, Longitude,
+       Coordinate_precision, Elevation, Day, Month, Year, Comments,
+       Temperature, Humidity) %>% 
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
+
+#Convert coordinates to lat/lon
+library(rgdal)
+coordinates = data %>% select(Longitude, Latitude)
+coordinates <- SpatialPoints (coordinates, proj4string = CRS ('+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs'))
+cord.WGS84 <- spTransform(coordinates, CRS('+proj=longlat +datum=WGS84 +no_defs'))
+#Add new cols to the tibble
+data = data %>% 
+mutate(Latitude = cord.WGS84$Latitude) %>% 
+mutate(Longitude = cord.WGS84$Longitude)
+#Split by Site_id
+InteractionData <- split(data, data$Site_id)
+
+
+#Prepare flower count data ----
+flower_count = read_csv("Data/Raw_data/9_Clough/flower_count_clough.csv")
+#Split data into different dataframes based on survey name
+FlowerCount <- split(flower_count, flower_count$Site_id)
+
+
+#Prepare metadata data ----
+
+#Store unique cases of plants and polls
+plant_single_cases = data %>% distinct(Plant_species)
+pollinator_single_cases = data %>%distinct(Pollinator_species)
+
+Metadata <- tibble(
+Doi = NA,
+Dataset_description = "This dataset documents 18 different sites in southernmost Sweden
+(Scania) along a gradient of landscape scale land-use composition.
+The focal habitat are Fennoscandian lowland species-rich dry to mesic grasslands.
+In each site a pair of 3x3 m study plots were established in temporary (April to August)
+5x9m cattle exclosures (electric fence). One plot was subjected to pollinator reduction
+treatment and is not included here. The control plot is completely open and it is the data
+from those plots that are inlcuded here. Three to four visits to each site were performed,
[…]

### 2022-11-06 Scripts/Processing/11_Karise.R

@@ -1,91 +1,0 @@
-#DATASET NUMBER 13; Karise
-#Just bumblebees
-
-#Load libraries
-library(tidyverse)
-
-#Prepare interaction data ----
-#Load interaction data
-data = read_csv("Data/Raw_data/11_Karise/interaction_data_karise.csv")
-
-#Drop some rows with na's in plants and polls
-data = data %>% 
-drop_na(Plant_species) %>% 
-drop_na(Pollinator_species)
-
-#lat and long cols are swapped, fix
-#Create new dataframe with coordinates 
-coord = data %>% 
-mutate(Latitude1 = Longitude) %>% 
-mutate(Longitude1 = Latitude) %>% 
-select(Latitude1, Longitude1) %>% 
-rename(Latitude = Latitude1) %>% 
-rename(Longitude = Longitude1)
-#Mutate cols now with correct coordinates
-data =  data %>% 
-mutate(Latitude = coord$Latitude) %>% 
-mutate(Longitude = coord$Longitude) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
-
-
-#Split by site, just for createing the listed name in this case
-InteractionData <- split(data, data$Site_id)
-
-#Prepare flower count data ----
-FlowerCount = tibble(Day = NA, Month = NA, Year = NA, Site_id = NA, Plant_species = NA,
-                     Flower_count = NA, Units = NA, Comment = NA)
-
-
-#Prepare metadata data ----
-
-#Select unique cases of plants and poll
-plant_single_cases = data %>% distinct(Plant_species)
-pollinator_single_cases = data %>%distinct(Pollinator_species)
-
-Metadata <- tibble(
-Doi = NA,
-Dataset_description = "66 different sites, from these 33 in Southern Estonia and 33 in
-Northern-Central Estonia. Bumble bees on field edges (500 m x 2 m walking transect.
-The entire transect on field edges or 400 m on field edges and 100 m inside entomophilous
-crop if this excisted in the area). Monitoring is carried out annually since 2006,
-but forage plant species was named only in years indicated. Also current year,
-the dataset will be created.",
-Taxa_recorded = "Just bumblebees",
-Sampling_year = "2014 to 2021",
-Country = "Estonia.",
-Habitat = "Field edges on agricultutal land",
-Sampling_sites = 66,
[…]

### 2022-11-06 Scripts/Processing/19_Hopfenmuller.R

@@ -1,95 +1,0 @@
-#DATASET NUMBER 21; Bernhard Hopfenmuller (script 19)
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(tidyverse)
-library(lubridate)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/19_Hopfenmuller/Interaction_data.csv")
-
-#Compare vars
-compare_variables(check_interaction_data, data)
-
-#Data
-data = data %>% 
-mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
-mutate(Year = year(ymd(Date))) %>% 
-mutate(Month = month(ymd(Date))) %>% 
-mutate(Day = day(ymd(Date))) 
-
-#Add interaction col
-data = data %>% 
-mutate(Interaction = 1) %>% 
-mutate(Sampling_method = "Variable transect walks") %>% 
-mutate(Sampling_effort_minutes = NA) %>% 
-mutate(Sampling_area_square_meters = NA) %>% 
-mutate(Habitat = "Calcareous grassland") %>% 
-mutate(Locality = Site_id) 
-
-#Add missing vars
-data = add_missing_variables(check_interaction_data, data) 
-
-#Reorder variables
-data = drop_variables(check_interaction_data, data) 
-
-#Drop this last two columns that are going to be at the metadata
-data = data %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
-
-#Split interaction data into dataframes within a list
-InteractionData <- split(data,data$Site_id)
-
-
-#Prepare flower count data ----
-flower_count <- read_csv("Data/Raw_data/19_Hopfenmuller/Flower_count.csv")
-
-#Compare vars
-compare_variables(check_flower_count_data, flower_count)
-
-#Fix names and variable
-flower_count = flower_count %>% 
-rename(Flower_count = Flower_units) %>% 
-mutate(Units = "Flower number")
-
-#Add variables
-flower_count = add_missing_variables(check_flower_count_data, flower_count) 
-
-#Order data as template
-flower_count = drop_variables(check_flower_count_data, flower_count) 
-
-#Split interaction data into dataframes within a list
-FlowerCount <- split(flower_count, flower_count$Site_id)
-
-#Prepare metadata data ----
-#Select unique cases of plants and poll
[…]

### 2022-11-06 Scripts/Processing/20_Hoiss.R

@@ -1,0 +1,114 @@
+#DATASET NUMBER 20; Bernhard Hoiss 
+source("Scripts/Empty_templates.R") #Read empty templates to compare with
+
+#Load libraries
+library(tidyverse)
+library(lubridate)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/20_Hoiss/Interaction_data.csv")
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+
+#Covert day into three separate cols
+data = data %>% 
+mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
+mutate(Year = year(ymd(Date))) %>% 
+mutate(Month = month(ymd(Date))) %>% 
+mutate(Day = day(ymd(Date))) 
+
+#Add interaction col
+data = data %>% 
+mutate(Interaction = 1) %>% 
+mutate(Sampling_method = "Transect") %>% 
+mutate(Habitat = "Grassland") %>% 
+mutate(Locality = NA) %>% 
+mutate(Coordinate_precision = NA) %>% 
+mutate(Comments = NA) %>% 
+mutate(Temperature = NA) %>% 
+mutate(Humidity = NA) %>% 
+mutate(Sampling_effort_minutes = NA) %>% 
+mutate(Sampling_area_square_meters = NA)
+  
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+#Drop this last two columns that are going to be at the metadata
+data = data %>% 
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
+
+#Split interaction data into dataframes within a list
+InteractionData <- split(data,data$Site_id)
+
+#Prepare flower count data ----
+flower_count <- read_csv("Data/Raw_data/18_Hoiss/Flower_count.csv")
+
+#Compare vars
+#compare_variables(check_flower_count_data, flower_count)
+
+#Covert day into three separate cols
+flower_count = flower_count %>% 
+mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
+mutate(Year = year(ymd(Date))) %>% 
+mutate(Month = month(ymd(Date))) %>% 
+mutate(Day = day(ymd(Date))) %>% 
+rename(Flower_count = Flower_percent_cover) %>% 
+mutate(Units = "Percent cover") %>% 
+mutate(Comment = NA)
+
+#Order data as template
+flower_count = drop_variables(check_flower_count_data, flower_count) 
+
[…]

### 2022-11-06 Scripts/Processing/4_Marini.R

@@ -1,63 +1,54 @@
 #DATASET NUMBER 4; Cappellari & Marini
-#DATASET NUMBER 5; Cappellari & Marini
-#DATASET NUMBER 6; Cappellari, Gazzea & Marini
 
 #Load libraries
 library(tidyverse)
 
 #Prepare interaction data ----
-data = read.csv("Data/Raw_data/4_Marini/interaction_data_marini.csv")
+data = read.csv("Data/Raw_data/4_5_6_Marini/interaction_data_marini.csv")
 
 #Delete all underscores and space in one column
 data = data %>% 
-mutate(across(everything(), function(x) str_replace_all(x,"_", " "))) %>% 
-mutate(Coordinate_precision = str_replace(Coordinate_precision, " ", "")) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
-mutate(Latitude = as.numeric(Latitude)) %>% 
-mutate(Longitude = as.numeric(Longitude))
+  mutate(across(everything(), function(x) str_replace_all(x,"_", " "))) %>% 
+  mutate(Coordinate_precision = str_replace(Coordinate_precision, " ", "")) %>% 
+  select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
+  mutate(Latitude = as.numeric(Latitude)) %>% 
+  mutate(Longitude = as.numeric(Longitude))
 
 #Split data into different dataframes based on survey name
 split_intdata <- split(data, data$Survey)
 #Convert to tibbles
 data1 <- as_tibble(split_intdata[[1]])
-data2 <- as_tibble(split_intdata[[2]])
-data3 <- as_tibble(split_intdata[[3]])
+
 #Now create a list of the different networks (unique Site_id) for each survey
 InteractionData1 <- split(data1, data1$Site_id)
-InteractionData2 <- split(data2, data2$Site_id)
-InteractionData3 <- split(data3, data3$Site_id)
 
 #Prepare flower count data ----
 flower_count = read.csv("Data/Raw_data/4_Marini/flower_count_marini.csv")
 
 #Delete all underscores
 FlowerCount = flower_count %>% 
-mutate(across(everything(), function(x) str_replace_all(x, "_", " "))) %>% 
-select(!c(Total_flower_cover, Collected_insects))
+  mutate(across(everything(), function(x) str_replace_all(x, "_", " "))) %>% 
+  select(!c(Total_flower_cover, Collected_insects))
 
 #Split data into different dataframes based on survey name
[…]

### 2022-11-06 Scripts/Processing/5_Marini.R

@@ -1,0 +1,103 @@
+#DATASET NUMBER 5; Cappellari & Marini
+
+#Load libraries
+library(tidyverse)
+
+#Prepare interaction data ----
+data = read.csv("Data/Raw_data/4_5_6_Marini/interaction_data_marini.csv")
+
+#Delete all underscores and space in one column
+data = data %>% 
+mutate(across(everything(), function(x) str_replace_all(x,"_", " "))) %>% 
+mutate(Coordinate_precision = str_replace(Coordinate_precision, " ", "")) %>% 
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
+mutate(Latitude = as.numeric(Latitude)) %>% 
+mutate(Longitude = as.numeric(Longitude))
+
+#Split data into different dataframes based on survey name
+split_intdata <- split(data, data$Survey)
+#Convert to tibbles
+data2 <- as_tibble(split_intdata[[2]])
+#Now create a list of the different networks (unique Site_id) for each survey
+InteractionData2 <- split(data2, data2$Site_id)
+
+#Prepare flower count data ----
+flower_count = read.csv("Data/Raw_data/4_Marini/flower_count_marini.csv")
+
+#Delete all underscores
+FlowerCount = flower_count %>% 
+mutate(across(everything(), function(x) str_replace_all(x, "_", " "))) %>% 
+select(!c(Total_flower_cover, Collected_insects))
+
+#Split data into different dataframes based on survey name
+split_flwdata <- split(FlowerCount, FlowerCount$Survey)
+#Convert to tibbles
+FlowerCount2 <- as_tibble(split_flwdata[[2]]) %>%  select(!Survey)
+#Split by Site_id
+FlowerCount2 = split(FlowerCount2, FlowerCount2$Site_id)
+
+#Prepare metadata data ----
+
+#Plants sampled in each site
+for (i in InteractionData2) {
+#Generate sum of distinct plants per site
+plant_sum2 <- bind_rows(lapply(InteractionData2, function(x) x %>% 
+                      select(Plant_species, Site_id) %>% 
+                      group_by(Site_id) %>% 
+                      summarise(Sum = n_distinct(Plant_species))))
+
+}
+#Plant sum should be the total number of plants sampled
+plant_sum2 = sum(plant_sum2$Sum)
+
+
+#Store unique cases of plants and polls
+plant_single_cases2 = data2 %>% distinct(Plant_species)
[…]

### 2022-11-08 Scripts/Map_study_locations.Rmd
file organization and renaming order
shiny app metadata

@@
 title: "Study locations"
-output: html_document
+output:
+  html_document: default
+  pdf_document: default
 ---
@@
             "#993F00",  "#5EF1F2", "#005C31", "#74C5CE", "black", "lightblue",  "darkorange3","#8F7C00",
-            "#8F7C00","#8F7C00", "#9DCC00", "#C20088", "gold1","#FFA8BB","#426600","94FFB5","#003380",  
+            "#8F7C00","#8F7C00", "#9DCC00", "#C20088", "gold1","#FFA8BB","#426600","#94FFB5","#003380",  
             "#FF0010","#FF0010", "salmon", "seashell3","greenyellow","mediumaquamarine")

### 2022-11-08 Scripts/Processing/26.R

@@ -1,39 +1,0 @@
-#DATASET NUMBER 26; Albrecht (year 2014)
-
-#Read templates to compare with
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(tidyverse)
-library(stringr)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/26_27_Albrecht/Interaction_data1.csv")
-
-#FIRST fix pollinator species (substitute dot by space)
-data = data %>% 
-mutate(Pollinator_species = str_replace_all(Pollinator_species, "\\.", " "))
-#Second fix plant species (This seems more tricky) 
-#I have to read the flower count data and see if they match after some preparation
-Plant_species <- read_csv("Data/Raw_data/26_27_Albrecht/Flower_count1.csv") %>% select(Plant_species)
-#Now create a column with the 3 first letters of genus and specific epithet separated by a point :)
-Plant_species = Plant_species %>%
-mutate(Complicated_name = str_replace_all(Plant_species, "\\.", " ")) %>% 
-mutate(Complicated_name1 = word(Complicated_name, 1)) %>% 
-mutate(Complicated_name2 = word(Complicated_name, 2)) %>% 
-mutate(Complicated_name1 = substr(Complicated_name1 , start = 1 , stop = 3 )) %>% 
-mutate(Complicated_name2 = substr(Complicated_name2 , start = 1 , stop = 3 )) %>% 
-mutate(Complicated_name = paste0(Complicated_name1, "." ,Complicated_name2)) %>% 
-distinct(Plant_species, Complicated_name) %>% #Now data is ready
-mutate(Complicated_name = trimws(Complicated_name))
-
-#Make first letter to capital in the original interaction data
-data = data %>% 
-mutate(Plant_species = str_to_title(Plant_species)) %>% 
-rename(Complicated_name = Plant_species) %>%#make sure there are not leading or trailing spaces
-mutate(Complicated_name = trimws(Complicated_name))
-
-#Now make a left join of the curated species name with the original dataset
-data = left_join(data, Plant_species)
-
-setdiff(data$Complicated_name, Plant_species$Complicated_name)

### 2022-12-15 Scripts/trial.Rmd
Datasets

---
title: "TidyTuesdayCoffeeRatings"
author: "Andrew Couch"
date: "7/7/2020"
output: html_document
---

Video: https://www.youtube.com/watch?v=XoCgToAuzsU&list=PLJfshcspBCYeJeO8YFT5e5HxuYOb5a_1W&index=35&t=1187s

```{r}
library(tidyverse)
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
coffee_ratings %>% glimpse()
```

```{r}
coffee_ratings %>% ggplot(aes(total_cup_points)) + geom_histogram() + scale_x_log10()
```

```{r}
#Owners can own multiple companies and companies can own multiple farms 
coffee_ratings %>%
  select(owner, company, farm_name) %>% 
  count(company, farm_name, sort = TRUE) %>% 
  drop_na()
```
```{r}
#Farms can have multiple owners 
coffee_ratings %>% count(farm_name, owner) %>% 
  drop_na() %>% 
  arrange(farm_name) %>% 
  count(farm_name, sort = TRUE)
```

```{r}
#Verify that we selected the correct features for point analysis
coffee_ratings %>% 
  select(total_cup_points, country_of_origin, aroma:cupper_points) %>% 
  mutate(row_num = row_number()) %>% 
  gather(key = "key", value = "value", -total_cup_points, -country_of_origin, -row_num) %>% 
  group_by(row_num) %>% 
  summarise(value = sum(value),
            total_cup_points = mean(total_cup_points))
```

```{r}
coffee_ratings %>% 
  filter(country_of_origin == "Ethiopia") %>% 
  select(aroma:cupper_points) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarise(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(key = str_replace(key, "_", " ") %>% str_to_title()) %>% 
  mutate(key = fct_reorder(key, value)) %>% 
  ggplot(aes(x = key, y = value, color = key)) + 
  geom_point(size = 5) + 
  geom_segment(aes(x = key, xend = key, y = value, yend = 0)) + 
  theme(legend.position = "none") + 
  ylab("") + 
  xlab("") + 
  coord_flip() + 
  labs(title = "Avg Flavour Profile")
```

```{r}
coffee_ratings %>% 
  filter(country_of_origin == "Ethiopia") %>% 
  select(variety) %>% 
  drop_na() %>% 
  count(variety) %>% 
  mutate(variety = fct_reorder(variety, n)) %>% 
  ggplot(aes(x = n, y = variety, fill = variety)) + 
  geom_col() + 
  ylab("") + 
  labs(title = "Bean Variety") + 
  theme(legend.position = "none")
```

```{r}
library(kableExtra)
coffee_ratings %>% 
  filter(country_of_origin == "Ethiopia") %>% 
  select(total_cup_points, species, country_of_origin, region) %>% 
  group_by(species, region) %>% 
  top_n(total_cup_points, n = 1) %>% 
  ungroup() %>% 
  mutate(region = str_trunc(region, 12, "right")) %>% 
  arrange(desc(total_cup_points)) %>% 
  kable()
```

```{r}
library(broom)
coffee_ratings %>% 
  select(country_of_origin, aroma:cupper_points) %>% 
  mutate(highlight = if_else(country_of_origin == "Ethiopia", "Highlight", "No-Highlight")) %>% 
  select(-country_of_origin) %>% 
  gather(key = "key", value = "value", -highlight) %>% 
  group_by(key) %>% 
  do(t_test = t.test(value~highlight, data = .) %>% tidy()) %>% 
  unnest(t_test) %>% 
  mutate(diffference = case_when(
    conf.low < 0 & conf.high < 0 ~ "Different",
    conf.low > 0 & conf.high > 0 ~ "Different",
    TRUE ~ "Not-Different"
  )) %>% 
  mutate(key = str_replace(key, "_", " ") %>% str_to_title()) %>% 
  mutate(key = fct_reorder(key, -estimate)) %>% 
  ggplot(aes(x = key, y = estimate, color = diffference)) + 
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  coord_flip() + 
  theme(legend.position = "none") + 
  xlab("") + 
  labs(title = "How different are the flavour profiles?")
```

### 2022-12-17 Scripts/Map_study_locations.Rmd
Adding STEP datasets

@@
 colors <- c("#8AD69E", "#74C5CE",  "#FDE9DE",  "#45A631", "#95F7F0", "#CB53E1",
-            "#FE90BB", "#B187E1", "#F8A982", "#0075DC", 
-            "#993F00",  "#5EF1F2", "#005C31", "#74C5CE", "black", "lightblue",  "darkorange3","#8F7C00",
-            "#8F7C00","#8F7C00", "#9DCC00", "#C20088", "gold1","#FFA8BB","#426600","#94FFB5","#003380",  
-            "#FF0010","#FF0010", "salmon", "seashell3","greenyellow","mediumaquamarine")
+"#FE90BB", "#B187E1", "#F8A982", 
+"#0075DC", "#993F00",  "#5EF1F2", 
+"#005C31", "#74C5CE", "black", 
+"lightblue",  "darkorange3","#8F7C00",
+"#8F7C00","#8F7C00", "#9DCC00", 
+"#C20088", "gold1","#FFA8BB",
+"#426600","#94FFB5","#003380",  
+"salmon", "palegreen","greenyellow",
+"mediumaquamarine")
 #To generate more random colors 

### 2022-12-17 Scripts/Processing/31_Roberts.R

@@ -1,0 +1,99 @@
+#DATASET NUMBER 31; STEP PROJECT: UK dataset
+#This metadataset was sent by A. Magrach
+#The dataset of Germany was already sent by other colleagues
+#So I won't add this country
+
+#Read templates to compare with
+source("Scripts/Empty_templates.R") #Read empty templates to compare with
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/29_30_31_STEP/Interaction_data_STEP.csv")
+
+#Check col names with template
+compare_variables(check_interaction_data, data)
+
+#Check levels per country
+levels(factor(data$Country))
+
+#Select 1st level that we don´t have: SPAIN
+data = data %>% filter(Country == "UK")
+
+#Select main cols
+data = data %>% 
+  select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
+#Check number of sites
+levels(factor(data$Site_id))
+
+#Split interaction data into dataframes within a list (separate now by site)
+InteractionData <- split(data, data$Site_id)
+
+#Prepare flower count data ----
+flower_count <- read_csv("Data/Raw_data/29_30_31_STEP/Flower_count_STEP.csv")
+
+#Compare vars
+compare_variables(check_flower_count_data, flower_count)
+#Select just Spain
+flower_count = flower_count %>% filter(country == "UK")
+#Order data as template and drop variables
+flower_count = drop_variables(check_flower_count_data, flower_count) 
+
+#Split interaction data into dataframes within a list
+FlowerCount <- split(flower_count, flower_count$Site_id)
+
+#Prepare metadata data ----
+#Store unique cases of plants and polls
+plant_single_cases = data %>% distinct(Plant_species)
+pollinator_single_cases = data %>%distinct(Pollinator_species)
+
+#Create metadata ordered
+Metadata <- tibble(
+  Doi = "https://doi.org/10.1111/ele.12657",
+  Dataset_description = "This is the Swedish dataset from the project
+Status and Trends of European Pollinators (STEP)",
+  Taxa_recorded = "Bees and syrphids",
+  Sampling_year = "2012 and 2013",
+  Country = "UK",
[…]

### 2022-12-17 Scripts/Processing/STEP.R

@@ -1,23 +1,0 @@
-#DATASET NUMBER; STEP PROJECT
-#This metadataset was sent by A. Magrach
-#The dataset of Germany was already sent by other colleagues
-#So I won't add it here
-
-#Read templates to compare with
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(dplyr)
-library(stringr)
-library(readr)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/29_STEP/Interaction_data_STEP.csv")
-
-#Check col names with template
-compare_variables(check_interaction_data, data)
-
-#Check levels per country
-levels(factor(data$Country))
-
-

### 2022-12-19 Scripts/Map_study_locations.Rmd
edits

@@
 "salmon", "palegreen","greenyellow",
-"mediumaquamarine")
+"mediumaquamarine", "tomato", "orange")
 #To generate more random colors 

### 2022-12-19 Scripts/Processing/34_Stanley.R

@@ -1,0 +1,162 @@
+#DATASET NUMBER 34; Stanley
+#ALL datasets from RUSSO (There are 6 in total)
+#One is an unpublished thesis
+
+#Read empty templates to compare with
+source("Scripts/Empty_templates.R") 
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
+
+#Divide col into cols
+levels(factor(data$Site_id))
+data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
+levels(factor(data$Name))
+
+#Recode authors to just their surname
+data = data %>% 
+mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
+                            "EileenPower" = "Power",
+                            "DaraStanley" = "Stanley",
+                            "SarahMullen" = "Mullen",
+                            "MichelleLarkin" = "Larkin",
+                            "CianWhite" = "White"))
+
+#Select first the one of Power
+data = data %>% filter(Name == "Stanley")
+
+#To calculate sampling effort later on
+data_time = data %>% select(Sampling_effort_minutes)
+
+#Compare vars
+compare_variables(check_interaction_data, data)
+
+#Prepare data
+data = data %>% 
+rename(Plant_species = Plant_Species) %>% 
+mutate(Habitat = "Farming grassland") %>% 
+mutate(Locality = "Farms in south-east Ireland") %>% 
+rename(Latitude = "WGS84/ETRS89") %>% 
+rename(Longitude = "WGS84/ETRS89_2") %>% 
+mutate(Site_id = Site) %>% 
+mutate(Sampling_method = "Transects") 
+
+#Add missing vars
+data = add_missing_variables(check_interaction_data, data) 
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+
+#Convert degree minute second to decimal degree coordinates
+library(parzer)
+
+#Prepare data before convert coordinates
+data = data %>% 
+mutate(Latitude = gsub("\\°.", "° ", Latitude)) %>% 
+mutate(Latitude = gsub("\\'.", "' ", Latitude)) %>% 
+mutate(Longitude = gsub("\\°.", "° ", Longitude)) %>% 
+mutate(Longitude = gsub("\\'.", "' ", Longitude)) 
[…]

### 2022-12-19 Scripts/Map_study_locations.Rmd
edits

--- 2022-12-17 Scripts/Map_study_locations.Rmd
@@
 "salmon", "palegreen","greenyellow",
-"mediumaquamarine")
+"mediumaquamarine", "tomato", "orange", "blue")
 #To generate more random colors 

### 2022-12-20 Scripts/Map_study_locations.Rmd
Datasets from Ireland

--- 2022-12-17 Scripts/Map_study_locations.Rmd
@@
 "salmon", "palegreen","greenyellow",
-"mediumaquamarine")
+"mediumaquamarine", "tomato", "orange", 
+"blue", "black", "green")
 #To generate more random colors 

### 2022-12-22 Scripts/Map_study_locations.Rmd
Maurer dataset and additional edits

--- 2022-12-19 Scripts/Map_study_locations.Rmd
@@
 "salmon", "palegreen","greenyellow",
-"mediumaquamarine", "tomato", "orange")
+"mediumaquamarine", "tomato", "orange", 
+"blue", "black", "green", "purple")
 #To generate more random colors 

### 2022-12-22 Scripts/Map_study_locations.Rmd
Edit on Maurer metadata
edits on map

@@
 output:
+  pdf_document: default
   html_document: default
-  pdf_document: default
 ---
@@
 g <- ggplot() +
-geom_sf(data = cntries, fill = "grey80", color = NA) +
+geom_sf(data = cntries, fill = "grey85", color = "grey60",size = 0.1) +
 xlim(c(2200000, 7150000)) +

### 2022-12-23 Scripts/Map_study_locations.Rmd
All datasets added!

@@
 
-colors <- c("#8AD69E", "#74C5CE",  "#FDE9DE",  "#45A631", "#95F7F0", "#CB53E1",
+colors <- c(
+"#8AD69E", "#74C5CE",  "#FDE9DE",
+"#45A631", "#95F7F0", "#CB53E1",
 "#FE90BB", "#B187E1", "#F8A982", 
@@
 "mediumaquamarine", "tomato", "orange", 
-"blue", "black", "green", "purple")
+"blue", "black", "green",
+"purple", "red")
 #To generate more random colors 

### 2022-12-23 Scripts/Processing/28_Albrecth.R

@@ -1,188 +1,0 @@
-#DATASET NUMBER 28; Albrecht 
-#Some species can be recovered from this dataset
-#There are some unmatching plant names
-#I have done the best I can but there are somoe missing ones
-#As far as I can read most are non-relevant poaceae for pollination
-#For instance, dactylis glomerolata, poa, lolium, etc...
-
-#Read templates to compare with
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(dplyr)
-library(stringr)
-library(readr)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/28_Albrecht/Interaction_data.csv")
-
-#Check col names with template
-compare_variables(check_interaction_data, data)
-
-#Rename variables accordingly
-data = data %>% 
-rename(Plant_species = Plant_species_corrected) %>% 
-rename(Habitat = Focal_habitat) %>% 
-mutate(Habitat =  as.factor(Habitat))  %>% 
-mutate(Habitat = recode_factor(Habitat, "HA" = "Herbaceous areal 1",
-                              "HA2" = "Herbaceous areal 2",
-                              "HA3" = "Herbaceous areal 3",
-                              "HL" = "Herbaceous linear",
-                              "HL2" = "Herbaceous linear 2",
-                              "WA" = "Woody areal",
-                              "WL" = "Woody linear",
-                              "WL2" = "Woody linear 2"))
-
-#Add missing variables
-data = add_missing_variables(check_interaction_data, data)
-
-#Plant species are just codes at the moment
-#Load flower count data which has the correct species names 
-plant_names <- read_csv("Data/Raw_data/28_Albrecht/Flower_count.csv") %>% 
-select(Plant_species, Species_abbreviation) %>% 
-rename(Plant_names_new = Plant_species, Plant_species = Species_abbreviation)
-
-#Recode some factors
-plant_names = plant_names %>% 
-mutate(Plant_species = case_when(
-Plant_names_new == "Taraxacum.campylodes" ~ "tar.off", 
-T ~ Plant_species))
-#Recode Silene
-plant_names = plant_names %>% 
[…]

### 2022-12-28 index.qmd
quarto

---
title: "Punk Title Here"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r load_things, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(gh) #How heavy would this library be?

releases <- gh("GET /repos/PunkUserName/PunkRepoName/releases",
   owner = "PunkUserName",
   repo = "PunkRepoName")
tag <- releases[[1]][["tag_name"]]
```

## Authors

Punk Author ^1^, ...

> (1) Does Punks care about their institutions?

**Last update:** `r Sys.Date()`

## Abstract

Tell me why I need to read this, and do it fast. Also, this document is is version `r tag`.

### How to cite this:

You can cite this dynamic document: P. Author. `r substr(Sys.Date(), 1, 4)` Punk Title. version `r tag`. DOI: [10.0000/zenodo.0000000](https://doi.org/10.0000/zenodo.0000000).

### Download in PDF:

### Source code:

You can find the source code, as well as previous releases of this repositoty at: [https://github.com/PunkUserName/PunkRepoName](https://github.com/PunkUserName/PunkRepoName)

### 2022-12-28 Scripts/Processing/10_Vanbergen.R

@@ -1,83 +1,0 @@
-#DATASET NUMBER 10; Vanbergen
-
-#Load libraries
-library(tidyverse)
-
-#Prepare interaction data ----
-data = read.csv("Data/Raw_data/10_Vanbergen/Interaction_data.csv")
-#Create cols of Temperature and humidity
-data = data %>% 
-mutate(Temperature = NA) %>% 
-mutate(Humidity = NA) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters))
-
-#Split by Site_id
-InteractionData <- split(data, data$Site_id)
-
-
-#Prepare flower count data ----
-flower_count = read.csv("Data/Raw_data/10_Vanbergen/Flower_count.csv")
-
-#Filter out one column but saved the data
-flower_count = flower_count %>% 
-mutate(Units = paste0(Units, "; ", Inflorescence.type)) %>% 
-select(!Inflorescence.type)
-#Split data into different dataframes based on survey name
-FlowerCount <- split(flower_count, flower_count$Site_id)
-
-#Prepare metadata data ----
-
-#Store unique cases of plants and polls
-plant_single_cases = data %>% distinct(Plant_species)
-pollinator_single_cases = data %>%distinct(Pollinator_species)
-
-Metadata <- tibble(
-Doi = NA,
-Dataset_description = "This dataset is part of the VOODOO-project.
-It consists of plant-pollinator networks in 12 different landscapes
-(4 agricultural, 4 rural, 4 urban), sampled over 3 periods in 2021
-(mid april, end may and mid july). Coordinates are the centroids of each landscape.
-Transects were stratified in relation with the area of each type of pollinator habitat
-within a 500 m radius around the centroid. Habitat is given for each individual catch,
-not for the landscape as a whole. Transect length and duration are therefore pooled in
-this file, rather than giving the sampling time in each habitat
-(in total: 120 min per site visit, so 360 min in total, if the networks are pooled
-over the 3 sampling periods). If requested, we can also provide sampling time from
-each subtransect in different habitat types.",
-Taxa_recorded = "All flying insects, except really small stuff (< 1 mm)",
-Sampling_year = 2021,
-Country = "France",
-Habitat = "4 agricultural, 4 rural, 4 urban",
-Sampling_sites = 12,
[…]

### 2022-12-28 Scripts/Processing/34_Stanley.R

@@ -1,163 +1,0 @@
-#DATASET NUMBER 34; Stanley
-#ALL datasets were gathered by Laura Russo who organised this
-#(There are 6 in total)
-#One is an unpublished thesis
-
-#Read empty templates to compare with
-source("Scripts/Empty_templates.R") 
-
-#Load libraries
-library(dplyr)
-library(stringr)
-library(readr)
-library(tibble)
-library(tidyr)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
-
-#Divide col into cols
-levels(factor(data$Site_id))
-data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
-levels(factor(data$Name))
-
-#Recode authors to just their surname
-data = data %>% 
-mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
-                            "EileenPower" = "Power",
-                            "DaraStanley" = "Stanley",
-                            "SarahMullen" = "Mullen",
-                            "MichelleLarkin" = "Larkin",
-                            "CianWhite" = "White"))
-
-#Select first the one of Power
-data = data %>% filter(Name == "Stanley")
-
-#To calculate sampling effort later on
-data_time = data %>% select(Sampling_effort_minutes)
-
-#Compare vars
-#compare_variables(check_interaction_data, data)
-
-#Prepare data
-data = data %>% 
-rename(Plant_species = Plant_Species) %>% 
-mutate(Habitat = "Farming grassland") %>% 
-mutate(Locality = "Farms in south-east Ireland") %>% 
-rename(Latitude = "WGS84/ETRS89") %>% 
-rename(Longitude = "WGS84/ETRS89_2") %>% 
-mutate(Site_id = Site) %>% 
-mutate(Sampling_method = "Transects") 
-
-#Add missing vars
-data = add_missing_variables(check_interaction_data, data) 
-#Reorder variables
-data = drop_variables(check_interaction_data, data) 
-
-
-#Convert degree minute second to decimal degree coordinates
-library(parzer)
-
-#Prepare data before convert coordinates
-data = data %>% 
-mutate(Latitude = gsub("\\°.", "° ", Latitude)) %>% 
-mutate(Latitude = gsub("\\'.", "' ", Latitude)) %>% 
-mutate(Longitude = gsub("\\°.", "° ", Longitude)) %>% 
[…]

### 2022-12-28 Scripts/Processing/37_White.R

@@ -1,169 +1,0 @@
-#DATASET NUMBER 37; White
-#ALL datasets from RUSSO (There are 6 in total)
-#One is an unpublished thesis
-
-#Read empty templates to compare with
-source("Scripts/Empty_templates.R") 
-
-#Load libraries
-library(dplyr)
-library(stringr)
-library(readr)
-library(tibble)
-library(tidyr)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
-
-#Divide col into cols
-levels(factor(data$Site_id))
-data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
-levels(factor(data$Name))
-
-#Recode authors to just their surname
-data = data %>% 
-  mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
-                              "EileenPower" = "Power",
-                              "DaraStanley" = "Stanley",
-                              "SarahMullen" = "Mullen",
-                              "MichelleLarkin" = "Larkin",
-                              "CianWhite" = "White"))
-
-#Select first the one of Power
-data = data %>% filter(Name == "White")
-
-#To calculate sampling effort later on
-data_time = data %>% select(Sampling_effort_minutes)
-
-#Compare vars
-#compare_variables(check_interaction_data, data)
-
-#Prepare data
-data = data %>% 
-  rename(Plant_species = Plant_Species) %>% 
-  mutate(Habitat = "Agricultrual and urban sites with different intensity") %>% 
-  mutate(Locality = "East Leinster (Ireland)") %>% 
-  rename(Latitude = "WGS84/ETRS89") %>% 
-  rename(Longitude = "WGS84/ETRS89_2") %>% 
-  mutate(Site_id = Site) %>% 
-  mutate(Sampling_method = "Transects") 
-
-#Add missing vars
-data = add_missing_variables(check_interaction_data, data) 
-#Reorder variables
-data = drop_variables(check_interaction_data, data) 
-
-
-#Convert degree minute second to decimal degree coordinates
-library(parzer)
-
-#Prepare data before convert coordinates
-data = data %>% 
-  mutate(Latitude = gsub("\\°.", "° ", Latitude)) %>% 
-  mutate(Latitude = gsub("\\'.", "' ", Latitude)) %>% 
-  mutate(Longitude = gsub("\\°.", "° ", Longitude)) %>% 
[…]

### 2022-12-28 Scripts/Processing/8_Biella.R

@@ -1,74 +1,0 @@
-#DATASET NUMBER 8; Biella
-
-#Load libraries
-library(tidyverse)
-
-#Prepare interaction data ----
-data = read.csv("Data/Raw_data/8_Biella/Interaction_data.csv")
-
-#Standardize data
-data = data %>% 
-mutate(Plant_species = str_replace(Plant_species, "_", " ")) %>% 
-mutate(Pollinator_species = str_replace(Pollinator_species, "_", " ")) %>% 
-mutate(Pollinator_species = str_to_sentence(Pollinator_species)) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) 
-
-
-#Split data into different dataframes based on survey name
-InteractionData <- split(data, data$Site_id)
-
-#Prepare flower count data ----
-flower_count = read.csv("Data/Raw_data/8_Biella/Flower_count.csv")
-#Split data into different dataframes based on survey name
-FlowerCount <- split(flower_count, flower_count$Site_id)
-
-#Prepare metadata data ----
-
-#Store unique cases of plants and polls
-plant_single_cases = data %>% distinct(Plant_species)
-pollinator_single_cases = data %>%distinct(Pollinator_species)
-
-#Generate metadata
-Metadata <- tibble(
-Doi = "https://doi.org/10.1556/168.2017.18.1.1",
-Dataset_description = "This dataset documents 2 communities on different slopes
-of a mountain in the Northen Apennine",
-Taxa_recorded = "All flower visitors",
-Sampling_sites = 6,
-Sampling_rounds = "Weekly sampling",
-Year = 2013,
-Country = "Italia",
-Sampling_method = "Plots",
-Sampling_area_details = "2.5 * 2.5 m plots",
-Sampling_area_species_m2 = NA,
-Sampling_area_total_m2 = 2.5 * 2.5 * 6,
-Sampling_time_details = "20 minutes / plot and round",
-Sampling_time_species_round_min = NA,
-Sampling_time_total_min = NA,
-Total_plant_species = nrow(plant_single_cases),
-Total_pollinator_species = nrow(pollinator_single_cases),
-Floral_counts =  "Yes")
- 
-
-#Transpose metadata
-Metadata = as.data.frame(t(Metadata)) %>%  
-rownames_to_column() %>% 
-rename(Metadata_fields = rowname, Metadata_info= V1) %>% as_tibble()
-
-
-#Prepare authorship data ----
[…]

### 2022-12-28 Scripts/Read_all.R

@@ -1,94 +1,0 @@
-library('gtools') #Library that allows to sort files by numerical order
-library(tidyverse)
-# install.packages("sf")
-library(sf)
-# install.packages("giscoR")
-library(giscoR)
-
-#Read all files with their paths
-files <- dir("Data/Clean_data", pattern="*.rds", full.names=T)
-files <- mixedsort(files)
-
-#Generate list of lists
-all_data = files %>%
-map(readRDS) 
-
-#Read all file names
-file_names <- dir("Data/Clean_data", pattern="*.rds", full.names=F)
-file_names <- mixedsort(file_names)
-
-#Delete extension
-file_names = str_replace(file_names, ".rds", "")
-#Rename all elements of the list
-names(all_data) <- file_names
-
-
-#Select coordinates from the list of lists----
-data <- list()
-
-for (i in 1:length(file_names)) {
-  
-  int_list <- all_data[[file_names[i]]]$InteractionData
-  data[[i]] <- bind_rows(map(int_list, function(x) x %>% #map here is the same as a lapply
-                dplyr::select(Longitude, Latitude)),  .id = 'Network_id')
-}
-
-
-
-
-#Rename list and 
-names(data) <- file_names
-
-#Prepare data to save it
-coord_list = data
-coord_long_format = bind_rows(data,  .id = 'Study_id')
-
-save(coord_list, coord_long_format, file = "Data/Processing/Coordinates.RData")
-save(coord_list, coord_long_format, file = "~/R_projects/SafeNet/Data/Coordinates.RData")
-
-load("Data/Processing/Coordinates.RData")
-
-#Prepare data to plot it
-data = bind_rows(data,  .id = 'Study_id')
-data_coord <- st_as_sf(data, coords = c(3:4)) %>% 
-st_set_crs(4326) %>% 
-st_transform(3035)
-
-data_coord$Study_id <- factor(data_coord$Study_id, levels = file_names)
-#Save data
-st_write(data_coord, "Data/Processing/data_coord.shp", delete_layer = TRUE)
-st_write(data_coord, "~/R_projects/SafeNet/Data/data_coord.shp", delete_layer = TRUE)
-
-# Year
-year_ref <- 2016
-
-# Get all countries and transform to the same CRS
-cntries <- gisco_get_countries(year = year_ref,
-        resolution = 03) %>%
-        st_transform(3035)
[…]

### 2022-12-29 Scripts/Example_select_one_column.R
fixing repo mistake

@@ -1,0 +1,39 @@
+
+library(tidyverse)
+
+#Generate datasets to create example
+dat1 <- tibble(col1 = runif(10), col2 = stringi::stri_rand_strings(10, 5))
+dat2 <- tibble(col1 = runif(10), col2 = stringi::stri_rand_strings(10, 5))
+dat3 <- tibble(col1 = runif(10), col2 = stringi::stri_rand_strings(10, 5))
+dat4 <- tibble(col1 = runif(10), col2 = stringi::stri_rand_strings(10, 5))
+
+#Generate lists (3 of them for example)
+list1 <- list(id1 = dat1, id2 = dat2)
+list2 <- list(id1 = dat3)
+list3 <- list(id1 = dat4)
+#Generate list of lists
+final_list <- list(dataset1 = list1, dataset2 = list2, dataset3 = list3)
+
+
+#select unique cases from the list of lists
+data <- NULL
+
+for (i in final_list) {
+
+single_cases <- bind_rows(lapply(i, function(x) x %>% select(col2) %>% distinct(col2)))
+
+data <- rbind(data, single_cases)  
+  
+}
+
+
+#select unique cases from one of the lists
+single_cases <- bind_rows(lapply(list1, function(x) x %>% select(col2) %>% distinct(col2)))
+
+str(data1)
+
+data1 <- c(list1, list2, list3) %>%
+  setNames(paste0("id", seq_along(.))) %>%
+  bind_rows(.id = "id")  %>%
+  distinct(id, col2) %>%
+  select(-id)

### 2022-12-29 Scripts/Processing/21_Hopfenmuller.R

@@ -1,0 +1,110 @@
+#DATASET NUMBER 21; Bernhard Hopfenmuller 
+source("Scripts/Empty_templates.R") #Read empty templates to compare with
+
+#Load libraries
+library(tidyverse)
+library(lubridate)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/21_Hopfenmuller/Interaction_data.csv")
+
+#Compare vars
+compare_variables(check_interaction_data, data)
+
+#Data
+data = data %>% 
+mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
+mutate(Year = year(ymd(Date))) %>% 
+mutate(Month = month(ymd(Date))) %>% 
+mutate(Day = day(ymd(Date))) 
+
+#Add interaction col
+data = data %>% 
+mutate(Interaction = 1) %>% 
+mutate(Sampling_method = "Variable transect walks") %>% 
+mutate(Sampling_effort_minutes = NA) %>% 
+mutate(Sampling_area_square_meters = NA) %>% 
+mutate(Habitat = "Calcareous grassland") %>% 
+mutate(Locality = Site_id) 
+
+#Add missing vars
+data = add_missing_variables(check_interaction_data, data) 
+
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+#Drop this last two columns that are going to be at the metadata
+data = data %>% 
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
+
+#Split interaction data into dataframes within a list
+InteractionData <- split(data,data$Site_id)
+
+
+#Prepare flower count data ----
+flower_count <- read_csv("Data/Raw_data/21_Hopfenmuller/Flower_count.csv")
+
+#Compare vars
+compare_variables(check_flower_count_data, flower_count)
+
+#Fix names and variable
+flower_count = flower_count %>% 
+rename(Flower_count = Flower_units) %>% 
+mutate(Units = "Flower number")
+
+#Add variables
+flower_count = add_missing_variables(check_flower_count_data, flower_count) 
+
+#Order data as template
+flower_count = drop_variables(check_flower_count_data, flower_count) 
+
+#Split interaction data into dataframes within a list
+FlowerCount <- split(flower_count, flower_count$Site_id)
+
+#Prepare metadata data ----
+#Select unique cases of plants and poll
[…]

### 2022-12-29 Scripts/Processing/28_Sutter.R

@@ -1,0 +1,286 @@
+#DATASET NUMBER 28; Sutter 
+#There are some unmatching plant names
+
+
+#Read templates to compare with
+source("Scripts/Empty_templates.R") #Read empty templates to compare with
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/28_Sutter/Interaction_data.csv")
+
+#Filter interactions greater than 0
+#And filter out insects caught flying
+data = data %>% 
+filter(Interaction > 0) %>% 
+filter(!Plant_species_corrected == "transect")
+
+#Check col names with template
+compare_variables(check_interaction_data, data)
+
+#Rename variables accordingly
+data = data %>% 
+rename(Plant_species = Plant_species_corrected) %>% 
+rename(Habitat = Focal_habitat) %>% 
+mutate(Habitat =  as.factor(Habitat))  %>% 
+mutate(Habitat = recode_factor(Habitat, "HA" = "Herbaceous areal 1",
+                              "HA2" = "Herbaceous areal 2",
+                              "HA3" = "Herbaceous areal 3",
+                              "HL" = "Herbaceous linear",
+                              "HL2" = "Herbaceous linear 2",
+                              "WA" = "Woody areal",
+                              "WL" = "Woody linear",
+                              "WL2" = "Woody linear 2"))
+
+#Add missing variables
+data = add_missing_variables(check_interaction_data, data)
+
+#Plant species are just codes at the moment
+#Load flower count data which has the correct species names 
+plant_names <- read_csv("Data/Raw_data/28_Sutter/Flower_count.csv") %>% 
+select(Plant_species, Species_abbreviation) %>% 
+rename(Plant_names_new = Plant_species, Plant_species = Species_abbreviation)
+
+#Recode some factors
+plant_names = plant_names %>% 
+mutate(Plant_species = case_when(
+Plant_names_new == "Taraxacum.campylodes" ~ "tar.off", 
+T ~ Plant_species))
+#Recode Silene
+plant_names = plant_names %>% 
+mutate(Plant_species = case_when(
[…]

### 2022-12-29 Scripts/Processing/35_Mullen.R

@@ -1,0 +1,162 @@
+#DATASET NUMBER 35; Mullen
+#ALL datasets were gathered by Laura Russo who organised this
+#(There are 6 in total)
+#One is an unpublished thesis
+
+#Read empty templates to compare with
+source("Scripts/Empty_templates.R") 
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
+
+#Divide col into cols
+levels(factor(data$Site_id))
+data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
+levels(factor(data$Name))
+
+#Recode authors to just their surname
+data = data %>% 
+  mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
+                              "EileenPower" = "Power",
+                              "DaraStanley" = "Stanley",
+                              "SarahMullen" = "Mullen",
+                              "MichelleLarkin" = "Larkin",
+                              "CianWhite" = "White"))
+
+#Select first the one of Power
+data = data %>% filter(Name == "Mullen")
+
+#To calculate sampling effort later on
+data_time = data %>% select(Sampling_effort_minutes)
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+
+#Prepare data
+data = data %>% 
+  rename(Plant_species = Plant_Species) %>% 
+  mutate(Habitat = "Farming grassland") %>% 
+  mutate(Locality = "Farms in south-east Ireland") %>% 
+  rename(Latitude = "WGS84/ETRS89") %>% 
+  rename(Longitude = "WGS84/ETRS89_2") %>% 
+  mutate(Site_id = Site) %>% 
+  mutate(Sampling_method = "Transects") 
+
+#Add missing vars
+data = add_missing_variables(check_interaction_data, data) 
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+#Convert degree minute second to decimal degree coordinates
+library(parzer)
+
+#Prepare data before convert coordinates
+data = data %>% 
+  mutate(Latitude = gsub("\\°.", "° ", Latitude)) %>% 
+  mutate(Latitude = gsub("\\'.", "' ", Latitude)) %>% 
[…]

### 2023-09-23 Manuscript/Ms.qmd
Adding Ms folder

---
title: "Mapping plant-pollinator interactions across Europe"
format: pdf
editor: visual
---

# Ideas/objectives:

- Write article point by point

- Ecography could be our potential journal (think about it, high fees)

- Finish paper by Christmas

- Create workflow graph

### 2023-09-25 Scripts/1_Visualisation/Map_study_locations.Rmd
Organising files

--- 2022-12-23 Scripts/Map_study_locations.Rmd
@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 #Read all files with their paths
-files <- dir("../Data/Clean_data", pattern="*.rds", full.names=T)
+files <- dir("../../Data/Clean_data", pattern="*.rds", full.names=T)
 files <- mixedsort(files)
@@
 #Read all file names
-file_names <- dir("../Data/Clean_data", pattern="*.rds", full.names=F)
+file_names <- dir("../../Data/Clean_data", pattern="*.rds", full.names=F)
 file_names <- mixedsort(file_names)

### 2023-09-25 Scripts/1_Processing_networks/20_Hoiss.R

@@ -1,0 +1,114 @@
+#DATASET NUMBER 20; Bernhard Hoiss 
+source("Scripts/Empty_templates.R") #Read empty templates to compare with
+
+#Load libraries
+library(tidyverse)
+library(lubridate)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/20_Hoiss/Interaction_data.csv")
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+
+#Covert day into three separate cols
+data = data %>% 
+mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
+mutate(Year = year(ymd(Date))) %>% 
+mutate(Month = month(ymd(Date))) %>% 
+mutate(Day = day(ymd(Date))) 
+
+#Add interaction col
+data = data %>% 
+mutate(Interaction = 1) %>% 
+mutate(Sampling_method = "Transect") %>% 
+mutate(Habitat = "Grassland") %>% 
+mutate(Locality = NA) %>% 
+mutate(Coordinate_precision = NA) %>% 
+mutate(Comments = NA) %>% 
+mutate(Temperature = NA) %>% 
+mutate(Humidity = NA) %>% 
+mutate(Sampling_effort_minutes = NA) %>% 
+mutate(Sampling_area_square_meters = NA)
+  
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+#Drop this last two columns that are going to be at the metadata
+data = data %>% 
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
+
+#Split interaction data into dataframes within a list
+InteractionData <- split(data,data$Site_id)
+
+#Prepare flower count data ----
+flower_count <- read_csv("Data/Raw_data/20_Hoiss/Flower_count.csv")
+
+#Compare vars
+#compare_variables(check_flower_count_data, flower_count)
+
+#Covert day into three separate cols
+flower_count = flower_count %>% 
+mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
+mutate(Year = year(ymd(Date))) %>% 
+mutate(Month = month(ymd(Date))) %>% 
+mutate(Day = day(ymd(Date))) %>% 
+rename(Flower_count = Flower_percent_cover) %>% 
+mutate(Units = "Percent cover") %>% 
+mutate(Comment = NA)
+
+#Order data as template
+flower_count = drop_variables(check_flower_count_data, flower_count) 
+
[…]

### 2023-09-25 Scripts/1_Processing_networks/28_Sutter.R

@@ -1,0 +1,287 @@
+#DATASET NUMBER 28; Sutter 
+#There are some unmatching plant names
+
+
+#Read templates to compare with
+source("Scripts/Empty_templates.R") #Read empty templates to compare with
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/28_Sutter/Interaction_data.csv")
+
+#Filter interactions greater than 0
+#And filter out insects caught flying
+data = data %>% 
+filter(Interaction > 0) %>% 
+filter(!Plant_species_corrected == "transect")
+
+#Check col names with template
+compare_variables(check_interaction_data, data)
+
+#Rename variables accordingly
+data = data %>% 
+rename(Plant_species = Plant_species_corrected) %>% 
+rename(Habitat = Focal_habitat) %>% 
+mutate(Habitat =  as.factor(Habitat))  %>% 
+mutate(Habitat = recode_factor(Habitat, "HA" = "Herbaceous areal 1",
+                              "HA2" = "Herbaceous areal 2",
+                              "HA3" = "Herbaceous areal 3",
+                              "HL" = "Herbaceous linear",
+                              "HL2" = "Herbaceous linear 2",
+                              "WA" = "Woody areal",
+                              "WL" = "Woody linear",
+                              "WL2" = "Woody linear 2")) %>% 
+mutate(Country = "Switzerland")
+
+#Add missing variables
+data = add_missing_variables(check_interaction_data, data)
+
+#Plant species are just codes at the moment
+#Load flower count data which has the correct species names 
+plant_names <- read_csv("Data/Raw_data/28_Sutter/Flower_count.csv") %>% 
+select(Plant_species, Species_abbreviation) %>% 
+rename(Plant_names_new = Plant_species, Plant_species = Species_abbreviation)
+
+#Recode some factors
+plant_names = plant_names %>% 
+mutate(Plant_species = case_when(
+Plant_names_new == "Taraxacum.campylodes" ~ "tar.off", 
+T ~ Plant_species))
+#Recode Silene
[…]

### 2023-09-25 Scripts/1_Processing_networks/32_ORourke.R

@@ -1,0 +1,142 @@
+#DATASET NUMBER 32; O’Rourke
+#ALL datasets were gathered by Laura Russo who organised this
+#(There are 6 in total)
+#One is an unpublished thesis
+#Although Russo is not an explicit author of any of them
+#I have added her as author in this first paper so 
+#I don't forget her authorship
+
+#Read empty templates to compare with
+source("Scripts/Empty_templates.R") 
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
+
+#Divide col into cols
+levels(factor(data$Site_id))
+data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
+levels(factor(data$Name))
+
+#Recode authors to just their surname
+data = data %>% 
+mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
+                            "EileenPower" = "Power",
+                            "DaraStanley" = "Stanley",
+                            "SarahMullen" = "Mullen",
+                            "MichelleLarkin" = "Larkin",
+                            "CianWhite" = "White"))
+
+#Select first the one of O’Rourke
+data = data %>% filter(Name == "O’Rourke")
+
+#To calculate sampling effor later on
+data_time = data %>% select(Sampling_effort_minutes)
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+
+data = data %>% 
+rename(Plant_species = Plant_Species) %>% 
+mutate(Habitat = "Dune ecosystem") %>% 
+mutate(Locality = "Along south-eastern Ireland") %>% 
+rename(Latitude = "WGS84/ETRS89") %>% 
+rename(Longitude = "WGS84/ETRS89_2") %>% 
+mutate(Site_id = Site_number) %>% 
+mutate(Sampling_method = "Focal_observations") #From the paper I understand that is focal and not transect
+
+#Add missing vars
+data = add_missing_variables(check_interaction_data, data) 
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+
+#Convert degree minute second to decimal degree coordinates
+library(parzer)
+
[…]

### 2023-09-25 Scripts/1_Processing_networks/33_Power.R

@@ -1,0 +1,164 @@
+#DATASET NUMBER 33; Power
+#ALL datasets were gathered by Laura Russo who organised this
+#(There are 6 in total)
+#One is an unpublished thesis
+
+#Read empty templates to compare with
+source("Scripts/Empty_templates.R") 
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
+
+#Divide col into cols
+levels(factor(data$Site_id))
+data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
+levels(factor(data$Name))
+
+#Recode authors to just their surname
+data = data %>% 
+mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
+                            "EileenPower" = "Power",
+                            "DaraStanley" = "Stanley",
+                            "SarahMullen" = "Mullen",
+                            "MichelleLarkin" = "Larkin",
+                            "CianWhite" = "White"))
+
+#Select first the one of Power
+data = data %>% filter(Name == "Power")
+
+#To calculate sampling effort later on
+data_time = data %>% select(Sampling_effort_minutes)
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+
+#Prepare data
+data = data %>% 
+rename(Plant_species = Plant_Species) %>% 
+mutate(Habitat = "Farming grassland") %>% 
+mutate(Locality = "Along south-eastern Ireland") %>% 
+rename(Latitude = "WGS84/ETRS89") %>% 
+rename(Longitude = "WGS84/ETRS89_2") %>% 
+mutate(Site_id = Site) %>% 
+mutate(Sampling_method = "Transects") 
+
+#Add missing vars
+data = add_missing_variables(check_interaction_data, data) 
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+
+#Convert degree minute second to decimal degree coordinates
+library(parzer)
+
+#Prepare data before convert coordinates
+data = data %>% 
+mutate(Latitude = gsub("\\°.", "° ", Latitude)) %>% 
+mutate(Latitude = gsub("\\'.", "' ", Latitude)) %>% 
+mutate(Longitude = gsub("\\°.", "° ", Longitude)) %>% 
[…]

### 2023-09-25 Scripts/1_Processing_networks/37_White.R

@@ -1,0 +1,169 @@
+#DATASET NUMBER 37; White
+#ALL datasets from RUSSO (There are 6 in total)
+#One is an unpublished thesis
+
+#Read empty templates to compare with
+source("Scripts/Empty_templates.R") 
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
+
+#Divide col into cols
+levels(factor(data$Site_id))
+data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
+levels(factor(data$Name))
+
+#Recode authors to just their surname
+data = data %>% 
+  mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
+                              "EileenPower" = "Power",
+                              "DaraStanley" = "Stanley",
+                              "SarahMullen" = "Mullen",
+                              "MichelleLarkin" = "Larkin",
+                              "CianWhite" = "White"))
+
+#Select first the one of Power
+data = data %>% filter(Name == "White")
+
+#To calculate sampling effort later on
+data_time = data %>% select(Sampling_effort_minutes)
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+
+#Prepare data
+data = data %>% 
+  rename(Plant_species = Plant_Species) %>% 
+  mutate(Habitat = "Agricultrual and urban sites with different intensity") %>% 
+  mutate(Locality = "East Leinster (Ireland)") %>% 
+  rename(Latitude = "WGS84/ETRS89") %>% 
+  rename(Longitude = "WGS84/ETRS89_2") %>% 
+  mutate(Site_id = Site) %>% 
+  mutate(Sampling_method = "Transects") 
+
+#Add missing vars
+data = add_missing_variables(check_interaction_data, data) 
+#Reorder variables
+data = drop_variables(check_interaction_data, data) 
+
+
+#Convert degree minute second to decimal degree coordinates
+library(parzer)
+
+#Prepare data before convert coordinates
+data = data %>% 
+  mutate(Latitude = gsub("\\°.", "° ", Latitude)) %>% 
+  mutate(Latitude = gsub("\\'.", "' ", Latitude)) %>% 
+  mutate(Longitude = gsub("\\°.", "° ", Longitude)) %>% 
[…]

### 2023-09-25 Scripts/1_Read_all.R

@@ -1,0 +1,103 @@
+library('gtools') #Library that allows to sort files by numerical order
+library(tidyverse)
+# install.packages("sf")
+library(sf)
+# install.packages("giscoR")
+library(giscoR)
+
+#Read all files with their paths
+files <- dir("Data/Clean_data", pattern="*.rds", full.names=T)
+files <- mixedsort(files)
+
+#Generate list of lists
+all_data = files %>%
+map(readRDS) 
+
+#Read all file names
+file_names <- dir("Data/Clean_data", pattern="*.rds", full.names=F)
+file_names <- mixedsort(file_names)
+
+#Delete extension
+file_names = str_replace(file_names, ".rds", "")
+#Rename all elements of the list
+names(all_data) <- file_names
+
+
+#Select coordinates from the list of lists----
+data <- list()
+
+for (i in 1:length(file_names)) {
+  
+  int_list <- all_data[[file_names[i]]]$InteractionData
+  data[[i]] <- bind_rows(map(int_list, function(x) x %>% #map here is the same as a lapply
+                dplyr::select(Longitude, Latitude, Country)),  .id = 'Network_id')
+}
+
+#Rename list and 
+names(data) <- file_names
+
+#Prepare data to save it
+coord_list = data
+coord_long_format = bind_rows(data,  .id = 'Study_id')
+
+#Check country levels
+levels(factor(coord_long_format$Country))
+
+#Create another col with study id and network id together so we can count all individuals networks
+coord_long_format = coord_long_format %>% 
+mutate(Study_network_id = paste0(Study_id, Network_id))
+
+#Save data
+save(coord_list, coord_long_format, file = "Data/Processing/Coordinates.RData")
+save(coord_list, coord_long_format, file = "~/R_projects/SafeNetApp/Data/Coordinates.RData")
+save(coord_list, coord_long_format, file = "~/R_projects/SafeNetWeb/Data/Coordinates.RData")
+
+#Load data
+load("Data/Processing/Coordinates.RData")
+
+#Prepare data to plot it
+data = bind_rows(data,  .id = 'Study_id')
+data_coord <- st_as_sf(data, coords = c(3:4)) %>% 
+st_set_crs(4326) %>% 
+st_transform(3035)
+
+data_coord$Study_id <- factor(data_coord$Study_id, levels = file_names)
[…]

### 2023-09-25 Scripts/2_Processing_metaweb/prep_data.R

@@ -1,0 +1,246 @@
+#create a EU metaweb by 
+  #complie all interactions in a single file from
+    #safenet  
+    #Alham
+    #Globi: https://github.com/Big-Bee-Network/global-bee-interaction-data
+    #Literature: https://github.com/Big-Bee-Network/bee-interaction-database
+  #transform it into a matrix
+  #Plot it nicely (if this is possible?)
+
+
+#create the structure of Plants and pol to be considered:-----
+#Polinators: Bee + sirphids 
+#Pollinators: Bombilids + beetles? Skip?
+#Plants: ??
+
+
+#start by reading SafeNet data----
+#Ideally from the repo once public. Now I made a copy in my personal computer
+
+#Example of one file.
+temp <- readRDS("scripts/temp_data/Clean_data/1_Bartomeus.rds")
+#str(temp)
+#temp$Metadata #Most fields already in interaction data (e.g. where and when)
+#temp$Authorship #Most fields already in interaction data (e.g. where and when)
+#merge all sites
+out<- temp$InteractionData[[1]]
+for(i in 2:length(temp$InteractionData)){
+  temp2 <- temp$InteractionData[[i]]
+  out <- rbind(temp2, out)
+}
+#head(out)
+#dim(out)
+colnames(out)
+#loop through all files
+#Read all file names to loop.
+files <- list.files(path = "scripts/temp_data/Clean_data")
+master <- data.frame(Plant_species = NA   ,Pollinator_species = NA   ,Interaction = NA   ,      
+                     Sampling_method = NA ,Site_id = NA              ,Habitat = NA       ,     
+                     Country = NA         ,Locality = NA             ,Latitude = NA      ,      
+                     Longitude = NA       ,Coordinate_precision = NA ,Elevation = NA     ,        
+                     Day = NA             ,Month = NA                ,Year = NA          ,      
+                     Comments = NA        ,Temperature = NA          ,Humidity = NA      ,
+                     Study = NA)
+for(k in 1:length(files)){
+  temp <- readRDS(paste0("scripts/temp_data/Clean_data/", files[k]))
+  #merge all sites
+  out<- temp$InteractionData[[1]]
+  if(length(temp$InteractionData) > 1){
+    for(i in 2:length(temp$InteractionData)){
[…]

### 2023-09-25 Scripts/Processing/17_Manincor.R

@@ -1,91 +1,0 @@
-#DATASET NUMBER 17; Manincor
-
-#Load libraries
-library(tidyverse)
-
-#Prepare interaction data ----
-#Load interaction data
-data = read_csv("Data/Raw_data/16_17_Manincor/Interaction_data2.csv", col_names = T)
-
-#Convert coordinates in degree mins and seconds to lat long in decimal degrees
-data = data %>%
-  mutate(Longitude = parzer::parse_lon(Longitude),
-         Latitude = parzer::parse_lat(Latitude)) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
-
-
-unique(factor(data$Latitude))
-unique(factor(data$Longitude))
-
-#Split data into different dataframes based on site id
-InteractionData <- split(data, data$Site_id)
-
-#Prepare flower count data ----
-flower_count = read_csv("Data/Raw_data/16_17_Manincor/Flower_count2.csv", col_names = T)
-
-#Delete extra column that indicated that this dataset was number one
-flower_count = flower_count %>% 
-  select(!c(Comment...8, Comment...9))
-
-#Split data into different dataframes based on survey name
-FlowerCount <- split(flower_count, flower_count$Site_id)
-
-#Prepare metadata data ----
-#Select unique cases of plants and poll
-plant_single_cases = data %>% distinct(Plant_species)
-pollinator_single_cases = data %>%distinct(Pollinator_species)
-
-
-Metadata <- tibble(
-Doi = "10.1016/j.actao.2020.103551",
-Dataset_description = "These datasets document 3 bee-plant networks observed in 3
-calcareaous grasslands in France during the month of July 2016.
-The current data only present the observed interactions (visit-based network
-in the mentioned paper published in Acta Oecologica). These data are part of the
-ANR ARSENIC project (grant no. 14-CE02-0012).",
-Taxa_recorded = "Only female bee visitors have been used for these datasets",
-Sampling_year = "2016",
-Country = "France",
-Habitat = "Calcareous grasslands",
-Sampling_sites = 3,
-Sampling_rounds = "3 days",
-Sampling_method = "Random walk",
-Sampling_area_details = "1 hectare per site",
-Sampling_area_species_m2 = NA,
-Sampling_area_total_m2 = "30000",
[…]

### 2023-09-25 Scripts/Processing/21_Hopfenmuller.R

@@ -1,110 +1,0 @@
-#DATASET NUMBER 21; Bernhard Hopfenmuller 
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(tidyverse)
-library(lubridate)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/21_Hopfenmuller/Interaction_data.csv")
-
-#Compare vars
-compare_variables(check_interaction_data, data)
-
-#Data
-data = data %>% 
-mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
-mutate(Year = year(ymd(Date))) %>% 
-mutate(Month = month(ymd(Date))) %>% 
-mutate(Day = day(ymd(Date))) 
-
-#Add interaction col
-data = data %>% 
-mutate(Interaction = 1) %>% 
-mutate(Sampling_method = "Variable transect walks") %>% 
-mutate(Sampling_effort_minutes = NA) %>% 
-mutate(Sampling_area_square_meters = NA) %>% 
-mutate(Habitat = "Calcareous grassland") %>% 
-mutate(Locality = Site_id) 
-
-#Add missing vars
-data = add_missing_variables(check_interaction_data, data) 
-
-#Reorder variables
-data = drop_variables(check_interaction_data, data) 
-
-#Drop this last two columns that are going to be at the metadata
-data = data %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
-
-#Split interaction data into dataframes within a list
-InteractionData <- split(data,data$Site_id)
-
-
-#Prepare flower count data ----
-flower_count <- read_csv("Data/Raw_data/21_Hopfenmuller/Flower_count.csv")
-
-#Compare vars
-compare_variables(check_flower_count_data, flower_count)
-
-#Fix names and variable
-flower_count = flower_count %>% 
-rename(Flower_count = Flower_units) %>% 
-mutate(Units = "Flower number")
-
-#Add variables
-flower_count = add_missing_variables(check_flower_count_data, flower_count) 
-
-#Order data as template
-flower_count = drop_variables(check_flower_count_data, flower_count) 
-
-#Split interaction data into dataframes within a list
-FlowerCount <- split(flower_count, flower_count$Site_id)
-
-#Prepare metadata data ----
-#Select unique cases of plants and poll
[…]

### 2023-09-25 Scripts/Processing/24_Holzschuh.R

@@ -1,122 +1,0 @@
-#DATASETS NUMBER (23)-24-(25); Holzschuh
-#To keep things clear here I have separated this in 3 different scripts (dataset 24 here)
-#I have separated this dataset by habitat. This one compiles Mass flowering crop
-
-#Read source to compare with templates later on
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(tidyverse)
-library(lubridate)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/23_24_25_Holzschuh/Interaction_data.csv")
-
-#Compare vars
-compare_variables(check_interaction_data, data)
-
-#Add interaction col
-data = data %>% 
-mutate(Interaction = 1) %>% 
-mutate(Sampling_method = "Transect") %>% 
-mutate(Sampling_effort_minutes = NA) %>% 
-mutate(Sampling_area_square_meters = NA) %>% 
-mutate(Habitat = recode_factor(Habitat, "FB" = "Field boundaries", 
-      "MFC" = "Mass flowering crop", "SNH" = "Seminatural")) %>% 
-mutate(Comment = "Location col indicates transects rounds (1 and 2)")
-
-#Add missing vars
-data = add_missing_variables(check_interaction_data, data) 
-#Reorder variables
-data = drop_variables(check_interaction_data, data) 
-#Drop this last two columns that are going to be at the metadata
-data = data %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
-
-#Now I split data into different dataframes based on habitat
-split_intdata <- split(data, data$Habitat)
-#Convert to tibbles
-data2 <- as_tibble(split_intdata[[2]])
-
-#Split interaction data into dataframes within a list (separate now by site)
-InteractionData2 <- split(data2, data2$Site_id)
-
-#Prepare flower count data ----
-flower_count <- read_csv("Data/Raw_data/23_24_25_Holzschuh/Flower_count.csv")
-
-#Compare vars
-compare_variables(check_flower_count_data, flower_count)
-
-#Split date col
-flower_count = flower_count %>% 
-mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
-mutate(Year = year(ymd(Date))) %>% 
-mutate(Month = month(ymd(Date))) %>% 
[…]

### 2023-09-25 Scripts/Processing/25_Holzschuh.R

@@ -1,122 +1,0 @@
-#DATASETS NUMBER (23-24)-25; Holzschuh
-#To keep things clear here I have separated this in 3 different scripts (dataset 25 here)
-#I have separated this dataset by habitat. This one compiles seminatural habitat
-
-#Read source to compare with template later on
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(tidyverse)
-library(lubridate)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/23_24_25_Holzschuh/Interaction_data.csv")
-
-#Compare vars
-compare_variables(check_interaction_data, data)
-
-#Add interaction col
-data = data %>% 
-mutate(Interaction = 1) %>% 
-mutate(Sampling_method = "Transect") %>% 
-mutate(Sampling_effort_minutes = NA) %>% 
-mutate(Sampling_area_square_meters = NA) %>% 
-mutate(Habitat = recode_factor(Habitat, "FB" = "Field boundaries", 
-      "MFC" = "Mass flowering crop", "SNH" = "Seminatural")) %>% 
-mutate(Comment = "Location col indicates transects rounds (1 and 2)")
-
-#Add missing vars
-data = add_missing_variables(check_interaction_data, data) 
-#Reorder variables
-data = drop_variables(check_interaction_data, data) 
-#Drop this last two columns that are going to be at the metadata
-data = data %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
-
-#Now I split data into different dataframes based on habitat
-split_intdata <- split(data, data$Habitat)
-#Convert to tibbles
-data3 <- as_tibble(split_intdata[[3]])
-
-#Split interaction data into dataframes within a list (separate now by site)
-InteractionData3 <- split(data3, data3$Site_id)
-
-#Prepare flower count data ----
-flower_count <- read_csv("Data/Raw_data/23_24_25_Holzschuh/Flower_count.csv")
-
-#Compare vars
-compare_variables(check_flower_count_data, flower_count)
-
-#Split date col
-flower_count = flower_count %>% 
-mutate(Date = as.Date(Date, format="%m-%d-%y")) %>% 
-mutate(Year = year(ymd(Date))) %>% 
-mutate(Month = month(ymd(Date))) %>% 
[…]

### 2023-09-25 Scripts/Processing/28_Sutter.R

@@ -1,287 +1,0 @@
-#DATASET NUMBER 28; Sutter 
-#There are some unmatching plant names
-
-
-#Read templates to compare with
-source("Scripts/Empty_templates.R") #Read empty templates to compare with
-
-#Load libraries
-library(dplyr)
-library(stringr)
-library(readr)
-library(tibble)
-library(tidyr)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/28_Sutter/Interaction_data.csv")
-
-#Filter interactions greater than 0
-#And filter out insects caught flying
-data = data %>% 
-filter(Interaction > 0) %>% 
-filter(!Plant_species_corrected == "transect")
-
-#Check col names with template
-compare_variables(check_interaction_data, data)
-
-#Rename variables accordingly
-data = data %>% 
-rename(Plant_species = Plant_species_corrected) %>% 
-rename(Habitat = Focal_habitat) %>% 
-mutate(Habitat =  as.factor(Habitat))  %>% 
-mutate(Habitat = recode_factor(Habitat, "HA" = "Herbaceous areal 1",
-                              "HA2" = "Herbaceous areal 2",
-                              "HA3" = "Herbaceous areal 3",
-                              "HL" = "Herbaceous linear",
-                              "HL2" = "Herbaceous linear 2",
-                              "WA" = "Woody areal",
-                              "WL" = "Woody linear",
-                              "WL2" = "Woody linear 2")) %>% 
-mutate(Country = "Switzerland")
-
-#Add missing variables
-data = add_missing_variables(check_interaction_data, data)
-
-#Plant species are just codes at the moment
-#Load flower count data which has the correct species names 
-plant_names <- read_csv("Data/Raw_data/28_Sutter/Flower_count.csv") %>% 
-select(Plant_species, Species_abbreviation) %>% 
-rename(Plant_names_new = Plant_species, Plant_species = Species_abbreviation)
-
-#Recode some factors
-plant_names = plant_names %>% 
-mutate(Plant_species = case_when(
-Plant_names_new == "Taraxacum.campylodes" ~ "tar.off", 
-T ~ Plant_species))
-#Recode Silene
[…]

### 2023-09-25 Scripts/Processing/34_Stanley.R

@@ -1,163 +1,0 @@
-#DATASET NUMBER 34; Stanley
-#ALL datasets were gathered by Laura Russo who organised this
-#(There are 6 in total)
-#One is an unpublished thesis
-
-#Read empty templates to compare with
-source("Scripts/Empty_templates.R") 
-
-#Load libraries
-library(dplyr)
-library(stringr)
-library(readr)
-library(tibble)
-library(tidyr)
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
-
-#Divide col into cols
-levels(factor(data$Site_id))
-data = data %>% separate(Site_id, c("Name", "Site", "Site_number"), remove = F)
-levels(factor(data$Name))
-
-#Recode authors to just their surname
-data = data %>% 
-mutate(Name = recode_factor(Name,  "AoifeO" = "O’Rourke",
-                            "EileenPower" = "Power",
-                            "DaraStanley" = "Stanley",
-                            "SarahMullen" = "Mullen",
-                            "MichelleLarkin" = "Larkin",
-                            "CianWhite" = "White"))
-
-#Select first the one of Power
-data = data %>% filter(Name == "Stanley")
-
-#To calculate sampling effort later on
-data_time = data %>% select(Sampling_effort_minutes)
-
-#Compare vars
-#compare_variables(check_interaction_data, data)
-
-#Prepare data
-data = data %>% 
-rename(Plant_species = Plant_Species) %>% 
-mutate(Habitat = "Farming grassland") %>% 
-mutate(Locality = "Farms in south-east Ireland") %>% 
-rename(Latitude = "WGS84/ETRS89") %>% 
-rename(Longitude = "WGS84/ETRS89_2") %>% 
-mutate(Site_id = Site) %>% 
-mutate(Sampling_method = "Transects") 
-
-#Add missing vars
-data = add_missing_variables(check_interaction_data, data) 
-#Reorder variables
-data = drop_variables(check_interaction_data, data) 
-
-
-#Convert degree minute second to decimal degree coordinates
-library(parzer)
-
-#Prepare data before convert coordinates
-data = data %>% 
-mutate(Latitude = gsub("\\°.", "° ", Latitude)) %>% 
-mutate(Latitude = gsub("\\'.", "' ", Latitude)) %>% 
-mutate(Longitude = gsub("\\°.", "° ", Longitude)) %>% 
[…]

### 2023-09-25 Scripts/Processing/5_Marini.R

@@ -1,103 +1,0 @@
-#DATASET NUMBER 5; Cappellari & Marini
-
-#Load libraries
-library(tidyverse)
-
-#Prepare interaction data ----
-data = read.csv("Data/Raw_data/4_5_6_Marini/Interaction_data.csv")
-
-#Delete all underscores and space in one column
-data = data %>% 
-mutate(across(everything(), function(x) str_replace_all(x,"_", " "))) %>% 
-mutate(Coordinate_precision = str_replace(Coordinate_precision, " ", "")) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
-mutate(Latitude = as.numeric(Latitude)) %>% 
-mutate(Longitude = as.numeric(Longitude)) 
-
-#Split data into different dataframes based on survey name
-split_intdata <- split(data, data$Survey)
-#Convert to tibbles
-data2 <- as_tibble(split_intdata[[2]])
-#Now create a list of the different networks (unique Site_id) for each survey
-InteractionData2 <- split(data2, data2$Site_id)
-
-#Prepare flower count data ----
-flower_count = read.csv("Data/Raw_data/4_5_6_Marini/Flower_count.csv")
-
-#Delete all underscores
-FlowerCount = flower_count %>% 
-mutate(across(everything(), function(x) str_replace_all(x, "_", " "))) %>% 
-select(!c(Total_flower_cover, Collected_insects))
-
-#Split data into different dataframes based on survey name
-split_flwdata <- split(FlowerCount, FlowerCount$Survey)
-#Convert to tibbles
-FlowerCount2 <- as_tibble(split_flwdata[[2]]) %>%  select(!Survey)
-#Split by Site_id
-FlowerCount2 = split(FlowerCount2, FlowerCount2$Site_id)
-
-#Prepare metadata data ----
-
-#Plants sampled in each site
-for (i in InteractionData2) {
-#Generate sum of distinct plants per site
-plant_sum2 <- bind_rows(lapply(InteractionData2, function(x) x %>% 
-                      select(Plant_species, Site_id) %>% 
-                      group_by(Site_id) %>% 
-                      summarise(Sum = n_distinct(Plant_species))))
-
-}
-#Plant sum should be the total number of plants sampled
-plant_sum2 = sum(plant_sum2$Sum)
-
-
-#Store unique cases of plants and polls
-plant_single_cases2 = data2 %>% distinct(Plant_species)
[…]

### 2023-09-29 Scripts/1_Visualisation/Map_study_locations.Rmd
cleaning metaweb script

@@
 # Get all countries and transform to the same CRS
-cntries <- gisco_get_countries(year = year_ref,
-        resolution = 03) %>%
+cntries <- gisco_get_countries(region = "Europe", year = "2016") %>% 
           st_transform(3035)
+
+cntries = sf::st_cast(cntries, "MULTIPOLYGON")
 
@@
 "blue", "black", "green",
-"purple", "red")
+"purple", "red", "gray")
 #To generate more random colors 

### 2023-09-29 Scripts/1_Processing_networks/11_Clough.R

@@ -2,6 +2,8 @@
 
 #Load libraries
 library(tidyverse)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
 
 #Prepare interaction data ----
 data = read_csv("Data/Raw_data/11_Clough/Interaction_data.csv") %>% 
@@ -13,6 +15,9 @@ select(Plant_species, Pollinator_species, Interaction, Sampling_method,
 select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>%  #Including this info in the metadata
 filter(!is.na(Plant_species)) %>% 
 filter(!is.na(Pollinator_species)) 
+
+#Unify structure of data
+data = change_str(data)
 
 #Convert coordinates to lat/lon
 library(rgdal)

### 2023-09-29 Scripts/1_Processing_networks/17_Manincor.R

@@ -2,6 +2,8 @@
 
 #Load libraries
 library(tidyverse)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
 
 #Prepare interaction data ----
 #Load interaction data
@@ -13,6 +15,8 @@ data = data %>%
          Latitude = parzer::parse_lat(Latitude)) %>% 
 select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
 
+#Unify structure of data
+data = change_str(data)
 
 unique(factor(data$Latitude))
 unique(factor(data$Longitude))

### 2023-09-29 Scripts/1_Processing_networks/27_Castro.R

@@ -6,6 +6,8 @@ source("Scripts/Empty_templates.R") #Read empty templates to compare with
 #Load libraries
 library(tidyverse)
 library(stringr)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
 
 #Prepare interaction data ----
 data <- read_csv("Data/Raw_data/26_27_Castro/Interaction_data2.csv")
@@ -16,6 +18,9 @@ compare_variables(check_interaction_data, data)
 #Select main cols
 data = data %>% 
 select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this info in the metadata
+
+#Unify structure of data
+data = change_str(data)
 
 #Split interaction data into dataframes within a list (separate now by site)
 InteractionData2 <- split(data, data$Habitat)

### 2023-09-29 Scripts/1_Processing_networks/33_Power.R

@@ -12,6 +12,8 @@ library(stringr)
 library(readr)
 library(tibble)
 library(tidyr)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
 
 #Prepare interaction data ----
 data <- read_csv("Data/Raw_data/32_to_37_Russo/Interaction_data.csv")
@@ -81,6 +83,9 @@ data = data %>%
 mutate(Day = as.Date(Day, origin = "2009-01-01")) %>% 
 mutate(Day = format(as.Date(Day,format="%Y-%m-%d"), format = "%d"))
 #it seems ok
+
+#Unify structure of data
+data = change_str(data)
 
 #Split interaction data into dataframes within a list
 InteractionData <- split(data, data$Site_id)

### 2023-09-29 Scripts/1_Processing_networks/39_Umons.R

@@ -8,16 +8,23 @@ library(stringr)
 library(readr)
 library(tibble)
 library(tidyr)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
 
 #Prepare interaction data ----
 data <- read_csv("Data/Raw_data/39_UMONS/Network_data.csv")
 
-#There are some sites without coordinates: I haven't tried to work that out
-
 #Filter interactions equal to 0
 data = data %>% filter(Interaction > 0)
 
-#Compare vars
+#Filter out any row that contains alphatic letter in latitude
+#This can be recovered but exclude it for now
+data = data %>%
+  filter(!grepl("^[A-Za-z]", Latitude)) %>% 
+    filter(!grepl("^[A-Za-z]", Longitude))
+
+#Unify structure of data
+data = change_str(data)
 
 #Split interaction data into dataframes within a list
 InteractionData <- split(data, data$Site_id)
@@ -59,5 +66,6 @@ UMONS <- list(InteractionData,  Metadata, Authorship)
 #Rename list elements
 names(UMONS) <- c("InteractionData","Metadata", "Authorship")
 #Save data
-saveRDS(UMONS, file="Data/Clean_data/39_Umons.rds")
+#I comment it for now as thee data needs more work
+#saveRDS(UMONS, file="Data/Clean_data/39_Umons.rds")
 

### 2023-09-29 Scripts/1_Processing_networks/4_Marini.R

@@ -1,7 +1,11 @@
 #DATASET NUMBER 4; Cappellari & Marini
+#It has 3 datasets that are prepared in 3 different scripts 
+#Those are: 4_Marini, 5_Marini, 6_Marini
 
 #Load libraries
 library(tidyverse)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
 
 #Prepare interaction data ----
 data = read.csv("Data/Raw_data/4_5_6_Marini/Interaction_data.csv")
@@ -10,12 +14,14 @@ data = read.csv("Data/Raw_data/4_5_6_Marini/Interaction_data.csv")
 data = data %>% 
 mutate(across(everything(), function(x) str_replace_all(x,"_", " "))) %>% 
 mutate(Coordinate_precision = str_replace(Coordinate_precision, " ", "")) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
-mutate(Latitude = as.numeric(Latitude)) %>% 
-mutate(Longitude = as.numeric(Longitude)) 
+mutate(Comments = Survey) %>%   
+select(!c(Survey, Sampling_effort_minutes, Sampling_area_square_meters)) 
+
+#Unify structure of data
+data = change_str(data)
 
 #Split data into different dataframes based on survey name
-split_intdata <- split(data, data$Survey)
+split_intdata <- split(data, data$Comments)
 #Convert to tibbles
 data1 <- as_tibble(split_intdata[[1]])
 
@@ -93,124 +99,11 @@ Metadata1 = as.data.frame(t(Metadata1)) %>%
   rename(Metadata_fields = rowname, Metadata_info= V1) %>% as_tibble()
 
 
-#Plants sampled in each site
-for (i in InteractionData2) {
-  #Generate sum of distinct plants per site
-  plant_sum2 <- bind_rows(lapply(InteractionData2, function(x) x %>% 
-                                   select(Plant_species, Site_id) %>% 
-                                   group_by(Site_id) %>% 
-                                   summarise(Sum = n_distinct(Plant_species))))
-  
-}
-#Plant sum should be the total number of plants sampled
-plant_sum2 = sum(plant_sum2$Sum)
-
-
-#Store unique cases of plants and polls
-plant_single_cases2 = data2 %>% distinct(Plant_species)
-pollinator_single_cases2 = data2 %>%distinct(Pollinator_species)
-
-Metadata2 <- tibble(
-  Doi = "https://doi.org/10.1007/s00442-022-05151-6",
[…]

### 2023-09-29 Scripts/1_Processing_networks/6_Marini.R

@@ -2,6 +2,8 @@
 
 #Load libraries
 library(tidyverse)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
 
 #Prepare interaction data ----
 data = read.csv("Data/Raw_data/4_5_6_Marini/Interaction_data.csv")
@@ -10,13 +12,16 @@ data = read.csv("Data/Raw_data/4_5_6_Marini/Interaction_data.csv")
 data = data %>% 
 mutate(across(everything(), function(x) str_replace_all(x,"_", " "))) %>% 
 mutate(Coordinate_precision = str_replace(Coordinate_precision, " ", "")) %>% 
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
+mutate(Comments = Survey) %>%   
+select(!c(Survey, Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
 mutate(Latitude = as.numeric(Latitude)) %>% 
 mutate(Longitude = as.numeric(Longitude)) 
 
+#Unify structure of data
+data = change_str(data)
 
 #Split data into different dataframes based on survey name
-split_intdata <- split(data, data$Survey)
+split_intdata <- split(data, data$Comments)
 #Convert to tibbles
 data3 <- as_tibble(split_intdata[[3]])
 #Now create a list of the different networks (unique Site_id) for each survey

### 2023-10-15 Scripts/1_Visualisation/Map_study_locations.Rmd
preparing taxonomic info of polls

@@
 "blue", "black", "green",
-"purple", "red", "gray")
+"purple", "red", "gray","violetred",
+"coral4", "plum1")
 #To generate more random colors 

### 2023-10-15 Scripts/1_Processing_networks/39_Schweiger.R

@@ -1,0 +1,117 @@
+#DATASET NUMBER 39; Schweiger
+
+#Data sent by Oliver Schweiger. This dataset belongs to the VOODOO project
+#https://voodoo-project.eu/about
+
+#Read empty templates to compare with
+source("Scripts/Empty_templates.R") 
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/39_Schweiger/Interaction_data.csv")
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+
+#Because we only consider pollinators that visited flowers
+#We filter out the ones that do not have listed a plant species  
+data = data %>% 
+filter(Plant_species != "no flower") %>% 
+filter(Plant_species != "missing")
+
+#Exclude 5 rows with missing coordinates for now
+data = data %>% 
+filter(!is.na(Latitude))
+
+#Drop sampling effort and square area (it would be added on the metadata)
+data = data %>%
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters))
+
+#Unify structure of data
+data = change_str(data)
+
+#Split interaction data into dataframes within a list
+InteractionData <- split(data, data$Site_id)
+
+#Prepare flower count data ---- 
+flower_count <- read_csv("Data/Raw_data/39_Schweiger/Flower_count.csv")
+
+#Check vars
+#compare_variables(check_flower_count_data, flower_count)
+
+#Split interaction data into dataframes within a list
+FlowerCount <- split(flower_count, flower_count$Site_id)
+
+#Prepare metadata data ----
+#Select unique cases of plants and poll
+plant_single_cases = data %>% distinct(Plant_species)
+pollinator_single_cases = data %>%distinct(Pollinator_species)
+
+#Build metadata
+Metadata <- tibble(
+  Doi = "Unpublished",
+  Dataset_description = "This dataset was generated within the VOODOO project
+  and documents 16 different landscapes in agricultural and urban areas of Central Germany.
+  We followed a landscape-level approach, where a site is defined by a 1km radius.
+  Within each site we placed transects on a stratified random basis in each of present habitat types.
+  The total length of the transect was 1km per site, but split according to the relative proportions of the different habitat types.
[…]

### 2023-10-15 Scripts/1_Processing_networks/40_Schweiger.R

@@ -1,117 +1,0 @@
-#DATASET NUMBER 40; Schweiger
-
-#Data sent by Oliver Schweiger. This dataset belongs to the VOODOO project
-#https://voodoo-project.eu/about
-
-#Read empty templates to compare with
-source("Scripts/Empty_templates.R") 
-
-#Load libraries
-library(dplyr)
-library(stringr)
-library(readr)
-library(tibble)
-library(tidyr)
-#Load function to unify structure of data
-source("Scripts/Change_str.R")
-
-#Prepare interaction data ----
-data <- read_csv("Data/Raw_data/40_Schweiger/Interaction_data.csv")
-
-#Compare vars
-#compare_variables(check_interaction_data, data)
-
-#Because we only consider pollinators that visited flowers
-#We filter out the ones that do not have listed a plant species  
-data = data %>% 
-filter(Plant_species != "no flower") %>% 
-filter(Plant_species != "missing")
-
-#Exclude 5 rows with missing coordinates for now
-data = data %>% 
-filter(!is.na(Latitude))
-
-#Drop sampling effort and square area (it would be added on the metadata)
-data = data %>%
-select(!c(Sampling_effort_minutes, Sampling_area_square_meters))
-
-#Unify structure of data
-data = change_str(data)
-
-#Split interaction data into dataframes within a list
-InteractionData <- split(data, data$Site_id)
-
-#Prepare flower count data ---- 
-flower_count <- read_csv("Data/Raw_data/40_Schweiger/Flower_count.csv")
-
-#Check vars
-#compare_variables(check_flower_count_data, flower_count)
-
-#Split interaction data into dataframes within a list
-FlowerCount <- split(flower_count, flower_count$Site_id)
-
-#Prepare metadata data ----
-#Select unique cases of plants and poll
-plant_single_cases = data %>% distinct(Plant_species)
-pollinator_single_cases = data %>%distinct(Pollinator_species)
-
-#Build metadata
-Metadata <- tibble(
-  Doi = "Unpublished",
-  Dataset_description = "This dataset was generated within the VOODOO project
-  and documents 16 different landscapes in agricultural and urban areas of Central Germany.
-  We followed a landscape-level approach, where a site is defined by a 1km radius.
-  Within each site we placed transects on a stratified random basis in each of present habitat types.
-  The total length of the transect was 1km per site, but split according to the relative proportions of the different habitat types.
[…]

### 2023-10-15 Scripts/2_Processing_metaweb/1_Bind_networks.R

@@ -1,0 +1,48 @@
+########################################################################
+#1) Read all networks and create a single dataset
+########################################################################
+
+#Load libraries
+library(dplyr) 
+library(stringr)
+
+#Read all files with their paths
+files <- dir("Data/Clean_data", pattern="*.rds", full.names=T)
+files <- mixedsort(files)
+
+#Generate list of lists
+all_data = files %>%
+map(readRDS) 
+
+#Read all file names
+file_names <- dir("Data/Clean_data", pattern="*.rds", full.names=F)
+file_names <- mixedsort(file_names)
+
+#Delete extension
+file_names = str_replace(file_names, ".rds", "")
+#Rename all elements of the list
+names(all_data) <- file_names
+
+#Select all cols from the list of lists----
+data <- list()
+
+for (i in 1:length(file_names)) {
+  
+  int_list <- all_data[[file_names[i]]]$InteractionData
+  data[[i]] <- bind_rows(map(int_list, function(x) x %>% #map here is the same as a lapply
+                dplyr::select(everything())),  .id = 'Network_id')
+}
+
+#Rename files from list  
+names(data) <- file_names
+#Prepare data to save it
+all_list = data
+all = bind_rows(data,  .id = 'Study_id')
+#Check country levels
+levels(factor(all$Country))
+#Create another col with study id and network id together so we can count all individuals networks
+master = all %>% 
+mutate(Study_network_id = paste0(Study_id, Network_id))
+
+#Save master file with all networks
+saveRDS(master, "Data/Processing/Building_metaweb.rds")

### 2023-10-15 Scripts/2_Processing_metaweb/2_Clean_pollinator_taxonomy.R

@@ -1,0 +1,602 @@
+########################################################################
+#2) Clean pollinator taxonomy----
+########################################################################
+
+#1)Use fuzzy matching to recover species names from Thesaurus master list
+#2)Conduct manual fixes from the non-recovered records
+
+#Load libraries
+library(dplyr) 
+library(stringr)
+library(purrr)
+library(gtools)
+library(reshape2) #to use melt function later on
+
+#Read raw metaweb
+master = readRDS("Data/Processing/Building_metaweb.rds")
+#Read master list of unique cases of pollinator species
+species_tesaurus <- readRDS(file = "Data/Species_thesaurus/taxonomy.rds")
+
+#Explore poll data
+#Check total number of distinct poll species names
+n_distinct(master$Pollinator_species)
+#Check levels
+levels(factor(master$Pollinator_species))
+
+########################################################################
+#1)Use fuzzy matching to recover species names from Thesaurus master list
+########################################################################
+#Create unique list of pollinator species
+Gen_sp = master %>% 
+select(Pollinator_species) %>% 
+distinct() %>% 
+pull()
+#Generate col with genus and species names from Thesaurus
+species_tesaurus = species_tesaurus %>% 
+mutate(Gen_sp = paste(trimws(species_tesaurus$Genus),
+                trimws(species_tesaurus$Species)))
+
+#Select the matching and the unmatching records
+matching <- Gen_sp[which(Gen_sp %in% species_tesaurus$Gen_sp)]
+unmatching <- Gen_sp[which(!Gen_sp %in% species_tesaurus$Gen_sp)]
+mismatches <- unique(unmatching) #speed up the process (just in case of duplicates)
+#As this takes long 
+#let's exclude the unidentified records to speed up more the process
+#There are few records that get lost that start with sp
+#Minor issue but keep in mind
[…]

### 2023-10-26 Scripts/1_Visualisation/Map_study_locations.Rmd
New datasets added and pollinator taxonomy updated

--- 2023-09-29 Scripts/1_Visualisation/Map_study_locations.Rmd
@@
 "blue", "black", "green",
-"purple", "red", "gray")
+"purple", "red", "gray","violetred",
+"coral4", "plum1", "yellow",
+"white")
 #To generate more random colors 

### 2023-10-26 Scripts/1_Processing_networks/44_Knight.R

@@ -1,0 +1,97 @@
+#DATASET NUMBER 44; Knight-Romania
+#Dataset sent by Tiffany
+
+source("Scripts/Empty_templates.R") 
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
+
+#Load function to unify structure of data
+source("Scripts/Change_str.R")
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/44_Knight/Interaction_data.csv", locale = locale(encoding = "latin1"))
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+#Nothing missing!
+
+#Drop sampling effort and square area (it would be added on the metadata)
+data = data %>%
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters))
+
+#Split interaction data into dataframes within a list
+InteractionData <- split(data, data$Site_id)
+
+#Prepare flower count data ---- 
+flower_count <- read_csv("Data/Raw_data/44_Knight/Flower_count.csv")
+
+#Check vars
+#compare_variables(check_flower_count_data, flower_count)
+#No misisng vars
+
+#Split interaction data into dataframes within a list
+FlowerCount <- split(flower_count, flower_count$Site_id)
+
+#Prepare metadata data ----
+
+#Store unique cases of plants and polls
+plant_single_cases = data %>% distinct(Plant_species)
+pollinator_single_cases = data %>%distinct(Pollinator_species)
+
+#Build metadata
+Metadata <- tibble(
+Doi = NA,
+Dataset_description = "We sample plants, pollinators and
+their interactions in 30 by 2 meter transects across 18
+sites in translyvania. We choose sites that were all
+semi-natural grasslands that varied in soil type, moisture
+(wet, mesophilic and dry meadows were considered) and
+elevation",
+Taxa_recorded = "Diptera, Hymenoptera, Lepidoptera",
+Sampling_year = "2018",
+Country = "Romania",
+Habitat = "Grassland",
+Sampling_sites = "18",
+Sampling_rounds = "",
+Sampling_method = "Transect",
+Sampling_area_details = "30 by 2 meter transects",
+Sampling_area_species_m2 = NA,
+Sampling_area_total_m2 = NA ,
[…]

### 2023-10-26 Scripts/2_Processing_metaweb/2_1_Clean_pollinator_taxonomy.R

@@ -541,6 +541,9 @@ mutate(Fixed = case_when(
   str_detect(Mismatch, "Vermilionidae") ~ "Vermileonidae",
   str_detect(Mismatch, "Tenthridae") ~ "Tenthredinidae",
   str_detect(Fixed, "Hoplitis scutellaris") ~ "Osmia scutellaris",
+  str_detect(Mismatch, "Lycaena corydon") ~ "Lysandra hispana",
+  Mismatch == "Anthidium scapulare" ~ "Pseudoanthidium scapulare",
+  Mismatch == "Eoseristalis lineata" ~ "Cheilosia morio",
 
   T ~ Fixed)) %>% 
   rename(Old_name = Mismatch, Name = Fixed) 
@@ -555,20 +558,19 @@ mutate(Name = case_when(
   Old_name == "Microgasterinae" ~ "Braconidae",
   T ~ Name))
 
-#Divide dataset in fixed and no fixed names
-to_recover1 =  to_recover %>% filter(!is.na(Name))
-to_recover2 = to_recover %>% filter(is.na(Name)) %>% 
+#Add uncertainty cols (default=No)
+to_recover1 = to_recover %>% 
 mutate(Unsure_id = "No") %>% 
-mutate(Uncertainty_type = NA)
-  
+mutate(Uncertainty_type = NA)  
+
 #For the fixed ones, add additional info
 #For instance, unsure id. Then add another col where we explain
 #the original status
-unsure_id = c("type","aggr.","agg.","[/]", 
+unsure_id = c(" type","aggr.","agg.","[/]", 
               "aff(?![:alpha:])", " cf", " Cf", "[?]",
               "_agg", "_group", "-group",
               "_c.f", "malachurum-alike",
-              "-.*-", " agg$")
+              "-.*-", " agg$", " complex")
 
 to_recover1 = to_recover1 %>% 
 mutate(Unsure_id = case_when(
@@ -588,9 +590,7 @@ mutate(Unsure_id = case_when(
  Old_name == "Lasioglossum sp_immunitum-alike" ~ "Yes",
  Old_name == "Halictus gr simplex" ~ "Yes",
  Old_name == "Lasioglossum immunitum-type" ~ "Yes",
- Old_name == "Bombus terrestris complex" ~ "Yes",
  Old_name == "Chrysosomoxys macrocercus" ~ "Yes",
-
  T ~Unsure_id
 
 ))
[…]

### 2023-10-26 Scripts/2_Processing_metaweb/2_Pollinator_processing/1_Build_pollinator_masterlist.R

@@ -1,0 +1,89 @@
+#Create master taxonomic and distribution files:
+
+#Load libraries
+library(reshape2)
+
+#Bees and syrphids----
+#Load data
+polBS <- read.csv(file = "Data/Species_thesaurus/Master_bees_syrphids.csv")
+head(polBS)
+#extract taxonomy
+taxonomy <- polBS[,1:9]
+#extract distribution
+colnames(polBS) #countries to match.
+#write.csv(colnames(polBS), "scripts/temp_data/country_tesaurus_init.csv")
+#This file form the bases to create manually country_tesaurus.csv
+#Make long format
+distribution1 <- melt(data = polBS, id.vars = "GenusAndSpecies", measure.vars = 10:ncol(polBS), 
+      value.name = "status", variable.name = "country")
+head(distribution1)
+#We remove absences
+distribution2 <- subset(distribution1, status != "A")
+head(distribution2)
+#we unify country names according to tesaurus
+tesaurus <- read.csv("Data/Species_thesaurus/country_tesaurus.csv")
+head(tesaurus)
+distribution <- merge(distribution2, tesaurus[,2:3], by.x = "country", by.y = "Bees_syrphids")
+head(distribution)
+distribution <- distribution[,-1]
+head(distribution)
+
+#Buterflies----
+#taxonomy
+taxBt <- read.delim(file = "Data/Species_thesaurus/Butterfly_taxon.txt",
+                    header = T)
+head(taxBt) #can be merged by id
+#subset species as taxonRank
+unique(taxBt$taxonRank) #unecessary, but alas...
+taxBt <- subset(taxBt, taxonRank == "species")
+taxonomyBT <-  taxBt[, c("order", "family", "genus",
+                         "specificEpithet", "scientificName", 
+                         "scientificNameAuthorship")] 
+#merge with taxonomy
+head(taxonomy)
+taxonomyBT$Subfamily <- NA
+taxonomyBT$Tribe <- NA
+taxonomyBT$Subgenus <- NA
+taxonomyBT <-  taxonomyBT[, c("order", "family", "Subfamily", "Tribe", "genus",
+                              "Subgenus", "specificEpithet", "scientificName", 
+                              "scientificNameAuthorship")] 
+colnames(taxonomyBT) <- colnames(taxonomy)
+taxonomy <- rbind(taxonomy, taxonomyBT)
+head(taxonomy); tail(taxonomy)
+
+#distribution
[…]

### 2023-10-26 Scripts/2_Processing_metaweb/Old_prep_data.R

@@ -1,0 +1,465 @@
+# ---- EU METAWEB PREPARATION PROCESS ----
+# A master file is created combining SafeNet (JBL data compilation) and other datasets.
+# Overview of datasets included (29/sep)----
+    # - SafeNet (40 networks, years?, countries?)
+    # - Umons (Alham, Museum collections and other pairwise interactions, 1850-2021, countries?)
+    # - Globi: https://github.com/Big-Bee-Network/global-bee-interaction-data (Not done)
+    # - Literature: https://github.com/Big-Bee-Network/bee-interaction-database
+
+# Overview of the structure of Plants and Pollinator to be considered-----
+# Pollinators1: Bee + Syrphids + Butterflies (with taxonomy and reference species names)
+# Pollinators2: Bombilids + Beetles + Sawflies (missing taxonomy and reference species names) (not included, 29/sep)
+# Plants: first taxonomic reference obtained, missing synonyms? 
+
+
+# 0 Read SafeNet data----
+#Ideally from the repo once public. Now I made a copy in my personal computer
+
+#Example of one file.
+temp <- readRDS("scripts/temp_data/Clean_data/1_Bartomeus.rds")
+#str(temp)
+#temp$Metadata #Most fields already in interaction data (e.g. where and when)
+#temp$Authorship #Most fields already in interaction data (e.g. where and when)
+#merge all sites
+out<- temp$InteractionData[[1]]
+for(i in 2:length(temp$InteractionData)){
+  temp2 <- temp$InteractionData[[i]]
+  out <- rbind(temp2, out)
+}
+#head(out)
+#dim(out)
+colnames(out)
+
+library(dplyr)
+library(stringr)
+library(data.table)
+library(taxize)
+library(cleanR)
+#loop through all files
+#Read all file names to loop.
+files <- list.files(path = "Data/Clean_data")
+master <- data.frame(Plant_species = NA   ,Pollinator_species = NA   ,Interaction = NA   ,      
+                     Sampling_method = NA ,Site_id = NA              ,Habitat = NA       ,     
+                     Country = NA         ,Locality = NA             ,Latitude = NA      ,      
+                     Longitude = NA       ,Coordinate_precision = NA ,Elevation = NA     ,        
[…]

### 2023-11-01 Scripts/1_Visualisation/Map_study_locations.Rmd
int data updated

--- 2023-09-29 Scripts/1_Visualisation/Map_study_locations.Rmd
@@
 "blue", "black", "green",
-"purple", "red", "gray")
+"purple", "red", "gray","violetred",
+"coral4", "plum1", "yellow",
+"brown3", "orange3", "red4")
 #To generate more random colors 

### 2023-11-01 Scripts/2_Build_metaweb/2_Pollinator_processing/2_2_Rename_unmatched_pollinators.R

@@ -1307,7 +1307,7 @@ mutate(Rank = case_when(
    Fixed_name == "Gymnodia" ~ "GENUS",
    T ~ Rank)) %>% 
 mutate(Matchtype = case_when(
-   Fixed_name == "Gymnodia" ~ "HIGHERRANK",
+   Fixed_name == "Gymnodia" ~ "EXACT",
    T ~ Matchtype))
 
 #Thecophora
@@ -1338,7 +1338,7 @@ mutate(Rank = case_when(
    Fixed_name == "Thecophora" ~ "GENUS",
    T ~ Rank)) %>% 
 mutate(Matchtype = case_when(
-   Fixed_name == "Thecophora" ~ "HIGHERRANK",
+   Fixed_name == "Thecophora" ~ "EXACT",
    T ~ Matchtype))
 
 #Rhinophora
@@ -1369,7 +1369,7 @@ mutate(Rank = case_when(
    Fixed_name == "Rhinophora" ~ "GENUS",
    T ~ Rank)) %>% 
 mutate(Matchtype = case_when(
-   Fixed_name == "Rhinophora" ~ "HIGHERRANK",
+   Fixed_name == "Rhinophora" ~ "EXACT",
    T ~ Matchtype))
 
 #Thecophora
@@ -1400,7 +1400,7 @@ mutate(Rank = case_when(
    Fixed_name == "Thecophora" ~ "GENUS",
    T ~ Rank)) %>% 
 mutate(Matchtype = case_when(
-   Fixed_name == "Thecophora" ~ "HIGHERRANK",
+   Fixed_name == "Thecophora" ~ "EXACT",
    T ~ Matchtype))
 
 #Potamia
@@ -1431,7 +1431,7 @@ mutate(Rank = case_when(
    Fixed_name == "Potamia" ~ "GENUS",
    T ~ Rank)) %>% 
 mutate(Matchtype = case_when(
-   Fixed_name == "Potamia" ~ "HIGHERRANK",
+   Fixed_name == "Potamia" ~ "EXACT",
    T ~ Matchtype))
 
 #Ctenopelmatinae
@@ -1462,5 +1462,68 @@ mutate(Rank = case_when(
    Fixed_name == "Ctenopelmatinae" ~ "SUBFAMILY",
    T ~ Rank)) %>% 
 mutate(Matchtype = case_when(
-   Fixed_name == "Ctenopelmatinae" ~ "HIGHERRANK",
+   Fixed_name == "Ctenopelmatinae" ~ "EXACT",
[…]

### 2023-11-06 Manuscript/Ms.qmd
Minor update
Benadi dataset ready to prepare

@@
 
-# Ideas/objectives:
+# Abstract
 
-- Write article point by point
+Pollinators play a key role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
+However, their diversity and role as pollinators is being impacted by global change (e.g., habitat loss or climate change).
+One of the major drawbacks to understand the conservation status and the underlying ecological processes of plant-pollinator communities is the lack of robust generalisable data across space and time.
+To prevent this limitation and set a baseline for future studies, we provide the first European metaweb of plant-pollinator interactions.
+This continent level metaweb consist of xxxx plant and xxx pollinator species with a total of xxxx interactions covering a total of XXX countries.
 
-- Ecography could be our potential journal (think about it, high fees)
-
-- Finish paper by Christmas
-
-- Create workflow graph
+Other things to do?

### 2023-11-13 Manuscript/Ms.qmd
More file updates

@@
 Pollinators play a key role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-However, their diversity and role as pollinators is being impacted by global change (e.g., habitat loss or climate change).
+However, their diversity and role as pollinators is being compromised by the different pressures of global change (e.g., habitat loss or climate change).
 One of the major drawbacks to understand the conservation status and the underlying ecological processes of plant-pollinator communities is the lack of robust generalisable data across space and time.
-To prevent this limitation and set a baseline for future studies, we provide the first European metaweb of plant-pollinator interactions.
-This continent level metaweb consist of xxxx plant and xxx pollinator species with a total of xxxx interactions covering a total of XXX countries.
+Although recent efforts have progressed towards this goal, Europe, one of the continents with higher number of plant-pollinator studies, still lacks an open access and homogenised dataset plant-pollinator interactions that serves as a baseline for future ecological research.
+Here, we provide and explore the first European metaweb of plant-pollinator interactions.
+This is a continent-level metaweb that comprised of xxx plants xxx pollinators and xxxx interactions which covers xxxx different countries.
+We show that plant and pollinator species are xxxxx and have xxxxxxx.
+For instance, we cover xxx% of European bee species and xxx% of flowering plants that rely on pollinators.
 
-Other things to do?
+# Introduction
+
+# Methods
+
+# Results
+
+# Discussion

### 2023-11-13 Scripts/3_1_Metaweb_taxa_richness.R

@@ -1,120 +1,0 @@
-#Metaweb checks of richness per taxonomic rank
-#For both plant and pollinator species
-
-#Workflow:
-#1)Read master data with accepted taxonomic data
-#2)Check number of spp, genera, family and order 
-#3)Unify and plot
-
-#Load libraris
-library(dplyr)
-library(ggplot2)
-
-#1)READ DATA
-data = readRDS("Data/Interactions_accepted_names.rds")
-#Check colnames
-colnames(data)
-###############################################
-#2)NUMBER OF SPECIES, GENERA, FAMILIES AND ORDERS
-###############################################
-##################
-#Polinator species
-##################
-#Check unique number of pollinator species
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-pollinator_spp_number #check number
-#Make sure that we do not miss any species
-poll_check = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name) %>% 
-filter(Pollinator_rank != "SPECIES") %>% 
-select(Pollinator_accepted_name)  %>% 
-distinct() #seems that we havent miss anything
-#Check unique number of pollinator genera
-pollinator_genus_number = data %>% 
-select(Pollinator_genus) %>% 
-n_distinct()
-pollinator_genus_number
-#Check unique number of pollinator families
-pollinator_families_number = data %>% 
-select(Pollinator_family) %>% 
-n_distinct()
-pollinator_families_number
-#Check unique number of pollinator orders
-pollinator_order_number = data %>% 
-select(Pollinator_order) %>% 
-n_distinct()
-pollinator_order_number
-##################
-#Plant species
-##################
-#Check unique number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-plant_spp_number #check number
-#Make sure that we do not miss any species
-poll_check = data %>% 
-select(Pollinator_rank, Pollinator_status, 
-Pollinator_matchtype,Pollinator_accepted_name) %>% 
-filter(Pollinator_rank != "SPECIES") %>% 
[…]

### 2023-11-13 Scripts/Visualisation/Map_study_locations.Rmd
Organisation

--- 2023-10-15 Scripts/1_Visualisation/Map_study_locations.Rmd
@@
 "purple", "red", "gray","violetred",
-"coral4", "plum1")
+"coral4", "plum1", "yellow",
+"brown3", "orange3", "red4", "darkviolet")
 #To generate more random colors 

### 2023-11-22 Manuscript/Ms.qmd
Updating pdf and scritps

--- 2023-09-23 Manuscript/Ms.qmd
@@
 
-# Ideas/objectives:
+# Abstract
 
-- Write article point by point
+Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
+However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
+One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time that allows to evaluate the consequences of environmental change on their conservation status and
 
-- Ecography could be our potential journal (think about it, high fees)
+conservation status of the different species and their responses to the different environmental changes.
+Although recent efforts have progressed towards this goal, Europe, one of the continents with higher number of plant-pollinator studies, still lacks
 
-- Finish paper by Christmas
+an open access and homogenised dataset of plant-pollinator interactions that serves as a baseline for future ecological research. Here, we provide
 
-- Create workflow graph
+and explore the first European metaweb of plant-pollinator interactions.
+This is a continent-level metaweb that comprised of xxx plants xxx pollinators and xxxx interactions which covers xxxx different countries.
+We show that plant and pollinator species are xxxxx and have xxxxxxx.
+For instance, we cover xxx% of European bee species and xxx% of flowering plants that rely on pollinators.
+
+# Introduction
+
+1st paragraph
+
+General introduction of how global change impacts plant-pollinator interactions
+
+Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions
+
+Highlight the relevance of large scale datasets
+
+2nd paragraph
+
+3rd paragraph
+
+4paragraph
+
+Introduce research questions
+
+LIST THEM HERE (Main ideas so far)
+
+Questions that we would like to answer:
+
+1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity
+
+2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?
+
+# Methods
+
+# Results
+
+# Discussion

### 2023-11-28 Manuscript/Ms.qmd
edits

@@
 However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time that allows to evaluate the consequences of environmental change on their conservation status and
+One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time.
 
-conservation status of the different species and their responses to the different environmental changes.
-Although recent efforts have progressed towards this goal, Europe, one of the continents with higher number of plant-pollinator studies, still lacks
+Here, we have tried to generate a comprehensive dataset of plant-pollination interactions to build the first European metaweb of plant and pollinators and explore their general patterns.
 
-an open access and homogenised dataset of plant-pollinator interactions that serves as a baseline for future ecological research. Here, we provide
+we have done a strong collaborative effort in unifying plant-pollinator studies at continent level to elucidate general patterns of plant-pollinator interactions and set a baseline for future ecological research.
+Thus, we explore the first time an European metaweb of plant-pollinator interactions
 

### 2023-11-28 Scripts/4_1_Add_taxonomy_to_master.R

@@ -2,8 +2,11 @@
 #Read and incorporate plant and pollinator names in master
 #-------------------------------------------------------#
 
+library(dplyr)
+library(tidyr)
+
 #----------------------
-#Read taxonomic tada
+#Read taxonomic data
 #----------------------
 plant_taxo = readRDS("Data/Species_taxonomy/Plant_taxonomy.rds")
 poll_taxo = readRDS("Data/Species_taxonomy/Pollinator_taxonomy.rds")
@@ -33,9 +36,20 @@ colnames(poll_taxo1)
 plant_taxo1
 data = left_join(master1, plant_taxo1)
 data1 = left_join(data, poll_taxo1)
+
+saveRDS(data1, "Data/Interactions.rds")
+
+#Important! Make all rows one single interaction
+#1st uncount interactions
+data2 = data1 %>% 
+uncount(Interaction) %>% 
+mutate(Interaction = 1)
+
 #----------------------
 #Save data
 #----------------------
 #Keep all cols for now
-saveRDS(data1, "Data/Interactions_accepted_names.rds")
+saveRDS(data2, "Data/Interactions_uncounted.rds")
+
+
 

### 2023-12-11 Manuscript/Ms.qmd
map

@@
 
+```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+library(dplyr)
+library(ggplot2)
+library(giscoR)
+library(sf)
+```
 # Abstract
@@
 One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time.
+Here, we present a dataset of plant-pollinator interactions at European level that consists in XXX studies compiled during the years xxx and xxxx.
 
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+#Print number of countries
+data = readRDS("../Data/Interactions_uncounted.rds")
+#Country number
+country_number = data %>% 
+select(Country) %>% 
+n_distinct()
+#Study number
+study_number = data %>% 
+select(Study_id) %>% 
+n_distinct()
+#Minimum year
+min_year = data %>% 
+select(Year) %>% 
+slice(which.min(Year)) %>% 
+pull()
+#Maximum year
+max_year = data %>% 
+select(Year) %>% 
+slice(which.max(Year)) %>% 
+pull()
+#Number of interactions
+interaction_number = nrow(data)
+#Check number of accepted pollinators
+pollinator_spp_number = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name)%>%
+filter(Pollinator_rank == "SPECIES") %>% 
+select(Pollinator_accepted_name) %>% 
+n_distinct()
+#Check number of plant species
+plant_spp_number = data %>% 
+select(Plant_rank, Plant_status, 
+Plant_matchtype, Plant_accepted_name) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+select(Plant_accepted_name) %>% 
+n_distinct()
+
+```
+
+## Dataset description
+
+This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
+This dataset covers `r country_number` countries and consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year`, and accounts for a total of `r prettyNum(interaction_number, big.mark=",")` interactions from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+
+## Species coverage
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%"}
+#Prepare coordinates
+for_plotting = data %>% 
+select(Study_id, Latitude, Longitude) %>% 
+distinct()
+#Get countries and transfrom coordinates
+countries <- gisco_get_countries(
+resolution = 20) %>%
+st_transform(3035) %>% 
+sf::st_as_sf(crs = 4326)
+
+#Number of plant and poll species per network
+#-----------------------#
+#Pollinators------
+#-----------------------#
+#Check unique number of pollinator species
+PollNumber_by_ID = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+select(Pollinator_accepted_name, Study_id) %>% 
+group_by(Study_id) %>% 
+summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
+#Merge now with spp counts by network dataset
+poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
+#Create duplicate cord columns to operate on them
+poll_data$coords <- data.frame(
+Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
+#Set coordinates
+poll_data$coords_sf = st_as_sf(poll_data$coords, 
+  coords = c("Longitude", "Latitude"), crs = 4326)
+#Plot map of Europe
+#Very noisy with many coordinates per point,
+#Let's select a unique coordinate by study
+poll_data1 =  poll_data %>% 
+distinct(Study_id, .keep_all = TRUE)
+#-----------------------#
+#Plants-----
+#-----------------------#
+#Check unique number of plant species
+PlantNumber_by_ID = data %>% 
+select(Plant_rank, Plant_accepted_name, Study_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+select(Plant_accepted_name, Study_id) %>% 
+group_by(Study_id) %>% 
+summarise(Plant_counts = n_distinct(Plant_accepted_name))
+#Merge now with spp counts by network dataset
+plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
+#Create duplicate cord columns to operate on them
+plant_data$coords <- data.frame(
+Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
+#Set coordinates
+plant_data$coords_sf = st_as_sf(plant_data$coords, 
+  coords = c("Longitude", "Latitude"), crs = 4326)
+#Plot map of Europe
+#Very noisy with many coordinates per point,
+#Let's select a unique coordinate by study
+plant_data1 =  plant_data %>% 
+distinct(Study_id, .keep_all = TRUE)
+#Add vars
+plant_data1$Group = "Plant"
+poll_data1$Group = "Pollinator"
+#Rename cols
+plant_data2 = plant_data1 %>% 
+rename(Counts = Plant_counts)
+poll_data2 = poll_data1 %>% 
+rename(Counts = Pollinator_counts)
+#Bind rows
+all_data = bind_rows(plant_data2, poll_data2)
+
+#Try to plot both together (same plot)
+ggplot(countries) +
+geom_sf(fill="gray80",color = "black", 
+        alpha = 1,linewidth=0.1) + 
+geom_point(data = all_data, alpha=1,shape=21,
+           aes(x= Longitude, y=Latitude,
+           size=Counts, group=Group, colour = Group))+  
+coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
+         xlim = c(-14, 39), ylim = c(37, 70))+
+theme(panel.grid.major = element_blank(), 
+      legend.position = ("right"),
+         legend.background = element_blank(),
+         legend.box.background = element_blank(),
+         legend.key = element_blank(), 
+         legend.title = element_text(face = "bold",  size = 8),
+      axis.text = element_text(size=6, face="bold"),
+         axis.title = element_text(face="bold", size = 10),
+         plot.title = element_text(face="bold", size = 10),
+         legend.text = element_text(face = "bold",
+         size = 6), legend.direction="vertical",
+         legend.spacing.y = unit(-0, "cm"),
+      legend.title.align=0.5) +
+theme(panel.background = element_rect(fill = "white"),
+        panel.border = element_rect(colour = "black",
+        fill=NA, size=1))+
+  scale_color_manual(name = "Taxonomic\ngroup", values = c("#009E73", "#D55E00")) +
+  scale_size(name = "Species\ncounts") +
+  ylab("Latitude") +
+xlab("Longitude") +
+guides(color = guide_legend(override.aes = list(size=3)))+
+scale_x_continuous(breaks = seq(0, 20, by = 10)) +
+scale_y_continuous(breaks = seq(40, 60, by = 10))
+
+```
 # Results

### 2023-12-12 Manuscript/Ms.qmd
Updating author list from Hervias-Parejo

@@
 ```
-# Abstract
-
-Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time.
-Here, we present a dataset of plant-pollinator interactions at European level that consists in XXX studies compiled during the years xxx and xxxx.
-
-Here, we have tried to generate a comprehensive dataset of plant-pollination interactions to build the first European metaweb of plant and pollinators and explore their general patterns.
-
-we have done a strong collaborative effort in unifying plant-pollinator studies at continent level to elucidate general patterns of plant-pollinator interactions and set a baseline for future ecological research.
-Thus, we explore the first time an European metaweb of plant-pollinator interactions
-
-and explore the first European metaweb of plant-pollinator interactions.
-This is a continent-level metaweb that comprised of xxx plants xxx pollinators and xxxx interactions which covers xxxx different countries.
-We show that plant and pollinator species are xxxxx and have xxxxxxx.
-For instance, we cover xxx% of European bee species and xxx% of flowering plants that rely on pollinators.
-
-# Introduction
-
-1st paragraph
-
-General introduction of how global change impacts plant-pollinator interactions
-
-Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions
-
-Highlight the relevance of large scale datasets
-
-2nd paragraph
-
-3rd paragraph
-
-4paragraph
-
-Introduce research questions
-
-LIST THEM HERE (Main ideas so far)
-
-Questions that we would like to answer:
-
-1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity
-
-2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?
-
-# Methods
 
@@
 ```
+
+# Abstract
+
+Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
+However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
+One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time.
+Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies compiled during the years xxx and xxxx.
+
+Here, we have tried to generate a comprehensive dataset of plant-pollination interactions to build the first European metaweb of plant and pollinators and explore their general patterns.
+
+we have done a strong collaborative effort in unifying plant-pollinator studies at continent level to elucidate general patterns of plant-pollinator interactions and set a baseline for future ecological research.
+Thus, we explore the first time an European metaweb of plant-pollinator interactions
+
+and explore the first European metaweb of plant-pollinator interactions.
+This is a continent-level metaweb that comprised of xxx plants xxx pollinators and xxxx interactions which covers xxxx different countries.
+We show that plant and pollinator species are xxxxx and have xxxxxxx.
+For instance, we cover xxx% of European bee species and xxx% of flowering plants that rely on pollinators.
+
+# Introduction
+
+1st paragraph
+
+General introduction of how global change impacts plant-pollinator interactions
+
+Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions
+
+Highlight the relevance of large scale datasets
+
+2nd paragraph
+
+3rd paragraph
+
+4paragraph
+
+Introduce research questions
+
+LIST THEM HERE (Main ideas so far)
+
+Questions that we would like to answer:
+
+1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity
+
+2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?
+
+# Methods
 

### 2023-12-12 Manuscript/Ms.qmd
adding levels of site id progamatically in metadata

@@
 One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time.
-Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies compiled during the years xxx and xxxx.
-
-Here, we have tried to generate a comprehensive dataset of plant-pollination interactions to build the first European metaweb of plant and pollinators and explore their general patterns.
-
-we have done a strong collaborative effort in unifying plant-pollinator studies at continent level to elucidate general patterns of plant-pollinator interactions and set a baseline for future ecological research.
-Thus, we explore the first time an European metaweb of plant-pollinator interactions
+Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies that cover `r country_number` countries and comprised over a million of interactions between plants and pollinators.
 

### 2023-12-12 Scripts/1_Processing_networks/14_Dupont.R

@@ -104,7 +104,7 @@ Taxa_recorded = "All flower visitors",
 Sampling_year = "2004 and 2005",
 Country = "Denmark",
 Habitat = "Heathland",
-Sampling_sites = 3,
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = "Full flowering season",
 Sampling_method = "Plots of 1*1m",
 Sampling_area_details = NA,

### 2023-12-12 Scripts/1_Processing_networks/1_Bartomeus.R

@@ -139,7 +139,7 @@ Taxa_recorded = "All floral visitors",
 Sampling_year = "2015",
 Country = "Spain",
 Habitat = "Mediterranean shrubland",
-Sampling_sites = "12",
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = "6",
 Sampling_method = "Focal observations",
 Sampling_area_details = "6 observation areas of 0.3 * 0.3 m / species and site",

### 2023-12-12 Scripts/1_Processing_networks/21_Hopfenmuller.R

@@ -82,7 +82,7 @@ terms of flowering units, not area",
 Sampling_year = 2010,
 Country = "Germany",
 Habitat = "Calcareous grassland",
-Sampling_sites = 23,
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = 5,
 Sampling_method = "Variable transect walks",
 Sampling_area_details = "0.1 ha per site (maybe 250 m long * 4 m wide?)",

### 2023-12-12 Scripts/1_Processing_networks/23_Holzschuh.R

@@ -94,7 +94,7 @@ Taxa_recorded = "Bees and hoverflies",
 Sampling_year = "2011 and 2012",
 Country = "Germany",
 Habitat = "Field boundary",
-Sampling_sites = "16",
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = 2,
 Sampling_method = "Transect",
 Sampling_area_details = "Two 150 x 1 m transects per round and site",

### 2023-12-12 Scripts/1_Processing_networks/27_Castro.R

@@ -45,7 +45,7 @@ Taxa_recorded = "All flower visitors",
 Sampling_year = 2021,
 Country = "Portugal",
 Habitat = "Pasture",
-Sampling_sites = "4",
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = NA,
 Sampling_method = "Random walks",
 Sampling_area_details = NA,

### 2023-12-12 Scripts/1_Processing_networks/30_Smith.R

@@ -65,7 +65,7 @@ Taxa_recorded = "Bees and syrphids",
 Sampling_year = "2011 and 2012",
 Country = "Sweden",
 Habitat = "Grassland",
-Sampling_sites = "16",
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = "4 per year",
 Sampling_method = "Transect",
 Sampling_area_details = "Two transects per site of 150 * 1m",

### 2023-12-12 Scripts/1_Processing_networks/31_Roberts.R

@@ -66,7 +66,7 @@ Status and Trends of European Pollinators (STEP)",
   Sampling_year = "2012 and 2013",
   Country = "England",
   Habitat = "Grassland",
-  Sampling_sites = "15",
+  Sampling_sites = nlevels(factor(data$Site_id)),
   Sampling_rounds = "4 per year",
   Sampling_method = "Transect",
   Sampling_area_details = "Two transects per site of 150 * 1m",

### 2023-12-12 Scripts/1_Processing_networks/34_Stanley.R

@@ -138,7 +138,7 @@ Taxa_recorded = "Bees, hoverflies and butterflies",
 Sampling_year = "2009",
 Country = "Ireland",
 Habitat = "Farming grassland",
-Sampling_sites = "25",
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = "3",
 Sampling_method = "Transects",
 Sampling_area_details = "4 transects of 100 x 2 m per site",

### 2023-12-12 Scripts/1_Processing_networks/37_White.R

@@ -144,7 +144,7 @@ Metadata <- tibble(
   Sampling_year = "2018",
   Country = "Ireland",
   Habitat = "Urban and agricultural landscapes",
-  Sampling_sites = "21",
+  Sampling_sites = nlevels(factor(data$Site_id)),
   Sampling_rounds = "4",
   Sampling_method = "Transects",
   Sampling_area_details = "4 transects of 100 x 2 m per site",

### 2023-12-12 Scripts/1_Processing_networks/45_Knight.R

@@ -72,7 +72,7 @@ Taxa_recorded = "Hymenoptera, Diptera and Lepidoptera",
 Sampling_year = "2018",
 Country = "Poland and Czech Republic",
 Habitat = "Grassland",
-Sampling_sites = "18",
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = "5",
 Sampling_method = "Transects",
 Sampling_area_details = "30 by 2 meter transects (10 to 6 per site)",

### 2023-12-12 Scripts/1_Processing_networks/6_Marini.R

@@ -79,7 +79,7 @@ the total percentage cover of all flowering plant species at each site.",
   Sampling_year = 2020,
   Country = "Italy",
   Habitat = "Agricultural",
-  Sampling_sites = 108,
+  Sampling_sites = nlevels(factor(data$Site_id)),
   Sampling_rounds = 4,
   Sampling_method = "Focal observations",
   Sampling_area_details = NA,

### 2023-12-12 Scripts/1_Processing_networks/7_Scheper.R

@@ -74,7 +74,7 @@ Taxa_recorded = "Bees and hoverflies",
 Sampling_year = 2015,
 Country = "France",
 Habitat = "Grassland",
-Sampling_sites = 25,
+Sampling_sites = nlevels(factor(data$Site_id)),
 Sampling_rounds = 4,
 Sampling_method = "Transects",
 Sampling_area_details = "2 transects per round (transect size 1 or 2 m × 75 m; we use 1.5 m to calculate the total area)",

### 2023-12-14 Manuscript/Figure1.qmd
overpower graph 1, take it or leave it

---
latex-tinytex: false
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
---

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr) #For handling fata 
library(ggplot2) #For plotting
library(giscoR) #For plotting
library(sf) #For handling coordinates and plotting
library(patchwork)#For plotting (binding plots)
library(ggstar) #For plotting (cool shapes)
library(scales) #For plotting (decimals on axes)
```

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}

#Load data
data = readRDS("../Data/Interactions_uncounted.rds")

#Prepare coordinates
for_plotting = data %>% 
select(Study_id, Latitude, Longitude) %>% 
distinct()
#Get countries and transfrom coordinates
countries <- gisco_get_countries(
resolution = 20) %>%
st_transform(3035) %>% 
sf::st_as_sf(crs = 4326)

#Number of plant and poll species per network
#-----------------------#
#Pollinators------
#-----------------------#
#Check unique number of pollinator species
PollNumber_by_ID = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
select(Pollinator_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
#Merge now with spp counts by network dataset
poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
poll_data$coords <- data.frame(
Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
#Set coordinates
poll_data$coords_sf = st_as_sf(poll_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
poll_data1 =  poll_data %>% 
distinct(Study_id, .keep_all = TRUE)

#-----------------------#
#Plants-----
#-----------------------#
#Check unique number of plant species
PlantNumber_by_ID = data %>% 
select(Plant_rank, Plant_accepted_name, Study_id) %>% 
filter(Plant_rank == "SPECIES") %>% 
select(Plant_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Plant_counts = n_distinct(Plant_accepted_name))
#Merge now with spp counts by network dataset
plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
plant_data$coords <- data.frame(
Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
#Set coordinates
plant_data$coords_sf = st_as_sf(plant_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
plant_data1 =  plant_data %>% 
distinct(Study_id, .keep_all = TRUE)

#Add vars
plant_data1$Group = "Plant"
poll_data1$Group = "Pollinator"
#Rename cols
plant_data2 = plant_data1 %>% 
rename(Counts = Plant_counts)
poll_data2 = poll_data1 %>% 
rename(Counts = Pollinator_counts)
#Bind rows
all_data = bind_rows(plant_data2, poll_data2)

#Try to plot both together (same plot)
map = ggplot(countries) +
geom_sf(fill="floralwhite",color = "black", 
        alpha = 1,linewidth=0.1) + 
geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
           size=Counts, starshape = Group, color= Group))+  
coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
         xlim = c(-14, 39), ylim = c(37, 70))+
theme(panel.grid.major = element_blank(), 
      legend.position = ("top"),
      legend.box = "vertical",
      legend.background = element_blank(),
      legend.box.background = element_blank(),
      legend.key = element_blank(), 
      legend.title = element_text(face =                "bold",  size = 6),
      legend.box.spacing = unit(0, "pt"),
      axis.text = element_text(size=6,                  face="bold"),
      axis.title = element_text(face="bold", size = 8),
         legend.text = element_text(face = "bold",
         size = 6), legend.direction="horizontal",
         legend.spacing.y = unit(-0.4, "cm"),
      legend.title.align=0.5) +
theme(panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(colour = "black",
        fill=NA, size=1)) + 
 scale_size_continuous(range = c(0, 4))+
scale_color_manual(values=c("#009E73", "#D55E00"),
         guide="none") +
scale_starshape_manual(name = "Taxonomic\ngroup",             values = c(15, 7),
             guide=guide_legend(
             override.aes=list(
                 size=2,
                 color=c("Plant"="#009E73", "Pollinator"= "#D55E00")))) +
  ylab("Latitude") +
xlab("Longitude") +
scale_x_continuous(breaks = seq(0, 20, by = 10)) +
scale_y_continuous(breaks = seq(40, 60, by = 10)) +
guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,7, 15)))) + 
ggtitle("(a)") +
theme(plot.title = element_text(vjust = -18, hjust=0.12, size=10))

#----------------#
#Pollinators
#----------------#
#First select cols of rank, accepted name and study ID
#Then select unique levels by study ID
#Now sum levels (maximum can be as maximum number of studies)
poll_spread = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
distinct(Pollinator_accepted_name) %>% 
ungroup() %>% 
count(Pollinator_accepted_name) %>% 
rename(n_studies_shared = n) %>% 
arrange(-n_studies_shared) %>% 
mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))

#Seelect quantile 50% from the total number of species
spp_number = poll_spread %>%  
select(Pollinator_accepted_name) %>% 
n_distinct()
df = seq(1, spp_number)
quantile_50 = quantile(df, 0.5)
#Plot
p1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.015, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  plot.title = element_text(size = 8, face = "bold"), 
  axis.text.y = element_text(size=6),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
xlab(NULL) +
ylab(NULL) + 
ggtitle("Pollinators")+
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.25) + 
annotate(geom="text", x=1280, y=0.83, label="50th percentile", color="black", size = 1.5) +
scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
annotate(geom="text", x=-500, y=1.71, label="(b)",color="black", size = 3.5) +
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")

#----------------#
#Plants
#----------------#
plant_spread = data %>% 
select(Plant_rank, Plant_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
filter(Plant_rank == "SPECIES") %>% 
distinct(Plant_accepted_name) %>% 
ungroup() %>% 
count(Plant_accepted_name) %>% 
rename(n_studies_shared = n) %>% 
arrange(-n_studies_shared) %>% 
mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))

#Seelect quantile 50% from the total number of species
spp_number = plant_spread %>%  
select(Plant_accepted_name) %>% 
n_distinct()
df = seq(1, spp_number)
quantile_50 = quantile(df, 0.5)
#Plot
p2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.015, fill = "#009E73") +
ylab("Shared plants (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         axis.title.x = element_text(face = "bold", size = 8),            plot.title = element_text(size = 8, face = "bold"),
         axis.text.y = element_text(size=6),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
xlab("Species") +
ylab(NULL) + 
ggtitle("Plants") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.25) + 
annotate(geom="text", x=830, y=0.7, label="50th percentile",color="black", size = 1.5) +
scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")

#Prepare panel with plots
plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Percentage of studies              \n sharing a species              ") + 
theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
#Plot everything
map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.65, 1))

```

### 2023-12-15 Manuscript/Figure1.qmd
Minor data edits

@@
       legend.key = element_blank(), 
-      legend.title = element_text(face =                "bold",  size = 6),
-      legend.box.spacing = unit(0, "pt"),
+      legend.title = element_text(face ="bold",  size = 6),
+      legend.box.spacing = unit(0.2, "cm"),
+      legend.margin=margin(0,0,0,0),
+        legend.box.margin=margin(-10,-10,-10,-10),
       axis.text = element_text(size=6,                  face="bold"),
       axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(face = "bold",
-         size = 6), legend.direction="horizontal",
-         legend.spacing.y = unit(-0.4, "cm"),
+         legend.text = element_text(size = 6),                           legend.direction="horizontal",
+         legend.spacing.y = unit(-0.3, "cm"),
       legend.title.align=0.5) +
@@
         fill=NA, size=1)) + 
- scale_size_continuous(range = c(0, 4))+
+ scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
 scale_color_manual(values=c("#009E73", "#D55E00"),
          guide="none") +
-scale_starshape_manual(name = "Taxonomic\ngroup",             values = c(15, 7),
-             guide=guide_legend(
+scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
+             guide=guide_legend(reverse=TRUE,
              override.aes=list(
                  size=2,
-                 color=c("Plant"="#009E73", "Pollinator"= "#D55E00")))) +
+                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
   ylab("Latitude") +
@@
 scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,7, 15)))) + 
+guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
 ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -18, hjust=0.12, size=10))
+theme(plot.title = element_text(vjust = -12, hjust=0.1, size=10))
 
@@
 quantile_50 = quantile(df, 0.5)
+
+#Obtain most common pollinator for plotting
+most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
+select(Pollinator_accepted_name) %>% pull()
+
 #Plot
@@
 color = "black", size=0.25) + 
-annotate(geom="text", x=1280, y=0.83, label="50th percentile", color="black", size = 1.5) +
+annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
+annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
 scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-annotate(geom="text", x=-500, y=1.71, label="(b)",color="black", size = 3.5) +
+annotate(geom="text", x=-500, y=1.15, label="(b)",color="black", size = 3.5) +
 coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
@@
 quantile_50 = quantile(df, 0.5)
+
+#Obtain most common plant for plotting
+most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
+select(Plant_accepted_name) %>% pull()
+
 #Plot
@@
           color = "black", size=0.25) + 
-annotate(geom="text", x=830, y=0.7, label="50th percentile",color="black", size = 1.5) +
+annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
+annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
+      ,color="black",  fontface = 'italic', size = 1.5)+
+annotate(geom="text", x=20, y=-0.08, label="Common spp."
+      ,color="black",  size = 1.5) +
+annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
+      ,color="black",  size = 1.5) + 
 scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
@@
 #Prepare panel with plots
-plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Percentage of studies              \n sharing a species              ") + 
+plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Occurrence across studies (%)") + 
 theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
 #Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.65, 1))
+map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 1))
 

### 2023-12-15 Manuscript/Ms.qmd
Minor data edits

@@
 title: "Mapping plant-pollinator interactions across Europe"
-format: pdf
-editor: visual
+latex-tinytex: false
+format:
+  pdf:
+    include-in-header: 
+      text: |
+        \usepackage{lscape}
+        \newcommand{\blandscape}{\begin{landscape}}
+        \newcommand{\elandscape}{\end{landscape}}
 ---
@@
 ```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr)
-library(ggplot2)
-library(giscoR)
-library(sf)
+library(dplyr) #For handling fata 
+library(ggplot2) #For plotting
+library(giscoR) #For plotting
+library(sf) #For handling coordinates and plotting
+library(patchwork)#For plotting
+library(ggstar) #For plotting (cool shapes)
+library(scales) #For plotting (decimals on axes)
 ```
@@
 However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time.
-Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies that cover `r country_number` countries and comprised over a million of interactions between plants and pollinators.
-
-and explore the first European metaweb of plant-pollinator interactions.
-This is a continent-level metaweb that comprised of xxx plants xxx pollinators and xxxx interactions which covers xxxx different countries.
-We show that plant and pollinator species are xxxxx and have xxxxxxx.
-For instance, we cover xxx% of European bee species and xxx% of flowering plants that rely on pollinators.
+One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
+Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies distributed across `r country_number` countries that comprise over a million of interactions between plants and pollinators.
+The dataset includes a total of `r pollinator_spp_number` pollinator and `r plant_spp_number` plant species that accounts approximately for 30% of each of the main pollinator groups (i.e., bees, syrphids and butterflies) and 5% of flowering plants that inhabit the European continent.
 
@@
 
+Shared species across studies (describe trend), added now in graph 1?
+
+Update graph 1 (Add shared species of plants and pollinators?)
+
 ## Species coverage
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%"}
+To calculate the completeness of the dataset
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}
+
+#Load data
+data = readRDS("../Data/Interactions_uncounted.rds")
+
 #Prepare coordinates
@@
 distinct(Study_id, .keep_all = TRUE)
+
 #-----------------------#
@@
 distinct(Study_id, .keep_all = TRUE)
+
 #Add vars
@@
 #Try to plot both together (same plot)
-ggplot(countries) +
-geom_sf(fill="gray80",color = "black", 
+map = ggplot(countries) +
+geom_sf(fill="floralwhite",color = "black", 
         alpha = 1,linewidth=0.1) + 
-geom_point(data = all_data, alpha=1,shape=21,
-           aes(x= Longitude, y=Latitude,
-           size=Counts, group=Group, colour = Group))+  
+geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
+           size=Counts, starshape = Group, color= Group))+  
 coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
@@
 theme(panel.grid.major = element_blank(), 
-      legend.position = ("right"),
+      legend.position = ("top"),
+      legend.box = "vertical",
       legend.background = element_blank(),
@@
       legend.key = element_blank(), 
-         legend.title = element_text(face = "bold",  size = 8),
+      legend.title = element_text(face ="bold",  size = 6),
+      legend.box.spacing = unit(0.2, "cm"),
+      legend.margin=margin(0,0,0,0),
+        legend.box.margin=margin(-10,-10,-10,-10),
       axis.text = element_text(size=6,                  face="bold"),
-         axis.title = element_text(face="bold", size = 10),
-         plot.title = element_text(face="bold", size = 10),
-         legend.text = element_text(face = "bold",
-         size = 6), legend.direction="vertical",
-         legend.spacing.y = unit(-0, "cm"),
+      axis.title = element_text(face="bold", size = 8),
+         legend.text = element_text(size = 6),                           legend.direction="horizontal",
+         legend.spacing.y = unit(-0.3, "cm"),
       legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "white"),
+theme(panel.background = element_rect(fill = "aliceblue"),
         panel.border = element_rect(colour = "black",
-        fill=NA, size=1))+
-  scale_color_manual(name = "Taxonomic\ngroup", values = c("#009E73", "#D55E00")) +
-  scale_size(name = "Species\ncounts") +
+        fill=NA, size=1)) + 
+ scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
+scale_color_manual(values=c("#009E73", "#D55E00"),
+         guide="none") +
+scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
+             guide=guide_legend(reverse=TRUE,
+             override.aes=list(
+                 size=2,
+                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
   ylab("Latitude") +
 xlab("Longitude") +
-guides(color = guide_legend(override.aes = list(size=3)))+
 scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10))
+scale_y_continuous(breaks = seq(40, 60, by = 10)) +
+guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
+ggtitle("(a)") +
+theme(plot.title = element_text(vjust = -12, hjust=0.1, size=10))
+
+#----------------#
+#Pollinators
+#----------------#
+#First select cols of rank, accepted name and study ID
+#Then select unique levels by study ID
+#Now sum levels (maximum can be as maximum number of studies)
+poll_spread = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
+group_by(Study_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+ungroup() %>% 
+count(Pollinator_accepted_name) %>% 
+rename(n_studies_shared = n) %>% 
+arrange(-n_studies_shared) %>% 
+mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = poll_spread %>%  
+select(Pollinator_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common pollinator for plotting
+most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
+select(Pollinator_accepted_name) %>% pull()
+
+#Plot
+p1 = ggplot(poll_spread, 
+aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
+geom_bar(stat = "identity", 
+           color = "#D55E00",
+           lwd = 0.015, fill = "#D55E00") +
+ylab("Shared pollinators (%)") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+  axis.ticks.x=element_blank(),
+  axis.ticks.y = element_line(size = 0.2),
+  plot.title = element_text(size = 8, face = "bold"), 
+  axis.text.y = element_text(size=6),
+  panel.grid.major.x = element_blank(),
+  panel.grid.minor.y = element_blank(),
+  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
+xlab(NULL) +
+ylab(NULL) + 
+ggtitle("Pollinators")+
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+color = "black", size=0.25) + 
+annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
+annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
+scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
+annotate(geom="text", x=-500, y=1.15, label="(b)",color="black", size = 3.5) +
+coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
+
+#----------------#
+#Plants
+#----------------#
+plant_spread = data %>% 
+select(Plant_rank, Plant_accepted_name, Study_id) %>% 
+group_by(Study_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+ungroup() %>% 
+count(Plant_accepted_name) %>% 
+rename(n_studies_shared = n) %>% 
+arrange(-n_studies_shared) %>% 
+mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = plant_spread %>%  
+select(Plant_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common plant for plotting
+most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
+select(Plant_accepted_name) %>% pull()
+
+#Plot
+p2 = ggplot(plant_spread, 
+aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
+geom_bar(stat = "identity", 
+           color = "#009E73",
+           lwd = 0.015, fill = "#009E73") +
+ylab("Shared plants (%)") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+         axis.ticks.x=element_blank(),
+         axis.ticks.y = element_line(size = 0.2),
+         axis.title.x = element_text(face = "bold", size = 8),            plot.title = element_text(size = 8, face = "bold"),
+         axis.text.y = element_text(size=6),
+         panel.grid.major.x = element_blank(),
+         panel.grid.minor.y = element_blank(),
+         panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
+xlab("Species") +
+ylab(NULL) + 
+ggtitle("Plants") + 
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+          color = "black", size=0.25) + 
+annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
+annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
+      ,color="black",  fontface = 'italic', size = 1.5)+
+annotate(geom="text", x=20, y=-0.08, label="Common spp."
+      ,color="black",  size = 1.5) +
+annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
+      ,color="black",  size = 1.5) + 
+scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
+coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")
+
+#Prepare panel with plots
+plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Occurrence across studies (%)") + 
+theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
+#Plot everything
+map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 1))
 
 ```
+
 # Results

### 2023-12-15 Scripts/Metaweb_info/Working_metadata.R

@@ -6,11 +6,34 @@ data = readRDS("Data/Interactions_uncounted.rds")
 #Select cols of interest
 metadata = data %>% 
 select(Study_id, Network_id, Sampling_method) %>% 
-distinct() %>% 
-mutate(Sampling_time = "") %>% 
-mutate(Sampling_area = "") 
+distinct()
 
 #Load library
-#library(googlesheets4)
+library(googlesheets4)
 #Load it to google sheets
 #(ss <- gs4_create("Sampling_details", sheets = metadata))
+
+library(lubridate)
+#Select cols of interest
+dates = data %>% 
+select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
+mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
+#Obtain date information
+dates1 = dates %>% 
+group_by(Study_id, Network_id, Sampling_method) %>% 
+summarise(Min_date = min(Date), 
+          Max_date= max(Date), 
+          Sampling_days = n_distinct(Date)) %>% 
+mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
+mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
+       .before=Study_id) %>% 
+arrange(Study_number) %>% 
+mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
+TRUE ~ NA)) %>% 
+mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
+TRUE ~ NA))
+
+(ss <- gs4_create("Sampling_details", sheets = dates1))
+
+
+

### 2023-12-18 Scripts/1_Processing_networks/19_Jauker.R
Updating scripts and processing files

@@ -53,6 +53,8 @@ mutate(Sampling_method = "Transect")
 #Unify structure of data
 data = change_str(data)
 
+levels(factor(data$Site_id))
+
 #Split interaction data into dataframes within a list
 InteractionData <- split(data,data$Site_id)
 

### 2023-12-18 Scripts/1_Processing_networks/20_Hoiss.R

@@ -42,7 +42,7 @@ select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) #Including this
 
 #Unify level
 data = data %>% 
-mutate(Sampling_method = "Transect")
+mutate(Sampling_method = "Plot")
 
 #Unify structure of data
 data = change_str(data)

### 2023-12-18 Scripts/3_1_Pollinator_taxonomy.R

@@ -578,6 +578,7 @@ mutate(Fixed = case_when(
   Mismatch == "Deraeocoris pratensis" ~ "Lygus pratensis", #fix
   Mismatch == "Axinotarsus varitarsis" ~ "Attalus varitarsis", #fix
   Mismatch == "Heterocapillus nigripes" ~ "Heterocapillus tigripes", #fix
+  Fixed == "Arachnopila" ~ "Arachnospila", #fix
 
   T ~ Fixed)) %>% 
   rename(Old_name = Mismatch, Name = Fixed) 
@@ -726,6 +727,14 @@ unmatched_gbif = name_backbone_checklist(name= name1, kingdom='animals')
 #Rename and filter out exact matches
 unmatched_gbif1 = change_str1(unmatched_gbif)
 clean = c("EXACT", "FUZZY")
+
+#Exclude some species that are not correctly identified
+unmatched_gbif1 = unmatched_gbif1 %>% 
+mutate(Matchtype = case_when(Fixed_name == "Zygoptera" ~ NA_character_,
+                             Fixed_name == "Podarcis lilfordi" ~ NA_character_,
+                             Fixed_name == "Cheloninae" ~ NA_character_,
+                           TRUE ~  Matchtype))
+
 #Select matched records and pass Canonical_name to accepted
 #When necessary
 unmatched_gbif1_found = unmatched_gbif1 %>% 
@@ -919,7 +928,13 @@ levels(factor(all$Study_id))
 #Do this fo every new dataset that we add
 #Last one being checked is written within the filter argument
 subset_check = all %>% 
-filter(Study_id == "50_Hervias-Parejo") %>% 
+filter(Study_id == "49_Hervias-Parejo") %>% 
 select(Old_name, Fixed_name, Rank, Status, Matchtype, Accepted_name, Unsure_id) %>% 
 distinct()
 
+subset_check = all %>% 
+filter(Study_id == "40_Knight") %>% 
+distinct()
+
+s = subset_check %>% 
+filter(is.na(Order))

### 2023-12-19 Manuscript/Ms.qmd
Updating scripts and processing files
Minor edit

@@
 n_distinct()
+#Average study number
+studies_by_country = data %>% 
+group_by(Country) %>% 
+summarise(study_number = n_distinct(Study_id)) %>% 
+summarise(mean(study_number)) %>% 
+pull()
+#Sampling method
+Sampling_method = data %>% 
+group_by(Study_id) %>% 
+distinct(Sampling_method) %>%  
+ungroup() %>% 
+count(Sampling_method) %>% 
+mutate(percent_method = n/sum(n) *100) %>% 
+mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
+#1st most common sampling methodology
+most_common_method = Sampling_method %>% 
+filter(percent_method == max(percent_method)) 
 #Minimum year
@@
 This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
-This dataset covers `r country_number` countries and consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year`, and accounts for a total of `r prettyNum(interaction_number, big.mark=",")` interactions from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+This dataset consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (mean studies by country = `r studies_by_country`).
+The dataset contains a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+These different studies differ in sampling effort and methodology, although most studies cover only a single year or flowering season (PERCENTAGE) with an average of XXX sampling days (range XXX to XXXX) over a time span of XXXX days, and documented interactions by using `r  most_common_method[[1]]` (`r  most_common_method[[3]]`%).
+
+Explain % of taxonomic ranks at order level that constitute the included pollinators.
 
 Shared species across studies (describe trend), added now in graph 1?
-
-Update graph 1 (Add shared species of plants and pollinators?)
 

### 2024-01-04 Manuscript/Ms.qmd
Relatively large update (Manincor, Fisogni and Petanidou added)

--- 2023-12-15 Manuscript/Ms.qmd
@@
 n_distinct()
+#Average study number
+studies_by_country = data %>% 
+group_by(Country) %>% 
+summarise(study_number = n_distinct(Study_id)) %>% 
+summarise(mean(study_number)) %>% 
+pull()
+#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
+#Sampling method
+Sampling_method = data %>% 
+group_by(Study_id) %>% 
+distinct(Sampling_method) %>%  
+ungroup() %>% 
+count(Sampling_method) %>% 
+mutate(percent_method = n/sum(n) *100) %>% 
+mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
+#1st most common sampling methodology
+most_common_method = Sampling_method %>% 
+filter(percent_method == max(percent_method)) 
 #Minimum year
@@
 This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
-This dataset covers `r country_number` countries and consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year`, and accounts for a total of `r prettyNum(interaction_number, big.mark=",")` interactions from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+This dataset consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
+The dataset contains a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (PERCENTAGE) with an average of XXX sampling days (range XXX to XXXX) over a time span of XXXX days, and documented interactions mostly by using `r  most_common_method[[1]]` (`r  most_common_method[[3]]`%).
+
+Explain % of taxonomic ranks at order level that constitute the included pollinators. Something with the different pollinators is going to be graph 2.
 
 Shared species across studies (describe trend), added now in graph 1?
-
-Update graph 1 (Add shared species of plants and pollinators?)
 

### 2024-01-04 Scripts/1_Processing_networks/51_Petanidou.R

@@ -1,0 +1,125 @@
+#DATASET NUMBER 51; Petanidou
+#Dataset sent by T. Petanidou
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+
+#Read empty templates to compare with
+source("Scripts/Processing/Functions/Empty_templates.R") 
+#Load function to unify structure of data
+source("Scripts/Processing/Functions/Change_str.R")
+
+#Prepare interaction data ----
+data <- read_csv("Data/Raw_data/51_Petanidou/Interaction_data.csv", locale = locale(encoding = "latin1"))
+
+#Compare vars
+#compare_variables(check_interaction_data, data)
+#Nothing missing!
+
+#Drop sampling effort and square area (it would be added on the metadata)
+#Replace "m" for empty space in coordinate precision
+data = data %>%
+select(!c(Sampling_effort_minutes, Sampling_area_square_meters)) %>% 
+mutate(Coordinate_precision = str_replace(Coordinate_precision, "m", ""))
+
+
+#Some plant names include 2 spp
+#After asking authors we need to divide those in 2 rows
+data_1spp = data %>% 
+filter(!str_detect(Plant_species, "[|]"))
+#Now work on those
+data_2spp = data %>% 
+filter(str_detect(Plant_species, "[|]"))
+#Split col into 2 cols
+data_2spp = data_2spp %>%
+separate(Plant_species, c("A", "B"), "[|]")
+#Create dataframe 1
+data_2spp_1 = data_2spp %>% 
+select(-B) %>% 
+rename(Plant_species = A) %>% 
+mutate(Plant_species = word(Plant_species, 1,2))
+#Create dataframe 2
+data_2spp_2 = data_2spp %>% 
+select(-A) %>% 
+rename(Plant_species = B) %>% 
+mutate(Plant_species = word(Plant_species, 1,2))
+#Bind dataframes (recovered spp and records without mistakes)
+data = bind_rows(data_2spp_1, data_2spp_2, data_1spp)
+
+#Select just 2 words for the rest of records too
+data = data %>% 
+mutate(Plant_species = word(Plant_species, 1,2))
+
+#Unify structure of data
+data = change_str(data)
+
+#Split interaction data into dataframes within a list
+InteractionData <- split(data, data$Site_id)
+
+#Prepare flower count data ---- 
[…]

### 2024-01-04 Scripts/3_2_Pollinator_coverage.R

@@ -158,9 +158,7 @@ lepidoptera_coverage
 #Save coverage values---- 
 #----------------------------#
 #Final checks before saving!
-bee_checks
-syrphid_checks
-lepidoptera_checks
+c(bee_checks, syrphid_checks, lepidoptera_checks)
 
 #Save
 saveRDS(bee_coverage, "Data/Manuscript_info/bee_coverage.RData")

### 2024-01-04 Scripts/Processing/Pollinators/2_3_Rename_unmatched_pollinators.R

@@ -1643,3 +1643,61 @@ mutate(Matchtype = case_when(
    Fixed_name == "Cheloninae" ~ "EXACT",
    T ~ Matchtype)) 
 
+#Eumerus clavatus uncipes
+unmatched_gbif1_not_found = 
+unmatched_gbif1_not_found %>% 
+mutate(Rank = case_when(
+   Fixed_name == "Eumerus clavatus uncipes" ~ "SPECIES",
+   T ~ Rank)) %>% 
+mutate(Matchtype = case_when(
+   Fixed_name == "Eumerus clavatus uncipes" ~ "EXACT",
+   T ~ Matchtype)) 
+
+#Cheilosia albitarsis ranunculi
+unmatched_gbif1_not_found = 
+unmatched_gbif1_not_found %>% 
+mutate(Matchtype = case_when(
+   Fixed_name == "Cheilosia albitarsis ranunculi" ~ "EXACT",
+   T ~ Matchtype)) 
+
+#Platycheirus albimanus muelleri
+unmatched_gbif1_not_found = 
+unmatched_gbif1_not_found %>% 
+mutate(Matchtype = case_when(
+   Fixed_name == "Platycheirus albimanus muelleri" ~ "EXACT",
+   T ~ Matchtype)) 
+
+#Meria
+unmatched_gbif1_not_found = 
+unmatched_gbif1_not_found %>% 
+mutate(Accepted_name = case_when(
+   Fixed_name == "Meria" ~ "Meria",
+   T ~ Accepted_name)) %>% 
+mutate(Scientific_name = case_when(
+   Fixed_name == "Meria" ~ "Meria (Illiger, 1807)",
+   T ~ Scientific_name)) %>% 
+mutate(Canonical_name = case_when(
+   Fixed_name == "Meria" ~ "Meria",
+   T ~ Canonical_name)) %>% 
+mutate(Phylum = case_when(
+   Fixed_name == "Meria" ~ "Arthropoda",
+   T ~ Phylum)) %>% 
+mutate(Order = case_when(
+   Fixed_name == "Meria" ~ "Hymenoptera",
+   T ~ Order)) %>% 
+mutate(Family = case_when(
+   Fixed_name == "Meria" ~ "Tiphiidae",
+   T ~ Family)) %>% 
+mutate(Genus = case_when(
+   Fixed_name == "Meria" ~ "Meria",
+   T ~ Genus)) %>% 
[…]

### 2024-01-04 Manuscript/Ms.qmd
Minor edits on Ms

--- 2023-12-15 Manuscript/Ms.qmd
@@
 library(scales) #For plotting (decimals on axes)
+library(lubridate) #To operate with dates
+library(tidyr) #Reshape data (wide format)
+library(stringr) #Fix/modify strings
+
 ```
@@
 n_distinct()
+#Extract percentage of studies conducted during a single year
+dates = data %>% 
+select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
+mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
+group_by(Study_id) %>% 
+summarise(Count = n_distinct(Year))
+#Filter studies with only one year
+one_year = dates %>% 
+filter(Count == 1) %>% 
+pull() %>% 
+length() / nrow(dates) * 100
+#Obtain date information
+dates = data %>% 
+select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
+mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
 
+dates1 = dates %>% 
+group_by(Study_id, Network_id, Sampling_method) %>% 
+summarise(Min_date = min(Date), 
+          Max_date= max(Date), 
+          Sampling_days = n_distinct(Date)) %>% 
+mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
+mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
+       .before=Study_id) %>% 
+arrange(Study_number) %>% 
+mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
+TRUE ~ NA)) %>% 
+mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
+TRUE ~ NA))
+#Extract min and max number of sampling days
+min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
+max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
+#Extract average of sampling days across networks and studies
+average_sampling_days = dates1 %>% 
+ungroup() %>% 
+group_by(Study_id) %>% 
+summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
+ungroup() %>% 
+summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
+pull()
 ```
@@
 This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
-This dataset covers `r country_number` countries and consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year`, and accounts for a total of `r prettyNum(interaction_number, big.mark=",")` interactions from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+This dataset consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
+The dataset contains a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
+These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
 
-Shared species across studies (describe trend), added now in graph 1?
-
-Update graph 1 (Add shared species of plants and pollinators?)
-
-## Species coverage
+## Taxonomic coverage
 
 To calculate the completeness of the dataset
+
+Check presentation from Herrera for cool graphs summarising patterns
 

### 2024-01-05 Data/1_Raw_data/18_Bartomeus/Protocols.rmd
Organisation of files

---
title: "Protocols"
author: "Bartomeus, Molina, ..."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Core observations

Aim: Understand the network of interactions between plants and pollinators and 
consequences for plant reproduction.

Data collection is done in 12 to 16 independent sites in a gradient of land use intensity across the Guadalquivir valley. All sites are sampled every ~ 15 days (weather permitting) for a total of 6 to 8 rounds per year. All sites are situated in open pine woodlands with an understory dominated by Lavender, Rosemary and Cistaceas. Data collection covers the whole  season from early spring (February) to Summer (June). We only sample on good weather conditions (no rain, extreme cold or high wind speeds).

Monitoring protocols:

1)	**Transect walks**:A fixed 100 meters straight line within each plot is walked at a slow pace in 30 minutes. During that time,  we record all the interactions that are observed between pollinators and flowering plants. In the case that we do not know the pollinator, it is captured with a hand net and stored in a plastic jar, which will be numbered and noted in a central notebook. Time devoted to capturing insects is deducted from the total sampling time. Since 2020 we repeat this transect twice to estimate detectability. In 2021 we ran this protocol weekly (12 sampling rounds).

After the 30 minutes transect walk, we conduct a __flower count__ along the same 100 meter transect. Every 10 meters we count all the species of flowering plants in about 1 meter around us (~3 m2). Overall, we do 10 counts along the transect.

-	**Focal census**: In addition, we perform focal census on specific plants. Those consist of choosing a flowering plant and observing all visitors during a 3 minute census, except the years 2015 and 2016, where we observed all visitors during 5 minutes. In the case that we do not know the pollinator species, it is captured with a hand net and stored in a plastic jar, which will be numbered and noted in a central notebook. Time devoted to capturing insects is deducted from the total sampling time. Usually, the most common plant species are selected in each site.

- **Fruit and seed set**: 13 plots were sampled between 2015 and 2019 and the fruit set and seed set were obtained for 13 different plant species (3-4 individuals per plant species). For each individual plant, the number of developed fruits (fruit yes), not developed fruits (fruits no), predated fruit, and dispersed fruit were recorded. In 2015 and 2016 fruits were counted per branch, whereas 2017, 2018 and 2019 fruits were aggregated for the whole plant. The fruit formed was obtained as the sum of fruit yes + predated fruit +  dispersed fruit, and the total potential fruits as the sum of the fruit formed + fruit no. The fruit proportion was calculated as formed fruit divided by total fruit for each individual per plant species at each site and year.

For the seed set, the number of seeds per fruit was counted (at least 3 fruit per individual) and each seed was weighed. In some fruits, more than one seed was weighed (for example, for individual A1, fruit "a" has 2 seeds and both are weighed), so the average weight of the seed was calculated per fruit in each individual. The average number of seeds per fruit and the average weight of the seed per fruit were obtained using the total amount of fruits per individual per plant species.

-	**Outside transect**: We also opportunistically capture rare pollinators outside of our transect, so that we can collect more complete information about the pollinators that we have in the plots at that time. Captured pollinators are placed in labeled jars for later identification in the lab and registered in the notebook.

-	**Identification process**: Due to the amount of bees captured we conserved them in a cooler at -20ºC. At the lab, the bees are pinned and labeled. Later, we identify them using the scarce taxonomic information available through determination keys. All the specimens were sent to a taxonomic expert to check the identification done and to identify the remaining unidentified species. The labels always contains the following information: Site, date, jar number, plant, family or genus if we know.

- **single visit deposition**: TBA

Departures from common protocols and important changes.

- Since 2020 we conduct each transect twice.
- For 2022, we have extra data of butterfly transects and 18 new sites (Nerea).
- We have extra data on transects from David Paz for 2015 year.
- We have extra data on cavity nesting bees for XXXX year.
- We have extra data on butterflies transects from XXXX year (Mark de Jong)
- Pairing between focal plants observed and fruit set?

### 2024-01-10 Scripts/1_Processing_networks/36_Larkin.R
Exploring habitats

@@ -98,6 +98,30 @@ mutate(Sampling_method = "Transect")
 data = data %>% 
 filter(!Year == 2018)
 
+#There is one network (site_id) that has 2 coordinates
+#Unify for simplicity as they are practically identical
+#latitude
+latitude = data %>% 
+select(Site_id, Latitude)  %>% 
+filter(Site_id == "GlencolumbkilleSouthCarronCountyClare") %>% 
+pull(Latitude)
+#longitude
+longitude = data %>% 
+select(Site_id, Longitude)  %>% 
+filter(Site_id == "GlencolumbkilleSouthCarronCountyClare") %>% 
+pull(Longitude)
+#Add coordinates
+#1st latitude
+data = data %>% 
+mutate(Latitude = 
+  case_when(Site_id == "GlencolumbkilleSouthCarronCountyClare" ~ latitude[1],
+  TRUE ~ Latitude))
+#2nd longitude
+data = data %>% 
+mutate(Longitude = 
+  case_when(Site_id == "GlencolumbkilleSouthCarronCountyClare" ~ longitude[1],
+  TRUE ~ Longitude))
+
 #Unify structure of data
 data = change_str(data)
 

### 2024-01-24 Manuscript/Ms.qmd
Exploring habitats
adding graph

@@
         \newcommand{\elandscape}{\end{landscape}}
+        \usepackage{caption}
+        \captionsetup[figure]{labelformat=empty}
 ---
@@
 library(stringr) #Fix/modify strings
+library(forcats)
 
@@
 #Print number of countries
-data = readRDS("../Data/Interactions_uncounted.rds")
+data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
 #Country number
@@
 
-Highlight the relevance of large scale datasets
+Highlight the relevance of large scale datasets to understand biological phenomena
 
@@
 
+Explore habitat types! (I think I need to fix some)
+
+## Taxonomy
+
+All plant and pollinator species names were checked and standardised according to the most recent
+
 ## Taxonomic coverage
@@
 
+Phylo for plants and pollinators!
+
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}
 
-#Load data
-data = readRDS("../Data/Interactions_uncounted.rds")
+#Load data (done it in previous chunk)
 
@@
 geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
+        alpha = 1,linewidth=0.01) + 
 geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
@@
 annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
-annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
+annotate(geom="text", x=20, y=0.89, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
 scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
@@
 annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
+annotate(geom="text", x=20, y=0.73, label=most_common_plant[1]
       ,color="black",  fontface = 'italic', size = 1.5)+
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%", fig.cap= "\\textbf{Figure 2}.
+Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
+The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
+The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
+Note that a single study can countribute to more than one habitat or bioclimatic region.
+Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=4, fig.height=4}
+#Load data (done it in earlier chunks)
+#Select cols of interest
+data1 = data %>% 
+select(Study_id, Network_id, Authors_habitat,
+  Corine_land_cover, SafeNet_habitat, Bioregion,
+  Latitude, Longitude, Day, Month, Year, Plant_rank,
+  Plant_status, Plant_matchtype, Plant_accepted_name,Plant_order,
+  Pollinator_rank, Pollinator_status, Pollinator_matchtype,
+  Pollinator_accepted_name, Pollinator_order, Interaction)
+
+#Prepare orders for filtering
+main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
+
+#Check number of studies by habitat
+habitat = data %>% 
+select(Study_id, SafeNet_habitat) %>% 
+distinct() %>%  
+group_by(SafeNet_habitat) %>% 
+summarise(Habitat_count = n()) 
+#Check number of studies by bioregion
+bioregion = data %>% 
+select(Study_id, Bioregion) %>% 
+distinct() %>%  
+group_by(Bioregion) %>% 
+summarise(Bioregion_count = n()) 
+#Check pols by habitat
+pollinator_percent = data1 %>%
+group_by(SafeNet_habitat, Pollinator_order) %>%
+filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+summarise(count = n()) %>%
+group_by(SafeNet_habitat) %>%
+mutate(Proportion = count/sum(count)) %>%
+select(-count) 
+#Check pols by bioregion
+pollinator_percent_bio = data1 %>%
+group_by(Bioregion, Pollinator_order) %>%
+filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+summarise(count = n()) %>%
+group_by(Bioregion) %>%
+mutate(Proportion = count/sum(count)) %>%
+select(-count) 
+#Preparing datasets for horizontal plot
+a1 = habitat %>% 
+mutate(Group = "Habitat") %>% 
+rename(Habitat = SafeNet_habitat) %>% 
+rename(Count = Habitat_count)
+
+b1 = bioregion %>% 
+mutate(Group = "Bioregion") %>% 
+rename(Habitat = Bioregion) %>% 
+rename(Count = Bioregion_count)
+
+#Bind datasets
+c1 = bind_rows(a1,b1)
+
+#Create datasets and rename cols
+a2 = pollinator_percent_bio %>% 
+mutate(Group = "Bioregion") %>% 
+rename(Habitat = Bioregion)
+
+b2 = pollinator_percent %>% 
+mutate(Group = "Habitat") %>% 
+rename(Habitat = SafeNet_habitat)
+#Bind rows
+c2 = bind_rows(a2, b2)
+
+#Prepare to set ordered labels
+l2 = c2 %>% 
+filter(Pollinator_order == "Hymenoptera") %>% 
+arrange((Proportion)) %>% 
+pull(Habitat)
+#Set order of labels
+c2$Habitat = factor(c2$Habitat, levels = l2)
+
+#Set the right order of labels
+c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
+
+p1 = ggplot(c2, 
+aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
+geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
+theme_bw() +
+theme(legend.position = "bottom",
+      plot.margin=unit(c(0,0,0,0), "mm"),
+      legend.text = element_text(size=6),
+      legend.key.size = unit(2, "mm"),
+      legend.spacing.y = unit(0.1, "cm"),
+       legend.box.spacing = unit(0, "pt"),
+      axis.title = element_text(face = "bold", size = 8),
+      axis.text = element_text(size=6)) +
+xlab(NULL) +
+theme(strip.text.x = element_text(size = 12, face = "bold"),
+    strip.background = element_rect(
+color=NA, fill=NA, size=0.5, linetype="solid")) +
+scale_fill_manual(values = c("#ffd67a","skyblue", "lightpink2", "tomato3"),
+                  name = NULL) +
+coord_flip(expand = FALSE) +
+ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
+guides(fill=guide_legend(nrow=2,byrow=TRUE))
+
+#Unify labels
+c1$Habitat = factor(c1$Habitat, levels = l2)
+#Create barplot on the side
+p2 =ggplot(c1, aes(Habitat, Count)) +
+geom_col(fill = "black",width = 0.9) +
+theme_light() +
+theme(plot.margin=unit(c(0,0,0,0), "mm"),
+      axis.ticks.y = element_blank(), 
+  axis.text.y = element_blank(), 
+      strip.text.x = element_text(size=12),
+      strip.background = element_rect(
+color=NA, fill=NA, size=0.5, linetype="solid"),
+  axis.title = element_text(face = "bold", size = 8),
+axis.text = element_text(size=6)) +
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+panel.border = element_blank(),
+axis.line = element_line(color = 'black')) +
+xlab(NULL) +
+ylab("Number of studies")+
+ylim(0, 30) + 
+theme(legend.position = "none")+
+coord_flip(expand = F) +
+ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
+
+#Plot panel
+p1 + p2 +plot_layout(widths = c(4,2))
+
+```
 # Results

### 2024-01-24 Scripts/4_Bioregion_information.R

@@ -1,0 +1,58 @@
+#Extract biogeographical regions based on coordinates
+# I have followed these posts (one is mine, the downvoted one :)
+#https://stackoverflow.com/questions/74723290/extract-biogeographical-genions-based-on-coodinates
+#https://stackoverflow.com/questions/77815238/speed-up-extraction-of-biogeographical-regions-from-shapefile-based-on-coordinat
+#Load libraries
+library(sf)
+library(dplyr)
+
+#Load data 
+data = readRDS("Data/3_Final_data/Interactions_uncounted.rds")
+
+##Read shapefile:
+bioregions <- sf::read_sf('Data/Working_files/Bio_Regions/BiogeoRegions2016.shp')
+bioregions <- st_transform(bioregions, crs = 4326) %>%  st_transform(3035)
+
+#Generate a tibble with coordinates
+coords = data %>% 
+select(Study_id, Longitude, Latitude) %>% 
+distinct() 
+
+#Prepare points for extraction
+pt <-  data.frame(x = coords$Longitude, y = coords$Latitude, id = 1:nrow(coords)) %>% 
+## add geometry column:
+st_as_sf(coords = c('x', 'y'))  %>%  ## set CRS:
+st_set_crs(4326) %>% 
+st_transform(3035)
+
+#Spatial join, by default a left join: 
+#All point are returned, those with no polygon match will have NA values
+result <- st_join(pt, bioregions)
+#Bind cols
+coords_result = bind_cols(coords, result) 
+
+#Select main cols
+#Check for NA's and fix 
+#Those are coordinates located on the Ocean
+coords_result = coords_result %>% 
+select(Study_id, Longitude, Latitude, code)
+
+#Filter NA's
+coords_result %>% 
+filter(is.na(code))
+#Fix manually here, just 5 coordinates
+coords_result = coords_result %>% 
+mutate(code = case_when(
+       Study_id == "34_Stanley" ~ "Atlantic",
+       Study_id == "37_White" ~ "Atlantic",
+       Study_id == "50_Hervias-Parejo" ~ "Mediterranean",
+        TRUE ~ code))  
+
+#Rename columns and save data
+coords_result = coords_result %>%
+rename(Bioregions = code)
+
+#Save Bioregions
+saveRDS(coords_result, "Data/Working_files/Bioregion.rds")
+
+

### 2024-01-24 Manuscript/Ms.qmd
minor edits

@@
 xlab(NULL) +
+scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
 theme(strip.text.x = element_text(size = 12, face = "bold"),
@@
 axis.line = element_line(color = 'black')) +
+scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
 xlab(NULL) +
 ylab("Number of studies")+
-ylim(0, 30) + 
 theme(legend.position = "none")+

### 2024-01-25 Manuscript/Ms.qmd
Preparing methods and fixing one species typo (it was a molusc!)

@@
         \captionsetup[figure]{labelformat=empty}
+bibliography: references.bib
 ---
@@
 
-Explore habitat types! (I think I need to fix some)
+## Taxonomy and taxonomic coverage
 
-## Taxonomy
+All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
+To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
+The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
+For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone (WFO, July, 07, 2022) as final filter of the taxonomic information -given that we use it to evaluate the approximate plant taxonomic coverage at European level- and corrected synonyms accordingly.
 
-All plant and pollinator species names were checked and standardised according to the most recent
+For pollinators:
+
+Explain unresolved species names and that we keep old names on the dataset for transparency.
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
+(\\textbf{a}) Approximate locations of the studies of the SafeNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
+The sizes of these shapes are proportional to the respective species counts.
+For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
+(\\textbf{b}) Percentage of shared occurrences of pollinators and plants across studies.
+Species on the left are highly shared across studies, while species on the right are found only in a few or a single study.
+As a reference, we have indicated the 50th percentile with a black dashed vertical line."}
 
@@
 ```
+
 # Results

### 2024-01-25 Manuscript/Ms.qmd
minor organisation edits

@@
 
+\newpage
+
 # Methods
@@
 These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-
-## Taxonomy and taxonomic coverage
-
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone (WFO, July, 07, 2022) as final filter of the taxonomic information -given that we use it to evaluate the approximate plant taxonomic coverage at European level- and corrected synonyms accordingly.
-
-For pollinators:
-
-Explain unresolved species names and that we keep old names on the dataset for transparency.
-
-## Taxonomic coverage
-
-To calculate the completeness of the dataset
-
-Check presentation from Herrera for cool graphs summarising patterns
-
-Phylo for plants and pollinators!
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%", fig.cap= "\\textbf{Figure 2}.
+## Taxonomy and taxonomic coverage
+
+All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
+To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
+The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
+For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone (WFO, July, 07, 2022) as final filter of the taxonomic information -given that we use it to evaluate the approximate plant taxonomic coverage at European level- and corrected synonyms accordingly.
+
+For pollinators:
+
+Explain unresolved species names and that we keep old names on the dataset for transparency.
+
+## Habitat type and bioclimatic region
+
+To calculate the completeness of the dataset
+
+Check presentation from Herrera for cool graphs summarising patterns
+
+Phylo for plants and pollinators!
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
 Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
@@
 Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=4, fig.height=4}
+Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=3.8, fig.height=3.8}
 #Load data (done it in earlier chunks)
@@
 
+\newpage
+
 # Results
 
+## Taxonomy and taxonomic coverage
+
+## Habitat type and bioclimatic region
+
 # Discussion
+
+\newpage
+
+# References

### 2024-01-31 Manuscript/Ms.qmd
Some progress

@@
 
-## Taxonomy and taxonomic coverage
+## Taxonomic harmonization
 
@@
 The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone (WFO, July, 07, 2022) as final filter of the taxonomic information -given that we use it to evaluate the approximate plant taxonomic coverage at European level- and corrected synonyms accordingly.
+For transparency, we have included in the database the old species name, the new assigned name, and, if the species name is uncertain (e.g., species complex or species alike).
 
-For pollinators:
+For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
 
-Explain unresolved species names and that we keep old names on the dataset for transparency.
+For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at European level by combining the recently published checklists of bees [@ghisbain2023], syrphids [@kovcic2023] and butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+
+## Taxonomic coverage
+
+To assess the completeness of plant and pollinator species in the SafeNet database at European level, we used the aforementioned checklists for plants and pollinators.
+Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
+We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
+Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
+For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
 
@@
 
-To calculate the completeness of the dataset
-
-Check presentation from Herrera for cool graphs summarising patterns
+The different sites per study were classified with an habitat by the authors.
+As these habitats are not standardize across studies, we homogenized the habitat classification with the help of land cover information and visual checks on current satellite imagery.
+The land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
+Then, we checked the coordinate sites on Google Earth to double check that the categories that we are assigning after evaluating the author's habitat classification and the land cover classification are meaningful.
+However, this is a non fully objective process and we advise authors to revise this classification if they intend to rely on this field for their analyses.
+Moreover, we assigned each study site to a biogeographical region, which are areas with similar environmental conditions that harbor similar groups of species.
+We believe this could be a useful field for organizing the different studies given the different environmental conditions and pool of species.
+The biogeographical regions were downloaded from the European Environment Agency (version 2016) and match with the different study site coordinates.
 

### 2024-02-01 Manuscript/Figure_2a.Rmd
Trying to show the year of the studies on figure 2. Not sure if it contains too much info now

--- 2024-01-25 Manuscript/Ms.qmd
@@
 ---
-title: "Mapping plant-pollinator interactions across Europe"
-latex-tinytex: false
-format:
-  pdf:
-    include-in-header: 
-      text: |
-        \usepackage{lscape}
-        \newcommand{\blandscape}{\begin{landscape}}
-        \newcommand{\elandscape}{\end{landscape}}
-        \usepackage{caption}
-        \captionsetup[figure]{labelformat=empty}
-bibliography: references.bib
+title: "Untitled"
+output: pdf_document
+date: "2024-02-01"
 ---
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Print number of countries
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-```
-
-# Abstract
-
-Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
-Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies distributed across `r country_number` countries that comprise over a million of interactions between plants and pollinators.
-The dataset includes a total of `r pollinator_spp_number` pollinator and `r plant_spp_number` plant species that accounts approximately for 30% of each of the main pollinator groups (i.e., bees, syrphids and butterflies) and 5% of flowering plants that inhabit the European continent.
-
-# Introduction
-
-1st paragraph
-
-General introduction of how global change impacts plant-pollinator interactions
-
-Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions
-
-Highlight the relevance of large scale datasets to understand biological phenomena
-
-2nd paragraph
-
-3rd paragraph
-
-4paragraph
-
-Introduce research questions
-
-LIST THEM HERE (Main ideas so far)
-
-Questions that we would like to answer:
-
-1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity
-
-2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?
-
-\newpage
-
-# Methods
-
-## Dataset description
-
-This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
-This dataset consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
-The dataset contains a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
-These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the SafeNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Percentage of shared occurrences of pollinators and plants across studies.
-Species on the left are highly shared across studies, while species on the right are found only in a few or a single study.
-As a reference, we have indicated the 50th percentile with a black dashed vertical line."}
-
-#Load data (done it in previous chunk)
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.01) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -12, hjust=0.1, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-p1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.015, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  plot.title = element_text(size = 8, face = "bold"), 
-  axis.text.y = element_text(size=6),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-xlab(NULL) +
-ylab(NULL) + 
-ggtitle("Pollinators")+
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.25) + 
-annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
-annotate(geom="text", x=20, y=0.89, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
-scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-annotate(geom="text", x=-500, y=1.15, label="(b)",color="black", size = 3.5) +
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-#Plot
-p2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.015, fill = "#009E73") +
-ylab("Shared plants (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         axis.title.x = element_text(face = "bold", size = 8),            plot.title = element_text(size = 8, face = "bold"),
-         axis.text.y = element_text(size=6),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-xlab("Species") +
-ylab(NULL) + 
-ggtitle("Plants") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.25) + 
-annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-annotate(geom="text", x=20, y=0.73, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 1.5)+
-annotate(geom="text", x=20, y=-0.08, label="Common spp."
-      ,color="black",  size = 1.5) +
-annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
-      ,color="black",  size = 1.5) + 
-scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")
-
-#Prepare panel with plots
-plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Occurrence across studies (%)") + 
-theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 1))
-
-```
-
-## Taxonomy and taxonomic coverage
-
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone (WFO, July, 07, 2022) as final filter of the taxonomic information -given that we use it to evaluate the approximate plant taxonomic coverage at European level- and corrected synonyms accordingly.
-
-For pollinators:
-
-Explain unresolved species names and that we keep old names on the dataset for transparency.
-
-## Habitat type and bioclimatic region
-
-To calculate the completeness of the dataset
-
-Check presentation from Herrera for cool graphs summarising patterns
-
-Phylo for plants and pollinators!
-
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
@@
 Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=3.8, fig.height=3.8}
+Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=6, fig.height=8}
+
+#Read interaction data 
+data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
 #Load data (done it in earlier chunks)
@@
 ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
+guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
+ggtitle("(a)") +
+theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
 
@@
 
+data2 = data %>%  select(Study_id, Year) %>% 
+distinct() %>% 
+group_by(Year) %>% 
+summarise(Counts = n())
+
+p3 = ggplot(data2, aes(Year, Counts)) +
+geom_col(fill = "gray10", width = 0.9) +
+theme_light() +
+theme(plot.margin=unit(c(0,0,0,0), "mm"),
+  axis.title.y = element_text(face = "bold", size = 8, vjust = -2),
+  axis.title.x = element_text(face = "bold", size = 8),
+  axis.text.y = element_text(size = 6),
+  axis.text.x = element_text(size = 6,  angle = 90), 
+  strip.text.x = element_text(size=12),
+  strip.background = element_rect(
+  color="black", fill=NA, size=0.5, linetype="solid")) + 
+ylab("Number of studies") +
+coord_cartesian(expand = F, clip = "off") + 
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(),
+  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+panel.border = element_blank(),
+axis.line = element_line(color = 'black')) +
+scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
+scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
+ggtitle("(b)") +
+theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
+
+a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
 #Plot panel
-p1 + p2 +plot_layout(widths = c(4,2))
+p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)
 
 ```
-
-\newpage
-
-# Results
-
-## Taxonomy and taxonomic coverage
-
-## Habitat type and bioclimatic region
-
-# Discussion
-
-\newpage
-
-# References

### 2024-02-01 Manuscript/Ms.qmd
Trying to show the year of the studies on figure 2. Not sure if it contains too much info now

@@
 
-The different sites per study were classified with an habitat by the authors.
-As these habitats are not standardize across studies, we homogenized the habitat classification with the help of land cover information and visual checks on current satellite imagery.
-The land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
-Then, we checked the coordinate sites on Google Earth to double check that the categories that we are assigning after evaluating the author's habitat classification and the land cover classification are meaningful.
-However, this is a non fully objective process and we advise authors to revise this classification if they intend to rely on this field for their analyses.
-Moreover, we assigned each study site to a biogeographical region, which are areas with similar environmental conditions that harbor similar groups of species.
-We believe this could be a useful field for organizing the different studies given the different environmental conditions and pool of species.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and match with the different study site coordinates.
-
-Phylo for plants and pollinators!
+The different sites per study were classified with a habitat type by the authors.
+As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
+For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
+Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the SafeNet database (see habitat type definition in supplementary material).
+These categories allow a quick comparison and understanding of the habitat types from the database.
+However, we advise authors to revise this classification if they intend to rely on this field for their analyses as this is a non-fully objective process.
+Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
+Thus, to allow authors to explore set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
+The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
 
@@
 ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
+guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
+ggtitle("(a)") +
+theme(plot.title = element_text(size = 14 ,hjust = -0.25, vjust = -2.5))
 
@@
 
+data2 = data %>%  select(Study_id, Year) %>% 
+distinct() %>% 
+group_by(Year) %>% 
+summarise(Counts = n())
+
+p3 = ggplot(data2, aes(Year, Counts)) +
+geom_col(fill = "gray10", width = 0.9) +
+theme_light() +
+theme(plot.margin=unit(c(0,0,0,0), "mm"),
+  axis.title.y = element_text(face = "bold", size = 8, vjust = -1),
+  axis.title.x = element_text(face = "bold", size = 8),
+  axis.text.y = element_text(size = 6),
+  axis.text.x = element_text(size = 6,  angle = 90), 
+  strip.text.x = element_text(size=12),
+  strip.background = element_rect(
+  color="black", fill=NA, size=0.5, linetype="solid")) + 
+ylab("Number of studies") +
+coord_cartesian(expand = F, clip = "off") + 
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(),
+  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+panel.border = element_blank(),
+axis.line = element_line(color = 'black')) +
+scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
+scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
+ggtitle("(b)") +
+theme(plot.title = element_text(size = 14,hjust = -0.3, vjust = 2))
+
+a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
 #Plot panel
-p1 + p2 +plot_layout(widths = c(4,2))
+p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)
 
@@
 # References
+
+::: {#refs}
+:::
+
+# Supplementary material
+
+Create definitions here based on Corine
+
+1)Pastures: Any type of low growing plant community that is highly influenced by human disturbance.
+For instance, agriculture, mowing, moderate to high grazing or urban environment.
+Note that this category also includes old pastures with regrowth of woody vegetation.
+
+2)Agricultural margings: Sides of crops that can include any type of vegetation from low growing plants to trees.
+
+3)Agricultural land: Includes any type of crop and any type of vegetation growing within them.
+
+4)Semi-natural grassland: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
+
+5)Ruderal vegetation: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
+
+6)Sclerophyllous vegetation: Any type of system with a dominant shrub community adapted to drought.
+Typical of the Mediterranean region.
+Note, that we have include in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
+
+7) Green urban areas: Parks, private gardens or small pastures within an  urban setting. For simplicity, we have also included botanical gardens.
+
+8) Forest/woodland understory: Any plant community sampled under a wooded group of plants 
+The forest could be embedded in an agricultural setting or in a fully natural scenario.
+We have included here agro-forestry areas, open forest to dense forest but try to exclude forest that contains typical sclerophyllous vegetation.
+
+9) Riparian vegetation: Plant communities growing on river margins.
+
+10): Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
+
+11) Beaches, dunes, sands: Plant communities growing on sandy soil.
+
+12) Moors and heathland: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.

### 2024-02-05 Manuscript/Untitled.Rmd
phylo plants

--- 2024-02-01 Manuscript/Figure_2a.Rmd
@@
 
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
+p1 = ggplot(c2, aes(fill = Pollinator_order, x = Habitat, y = Proportion)) + 
+  geom_bar(position = position_stack(reverse = TRUE), stat = "identity") +
   theme_bw() +
   theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
+        plot.margin = unit(c(0, 0, 0, 0), "mm"),
+        legend.text = element_text(size = 6),
         legend.key.size = unit(2, "mm"),
@@
         axis.title = element_text(face = "bold", size = 8),
-      axis.text = element_text(size=6)) +
-xlab(NULL) +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
+        axis.text = element_text(size = 6)) +
+  ylab(NULL) +
+  scale_x_continuous(breaks = c(0.25, 0.5, 0.75), labels = c("0.25", "0.5", "0.75")) +
   theme(strip.text.x = element_text(size = 12, face = "bold"),
         strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = c("#ffd67a","skyblue", "lightpink2", "tomato3"),
+          color = NA, fill = NA, size = 0.5, linetype = "solid")) +
+  scale_fill_manual(values = c("#ffd67a", "skyblue", "lightpink2", "tomato3"),
                     name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
+  ggforce::facet_col(~ factor(Group, levels = c("Habitat", "Bioregion")), space = "free", scales = "free_x") +
+  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
   ggtitle("(a)") +
-theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
-
+  theme(plot.title = element_text(size = 10, face = "bold", hjust = -0.25, vjust = 1))
 #Unify labels
@@
 #Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "black",width = 0.9) +
+p2 = ggplot(c1, aes(x = Habitat, y = Count)) +
+  geom_col(fill = "black", width = 0.9) +
   theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
+  theme(plot.margin = unit(c(0, 0, 0, 0), "mm"),
+        axis.ticks.x = element_blank(), 
+        axis.text.x = element_blank(), 
+        strip.text.y = element_text(size = 12),
         strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
+          color = NA, fill = NA, size = 0.5, linetype = "solid"),
         axis.title = element_text(face = "bold", size = 8),
-axis.text = element_text(size=6)) +
+        axis.text = element_text(size = 6)) +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
@@
         axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
+  scale_x_continuous(breaks = c(0, 10, 20), labels = c("0", "10", "20"), limits = c(0, 30)) +
+  ylab(NULL) +
+  xlab("Number of studies") +
+  theme(legend.position = "none") +
+  ggforce::facet_row(~ factor(Group, levels = c("Habitat", "Bioregion")), space = "free", scales = "free_x")
 
@@
 
-a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
+p1/p2 + p3
+
+#a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
 #Plot panel
-p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)
+#p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)
 

### 2024-02-05 Scripts/3_1_Plant_taxonomy_GBIF.R

@@ -149,7 +149,7 @@ mutate(Fixed_name = case_when(
   Fixed_name == "Rosa pimpernellifolia" ~ "Rosa spinosissima", #synonym
   Fixed_name == "Sysimbrium officinale" ~ "Sisymbrium officinale", #typo
   Fixed_name == "Escallonia macrantha" ~ "Escallonia rubra", #synonym
-  Fixed_name == "Petunia xatkinsiana" ~ "Petunia ×atkinsiana", #typo
+  Fixed_name == "Petunia xatkinsiana" ~ "Petunia hybrida", #synonym
   Fixed_name == "Malus pumila" ~ "Malus domestica", #synonym
   Fixed_name == "Weigela kosteriana_variegata" ~ "Weigela", #unknown, fix to higherank...
   Fixed_name == "Polygonum lapathifolium" ~ "Persicaria lapathifolia", #synonym
@@ -414,6 +414,13 @@ mutate(Genus =
     if_else(Genus == word(Plant_data1$Accepted_name, 1), 
             Genus, word(Plant_data1$Accepted_name, 1)))
 
+#Some genera missing, recover form species names when rank genus
+Plant_data1 = Plant_data1 %>% 
+mutate(Genus = 
+if_else(is.na(Genus) & Rank =="GENUS", 
+            word(Plant_data1$Accepted_name, 1), Genus))
+
+
 #Similar case
 #Genus level records are not added as accepted
 Plant_data1 = Plant_data1 %>% 
@@ -421,12 +428,20 @@ mutate(Accepted_name =
 if_else(is.na(Accepted_name) & Matchtype== "EXACT" & Rank =="GENUS", 
             Canonical_name, Accepted_name))
 
-
 #Keep subspecies rank as species
 Plant_data1 = Plant_data1 %>% 
 mutate(Rank = case_when(Rank == "SUBSPECIES" ~ "SPECIES",
                         T ~ Rank))
 
+#Add for Thesium genus, Santalaceae as accepted family
+Plant_data1 = Plant_data1 %>% 
+mutate(Family = case_when(Genus == "Thesium" ~ "Santalaceae",
+                        T ~ Family))
+
+#Add for Heliotropium genus, Boraginacea as accepted family
+Plant_data1 = Plant_data1 %>% 
+mutate(Family = case_when(Genus == "Heliotropium" ~ "Boraginaceae",
+                        T ~ Family))
 #--------------------------------------#
 #8) Save pollinator taxonomy-------
 #--------------------------------------#

### 2024-02-07 Manuscript/Figure3.Rmd
Good day, aiming for Global Ecol and Biogeography. And probably lacking 2 graphs.

---
output: pdf_document
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=8, fig.height=8}

library(ggplot2)
library(ggtree)
library(ggnewscale)
library(ggtreeExtra)
library(cowplot)

#Save data to plot it in markdown file
c = readRDS("../Data/Working_files/Figures/data_figure3b.rds")
tree_family_interactions = readRDS("../Data/Working_files/Figures/data_figure3a1.rds")
plant_fam_interactions = readRDS("../Data/Working_files/Figures/data_figure3a2.rds")

# The circular layout tree
p = ggtree(tree_family_interactions, 
    layout="fan", size=0.1, open.angle=5, alpha=0.5)+ 
geom_tippoint(aes(size= Family_species), colour='cyan4') +
geom_tiplab(linetype='dashed', linesize=.05, 
    size=1.75, color= "black", offset = -45, fontface="bold") +
labs("log(Species)")+
scale_size(name = "log(Species #)") +
ggtitle("a)") +
theme(plot.title = element_text(vjust = -31, hjust = 0.12),
      legend.position = "bottom")

#Create barplot around circular tree
p1 = p + 
new_scale_fill() +
geom_fruit(data=plant_fam_interactions, geom=geom_bar,
   mapping=aes(y=Plant_family1, x=Family_interactions1),
   pwidth=0.38, 
   orientation="y", 
   stat="identity", 
   color = "black") +
theme(legend.margin=margin(0,0,0,0),
   legend.box.margin=margin(-50,-50,-50,-50),
   plot.margin = margin(-15, -15, -10, -10, "pt")) +
scale_size(name = "log(species number)", range=c(0,4))

p2 = ggplot() + 
geom_col(data = c, 
  aes(x = reorder(Family_names, Spp_per_family), 
  y = log(Spp_per_family), fill = Group), position=position_dodge(0.1),width=1) +
xlab(NULL) +
ylab("log(species number)") +
theme_bw() +
theme(axis.text.y = element_text(size = 5, face = "bold", color = "black"),
      panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
coord_flip(expand = F) +
scale_fill_manual(name="", values = c("black", "cyan4")) +
  theme(plot.margin = margin(-10, 0, -10, -10, "pt")) +
ggtitle("b)")+
theme(legend.margin=margin(0,0,0,0),
   legend.box.margin=margin(-10,-10,-10,-10)) +
  theme(legend.key.size = unit(0.3, 'cm'),
        legend.position = "bottom")

p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
plot_grid(p1,p2_1, rel_widths = c(16, 7), rel_heights  = c(9, 1), align = "none")

```

### 2024-02-07 Manuscript/Ms.qmd
Good day, aiming for Global Ecol and Biogeography. And probably lacking 2 graphs.

@@
 These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+
+## Data structure?
+
+TABLE? Describe cols?
 
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
+a) Phylogenetic tree of plant families in the Safenet database, showing the number of species per family in logarithmic scale (blue circles) and the total number of interactions recorded per family (gray bars).
+b) Comparison of the top 50 families with the highest number of species in the SafeNet database against the total number of species recorded for those families at the European level.
+Species counts are displayed in logarithmic scale for better visualization.", fig.width=8, fig.height=8}
+
+library(ggplot2)
+library(ggtree)
+library(ggnewscale)
+library(ggtreeExtra)
+library(cowplot)
+
+#Save data to plot it in markdown file
+c = readRDS("../Data/Working_files/Figures/data_figure3b.rds")
+tree_family_interactions = readRDS("../Data/Working_files/Figures/data_figure3a1.rds")
+plant_fam_interactions = readRDS("../Data/Working_files/Figures/data_figure3a2.rds")
+
+# The circular layout tree
+p = ggtree(tree_family_interactions, 
+    layout="fan", size=0.1, open.angle=5, alpha=0.5)+ 
+geom_tippoint(aes(size= Family_species), colour='cyan4') +
+geom_tiplab(linetype='dashed', linesize=.05, 
+    size=1.75, color= "black", offset = -45, fontface="bold") +
+labs("log(Species)")+
+scale_size(name = "log(Species #)") +
+ggtitle("a)") +
+theme(plot.title = element_text(vjust = -31, hjust = 0.12, size = 14),
+      legend.position = "bottom")
+
+#Create barplot around circular tree
+p1 = p + 
+new_scale_fill() +
+geom_fruit(data=plant_fam_interactions, geom=geom_bar,
+   mapping=aes(y=Plant_family1, x=Family_interactions1),
+   pwidth=0.38, 
+   orientation="y", 
+   stat="identity", 
+   color = "black") +
+theme(legend.margin=margin(0,0,0,0),
+   legend.box.margin=margin(-50,-50,-50,-50),
+   plot.margin = margin(-15, -15, -10, -10, "pt")) +
+scale_size(name = "log(species number)", range=c(0,4))
+
+p2 = ggplot() + 
+geom_col(data = c, 
+  aes(x = reorder(Family_names, Spp_per_family), 
+  y = log(Spp_per_family), fill = Group), position=position_dodge(0.1),width=1) +
+xlab(NULL) +
+ylab("log(species number)") +
+theme_bw() +
+theme(axis.text.y = element_text(size = 5, face = "bold", color = "black"),
+      panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+coord_flip(expand = F) +
+scale_fill_manual(name="", values = c("black", "cyan4")) +
+  theme(plot.margin = margin(-10, 0, -10, -10, "pt")) +
+ggtitle("b)")+
+theme(legend.margin=margin(0,0,0,0),
+   legend.box.margin=margin(-10,-10,-10,-10)) +
+  theme(legend.key.size = unit(0.3, 'cm'),
+        legend.position = "bottom",
+        plot.title = element_text(size=14))
+
+p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
+plot_grid(p1,p2_1, rel_widths = c(16, 7), rel_heights  = c(9, 1), align = "none")
+
+```
+
+FIGURE 4 POLLINATOR COVERAGE
+
+FIGURE 5? CLIFFHANGER OF NETWORK METRICS
+
 ## Taxonomy and taxonomic coverage
@@
 # Discussion
+
+Geographic and taxonomic biases!
 
@@
 
-Create definitions here based on Corine
+Supplementary text 1
 
-1)Pastures: Any type of low growing plant community that is highly influenced by human disturbance.
-For instance, agriculture, mowing, moderate to high grazing or urban environment.
+Habitat definitions:
+
+1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
+
+2) **Agricultural margings**: Sides of crops that can include any type of vegetation from low growing plants to trees.
+
+3) **Green urban areas**: Parks, private gardens or small pastures within an  urban setting. Botanical gardens are included in this category.
+
+4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
+
+5) **Forest/woodland understory**: Any plant community sampled under a wooded group of plants.
+The forest could be embedded in an agricultural setting or in a fully natural scenario.
+We have included here agro-forestry areas and open to dense forest.
+Note that we have excluded from this category forest that contains sclerophyllous vegetation.
+
+6) **Semi-natural grassland**: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
+
+7) **Pastures**: Any type of low growing plant community that is highly influenced by human disturbance.
+For instance, agriculture, mowing, moderate to high grazing or urban environments.
 Note that this category also includes old pastures with regrowth of woody vegetation.
 
-2)Agricultural margings: Sides of crops that can include any type of vegetation from low growing plants to trees.
-
-3)Agricultural land: Includes any type of crop and any type of vegetation growing within them.
-
-4)Semi-natural grassland: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
-
-5)Ruderal vegetation: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-
-6)Sclerophyllous vegetation: Any type of system with a dominant shrub community adapted to drought.
+8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
 Typical of the Mediterranean region.
@@
 
-7) Green urban areas: Parks, private gardens or small pastures within an  urban setting. For simplicity, we have also included botanical gardens.
+9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
 
-8) Forest/woodland understory: Any plant community sampled under a wooded group of plants 
-The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included here agro-forestry areas, open forest to dense forest but try to exclude forest that contains typical sclerophyllous vegetation.
+10) **Riparian vegetation**: Plant communities growing on river margins.
 
-9) Riparian vegetation: Plant communities growing on river margins.
+10) **Natural grassslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
 
-10): Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
-
-11) Beaches, dunes, sands: Plant communities growing on sandy soil.
-
-12) Moors and heathland: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
+12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.

### 2024-02-08 Manuscript/Ms.qmd
draft 1st paragraph intro

@@
 
-1st paragraph
-
-General introduction of how global change impacts plant-pollinator interactions
-
-Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions
-
-Highlight the relevance of large scale datasets to understand biological phenomena
+The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
+Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity [REF?] and human economy [REF?].
+The synergistic effect of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016].
+Documenting plant and pollinator diversity is the first step to conserve their biodiversity [REF].
+Europe is one of the continents with more information about their plant and pollinator species as indicated by the recently available checklists of plants [REF] and pollinators [REF].
+However, despite the increasing number of studies documenting their interactions at European level, we lack robust datasets of their ecological interactions at European level, which has crucial implications to understand the different biological phenomena involved around these twp trophic levels.
 
@@
 
+Plant-pollinator databases?
+
+Mangal/Web of life/the new one that has a bit of everything...
+
 3rd paragraph
 
+Threatened species? Red lists?
+
+Discuss actual TRENDS!
+
 4paragraph
-
-Introduce research questions
-
-LIST THEM HERE (Main ideas so far)
-
-Questions that we would like to answer:
-
-1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity
-
-2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?
 
@@
 
+Something always of interest: provide % of interactions of _Apis mellifera_?
+
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.

### 2024-02-09 Manuscript/Ms.qmd
Paragraph 2 outline

--- 2024-02-07 Manuscript/Ms.qmd
@@
 
-1st paragraph
+The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
+Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
+The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016].
+Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning.
+However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016;  @poisot2021] that limit our ability to effectively protect their biodiversity.
 
-General introduction of how global change impacts plant-pollinator interactions
+Integrating ecological data is allowing researchers to understand biological patterns [REF].
+Europe is one of the continents with larger information of plant and pollinator biodiversity data available that is proven by the different checklists of species of plants [ref], bees [ref], syrphids [REF] or butterflies [REF].
+However,
 
-Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions
+SOMETHING ABOUT THE DIFFICULTY OF CAPTURING ECOLOGICAL INTERACTIONS
+Jordano Funct Ecol?
 
-Highlight the relevance of large scale datasets to understand biological phenomena
+connectance/ecological networks
 
-2nd paragraph
+what we know-->Plant-pollinator databases? -->Mangal/Web of life/the new one that has a bit of everything...
+
+Maybe add another sentence about data integration
+https://doi.org/10.1073/pnas.2018093118
 
@@
 
+Threatened species? Red lists?
+
+Discuss actual TRENDS!
+
 4paragraph
-
-Introduce research questions
-
-LIST THEM HERE (Main ideas so far)
-
-Questions that we would like to answer:
-
-1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity
-
-2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?
 
@@
 
+NOT FULLY HAPPY with it
+
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.

### 2024-02-12 Manuscript/Ms.qmd
adding idea
2nd paragraph done, now I need something a bit more into detail in the 3rd...

--- 2024-02-07 Manuscript/Ms.qmd
@@
 title: "Mapping plant-pollinator interactions across Europe"
-latex-tinytex: false
+
 format:
@@
 bibliography: references.bib
+latex-tinytex: false
+link-citations: TRUE
+linkcolor: RoyalBlue
+urlcolor: RoyalBlue
 ---
@@
 
-1st paragraph
+The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
+Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
+The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016; @eichenberg2021].
+Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
+However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024] that limit our ability to effectively protect their biodiversity.
 
-General introduction of how global change impacts plant-pollinator interactions
-
-Maybe expand on some drivers? Climate change, habitat fragmentation, agricultural intensification, urbanization, pollution, pesticides and species' invasions
-
-Highlight the relevance of large scale datasets to understand biological phenomena
-
-2nd paragraph
+Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
+While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators.
+However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
+Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
+Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
+A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., sampling effort or methodology), which hampers our understanding of plant-pollinator interactions.
+Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.
 
@@
 
+Low connectance?
+
+Species occurrences
+
+Threatened species? Red lists?
+
+Discuss actual TRENDS!
+
 4paragraph
 
-Introduce research questions
-
-LIST THEM HERE (Main ideas so far)
-
-Questions that we would like to answer:
-
-1\) What are the most common plant a pollinator species? Are those shared across networks? Most common interactions across Europe? Interaction fidelity
-
-2\) Is generalization the rule? Or specialization? How this impacts indirect interactions? Go in the direction of pollinator importance?
+Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonize interaction data of plants and pollinators at European level.
 
@@
 These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+
+Define study area!
 
@@
 
+Contextualise European diversity of plant and pollinator species with global biodiveristy
+
 \newpage

### 2024-02-13 Manuscript/Ms.qmd
preparing authors list

@@
 ---
-title: "Mapping plant-pollinator interactions across Europe"
-
+title: "SafeNet: A European metaweb of plant-pollinator interactions"
 format:
@@
 pull()
+#Check number of distinct sampling sites
+n_sites = data %>% 
+select(Latitude, Longitude) %>% 
+distinct() %>% 
+nrow(.)
+#Check number of distinct bioregions
+n_bioregion = data %>% 
+select(Bioregion) %>% 
+distinct()  %>% 
+nrow(.)
 ```
@@
 
+## Motivation
+
 Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
@@
 One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
+Here, we present the SafeNet database, a continent-level database of geo-referenced plant-pollinator interactions with the aim to help researchers to explore the impacts of global change and guide future conservation planning.
+
+## Main Types of Variables Included
+
+SafeNet contains over a million of interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
+In addition, information about sampling methodology, habitat, type of bio-climatic region and further taxonomic information for both plant and pollinator species are also provided.
+
+## Spatial location and grain
+
+The database contains `r n_sites` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bioclimatic regions.
+All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
+
+## Time period and grain
+
+## Major taxa and level of measurement
+
+The database includes...
+
+## Software format
+
+The database was built with R software and is stored as ".rds" and ".csv" formats. The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINK.
+
 Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies distributed across `r country_number` countries that comprise over a million of interactions between plants and pollinators.
@@
 Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024] that limit our ability to effectively protect their biodiversity.
+However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
 
@@
 
-Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonize interaction data of plants and pollinators at European level.
+Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
+The primary focus of the pollinator taxonomic groups of this database are Hymenoptera, Diptera, Lepidoptera and Coleoptera.
 
@@
 
-Define study area!
-
-## Data structure?
+Define study area//Data acquisition//Data structure?
 

### 2024-02-13 Scripts/1_Processing_networks/48_Lara-Romero.R

@@ -98,7 +98,7 @@ rename(Metadata_fields = rowname, Metadata_info= V1) %>% as_tibble()
 
 #Prepare authorship data ----
 Authorship <- data.frame(
-  Coauthor_name = c("Carlos Lara-Romero", "Jose M iriondo", "Javier Morente-López"),
+  Coauthor_name = c("Carlos Lara-Romero", "José M. Iriondo", "Javier Morente-López"),
   Orcid = c("0000-0003-0962-0567", "0000-0003-2710-3889", "0000-0001-9141-8581"),
   E_mail = c("carlos.lara@urjc.es", "carlos.lara@urjc.es", "carlos.lara@urjc.es"))
 

### 2024-02-14 Manuscript/Ms.qmd
List of authors added to MS. Still need affiliations but good for now.

@@
 ---
-title: "SafeNet: A European metaweb of plant-pollinator interactions"
+title: "SafeNet: A European database of plant-pollinator interactions"
 format:
@@
 nrow(.)
+#Read author list 
+authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
+#Print authors
+authors = authors %>% 
+summarise(col = paste(Coauthor_name, collapse=", ")) %>% 
+pull()
 ```
+
+`r authors`
+
+**Corresponding author**= barragansljose@gmail.com
+*All authors excluding first, second and last are ordered alphabetically
+
+\newpage
 

### 2024-02-16 Manuscript/Ms.qmd
Author list ready

--- 2024-02-13 Manuscript/Ms.qmd
@@
 nrow(.)
+#Read author list 
+authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
+#Print authors
+authors = authors %>% 
+summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
+pull()
+#Read affiliation list 
+aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
+aff_list = aff_list %>% 
+summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
+pull()
+
 ```
+
+\vspace*{-12mm}
+
+`r authors`
+
+**Corresponding author**= barragansljose@gmail.com\
+*All authors excluding the first three and last are ordered alphabetically*
+
+\begingroup
+\small
+
+`r aff_list`
+
+\small
+\endgroup
 

### 2024-02-16 Scripts/1_Processing_networks/30_Smith.R

@@ -98,7 +98,7 @@ Metadata = as.data.frame(t(Metadata)) %>%
 Authorship <- tibble(
   Coauthor_name = c("Henrik G. Smith", "Maj Rundlöf",
                     "Ante Vujić", "Ricardo Bommarco"),
-  Orcid = c("", "0000-0003-3014-1544", "0000-0002-8819-8079",
+  Orcid = c("0000-0002-2289-889X", "0000-0003-3014-1544", "0000-0002-8819-8079",
             "0000-0001-8888-0476"),
   E_mail = c("henrik.smith@biol.lu.se", "maj.rundlof@biol.lu.se",
              "vujic_ante@yahoo.com", "Riccardo.Bommarco@slu.se"))

### 2024-02-16 Scripts/1_Processing_networks/37_White.R

@@ -191,8 +191,8 @@ Metadata = as.data.frame(t(Metadata)) %>%
 
 #Prepare authorship data ----
 Authorship <- data.frame(
-  Coauthor_name = c("Cian White", "Jane Stout"),
-  Orcid = c(NA, "0000-0002-2027-0863"),
+  Coauthor_name = c("Cian White", "Jane C. Stout"),
+  Orcid = c("0000-0002-7126-1916", "0000-0002-2027-0863"),
   E_mail = c("ciwhite@tcd.ie", "stoutj@tcd.ie"))
 
 #Save data ----

### 2024-02-19 Manuscript/Ms.qmd
Abstract ready!

@@
         \captionsetup[figure]{labelformat=empty}
+
 bibliography: references.bib
@@
 
+#Interactions at species level
+species_level = data %>% 
+filter(Plant_rank == "SPECIES") %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Plant_status == "ACCEPTED") %>% 
+filter(Pollinator_status == "ACCEPTED") %>% 
+nrow(.)
+species_level_int = species_level / nrow(data) * 100
+#Read here plant and pollinator coverage
+bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
+lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
+syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
+plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
+
 ```
 
-\vspace*{-12mm}
+\vspace*{-14mm}
 
@@
 
+\pagebreak
+
 # Abstract
+\vspace*{-5mm}
 
-## Motivation
-
-Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
+**Motivation:** Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
 However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
 One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
-Here, we present the SafeNet database, a continent-level database of geo-referenced plant-pollinator interactions with the aim to help researchers to explore the impacts of global change and guide future conservation planning.
+Here, we present the SafeNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
+This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
+We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species and their interactions.
 
-## Main Types of Variables Included
+**Main Types of Variables Included:** SafeNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
+In addition, information about sampling methodology, habitat type, bio-climatic region and further taxonomic rank information for both plant and pollinator species are also provided.
 
-SafeNet contains over a million of interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat, type of bio-climatic region and further taxonomic information for both plant and pollinator species are also provided.
-
-## Spatial location and grain
-
-The database contains `r n_sites` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bioclimatic regions.
+**Spatial location and grain:** The database contains `r n_sites` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
 All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
 
-## Time period and grain
+**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`.
+All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
 
-## Major taxa and level of measurement
+**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
+The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants.
+The main focus of this database are insect pollinators and accounts for `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
 
-The database includes...
+**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats.
+The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
 
-## Software format
-
-The database was built with R software and is stored as ".rds" and ".csv" formats. The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINK.
-
-Here, we present a dataset of plant-pollinator interactions at European level that consists of `r study_number` studies distributed across `r country_number` countries that comprise over a million of interactions between plants and pollinators.
-The dataset includes a total of `r pollinator_spp_number` pollinator and `r plant_spp_number` plant species that accounts approximately for 30% of each of the main pollinator groups (i.e., bees, syrphids and butterflies) and 5% of flowering plants that inhabit the European continent.
+**KEYWORDS**\
+\begingroup
+\small
+plant-pollinator interactions, flowering plants, Angiosperms, pollinators, Hymenoptera, Diptera, Lepidoptera, Coleoptera, Europe
+\endgroup
+\pagebreak
 

### 2024-02-19 Scripts/3_2_Plant_coverage.R

@@ -363,6 +363,8 @@ total_coverage_non_wind =
   plant_spp_list/all_spp_filtered *100
 total_coverage_non_wind
 
+saveRDS(total_coverage_non_wind, "Data/Manuscript_info/plant_coverage.RData")
+
 #Remember to trust this if:
 plant_checks
 

### 2024-02-20 Manuscript/Ms.qmd
3rd paragraph... some progress

@@
 
+#Main orders percentage
+o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
+main_orders = data %>% 
+filter(Pollinator_order %in% o)
+main_orders_percent = nrow(main_orders) / nrow(data) * 100
+
 ```
@@
 Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., sampling effort or methodology), which hampers our understanding of plant-pollinator interactions.
+A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., habitat type or floral availability), which hampers our understanding of plant-pollinator interactions.
 Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.
 
-3rd paragraph
-
-Low connectance?
+The interactions between different plant and pollinator species within a community form complex networks.
+Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of the interaction ecology of the different taxonomic groups across space [REF] and environmental gradients [@saunders2023].
+In addition, environmental gradients has served as useful tool to explore the impacts of climate on phenological change [REF:Byers 2017].
+Importantly, macro-ecological analyses of plant-pollinator networks rarely account for differences in sampling effort [REF] or sampling methodology [REF], ignoring important components that can largely influence our view of plant-pollinator networks [REF].
+In addition, despite how relevant are interactions to understand plant-pollinator communities, most communities remain under-sampled [REFERENCES] due to XXXXXXXSpecies occurrences?.
+Thus, large scale datasets of plant-pollinator interactions so researchers can account for differences across studies...
 
@@
 Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups of this database are Hymenoptera, Diptera, Lepidoptera and Coleoptera.
+The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
+These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, which comprise almost the totality of recorded interactions in the SafeNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+
+SafeNet provides a comprehensive overview of plant-pollinator interactions at European level
 
@@
 
-This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
-This dataset consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
-The dataset contains a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+This European metaweb consist of phyto-centric species level networks of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
+This dataset contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
+The dataset includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
 The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).

### 2024-02-21 Manuscript/Ms.qmd
edits

@@
 Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., habitat type or floral availability), which hampers our understanding of plant-pollinator interactions.
+A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., habitat type or floral availability), which hampers our understanding of plant-pollinator interactions within their ecological context.
 Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.
@@
 The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of the interaction ecology of the different taxonomic groups across space [REF] and environmental gradients [@saunders2023].
-In addition, environmental gradients has served as useful tool to explore the impacts of climate on phenological change [REF:Byers 2017].
-Importantly, macro-ecological analyses of plant-pollinator networks rarely account for differences in sampling effort [REF] or sampling methodology [REF], ignoring important components that can largely influence our view of plant-pollinator networks [REF].
-In addition, despite how relevant are interactions to understand plant-pollinator communities, most communities remain under-sampled [REFERENCES] due to XXXXXXXSpecies occurrences?.
-Thus, large scale datasets of plant-pollinator interactions so researchers can account for differences across studies...
-
-Species occurrences
-
-Threatened species? Red lists?
-
-Discuss actual TRENDS!
-
-4paragraph
+Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @saunders2023].
+Undoubtedly, macro-ecological studies of plant-pollinator networks makes significant contributions to knowledge.
+However, they often fail to account for differences in sampling effort and sampling methodology, impacting our view and understanding of plant-pollinator interactions [REFS].
+In addition, these networks tend to be incomplete snapshots of the reality because insufficient sampling [@olesen2011; @chacoff2012; @jordano2016], highlighting that strong sampling efforts are needed to capture the missing links of these ecological networks.
+Thus, providing a high number of informative plant-pollinator networks can help researchers to control for a larger amount of ecological variables when exploring macro-ecological patterns.
+Further, these different networks require from a taxonomic harmonization process that is generally not shared or reusable [REF].
+Providing open, informative and reusable data is neccesary to set the foundation of future plant-pollinator research.
 

### 2024-02-23 Manuscript/Ms.qmd
Intro-->Ready. Polishing methods and jumping to results

--- 2024-02-19 Manuscript/Ms.qmd
@@
 This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
-We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species and their interactions.
+We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.
 
@@
 
-**Spatial location and grain:** The database contains `r n_sites` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
+**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
 All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
@@
 Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-A major drawback of these global databases is the limited number of studies at regional scales and the lack of additional information (e.g., sampling effort or methodology), which hampers our understanding of plant-pollinator interactions.
-Thus, in a scenario of rapid environmental change, integrating robust informative plant-pollinator interaction data is essential for evaluating our current knowledge at the European level, guiding conservation planning, and establishing a solid foundation for global change research.
+Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
+Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
 
-3rd paragraph
-
-Low connectance?
-
-Species occurrences
-
-Threatened species? Red lists?
-
-Discuss actual TRENDS!
-
-4paragraph
+The interactions between different plant and pollinator species within a community form complex networks.
+Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023].
+Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
+For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
+Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
+Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes at large scales.
 
 Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups of this database are Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-
-\newpage
+The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
+These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the SafeNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+SafeNet aims to cover a wide range of taxonomic groups and habitats while also provides other variables of interest that allow to better control for the ecological context and sampling methods.
+In addition, SafeNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
+We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
 
@@
 
+## Data acquisition
+
+This database is the result of one of the working packages of the European project Safeguard.
+The SafeNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
+Data acquisition followed a non-systematic approach.
+First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
+These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions or countries within the database.
+To keep high quality standards that will allow robust future ecological research, we only included: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
+The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
+
 ## Dataset description
 
-This European metaweb consist of datasets of plant-pollinator interactions compiled initially by a wide number of researchers and institutions within the European continent.
-This dataset consist of `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
-The dataset contains a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+SafeNet includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
 The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
@@
 
-Define study area//Data acquisition//Data structure?
-
-TABLE? Describe cols?
+Define study area//Data structure? TABLE? Describe cols?
 

### 2024-02-26 Manuscript/Ms.qmd
preparing fig 2

@@
 First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions or countries within the database.
+These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions of countries within the database.
 To keep high quality standards that will allow robust future ecological research, we only included: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
@@
 The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if the species name is uncertain (e.g., species complex or species alike).
+For transparency, we have included in the database the old species name, the new assigned name, and, if name of the species is uncertain (e.g., species complex or species alike).
 
@@
 
-NOT FULLY HAPPY with it
+\newpage
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
+# Results
+
+PART A
+
+Provide number of species per taxonomic group and number of interactions
+Something always of interest: provide % of interactions of _Apis mellifera_?
+
+FIGURE 3 POLLINATOR COVERAGE
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
+a) Phylogenetic tree of plant families in the Safenet database, showing the number of species per family in logarithmic scale (blue circles) and the total number of interactions recorded per family (gray bars).
+b) Comparison of the top 50 families with the highest number of species in the SafeNet database against the total number of species recorded for those families at the European level.
+Species counts are displayed in logarithmic scale for better visualization.", fig.width=8, fig.height=8}
+
+library(ggplot2)
+library(ggtree)
+library(ggnewscale)
+library(ggtreeExtra)
+library(cowplot)
+
+#Save data to plot it in markdown file
+c = readRDS("../Data/Working_files/Figures/data_figure3b.rds")
+tree_family_interactions = readRDS("../Data/Working_files/Figures/data_figure3a1.rds")
+plant_fam_interactions = readRDS("../Data/Working_files/Figures/data_figure3a2.rds")
+
+# The circular layout tree
+p = ggtree(tree_family_interactions, 
+    layout="fan", size=0.1, open.angle=5, alpha=0.5)+ 
+geom_tippoint(aes(size= Family_species), colour='cyan4') +
+geom_tiplab(linetype='dashed', linesize=.05, 
+    size=1.75, color= "black", offset = -45, fontface="bold") +
+labs("log(Species)")+
+scale_size(name = "log(Species #)") +
+ggtitle("a)") +
+theme(plot.title = element_text(vjust = -31, hjust = 0.12, size = 14),
+      legend.position = "bottom")
+
+#Create barplot around circular tree
+p1 = p + 
+new_scale_fill() +
+geom_fruit(data=plant_fam_interactions, geom=geom_bar,
+   mapping=aes(y=Plant_family1, x=Family_interactions1),
+   pwidth=0.38, 
+   orientation="y", 
+   stat="identity", 
+   color = "black") +
+theme(legend.margin=margin(0,0,0,0),
+   legend.box.margin=margin(-50,-50,-50,-50),
+   plot.margin = margin(-15, -15, -10, -10, "pt")) +
+scale_size(name = "log(species number)", range=c(0,4))
+
+p2 = ggplot() + 
+geom_col(data = c, 
+  aes(x = reorder(Family_names, Spp_per_family), 
+  y = log(Spp_per_family), fill = Group), position=position_dodge(0.1),width=1) +
+xlab(NULL) +
+ylab("log(species number)") +
+theme_bw() +
+theme(axis.text.y = element_text(size = 5, face = "bold", color = "black"),
+      panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
+coord_flip(expand = F) +
+scale_fill_manual(name="", values = c("black", "cyan4")) +
+  theme(plot.margin = margin(-10, 0, -10, -10, "pt")) +
+ggtitle("b)")+
+theme(legend.margin=margin(0,0,0,0),
+   legend.box.margin=margin(-10,-10,-10,-10)) +
+  theme(legend.key.size = unit(0.3, 'cm'),
+        legend.position = "bottom",
+        plot.title = element_text(size=14))
+
+p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
+plot_grid(p1,p2_1, rel_widths = c(16, 7), rel_heights  = c(9, 1), align = "none")
+
+```
+
+PART B
+
+Explain habitat and bioregion
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
 Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
@@
 
-\newpage
-
-# Results
-
-Something always of interest: provide % of interactions of _Apis mellifera_?
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-a) Phylogenetic tree of plant families in the Safenet database, showing the number of species per family in logarithmic scale (blue circles) and the total number of interactions recorded per family (gray bars).
-b) Comparison of the top 50 families with the highest number of species in the SafeNet database against the total number of species recorded for those families at the European level.
-Species counts are displayed in logarithmic scale for better visualization.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data to plot it in markdown file
-c = readRDS("../Data/Working_files/Figures/data_figure3b.rds")
-tree_family_interactions = readRDS("../Data/Working_files/Figures/data_figure3a1.rds")
-plant_fam_interactions = readRDS("../Data/Working_files/Figures/data_figure3a2.rds")
-
-# The circular layout tree
-p = ggtree(tree_family_interactions, 
-    layout="fan", size=0.1, open.angle=5, alpha=0.5)+ 
-geom_tippoint(aes(size= Family_species), colour='cyan4') +
-geom_tiplab(linetype='dashed', linesize=.05, 
-    size=1.75, color= "black", offset = -45, fontface="bold") +
-labs("log(Species)")+
-scale_size(name = "log(Species #)") +
-ggtitle("a)") +
-theme(plot.title = element_text(vjust = -31, hjust = 0.12, size = 14),
-      legend.position = "bottom")
-
-#Create barplot around circular tree
-p1 = p + 
-new_scale_fill() +
-geom_fruit(data=plant_fam_interactions, geom=geom_bar,
-   mapping=aes(y=Plant_family1, x=Family_interactions1),
-   pwidth=0.38, 
-   orientation="y", 
-   stat="identity", 
-   color = "black") +
-theme(legend.margin=margin(0,0,0,0),
-   legend.box.margin=margin(-50,-50,-50,-50),
-   plot.margin = margin(-15, -15, -10, -10, "pt")) +
-scale_size(name = "log(species number)", range=c(0,4))
-
-p2 = ggplot() + 
-geom_col(data = c, 
-  aes(x = reorder(Family_names, Spp_per_family), 
-  y = log(Spp_per_family), fill = Group), position=position_dodge(0.1),width=1) +
-xlab(NULL) +
-ylab("log(species number)") +
-theme_bw() +
-theme(axis.text.y = element_text(size = 5, face = "bold", color = "black"),
-      panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-coord_flip(expand = F) +
-scale_fill_manual(name="", values = c("black", "cyan4")) +
-  theme(plot.margin = margin(-10, 0, -10, -10, "pt")) +
-ggtitle("b)")+
-theme(legend.margin=margin(0,0,0,0),
-   legend.box.margin=margin(-10,-10,-10,-10)) +
-  theme(legend.key.size = unit(0.3, 'cm'),
-        legend.position = "bottom",
-        plot.title = element_text(size=14))
-
-p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
-plot_grid(p1,p2_1, rel_widths = c(16, 7), rel_heights  = c(9, 1), align = "none")
-
-```
-
-FIGURE 4 POLLINATOR COVERAGE
-
 FIGURE 5? CLIFFHANGER OF NETWORK METRICS
@@
 :::
+
+# Acknowledgments
+
+We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data. This research was funded by the H2020 European project Safeguard (101003476).
+
+# Author contributions
 

### 2024-02-29 Manuscript/Figure_2.Rmd
progress

---
output: pdf_document
date: "2024-02-29"
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=10, fig.height=10}

#Load libraries
library(ggplot2)
library(ggnewscale)
library(ggtreeExtra)
library(ggtree)
library(ggpattern)
library(patchwork)
library(cowplot)

#Read data
tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
family_level1 = readRDS("../Data/Manuscript_info/bee_family_info.RData")
bee_plotting = readRDS("../Data/Manuscript_info/bee_plotting.RData")

#Plot
p = ggtree(tree_family_interactions,  size=0.5,  alpha=1)+ 
geom_tippoint(aes(size= Interactions, fill = Pollinator_family), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = 0.19, size=2.4, fontface=4) +
labs("log(Species)")+
scale_size(name = "Species", range = c(0, 5.5)) +
guides(fill = "none") +
theme(legend.position = "bottom")

p1 = p + 
new_scale_fill() +
geom_fruit(data=family_level1, geom=geom_bar,
      mapping=aes(y=Pollinator_genus, x=Interactions1, 
      fill = Pollinator_family1),
      pwidth=0.19, 
      orientation="y", 
      stat="identity",
      color="black",
      offset = 0.23,
      axis.params=list(axis       = "x",
                         text.size= 1.8,
                         hjust    = 1,
                         vjust      = 0.5,
                         nbreak     = 2),grid.params=list())  +
theme(legend.position = "bottom")

p2 = ggplot(data = bee_plotting, aes(reorder(Pollinator_family, -Spp_number), 
       Spp_number, fill = Group, pattern = Group)) +
geom_col_pattern(position = "dodge",color = "black", pattern_fill = "black") +
scale_pattern_manual(values = c(A = "stripe", B = "none"), labels= c(A = "SafeNet spp", B = "European spp")) +
scale_fill_manual(values = c(A = "tan2", B = "black"), labels= c(A = "SafeNet spp", B = "European spp")) + 
coord_cartesian(expand = FALSE) +
theme_bw() + 
ylab("Number of species") +
xlab("")+
ggtitle("Bee family coverage") +
theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, face = "bold"),
 legend.box.margin=margin(40,40,40,40),
 legend.position = "bottom")

p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
plot_grid(p1,p2_1, rel_widths = c(2, 2), rel_heights  = c(4, 0.4), align = "none")

```

### 2024-02-29 Scripts/Figure2_bees.R

@@ -1,0 +1,208 @@
+#Code to explore graphically pollinator coverage of the different species
+
+#Load libraris
+library(data.tree)
+library(ape)
+library(phytools)
+library(readr)
+library(stringr)
+library(dplyr)
+library(ggplot2)
+library(ggtree)
+library(ggpattern)
+
+#------------------------------------------------------------------#
+#Load data----
+#------------------------------------------------------------------#
+#Safenet interactions
+data = readRDS("Data/3_Final_data/Interactions_uncounted.rds")
+#Bee-syrphid Masterlist
+master_list = read_csv("Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
+#Bee phylo
+bee.trees=read.tree(file="Data/Working_files/phylogeny_genus_level.txt")
+
+#------------------------------------------------------------------#
+#Explore the number of interactions and species per bee family-----
+#------------------------------------------------------------------#
+#Bee families
+bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
+            "Andrenidae", "Colletidae","Melittidae")
+#Bee family interactions
+bee_family_interactions = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Pollinator_order == "Hymenoptera") %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+filter(Pollinator_family %in% bee_fam) %>% 
+group_by(Pollinator_family) %>% 
+summarise(Interactions = length(Pollinator_family))
+#Bee family species
+bee_family_species = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Pollinator_order == "Hymenoptera") %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+filter(Pollinator_family %in% bee_fam) %>% 
+group_by(Pollinator_family) %>% 
+summarise(Spp_number = n_distinct(Pollinator_accepted_name))
[…]

### 2024-03-01 Manuscript/Figure_2.Rmd
Figure phylo and coverage of bees done!

@@
 tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-family_level1 = readRDS("../Data/Manuscript_info/bee_family_info.RData")
-bee_plotting = readRDS("../Data/Manuscript_info/bee_plotting.RData")
+d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
+n = readRDS("../Data/Manuscript_info/node_data.RData")
 
-#Plot
-p = ggtree(tree_family_interactions,  size=0.5,  alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Pollinator_family), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.19, size=2.4, fontface=4) +
-labs("log(Species)")+
-scale_size(name = "Species", range = c(0, 5.5)) +
-guides(fill = "none") +
-theme(legend.position = "bottom")
-
-p1 = p + 
+#Circular layour
+ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
+geom_tippoint(aes(size= Interactions), fill = "tan2", color="black", shape=21) +
+geom_tiplab(linetype='dashed', linesize=.05, offset = 0.062, size=2.5, fontface=1) +
+scale_size(name = "Log(Interactions)", range = c(0,4.5))  +
+theme(legend.position = "bottom") +
 new_scale_fill() +
-geom_fruit(data=family_level1, geom=geom_bar,
-      mapping=aes(y=Pollinator_genus, x=Interactions1, 
-      fill = Pollinator_family1),
-      pwidth=0.19, 
+geom_fruit(data = d1, geom=geom_bar,
+                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
+                    pwidth=0.1, 
                     orientation="y", 
@@
                     color="black",
-      offset = 0.23,
-      axis.params=list(axis       = "x",
-                         text.size= 1.8,
+                    offset = 0.128,
+            axis.params=list(
+                         axis = "x",
+                         text.size  = 2.5,
                          hjust  = 1,
                          vjust = 0.5,
-                         nbreak     = 2),grid.params=list())  +
-theme(legend.position = "bottom")
-
-p2 = ggplot(data = bee_plotting, aes(reorder(Pollinator_family, -Spp_number), 
-       Spp_number, fill = Group, pattern = Group)) +
-geom_col_pattern(position = "dodge",color = "black", pattern_fill = "black") +
-scale_pattern_manual(values = c(A = "stripe", B = "none"), labels= c(A = "SafeNet spp", B = "European spp")) +
-scale_fill_manual(values = c(A = "tan2", B = "black"), labels= c(A = "SafeNet spp", B = "European spp")) + 
-coord_cartesian(expand = FALSE) +
-theme_bw() + 
-ylab("Number of species") +
-xlab("")+
-ggtitle("Bee family coverage") +
-theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, face = "bold"),
- legend.box.margin=margin(40,40,40,40),
- legend.position = "bottom")
-
-p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
-plot_grid(p1,p2_1, rel_widths = c(2, 2), rel_heights  = c(4, 0.4), align = "none")
+                         nbreak = 3,
+                         fontface = 2
+                     ),
+         grid.params=list()) +
+guides(fill = "none")+
+scale_fill_manual(values = c(a = "tan2", b = "lightblue3"), guide = "none")+
+geom_fruit(geom=geom_bar,
+                    mapping=aes(y=label, x=Spp_number_Europe),
+                    pwidth=0.1, 
+                    orientation="y", 
+                    stat="identity",
+                    color="black",
+                    offset = 0,
+            axis.params=list(
+                         axis = "x",
+                         text.size  = 2.5,
+                         hjust  = 1,
+                         vjust = 0.5,
+                         nbreak = 1,
+                         fontface = 2
+                     ),
+         grid.params=list()) +
+annotate("text", x = 14, y = 70, label = "SafeNet spp\n coverage (%)", size= 2.65, fontface= "bold") +
+annotate("text", x = 15.2, y = 70, label = "European\n species", size= 2.65, fontface= "bold")+
+annotate("text", x = 12.1, y = 70, label = "SafeNet\n interactions", size= 2.65, fontface= "bold")+ 
+new_scale_fill() +
+geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
+annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
+annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
+annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
+annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
+annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
+annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
+coord_cartesian(clip = "off")+
+theme(legend.margin=unit(1,"cm"),
+      legend.spacing = unit(1, "cm"),
+      plot.margin = margin(4,0,4,0), 
+      legend.position = c(0.6,0),
+      legend.direction = "horizontal")
 

### 2024-03-01 Manuscript/Ms.qmd
Figure phylo and coverage of bees done!

@@
 
-FIGURE 3 POLLINATOR COVERAGE
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
+Phylogenetic representation of the bee genera in the SafeNet database within the context of all European bee genera.
+The number of interactions recorded in the database per genus is depicted with circles of proportional size corresponding to the number of interactions.
+The coverage of species per genus is depicted with orange and light blue bars, representing the percentage of species included and not included in the database, respectively, out of the total number of species per genus at the European level.
+Grey bars indicate the total number of species per genus at the European level."}
+
+#Load libraries
+library(ggplot2)
+library(ggnewscale)
+library(ggtreeExtra)
+library(ggtree)
+library(ggpattern)
+library(patchwork)
+library(cowplot)
+
+#Read data
+tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
+d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
+n = readRDS("../Data/Manuscript_info/node_data.RData")
+
+#Circular layour
+ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
+geom_tippoint(aes(size= Interactions), fill = "tan2", color="black", shape=21) +
+geom_tiplab(linetype='dashed', linesize=.05, offset = 0.063, size=2.9, fontface=1) +
+scale_size(name = "Log(Interactions)", range = c(0,4.5))  +
+theme(legend.position = "bottom") +
+new_scale_fill() +
+geom_fruit(data = d1, geom=geom_bar,
+                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
+                    pwidth=0.12, 
+                    orientation="y", 
+                    stat="identity",
+                    color="black",
+                    offset = 0.137,
+            axis.params=list(
+                         axis = "x",
+                         text.size  = 2.5,
+                         hjust  = 1,
+                         vjust = 0.5,
+                         nbreak = 3,
+                         fontface = 2
+                     ),
+         grid.params=list()) +
+guides(fill = "none")+
+scale_fill_manual(values = c(a = "tan2", b = "lightblue3"), guide = "none")+
+geom_fruit(geom=geom_bar,
+                    mapping=aes(y=label, x=Spp_number_Europe),
+                    pwidth=0.12, 
+                    orientation="y", 
+                    stat="identity",
+                    color="black",
+                    offset = 0,
+            axis.params=list(
+                         axis = "x",
+                         text.size  = 2.5,
+                         hjust  = 1,
+                         vjust = 0.5,
+                         nbreak = 1,
+                         fontface = 2
+                     ),
+         grid.params=list()) +
+annotate("text", x = 14.2, y = 70, label = "SafeNet spp\n coverage (%)", size= 3, fontface= "bold") +
+annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
+annotate("text", x = 12.1, y = 70, label = "SafeNet\n interactions", size= 3, fontface= "bold")+ 
+new_scale_fill() +
+geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
+annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
+annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
+annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
+annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
+annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
+annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
+coord_cartesian(clip = "off")+
+theme(legend.margin=unit(1,"cm"),
+      legend.spacing = unit(1, "cm"),
+      plot.margin = margin(4,0,4,0), 
+      legend.position = c(0.6,0),
+      legend.direction = "horizontal")
+
+```
 

### 2024-03-04 Manuscript/Figure_3.Rmd
phylo plants updated

---
output: pdf_document
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=8, fig.height=8}

library(ggplot2)
library(ggtree)
library(ggnewscale)
library(ggtreeExtra)
library(viridis)
#Save data and plot it in an Rmarkdown
tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")

#Check if there are differences in labels and order
ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
scale_size(range = c(0,4), guide = 'none') +
scale_fill_continuous(type = "viridis", direction=-1) + 
theme(legend.position = "bottom") +
guides(fill=guide_legend(), size = guide_legend()) +
new_scale_fill() +
geom_fruit(data = d1, geom=geom_bar,
   mapping=aes(y=label, x=Percent, fill = fam_group),
   pwidth=0.1, 
   orientation="y", 
   stat="identity",
   color="black",
   offset = 0.028,
   linewidth=0.35,
   axis.params=list(
     axis = "x",
     text.size  = 1.5,
     hjust  = 1,
     vjust = 0.5,
     nbreak = 3,
     fontface = 2),
  grid.params=list())+
guides(fill = "none")+
scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
geom_fruit(geom=geom_bar,
    mapping=aes(y=label, x=log(Spp_Europe +1)),
    pwidth=0.1, 
    orientation="y", 
    stat="identity",
    color="black",
    offset = 0,
    linewidth=0.35) +
theme(legend.margin=unit(1,"cm"),
      legend.spacing = unit(1, "cm"),
      plot.margin = margin(4,0,4,0), 
      legend.position = c(0.5,0.09),
      legend.direction = "horizontal")

```

### 2024-03-04 Manuscript/Ms.qmd
phylo plants updated

--- 2024-02-26 Manuscript/Ms.qmd
@@
 
-FIGURE 3 POLLINATOR COVERAGE
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
+Phylogenetic representation of bee genera in the SafeNet database within the context of all European bee genera.
+The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
+Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
+The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
+Grey bars indicate the total number of species per genus at the European level."}
+
+#Load libraries
+library(ggplot2)
+library(ggnewscale)
+library(ggtreeExtra)
+library(ggtree)
+library(ggpattern)
+library(patchwork)
+library(cowplot)
+
+#Read data
+tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
+d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
+n = readRDS("../Data/Manuscript_info/node_data.RData")
+
+#Circular layour
+ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
+geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
+geom_tiplab(linetype='dashed', linesize=.05, offset = 0.063, size=2.9, fontface=1) +
+scale_size(range = c(0,4.5), guide = 'none')  +
+scale_fill_continuous(type = "viridis", direction=-1) + 
+guides(fill=guide_legend(), size = guide_legend()) +
+theme(legend.position = "bottom") +
+new_scale_fill() +
+geom_fruit(data = d1, geom=geom_bar,
+                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
+                    pwidth=0.13, 
+                    orientation="y", 
+                    stat="identity",
+                    color="black",
+                    offset = 0.137,
+            axis.params=list(
+                         axis = "x",
+                         text.size  = 2.5,
+                         hjust  = 1,
+                         vjust = 0.5,
+                         nbreak = 3,
+                         fontface = 2
+                     ),
+         grid.params=list()) +
+guides(fill = "none")+
+scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
+geom_fruit(geom=geom_bar,
+                    mapping=aes(y=label, x=Spp_number_Europe),
+                    pwidth=0.12, 
+                    orientation="y", 
+                    stat="identity",
+                    color="black",
+                    offset = 0,
+            axis.params=list(
+                         axis = "x",
+                         text.size  = 2.5,
+                         hjust  = 1,
+                         vjust = 0.5,
+                         nbreak = 1,
+                         fontface = 2
+                     ),
+         grid.params=list()) +
+annotate("text", x = 14.2, y = 70, label = "SafeNet spp\n coverage (%)", size= 3, fontface= "bold") +
+annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
+annotate("text", x = 12.1, y = 70, label = "SafeNet\n interactions", size= 3, fontface= "bold")+ 
+new_scale_fill() +
+geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
+annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
+annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
+annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
+annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
+annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
+annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
+coord_cartesian(clip = "off")+
+theme(legend.margin=unit(1,"cm"),
+      legend.spacing = unit(1, "cm"),
+      plot.margin = margin(4,0,4,0), 
+      legend.position = c(0.6,0),
+      legend.direction = "horizontal")
+
+```
 
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-a) Phylogenetic tree of plant families in the Safenet database, showing the number of species per family in logarithmic scale (blue circles) and the total number of interactions recorded per family (gray bars).
-b) Comparison of the top 50 families with the highest number of species in the SafeNet database against the total number of species recorded for those families at the European level.
-Species counts are displayed in logarithmic scale for better visualization.", fig.width=8, fig.height=8}
+Phylogenetic representation of the plant families in the SafeNet database within the context of all flowering plant families occurring in Europe.
+The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
+Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
+The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
+Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
 
@@
 
-#Save data to plot it in markdown file
-c = readRDS("../Data/Working_files/Figures/data_figure3b.rds")
-tree_family_interactions = readRDS("../Data/Working_files/Figures/data_figure3a1.rds")
-plant_fam_interactions = readRDS("../Data/Working_files/Figures/data_figure3a2.rds")
+#Save data and plot it in an Rmarkdown
+tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
+d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
 
-# The circular layout tree
-p = ggtree(tree_family_interactions, 
-    layout="fan", size=0.1, open.angle=5, alpha=0.5)+ 
-geom_tippoint(aes(size= Family_species), colour='cyan4') +
-geom_tiplab(linetype='dashed', linesize=.05, 
-    size=1.75, color= "black", offset = -45, fontface="bold") +
-labs("log(Species)")+
-scale_size(name = "log(Species #)") +
-ggtitle("a)") +
-theme(plot.title = element_text(vjust = -31, hjust = 0.12, size = 14),
-      legend.position = "bottom")
-
-#Create barplot around circular tree
-p1 = p + 
+#Check if there are differences in labels and order
+ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
+geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
+geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
+scale_size(range = c(0,4), guide = 'none') +
+scale_fill_continuous(type = "viridis", direction=-1) + 
+theme(legend.position = "bottom") +
+guides(fill=guide_legend(), size = guide_legend()) +
 new_scale_fill() +
-geom_fruit(data=plant_fam_interactions, geom=geom_bar,
-   mapping=aes(y=Plant_family1, x=Family_interactions1),
-   pwidth=0.38, 
+geom_fruit(data = d1, geom=geom_bar,
+   mapping=aes(y=label, x=Percent, fill = fam_group),
+   pwidth=0.1, 
    orientation="y", 
    stat="identity",
-   color = "black") +
-theme(legend.margin=margin(0,0,0,0),
-   legend.box.margin=margin(-50,-50,-50,-50),
-   plot.margin = margin(-15, -15, -10, -10, "pt")) +
-scale_size(name = "log(species number)", range=c(0,4))
-
-p2 = ggplot() + 
-geom_col(data = c, 
-  aes(x = reorder(Family_names, Spp_per_family), 
-  y = log(Spp_per_family), fill = Group), position=position_dodge(0.1),width=1) +
-xlab(NULL) +
-ylab("log(species number)") +
-theme_bw() +
-theme(axis.text.y = element_text(size = 5, face = "bold", color = "black"),
-      panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
-coord_flip(expand = F) +
-scale_fill_manual(name="", values = c("black", "cyan4")) +
-  theme(plot.margin = margin(-10, 0, -10, -10, "pt")) +
-ggtitle("b)")+
-theme(legend.margin=margin(0,0,0,0),
-   legend.box.margin=margin(-10,-10,-10,-10)) +
-  theme(legend.key.size = unit(0.3, 'cm'),
-        legend.position = "bottom",
-        plot.title = element_text(size=14))
-
-p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
-plot_grid(p1,p2_1, rel_widths = c(16, 7), rel_heights  = c(9, 1), align = "none")
+   color="black",
+   offset = 0.028,
+   linewidth=0.35,
+   axis.params=list(
+     axis = "x",
+     text.size  = 1.5,
+     hjust  = 1,
+     vjust = 0.5,
+     nbreak = 3,
+     fontface = 2),
+  grid.params=list())+
+guides(fill = "none")+
+scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
+geom_fruit(geom=geom_bar,
+    mapping=aes(y=label, x=log(Spp_Europe +1)),
+    pwidth=0.1, 
+    orientation="y", 
+    stat="identity",
+    color="black",
+    offset = 0,
+    linewidth=0.35) +
+theme(legend.margin=unit(1,"cm"),
+      legend.spacing = unit(1, "cm"),
+      plot.margin = margin(4,0,4,0), 
+      legend.position = c(0.5,0.09),
+      legend.direction = "horizontal")
 

### 2024-03-05 Manuscript/Figure_4.Rmd
Edits on Figure 4 and Figure 2. Figure S1 added to supplementary. Think about Figure 5...

--- 2024-02-01 Manuscript/Figure_2a.Rmd
@@
 ---
-title: "Untitled"
 output: pdf_document
-date: "2024-02-01"
 ---
@@
 library(forcats)
+library(viridis)
 
@@
       axis.title = element_text(face = "bold", size = 8),
-      axis.text = element_text(size=6)) +
+      axis.text = element_text(size=6),
+      panel.border = element_rect(size=1)) +
 xlab(NULL) +
@@
 color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = c("#ffd67a","skyblue", "lightpink2", "tomato3"),
+scale_fill_manual(values = viridis(4),
                   name = NULL) +
@@
 
+#c("azure3","steelblue", "wheat2", "tomato3")
+
 #Unify labels
@@
 p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "black",width = 0.9) +
+geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
 theme_light() +
@@
       axis.ticks.y = element_blank(), 
+      axis.ticks.x = element_line(color="black"),
   axis.text.y = element_blank(), 
@@
 panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
+axis.line = element_line(color = 'black'),
+plot.margin = margin(0,0,0,0)) +
 scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
@@
 p3 = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray10", width = 0.9) +
+geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
 theme_light() +
@@
   strip.text.x = element_text(size=12),
+   axis.ticks.x = element_line(color="black"),
   strip.background = element_rect(

### 2024-03-05 Manuscript/Figure_S1.Rmd
Edits on Figure 4 and Figure 2. Figure S1 added to supplementary. Think about Figure 5...

---
output: pdf_document
date: "2024-02-29"
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=10, fig.height=4}
#Prepare figure for supplementary material
#Coverage across high taxonomic ranks of bees, syrphids and butterflies

#Load libraries
library(dplyr)
library(readr)
library(viridis)
library(ggplot2)
#--------------------------------------------------------------------------------------------#
#Start from bees----
#Load data
#Safenet interactions
data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
#Bee-syrphid Masterlist
master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
#Load butterfly masterlist
master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
                    header = T)
#Bee families
bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
            "Andrenidae", "Colletidae","Melittidae")

#Species per family within SafeNet
bee_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Hymenoptera") %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
filter(Pollinator_family %in% bee_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Species per family within Europe
bee_spp_europe =  master_list %>% 
select(Family, Genus, Species, GenusAndSpecies) %>% 
filter(Family %in% bee_fam) %>% 
group_by(Family) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
rename(Pollinator_family = Family) %>% 
mutate(Group = "Europe")

#Bind both datasets for plotting
bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)

p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
ylab("Number of species") +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Bee family coverage")
#--------------------------------------------------------------------------------------------#
#Continue with syrphids----
#As we don't have subfamilies, let's obtained them from the master list
syrphids = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
rename(Pollinator_accepted_name = GenusAndSpecies)
#Syrphid species safenet
syrphid_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_family== "Syrphidae") %>% 
distinct()
#Add subfamily info
syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
#Summarise spp per subfamily
syrphid_spp_safenet = syrphid_spp_safenet %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Get spp per subfamily at european level
syrphid_spp_europe = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
#Plot
p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Syrphid subfamily coverage")
#--------------------------------------------------------------------------------------------#
#Continue with butterflies----
#Do it at family level
#Create vector with butterfly families
butt_fam = c(unique(master_list1$family))
#Species per family within SafeNet
butterfly_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Lepidoptera") %>% 
filter(Pollinator_family %in% butt_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Species per family within Europe
butterfly_spp_europe =  master_list1 %>% 
select(family, genus, scientificName) %>% 
group_by(family) %>% 
summarise(Spp_number = n_distinct(scientificName)) %>% 
rename(Pollinator_family = family) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
#Plot
p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Butterfly family coverage")

#--------------------------------------------------------------------------------------------#
library(patchwork)
p1+p2+p3 + plot_layout(guides = "collect") & 
theme(plot.title = element_text(size = 12, face = "bold"),
      axis.title.y = element_text(face = "bold"))

```

### 2024-03-05 Manuscript/Ms.qmd
Edits on Figure 4 and Figure 2. Figure S1 added to supplementary. Think about Figure 5...

@@
 library(dplyr) #For handling fata 
-library(ggplot2) #For plotting
-library(giscoR) #For plotting
 library(sf) #For handling coordinates and plotting
-library(patchwork)#For plotting
-library(ggstar) #For plotting (cool shapes)
 library(scales) #For plotting (decimals on axes)
@@
 library(forcats)
+library(readr)
+#Load libraries for plotting
+library(giscoR) #Maps
+library(ggstar) #cool shapes as points
+library(ggplot2) #general plotting
+library(ggtree) #phylo trees
+library(ggpattern) #cool shapes in barplots
+library(patchwork) #plot organisation
+library(cowplot) #plot organisation
+library(viridis) #colours
+library(ggnewscale)
+library(ggtreeExtra)
 
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Print number of countries
+#Load data
 data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
+#Bee-syrphid Masterlist
+master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
+#Load butterfly masterlist
+master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
+                    header = T)
 #Country number
@@
 
-PART A
+## Taxonomy and taxonomic coverage
 
@@
 Grey bars indicate the total number of species per genus at the European level."}
-
-#Load libraries
-library(ggplot2)
-library(ggnewscale)
-library(ggtreeExtra)
-library(ggtree)
-library(ggpattern)
-library(patchwork)
-library(cowplot)
 
@@
 
-PART B
+\newpage
 
-Explain habitat and bioregion
+## Habitat type and bioclimatic region
 
@@
 Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=6, fig.height=8}
+Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=6, fig.height=8}
 
-#Read interaction data 
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
 #Load data (done it in earlier chunks)
@@
 ggtitle("(a)") +
-theme(plot.title = element_text(size = 14 ,hjust = -0.25, vjust = -2.5))
+theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
+
+#c("azure3","steelblue", "wheat2", "tomato3")
 
@@
 p3 = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray10", width = 0.9) +
+geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
 theme_light() +
 theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 8, vjust = -1),
+  axis.title.y = element_text(face = "bold", size = 8, vjust = -2),
   axis.title.x = element_text(face = "bold", size = 8),
@@
 ggtitle("(b)") +
-theme(plot.title = element_text(size = 14,hjust = -0.3, vjust = 2))
+theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
 
@@
 
-FIGURE 5? CLIFFHANGER OF NETWORK METRICS
+**FIGURE 5?** CLIFFHANGER OF NETWORK METRICS
 
-## Taxonomy and taxonomic coverage
-
-## Habitat type and bioclimatic region
+\newpage
 
@@
 
-Supplementary text 1
+**Supplementary text 1**
 
@@
 12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
+
+\newpage
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure S1}.
+Coverage of the SafeNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
+
+#Prepare figure for supplementary material
+#Coverage across high taxonomic ranks of bees, syrphids and butterflies
+#--------------------------------------------------------------------------------------------#
+#Start from bees----
+#Bee families
+bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
+            "Andrenidae", "Colletidae","Melittidae")
+
+#Species per family within SafeNet
+bee_spp_safenet = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Pollinator_order == "Hymenoptera") %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+filter(Pollinator_family %in% bee_fam) %>% 
+group_by(Pollinator_family) %>% 
+summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
+mutate(Group = "SafeNet")
+#Species per family within Europe
+bee_spp_europe =  master_list %>% 
+select(Family, Genus, Species, GenusAndSpecies) %>% 
+filter(Family %in% bee_fam) %>% 
+group_by(Family) %>% 
+summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
+rename(Pollinator_family = Family) %>% 
+mutate(Group = "Europe")
+
+#Bind both datasets for plotting
+bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
+
+p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
+geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
+coord_cartesian(expand = FALSE) +
+xlab(NULL) +
+theme_bw() +
+theme(axis.text.x = element_text(angle = 45, hjust =1),
+      panel.border = element_rect(size=1)) +
+ylab("Number of species") +
+scale_fill_manual(values = c(viridis(2)))+
+ggtitle("Bee family coverage")
+#--------------------------------------------------------------------------------------------#
+#Continue with syrphids----
+#As we don't have subfamilies, let's obtained them from the master list
+syrphids = master_list %>% 
+filter(Family == "Syrphidae") %>% 
+select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
+rename(Pollinator_accepted_name = GenusAndSpecies)
+#Syrphid species safenet
+syrphid_spp_safenet = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Pollinator_family== "Syrphidae") %>% 
+distinct()
+#Add subfamily info
+syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
+#Summarise spp per subfamily
+syrphid_spp_safenet = syrphid_spp_safenet %>% 
+group_by(Subfamily) %>% 
+summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
+mutate(Group = "SafeNet")
+#Get spp per subfamily at european level
+syrphid_spp_europe = master_list %>% 
+filter(Family == "Syrphidae") %>% 
+select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
+group_by(Subfamily) %>% 
+summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
+mutate(Group = "Europe")
+#Bind both datasets for plotting
+syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
+#Plot
+p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
+geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
+coord_cartesian(expand = FALSE) +
+xlab(NULL) +
+ylab(NULL) +
+theme_bw() +
+theme(axis.text.x = element_text(angle = 45, hjust =1),
+      panel.border = element_rect(size=1)) +
+scale_fill_manual(values = c(viridis(2)))+
+ggtitle("Syrphid subfamily coverage")
+#--------------------------------------------------------------------------------------------#
+#Continue with butterflies----
+#Do it at family level
+#Create vector with butterfly families
+butt_fam = c(unique(master_list1$family))
+#Species per family within SafeNet
+butterfly_spp_safenet = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Pollinator_order == "Lepidoptera") %>% 
+filter(Pollinator_family %in% butt_fam) %>% 
+group_by(Pollinator_family) %>% 
+summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
+mutate(Group = "SafeNet")
+#Species per family within Europe
+butterfly_spp_europe =  master_list1 %>% 
+select(family, genus, scientificName) %>% 
+group_by(family) %>% 
+summarise(Spp_number = n_distinct(scientificName)) %>% 
+rename(Pollinator_family = family) %>% 
+mutate(Group = "Europe")
+#Bind both datasets for plotting
+butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
+#Plot
+p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
+geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
+coord_cartesian(expand = FALSE) +
+xlab(NULL) +
+ylab(NULL) +
+theme_bw() +
+theme(axis.text.x = element_text(angle = 45, hjust =1),
+      panel.border = element_rect(size=1)) +
+scale_fill_manual(values = c(viridis(2)))+
+ggtitle("Butterfly family coverage")
+
+#--------------------------------------------------------------------------------------------#
+p1+p2+p3 + plot_layout(guides = "collect") & 
+theme(plot.title = element_text(size = 12, face = "bold"),
+      axis.title.y = element_text(face = "bold"))
+
+```

### 2024-03-06 Manuscript/Figure1.qmd
Table with data sturcture added to MS

@@
 #Load data
-data = readRDS("../Data/Interactions_uncounted.rds")
+data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
 
@@
   plot.title = element_text(size = 8, face = "bold"), 
-  axis.text.y = element_text(size=6),
+  axis.text.y = element_text(size=5),
   panel.grid.major.x = element_blank(),
@@
 scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-annotate(geom="text", x=-500, y=1.15, label="(b)",color="black", size = 3.5) +
+annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
 coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
@@
            lwd = 0.015, fill = "#009E73") +
-ylab("Shared plants (%)") +
 theme_minimal()+
@@
          axis.ticks.y = element_line(size = 0.2),
-         axis.title.x = element_text(face = "bold", size = 8),            plot.title = element_text(size = 8, face = "bold"),
-         axis.text.y = element_text(size=6),
+         axis.title.x = element_text(face = "bold", size = 7),            plot.title = element_text(size = 8, face = "bold"),
+         axis.text.y = element_text(size=5),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
+         panel.grid.major.y = element_line(size=0.2, linetype =  2),
+      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
 xlab("Species") +
-ylab(NULL) + 
+ylab("Occurrence across studies") + 
 ggtitle("Plants") + 
@@
 
-#Prepare panel with plots
-plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Occurrence across studies (%)") + 
-theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
+data2 = data %>%  select(Study_id, Year) %>% 
+distinct() %>% 
+group_by(Year) %>% 
+summarise(Counts = n())
+
+p3 = ggplot(data2, aes(Year, Counts)) +
+geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
+theme_light() +
+theme(plot.margin=unit(c(0,0,0,0), "mm"),
+  axis.title.y = element_text(face = "bold", size = 7, vjust = -1.1),
+  axis.title.x = element_text(face = "bold", size = 8),
+  axis.text.y = element_text(size = 6),
+  axis.text.x = element_text(size = 6,  angle = 90), 
+  strip.text.x = element_text(size=12),
+   axis.ticks.x = element_line(color="black"),
+  strip.background = element_rect(
+  color="black", fill=NA, size=0.5, linetype="solid")) + 
+ylab("Number of studies") +
+coord_cartesian(expand = F, clip = "off") + 
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(),
+  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+panel.border = element_blank(),
+axis.line = element_line(color = 'black')) +
+scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
+scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
+ggtitle("(b)") +
+theme(plot.title = element_text(size = 10,hjust = -0.35, vjust = 0.8))
+
+plot2 = p3 / p1 / p2
+
 #Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 1))
+map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 0.8))
 

### 2024-03-06 Manuscript/Figure_4.Rmd
Table with data sturcture added to MS

@@
 Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=6, fig.height=8}
+Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
 
@@
 ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
-ggtitle("(a)") +
-theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
+guides(fill=guide_legend(nrow=2,byrow=TRUE))
 
@@
 
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p3 = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 8, vjust = -2),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-   axis.ticks.x = element_line(color="black"),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
-
-a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
 #Plot panel
-p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)
+p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
 

### 2024-03-06 Manuscript/Ms.qmd
Table with data sturcture added to MS

@@
 library(ggtreeExtra)
+library(kableExtra)
 
@@
                     header = T)
+#Prepare table with data structure (as it would be made available to others)
+safenet_database = readRDS("../Data/3_Final_data/safenet_database.rds")
+
+#Work around to get the three cols
+#Note that month 1 and day 1 indicates the presence of a missing value
+#At least in Europe :)
+data = data %>%
+dplyr::mutate(Year = lubridate::year(Date), 
+                Month = lubridate::month(Date), 
+                Day = lubridate::day(Date))
+
 #Country number
@@
 To keep high quality standards that will allow robust future ecological research, we only included: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
+The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**).
 
@@
 These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-
-Define study area//Data structure? TABLE? Describe cols?
 
@@
 For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Percentage of shared occurrences of pollinators and plants across studies.
+(\\textbf{b}) Number of studies by year within the database.
+(\\textbf{c}) Percentage of shared occurrences of pollinators and plants across studies.
 Species on the left are highly shared across studies, while species on the right are found only in a few or a single study.
 As a reference, we have indicated the 50th percentile with a black dashed vertical line."}
-
-#Load data (done it in previous chunk)
 
@@
 geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.01) + 
+        alpha = 1,linewidth=0.1) + 
 geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
@@
 ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -12, hjust=0.1, size=10))
+theme(plot.title = element_text(vjust = -12, hjust=0.2, size=10))
 
@@
 annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
-annotate(geom="text", x=20, y=0.89, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
+annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
 scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-annotate(geom="text", x=-500, y=1.15, label="(b)",color="black", size = 3.5) +
+annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
 coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
@@
 annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-annotate(geom="text", x=20, y=0.73, label=most_common_plant[1]
+annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
       ,color="black",  fontface = 'italic', size = 1.5)+
@@
 
-#Prepare panel with plots
-plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Occurrence across studies (%)") + 
-theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
+data2 = data %>%  select(Study_id, Year) %>% 
+distinct() %>% 
+group_by(Year) %>% 
+summarise(Counts = n())
+
+p3 = ggplot(data2, aes(Year, Counts)) +
+geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
+theme_light() +
+theme(plot.margin=unit(c(0,0,0,0), "mm"),
+  axis.title.y = element_text(face = "bold", size = 7, vjust = -1.1),
+  axis.title.x = element_text(face = "bold", size = 8),
+  axis.text.y = element_text(size = 6),
+  axis.text.x = element_text(size = 6,  angle = 90), 
+  strip.text.x = element_text(size=12),
+   axis.ticks.x = element_line(color="black"),
+  strip.background = element_rect(
+  color="black", fill=NA, size=0.5, linetype="solid")) + 
+ylab("Number of studies") +
+coord_cartesian(expand = F, clip = "off") + 
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  panel.background = element_blank(),
+  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+panel.background = element_blank(),
+panel.border = element_blank(),
+axis.line = element_line(color = 'black')) +
+scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
+scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
+ggtitle("(b)") +
+theme(plot.title = element_text(size = 10,hjust = -0.35, vjust = 0.8))
+
+plot2 = p3 / p1 / p2
+
 #Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 1))
+map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.58, 0.4))
+
+```
+
+\newpage
+
+#Data structure
+
+EXPLAIN!!!
+
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
+
+#Create variables
+vars = colnames(safenet_database)
+#Create table
+d = tibble(Variable = vars)
+
+d = d %>% 
+mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
+filter(!Variable == "Longitude")
+
+#Add descriptor
+my_tibble = d %>%
+  mutate(Description = case_when(
+    Variable == "Study_id" ~ "Identifier of the study",
+    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
+    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
+    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
+    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
+    Variable == "Bioregion" ~ "European biogiographical regions",
+    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
+    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
+    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
+    Variable == "Date" ~ "Year, month and day when the observation took place",
+    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
+    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
+    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
+    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
+    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
+    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
+    Variable == "Plant_unsure_id" ~ "Binary category to indicate if the plant species name is unsure (Yes) or not (No)",
+    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
+    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
+    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_unsure_id" ~ "Binary category to indicate if the pollinator species name is unsure (Yes) or not (No)",
+    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+    TRUE ~ "Other"
+  ))
+
+kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.}") %>%
+  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
+kable_styling(position = "center")
 
@@
 Something always of interest: provide % of interactions of _Apis mellifera_?
+
+phylogenetic and geographical biases
 
@@
 Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=6, fig.height=8}
+Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
 
@@
 select(Study_id, Network_id, Authors_habitat,
-  Corine_land_cover, SafeNet_habitat, Bioregion,
+  SafeNet_habitat, Bioregion,
   Latitude, Longitude, Day, Month, Year, Plant_rank,
@@
 ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
-ggtitle("(a)") +
-theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))
-
-#c("azure3","steelblue", "wheat2", "tomato3")
+guides(fill=guide_legend(nrow=2,byrow=TRUE))
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure S1}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S1}.
 Coverage of the SafeNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}

### 2024-03-06 Manuscript/Table_data_structure.Rmd
Table with data sturcture added to MS

--- 2024-03-04 Manuscript/Ms.qmd
@@
 ---
-title: "SafeNet: A European database of plant-pollinator interactions"
-format:
-  pdf:
-    include-in-header: 
-      text: |
-        \usepackage{lscape}
-        \newcommand{\blandscape}{\begin{landscape}}
-        \newcommand{\elandscape}{\end{landscape}}
-        \usepackage{caption}
-        \captionsetup[figure]{labelformat=empty}
-
-bibliography: references.bib
-latex-tinytex: false
-link-citations: TRUE
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
+output: pdf_document
 ---
 
-```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr) #For handling fata 
-library(ggplot2) #For plotting
-library(giscoR) #For plotting
-library(sf) #For handling coordinates and plotting
-library(patchwork)#For plotting
-library(ggstar) #For plotting (cool shapes)
-library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
+```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
+
+library(dplyr)
+library(kableExtra)
+library(stringr)
+
+#Prepare table with data structure
+data = readRDS("../Data/3_Final_data/safenet_database.rds")
+
+#Create variables
+vars = colnames(data)
+#Create table
+d = tibble(Variable = vars)
+
+d = d %>% 
+mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
+filter(!Variable == "Longitude")
+
+#Add descriptor
+my_tibble = d %>%
+  mutate(Description = case_when(
+    Variable == "Study_id" ~ "Identifier of the study",
+    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
+    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
+    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
+    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
+    Variable == "Bioregion" ~ "European biogiographical regions",
+    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
+    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
+    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
+    Variable == "Date" ~ "Year, month and day when the observation took place",
+    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
+    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
+    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
+    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
+    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
+    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
+    Variable == "Plant_unsure_id" ~ "Binary category to indicate if the plant species name is unsure (Yes) or not (No)",
+    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
+    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
+    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_unsure_id" ~ "Binary category to indicate if the pollinator species name is unsure (Yes) or not (No)",
+    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+    TRUE ~ "Other"
+  ))
+
+#Next step add it to an R markdown
+
+kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.}") %>%
+  kable_styling(full_width = F,font_size = 10) %>% row_spec(0, bold=T) %>% 
+kable_styling(position = "center")
 
 ```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Print number of countries
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Read author list 
-authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
-#Print authors
-authors = authors %>% 
-summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
-pull()
-#Read affiliation list 
-aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
-aff_list = aff_list %>% 
-summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
-pull()
-
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first three and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
-Here, we present the SafeNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
-This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
-We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.
-
-**Main Types of Variables Included:** SafeNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region and further taxonomic rank information for both plant and pollinator species are also provided.
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants.
-The main focus of this database are insect pollinators and accounts for `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
-
-**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats.
-The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
-
-**KEYWORDS**\
-\begingroup
-\small
-plant-pollinator interactions, flowering plants, Angiosperms, pollinators, Hymenoptera, Diptera, Lepidoptera, Coleoptera, Europe
-\endgroup
-\pagebreak
-
-# Introduction
-
-The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023].
-Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes at large scales.
-
-Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the SafeNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-SafeNet aims to cover a wide range of taxonomic groups and habitats while also provides other variables of interest that allow to better control for the ecological context and sampling methods.
-In addition, SafeNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
-We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
-
-# Methods
-
-## Data acquisition
-
-This database is the result of one of the working packages of the European project Safeguard.
-The SafeNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-Data acquisition followed a non-systematic approach.
-First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions of countries within the database.
-To keep high quality standards that will allow robust future ecological research, we only included: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a**).
-
-## Dataset description
-
-SafeNet includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
-These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-
-Define study area//Data structure? TABLE? Describe cols?
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the SafeNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Percentage of shared occurrences of pollinators and plants across studies.
-Species on the left are highly shared across studies, while species on the right are found only in a few or a single study.
-As a reference, we have indicated the 50th percentile with a black dashed vertical line."}
-
-#Load data (done it in previous chunk)
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.01) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -12, hjust=0.1, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-p1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.015, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  plot.title = element_text(size = 8, face = "bold"), 
-  axis.text.y = element_text(size=6),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-xlab(NULL) +
-ylab(NULL) + 
-ggtitle("Pollinators")+
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.25) + 
-annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
-annotate(geom="text", x=20, y=0.89, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
-scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-annotate(geom="text", x=-500, y=1.15, label="(b)",color="black", size = 3.5) +
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-#Plot
-p2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.015, fill = "#009E73") +
-ylab("Shared plants (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         axis.title.x = element_text(face = "bold", size = 8),            plot.title = element_text(size = 8, face = "bold"),
-         axis.text.y = element_text(size=6),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-xlab("Species") +
-ylab(NULL) + 
-ggtitle("Plants") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.25) + 
-annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-annotate(geom="text", x=20, y=0.73, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 1.5)+
-annotate(geom="text", x=20, y=-0.08, label="Common spp."
-      ,color="black",  size = 1.5) +
-annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
-      ,color="black",  size = 1.5) + 
-scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")
-
-#Prepare panel with plots
-plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Occurrence across studies (%)") + 
-theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 1))
-
-```
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if name of the species is uncertain (e.g., species complex or species alike).
-
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at European level by combining the recently published checklists of bees [@ghisbain2023], syrphids [@kovcic2023] and butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the SafeNet database at European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-
-## Habitat type and bioclimatic region
-
-The different sites per study were classified with a habitat type by the authors.
-As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
-For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
-Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the SafeNet database (see habitat type definition in supplementary material).
-These categories allow a quick comparison and understanding of the habitat types from the database.
-However, we advise authors to revise this classification if they intend to rely on this field for their analyses as this is a non-fully objective process.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
-
-\newpage
-
-# Results
-
-PART A
-
-Provide number of species per taxonomic group and number of interactions
-Something always of interest: provide % of interactions of _Apis mellifera_?
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic representation of bee genera in the SafeNet database within the context of all European bee genera.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Grey bars indicate the total number of species per genus at the European level."}
-
-#Load libraries
-library(ggplot2)
-library(ggnewscale)
-library(ggtreeExtra)
-library(ggtree)
-library(ggpattern)
-library(patchwork)
-library(cowplot)
-
-#Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
-
-#Circular layour
-ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.063, size=2.9, fontface=1) +
-scale_size(range = c(0,4.5), guide = 'none')  +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-guides(fill=guide_legend(), size = guide_legend()) +
-theme(legend.position = "bottom") +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
-                    pwidth=0.13, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0.137,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-guides(fill = "none")+
-scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
-geom_fruit(geom=geom_bar,
-                    mapping=aes(y=label, x=Spp_number_Europe),
-                    pwidth=0.12, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1,
-                         vjust = 0.5,
-                         nbreak = 1,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-annotate("text", x = 14.2, y = 70, label = "SafeNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.1, y = 70, label = "SafeNet\n interactions", size= 3, fontface= "bold")+ 
-new_scale_fill() +
-geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
-annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
-annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
-annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
-annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
-annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
-annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
-coord_cartesian(clip = "off")+
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.6,0),
-      legend.direction = "horizontal")
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-Phylogenetic representation of the plant families in the SafeNet database within the context of all flowering plant families occurring in Europe.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-PART B
-
-Explain habitat and bioregion
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=6, fig.height=8}
-
-#Read interaction data 
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
-#Load data (done it in earlier chunks)
-#Select cols of interest
-data1 = data %>% 
-select(Study_id, Network_id, Authors_habitat,
-  Corine_land_cover, SafeNet_habitat, Bioregion,
-  Latitude, Longitude, Day, Month, Year, Plant_rank,
-  Plant_status, Plant_matchtype, Plant_accepted_name,Plant_order,
-  Pollinator_rank, Pollinator_status, Pollinator_matchtype,
-  Pollinator_accepted_name, Pollinator_order, Interaction)
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Study_id, SafeNet_habitat) %>% 
-distinct() %>%  
-group_by(SafeNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(SafeNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(SafeNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 8),
-      axis.text = element_text(size=6)) +
-xlab(NULL) +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 12, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = c("#ffd67a","skyblue", "lightpink2", "tomato3"),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
-ggtitle("(a)") +
-theme(plot.title = element_text(size = 14 ,hjust = -0.25, vjust = -2.5))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "black",width = 0.9) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 8),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p3 = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray10", width = 0.9) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 8, vjust = -1),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 14,hjust = -0.3, vjust = 2))
-
-a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
-#Plot panel
-p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)
-
-```
-
-FIGURE 5? CLIFFHANGER OF NETWORK METRICS
-
-## Taxonomy and taxonomic coverage
-
-## Habitat type and bioclimatic region
-
-# Discussion
-
-Geographic and taxonomic biases!
-
-Contextualise European diversity of plant and pollinator species with global biodiveristy
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-# Acknowledgments
-
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data. This research was funded by the H2020 European project Safeguard (101003476).
-
-# Author contributions
-
-# Supplementary material
-
-Supplementary text 1
-
-Habitat definitions:
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-
-2) **Agricultural margings**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an  urban setting. Botanical gardens are included in this category.
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-
-5) **Forest/woodland understory**: Any plant community sampled under a wooded group of plants.
-The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included here agro-forestry areas and open to dense forest.
-Note that we have excluded from this category forest that contains sclerophyllous vegetation.
-
-6) **Semi-natural grassland**: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
-
-7) **Pastures**: Any type of low growing plant community that is highly influenced by human disturbance.
-For instance, agriculture, mowing, moderate to high grazing or urban environments.
-Note that this category also includes old pastures with regrowth of woody vegetation.
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have include in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-
-10) **Natural grassslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
-
-12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.

### 2024-03-07 Manuscript/Ms.qmd
Edits

@@
         \captionsetup[figure]{labelformat=empty}
+        \captionsetup[table]{labelformat=empty}
 
@@
 
+#Prepare table with all info 
+#Create variables
+vars = colnames(safenet_database)
+#Create table
+d = tibble(Variable = vars)
+
+d = d %>% 
+mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
+filter(!Variable == "Longitude")
+
+#Add descriptor
+my_tibble = d %>%
+  mutate(Description = case_when(
+    Variable == "Study_id" ~ "Identifier of the study",
+    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
+    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
+    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
+    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
+    Variable == "Bioregion" ~ "European biogiographical regions",
+    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
+    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
+    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees (two columns)",
+    Variable == "Date" ~ "Year, month and day when the observation took place",
+    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
+    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
+    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
+    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
+    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
+    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
+    Variable == "Plant_unsure_id" ~ "Binary category to indicate if the plant species name is unsure (Yes) or not (No)",
+    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
+    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
+    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_unsure_id" ~ "Binary category to indicate if the pollinator species name is unsure (Yes) or not (No)",
+    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+    TRUE ~ "Other"
+  ))
+
 ```
@@
 
-# Introduction
+# 1 | INTRODUCTION
 
@@
 
-# Methods
+# 2 | METHODS
 
@@
 
-#Data structure
-
-EXPLAIN!!!
+# Data structure
+
+The SafeNet database is available in both .csv and .rds formats.
+The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
+The data is organized in three different files named interaction data, flower count data and metadata.
+The interaction file contains `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
+To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
+To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study.
+In addition, information about habitat type and bioclimatic region are also provided in this file.
+The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the old plant name provided by the authors, site, and date if available.
+Only xxxx% of studies contain floral count information.
+The metadata file contains [...]
 
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-d = d %>% 
-mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
-filter(!Variable == "Longitude")
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Binary category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Binary category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    TRUE ~ "Other"
-  ))
 
@@
 
-# Results
+# 3 | RESULTS
 
@@
 
-# Discussion
+# 3 | DISCUSSION
 

### 2024-03-20 Manuscript/Ms.qmd
back to the start

--- 2024-03-06 Manuscript/Ms.qmd
@@
 #Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/safenet_database.rds")
+safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")
 
@@
 
+#Prepare table with all info 
+#Create variables
+vars = colnames(safenet_database)
+#Create table
+d = tibble(Variable = vars)
+
+d = d %>% 
+mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
+filter(!Variable == "Longitude")
+
+#Add descriptor
+my_tibble = d %>%
+  mutate(Description = case_when(
+    Variable == "Study_id" ~ "Identifier of the study",
+    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
+    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
+    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
+    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
+    Variable == "Bioregion" ~ "European biogiographical regions",
+    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
+    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
+    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
+    Variable == "Date" ~ "Year, month and day when the observation took place",
+    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
+    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
+    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
+    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
+    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
+    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
+    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
+    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
+    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
+    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
+    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
+    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
+    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
+        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
+
+    TRUE ~ "Other"
+  ))
+
 ```

### 2024-03-20 Manuscript/Table_data_structure.Rmd
back to the start

@@
 #Prepare table with data structure
-data = readRDS("../Data/3_Final_data/safenet_database.rds")
+data = readRDS("../Data/3_Final_data/Interaction_data.rds")
 
@@
     Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Binary category to indicate if the plant species name is unsure (Yes) or not (No)",
+    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
     Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
@@
     Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Binary category to indicate if the pollinator species name is unsure (Yes) or not (No)",
+    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
     Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
+        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
+        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
+
     TRUE ~ "Other"

### 2024-03-20 Scripts/1_Processing_networks/10_Vanbergen.R

@@ -9,6 +9,10 @@ source("Scripts/Processing/Functions/Change_str.R")
 #Prepare interaction data ----
 data = read.csv("Data/1_Raw_data/10_Vanbergen/Interaction_data.csv")
 
+#Fix dates
+data = data %>%
+separate(Date, into = c("Day", "Month", "Year"), sep = "/")
+
 #Check what is missing and what we have
 #compare_variables(check_interaction_data, data)
 
@@ -21,10 +25,6 @@ mutate(Flower_data_merger = NA) %>%
 rename(Plant_species = Best.plant.ID) %>% 
 rename(Pollinator_species = Best.ID) %>% 
 rename(Site_id = Site) %>% 
-mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date)) %>% 
-select(!Date) %>% 
 mutate(Interaction = 1) %>% 
 mutate(Country = "France") %>% 
 mutate(Locality = "Côte d'or") %>% 

### 2024-03-21 Manuscript/Ms.qmd
Edits on Figure 1

@@
 
+#Fix dates
+data = data %>%
+dplyr::mutate(Year = lubridate::year(Date), 
+                Month = lubridate::month(Date), 
+                Day = lubridate::day(Date))
+
 #Prepare coordinates
@@
 ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -12, hjust=0.2, size=10))
+theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))
 
@@
 #Plot
-p1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.015, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  plot.title = element_text(size = 8, face = "bold"), 
-  axis.text.y = element_text(size=5),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-xlab(NULL) +
-ylab(NULL) + 
-ggtitle("Pollinators")+
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.25) + 
-annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
-annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
-scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
+#p1 = ggplot(poll_spread, 
+#aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
+#geom_bar(stat = "identity", 
+#           color = "#D55E00",
+#           lwd = 0.015, fill = "#D55E00") +
+#ylab("Shared pollinators (%)") +
+#theme_minimal()+
+#theme(axis.text.x = element_blank(),
+#  axis.ticks.x=element_blank(),
+#  axis.ticks.y = element_line(size = 0.2),
+#  plot.title = element_text(size = 8, face = "bold"), 
+#  axis.text.y = element_text(size=5),
+#  panel.grid.major.x = element_blank(),
+#  panel.grid.minor.y = element_blank(),
+#  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
+#xlab(NULL) +
+#ylab(NULL) + 
+#ggtitle("Pollinators")+
+#geom_vline(xintercept = quantile_50, linetype="dashed", 
+#color = "black", size=0.25) + 
+#annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = #1.5) +
+#annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  #fontface = 'italic', size = 1.5)+
+#scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
+#annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
+#coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, #max(poll_spread$Percent_total)), clip="off")
+#
+##----------------#
+##Plants
+##----------------#
+#plant_spread = data %>% 
+#select(Plant_rank, Plant_accepted_name, Study_id) %>% 
+#group_by(Study_id) %>% 
+#filter(Plant_rank == "SPECIES") %>% 
+#distinct(Plant_accepted_name) %>% 
+#ungroup() %>% 
+#count(Plant_accepted_name) %>% 
+#rename(n_studies_shared = n) %>% 
+#arrange(-n_studies_shared) %>% 
+#mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
+#  
+##Seelect quantile 50% from the total number of species
+#spp_number = plant_spread %>%  
+#select(Plant_accepted_name) %>% 
+#n_distinct()
+#df = seq(1, spp_number)
+#quantile_50 = quantile(df, 0.5)
+#
+##Obtain most common plant for plotting
+#most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
+#select(Plant_accepted_name) %>% pull() 
+#
 #Plot
-p2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.015, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         axis.title.x = element_text(face = "bold", size = 7),            plot.title = element_text(size = 8, face = "bold"),
-         axis.text.y = element_text(size=5),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2),
-      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
-xlab("Species") +
-ylab("Occurrence across studies") + 
-ggtitle("Plants") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.25) + 
-annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 1.5)+
-annotate(geom="text", x=20, y=-0.08, label="Common spp."
-      ,color="black",  size = 1.5) +
-annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
-      ,color="black",  size = 1.5) + 
-scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")
+#p2 = ggplot(plant_spread, 
+#aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
+#geom_bar(stat = "identity", 
+#           color = "#009E73",
+#           lwd = 0.015, fill = "#009E73") +
+#theme_minimal()+
+#theme(axis.text.x = element_blank(),
+#         axis.ticks.x=element_blank(),
+#         axis.ticks.y = element_line(size = 0.2),
+#         axis.title.x = element_text(face = "bold", size = 7),            plot.title = #element_text(size = 8, face = "bold"),
+#         axis.text.y = element_text(size=5),
+#         panel.grid.major.x = element_blank(),
+#         panel.grid.minor.y = element_blank(),
+#         panel.grid.major.y = element_line(size=0.2, linetype =  2),
+#      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
+#xlab("Species") +
+#ylab("Occurrence across studies") + 
+#ggtitle("Plants") + 
+#geom_vline(xintercept = quantile_50, linetype="dashed", 
+#          color = "black", size=0.25) + 
+#annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
+#annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
+#      ,color="black",  fontface = 'italic', size = 1.5)+
+#annotate(geom="text", x=20, y=-0.08, label="Common spp."
+#      ,color="black",  size = 1.5) +
+#annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
+#      ,color="black",  size = 1.5) + 
+#scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = #label_number(accuracy = 0.01)) +
+#coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, #max(plant_spread$Percent_total)), clip="off")
+#
 
@@
 
-p3 = ggplot(data2, aes(Year, Counts)) +
+p2  = ggplot(data2, aes(Year, Counts)) +
 geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
@@
 theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 7, vjust = -1.1),
+  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
   axis.title.x = element_text(face = "bold", size = 8),
@@
 ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.35, vjust = 0.8))
-
-plot2 = p3 / p1 / p2
+theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -5))
+
+main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
+#Select main orders
+data1 = data %>% 
+select(Pollinator_order, Pollinator_accepted_name) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+group_by(Pollinator_order) %>% 
+summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.),
+          Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
+
+# Combine Interactions and Species into a single column
+data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
+
+#Set order of levels for plotting
+data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))
+
+#Plot
+p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
+geom_bar(stat = "identity", width = 0.7) +
+theme_classic() +
+scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator order") +
+xlab(NULL) +
+ylab("Percentage") + 
+scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
+theme(axis.text = element_text(size=6), 
+     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
+     axis.title.x = element_text(face="bold", size = 7),
+     legend.position = "top",
+     legend.title = element_text(size = 5, face = "bold"),
+     legend.text = element_text(size = 5),
+     legend.key.size = unit(3, "mm"),
+     legend.margin=margin(0,0,0,0),
+     legend.box.margin=margin(-8,-5,-5,-5)) +
+guides(fill=guide_legend(nrow=2,byrow=TRUE))+
+ggtitle("(c)") +
+theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6))
+
+#plot2 = p3 / p1 / p2 
+plot2 = p2 / p3
 
 #Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.58, 0.4))
+map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.485, 0.8))
 

### 2024-03-27 Manuscript/1_Figure1.qmd
edits

--- 2024-03-21 Manuscript/Ms.qmd
@@
 ---
-title: "SafeNet: A European database of plant-pollinator interactions"
+latex-tinytex: false
 format:
@@
         \newcommand{\elandscape}{\end{landscape}}
-        \usepackage{caption}
-        \captionsetup[figure]{labelformat=empty}
-        \captionsetup[table]{labelformat=empty}
-
-bibliography: references.bib
-latex-tinytex: false
-link-citations: TRUE
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
 ---
@@
 library(dplyr) #For handling fata 
+library(ggplot2) #For plotting
+library(giscoR) #For plotting
 library(sf) #For handling coordinates and plotting
+library(patchwork)#For plotting (binding plots)
+library(ggstar) #For plotting (cool shapes)
 library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
-library(readr)
-#Load libraries for plotting
-library(giscoR) #Maps
-library(ggstar) #cool shapes as points
-library(ggplot2) #general plotting
-library(ggtree) #phylo trees
-library(ggpattern) #cool shapes in barplots
-library(patchwork) #plot organisation
-library(cowplot) #plot organisation
-library(viridis) #colours
-library(ggnewscale)
-library(ggtreeExtra)
-library(kableExtra)
-
+library(tidyr)
+library(viridis)
 ```
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}
+
 #Load data
 data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
-#Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
-#Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
-                    header = T)
-#Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")
-
-#Work around to get the three cols
-#Note that month 1 and day 1 indicates the presence of a missing value
-#At least in Europe :)
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Read author list 
-authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
-#Print authors
-authors = authors %>% 
-summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
-pull()
-#Read affiliation list 
-aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
-aff_list = aff_list %>% 
-summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
-pull()
-
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-#Prepare table with all info 
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-d = d %>% 
-mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
-filter(!Variable == "Longitude")
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
-        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
-
-    TRUE ~ "Other"
-  ))
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first three and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
-Here, we present the SafeNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
-This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
-We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.
-
-**Main Types of Variables Included:** SafeNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region and further taxonomic rank information for both plant and pollinator species are also provided.
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants.
-The main focus of this database are insect pollinators and accounts for `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
-
-**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats.
-The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
-
-**KEYWORDS**\
-\begingroup
-\small
-plant-pollinator interactions, flowering plants, Angiosperms, pollinators, Hymenoptera, Diptera, Lepidoptera, Coleoptera, Europe
-\endgroup
-\pagebreak
-
-# 1 | INTRODUCTION
-
-The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023].
-Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes at large scales.
-
-Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the SafeNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-SafeNet aims to cover a wide range of taxonomic groups and habitats while also provides other variables of interest that allow to better control for the ecological context and sampling methods.
-In addition, SafeNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
-We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
-
-# 2 | METHODS
-
-## Data acquisition
-
-This database is the result of one of the working packages of the European project Safeguard.
-The SafeNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-Data acquisition followed a non-systematic approach.
-First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions of countries within the database.
-To keep high quality standards that will allow robust future ecological research, we only included: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**).
-
-## Dataset description
-
-SafeNet includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
-These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the SafeNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Number of studies by year within the database.
-(\\textbf{c}) Percentage of shared occurrences of pollinators and plants across studies.
-Species on the left are highly shared across studies, while species on the right are found only in a few or a single study.
-As a reference, we have indicated the 50th percentile with a black dashed vertical line."}
 
@@
 ```
-
-\newpage
-
-# Data structure
-
-The SafeNet database is available in both .csv and .rds formats.
-The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
-The data is organized in three different files named interaction data, flower count data and metadata.
-The interaction file contains `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
-To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study.
-In addition, information about habitat type and bioclimatic region are also provided in this file.
-The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the old plant name provided by the authors, site, and date if available.
-Only xxxx% of studies contain floral count information.
-The metadata file contains [...]
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.}") %>%
-  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
-kable_styling(position = "center")
-
-```
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if name of the species is uncertain (e.g., species complex or species alike).
-
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at European level by combining the recently published checklists of bees [@ghisbain2023], syrphids [@kovcic2023] and butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the SafeNet database at European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-
-## Habitat type and bioclimatic region
-
-The different sites per study were classified with a habitat type by the authors.
-As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
-For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
-Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the SafeNet database (see habitat type definition in supplementary material).
-These categories allow a quick comparison and understanding of the habitat types from the database.
-However, we advise authors to revise this classification if they intend to rely on this field for their analyses as this is a non-fully objective process.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
-
-\newpage
-
-# 3 | RESULTS
-
-## Taxonomy and taxonomic coverage
-
-Provide number of species per taxonomic group and number of interactions
-Something always of interest: provide % of interactions of _Apis mellifera_?
-
-phylogenetic and geographical biases
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic representation of bee genera in the SafeNet database within the context of all European bee genera.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Grey bars indicate the total number of species per genus at the European level."}
-
-#Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
-
-#Circular layour
-ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.063, size=2.9, fontface=1) +
-scale_size(range = c(0,4.5), guide = 'none')  +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-guides(fill=guide_legend(), size = guide_legend()) +
-theme(legend.position = "bottom") +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
-                    pwidth=0.13, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0.137,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-guides(fill = "none")+
-scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
-geom_fruit(geom=geom_bar,
-                    mapping=aes(y=label, x=Spp_number_Europe),
-                    pwidth=0.12, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1,
-                         vjust = 0.5,
-                         nbreak = 1,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-annotate("text", x = 14.2, y = 70, label = "SafeNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.1, y = 70, label = "SafeNet\n interactions", size= 3, fontface= "bold")+ 
-new_scale_fill() +
-geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
-annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
-annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
-annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
-annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
-annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
-annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
-coord_cartesian(clip = "off")+
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.6,0),
-      legend.direction = "horizontal")
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-Phylogenetic representation of the plant families in the SafeNet database within the context of all flowering plant families occurring in Europe.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-\newpage
-
-## Habitat type and bioclimatic region
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Load data (done it in earlier chunks)
-#Select cols of interest
-data1 = data %>% 
-select(Study_id, Network_id, Authors_habitat,
-  SafeNet_habitat, Bioregion,
-  Latitude, Longitude, Day, Month, Year, Plant_rank,
-  Plant_status, Plant_matchtype, Plant_accepted_name,Plant_order,
-  Pollinator_rank, Pollinator_status, Pollinator_matchtype,
-  Pollinator_accepted_name, Pollinator_order, Interaction)
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Study_id, SafeNet_habitat) %>% 
-distinct() %>%  
-group_by(SafeNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(SafeNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(SafeNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 8),
-      axis.text = element_text(size=6),
-      panel.border = element_rect(size=1)) +
-xlab(NULL) +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 12, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = viridis(4),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-      axis.ticks.x = element_line(color="black"),
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 8),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black'),
-plot.margin = margin(0,0,0,0)) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-#Plot panel
-p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
-
-```
-
-**FIGURE 5?** CLIFFHANGER OF NETWORK METRICS
-
-\newpage
-
-# 3 | DISCUSSION
-
-Geographic and taxonomic biases!
-
-Contextualise European diversity of plant and pollinator species with global biodiveristy
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-# Acknowledgments
-
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data. This research was funded by the H2020 European project Safeguard (101003476).
-
-# Author contributions
-
-# Supplementary material
-
-**Supplementary text 1**
-
-Habitat definitions:
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-
-2) **Agricultural margings**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an  urban setting. Botanical gardens are included in this category.
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-
-5) **Forest/woodland understory**: Any plant community sampled under a wooded group of plants.
-The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included here agro-forestry areas and open to dense forest.
-Note that we have excluded from this category forest that contains sclerophyllous vegetation.
-
-6) **Semi-natural grassland**: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
-
-7) **Pastures**: Any type of low growing plant community that is highly influenced by human disturbance.
-For instance, agriculture, mowing, moderate to high grazing or urban environments.
-Note that this category also includes old pastures with regrowth of woody vegetation.
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have include in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-
-10) **Natural grassslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
-
-12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S1}.
-Coverage of the SafeNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
-
-#Prepare figure for supplementary material
-#Coverage across high taxonomic ranks of bees, syrphids and butterflies
-#--------------------------------------------------------------------------------------------#
-#Start from bees----
-#Bee families
-bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
-            "Andrenidae", "Colletidae","Melittidae")
-
-#Species per family within SafeNet
-bee_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-filter(Pollinator_family %in% bee_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
-#Species per family within Europe
-bee_spp_europe =  master_list %>% 
-select(Family, Genus, Species, GenusAndSpecies) %>% 
-filter(Family %in% bee_fam) %>% 
-group_by(Family) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-rename(Pollinator_family = Family) %>% 
-mutate(Group = "Europe")
-
-#Bind both datasets for plotting
-bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
-
-p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-ylab("Number of species") +
-scale_fill_manual(values = c(viridis(2)))+
-ggtitle("Bee family coverage")
-#--------------------------------------------------------------------------------------------#
-#Continue with syrphids----
-#As we don't have subfamilies, let's obtained them from the master list
-syrphids = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
-rename(Pollinator_accepted_name = GenusAndSpecies)
-#Syrphid species safenet
-syrphid_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_family== "Syrphidae") %>% 
-distinct()
-#Add subfamily info
-syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
-#Summarise spp per subfamily
-syrphid_spp_safenet = syrphid_spp_safenet %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
-#Get spp per subfamily at european level
-syrphid_spp_europe = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
-#Plot
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c(viridis(2)))+
-ggtitle("Syrphid subfamily coverage")
-#--------------------------------------------------------------------------------------------#
-#Continue with butterflies----
-#Do it at family level
-#Create vector with butterfly families
-butt_fam = c(unique(master_list1$family))
-#Species per family within SafeNet
-butterfly_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Lepidoptera") %>% 
-filter(Pollinator_family %in% butt_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
-#Species per family within Europe
-butterfly_spp_europe =  master_list1 %>% 
-select(family, genus, scientificName) %>% 
-group_by(family) %>% 
-summarise(Spp_number = n_distinct(scientificName)) %>% 
-rename(Pollinator_family = family) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
-#Plot
-p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c(viridis(2)))+
-ggtitle("Butterfly family coverage")
-
-#--------------------------------------------------------------------------------------------#
-p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 12, face = "bold"),
-      axis.title.y = element_text(face = "bold"))
-
-```

### 2024-03-27 Manuscript/2_Figure_2.Rmd
edits

---
output:
  pdf_document:
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.", fig.width=10, fig.height=6}

#Load libraries
library(dplyr)
library(ggplot2)
library(scales)
library(patchwork)

#Pollinators-----
poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
#Create dataset
d_poll = poll_output$iNextEst$Network_id
#Check cols
d_poll$col = d_poll$method

#Now create tibble with same structure
a_poll = tibble(t = pollinators_curves$Study_sampled, 
           method = as.character(pollinators_curves$iteration), 
           qD = pollinators_curves$Unique_spp_cumulative,
           col = "collectors")

#Bind datasets 
d_poll1 = bind_rows(d_poll, a_poll)
d_poll1_point = d_poll1 %>%  filter(method == "observed")
d_poll1_lines = d_poll1 %>%  filter(!method == "observed")

#Order levels
d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
#Plot
p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
geom_line(aes(linetype = col),size = 0.65) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_poll1_point, col = "black", size=2) + 
ylab("Pollinator species") +
xlab("Sampling sites") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off")  +
xlim(0,3000)+
ylim(0, 3000) +
ggtitle("(a)") +
theme(plot.margin = margin(r = 12))

#Plants-----
plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
#Create dataset
d_plant = plant_output$iNextEst$Network_id
#Check cols
d_plant$col = d_plant$method

#Now create tibble with same structure
a_plant = tibble(t = plant_curves$Study_sampled, 
           method = as.character(plant_curves$iteration), 
           qD = plant_curves$Unique_spp_cumulative,
           col = "collectors")

#Bind datasets 
d_plant1 = bind_rows(d_plant, a_plant)
d_plant1_point = d_plant1 %>%  filter(method == "observed")
d_plant1_lines = d_plant1 %>%  filter(!method == "observed")

#Order levels
d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
#Plot
p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
geom_line(aes(linetype = col), size = 0.65) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_plant1_point, col = "black", size=2) + 
ylab("Plant species") +
xlab("Sampling sites") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off")  +
xlim(0,3000)+
ylim(0, 3000)+
ggtitle("(b)")+
theme(plot.margin = margin(r = 12))

#Interactions-----
int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
#Create dataset
d_int = int_output$iNextEst$Network_id
#Check cols
d_int$col = d_int$method

#Now create tibble with same structure
a_poll = tibble(t = interactions_curves$Study_sampled, 
           method = as.character(interactions_curves$iteration), 
           qD = interactions_curves$Unique_spp_cumulative,
           col = "collectors")

#Bind datasets 
d_int1 = bind_rows(d_int, a_poll)
d_int1_point = d_int1 %>%  filter(method == "observed")
d_int1_lines = d_int1 %>%  filter(!method == "observed")

#Plot
#Order levels
d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
#Plot
p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
geom_line(aes(linetype = col), size = 0.65) +
scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
geom_ribbon(aes(ymin = qD.LCL, 
    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
scale_fill_manual(values = c("gray", "black", "red")) +
geom_point(data = d_int1_point, col = "black", size=2) + 
ylab("Interactions") +
xlab("Sampling sites") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off")  +
xlim(0,3000) +
ylim(0, 30000)+
ggtitle("(c)")

#----------------#
#Pollinators
#----------------#
#Read data
data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")%>% 
mutate(Network_id = paste0(Study_id, Network_id))  
#First select cols of rank, accepted name and study ID
#Then select unique levels by study ID
#Now sum levels (maximum can be as maximum number of studies)
poll_spread = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
group_by(Network_id) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
distinct(Pollinator_accepted_name) %>% 
ungroup() %>% 
count(Pollinator_accepted_name) %>% 
rename(n_networks_shared = n) %>% 
arrange(-n_networks_shared) %>% 
mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))

#Seelect quantile 50% from the total number of species
spp_number = poll_spread %>%  
select(Pollinator_accepted_name) %>% 
n_distinct()
df = seq(1, spp_number)
quantile_50 = quantile(df, 0.5)

#Obtain most common pollinator for plotting
most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
select(Pollinator_accepted_name) %>% pull()

#Plot
o1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.03, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
xlab("Pollinator species") +
ylab("Occurrence across sampling sites") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.35) + 
annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=200, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3)  +
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
                ylim=c(0,0.7), clip="off") +
ggtitle("(d)")

#----------------#
#Plants
#----------------#
plant_spread = data %>% 
select(Plant_rank, Network_id, Plant_accepted_name) %>% 
group_by(Network_id) %>% 
filter(Plant_rank == "SPECIES") %>% 
distinct(Plant_accepted_name) %>% 
ungroup() %>% 
count(Plant_accepted_name) %>% 
rename(n_networks_shared = n) %>% 
arrange(-n_networks_shared) %>% 
mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))

#Seelect quantile 50% from the total number of species
spp_number = plant_spread %>%  
select(Plant_accepted_name) %>% 
n_distinct()
df = seq(1, spp_number)
quantile_50 = quantile(df, 0.5)

#Obtain most common plant for plotting
most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
select(Plant_accepted_name) %>% pull()

o2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.03, fill = "#009E73") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
xlab("Plant species") +
ylab("Occurrence across sampling sites") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.35) + 
annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 3)+
annotate(geom="text", x=150, y=-0.05, label="Common spp."
      ,color="black",  size = 3) +
annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
      ,color="black",  size = 3) + 
scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
                ylim=c(0, 0.7), clip="off") +
ggtitle("(e)")

plant_poll = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
rarefaction_curve1 = readRDS("../Data/Working_files/rarefaction_curve_plant_poll.rds")

#Plot curves
pp1 = plant_poll %>% 
ggplot(aes(x = Plant_sampled, y= Unique_spp_cumulative, group = iteration)) +
geom_line(color = "gray") +
geom_smooth(data = rarefaction_curve1, 
  aes(x = Plant_sampled, y = mean_curve), inherit.aes = FALSE,method = "loess", color = "black", size = 0.65, linetype = "solid", se = FALSE, span = 0.5) +  
geom_point(data = rarefaction_curve1, x = max(rarefaction_curve1$Plant_sampled), y = max(rarefaction_curve1$mean_curve), inherit.aes = FALSE, size=2) +
ylab("Pollinator species") +
xlab("Plant species") +
theme_bw() +
coord_cartesian(expand = FALSE, clip = "off") +
ggtitle("(f)") +
scale_x_continuous(breaks = seq(0, 1400, by = 400), limits=c(0,1400)) +
ylim(0,2500)

#Plot all
a = p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.title=element_blank(),
      axis.title = element_text(size=12, face = "bold"),
      axis.text = element_text(size=10),
      plot.title = element_text(size=18))
b = o1 + o2 + pp1 & 
  theme(axis.title = element_text(size=12, face = "bold"),
        axis.text = element_text(size=10),
      plot.title = element_text(size=18))

a/b

```

### 2024-03-27 Manuscript/3_Figure_3.Rmd
edits

--- 2024-03-07 Manuscript/Ms.qmd
@@
 ---
-title: "SafeNet: A European database of plant-pollinator interactions"
-format:
-  pdf:
-    include-in-header: 
-      text: |
-        \usepackage{lscape}
-        \newcommand{\blandscape}{\begin{landscape}}
-        \newcommand{\elandscape}{\end{landscape}}
-        \usepackage{caption}
-        \captionsetup[figure]{labelformat=empty}
-        \captionsetup[table]{labelformat=empty}
-
-bibliography: references.bib
-latex-tinytex: false
-link-citations: TRUE
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
+output:
+  pdf_document:
+header-includes:
+- \usepackage{caption}
+- \captionsetup[figure]{labelformat=empty}
+- \captionsetup[table]{labelformat=empty}
 ---
 
-```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr) #For handling fata 
-library(sf) #For handling coordinates and plotting
-library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
-library(readr)
-#Load libraries for plotting
-library(giscoR) #Maps
-library(ggstar) #cool shapes as points
-library(ggplot2) #general plotting
-library(ggtree) #phylo trees
-library(ggpattern) #cool shapes in barplots
-library(patchwork) #plot organisation
-library(cowplot) #plot organisation
-library(viridis) #colours
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=10, fig.height=10}
+
+#Load libraries
+library(ggplot2)
 library(ggnewscale)
 library(ggtreeExtra)
-library(kableExtra)
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Load data
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
-#Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
-#Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
-                    header = T)
-#Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/safenet_database.rds")
-
-#Work around to get the three cols
-#Note that month 1 and day 1 indicates the presence of a missing value
-#At least in Europe :)
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Read author list 
-authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
-#Print authors
-authors = authors %>% 
-summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
-pull()
-#Read affiliation list 
-aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
-aff_list = aff_list %>% 
-summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
-pull()
-
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-#Prepare table with all info 
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-d = d %>% 
-mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
-filter(!Variable == "Longitude")
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees (two columns)",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Binary category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Binary category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    TRUE ~ "Other"
-  ))
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first three and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
-Here, we present the SafeNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
-This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
-We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.
-
-**Main Types of Variables Included:** SafeNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region and further taxonomic rank information for both plant and pollinator species are also provided.
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants.
-The main focus of this database are insect pollinators and accounts for `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
-
-**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats.
-The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
-
-**KEYWORDS**\
-\begingroup
-\small
-plant-pollinator interactions, flowering plants, Angiosperms, pollinators, Hymenoptera, Diptera, Lepidoptera, Coleoptera, Europe
-\endgroup
-\pagebreak
-
-# 1 | INTRODUCTION
-
-The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023].
-Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes at large scales.
-
-Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the SafeNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-SafeNet aims to cover a wide range of taxonomic groups and habitats while also provides other variables of interest that allow to better control for the ecological context and sampling methods.
-In addition, SafeNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
-We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
-
-# 2 | METHODS
-
-## Data acquisition
-
-This database is the result of one of the working packages of the European project Safeguard.
-The SafeNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-Data acquisition followed a non-systematic approach.
-First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions of countries within the database.
-To keep high quality standards that will allow robust future ecological research, we only included: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**).
-
-## Dataset description
-
-SafeNet includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
-These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the SafeNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Number of studies by year within the database.
-(\\textbf{c}) Percentage of shared occurrences of pollinators and plants across studies.
-Species on the left are highly shared across studies, while species on the right are found only in a few or a single study.
-As a reference, we have indicated the 50th percentile with a black dashed vertical line."}
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -12, hjust=0.2, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-p1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.015, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  plot.title = element_text(size = 8, face = "bold"), 
-  axis.text.y = element_text(size=5),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-xlab(NULL) +
-ylab(NULL) + 
-ggtitle("Pollinators")+
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.25) + 
-annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
-annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
-scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-#Plot
-p2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.015, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         axis.title.x = element_text(face = "bold", size = 7),            plot.title = element_text(size = 8, face = "bold"),
-         axis.text.y = element_text(size=5),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2),
-      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
-xlab("Species") +
-ylab("Occurrence across studies") + 
-ggtitle("Plants") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.25) + 
-annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 1.5)+
-annotate(geom="text", x=20, y=-0.08, label="Common spp."
-      ,color="black",  size = 1.5) +
-annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
-      ,color="black",  size = 1.5) + 
-scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")
-
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p3 = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 7, vjust = -1.1),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-   axis.ticks.x = element_line(color="black"),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.35, vjust = 0.8))
-
-plot2 = p3 / p1 / p2
-
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.58, 0.4))
-
-```
-
-\newpage
-
-# Data structure
-
-The SafeNet database is available in both .csv and .rds formats.
-The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
-The data is organized in three different files named interaction data, flower count data and metadata.
-The interaction file contains `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
-To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study.
-In addition, information about habitat type and bioclimatic region are also provided in this file.
-The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the old plant name provided by the authors, site, and date if available.
-Only xxxx% of studies contain floral count information.
-The metadata file contains [...]
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.}") %>%
-  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
-kable_styling(position = "center")
-
-```
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if name of the species is uncertain (e.g., species complex or species alike).
-
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at European level by combining the recently published checklists of bees [@ghisbain2023], syrphids [@kovcic2023] and butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the SafeNet database at European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-
-## Habitat type and bioclimatic region
-
-The different sites per study were classified with a habitat type by the authors.
-As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
-For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
-Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the SafeNet database (see habitat type definition in supplementary material).
-These categories allow a quick comparison and understanding of the habitat types from the database.
-However, we advise authors to revise this classification if they intend to rely on this field for their analyses as this is a non-fully objective process.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
-
-\newpage
-
-# 3 | RESULTS
-
-## Taxonomy and taxonomic coverage
-
-Provide number of species per taxonomic group and number of interactions
-Something always of interest: provide % of interactions of _Apis mellifera_?
-
-phylogenetic and geographical biases
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic representation of bee genera in the SafeNet database within the context of all European bee genera.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Grey bars indicate the total number of species per genus at the European level."}
+library(ggtree)
+library(ggpattern)
+library(patchwork)
+library(cowplot)
 
@@
 ```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-Phylogenetic representation of the plant families in the SafeNet database within the context of all flowering plant families occurring in Europe.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-\newpage
-
-## Habitat type and bioclimatic region
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can countribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Load data (done it in earlier chunks)
-#Select cols of interest
-data1 = data %>% 
-select(Study_id, Network_id, Authors_habitat,
-  SafeNet_habitat, Bioregion,
-  Latitude, Longitude, Day, Month, Year, Plant_rank,
-  Plant_status, Plant_matchtype, Plant_accepted_name,Plant_order,
-  Pollinator_rank, Pollinator_status, Pollinator_matchtype,
-  Pollinator_accepted_name, Pollinator_order, Interaction)
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Study_id, SafeNet_habitat) %>% 
-distinct() %>%  
-group_by(SafeNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(SafeNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(SafeNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 8),
-      axis.text = element_text(size=6),
-      panel.border = element_rect(size=1)) +
-xlab(NULL) +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 12, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = viridis(4),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-      axis.ticks.x = element_line(color="black"),
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 8),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black'),
-plot.margin = margin(0,0,0,0)) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-#Plot panel
-p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
-
-```
-
-**FIGURE 5?** CLIFFHANGER OF NETWORK METRICS
-
-\newpage
-
-# 3 | DISCUSSION
-
-Geographic and taxonomic biases!
-
-Contextualise European diversity of plant and pollinator species with global biodiveristy
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-# Acknowledgments
-
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data. This research was funded by the H2020 European project Safeguard (101003476).
-
-# Author contributions
-
-# Supplementary material
-
-**Supplementary text 1**
-
-Habitat definitions:
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-
-2) **Agricultural margings**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an  urban setting. Botanical gardens are included in this category.
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-
-5) **Forest/woodland understory**: Any plant community sampled under a wooded group of plants.
-The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included here agro-forestry areas and open to dense forest.
-Note that we have excluded from this category forest that contains sclerophyllous vegetation.
-
-6) **Semi-natural grassland**: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
-
-7) **Pastures**: Any type of low growing plant community that is highly influenced by human disturbance.
-For instance, agriculture, mowing, moderate to high grazing or urban environments.
-Note that this category also includes old pastures with regrowth of woody vegetation.
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have include in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-
-10) **Natural grassslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
-
-12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S1}.
-Coverage of the SafeNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
-
-#Prepare figure for supplementary material
-#Coverage across high taxonomic ranks of bees, syrphids and butterflies
-#--------------------------------------------------------------------------------------------#
-#Start from bees----
-#Bee families
-bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
-            "Andrenidae", "Colletidae","Melittidae")
-
-#Species per family within SafeNet
-bee_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-filter(Pollinator_family %in% bee_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
-#Species per family within Europe
-bee_spp_europe =  master_list %>% 
-select(Family, Genus, Species, GenusAndSpecies) %>% 
-filter(Family %in% bee_fam) %>% 
-group_by(Family) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-rename(Pollinator_family = Family) %>% 
-mutate(Group = "Europe")
-
-#Bind both datasets for plotting
-bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
-
-p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-ylab("Number of species") +
-scale_fill_manual(values = c(viridis(2)))+
-ggtitle("Bee family coverage")
-#--------------------------------------------------------------------------------------------#
-#Continue with syrphids----
-#As we don't have subfamilies, let's obtained them from the master list
-syrphids = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
-rename(Pollinator_accepted_name = GenusAndSpecies)
-#Syrphid species safenet
-syrphid_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_family== "Syrphidae") %>% 
-distinct()
-#Add subfamily info
-syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
-#Summarise spp per subfamily
-syrphid_spp_safenet = syrphid_spp_safenet %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
-#Get spp per subfamily at european level
-syrphid_spp_europe = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
-#Plot
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c(viridis(2)))+
-ggtitle("Syrphid subfamily coverage")
-#--------------------------------------------------------------------------------------------#
-#Continue with butterflies----
-#Do it at family level
-#Create vector with butterfly families
-butt_fam = c(unique(master_list1$family))
-#Species per family within SafeNet
-butterfly_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Lepidoptera") %>% 
-filter(Pollinator_family %in% butt_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
-#Species per family within Europe
-butterfly_spp_europe =  master_list1 %>% 
-select(family, genus, scientificName) %>% 
-group_by(family) %>% 
-summarise(Spp_number = n_distinct(scientificName)) %>% 
-rename(Pollinator_family = family) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
-#Plot
-p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c(viridis(2)))+
-ggtitle("Butterfly family coverage")
-
-#--------------------------------------------------------------------------------------------#
-p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 12, face = "bold"),
-      axis.title.y = element_text(face = "bold"))
-
-```

### 2024-03-27 Manuscript/4_Figure_4.Rmd
edits

--- 2024-03-04 Manuscript/Figure_3.Rmd
@@
 ---
-output: pdf_document
+output:
+  pdf_document:
+header-includes:
+- \usepackage{caption}
+- \captionsetup[figure]{labelformat=empty}
+- \captionsetup[table]{labelformat=empty}
 ---
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=8, fig.height=8}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.", fig.width=8, fig.height=8}
 

### 2024-03-27 Manuscript/5_Figure_5.Rmd
edits

--- 2024-03-06 Manuscript/Figure_4.Rmd
@@
 ---
-output: pdf_document
+output:
+  pdf_document:
+header-includes:
+- \usepackage{caption}
+- \captionsetup[figure]{labelformat=empty}
+- \captionsetup[table]{labelformat=empty}
 ---
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
 Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
@@
 data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
+
+data = data %>%
+dplyr::mutate(Year = lubridate::year(Date), 
+                Month = lubridate::month(Date), 
+                Day = lubridate::day(Date))
+
 #Load data (done it in earlier chunks)
@@
 
-#c("azure3","steelblue", "wheat2", "tomato3")
-
 #Unify labels

### 2024-03-27 Manuscript/Ms.qmd
edits

@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Load data
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
+
+#Read data
+data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")%>% 
+mutate(Network_id = paste0(Study_id, Network_id))  
 #Bee-syrphid Masterlist
@@
 
+## Sampling coverage
+
+bootstraping, inext package
+
 ## Habitat type and bioclimatic region
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
+Grey solid lines represent 100 randomized accumulation curves, the black solid lines represents the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrates the extrapolated curve for approximately 3000 sampling sites.
+The solid black points indicate the number of species and interactions contained in the database.
+Graphs (d-e) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
+Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
+Graph (f) shows the accumulation curve of pollinator species across increasing number of plant species.
+This last graph uses the same color and shape structure as the ones in the top panel.", fig.width=10, fig.height=6, out.width="115%"}
+
+#Pollinators-----
+poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
+pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
+#Create dataset
+d_poll = poll_output$iNextEst$Network_id
+#Check cols
+d_poll$col = d_poll$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = pollinators_curves$Study_sampled, 
+           method = as.character(pollinators_curves$iteration), 
+           qD = pollinators_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_poll1 = bind_rows(d_poll, a_poll)
+d_poll1_point = d_poll1 %>%  filter(method == "observed")
+d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
+
+#Order levels
+d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col),size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_poll1_point, col = "black", size=2) + 
+ylab("Pollinator species") +
+xlab("Sampling sites") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000)+
+ylim(0, 3000) +
+ggtitle("(a)") +
+theme(plot.margin = margin(r = 12))
+
+#Plants-----
+plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
+plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
+#Create dataset
+d_plant = plant_output$iNextEst$Network_id
+#Check cols
+d_plant$col = d_plant$method
+
+#Now create tibble with same structure
+a_plant = tibble(t = plant_curves$Study_sampled, 
+           method = as.character(plant_curves$iteration), 
+           qD = plant_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_plant1 = bind_rows(d_plant, a_plant)
+d_plant1_point = d_plant1 %>%  filter(method == "observed")
+d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
+
+#Order levels
+d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col), size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_plant1_point, col = "black", size=2) + 
+ylab("Plant species") +
+xlab("Sampling sites") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000)+
+ylim(0, 3000)+
+ggtitle("(b)")+
+theme(plot.margin = margin(r = 12))
+
+#Interactions-----
+int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
+interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
+#Create dataset
+d_int = int_output$iNextEst$Network_id
+#Check cols
+d_int$col = d_int$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = interactions_curves$Study_sampled, 
+           method = as.character(interactions_curves$iteration), 
+           qD = interactions_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_int1 = bind_rows(d_int, a_poll)
+d_int1_point = d_int1 %>%  filter(method == "observed")
+d_int1_lines = d_int1 %>%  filter(!method == "observed")
+
+#Plot
+#Order levels
+d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col), size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_int1_point, col = "black", size=2) + 
+ylab("Interactions") +
+xlab("Sampling sites") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000) +
+ylim(0, 30000)+
+ggtitle("(c)")
+
+#----------------#
+#Pollinators
+#----------------#
+#Read data
+data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")%>% 
+mutate(Network_id = paste0(Study_id, Network_id))  
+#First select cols of rank, accepted name and study ID
+#Then select unique levels by study ID
+#Now sum levels (maximum can be as maximum number of studies)
+poll_spread = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
+group_by(Network_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+ungroup() %>% 
+count(Pollinator_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = poll_spread %>%  
+select(Pollinator_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common pollinator for plotting
+most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
+select(Pollinator_accepted_name) %>% pull()
+
+#Plot
+o1 = ggplot(poll_spread, 
+aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
+geom_bar(stat = "identity", 
+           color = "#D55E00",
+           lwd = 0.03, fill = "#D55E00") +
+ylab("Shared pollinators (%)") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+  axis.ticks.x=element_blank(),
+  axis.ticks.y = element_line(size = 0.2),
+  panel.grid.major.x = element_blank(),
+  panel.grid.minor.y = element_blank(),
+  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
+xlab("Pollinator species") +
+ylab("Occurrence across sampling sites") + 
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+color = "black", size=0.35) + 
+annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
+annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
+annotate(geom="text", x=200, y=-0.05, label="Common spp."
+      ,color="black",  size = 3) +
+annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
+      ,color="black",  size = 3)  +
+scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
+coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
+                ylim=c(0,0.7), clip="off") +
+ggtitle("(d)")
+
+#----------------#
+#Plants
+#----------------#
+plant_spread = data %>% 
+select(Plant_rank, Network_id, Plant_accepted_name) %>% 
+group_by(Network_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+ungroup() %>% 
+count(Plant_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = plant_spread %>%  
+select(Plant_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common plant for plotting
+most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
+select(Plant_accepted_name) %>% pull()
+
+o2 = ggplot(plant_spread, 
+aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
+geom_bar(stat = "identity", 
+           color = "#009E73",
+           lwd = 0.03, fill = "#009E73") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+         axis.ticks.x=element_blank(),
+         axis.ticks.y = element_line(size = 0.2),
+         panel.grid.major.x = element_blank(),
+         panel.grid.minor.y = element_blank(),
+         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
+xlab("Plant species") +
+ylab("Occurrence across sampling sites") + 
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+          color = "black", size=0.35) + 
+annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
+annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
+      ,color="black",  fontface = 'italic', size = 3)+
+annotate(geom="text", x=150, y=-0.05, label="Common spp."
+      ,color="black",  size = 3) +
+annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
+      ,color="black",  size = 3) + 
+scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
+coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
+                ylim=c(0, 0.7), clip="off") +
+ggtitle("(e)")
+
+plant_poll = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
+rarefaction_curve1 = readRDS("../Data/Working_files/rarefaction_curve_plant_poll.rds")
+
+#Plot curves
+pp1 = plant_poll %>% 
+ggplot(aes(x = Plant_sampled, y= Unique_spp_cumulative, group = iteration)) +
+geom_line(color = "gray") +
+geom_smooth(data = rarefaction_curve1, 
+  aes(x = Plant_sampled, y = mean_curve), inherit.aes = FALSE,method = "loess", color = "black", size = 0.65, linetype = "solid", se = FALSE, span = 0.5) +  
+geom_point(data = rarefaction_curve1, x = max(rarefaction_curve1$Plant_sampled), y = max(rarefaction_curve1$mean_curve), inherit.aes = FALSE, size=2) +
+ylab("Pollinator species") +
+xlab("Plant species") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off") +
+ggtitle("(f)") +
+scale_x_continuous(breaks = seq(0, 1400, by = 400), limits=c(0,1400)) +
+ylim(0,2500)
+
+#Plot all
+a = p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.title=element_blank(),
+      axis.title = element_text(size=12, face = "bold"),
+      axis.text = element_text(size=10),
+      plot.title = element_text(size=18))
+b = o1 + o2 + pp1 & 
+  theme(axis.title = element_text(size=12, face = "bold"),
+        axis.text = element_text(size=10),
+      plot.title = element_text(size=18))
+
+a/b
+
+```
+
 phylogenetic and geographical biases
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 3}.
 Phylogenetic representation of bee genera in the SafeNet database within the context of all European bee genera.
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
 Phylogenetic representation of the plant families in the SafeNet database within the context of all flowering plant families occurring in Europe.
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
 Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
@@
 Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Load data (done it in earlier chunks)
-#Select cols of interest
-data1 = data %>% 
-select(Study_id, Network_id, Authors_habitat,
-  SafeNet_habitat, Bioregion,
-  Latitude, Longitude, Day, Month, Year, Plant_rank,
-  Plant_status, Plant_matchtype, Plant_accepted_name,Plant_order,
-  Pollinator_rank, Pollinator_status, Pollinator_matchtype,
-  Pollinator_accepted_name, Pollinator_order, Interaction)
 
@@
 #Check pols by habitat
-pollinator_percent = data1 %>%
+pollinator_percent = data %>%
 group_by(SafeNet_habitat, Pollinator_order) %>%
@@
 #Check pols by bioregion
-pollinator_percent_bio = data1 %>%
+pollinator_percent_bio = data %>%
 group_by(Bioregion, Pollinator_order) %>%

### 2024-03-27 Scripts/6_1_Collectors_curve.R

@@ -23,7 +23,7 @@ distinct()
 #Select columns of interest
 data1_plant = data %>% 
 filter(Plant_rank == "SPECIES") %>% 
-select(Study_id,Network_id, Pollinator_accepted_name) %>% 
+select(Study_id,Network_id, Plant_accepted_name) %>% 
 distinct() 
 
 length(unique(data1_plant$Network_id))
@@ -41,7 +41,7 @@ distinct()
 #Data is our dataset with the cols of interest
 #Level: Study_id or Network_id
 #Taxa: Pollinator_accepted_name/Plant_accepted_name/Interaction_id  
-collect = function(data, level_id, taxa) {
+collect = function(data1, level_id, taxa) {
 #Filter studies randomly and sum species per study
 sampled_studies = sample(unique(data %>% pull({{ level_id }})))
 
@@ -94,6 +94,10 @@ xlab("Studies") +
 theme_bw() +
 coord_cartesian(expand = FALSE)
 s2
+
+
+colnames(data1_int)
+
 #Calculate N collectors curves for interactions
 collect_curves = map_dfr(1:100, ~ collect(data1_int, Study_id, Interaction_id), .id="iteration") 
 #Calculate mean across curves
@@ -177,38 +181,52 @@ n1 + n2 + n3
 
 
 #Try now to obtain accumulated plants per pollinator species----
-collect1 = function(data) {
-sampled_pollinator = sample(unique(data1$Pollinator_accepted_name))
+collect1 = function(data1) {
+sampled_plant = sample(unique(data1$Plant_accepted_name))
 data1 %>% 
-filter(Pollinator_accepted_name %in% sampled_pollinator) %>%
-arrange(match(Pollinator_accepted_name, sampled_pollinator)) %>%
-group_by(Plant_accepted_name) %>%
+filter(Plant_accepted_name %in% sampled_plant) %>%
+arrange(match(Plant_accepted_name, sampled_plant)) %>%
+group_by(Pollinator_accepted_name) %>%
 mutate(distinct = row_number() == 1) %>%
 ungroup() %>%
-group_by(Pollinator_accepted_name) %>%
+group_by(Plant_accepted_name) %>%
 mutate(Unique_spp = sum(distinct)) %>%
 ungroup() %>%
-select(Pollinator_accepted_name, Unique_spp) %>%
+select(Plant_accepted_name, Unique_spp) %>%
 distinct() %>%
[…]

### 2024-04-05 Manuscript/6_Figure_6.Rmd
edits

---
output:
  pdf_document:
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}. ", fig.width=10, fig.height=7, fig.align='center'}

# Load libraries
library(dplyr)
library(ggplot2)
library(stringr)
library(patchwork)
library(psych)
# Load data
data = readRDS("../Data/3_Final_data/Interaction_data.rds")
metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
null_networks = readRDS("../Data/Working_files/null_networks.rds")
null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks.rds")

# Prepare data
data = data %>%
dplyr::mutate(Year = lubridate::year(Date), 
              Month = lubridate::month(Date), 
              Day = lubridate::day(Date))

# Subset data for studies with only bumblebees
v = c("13_Karise", "22_Kallnik")
subsetting = unique(data$Study_id)

long_format = data %>% 
mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
filter(!Study_id %in% v) %>% 
select(Network_id, Plant_accepted_name, Pollinator_accepted_name)

# Filter networks larger than 4 plants and 4 pollinator species
v2 = long_format %>% 
group_by(Network_id) %>% 
summarise(Poll_spp = n_distinct(Plant_accepted_name),
          Plant_spp = n_distinct(Pollinator_accepted_name)) %>% 
ungroup() %>% 
filter(Poll_spp >= 4 & Plant_spp >= 4) %>% 
pull(Network_id)

long_format = long_format %>% 
filter(Network_id %in% v2)

# Filter networks with more links than species
links = long_format %>% 
group_by(Network_id, Plant_accepted_name, Pollinator_accepted_name) %>%
distinct() %>% 
group_by(Network_id) %>% 
summarise(links = length(Network_id))

spp = long_format %>% 
group_by(Network_id) %>% 
summarise(n_plants = n_distinct(Plant_accepted_name),
          n_polls = n_distinct(Pollinator_accepted_name)) %>% 
mutate(n_total = n_plants + n_polls)

vector_to_include = left_join(links, spp)

v3 = vector_to_include %>% 
filter(links > n_total) %>% 
pull(Network_id)

long_format = long_format %>% 
filter(Network_id %in% v3)

# Explore how NODFc changes across the Y-X AXES (latitude-Longitude)
long_format_coords = data %>% 
mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
select(Network_id, Latitude, Longitude, Bioregion) %>% 
distinct()

# Merge data
metrics_by_network_coords = left_join(metrics_by_network, long_format_coords)

# Connectance plots
# 1st Latitude
p1 = metrics_by_network_coords %>% 
ggplot(aes(Latitude, Connectance)) +
geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.15, size=2.5, alpha=1) +
theme_bw() +
ylab("Connectance") +
scale_fill_viridis_d() +
scale_colour_viridis_d() +
scale_shape_manual(values = c(
  "Alpine" = 21,  # Circle
  "Atlantic" = 22, # Triangle
  "Boreal" = 23,   # Square
  "Continental" = 24,  # Diamond
  "Mediterranean" = 25  # Cross
))+
ggtitle("(a)")

# 2nd Plot
p2 = metrics_by_network_coords %>% 
ggplot(aes(Latitude, Normalised_nestedness)) +
geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.15, size=2.5, alpha=1) +
theme_bw() +
ylab("NODFc") +
scale_fill_viridis_d() +
scale_colour_viridis_d() +
scale_shape_manual(values = c(
  "Alpine" = 21,  # Circle
  "Atlantic" = 22, # Triangle
  "Boreal" = 23,   # Square
  "Continental" = 24,  # Diamond
  "Mediterranean" = 25  # Cross
  )) +
ggtitle("(b)")

# Merge and process data
d = left_join(metrics_by_network, null_metrics_networks)
d = d %>% 
mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)

p = 0.05 #cutoff probability 95% confidence
critical_value = qnorm(p/2) #double tail probability divide by 2

d1 = d %>%
mutate(infra_over_represented = case_when(
z_score < -abs(critical_value) ~ "Under-represented",
between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
    z_score > abs(critical_value) ~ "Over-represented"
  ))

# 2nd Species (geometric mean)
species_number = long_format %>% 
group_by(Network_id) %>% 
summarise(N_plants = n_distinct(Plant_accepted_name),
          N_pollinators = n_distinct(Pollinator_accepted_name)) %>% 
group_by(Network_id) %>% 
summarise(geometric_mean_spp = geometric.mean(c(N_plants, N_pollinators)))

metrics_by_network_spp_number = left_join(metrics_by_network, species_number)

p3 = metrics_by_network_spp_number %>% 
ggplot(aes(geometric_mean_spp, Connectance)) +
geom_point(pch = 21, fill = "azure3", size=2.5, stroke=0.25, alpha=1) +
theme_bw() +
xlab("Species")+
ggtitle("(c)")

# Plot
p4 = d1 %>% 
ggplot(aes(x = z_score)) +
geom_histogram(aes(y = ..density.., fill = infra_over_represented), 
               bins = 100, alpha = 0.5, color = "black", position = "stack", size=0.1) +
stat_function(data = d1, fun = function(x) dnorm(x, mean = mean(d1$z_score), sd = sd(d1$z_score)) * 1.7, 
                n = 1000, inherit.aes = FALSE, color = "gray18", size = 1) +
theme_bw() +
coord_cartesian(expand = FALSE) +
geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red")+
geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
ylab("Density") +
xlab("Z scores (NODF)")  +
xlim(-6, 9) +
ylim(0, 0.8) +
scale_fill_manual(name = "Observed NODF against null",
                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
                    labels = c("Less nested", "No difference", "More nested"),
                    values = c("Under-represented" = "coral2", 
                               "No statistical difference" = "palegreen3", 
                               "Over-represented" = "cyan3")) +
ggtitle("(d)")

# Layout plots
a = p1 + p2 + plot_layout(guides = 'collect') & theme(legend.position = "right",
                                                      legend.justification = "left")
b = (p3 + p4)  & theme(legend.justification = "left")
a / b  & theme(axis.title =  element_text(size=16, face="bold"),
               plot.title = element_text(size=22),
               panel.border = element_rect(colour = "black",
               fill=NA, size=1),
               axis.text=element_text(size=12))

```

### 2024-04-05 Manuscript/Ms.qmd
edits

@@
 ---
-title: "SafeNet: A European database of plant-pollinator interactions"
+title: "EuPPollNet: A European database of plant-pollinator networks"
 format:
@@
 One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
-Here, we present the SafeNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
+Here, we present the EuPPollNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
 This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
@@
 
-**Main Types of Variables Included:** SafeNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
+**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
 In addition, information about sampling methodology, habitat type, bio-climatic region and further taxonomic rank information for both plant and pollinator species are also provided.
@@
 
-Here, we present the SafeNet database, an acronym derived from the European project Safeguard and plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
+Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
 The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the SafeNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-SafeNet aims to cover a wide range of taxonomic groups and habitats while also provides other variables of interest that allow to better control for the ecological context and sampling methods.
-In addition, SafeNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
+These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+EuPPollNet aims to cover a wide range of taxonomic groups and habitats while also provides other variables of interest that allow to better control for the ecological context and sampling methods.
+In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
 We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
@@
 This database is the result of one of the working packages of the European project Safeguard.
-The SafeNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
+The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
 Data acquisition followed a non-systematic approach.
@@
 
-SafeNet includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+EuPPollNet includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
 The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the SafeNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
+(\\textbf{a}) Approximate locations of the studies of the EuPPollNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
 The sizes of these shapes are proportional to the respective species counts.
@@
 
-The SafeNet database is available in both .csv and .rds formats.
+The EuPPollNet database is available in both .csv and .rds formats.
 The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
@@
 In addition, information about habitat type and bioclimatic region are also provided in this file.
-The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the old plant name provided by the authors, site, and date if available.
+The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the column of "Floral_data_merger.
 Only xxxx% of studies contain floral count information.
@@
 
-To assess the completeness of plant and pollinator species in the SafeNet database at European level, we used the aforementioned checklists for plants and pollinators.
+To assess the completeness of plant and pollinator species in the EuPPollNet database at European level, we used the aforementioned checklists for plants and pollinators.
 Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
@@
 For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
-Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the SafeNet database (see habitat type definition in supplementary material).
+Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the EuPPollNet database (see habitat type definition in supplementary material).
 These categories allow a quick comparison and understanding of the habitat types from the database.
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 3}.
-Phylogenetic representation of bee genera in the SafeNet database within the context of all European bee genera.
+Phylogenetic representation of bee genera in the EuPPollNet database within the context of all European bee genera.
 The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
 Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
+The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
 Grey bars indicate the total number of species per genus at the European level."}
@@
          grid.params=list()) +
-annotate("text", x = 14.2, y = 70, label = "SafeNet spp\n coverage (%)", size= 3, fontface= "bold") +
+annotate("text", x = 14.2, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
 annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.1, y = 70, label = "SafeNet\n interactions", size= 3, fontface= "bold")+ 
+annotate("text", x = 12.1, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
 new_scale_fill() +
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Phylogenetic representation of the plant families in the SafeNet database within the context of all flowering plant families occurring in Europe.
+Phylogenetic representation of the plant families in the EuPPollNet database within the context of all flowering plant families occurring in Europe.
 The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
 Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
+The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
 Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database.
+Proportion of the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
 The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
@@
        legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 8),
+      axis.title = element_text(face = "bold", size = 7),
       axis.text = element_text(size=6),
@@
 scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 12, face = "bold"),
+theme(strip.text.x = element_text(size = 10, face = "bold"),
     strip.background = element_rect(
@@
 color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 8),
+  axis.title = element_text(face = "bold", size = 7),
 axis.text = element_text(size=6)) +
@@
 
-**FIGURE 5?** CLIFFHANGER OF NETWORK METRICS
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
+Graphs (a-b) indicate network connectance and nestedness across the latitudinal range of the studies from the database.
+In addition, the bioclimatic region where each network is located is indicated with points of different shapes and colours.
+Graph (c) shows the the association between network connectances and the geometric mean of plant and pollinator species per network.
+Graph (d) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
+The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
+Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).", fig.width=10, fig.height=7, fig.align='center'}
+
+# Load libraries
+
+library(psych)
+# Load data
+data = readRDS("../Data/3_Final_data/Interaction_data.rds")
+metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
+null_networks = readRDS("../Data/Working_files/null_networks.rds")
+null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks.rds")
+
+# Prepare data
+data = data %>%
+dplyr::mutate(Year = lubridate::year(Date), 
+              Month = lubridate::month(Date), 
+              Day = lubridate::day(Date))
+
+# Subset data for studies with only bumblebees
+v = c("13_Karise", "22_Kallnik")
+subsetting = unique(data$Study_id)
+
+long_format = data %>% 
+mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
+mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
+filter(!Study_id %in% v) %>% 
+select(Network_id, Plant_accepted_name, Pollinator_accepted_name)
+
+# Filter networks larger than 4 plants and 4 pollinator species
+v2 = long_format %>% 
+group_by(Network_id) %>% 
+summarise(Poll_spp = n_distinct(Plant_accepted_name),
+          Plant_spp = n_distinct(Pollinator_accepted_name)) %>% 
+ungroup() %>% 
+filter(Poll_spp >= 4 & Plant_spp >= 4) %>% 
+pull(Network_id)
+
+long_format = long_format %>% 
+filter(Network_id %in% v2)
+
+# Filter networks with more links than species
+links = long_format %>% 
+group_by(Network_id, Plant_accepted_name, Pollinator_accepted_name) %>%
+distinct() %>% 
+group_by(Network_id) %>% 
+summarise(links = length(Network_id))
+
+spp = long_format %>% 
+group_by(Network_id) %>% 
+summarise(n_plants = n_distinct(Plant_accepted_name),
+          n_polls = n_distinct(Pollinator_accepted_name)) %>% 
+mutate(n_total = n_plants + n_polls)
+
+vector_to_include = left_join(links, spp)
+
+v3 = vector_to_include %>% 
+filter(links > n_total) %>% 
+pull(Network_id)
+
+long_format = long_format %>% 
+filter(Network_id %in% v3)
+
+# Explore how NODFc changes across the Y-X AXES (latitude-Longitude)
+long_format_coords = data %>% 
+mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
+mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
+select(Network_id, Latitude, Longitude, Bioregion) %>% 
+distinct()
+
+# Merge data
+metrics_by_network_coords = left_join(metrics_by_network, long_format_coords)
+
+# Connectance plots
+# 1st Latitude
+p1 = metrics_by_network_coords %>% 
+ggplot(aes(Latitude, Connectance)) +
+geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.15, size=2.5, alpha=1) +
+theme_bw() +
+ylab("Connectance") +
+scale_fill_viridis_d() +
+scale_colour_viridis_d() +
+scale_shape_manual(values = c(
+  "Alpine" = 21,  # Circle
+  "Atlantic" = 22, # Triangle
+  "Boreal" = 23,   # Square
+  "Continental" = 24,  # Diamond
+  "Mediterranean" = 25  # Cross
+))+
+ggtitle("(a)")
+
+# 2nd Plot
+p2 = metrics_by_network_coords %>% 
+ggplot(aes(Latitude, Normalised_nestedness)) +
+geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.15, size=2.5, alpha=1) +
+theme_bw() +
+ylab("NODFc") +
+scale_fill_viridis_d() +
+scale_colour_viridis_d() +
+scale_shape_manual(values = c(
+  "Alpine" = 21,  # Circle
+  "Atlantic" = 22, # Triangle
+  "Boreal" = 23,   # Square
+  "Continental" = 24,  # Diamond
+  "Mediterranean" = 25  # Cross
+  )) +
+ggtitle("(b)")
+
+# Merge and process data
+d = left_join(metrics_by_network, null_metrics_networks)
+d = d %>% 
+mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
+
+p = 0.05 #cutoff probability 95% confidence
+critical_value = qnorm(p/2) #double tail probability divide by 2
+
+d1 = d %>%
+mutate(infra_over_represented = case_when(
+z_score < -abs(critical_value) ~ "Under-represented",
+between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
+    z_score > abs(critical_value) ~ "Over-represented"
+  ))
+
+# 2nd Species (geometric mean)
+species_number = long_format %>% 
+group_by(Network_id) %>% 
+summarise(N_plants = n_distinct(Plant_accepted_name),
+          N_pollinators = n_distinct(Pollinator_accepted_name)) %>% 
+group_by(Network_id) %>% 
+summarise(geometric_mean_spp = geometric.mean(c(N_plants, N_pollinators)))
+
+metrics_by_network_spp_number = left_join(metrics_by_network, species_number)
+
+p3 = metrics_by_network_spp_number %>% 
+ggplot(aes(geometric_mean_spp, Connectance)) +
+geom_point(pch = 21, fill = "azure3", size=2.5, stroke=0.25, alpha=1) +
+theme_bw() +
+xlab("Species")+
+ggtitle("(c)")
+
+# Plot
+p4 = d1 %>% 
+ggplot(aes(x = z_score)) +
+geom_histogram(aes(y = ..density.., fill = infra_over_represented), 
+               bins = 100, alpha = 0.5, color = "black", position = "stack", size=0.1) +
+stat_function(data = d1, fun = function(x) dnorm(x, mean = mean(d1$z_score), sd = sd(d1$z_score)) * 1.7, 
+                n = 1000, inherit.aes = FALSE, color = "gray18", size = 1) +
+theme_bw() +
+coord_cartesian(expand = FALSE) +
+geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red")+
+geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
+ylab("Density") +
+xlab("Z scores (NODF)")  +
+xlim(-6, 9) +
+ylim(0, 0.8) +
+scale_fill_manual(name = "Observed NODF against null",
+                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
+                    labels = c("Less nested", "No difference", "More nested"),
+                    values = c("Under-represented" = "coral2", 
+                               "No statistical difference" = "palegreen3", 
+                               "Over-represented" = "cyan3")) +
+ggtitle("(d)")
+
+# Layout plots
+a = p1 + p2 + plot_layout(guides = 'collect') & theme(legend.position = "right",
+                                                      legend.justification = "left")
+b = (p3 + p4)  & theme(legend.justification = "left")
+a / b  & theme(axis.title =  element_text(size=16, face="bold"),
+               plot.title = element_text(size=22),
+               panel.border = element_rect(colour = "black",
+               fill=NA, size=1),
+               axis.text=element_text(size=12),
+               legend.text = element_text(size=10),
+               legend.key.size = unit(6, "mm"),
+               legend.spacing.y = unit(0.5, "cm"),
+               legend.box.spacing = unit(0, "pt"),
+               legend.title = element_text(face = "bold"))
+
+```
+
+The level of statistical difference is indicated with different colours.
 
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S1}.
-Coverage of the SafeNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
+Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
 
@@
 p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 12, face = "bold"),
+theme(plot.title = element_text(size = 14, face = "bold"),
       axis.title.y = element_text(face = "bold"))

### 2024-04-05 Manuscript/Ms.qmd
minor edits

@@
 In addition, information about habitat type and bioclimatic region are also provided in this file.
-The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the column of "Floral_data_merger.
+The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the column of "Flower_data_merger".
 Only xxxx% of studies contain floral count information.

### 2024-04-08 Manuscript/Ms.qmd
edits

@@
 #Read data
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")%>% 
+data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
 mutate(Network_id = paste0(Study_id, Network_id))  
@@
 
+main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
+
+order_info = data %>% 
+select(Pollinator_order, Pollinator_accepted_name) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+group_by(Pollinator_order) %>% 
+summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
+          Species = n_distinct(Pollinator_accepted_name))
+
+Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
+select(Interactions) %>% pull()
+
 ```
@@
 Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
+Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
 Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
@@
 These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats while also provides other variables of interest that allow to better control for the ecological context and sampling methods.
+EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow to better control for the ecological context and sampling methods.
 In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
@@
 First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions of countries within the database.
-To keep high quality standards that will allow robust future ecological research, we only included: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
+These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions within the database.
+To keep high quality standards that will allow robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
+
+## Dataset description
+
 The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**).
-
-## Dataset description
-
-EuPPollNet includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
-These different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+The different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species.
+However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
 
@@
 (\\textbf{b}) Number of studies by year within the database.
-(\\textbf{c}) Percentage of shared occurrences of pollinators and plants across studies.
-Species on the left are highly shared across studies, while species on the right are found only in a few or a single study.
-As a reference, we have indicated the 50th percentile with a black dashed vertical line."}
+(\\textbf{c}) Number of species and interactions across the 4 main pollinator orders at European level."}
 
@@
 group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.),
-          Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
+summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.),
+          Interactions = length(Pollinator_accepted_name)/ nrow(.))
 
@@
 The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
-The data is organized in three different files named interaction data, flower count data and metadata.
+The data is organized in two different files named interaction data and flower count data.
 The interaction file contains `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
@@
 The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the column of "Flower_data_merger".
-Only xxxx% of studies contain floral count information.
-The metadata file contains [...]
+
+EXPLAIN WHY IS IN A SEPARATE DATASET
+
+Only xxxx% of studies contain floral count information. The metadata file contains [...]
+
+Prepare final metadata
 
@@
 # 3 | RESULTS
+
+The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
 

### 2024-04-08 Scripts/1_Processing_networks/38_Maurer.R

@@ -18,7 +18,7 @@ data = read_csv("Data/1_Raw_data/38_Maurer/Interaction_data.csv")
 #Add flower info cols
 data = data %>% 
 mutate(Flower_data_merger = NA) %>% 
-mutate(Flower_data = "No") 
+mutate(Flower_data = "Yes") 
 
 #There are some sites without coordinates
 #I asked Corina and she provided those by mail

### 2024-04-08 Scripts/1_Processing_networks/41_Knight.R

@@ -102,7 +102,7 @@ Authorship <- data.frame(
                     "Lena Neuenkamp"),
   
   Orcid = c("0000-0003-0318-1567", "0000-0002-2407-9564",
-            "0000-0002-4757-8008", "", "0000-0001-6108-5720"),
+            "0000-0002-4757-8008", "0009-0009-6004-0702", "0000-0001-6108-5720"),
   E_mail = c("tiffany.knight@idiv.de", "elena.motivans@ufz.de",
              "valentin.stefan@idiv.de", "reinart.feldmann@ufz.de",
              "lena-neuenkamp@web.de"))

### 2024-04-08 Scripts/1_Processing_networks/45_Knight.R

@@ -122,7 +122,7 @@ Authorship <- data.frame(
                     "Sebastian Świerszcz", "Elisabeth Kühn"),
   Orcid = c("0000-0003-0318-1567", "0000-0001-8010-4990",
             "0000-0002-2407-9564", "0000-0002-4757-8008",
-            "", "0000-0001-8638-0208", "0000-0003-2035-0035",
+            "0009-0009-6004-0702", "0000-0001-8638-0208", "0000-0003-2035-0035",
             "0000-0001-8894-2462"),
   E_mail = c("tiffany.knight@idiv.de", "demetra.rakosy@gmail.com",
              "elena.motivans@ufz.de", "valentin.stefan@idiv.de",
@@ -131,7 +131,7 @@ Authorship <- data.frame(
 
 #Save data ----
 #Create metadata list
-Knight <- list(InteractionData, FlowerCount, Metadata, Authorship)
+Knight = list(InteractionData, FlowerCount, Metadata, Authorship)
 #Rename list elements
 names(Knight) <- c("InteractionData", "FlowerCount","Metadata", "Authorship")
 #Save data

### 2024-04-08 Scripts/1_Processing_networks/46_Knight.R

@@ -114,7 +114,7 @@ Authorship <- data.frame(
                     "Sebastian Świerszcz", "Elisabeth Kühn"),
   Orcid = c("0000-0003-0318-1567", "0000-0001-8010-4990",
             "0000-0002-2407-9564", "0000-0002-4757-8008",
-            "", "0000-0001-8638-0208", "0000-0003-2035-0035",
+            "0009-0009-6004-0702", "0000-0001-8638-0208", "0000-0003-2035-0035",
             "0000-0001-8894-2462"),
   E_mail = c("tiffany.knight@idiv.de", "demetra.rakosy@gmail.com",
              "elena.motivans@ufz.de", "valentin.stefan@idiv.de",

### 2024-04-09 Manuscript/Ms.qmd
methods ready or more or less

@@
 
+#Check number of species and percentage of interactions of poll. groups
 main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
@@
 select(Interactions) %>% pull()
+
+#Check percentage of studies with flower counts
+floral_counts_percent = data  %>%
+select(Study_id, Flower_data) %>% 
+distinct() %>% 
+group_by(Flower_data) %>% 
+summarise(Percentage = n()/ nrow(.)*100) %>% 
+filter(Flower_data == "Yes")
 
@@
 The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
-The data is organized in two different files named interaction data and flower count data.
-The interaction file contains `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
+The file contains a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
 To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
 To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study.
-In addition, information about habitat type and bioclimatic region are also provided in this file.
-The floral count file contains measurements of floral availability and the unit of measurement, and should be merged with the interaction file by using the column of "Flower_data_merger".
-
-EXPLAIN WHY IS IN A SEPARATE DATASET
-
-Only xxxx% of studies contain floral count information. The metadata file contains [...]
-
-Prepare final metadata
+Information about habitat type and bioclimatic region are also provided in this file.
+In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study.
+The flower count data is provided in a separate file given that one third of studies lack floral measurements and that the units differ greatly across studies.
+Finally, metadata at the study level is provided, including information about the authors, digital object identifier if available, sampling time and taxonomic coverage of the main pollinator groups for each study.
 
@@
 
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.}") %>%
+kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.} Column names and their descriptions within the EuPPollNet database.") %>%
   kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
@@
 ```
+
+## Sampling coverage
+
+The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
+In addition, the accumulation curve of pollinator species with increasing number plant species was also explored.
+The rarefied and extrapolated sampling curves were obtained with the help of the **iNEXT** package [@hsieh2016].
+This was complemented with 100 random accumulation curves obtained programmatically to guide the visualization of the different rarefied curves.
 
@@
 The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if name of the species is uncertain (e.g., species complex or species alike).
+For transparency, we have included in the database the old species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
 
@@
 For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-
-## Sampling coverage
-
-bootstraping, inext package
 
@@
 The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
+
+## Network analyses
+
+To get a general overview of some of the main aspects of these set of plant-pollinator networks, we evaluated how connectance and nestedness change across the latitudinal range of studies and bioclimatic regions.
+In addition, we explored the association between connectance and number of species per network (i.e., geometric mean of plants and pollinators), and if the different networks are more or less nested than expected by chance.
+Importantly, we implemented two different metrics of nestedness given the different analytical objectives.
+Common metrics of nestedness are highly correlated to network size, limiting our ability to compare this metric across networks.
+To avoid this, we implemented a normalised version of nestedness (NODFc) that corrects by connectance and the number of species [@song2017].
+This metric was calculated with the help of the **maxnodf** package [@hoeppke2021], which employs a computational efficient approach to calculate it.
+However, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
+We calculated 100 null models for each network with the help of the _vaznull_ function from the package **Bipartite** [@dormann2008] that implements the null model from @vazquez2005.
+These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
+In other words, interactions are reshuffled with almost equal probabilities to their relative abundances.
+Finally, connectance was estimated for each network by using the function _networklevel_ from **Bipartite**.
 
@@
 Geographic and taxonomic biases!
+Supporting reference:
+https://doi.org/10.1111/ecog.05926
+
+Networks, ref:
+
+https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
 

### 2024-04-10 Manuscript/2_Figure_2.Rmd
results, sampling coverage

@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.", fig.width=10, fig.height=6}
+```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+library(dplyr) #For handling fata 
+library(sf) #For handling coordinates and plotting
+library(scales) #For plotting (decimals on axes)
+library(lubridate) #To operate with dates
+library(tidyr) #Reshape data (wide format)
+library(stringr) #Fix/modify strings
+library(forcats)
+library(readr)
+#Load libraries for plotting
+library(giscoR) #Maps
+library(ggstar) #cool shapes as points
+library(ggplot2) #general plotting
+library(ggtree) #phylo trees
+library(ggpattern) #cool shapes in barplots
+library(patchwork) #plot organisation
+library(cowplot) #plot organisation
+library(viridis) #colours
+library(ggnewscale)
+library(ggtreeExtra)
+library(kableExtra)
 
-#Load libraries
-library(dplyr)
-library(ggplot2)
-library(scales)
-library(patchwork)
+```
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
+Grey solid lines represent 100 randomized accumulation curves, the black solid lines represents the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrates the extrapolated curve for approximately 3000 sampling sites.
+The solid black points indicate the number of species and interactions contained in the database.
+Graph (d) shows the accumulation curve of pollinator species across increasing number of plant species.
+This last graph uses the same color and shape structure as the ones in the top panel.
+Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
+Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.", fig.width=10, fig.height=6, out.width="115%"}
 
@@
                 ylim=c(0,0.7), clip="off") +
-ggtitle("(d)")
+ggtitle("(e)")
 
@@
                 ylim=c(0, 0.7), clip="off") +
-ggtitle("(e)")
+ggtitle("(f)")
 
-plant_poll = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
-rarefaction_curve1 = readRDS("../Data/Working_files/rarefaction_curve_plant_poll.rds")
+#Pollinator by plant species
+plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
+plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
 
-#Plot curves
-pp1 = plant_poll %>% 
-ggplot(aes(x = Plant_sampled, y= Unique_spp_cumulative, group = iteration)) +
-geom_line(color = "gray") +
-geom_smooth(data = rarefaction_curve1, 
-  aes(x = Plant_sampled, y = mean_curve), inherit.aes = FALSE,method = "loess", color = "black", size = 0.65, linetype = "solid", se = FALSE, span = 0.5) +  
-geom_point(data = rarefaction_curve1, x = max(rarefaction_curve1$Plant_sampled), y = max(rarefaction_curve1$mean_curve), inherit.aes = FALSE, size=2) +
+#Create dataset
+d_pp = plant_poll_output$iNextEst$Plant_id
+#Check cols
+d_pp$col = d_pp$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
+           method = as.character(plant_poll_curves$iteration), 
+           qD = plant_poll_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_pp1 = bind_rows(d_pp, a_poll)
+d_pp1_point = d_pp1 %>%  filter(method == "observed")
+d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
+
+#Plot
+#Order levels
+d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col)) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_pp1_point, col = "black") + 
 ylab("Pollinator species") +
@@
 coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(f)") +
-scale_x_continuous(breaks = seq(0, 1400, by = 400), limits=c(0,1400)) +
-ylim(0,2500)
+ggtitle("(d)") +
+scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
+scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
+theme(legend.position = "none")
 
 #Plot all
-a = p1 + p2 + p3 + plot_layout(guides = "collect") & theme(legend.title=element_blank(),
+p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
       axis.title = element_text(size=12, face = "bold"),
@@
       plot.title = element_text(size=18))
-b = o1 + o2 + pp1 & 
-  theme(axis.title = element_text(size=12, face = "bold"),
-        axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
 
-a/b
+#a/b
 

### 2024-04-10 Manuscript/Ms.qmd
results, sampling coverage

@@
 This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
+Furthermore, this work provides an in-depth assessment of the taxonomic and sampling coverage of the database at the European level, complemented by analyses of key structural properties in plant-pollinator networks.
 We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.
@@
 \small
-plant-pollinator interactions, flowering plants, Angiosperms, pollinators, Hymenoptera, Diptera, Lepidoptera, Coleoptera, Europe
+plant-pollinator networks, plant-pollinator interactions, flowering plants, Angiosperms, pollinators, nestedness, connectance
 \endgroup
@@
 Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes at large scales.
+Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
 
@@
 These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses.
+In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions.
+Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations.
 EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow to better control for the ecological context and sampling methods.
@@
 
-# Data structure
+## Data structure
 
@@
 
-The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figure 1b**).
-
-## Taxonomy and taxonomic coverage
-
-Provide number of species per taxonomic group and number of interactions
-Something always of interest: provide % of interactions of _Apis mellifera_?
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
+plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
+int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
+plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
+
+#Coverage
+poll_coverage = poll_output$DataInfo$SC *100
+plant_coverage = plant_output$DataInfo$SC *100
+int_coverage = int_output$DataInfo$SC *100
+plant_poll_coverage = plant_poll_output$DataInfo$SC *100
+
+#Increase doubling effort
+##Pollinators
+poll_increase = poll_output$iNextEst$Network_id
+poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
+poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
+poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
+##Plants
+plant_increase = plant_output$iNextEst$Network_id
+plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
+plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
+plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
+##Interactions
+int_increase = int_output$iNextEst$Network_id
+int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
+int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
+int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
+##Pollinator by plant
+plant_poll_increase = plant_poll_output$iNextEst$Plant_id
+plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
+plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
+plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
+```
+
+## Sampling coverage
+
+The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
+This suggest that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 2a-2b**).
+The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
+However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 2c**).
+Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
+The predicted increase of recorded pollinator species by sampling the double of plant species would be `r format(round(plant_poll_increase3, 2), nsmall=2)`%.
+
+Maybe fix last graph? And add it 1st of the last panel (**Figure 2d**)
+
+The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figures 2e-2f**).
+For instance, xxxxxxxxx.
+
+The estimated sampling coverage when we consider plants,
 
@@
 The solid black points indicate the number of species and interactions contained in the database.
-Graphs (d-e) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
-Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
-Graph (f) shows the accumulation curve of pollinator species across increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.", fig.width=10, fig.height=6, out.width="115%"}
+Graph (d) shows the accumulation curve of pollinator species across increasing number of plant species.
+This last graph uses the same color and shape structure as the ones in the top panel.
+Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
+Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.", fig.width=10, fig.height=6, out.width="115%"}
 
@@
                 ylim=c(0, 0.7), clip="off") +
-ggtitle("(e)")
-
-plant_poll = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
-rarefaction_curve1 = readRDS("../Data/Working_files/rarefaction_curve_plant_poll.rds")
-
-#Plot curves
-pp1 = plant_poll %>% 
-ggplot(aes(x = Plant_sampled, y= Unique_spp_cumulative, group = iteration)) +
-geom_line(color = "gray") +
-geom_smooth(data = rarefaction_curve1, 
-  aes(x = Plant_sampled, y = mean_curve), inherit.aes = FALSE,method = "loess", color = "black", size = 0.65, linetype = "solid", se = FALSE, span = 0.5) +  
-geom_point(data = rarefaction_curve1, x = max(rarefaction_curve1$Plant_sampled), y = max(rarefaction_curve1$mean_curve), inherit.aes = FALSE, size=2) +
+ggtitle("(f)")
+
+#Pollinator by plant species
+plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
+plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
+
+#Create dataset
+d_pp = plant_poll_output$iNextEst$Plant_id
+#Check cols
+d_pp$col = d_pp$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
+           method = as.character(plant_poll_curves$iteration), 
+           qD = plant_poll_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_pp1 = bind_rows(d_pp, a_poll)
+d_pp1_point = d_pp1 %>%  filter(method == "observed")
+d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
+
+#Plot
+#Order levels
+d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col)) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_pp1_point, col = "black") + 
 ylab("Pollinator species") +
@@
       plot.title = element_text(size=18))
-b = o1 + o2 + pp1 & 
-  theme(axis.title = element_text(size=12, face = "bold"),
-        axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-a/b
+
+#a/b
 
@@
 
-phylogenetic and geographical biases
+\newpage
+
+## Taxonomic coverage
+
+The database contains xxxx species of pollinators and xxxx species of plants. Most species of pollinators are Hymenoptera species (%), and most of these species are bees (%).
+
+Provide number of species per taxonomic group and number of interactions
+Something always of interest: provide % of interactions of _Apis mellifera_?
+
+## Habitat type and bioclimatic region
+
+## Network properties
 
@@
 
-## Habitat type and bioclimatic region
-
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
@@
 
-The level of statistical difference is indicated with different colours.
-
 \newpage
@@
 # 3 | DISCUSSION
+
+Although Europe...
 
@@
 summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
+mutate(Group = "EuPPollNet")
 #Species per family within Europe
@@
 ylab("Number of species") +
-scale_fill_manual(values = c(viridis(2)))+
+scale_fill_manual(values = c("tan2", "lightgrey"))+
 ggtitle("Bee family coverage")
@@
 summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "SafeNet")
+mutate(Group = "EuPPollNet")
 #Get spp per subfamily at european level
@@
       panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c(viridis(2)))+
+scale_fill_manual(values = c("tan2", "lightgrey"))+
 ggtitle("Syrphid subfamily coverage")
@@
       panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c(viridis(2)))+
+scale_fill_manual(values = c("tan2", "lightgrey"))+
 ggtitle("Butterfly family coverage")
@@
 p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 14, face = "bold"),
+theme(plot.title = element_text(size = 13, face = "bold"),
       axis.title.y = element_text(face = "bold"))

### 2024-04-11 Manuscript/Ms.qmd
some text of taxonomic coverage... Still thinking how to best explain it

@@
 The different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant, from `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant.
 Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species.
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Sampling coverage outputs
 poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
@@
 plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
+
+#Prepare results from common and rare poll across sampling sites
+#Pollinators
+poll_spread = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
+group_by(Network_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+ungroup() %>% 
+count(Pollinator_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
+#Check number of studies under 1%
+poll_total = nrow(poll_spread)
+polls_under_one_percent = poll_spread %>% 
+filter(Percent_total<1) %>% 
+nrow()
+polls_under_one_percent = polls_under_one_percent/poll_total * 100
+#Find most common pollinator
+most_common_poll = poll_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Pollinator_accepted_name)
+#Find most common poll percent
+most_common_poll_percent = poll_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Percent_total)
+#Plants
+plant_spread = data %>% 
+select(Plant_rank, Network_id, Plant_accepted_name) %>% 
+group_by(Network_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+ungroup() %>% 
+count(Plant_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
+#Check number of studies under 1%
+plant_total = nrow(plant_spread)
+plants_under_one_percent = plant_spread %>% 
+filter(Percent_total<1) %>% 
+nrow()
+plants_under_one_percent = plants_under_one_percent/plant_total * 100  
+#Find most common plant
+most_common_plant = plant_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Plant_accepted_name)
+#Find most common plant percent
+most_common_plant_percent = plant_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Percent_total)
 ```
@@
 Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted increase of recorded pollinator species by sampling the double of plant species would be `r format(round(plant_poll_increase3, 2), nsmall=2)`%.
-
-Maybe fix last graph? And add it 1st of the last panel (**Figure 2d**)
-
-The majority of plant and pollinator species tend to be regionally specific, with only a minor portion of them being shared across a broad range of studies (**Figures 2e-2f**).
-For instance, xxxxxxxxx.
-
-The estimated sampling coverage when we consider plants,
+The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 2d**).
+In addition, when we explore how common or rare are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 2e-2f**).
+For instance, most plants (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinators (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) are exclusively found in less than 1% of sampling sites; and the most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are only found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
 
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+#Taxonomic coverage outputs
+#Find number of plant and pollinator species
+##Pollinators
+poll_spp = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+nrow(.)
+##Plants
+plant_spp = data %>% 
+select(Plant_rank, Plant_accepted_name) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+nrow(.)
+
+#Overall plant coverage
+plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
+
+#Average plant family coverage
+d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
+mean_plant_fam_cov = d1 %>% 
+ungroup() %>% 
+filter(fam_group =="B") %>% 
+summarise(mean_plant_fam_cov = mean(Percent)) %>% 
+pull()
+
+```
+
 ## Taxonomic coverage
 
-The database contains xxxx species of pollinators and xxxx species of plants. Most species of pollinators are Hymenoptera species (%), and most of these species are bees (%).
+The database contains a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
+The database coverage of flowering plants occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
+There are x% of families without recorded interactions and the absence of information seems to not follow strongly phylogenetic constraints (OUTPUT lambda)
+
+The estimated sampling coverage of plant species at European level is x% (**Figure 4**).
+The three families with higher number of interactions within the database are xxxxx (%), xxxxxx (%) and xxxxx (%).
+Despite these species have a high number of interactions within the database, the number of species recorded within the database in comparison with the total number of species at European level reveals a low coverage of interactions for most of the species of these families.
+
+The most common order of plants is Asterales with Asteraceae being the family with highest number of species within the database.
+
+The database coverage of plants at European level is x% and the pollinator coverage when considering bees x%, syrphids x% and butterflies x%.
+
+Most species of pollinators are Hymenoptera species (%), and most of these species are bees (%).
+
+**Figure 3**
+**Figure 4**
+**Figure S1**
 

### 2024-04-12 Manuscript/Ms.qmd
edits

@@
 library(readr)
+library(phytools)
 #Load libraries for plotting
@@
 #Average plant family coverage
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
+d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData") %>% 
+filter(fam_group == "B")
+#Average coverage per family
 mean_plant_fam_cov = d1 %>% 
 ungroup() %>% 
-filter(fam_group =="B") %>% 
 summarise(mean_plant_fam_cov = mean(Percent)) %>% 
@@
 
+#Percent of families with interaction records
+all_fam = nrow(d1)
+fam_with_data = d1 %>% 
+filter(Percent > 0) %>% 
+nrow()
+percent_plant_fam_with_data = fam_with_data/all_fam * 100
+
+#Calculate phylogenetic signal of presence-absence information
+#Read data and plot it
+tree = readRDS("../Data/Manuscript_info/plant_phylo.RData")
+#Create binary trait to calculate phylo
+phylo_trait = d1 %>% 
+mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
+#select(label, categorical) %>% 
+ungroup() 
+#Check phylogenetic signal
+d1 = d1 %>% arrange(factor(label, levels = tree@phylo$tip.label))
+phylo_signal_plant = phylosig(tree@phylo, phylo_trait$categorical, method="lambda", test=TRUE)
+lambda_phylosig_plants = phylo_signal_plant$lambda
+p_phylosig_plants = phylo_signal_plant$P
+
 ```
@@
 The database coverage of flowering plants occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-There are x% of families without recorded interactions and the absence of information seems to not follow strongly phylogenetic constraints (OUTPUT lambda)
+Most plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%) and the presence or absence of interaction data seems not to follow a phylogenetic pattern (RESULT).
+
+`r lambda_phylosig_plants`
+`r p_phylosig_plants`
 

### 2024-04-14 Manuscript/Ms.qmd
Edits

@@
 
+#Check percentage of apis mellifera records
+all_records = data %>% nrow(.) 
+apis_records = data %>% 
+filter(Pollinator_accepted_name == "Apis mellifera") %>% 
+nrow(.)
+apis_percent = apis_records/all_records * 100
+#Check average pecentage of interactions of apis by net
+a = data %>% 
+group_by(Network_id) %>% 
+summarise(n_interactions = length(Pollinator_accepted_name))
+b = data %>% 
+group_by(Network_id) %>% 
+filter(Pollinator_accepted_name == "Apis mellifera") %>% 
+summarise(n_interactions_apis = length(Pollinator_accepted_name))
+c = left_join(a,b)
+apis_average_network = c %>% 
+mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
+ungroup() %>% 
+summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
+pull()
+
 ```
@@
 However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
+Notably,  _Apis mellifera_ represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
 
@@
 d1 = d1 %>% arrange(factor(label, levels = tree@phylo$tip.label))
-phylo_signal_plant = phylosig(tree@phylo, phylo_trait$categorical, method="lambda", test=TRUE)
+
+v = phylo_trait$categorical
+names(v) = tree@phylo$tip.label
+
+phylo_signal_plant = phylosig(tree@phylo, v, method="lambda", test=TRUE)
 lambda_phylosig_plants = phylo_signal_plant$lambda
@@
 The database contains a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database coverage of flowering plants occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Most plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%) and the presence or absence of interaction data seems not to follow a phylogenetic pattern (RESULT).
-
-`r lambda_phylosig_plants`
-`r p_phylosig_plants`
-
-The estimated sampling coverage of plant species at European level is x% (**Figure 4**).
-The three families with higher number of interactions within the database are xxxxx (%), xxxxxx (%) and xxxxx (%).
-Despite these species have a high number of interactions within the database, the number of species recorded within the database in comparison with the total number of species at European level reveals a low coverage of interactions for most of the species of these families.
-
-The most common order of plants is Asterales with Asteraceae being the family with highest number of species within the database.
-
-The database coverage of plants at European level is x% and the pollinator coverage when considering bees x%, syrphids x% and butterflies x%.
-
-Most species of pollinators are Hymenoptera species (%), and most of these species are bees (%).
+The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
+Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
+
+Pollinators:
 
 **Figure 3**
-**Figure 4**
+
 **Figure S1**
@@
 Provide number of species per taxonomic group and number of interactions
-Something always of interest: provide % of interactions of _Apis mellifera_?
 

### 2024-04-15 Manuscript/Ms.qmd
taxonomic coverage text

@@
 
+#Load pollinator coverage
+bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
+syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
+lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
+#Check percentage of interactions from bees with and without apis
+#Check percentage of bees in the database
+bee_fam = c("Apidae", "Megachilidae", "Halictidae",
+            "Andrenidae", "Colletidae", "Melittidae")
+all_rows = data %>% nrow(.)
+#
+bee_rows = data %>%
+filter(Pollinator_family %in% bee_fam) %>% 
+nrow(.)
+#Bee interactions
+bee_interactions_percent = bee_rows/all_rows *100
+#Eclude now apis
+bee_rows_non_apis = data %>%
+filter(Pollinator_family %in% bee_fam) %>% 
+filter(Pollinator_accepted_name == "Apis mellifera") %>% 
+nrow(.)
+#
+bee_interactions_non_apis_percent = bee_rows_non_apis/bee_rows *100
+
+#Coverage across bee genera
+d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
+all_rows_bees = d1 %>%  
+filter(fam_group == "a") %>% 
+nrow(.)
+#
+data_bees = d1 %>%  
+filter(fam_group == "a") %>% 
+mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
+filter(Percent > 0) %>% 
+nrow(.)
+#
+bee_genera_percent_coverage = data_bees/all_rows_bees
+#Check average coverage across bee genera
+average_coverage_bee_genera = d1 %>% 
+filter(fam_group == "a") %>% 
+mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
+ungroup() %>% 
+summarise(average_coverage_bee_genera = mean(Percent)) %>% 
+pull()
+
+#Calculate phylogenetic signal of presence-absence information
+##Read data and plot it
+tree = readRDS("../Data/Manuscript_info/bee.tree100.rds")
+##Create binary trait to calculate phylo
+phylo_trait = d1 %>% 
+filter(fam_group == "a") %>% 
+mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
+mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
+ungroup() 
+##Conduct left join to tibble of tip labels to organise order
+v = tibble(Genus1 = tree$tip.label) %>% 
+mutate(Genus1 = str_to_title(word(Genus1,1,sep = "_")))
+#
+v1 = left_join(v, phylo_trait)
+#
+v2 = v1$categorical
+names(v2) = tree$tip.label
+#
+phylo_signal_bees = phylosig(tree, v2, method="lambda", test=TRUE)
+lambda_phylosig_bees = phylo_signal_bees$lambda
+p_phylosig_bees = phylo_signal_bees$P
+
 ```
@@
 Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-
-Pollinators:
+The coverage of the main taxonomic groups of pollinators at European level is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
+Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions of the database, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
+
+`r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records within the database and the average coverage at the genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`%.
+Notably, the presence or absence of interactions records seems or not seems to follow any sort of phylogenetic signal (($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`)).
 
 **Figure 3**
-
-**Figure S1**
-
-Provide number of species per taxonomic group and number of interactions
 

### 2024-04-16 Manuscript/Ms.qmd
edits

@@
     Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "SafeNet_habitat" ~ "Type of habitat homogenized across studies",
+    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
     Variable == "Bioregion" ~ "European biogiographical regions",
@@
 #
-bee_genera_percent_coverage = data_bees/all_rows_bees
+bee_genera_percent_coverage = data_bees/all_rows_bees * 100
 #Check average coverage across bee genera
@@
 
-The database contains a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
+The database contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+The coverage of the main pollinators groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
+Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions of the database, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
+Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 3**).
+In addition, the presence or absence of interactions records for bees seems not to follow any sort of phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
 The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-The coverage of the main taxonomic groups of pollinators at European level is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions of the database, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-
-`r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records within the database and the average coverage at the genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`%.
-Notably, the presence or absence of interactions records seems or not seems to follow any sort of phylogenetic signal (($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`)).
-
-**Figure 3**
-
-## Habitat type and bioclimatic region
-
-## Network properties
+Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
 
@@
 
+## Habitat type and bioclimatic region
+
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
@@
 habitat = data %>% 
-select(Study_id, SafeNet_habitat) %>% 
+select(Study_id, EuPPollNet_habitat) %>% 
 distinct() %>%  
-group_by(SafeNet_habitat) %>% 
+group_by(EuPPollNet_habitat) %>% 
 summarise(Habitat_count = n()) 
@@
 pollinator_percent = data %>%
-group_by(SafeNet_habitat, Pollinator_order) %>%
+group_by(EuPPollNet_habitat, Pollinator_order) %>%
 filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
@@
 summarise(count = n()) %>%
-group_by(SafeNet_habitat) %>%
+group_by(EuPPollNet_habitat) %>%
 mutate(Proportion = count/sum(count)) %>%
@@
 mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat) %>% 
+rename(Habitat = EuPPollNet_habitat) %>% 
 rename(Count = Habitat_count)
@@
 mutate(Group = "Habitat") %>% 
-rename(Habitat = SafeNet_habitat)
+rename(Habitat = EuPPollNet_habitat)
 #Bind rows
@@
 ```
+
+\newpage
+
+## Network properties
 

### 2024-04-16 Scripts/Figures/Map_study_locations.Rmd
edits

--- 2023-10-26 Scripts/1_Visualisation/Map_study_locations.Rmd
@@
 "coral4", "plum1", "yellow",
-"white")
+"brown3", "orange3", "red4", "darkviolet",
+"darkgreen")
 #To generate more random colors 

### 2024-04-16 Manuscript/Ms.qmd
habitat type and bioclimatic region -results-

@@
 
+#Check number of studies by habitat
+studies_by_habitat = data %>% 
+select(Study_id, EuPPollNet_habitat) %>% 
+distinct() %>%  
+group_by(EuPPollNet_habitat) %>% 
+summarise(Habitat_count = n()) %>% 
+slice_max(Habitat_count, n = 3) %>% 
+mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
+#Check number of sampling sites by habitat
+networks_by_habitat = data %>% 
+select(Network_id, EuPPollNet_habitat) %>% 
+distinct() %>%  
+group_by(EuPPollNet_habitat) %>% 
+summarise(Habitat_count = n()) %>% 
+slice_max(Habitat_count, n = 3) %>% 
+mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
+#Check number of studies by bioregion
+studies_by_bioregion = data %>% 
+select(Study_id, Bioregion) %>% 
+distinct() %>%  
+group_by(Bioregion) %>% 
+summarise(Bioregion_count = n()) %>% 
+slice_max(Bioregion_count, n = 3) %>% 
+mutate(Bioregion = tolower(Bioregion))
+#Check number of sampling sites by bioregion
+networks_by_bioregion = data %>% 
+select(Network_id, Bioregion) %>% 
+distinct() %>%  
+group_by(Bioregion) %>% 
+summarise(Bioregion_count = n()) %>% 
+slice_max(Bioregion_count, n = 3) %>% 
+mutate(Bioregion = tolower(Bioregion))
+
 ```
@@
 ## Habitat type and bioclimatic region
+
+The proportion of the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
+As expected, Hymenoptera was the main taxonomic order on the majority of habitats, exceeded only by Diptera on the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
+Overall, the proportion of Lepidoptera and Coleoptera was low across all habitats but Coleoptera showed a notably increase in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
+Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
+In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
+Again, Lepidoptera shows low proportions across all bioclimatic regions and Coleptera seems to be only relevant in Mediterranean regions at European level.
+Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
+The habitats sampled by a higher number of studies in the database are `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
+However, the habitats that contain a higher number of sampling sites are `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
+The bioclimatic regions with a higher number of studies are `r studies_by_bioregion$Bioregion[[1]]` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r studies_by_bioregion$Bioregion[[2]]` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r studies_by_bioregion$Bioregion[[3]]` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r networks_by_bioregion$Bioregion[[1]]` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r networks_by_bioregion$Bioregion[[2]]` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r networks_by_bioregion$Bioregion[[3]]` (`r networks_by_bioregion$Bioregion_count[[3]]`).
 
@@
 ```
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%", fig.cap= "\\textbf{Figure S2}.
+Barplot indicating the number of sampling sites by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
+
+#Check number of studies by habitat
+habitat = data %>% 
+select(Network_id, EuPPollNet_habitat) %>% 
+distinct() %>%  
+group_by(EuPPollNet_habitat) %>% 
+summarise(Habitat_count = n()) %>% 
+mutate(Group = "Habitat") %>% 
+rename(Habitat_bioregion = EuPPollNet_habitat) %>% 
+rename(Count = Habitat_count)
+
+#Check number of studies by bioregion
+bioregion = data %>% 
+select(Network_id, Bioregion) %>% 
+distinct() %>%  
+group_by(Bioregion) %>% 
+summarise(Bioregion_count = n()) %>% 
+mutate(Group = "Bioregion") %>% 
+rename(Habitat_bioregion = Bioregion) %>% 
+rename(Count = Bioregion_count)
+
+c = bind_rows(habitat, bioregion)
+
+ggplot(c, aes(Habitat_bioregion, Count)) +
+geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
+theme_bw()+
+coord_flip(expand = F) +
+ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")),
+        space = "free", scales = "free_y") +
+ylab("Sampling sites") +
+xlab(NULL) +
+theme(plot.margin=unit(c(0,0,0,0), "mm"),
+      axis.ticks.x = element_line(color="black"),
+      strip.text.x = element_text(size=12),
+      strip.background = element_rect(size=0.5, linetype="solid"),
+  axis.title = element_text(face = "bold", size = 12),
+axis.text = element_text(size=10))
+
+```

### 2024-04-17 Manuscript/Ms.qmd
results section

@@
 library(kableExtra)
+
+#In case of conflicts select the main dplyr functions
+library(conflicted)
+conflicts_prefer(dplyr::filter)
+conflicts_prefer(dplyr::select)
 
@@
 For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
+Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated with the the phylogenetic signal with the help of the function _phylosig_ from the **phytools** package.
 
@@
 To get a general overview of some of the main aspects of these set of plant-pollinator networks, we evaluated how connectance and nestedness change across the latitudinal range of studies and bioclimatic regions.
-In addition, we explored the association between connectance and number of species per network (i.e., geometric mean of plants and pollinators), and if the different networks are more or less nested than expected by chance.
+We implemented linear models, with the network metrics as the response variable, and both latitude and bioclimatic region as predictors in two separate models.
+Finally, with the help of a Tukey test from the package _emmeans_, we evaluated if nestedness and connectance differ statistically across bioclimatic regions.
+In addition, we explored the association between connectance and nestedness with the number of species per network (i.e., geometric mean of plants and pollinators) by conducting Kendall rank correlation coefficient.
+Finally, we evaluated if the different networks are more or less nested than expected by chance.
 Importantly, we implemented two different metrics of nestedness given the different analytical objectives.
@@
 In other words, interactions are reshuffled with almost equal probabilities to their relative abundances.
-Finally, connectance was estimated for each network by using the function _networklevel_ from **Bipartite**.
+Connectance was estimated for each network by using the function _networklevel_ from **Bipartite**.
 
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Outputs calculated in script 7_3
+#Load lm output connectance ~ latitude
+connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
+connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
+#Load lm output nestedness ~ latitude
+normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
+normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
+#Load values of connectance
+min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
+max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
+mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
+#Load values of nestedness
+min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
+max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
+mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
+#Load kendall corr from connectance and spp
+connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
+connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
+#Load kendall corr from nestedness and spp
+normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
+normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
+#Load z values
+#Save data
+z_percent = readRDS("../Data/Manuscript_info/z_percent.rds")
+no_diff_z_percent = z_percent %>% 
+filter(infra_over_represented == "No statistical difference") %>% 
+pull(percent)
+over_z_percent = z_percent %>% 
+filter(infra_over_represented == "Over-represented") %>% 
+pull(percent)
+under_z_percent = z_percent %>% 
+filter(infra_over_represented == "Under-represented") %>% 
+pull(percent)
+
+```
+
 ## Network properties
+
+Connectance and nestedness did not change across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
+However, differences were found for connectance but not nestedness across bioclimatic regions.
+Specifically, the alpine region differed statistically in connectance to the other bioclimatic regions (Tukey test _P_ < 0.01).
+In addition, the atlantic and continental bioclimatic regions also showed statistical meaningful differences in connectance (Tukey test _P_ < 0.01).
+Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed an negative exponential relationship with the geometric mean of plant and pollinator species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6c**).
+Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
+Finally, we found that `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, `r round(no_diff_z_percent, 2)`% showed no statistical difference, and `r round(under_z_percent, 2)`% were less nested than null expectations (**Figure 6d**).
 
@@
 p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 13, face = "bold"),
-      axis.title.y = element_text(face = "bold"))
+theme(plot.title = element_text(size = 14, face = "bold"),
+      axis.title = element_text(face = "bold", size = 14),
+      axis.text = element_text(face = "bold", size = 14))
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="100%", fig.cap= "\\textbf{Figure S2}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S2}.
 Barplot indicating the number of sampling sites by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
@@
       axis.ticks.x = element_line(color="black"),
-      strip.text.x = element_text(size=12),
+      strip.text.x = element_text(size=14, face = "bold"),
       strip.background = element_rect(size=0.5, linetype="solid"),
@@
 ```
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="80%", fig.cap= "\\textbf{Figure S3}.
+Association between nestedness (NODFc) and the geometric mean of plan and pollinator species per network", fig.width=3, fig.height=3}
+
+metrics_by_network_spp_number %>% 
+ggplot(aes(geometric_mean_spp, Normalised_nestedness)) +
+geom_point(pch = 21, fill = "azure3", size=1.5, stroke=0.25, alpha=1) +
+theme_bw() +
+xlab("Species") +
+ylab("Normalised nestedness (NODFc)") +
+theme(
+    axis.title = element_text(face = "bold", size = 10),
+    axis.text = element_text(size = 8))
+
+```

### 2024-04-17 Scripts/7_3_Network_indices_statistics.R

@@ -1,0 +1,226 @@
+# Evaluate statistical differences between network indices
+
+#Load libraries
+library(ggplot2)
+library(emmeans)
+library(multcomp)
+library(broom)
+library(conflicted)
+conflict_prefer("select", "dplyr")
+# Load data
+data = readRDS("Data/3_Final_data/Interaction_data.rds")
+metrics_by_network = readRDS("Data/Working_files/metrics_by_network.rds")
+null_networks = readRDS("Data/Working_files/null_networks.rds")
+null_metrics_networks = readRDS("Data/Working_files/null_metrics_networks.rds")
+# Prepare data
+data = data %>%
+dplyr::mutate(Year = lubridate::year(Date), 
+              Month = lubridate::month(Date), 
+              Day = lubridate::day(Date))
+
+# Subset data for studies with only bumblebees
+v = c("13_Karise", "22_Kallnik")
+subsetting = unique(data$Study_id)
+
+long_format = data %>% 
+mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
+mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
+filter(!Study_id %in% v) %>% 
+select(Network_id, Plant_accepted_name, Pollinator_accepted_name)
+
+# Filter networks larger than 4 plants and 4 pollinator species
+v2 = long_format %>% 
+group_by(Network_id) %>% 
+summarise(Poll_spp = n_distinct(Plant_accepted_name),
+          Plant_spp = n_distinct(Pollinator_accepted_name)) %>% 
+ungroup() %>% 
+filter(Poll_spp >= 4 & Plant_spp >= 4) %>% 
+pull(Network_id)
+
+long_format = long_format %>% 
+filter(Network_id %in% v2)
+
+# Filter networks with more links than species
+links = long_format %>% 
+group_by(Network_id, Plant_accepted_name, Pollinator_accepted_name) %>%
+distinct() %>% 
+group_by(Network_id) %>% 
+summarise(links = length(Network_id))
+
+spp = long_format %>% 
+group_by(Network_id) %>% 
+summarise(n_plants = n_distinct(Plant_accepted_name),
+          n_polls = n_distinct(Pollinator_accepted_name)) %>% 
+mutate(n_total = n_plants + n_polls)
+
+vector_to_include = left_join(links, spp)
+
+v3 = vector_to_include %>% 
[…]

### 2024-04-17 Manuscript/EuPPollNet_Ms.qmd
edits

--- 2024-04-17 Manuscript/Ms.qmd
@@
 For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
+Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because of undersampling [@olesen2011; @chacoff2012].
+Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need to keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
 
@@
 Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations.
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow to better control for the ecological context and sampling methods.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
+EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods.
+In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time.
 We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
@@
 Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represents the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrates the extrapolated curve for approximately 3000 sampling sites.
+Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites.
 The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across increasing number of plant species.
+Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
 This last graph uses the same color and shape structure as the ones in the top panel.
@@
 In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
-Again, Lepidoptera shows low proportions across all bioclimatic regions and Coleptera seems to be only relevant in Mediterranean regions at European level.
+Again, Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera seems to be only relevant in Mediterranean regions at European level.
 Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
@@
 The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can countribute to more than one habitat or bioclimatic region.
+Note that a single study can contribute to more than one habitat or bioclimatic region.
 Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
@@
 In addition, the bioclimatic region where each network is located is indicated with points of different shapes and colours.
-Graph (c) shows the the association between network connectances and the geometric mean of plant and pollinator species per network.
+Graph (c) shows the association between network connectance and the geometric mean of plant and pollinator species per network.
 Graph (d) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
@@
       axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 14))
+      axis.text = element_text(face = "bold", size = 12))
 

### 2024-04-17 Manuscript/EuPPollNet_word.qmd
edits

--- 2024-04-17 Manuscript/Ms.qmd
@@
 format:
-  pdf:
+  docx:
     include-in-header: 
@@
 latex-tinytex: false
+prefer-html: true
 link-citations: TRUE
@@
       axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 14))
+      axis.text = element_text(face = "bold", size = 1))
 

### 2024-04-17 Manuscript/Figures_and_Tables/2_Figure_2.Rmd
edits

--- 2024-04-10 Manuscript/2_Figure_2.Rmd
@@
 #Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
+poll_output = readRDS("../../Data/Working_files/pollinator_sampling_coverage.rds")
+pollinators_curves = readRDS("../../Data/Working_files/collectors_curves_pollinators.rds")
 #Create dataset
@@
 #Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
+plant_output = readRDS("../../Data/Working_files/plant_sampling_coverage.rds")
+plant_curves = readRDS("../../Data/Working_files/collectors_curves_plants.rds")
 #Create dataset
@@
 #Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
+int_output = readRDS("../../Data/Working_files/interactions_sampling_coverage.rds")
+interactions_curves = readRDS("../../Data/Working_files/collectors_curves_interactions.rds")
 #Create dataset
@@
 #Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
+data = readRDS("../../Data/Working_files/Interactions_uncounted.rds")%>% 
 mutate(Network_id = paste0(Study_id, Network_id))  
@@
 #Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
+plant_poll_curves = readRDS("../../Data/Working_files/collectors_curves_plant_ploll.rds")
+plant_poll_output = readRDS("../../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
 

### 2024-04-17 Manuscript/Figures_and_Tables/3_Figure_3.Rmd
edits

--- 2024-03-27 Manuscript/3_Figure_3.Rmd
@@
 #Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
+tree_family_interactions= readRDS("../../Data/Manuscript_info/bee_phylo.RData")
+d1 = readRDS("../../Data/Manuscript_info/bee_data.RData")
+n = readRDS("../../Data/Manuscript_info/node_data.RData")
 

### 2024-04-17 Manuscript/Figures_and_Tables/4_Figure_4.Rmd
edits

--- 2024-03-27 Manuscript/4_Figure_4.Rmd
@@
 #Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
+tree_family_interactions = readRDS("../../Data/Manuscript_info/plant_phylo.RData")
+d1 = readRDS("../../Data/Manuscript_info/plant_coverage_phylo_RData")
 

### 2024-04-17 Manuscript/Figures_and_Tables/6_Figure_6.Rmd
edits

--- 2024-04-05 Manuscript/6_Figure_6.Rmd
@@
 # Load data
-data = readRDS("../Data/3_Final_data/Interaction_data.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-null_networks = readRDS("../Data/Working_files/null_networks.rds")
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks.rds")
+data = readRDS("../../Data/3_Final_data/Interaction_data.rds")
+metrics_by_network = readRDS("../../Data/Working_files/metrics_by_network.rds")
+null_networks = readRDS("../../Data/Working_files/null_networks.rds")
+null_metrics_networks = readRDS("../../Data/Working_files/null_metrics_networks.rds")
 

### 2024-04-17 Manuscript/Figures_and_Tables/Figure_S1.Rmd
edits

--- 2024-03-05 Manuscript/Figure_S1.Rmd
@@
 #Safenet interactions
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
+data = readRDS("../../Data/3_Final_data/Interactions_uncounted.rds")
 #Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
+master_list = read_csv("../../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
 #Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
+master_list1 <- read.delim(file = "../../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
                     header = T)

### 2024-05-08 Manuscript/EuPPollNet_Ms.qmd
Adding Nacho's comments to main text, some few things left. I left them in things to do in the discussion

@@
 
-**Motivation:** Pollinators play a crucial role in maintaining Earth's terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-However, their diversity and role as pollinators are increasingly compromised by rapid human-induced environmental changes.
-One of the major challenges for pollinator conservation, is the lack of robust generalisable data across space and time to comprehend the conservation status and population trends among different pollinator species.
-Here, we present the EuPPollNet database, a fully reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
-This database offers an open workflow that allows researchers to track decisions and edit them according to their preferences, while also providing other ecological variables of interest.
+**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
+Indeed, the intricate network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
+However, rapid human-induced environmental changes are compromising its long-term persistence.
+One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
+Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
+This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences, while also providing other ecological variables of interest.
 Furthermore, this work provides an in-depth assessment of the taxonomic and sampling coverage of the database at the European level, complemented by analyses of key structural properties in plant-pollinator networks.
-We hope this database can help researchers to: 1) identify taxonomic, ecological and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change to guide future conservation planning for both plant and pollinator species.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region and further taxonomic rank information for both plant and pollinator species are also provided.
+We hope this database can help researchers to: 1) identify taxonomic, ecological, and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change on plant-pollinator networks to guide future conservation planning for both plant and pollinator species.
+
+**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_sites, big.mark=",")` distinct locations (i.e., plant-pollinator networks), which belong to `r study_number` different studies distributed across `r country_number` European countries.
+In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided.
 
@@
 **Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants.
-The main focus of this database are insect pollinators and accounts for `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
+The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
 
@@
 The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016; @eichenberg2021].
+Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
+The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
 Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
@@
 The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023].
+Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
+In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023].
 Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
 For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because of undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need to keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
-
-Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
+Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
+Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
+
+Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level.
 The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
@@
 The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
-In addition, the accumulation curve of pollinator species with increasing number plant species was also explored.
+In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [@kleijn2015].
+To that end, we also calculated the accumulation curve of pollinator species with increasing number plant species.
 The rarefied and extrapolated sampling curves were obtained with the help of the **iNEXT** package [@hsieh2016].
@@
 
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
+For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) selected unmatched cases and fix orthographic errors; (iii) retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
 
@@
 For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated with the the phylogenetic signal with the help of the function _phylosig_ from the **phytools** package.
+Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
+The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
+We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it with the help of the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**; and for plants the phylogenetic tree was obtained from a species level PLANT phylogeny [@smith2018] with the help of **rtree** package [@li2023].
 
@@
 
-The different sites per study were classified with a habitat type by the authors.
+The different sites per study were described with a habitat type by the authors.
 As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
@@
 The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
-This suggest that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 2a-2b**).
+This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 2a-2b**).
 The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
@@
 
-Connectance and nestedness did not change across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
+Connectance and nestedness did not change across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
 However, differences were found for connectance but not nestedness across bioclimatic regions.
@@
 In addition, the atlantic and continental bioclimatic regions also showed statistical meaningful differences in connectance (Tukey test _P_ < 0.01).
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed an negative exponential relationship with the geometric mean of plant and pollinator species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6c**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
-Finally, we found that `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, `r round(no_diff_z_percent, 2)`% showed no statistical difference, and `r round(under_z_percent, 2)`% were less nested than null expectations (**Figure 6d**).
+Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the geometric mean of plant and pollinator species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6c**).
+Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
+Finally, we found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6d**).
 
@@
 # 3 | DISCUSSION
+
+Things to do:
+-Habitat Figure: Patterns of boreal are weird --> Comment in results or discussion
+-Figure 2 Bombus por encima de Apis, explain that Apis is not included
+-Intro: Somewhere, bur maybe later, we should mention what we add.
+1) Way more networks than those on Mangal or Webof life (which you should cite).
+2) curated metadata and harmonization, 3) Community level data and not only isolated interactions as in GloBI.
+-Network properties / significance issue
 

### 2024-05-08 Manuscript/EuPPollNet_word.qmd
Adding Nacho's comments to main text, some few things left. I left them in things to do in the discussion

@@
 Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@gonzalez2013; @settele2016; @eichenberg2021].
+The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
 Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
@@
 The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003], and has added relevant information of their interaction ecology of plant-pollinator networks across geographic regions [@olesen2002; @traveset2016] and environmental gradients [@ramos2010; @rech2016; @saunders2023].
+Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
+In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023].
 Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
@@
 Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage along the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
-
-Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized informative interaction data of plants and pollinators at European level.
+Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
+
+Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level.
 The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
@@
 
-For plants: (i) we initially verify exact matches with the GBIF species checklist; (ii) select unmatched cases and fix orthographic errors; (iii) retrieve again taxonomic information for those unmatched cases, evaluate accuracy of fuzzing matching and programmatically fix records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
+For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) selected unmatched cases and fix orthographic errors; (iii) retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that were still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
 
@@
 https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
+
+Major changes in NW Europe occurred pre-1990
+https://doi.org/10.1111/ele.12121
+
+EuPPollNet represents the largest database of plant-pollinator interactions compiled to date and is a work in progress database as there are more studies, unsampled habitats, species, etc...
 

### 2024-05-09 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time.
+This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
 We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
@@
 Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
-Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.", fig.width=10, fig.height=6, out.width="115%"}
+Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
+Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
 
@@
 
+#Karise is a very large study that only sampled bumblebees and
+#is providing unrealistic results for the Boreal region
+#Filter out
+#Maybe add to caption?
+data1 = data %>% 
+filter(!Study_id == "13_Karise")
+
 #Check number of studies by habitat
-habitat = data %>% 
+habitat = data1 %>% 
 select(Study_id, EuPPollNet_habitat) %>% 
@@
 #Check number of studies by bioregion
-bioregion = data %>% 
+bioregion = data1 %>% 
 select(Study_id, Bioregion) %>% 
@@
 #Check pols by habitat
-pollinator_percent = data %>%
+pollinator_percent = data1 %>%
 group_by(EuPPollNet_habitat, Pollinator_order) %>%
@@
 #Check pols by bioregion
-pollinator_percent_bio = data %>%
+pollinator_percent_bio = data1 %>%
 group_by(Bioregion, Pollinator_order) %>%
@@
 
-Connectance and nestedness did not change across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
+Connectance and nestedness did not show relevant trends across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
 However, differences were found for connectance but not nestedness across bioclimatic regions.
@@
 # 3 | DISCUSSION
-
-Things to do:
--Habitat Figure: Patterns of boreal are weird --> Comment in results or discussion
--Figure 2 Bombus por encima de Apis, explain that Apis is not included
--Intro: Somewhere, bur maybe later, we should mention what we add.
-1) Way more networks than those on Mangal or Webof life (which you should cite).
-2) curated metadata and harmonization, 3) Community level data and not only isolated interactions as in GloBI.
--Network properties / significance issue
 

### 2024-05-10 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 # 3 | DISCUSSION
+
+EuPPollNet offers the largest set of plant-pollinator networks and plant-pollinator studies compiled to date at European level.
+The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million of interaction records.
+Plant and pollinator species are rarely shared across multiple sites, indicating that their observed diversity strongly depends on the local context.
+While the overall sampling coverage of plant and pollinator species is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (e.g., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
+Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity [@hass2018; @martinez2022], adding studies on other habitat types is likely to rapidly increase the species coverage.
+
+Check for more references of habitat heterogeneity
+Interactions|Add reference
+
+Paragraph
+
+Sampling and taxonomic coverage
+Habitat type and bioclimatic region 
+Network properties
+
+Minor things to say
+-Presence of honey bees in context with previous studies
+Maybe cite Proc B and myself?
 

### 2024-05-12 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 While the overall sampling coverage of plant and pollinator species is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (e.g., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity [@hass2018; @martinez2022], adding studies on other habitat types is likely to rapidly increase the species coverage.
-
-Check for more references of habitat heterogeneity
-Interactions|Add reference
+Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to rapidly increase species coverage.
+
+Maybe something about: With relatively little sampling effort a large fraction of species are captured but strong efforts are needed to fully capture species within an habitat.
+
+Link with interactions
+
+Interactions:
+Sampling area, spp accumulation vs interactions:
+https://doi.org/10.1890/12-0367.1
+Sampling completeness
 

### 2024-05-13 Manuscript/EuPPollNet_Ms.qmd
1st paragraph discussion

--- 2024-05-09 Manuscript/EuPPollNet_Ms.qmd
@@
 
+EuPPollNet offers the largest set of plant-pollinator networks and plant-pollinator studies compiled to date at European level.
+The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million of interaction records.
+While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (e.g., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
+Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
+Plant and pollinator species are rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
+This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
+In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require of numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
+
+Taxonomy paragraph?
+
+Despite a similar number of Hymenoptera and Diptera species, most of the interactions were done by Hymenoptera species (i.e., mainly bees). Notably,
+
+Ollerton
+Contextualise European diversity of plant and pollinator species with global biodiveristy
+
+Network metric paragraph?
+This higher number of pollinator species compared to plants seems to be a common feature of phytocentric sampled plant-pollinator networks [@jordano2016].
+
+Link with interactions
+
+Interactions:
+Sampling area, spp accumulation vs interactions:
+https://doi.org/10.1890/12-0367.1
+Sampling completeness 
+ [e.g., @chacoff2012]
+
+Paragraph
+
+Habitat type and bioclimatic region
+
+Minor things to say
+-Presence of honey bees in context with previous studies
+Maybe cite Proc B and myself?
+
 Although Europe...
@@
 Networks, ref:
-
 https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
-
-Contextualise European diversity of plant and pollinator species with global biodiveristy
 

### 2024-05-14 Manuscript/EuPPollNet_Ms.qmd
paragraph 2 ongoing...

@@
 Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-Plant and pollinator species are rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
+Plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
 This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
@@
 
-Taxonomy paragraph?
+Bees conducted the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
+For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were especially fly species rich or fly-dominated.
+These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
+In addition, beetles were only species rich in the Mediterranean region.
+
+Comment on
+Coleoptera -sampling methodology-
+Lepidoptera -Ollerton ref, global context-
+
+Bee diversity peaked in the Mediterranean area as found in previous large scale comparisons [REF] and beetles were particularly relevant in this region
+
+, while non-bee insects such as syrphids were particularly relevant in colder 
+non-bee insect pollinators were particularly relevant
+
+Hymenoptera and Diptera had a similar number of species, the majority of interactions were conducted by bees.
+
+Bees conducted the majority of interactions within the database which is in agreement with previous large scale studies conducted in Europe [REF].
+
+and for the majority of Angiosperm species worldwide [@ollerton2017]. The proportion of the major pollinator orders within the database
+
+Despite Hymenoptera and Diptera species have a similar number of species within the database, bees conducted most of the interactions as found in previous large scale studies [@Lanuza2023].
+
+Ollerton
+Myself
+Someone else?
 

### 2024-05-15 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 In addition, beetles were only species rich in the Mediterranean region.
-
-Comment on
-Coleoptera -sampling methodology-
-Lepidoptera -Ollerton ref, global context-
-
-Bee diversity peaked in the Mediterranean area as found in previous large scale comparisons [REF] and beetles were particularly relevant in this region
-
-, while non-bee insects such as syrphids were particularly relevant in colder 
-non-bee insect pollinators were particularly relevant
-
-Hymenoptera and Diptera had a similar number of species, the majority of interactions were conducted by bees.
-
-Bees conducted the majority of interactions within the database which is in agreement with previous large scale studies conducted in Europe [REF].
-
-and for the majority of Angiosperm species worldwide [@ollerton2017]. The proportion of the major pollinator orders within the database
-
-Despite Hymenoptera and Diptera species have a similar number of species within the database, bees conducted most of the interactions as found in previous large scale studies [@Lanuza2023].
-
-Ollerton
-Myself
-Someone else?
-
-Despite a similar number of Hymenoptera and Diptera species, most of the interactions were done by Hymenoptera species (i.e., mainly bees). Notably,
-
-Ollerton
-Contextualise European diversity of plant and pollinator species with global biodiveristy
-
-Network metric paragraph?
+Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their role as pollinators in this region [@herrera2019; @leon2022].
+The number of butterfly species and interactions were relatively low compared to the other taxa.
+Although Europe contains a little number of butterfly species compared to other areas of the world [@ollerton2017], their relevance as pollinators is likely under-estimated within this database as sampling methodologies for other insects pollinators (e.g., bees or flies) generally require less sampling distance and cover less space.
+Honey bees did one third of the total recorded interactions per network, highlighting their high abundances in natural communities and possibly impacting negatively wild pollinators.
+
+-Presence of honey bees in context with previous studies
+Maybe cite Proc B and myself?
+
+**Network paragraph**
+
 This higher number of pollinator species compared to plants seems to be a common feature of phytocentric sampled plant-pollinator networks [@jordano2016].
@@
 
-Paragraph
-
-Habitat type and bioclimatic region
-
-Minor things to say
--Presence of honey bees in context with previous studies
-Maybe cite Proc B and myself?
-
-Although Europe...
-
+ Networks, ref:
+https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
+
+**Conclusion paragraph**
 Geographic and taxonomic biases!
@@
 https://doi.org/10.1111/ecog.05926
-
-Networks, ref:
-https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
 

### 2024-05-16 Manuscript/EuPPollNet_Ms.qmd
second paragraph

@@
 mutate(Bioregion = tolower(Bioregion))
+#Check percentage of Apis mellifera across networks
+apis_percent_networks = data %>% 
+filter(Pollinator_accepted_name == "Apis mellifera") %>%
+group_by(Study_id) %>% 
+mutate(Apis = "Yes") %>% 
+distinct(Study_id, Apis) %>% 
+nrow() / nlevels(factor(data$Study_id)) *100
 
@@
 The number of butterfly species and interactions were relatively low compared to the other taxa.
-Although Europe contains a little number of butterfly species compared to other areas of the world [@ollerton2017], their relevance as pollinators is likely under-estimated within this database as sampling methodologies for other insects pollinators (e.g., bees or flies) generally require less sampling distance and cover less space.
-Honey bees did one third of the total recorded interactions per network, highlighting their high abundances in natural communities and possibly impacting negatively wild pollinators.
-
--Presence of honey bees in context with previous studies
-Maybe cite Proc B and myself?
+While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
+This is because a large fraction (~40%) of studies did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
+Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted in average a third of the total interactions per network.
+This proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights the need of further work evaluating how honey bees and their management impact natural pollinator communities.
 
@@
 Link with interactions
+
+Connectance
+Nestedness
+Pattern across latitude
 

### 2024-05-17 Manuscript/EuPPollNet_Ms.qmd
edits
Starting 3rd paragraph

@@
 plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
+
+#Calculate mean coverage and deviation
+coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
+mean_coverage = mean(coverage)
+sd_coverage = sd(coverage)
 
@@
 For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
+The potential number of pollinator species at European level was estimated by adding the total number of species of bees, syrphids, and flies from the checklists, along with the extrapolated number of species from other insect pollinators.
+This extrapolated number was estimated by assuming that the sampling coverage of these other insect pollinators is equal to the average coverage across bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+
 Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
 The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it with the help of the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**; and for plants the phylogenetic tree was obtained from a species level PLANT phylogeny [@smith2018] with the help of **rtree** package [@li2023].
+We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it with the help of the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**; and for plants the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of **rtree** package [@li2023].
 
@@
 
-**Network paragraph**
-
-This higher number of pollinator species compared to plants seems to be a common feature of phytocentric sampled plant-pollinator networks [@jordano2016].
-
-Link with interactions
-
-Connectance
+Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of observed plant species.
+This seems a common feature of phytocentric plant-pollinator networks, which tend to overlook rare plant species and capture a higher proportion of pollinator species [REF].
+
+Discuss this
+
+The number of pollinator species almost doubled the number of plant species within the database.
+This is a common aspect in phytocentric networks [@jordano2016], where an observer records the different pollinators that land on the reproductive structures of plants, resulting in a higher number of pollinator species per plant.
+
+Link with connectance now?
+
+Discuss that other type of informative networks could be useful an increase connectance?
+
+visitation networks
+
+Link with interactions -->
+
+More species less connectance, which increases the number of missing links.
+
 Nestedness
 Pattern across latitude
+
+Missing links?
+
+networks condense information but this workflow allo possibility to explore species phenology... and more
 
@@
 **Conclusion paragraph**
+
+foundation for future work
+
 Geographic and taxonomic biases!

### 2024-05-20 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 
-Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of observed plant species.
-This seems a common feature of phytocentric plant-pollinator networks, which tend to overlook rare plant species and capture a higher proportion of pollinator species [REF].
-
-Discuss this
-
-The number of pollinator species almost doubled the number of plant species within the database.
-This is a common aspect in phytocentric networks [@jordano2016], where an observer records the different pollinators that land on the reproductive structures of plants, resulting in a higher number of pollinator species per plant.
-
-Link with connectance now?
+Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of plants in the database.
+This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
+While animal-centred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
+As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
+The less connected networks were the Alpine ones as they were the ones with higher number of species.
+
+#Task: Calcular residual connectance.
+
+with more species had lower connectance .
+
+Finish this paragraph with an idea of how sampling can bias our view and that is important to be aware of that when understanding plant-pollinator patterns
 

### 2024-05-22 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 
-To get a general overview of some of the main aspects of these set of plant-pollinator networks, we evaluated how connectance and nestedness change across the latitudinal range of studies and bioclimatic regions.
-We implemented linear models, with the network metrics as the response variable, and both latitude and bioclimatic region as predictors in two separate models.
-Finally, with the help of a Tukey test from the package _emmeans_, we evaluated if nestedness and connectance differ statistically across bioclimatic regions.
-In addition, we explored the association between connectance and nestedness with the number of species per network (i.e., geometric mean of plants and pollinators) by conducting Kendall rank correlation coefficient.
-Finally, we evaluated if the different networks are more or less nested than expected by chance.
-Importantly, we implemented two different metrics of nestedness given the different analytical objectives.
-Common metrics of nestedness are highly correlated to network size, limiting our ability to compare this metric across networks.
-To avoid this, we implemented a normalised version of nestedness (NODFc) that corrects by connectance and the number of species [@song2017].
-This metric was calculated with the help of the **maxnodf** package [@hoeppke2021], which employs a computational efficient approach to calculate it.
-However, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the _vaznull_ function from the package **Bipartite** [@dormann2008] that implements the null model from @vazquez2005.
+To provide a general overview of the main aspects of this set of plant-pollinator networks, we evaluated network patterns of connectance and nestedness across the latitudinal range of studies and bioclimatic regions.
+We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
+However, these network metrics are not independent of sampling effort (e.g., number of species) and to allow comparisons across them, we implemented “standardised” versions of connectance and nestedness.
+As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species  (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
+This was done with the help of a Beta regression implemented with the package **betareg** [@cribari2010].
+Second, we implemented a metric of nedstedness (i.e., NODFc) that allows the comparison across networks as it corrects by connectance and the number of species [@song2017].
+This metric was calculated with the help of the **maxnodf** package [@hoeppke2021], which employs a computational eﬀicient approach to calculate it.
+Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
+In addition, we explored the association between connectance and nestedness with the number of species per network by conducting Kendall rank correlation coefficient in order to be able to compare the strength of this association across both network metrics.
+
+Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
+We calculated 100 null models for each network with the help of the vaznull function from the package **bipartite** [@dormann2008] that implements the null model from @vazquez2005.
 These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
-In other words, interactions are reshuffled with almost equal probabilities to their relative abundances.
-Connectance was estimated for each network by using the function _networklevel_ from **Bipartite**.
+Both connectance and nestedness (NODF) were estimated for each network by using the function _networklevel_ from **bipartite**.
 
@@
 
-Connectance and nestedness did not show relevant trends across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
-However, differences were found for connectance but not nestedness across bioclimatic regions.
-Specifically, the alpine region differed statistically in connectance to the other bioclimatic regions (Tukey test _P_ < 0.01).
-In addition, the atlantic and continental bioclimatic regions also showed statistical meaningful differences in connectance (Tukey test _P_ < 0.01).
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the geometric mean of plant and pollinator species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6c**).
+Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
 Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
-Finally, we found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6d**).
+We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**).
+Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
+Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes.
 
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
-Graphs (a-b) indicate network connectance and nestedness across the latitudinal range of the studies from the database.
-In addition, the bioclimatic region where each network is located is indicated with points of different shapes and colours.
-Graph (c) shows the association between network connectance and the geometric mean of plant and pollinator species per network.
-Graph (d) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
+Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
+Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
 The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
-Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).", fig.width=10, fig.height=7, fig.align='center'}
+Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).
+Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
+In addition, the bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
 
 # Load libraries
-
-library(psych)
-# Load data
-data = readRDS("../Data/3_Final_data/Interaction_data.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-null_networks = readRDS("../Data/Working_files/null_networks.rds")
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks.rds")
-
-# Prepare data
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-              Month = lubridate::month(Date), 
-              Day = lubridate::day(Date))
-
-# Subset data for studies with only bumblebees
-v = c("13_Karise", "22_Kallnik")
-subsetting = unique(data$Study_id)
-
-long_format = data %>% 
-mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
-mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
-filter(!Study_id %in% v) %>% 
-select(Network_id, Plant_accepted_name, Pollinator_accepted_name)
-
-# Filter networks larger than 4 plants and 4 pollinator species
-v2 = long_format %>% 
-group_by(Network_id) %>% 
-summarise(Poll_spp = n_distinct(Plant_accepted_name),
-          Plant_spp = n_distinct(Pollinator_accepted_name)) %>% 
-ungroup() %>% 
-filter(Poll_spp >= 4 & Plant_spp >= 4) %>% 
-pull(Network_id)
-
-long_format = long_format %>% 
-filter(Network_id %in% v2)
-
-# Filter networks with more links than species
-links = long_format %>% 
-group_by(Network_id, Plant_accepted_name, Pollinator_accepted_name) %>%
-distinct() %>% 
-group_by(Network_id) %>% 
-summarise(links = length(Network_id))
-
-spp = long_format %>% 
-group_by(Network_id) %>% 
-summarise(n_plants = n_distinct(Plant_accepted_name),
-          n_polls = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(n_total = n_plants + n_polls)
-
-vector_to_include = left_join(links, spp)
-
-v3 = vector_to_include %>% 
-filter(links > n_total) %>% 
-pull(Network_id)
-
-long_format = long_format %>% 
-filter(Network_id %in% v3)
-
-# Explore how NODFc changes across the Y-X AXES (latitude-Longitude)
-long_format_coords = data %>% 
-mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
-mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
-select(Network_id, Latitude, Longitude, Bioregion) %>% 
-distinct()
-
-# Merge data
-metrics_by_network_coords = left_join(metrics_by_network, long_format_coords)
-
-# Connectance plots
-# 1st Latitude
-p1 = metrics_by_network_coords %>% 
-ggplot(aes(Latitude, Connectance)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.15, size=2.5, alpha=1) +
+library(dplyr)
+library(ggplot2)
+library(stringr)
+library(patchwork)
+
+# Load plots generated in scripts 7_2_Nestedness and 7_2_Connectance
+
+connectance_species = readRDS("../Data/Manuscript_info/Figure6_connectance_species.rds")
+connectance_latitude = readRDS("../Data/Manuscript_info/Figure6_connectance_latitude.rds")
+nestedness_z_scores = readRDS("../Data/Manuscript_info/nestedness_z_scores.rds")
+nestedness_latitude = readRDS("../Data/Manuscript_info/nestedness_latitude.rds")
+
+#Prepare first connectance plots
+library(betareg)
+model = betareg(Connectance ~ log_geometric_mean_spp, data = connectance_species)
+p1 = ggplot(connectance_species, aes(x = log_geometric_mean_spp, y = Connectance))+
+geom_point(pch = 21, fill = "azure3", size=2.5) +
+geom_line(aes(y = predict(model, connectance_species)), colour= "black")+
+xlab("log(Species)") +
 theme_bw() +
-ylab("Connectance") +
+ggtitle("(a)")
+
+p3 = connectance_latitude %>% 
+ggplot(aes(Latitude, residual_conectance)) +
+geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
 scale_fill_viridis_d() +
@@
   "Mediterranean" = 25  # Cross
-))+
-ggtitle("(a)")
+  )) +
+geom_line(data = connectance_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
+geom_ribbon(data = connectance_latitude, 
+    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
+    fill = "grey70", alpha= 0.3) +
+theme_bw()+
+ylab("Residual connectance") +
+ggtitle("(c)")
+
+#Prepare now nestedness plots
+# Plot
+p = 0.05 #cutoff probability 95% confidence
+critical_value = qnorm(p/2) #double tail probability divide by 2
+
+nestedness_z_scores = nestedness_z_scores %>%
+mutate(infra_over_represented = case_when(
+z_score < -abs(critical_value) ~ "Under-represented",
+between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
+    z_score > abs(critical_value) ~ "Over-represented"
+  ))
+
+p2 = nestedness_z_scores %>% 
+ggplot(aes(x = z_score)) +
+geom_histogram(aes(y = ..density.., fill = infra_over_represented), 
+               bins = 100, alpha = 0.5, color = "black", position = "stack") +
+stat_function(data = nestedness_z_scores, fun = function(x) dnorm(x, mean = mean(nestedness_z_scores$z_score), sd = sd(nestedness_z_scores$z_score)) * 1.7, 
+              n = 1000, inherit.aes = FALSE, color = "gray18", size = 1) +
+theme_bw() +
+coord_cartesian(expand = FALSE) +
+geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
+geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
+ylab("Density") +
+xlab("Z scores (NODF)")  +
+xlim(-6, 9) +
+ylim(0, 0.8) +
+scale_fill_manual(name = "Observed against null",
+                  limits = c("Under-represented", "No statistical difference", "Over-represented"),
+                  labels = c("Less nested", "No difference", "More nested"),
+                  values = c("Under-represented" = "coral2", 
+                             "No statistical difference" = "palegreen3", 
+                             "Over-represented" = "cyan3")) +
+ggtitle("(b)")
 
 # 2nd Plot
-p2 = metrics_by_network_coords %>% 
+p4 = nestedness_latitude %>% 
 ggplot(aes(Latitude, Normalised_nestedness)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.15, size=2.5, alpha=1) +
+geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
 theme_bw() +
@@
   )) +
-ggtitle("(b)")
-
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks)
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-# 2nd Species (geometric mean)
-species_number = long_format %>% 
-group_by(Network_id) %>% 
-summarise(N_plants = n_distinct(Plant_accepted_name),
-          N_pollinators = n_distinct(Pollinator_accepted_name)) %>% 
-group_by(Network_id) %>% 
-summarise(geometric_mean_spp = geometric.mean(c(N_plants, N_pollinators)))
-
-metrics_by_network_spp_number = left_join(metrics_by_network, species_number)
-
-p3 = metrics_by_network_spp_number %>% 
-ggplot(aes(geometric_mean_spp, Connectance)) +
-geom_point(pch = 21, fill = "azure3", size=2.5, stroke=0.25, alpha=1) +
-theme_bw() +
-xlab("Species")+
-ggtitle("(c)")
-
-# Plot
-p4 = d1 %>% 
-ggplot(aes(x = z_score)) +
-geom_histogram(aes(y = ..density.., fill = infra_over_represented), 
-               bins = 100, alpha = 0.5, color = "black", position = "stack", size=0.1) +
-stat_function(data = d1, fun = function(x) dnorm(x, mean = mean(d1$z_score), sd = sd(d1$z_score)) * 1.7, 
-                n = 1000, inherit.aes = FALSE, color = "gray18", size = 1) +
-theme_bw() +
-coord_cartesian(expand = FALSE) +
-geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red")+
-geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-ylab("Density") +
-xlab("Z scores (NODF)")  +
-xlim(-6, 9) +
-ylim(0, 0.8) +
-scale_fill_manual(name = "Observed NODF against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
+geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
+geom_ribbon(data = nestedness_latitude, 
+    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
+    fill = "grey70", alpha= 0.3) +
+theme_bw()+
 ggtitle("(d)")
@@
 # Layout plots
-a = p1 + p2 + plot_layout(guides = 'collect') & theme(legend.position = "right",
-                                                      legend.justification = "left")
-b = (p3 + p4)  & theme(legend.justification = "left")
+#a = (p1 + p2) 
+#b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.position = "right")
+#a / b
+
+# Layout plots
+a = (p1 + p2)  & theme(legend.position = "right", legend.justification = "left")
+b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.justification = "left")
 a / b  & theme(axis.title =  element_text(size=16, face="bold"),
@@
                fill=NA, size=1),
-               axis.text=element_text(size=12),
-               legend.text = element_text(size=10),
-               legend.key.size = unit(6, "mm"),
-               legend.spacing.y = unit(0.5, "cm"),
-               legend.box.spacing = unit(0, "pt"),
-               legend.title = element_text(face = "bold"))
+               axis.text=element_text(size=12))
 
@@
 
+metrics_by_network_spp_number = readRDS("../Data/Working_files/metrics_by_network_spp_number.rds")
+
 metrics_by_network_spp_number %>% 

### 2024-05-22 Manuscript/Figures_and_Tables/Figure_6_Updated.Rmd
edits

--- 2024-05-22 Manuscript/EuPPollNet_Ms.qmd
@@
 ---
-title: "EuPPollNet: A European database of plant-pollinator networks"
-format:
-  pdf:
-    include-in-header: 
-      text: |
-        \usepackage{lscape}
-        \newcommand{\blandscape}{\begin{landscape}}
-        \newcommand{\elandscape}{\end{landscape}}
-        \usepackage{caption}
-        \captionsetup[figure]{labelformat=empty}
-        \captionsetup[table]{labelformat=empty}
-
-bibliography: references.bib
-latex-tinytex: false
-link-citations: TRUE
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
+output:
+  pdf_document:
+header-includes:
+- \usepackage{caption}
+- \captionsetup[figure]{labelformat=empty}
+- \captionsetup[table]{labelformat=empty}
 ---
 
-```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr) #For handling fata 
-library(sf) #For handling coordinates and plotting
-library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
-library(readr)
-library(phytools)
-#Load libraries for plotting
-library(giscoR) #Maps
-library(ggstar) #cool shapes as points
-library(ggplot2) #general plotting
-library(ggtree) #phylo trees
-library(ggpattern) #cool shapes in barplots
-library(patchwork) #plot organisation
-library(cowplot) #plot organisation
-library(viridis) #colours
-library(ggnewscale)
-library(ggtreeExtra)
-library(kableExtra)
-
-#In case of conflicts select the main dplyr functions
-library(conflicted)
-conflicts_prefer(dplyr::filter)
-conflicts_prefer(dplyr::select)
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
-#Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
-                    header = T)
-#Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")
-
-#Work around to get the three cols
-#Note that month 1 and day 1 indicates the presence of a missing value
-#At least in Europe :)
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Read author list 
-authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
-#Print authors
-authors = authors %>% 
-summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
-pull()
-#Read affiliation list 
-aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
-aff_list = aff_list %>% 
-summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
-pull()
-
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Calculate mean coverage and deviation
-coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
-mean_coverage = mean(coverage)
-sd_coverage = sd(coverage)
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-#Prepare table with all info 
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-d = d %>% 
-mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
-filter(!Variable == "Longitude")
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
-        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
-
-    TRUE ~ "Other"
-  ))
-
-#Check number of species and percentage of interactions of poll. groups
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-order_info = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
-          Species = n_distinct(Pollinator_accepted_name))
-
-Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
-select(Interactions) %>% pull()
-
-#Check percentage of studies with flower counts
-floral_counts_percent = data  %>%
-select(Study_id, Flower_data) %>% 
-distinct() %>% 
-group_by(Flower_data) %>% 
-summarise(Percentage = n()/ nrow(.)*100) %>% 
-filter(Flower_data == "Yes")
-
-#Check percentage of apis mellifera records
-all_records = data %>% nrow(.) 
-apis_records = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-apis_percent = apis_records/all_records * 100
-#Check average pecentage of interactions of apis by net
-a = data %>% 
-group_by(Network_id) %>% 
-summarise(n_interactions = length(Pollinator_accepted_name))
-b = data %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-summarise(n_interactions_apis = length(Pollinator_accepted_name))
-c = left_join(a,b)
-apis_average_network = c %>% 
-mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
-ungroup() %>% 
-summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
-pull()
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first three and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the intricate network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising its long-term persistence.
-One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences, while also providing other ecological variables of interest.
-Furthermore, this work provides an in-depth assessment of the taxonomic and sampling coverage of the database at the European level, complemented by analyses of key structural properties in plant-pollinator networks.
-We hope this database can help researchers to: 1) identify taxonomic, ecological, and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change on plant-pollinator networks to guide future conservation planning for both plant and pollinator species.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_sites, big.mark=",")` distinct locations (i.e., plant-pollinator networks), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided.
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
-
-**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats.
-The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
-
-**KEYWORDS**\
-\begingroup
-\small
-plant-pollinator networks, plant-pollinator interactions, flowering plants, Angiosperms, pollinators, nestedness, connectance
-\endgroup
-\pagebreak
-
-# 1 | INTRODUCTION
-
-The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
-In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023].
-Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
-
-Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions.
-Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations.
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time.
-This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
-We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
-
-# 2 | METHODS
-
-## Data acquisition
-
-This database is the result of one of the working packages of the European project Safeguard.
-The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-Data acquisition followed a non-systematic approach.
-First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions within the database.
-To keep high quality standards that will allow robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
-
-## Dataset description
-
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**).
-The different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant.
-Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species.
-However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
-Notably,  _Apis mellifera_ represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the EuPPollNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Number of studies by year within the database.
-(\\textbf{c}) Number of species and interactions across the 4 main pollinator orders at European level."}
-
-#Fix dates
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-#p1 = ggplot(poll_spread, 
-#aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-#geom_bar(stat = "identity", 
-#           color = "#D55E00",
-#           lwd = 0.015, fill = "#D55E00") +
-#ylab("Shared pollinators (%)") +
-#theme_minimal()+
-#theme(axis.text.x = element_blank(),
-#  axis.ticks.x=element_blank(),
-#  axis.ticks.y = element_line(size = 0.2),
-#  plot.title = element_text(size = 8, face = "bold"), 
-#  axis.text.y = element_text(size=5),
-#  panel.grid.major.x = element_blank(),
-#  panel.grid.minor.y = element_blank(),
-#  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-#xlab(NULL) +
-#ylab(NULL) + 
-#ggtitle("Pollinators")+
-#geom_vline(xintercept = quantile_50, linetype="dashed", 
-#color = "black", size=0.25) + 
-#annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = #1.5) +
-#annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  #fontface = 'italic', size = 1.5)+
-#scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-#annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
-#coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, #max(poll_spread$Percent_total)), clip="off")
-#
-##----------------#
-##Plants
-##----------------#
-#plant_spread = data %>% 
-#select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-#group_by(Study_id) %>% 
-#filter(Plant_rank == "SPECIES") %>% 
-#distinct(Plant_accepted_name) %>% 
-#ungroup() %>% 
-#count(Plant_accepted_name) %>% 
-#rename(n_studies_shared = n) %>% 
-#arrange(-n_studies_shared) %>% 
-#mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-#  
-##Seelect quantile 50% from the total number of species
-#spp_number = plant_spread %>%  
-#select(Plant_accepted_name) %>% 
-#n_distinct()
-#df = seq(1, spp_number)
-#quantile_50 = quantile(df, 0.5)
-#
-##Obtain most common plant for plotting
-#most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-#select(Plant_accepted_name) %>% pull() 
-#
-#Plot
-#p2 = ggplot(plant_spread, 
-#aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-#geom_bar(stat = "identity", 
-#           color = "#009E73",
-#           lwd = 0.015, fill = "#009E73") +
-#theme_minimal()+
-#theme(axis.text.x = element_blank(),
-#         axis.ticks.x=element_blank(),
-#         axis.ticks.y = element_line(size = 0.2),
-#         axis.title.x = element_text(face = "bold", size = 7),            plot.title = #element_text(size = 8, face = "bold"),
-#         axis.text.y = element_text(size=5),
-#         panel.grid.major.x = element_blank(),
-#         panel.grid.minor.y = element_blank(),
-#         panel.grid.major.y = element_line(size=0.2, linetype =  2),
-#      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
-#xlab("Species") +
-#ylab("Occurrence across studies") + 
-#ggtitle("Plants") + 
-#geom_vline(xintercept = quantile_50, linetype="dashed", 
-#          color = "black", size=0.25) + 
-#annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-#annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
-#      ,color="black",  fontface = 'italic', size = 1.5)+
-#annotate(geom="text", x=20, y=-0.08, label="Common spp."
-#      ,color="black",  size = 1.5) +
-#annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
-#      ,color="black",  size = 1.5) + 
-#scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = #label_number(accuracy = 0.01)) +
-#coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, #max(plant_spread$Percent_total)), clip="off")
-#
-
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p2  = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-   axis.ticks.x = element_line(color="black"),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -5))
-
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-#Select main orders
-data1 = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.),
-          Interactions = length(Pollinator_accepted_name)/ nrow(.))
-
-# Combine Interactions and Species into a single column
-data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
-
-#Set order of levels for plotting
-data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))
-
-#Plot
-p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
-geom_bar(stat = "identity", width = 0.7) +
-theme_classic() +
-scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator order") +
-xlab(NULL) +
-ylab("Percentage") + 
-scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
-theme(axis.text = element_text(size=6), 
-     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
-     axis.title.x = element_text(face="bold", size = 7),
-     legend.position = "top",
-     legend.title = element_text(size = 5, face = "bold"),
-     legend.text = element_text(size = 5),
-     legend.key.size = unit(3, "mm"),
-     legend.margin=margin(0,0,0,0),
-     legend.box.margin=margin(-8,-5,-5,-5)) +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))+
-ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6))
-
-#plot2 = p3 / p1 / p2 
-plot2 = p2 / p3
-
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.485, 0.8))
-
-```
-
-\newpage
-
-## Data structure
-
-The EuPPollNet database is available in both .csv and .rds formats.
-The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
-The file contains a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
-To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study.
-Information about habitat type and bioclimatic region are also provided in this file.
-In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study.
-The flower count data is provided in a separate file given that one third of studies lack floral measurements and that the units differ greatly across studies.
-Finally, metadata at the study level is provided, including information about the authors, digital object identifier if available, sampling time and taxonomic coverage of the main pollinator groups for each study.
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.} Column names and their descriptions within the EuPPollNet database.") %>%
-  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
-kable_styling(position = "center")
-
-```
-
-## Sampling coverage
-
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
-In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [@kleijn2015].
-To that end, we also calculated the accumulation curve of pollinator species with increasing number plant species.
-The rarefied and extrapolated sampling curves were obtained with the help of the **iNEXT** package [@hsieh2016].
-This was complemented with 100 random accumulation curves obtained programmatically to guide the visualization of the different rarefied curves.
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) selected unmatched cases and fix orthographic errors; (iii) retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at European level by combining the recently published checklists of bees [@ghisbain2023], syrphids [@kovcic2023] and butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the EuPPollNet database at European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-The potential number of pollinator species at European level was estimated by adding the total number of species of bees, syrphids, and flies from the checklists, along with the extrapolated number of species from other insect pollinators.
-This extrapolated number was estimated by assuming that the sampling coverage of these other insect pollinators is equal to the average coverage across bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
-
-Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
-The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it with the help of the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**; and for plants the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of **rtree** package [@li2023].
-
-## Habitat type and bioclimatic region
-
-The different sites per study were described with a habitat type by the authors.
-As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
-For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
-Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the EuPPollNet database (see habitat type definition in supplementary material).
-These categories allow a quick comparison and understanding of the habitat types from the database.
-However, we advise authors to revise this classification if they intend to rely on this field for their analyses as this is a non-fully objective process.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
-
-## Network analyses
-
-To provide a general overview of the main aspects of this set of plant-pollinator networks, we evaluated network patterns of connectance and nestedness across the latitudinal range of studies and bioclimatic regions.
-We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
-However, these network metrics are not independent of sampling effort (e.g., number of species) and to allow comparisons across them, we implemented “standardised” versions of connectance and nestedness.
-As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species  (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-This was done with the help of a Beta regression implemented with the package **betareg** [@cribari2010].
-Second, we implemented a metric of nedstedness (i.e., NODFc) that allows the comparison across networks as it corrects by connectance and the number of species [@song2017].
-This metric was calculated with the help of the **maxnodf** package [@hoeppke2021], which employs a computational eﬀicient approach to calculate it.
-Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, we explored the association between connectance and nestedness with the number of species per network by conducting Kendall rank correlation coefficient in order to be able to compare the strength of this association across both network metrics.
-
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the vaznull function from the package **bipartite** [@dormann2008] that implements the null model from @vazquez2005.
-These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
-Both connectance and nestedness (NODF) were estimated for each network by using the function _networklevel_ from **bipartite**.
-
-\newpage
-
-# 3 | RESULTS
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Sampling coverage outputs
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Coverage
-poll_coverage = poll_output$DataInfo$SC *100
-plant_coverage = plant_output$DataInfo$SC *100
-int_coverage = int_output$DataInfo$SC *100
-plant_poll_coverage = plant_poll_output$DataInfo$SC *100
-
-#Increase doubling effort
-##Pollinators
-poll_increase = poll_output$iNextEst$Network_id
-poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
-poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
-poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
-##Plants
-plant_increase = plant_output$iNextEst$Network_id
-plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
-plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
-##Interactions
-int_increase = int_output$iNextEst$Network_id
-int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
-int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
-int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
-##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$Plant_id
-plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
-plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
-
-#Prepare results from common and rare poll across sampling sites
-#Pollinators
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-poll_total = nrow(poll_spread)
-polls_under_one_percent = poll_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-polls_under_one_percent = polls_under_one_percent/poll_total * 100
-#Find most common pollinator
-most_common_poll = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Pollinator_accepted_name)
-#Find most common poll percent
-most_common_poll_percent = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-#Plants
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-plant_total = nrow(plant_spread)
-plants_under_one_percent = plant_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-plants_under_one_percent = plants_under_one_percent/plant_total * 100  
-#Find most common plant
-most_common_plant = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Plant_accepted_name)
-#Find most common plant percent
-most_common_plant_percent = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-```
-
-## Sampling coverage
-
-The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 2a-2b**).
-The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 2c**).
-Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 2d**).
-In addition, when we explore how common or rare are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 2e-2f**).
-For instance, most plants (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinators (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) are exclusively found in less than 1% of sampling sites; and the most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are only found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites.
-The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
-Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
-Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
-
-#Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
-#Create dataset
-d_poll = poll_output$iNextEst$Network_id
-#Check cols
-d_poll$col = d_poll$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           method = as.character(pollinators_curves$iteration), 
-           qD = pollinators_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(method == "observed")
-d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
-
-#Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col),size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_poll1_point, col = "black", size=2) + 
-ylab("Pollinator species") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000) +
-ggtitle("(a)") +
-theme(plot.margin = margin(r = 12))
-
-#Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
-#Create dataset
-d_plant = plant_output$iNextEst$Network_id
-#Check cols
-d_plant$col = d_plant$method
-
-#Now create tibble with same structure
-a_plant = tibble(t = plant_curves$Study_sampled, 
-           method = as.character(plant_curves$iteration), 
-           qD = plant_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(method == "observed")
-d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
-
-#Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_plant1_point, col = "black", size=2) + 
-ylab("Plant species") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000)+
-ggtitle("(b)")+
-theme(plot.margin = margin(r = 12))
-
-#Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
-#Create dataset
-d_int = int_output$iNextEst$Network_id
-#Check cols
-d_int$col = d_int$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = interactions_curves$Study_sampled, 
-           method = as.character(interactions_curves$iteration), 
-           qD = interactions_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(method == "observed")
-d_int1_lines = d_int1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Interactions") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000) +
-ylim(0, 30000)+
-ggtitle("(c)")
-
-#----------------#
-#Pollinators
-#----------------#
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-o1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.03, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
-xlab("Pollinator species") +
-ylab("Occurrence across sampling sites") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.35) + 
-annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
-annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=200, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3)  +
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
-                ylim=c(0,0.7), clip="off") +
-ggtitle("(e)")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-o2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.03, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
-xlab("Plant species") +
-ylab("Occurrence across sampling sites") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.35) + 
-annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
-annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=150, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3) + 
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
-                ylim=c(0, 0.7), clip="off") +
-ggtitle("(f)")
-
-#Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Create dataset
-d_pp = plant_poll_output$iNextEst$Plant_id
-#Check cols
-d_pp$col = d_pp$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           method = as.character(plant_poll_curves$iteration), 
-           qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(method == "observed")
-d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col)) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_pp1_point, col = "black") + 
-ylab("Pollinator species") +
-xlab("Plant species") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(d)") +
-scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
-theme(legend.position = "none")
-
-#Plot all
-p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
-      axis.title = element_text(size=12, face = "bold"),
-      axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-#a/b
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Taxonomic coverage outputs
-#Find number of plant and pollinator species
-##Pollinators
-poll_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-##Plants
-plant_spp = data %>% 
-select(Plant_rank, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-nrow(.)
-
-#Overall plant coverage
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Average plant family coverage
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData") %>% 
-filter(fam_group == "B")
-#Average coverage per family
-mean_plant_fam_cov = d1 %>% 
-ungroup() %>% 
-summarise(mean_plant_fam_cov = mean(Percent)) %>% 
-pull()
-
-#Percent of families with interaction records
-all_fam = nrow(d1)
-fam_with_data = d1 %>% 
-filter(Percent > 0) %>% 
-nrow()
-percent_plant_fam_with_data = fam_with_data/all_fam * 100
-
-#Calculate phylogenetic signal of presence-absence information
-#Read data and plot it
-tree = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-#Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-#select(label, categorical) %>% 
-ungroup() 
-#Check phylogenetic signal
-d1 = d1 %>% arrange(factor(label, levels = tree@phylo$tip.label))
-
-v = phylo_trait$categorical
-names(v) = tree@phylo$tip.label
-
-phylo_signal_plant = phylosig(tree@phylo, v, method="lambda", test=TRUE)
-lambda_phylosig_plants = phylo_signal_plant$lambda
-p_phylosig_plants = phylo_signal_plant$P
-
-#Load pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-#Check percentage of interactions from bees with and without apis
-#Check percentage of bees in the database
-bee_fam = c("Apidae", "Megachilidae", "Halictidae",
-            "Andrenidae", "Colletidae", "Melittidae")
-all_rows = data %>% nrow(.)
-#
-bee_rows = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-nrow(.)
-#Bee interactions
-bee_interactions_percent = bee_rows/all_rows *100
-#Eclude now apis
-bee_rows_non_apis = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-#
-bee_interactions_non_apis_percent = bee_rows_non_apis/bee_rows *100
-
-#Coverage across bee genera
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-all_rows_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-nrow(.)
-#
-data_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-filter(Percent > 0) %>% 
-nrow(.)
-#
-bee_genera_percent_coverage = data_bees/all_rows_bees * 100
-#Check average coverage across bee genera
-average_coverage_bee_genera = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-ungroup() %>% 
-summarise(average_coverage_bee_genera = mean(Percent)) %>% 
-pull()
-
-#Calculate phylogenetic signal of presence-absence information
-##Read data and plot it
-tree = readRDS("../Data/Manuscript_info/bee.tree100.rds")
-##Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-ungroup() 
-##Conduct left join to tibble of tip labels to organise order
-v = tibble(Genus1 = tree$tip.label) %>% 
-mutate(Genus1 = str_to_title(word(Genus1,1,sep = "_")))
-#
-v1 = left_join(v, phylo_trait)
-#
-v2 = v1$categorical
-names(v2) = tree$tip.label
-#
-phylo_signal_bees = phylosig(tree, v2, method="lambda", test=TRUE)
-lambda_phylosig_bees = phylo_signal_bees$lambda
-p_phylosig_bees = phylo_signal_bees$P
-
-#Check number of studies by habitat
-studies_by_habitat = data %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of sampling sites by habitat
-networks_by_habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of studies by bioregion
-studies_by_bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check number of sampling sites by bioregion
-networks_by_bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check percentage of Apis mellifera across networks
-apis_percent_networks = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>%
-group_by(Study_id) %>% 
-mutate(Apis = "Yes") %>% 
-distinct(Study_id, Apis) %>% 
-nrow() / nlevels(factor(data$Study_id)) *100
-
-```
-
-## Taxonomic coverage
-
-The database contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinators groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions of the database, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 3**).
-In addition, the presence or absence of interactions records for bees seems not to follow any sort of phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
-The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 3}.
-Phylogenetic representation of bee genera in the EuPPollNet database within the context of all European bee genera.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Grey bars indicate the total number of species per genus at the European level."}
-
-#Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
-
-#Circular layour
-ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.063, size=2.9, fontface=1) +
-scale_size(range = c(0,4.5), guide = 'none')  +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-guides(fill=guide_legend(), size = guide_legend()) +
-theme(legend.position = "bottom") +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
-                    pwidth=0.13, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0.137,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-guides(fill = "none")+
-scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
-geom_fruit(geom=geom_bar,
-                    mapping=aes(y=label, x=Spp_number_Europe),
-                    pwidth=0.12, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1,
-                         vjust = 0.5,
-                         nbreak = 1,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-annotate("text", x = 14.2, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.1, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
-new_scale_fill() +
-geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
-annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
-annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
-annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
-annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
-annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
-annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
-coord_cartesian(clip = "off")+
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.6,0),
-      legend.direction = "horizontal")
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Phylogenetic representation of the plant families in the EuPPollNet database within the context of all flowering plant families occurring in Europe.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-\newpage
-
-## Habitat type and bioclimatic region
-
-The proportion of the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-As expected, Hymenoptera was the main taxonomic order on the majority of habitats, exceeded only by Diptera on the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
-Overall, the proportion of Lepidoptera and Coleoptera was low across all habitats but Coleoptera showed a notably increase in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
-Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
-Again, Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera seems to be only relevant in Mediterranean regions at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
-The habitats sampled by a higher number of studies in the database are `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
-However, the habitats that contain a higher number of sampling sites are `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies are `r studies_by_bioregion$Bioregion[[1]]` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r studies_by_bioregion$Bioregion[[2]]` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r studies_by_bioregion$Bioregion[[3]]` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r networks_by_bioregion$Bioregion[[1]]` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r networks_by_bioregion$Bioregion[[2]]` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r networks_by_bioregion$Bioregion[[3]]` (`r networks_by_bioregion$Bioregion_count[[3]]`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can contribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Karise is a very large study that only sampled bumblebees and
-#is providing unrealistic results for the Boreal region
-#Filter out
-#Maybe add to caption?
-data1 = data %>% 
-filter(!Study_id == "13_Karise")
-
-#Check number of studies by habitat
-habitat = data1 %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data1 %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(EuPPollNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(EuPPollNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 7),
-      axis.text = element_text(size=6),
-      panel.border = element_rect(size=1)) +
-xlab(NULL) +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 10, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = viridis(4),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-      axis.ticks.x = element_line(color="black"),
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 7),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black'),
-plot.margin = margin(0,0,0,0)) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-#Plot panel
-p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Outputs calculated in script 7_3
-#Load lm output connectance ~ latitude
-connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
-connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
-#Load lm output nestedness ~ latitude
-normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
-normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
-#Load values of connectance
-min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
-max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
-mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
-#Load values of nestedness
-min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
-max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
-mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
-#Load kendall corr from connectance and spp
-connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
-connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
-#Load kendall corr from nestedness and spp
-normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
-normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
-#Load z values
-#Save data
-z_percent = readRDS("../Data/Manuscript_info/z_percent.rds")
-no_diff_z_percent = z_percent %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent = z_percent %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent = z_percent %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-
-```
-
-## Network properties
-
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
-We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**).
-Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
-Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
-Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
-Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
-The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
-Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).
-Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
-In addition, the bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}. ", fig.width=10, fig.height=7, fig.align='center'}
 
@@
 
-connectance_species = readRDS("../Data/Manuscript_info/Figure6_connectance_species.rds")
-connectance_latitude = readRDS("../Data/Manuscript_info/Figure6_connectance_latitude.rds")
-nestedness_z_scores = readRDS("../Data/Manuscript_info/nestedness_z_scores.rds")
-nestedness_latitude = readRDS("../Data/Manuscript_info/nestedness_latitude.rds")
+connectance_species = readRDS("../../Data/Manuscript_info/Figure6_connectance_species.rds")
+connectance_latitude = readRDS("../../Data/Manuscript_info/Figure6_connectance_latitude.rds")
+nestedness_z_scores = readRDS("../../Data/Manuscript_info/nestedness_z_scores.rds")
+nestedness_latitude = readRDS("../../Data/Manuscript_info/nestedness_latitude.rds")
 
@@
 ```
-
-\newpage
-
-# 3 | DISCUSSION
-
-EuPPollNet offers the largest set of plant-pollinator networks and plant-pollinator studies compiled to date at European level.
-The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million of interaction records.
-While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (e.g., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-Plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
-This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
-In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require of numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
-
-Bees conducted the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
-For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were especially fly species rich or fly-dominated.
-These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only species rich in the Mediterranean region.
-Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their role as pollinators in this region [@herrera2019; @leon2022].
-The number of butterfly species and interactions were relatively low compared to the other taxa.
-While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
-This is because a large fraction (~40%) of studies did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted in average a third of the total interactions per network.
-This proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights the need of further work evaluating how honey bees and their management impact natural pollinator communities.
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of plants in the database.
-This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
-The less connected networks were the Alpine ones as they were the ones with higher number of species.
-
-#Task: Calcular residual connectance.
-
-with more species had lower connectance .
-
-Finish this paragraph with an idea of how sampling can bias our view and that is important to be aware of that when understanding plant-pollinator patterns
-
-Discuss that other type of informative networks could be useful an increase connectance?
-
-visitation networks
-
-Link with interactions -->
-
-More species less connectance, which increases the number of missing links.
-
-Nestedness
-Pattern across latitude
-
-Missing links?
-
-networks condense information but this workflow allo possibility to explore species phenology... and more
-
-Interactions:
-Sampling area, spp accumulation vs interactions:
-https://doi.org/10.1890/12-0367.1
-Sampling completeness 
- [e.g., @chacoff2012]
-
-Networks, ref:
-https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
-
-**Conclusion paragraph**
-
-foundation for future work
-
-Geographic and taxonomic biases!
-Supporting reference:
-https://doi.org/10.1111/ecog.05926
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-# Acknowledgments
-
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data. This research was funded by the H2020 European project Safeguard (101003476).
-
-# Author contributions
-
-# Supplementary material
-
-**Supplementary text 1**
-
-Habitat definitions:
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-
-2) **Agricultural margings**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an  urban setting. Botanical gardens are included in this category.
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-
-5) **Forest/woodland understory**: Any plant community sampled under a wooded group of plants.
-The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included here agro-forestry areas and open to dense forest.
-Note that we have excluded from this category forest that contains sclerophyllous vegetation.
-
-6) **Semi-natural grassland**: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
-
-7) **Pastures**: Any type of low growing plant community that is highly influenced by human disturbance.
-For instance, agriculture, mowing, moderate to high grazing or urban environments.
-Note that this category also includes old pastures with regrowth of woody vegetation.
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have include in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-
-10) **Natural grassslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
-
-12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S1}.
-Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
-
-#Prepare figure for supplementary material
-#Coverage across high taxonomic ranks of bees, syrphids and butterflies
-#--------------------------------------------------------------------------------------------#
-#Start from bees----
-#Bee families
-bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
-            "Andrenidae", "Colletidae","Melittidae")
-
-#Species per family within SafeNet
-bee_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-filter(Pollinator_family %in% bee_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-bee_spp_europe =  master_list %>% 
-select(Family, Genus, Species, GenusAndSpecies) %>% 
-filter(Family %in% bee_fam) %>% 
-group_by(Family) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-rename(Pollinator_family = Family) %>% 
-mutate(Group = "Europe")
-
-#Bind both datasets for plotting
-bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
-
-p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-ylab("Number of species") +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Bee family coverage")
-#--------------------------------------------------------------------------------------------#
-#Continue with syrphids----
-#As we don't have subfamilies, let's obtained them from the master list
-syrphids = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
-rename(Pollinator_accepted_name = GenusAndSpecies)
-#Syrphid species safenet
-syrphid_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_family== "Syrphidae") %>% 
-distinct()
-#Add subfamily info
-syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
-#Summarise spp per subfamily
-syrphid_spp_safenet = syrphid_spp_safenet %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Get spp per subfamily at european level
-syrphid_spp_europe = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
-#Plot
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Syrphid subfamily coverage")
-#--------------------------------------------------------------------------------------------#
-#Continue with butterflies----
-#Do it at family level
-#Create vector with butterfly families
-butt_fam = c(unique(master_list1$family))
-#Species per family within SafeNet
-butterfly_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Lepidoptera") %>% 
-filter(Pollinator_family %in% butt_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-butterfly_spp_europe =  master_list1 %>% 
-select(family, genus, scientificName) %>% 
-group_by(family) %>% 
-summarise(Spp_number = n_distinct(scientificName)) %>% 
-rename(Pollinator_family = family) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
-#Plot
-p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Butterfly family coverage")
-
-#--------------------------------------------------------------------------------------------#
-p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 14, face = "bold"),
-      axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 12))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S2}.
-Barplot indicating the number of sampling sites by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat_bioregion = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat_bioregion = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-c = bind_rows(habitat, bioregion)
-
-ggplot(c, aes(Habitat_bioregion, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_bw()+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")),
-        space = "free", scales = "free_y") +
-ylab("Sampling sites") +
-xlab(NULL) +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.x = element_line(color="black"),
-      strip.text.x = element_text(size=14, face = "bold"),
-      strip.background = element_rect(size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 12),
-axis.text = element_text(size=10))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="80%", fig.cap= "\\textbf{Figure S3}.
-Association between nestedness (NODFc) and the geometric mean of plan and pollinator species per network", fig.width=3, fig.height=3}
-
-metrics_by_network_spp_number = readRDS("../Data/Working_files/metrics_by_network_spp_number.rds")
-
-metrics_by_network_spp_number %>% 
-ggplot(aes(geometric_mean_spp, Normalised_nestedness)) +
-geom_point(pch = 21, fill = "azure3", size=1.5, stroke=0.25, alpha=1) +
-theme_bw() +
-xlab("Species") +
-ylab("Normalised nestedness (NODFc)") +
-theme(
-    axis.title = element_text(face = "bold", size = 10),
-    axis.text = element_text(size = 8))
-
-```

### 2024-05-23 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 xlab(NULL) +
+ylab("Proportion of species") +
 scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
@@
 Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes.
+Note that residual connectance and normalised nestedness have a moderate negative correlation (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`).
 
@@
 This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
+While animal-centred sampling is likely to increase the plant:pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
 As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
-The less connected networks were the Alpine ones as they were the ones with higher number of species.
-
-#Task: Calcular residual connectance.
-
-with more species had lower connectance .
+Overall, latitude did not explain major changes in residual connectance or normalised nestedness of the plant-pollinator networks.
+Consistent with @olesen2002, residual connectance was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
+Given their correlation, the potential ecological explanations that can help understanding these structural patterns can be similar.
+For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019]; and, the level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
+Therefore, these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones in comparison to networks at higher latitudes.
+In addition, while one third of networks were more nested than expected by chance, two thirds of them did not show statistical difference with null expectations.
+This goes against the common view that all plant-pollinator networks are nested [@bascompte2003], and supports the idea that despite a strong tendency towards it, being nested is not the rule in this type of mutualism.
+
+Despite this database covers a wide range of habitats across Europe, still contains major geographical gaps such as the East of Europe or the Mediterranean region.
+
+Maybe some refs about the above phrase?
+
+Add that
+
+--> Methodology, impact the structure of the network
+
+Finally, these networks have different sampling effort and employ different metodologies which can impact the resulting structure of interactions [REF].
 
 Finish this paragraph with an idea of how sampling can bias our view and that is important to be aware of that when understanding plant-pollinator patterns
-
-Discuss that other type of informative networks could be useful an increase connectance?
-
-visitation networks
+Discuss that other type of informative networks could be useful a increase connectance?
 
@@
 More species less connectance, which increases the number of missing links.
-
-Nestedness
-Pattern across latitude
-
-Missing links?
 

### 2024-05-24 Manuscript/EuPPollNet_Ms.qmd
paragraph 4 more or less on the way now

@@
 The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
 To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses.
@@
 While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
-This is because a large fraction (~40%) of studies did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
+This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
 Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted in average a third of the total interactions per network.
-This proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights the need of further work evaluating how honey bees and their management impact natural pollinator communities.
+This proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights the need of accounting for the potential negative impact of honey bees to wild pollinator communities when managing pollinator communities in Europe [@steffan2000; @herrera2020].
 
@@
 In addition, while one third of networks were more nested than expected by chance, two thirds of them did not show statistical difference with null expectations.
-This goes against the common view that all plant-pollinator networks are nested [@bascompte2003], and supports the idea that despite a strong tendency towards it, being nested is not the rule in this type of mutualism.
-
-Despite this database covers a wide range of habitats across Europe, still contains major geographical gaps such as the East of Europe or the Mediterranean region.
-
-Maybe some refs about the above phrase?
-
-Add that
+This goes against the common view that most plant-pollinator networks are nested [@bascompte2003], and supports the idea that despite a strong tendency towards it, being nested is not the rule fo this large set of plant-pollinator networks.
+
+Despite this database covering a wide range of habitats across `r country_number` European countries, some areas such as Eastern Europe and the Mediterranean region are evidently under-represented.
+This reflects the lack of plant-pollinator community data in highly relevant regions of pollinator diversity [REFs] and could help explain the limited taxonomic coverage found of plant and pollinator species at the European level.
+This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@bennett2018] and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [REFs], land use change [REFs], or the introduction of alien species [@vila2009; @vanbergen2018].
+
+Improve this idea but something like this: -make it more discussion style-
+
+Recording their interactions and understanding plant-pollinator community assembleges now is essential to understand the trends of biodiversity, as shown by the importance of historical or time series datasets.
 

### 2024-05-25 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 
-Despite this database covering a wide range of habitats across `r country_number` European countries, some areas such as Eastern Europe and the Mediterranean region are evidently under-represented.
-This reflects the lack of plant-pollinator community data in highly relevant regions of pollinator diversity [REFs] and could help explain the limited taxonomic coverage found of plant and pollinator species at the European level.
-This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@bennett2018] and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
-Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [REFs], land use change [REFs], or the introduction of alien species [@vila2009; @vanbergen2018].
-
-Improve this idea but something like this: -make it more discussion style-
-
-Recording their interactions and understanding plant-pollinator community assembleges now is essential to understand the trends of biodiversity, as shown by the importance of historical or time series datasets.
+Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
+For instance, most plant-pollinator networks belongs to central Europe, while Eastern Europe and the Mediterranean region are under-represented.
+This is consistent with previous studies which also report lack of plant-pollinator data for those regions [REFs], highlighting that this database reflects existing patterns in data availability despite the absence of a systematic search for studies.
+This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@bennett2018], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+These areas are well known for their rich pollinator diversity, and their under-representation is likely contributing to the low taxonomic coverage of this database.
 
@@
 
-Link with interactions -->
-
-More species less connectance, which increases the number of missing links.
-
-networks condense information but this workflow allo possibility to explore species phenology... and more
-
-Interactions:
-Sampling area, spp accumulation vs interactions:
-https://doi.org/10.1890/12-0367.1
-Sampling completeness 
- [e.g., @chacoff2012]
-
 Networks, ref:
@@
 
+**Maybe for the concluding paragraph**:
+
+Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [REFs], land use change [REFs], or the introduction of alien species [@vila2009; @vanbergen2018].
+
+Improve this idea but something like this: -make it more discussion style-
+
+Recording their interactions and understanding plant-pollinator community assembleges now is essential to understand the trends of biodiversity, as shown by the importance of historical or time series datasets.
+
 **Conclusion paragraph**
-
-foundation for future work
-
-Geographic and taxonomic biases!
-Supporting reference:
-https://doi.org/10.1111/ecog.05926
 

### 2024-05-26 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
-Overall, latitude did not explain major changes in residual connectance or normalised nestedness of the plant-pollinator networks.
-Consistent with @olesen2002, residual connectance was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
-Given their correlation, the potential ecological explanations that can help understanding these structural patterns can be similar.
-For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019]; and, the level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
+Overall, latitude did not explain major changes in corrected connectance or normalised nestedness.
+However, consistent with @olesen2002, corrected connectance was lower at higher latitudes, while normalised nestedness, which is negatively correlated with corrected connectance, increased towards higher latitudes.
+Thus, the potential ecological explanations that can help understanding these associations can be similar.
+For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019].
+Further, the level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
 Therefore, these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones in comparison to networks at higher latitudes.
@@
 This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@bennett2018], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
-These areas are well known for their rich pollinator diversity, and their under-representation is likely contributing to the low taxonomic coverage of this database.
+These areas are well known for their rich pollinator diversity [REFs], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
 

### 2024-05-27 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at European level by combining the recently published checklists of bees [@ghisbain2023], syrphids [@kovcic2023] and butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023] with the one of butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
 
@@
 The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million of interaction records.
-While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (e.g., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
+While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
 Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-Plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
+However, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
 This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
@@
 Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted in average a third of the total interactions per network.
-This proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights the need of accounting for the potential negative impact of honey bees to wild pollinator communities when managing pollinator communities in Europe [@steffan2000; @herrera2020].
+Indeed, this proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights their key role as potential pollinator of many flowering plants.
+However, high honey bee densities can have negative impacts on wild pollinators and the plants they visit [@steffan2000; @geldmann2018; @herrera2020], emphasising the need to evaluate these impacts and implement actions if necessary.
 
@@
 As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
-Overall, latitude did not explain major changes in corrected connectance or normalised nestedness.
-However, consistent with @olesen2002, corrected connectance was lower at higher latitudes, while normalised nestedness, which is negatively correlated with corrected connectance, increased towards higher latitudes.
-Thus, the potential ecological explanations that can help understanding these associations can be similar.
-For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019].
-Further, the level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
-Therefore, these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones in comparison to networks at higher latitudes.
-In addition, while one third of networks were more nested than expected by chance, two thirds of them did not show statistical difference with null expectations.
-This goes against the common view that most plant-pollinator networks are nested [@bascompte2003], and supports the idea that despite a strong tendency towards it, being nested is not the rule fo this large set of plant-pollinator networks.
+Overall, latitude did not explain major changes in residual connectance or normalised nestedness across networks.
+However, consistent with @olesen2002, residual connectance -the deviation from the expected connectance for a given network size- was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
+Thus, giving their association and that both metrics capture structural properties of species interactions, the potential ecological explanations that can help understanding their latitudinal trends can be similar.
+For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
+Therefore, we hypothesise that these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones in comparison to networks at higher latitudes.
+This higher proportion of generalists could result in a more even number of links per species, deviating more from a perfect nested or hierarchical structure of the network.
+Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
+Therefore, while this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks.
 
@@
 For instance, most plant-pollinator networks belongs to central Europe, while Eastern Europe and the Mediterranean region are under-represented.
-This is consistent with previous studies which also report lack of plant-pollinator data for those regions [REFs], highlighting that this database reflects existing patterns in data availability despite the absence of a systematic search for studies.
-This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@bennett2018], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
-These areas are well known for their rich pollinator diversity [REFs], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
+This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
+This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
+Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [REFs], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
+Therefore, continuous monitoring programs of plant-pollinator communities are needed in order to evaluate
+
+are essential to understand temporal and spatial changes in order to preserve their biodiversity.
 
@@
 
-Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [REFs], land use change [REFs], or the introduction of alien species [@vila2009; @vanbergen2018].
-
 Improve this idea but something like this: -make it more discussion style-
@@
 Recording their interactions and understanding plant-pollinator community assembleges now is essential to understand the trends of biodiversity, as shown by the importance of historical or time series datasets.
+
+Hence, this emphasize how this database can help understanding ecological questions from plant-pollinator networks and its potential implications for future research questions.
 

### 2024-05-28 Manuscript/EuPPollNet_Ms.qmd
Discussion

@@
 pull(percent)
+#Load correlation outputs of residual connectance and nestednes
+resid_connectance_nestedness_corr_tau = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_tau.rds")
+resid_connectance_nestedness_corr_pval = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_pval.rds")
 
@@
 Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes.
-Note that residual connectance and normalised nestedness have a moderate negative correlation (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`).
+Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
 
@@
 For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
-Therefore, we hypothesise that these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones in comparison to networks at higher latitudes.
-This higher proportion of generalists could result in a more even number of links per species, deviating more from a perfect nested or hierarchical structure of the network.
+Therefore, we hypothesise that these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
 Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
@@
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [REFs], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
-Therefore, continuous monitoring programs of plant-pollinator communities are needed in order to evaluate
-
-are essential to understand temporal and spatial changes in order to preserve their biodiversity.
-
---> Methodology, impact the structure of the network
-
-Finally, these networks have different sampling effort and employ different metodologies which can impact the resulting structure of interactions [REF].
-
-Finish this paragraph with an idea of how sampling can bias our view and that is important to be aware of that when understanding plant-pollinator patterns
-Discuss that other type of informative networks could be useful a increase connectance?
-
-Networks, ref:
-https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
-
-**Maybe for the concluding paragraph**:
-
-Improve this idea but something like this: -make it more discussion style-
-
-Recording their interactions and understanding plant-pollinator community assembleges now is essential to understand the trends of biodiversity, as shown by the importance of historical or time series datasets.
-
-Hence, this emphasize how this database can help understanding ecological questions from plant-pollinator networks and its potential implications for future research questions.
-
-**Conclusion paragraph**
+Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
+Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
+This will allow local and large scale analyses of the status and trends of plant-pollinator communities, effectively informing management and conservation actions.
+
+In conclusion, the EuPPollNet database enables researchers to explore spatial, taxonomic and structural properties of plant-pollinator networks within Europe.
+Importantly, the networks of the database differ in sampling effort and methodology, which can largely impact the resulting structure of interactions [@rivera2012; @chacoff2012].
+In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control and select the most suitable networks for their research questions.
+Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
+Although further evaluations of these findings are needed, this illustrates the numerous questions that can be answered with this set of networks.
+The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and spatio-temporal trends of plant-pollinator communities.
+Finally, we hope this database becomes an iterative resource that keeps growing and improving over time to better understand and conserve European biodiversity.
 
@@
 
-# References
+# REFERENCES
 
@@
 
-# Acknowledgments
-
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data. This research was funded by the H2020 European project Safeguard (101003476).
-
-# Author contributions
-
-# Supplementary material
+# ACKNOWLEDGEMENTS
+
+We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data.
+
+# FUNDING INFORMATION
+
+This research was funded by the H2020 European project Safeguard (101003476).
+
+# CONFLICT OF INTEREST
+
+None.
+
+# DATA AVAILABILITY
+
+All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (LINK).
+
+\newpage
+
+# SUPPORTING INFORMATION
+
+**Title**: EuPPollNet: A European database of plant-pollinator networks
+
+**Authors**: `r authors`
+
+**Contains**:
+
+- Supplementary text 1
+
+- Figure S1
+
+- Figure S2
+
+- Figure S3
+
+\newpage
 
@@
       strip.background = element_rect(size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 12),
+  axis.title = element_text(face = "bold", size = 14),
 axis.text = element_text(size=10))
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="80%", fig.cap= "\\textbf{Figure S3}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S3}.
 Association between nestedness (NODFc) and the geometric mean of plan and pollinator species per network", fig.width=3, fig.height=3}

### 2024-05-29 Manuscript/EuPPollNet_word.qmd
Minor typo edits

@@
 
+#Calculate mean coverage and deviation
+coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
+mean_coverage = mean(coverage)
+sd_coverage = sd(coverage)
+
 #Main orders percentage
@@
 The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutive history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
+Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
 The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
@@
 Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations.
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow to better control for the ecological context and sampling methods.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allow to being reused and keep building on it over time.
+EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods.
+In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time.
+This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
 We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
@@
 
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) selected unmatched cases and fix orthographic errors; (iii) retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that were still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at European level by combining the recently published checklists of bees [@ghisbain2023], syrphids [@kovcic2023] and butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) selected unmatched cases and fix orthographic errors; (iii) retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
+
+For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023] with the one of butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
 
@@
 For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated with the the phylogenetic signal with the help of the function _phylosig_ from the **phytools** package.
+The potential number of pollinator species at European level was estimated by adding the total number of species of bees, syrphids, and flies from the checklists, along with the extrapolated number of species from other insect pollinators.
+This extrapolated number was estimated by assuming that the sampling coverage of these other insect pollinators is equal to the average coverage across bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+
+Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
+The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
+We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it with the help of the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**; and for plants the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of **rtree** package [@li2023].
 
@@
 Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represents the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrates the extrapolated curve for approximately 3000 sampling sites.
+Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites.
 The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across increasing number of plant species.
+Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
 This last graph uses the same color and shape structure as the ones in the top panel.
 Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
-Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.", fig.width=10, fig.height=6, out.width="115%"}
+Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
+Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
 
@@
 
-Connectance and nestedness did not change across the latitudinal range of studies (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
-However, differences were found for connectance but not nestedness across bioclimatic regions.
-Specifically, the alpine region differed statistically in connectance to the other bioclimatic regions (Tukey test _P_ < 0.01).
-In addition, the atlantic and continental bioclimatic regions also showed statistical meaningful differences in connectance (Tukey test _P_ < 0.01).
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed an negative exponential relationship with the geometric mean of plant and pollinator species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6c**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
-Finally, we found that `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, `r round(no_diff_z_percent, 2)`% showed no statistical difference, and `r round(under_z_percent, 2)`% were less nested than null expectations (**Figure 6d**).
+Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
+Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
+We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**).
+Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
+Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes.
+Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
 
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
-Graphs (a-b) indicate network connectance and nestedness across the latitudinal range of the studies from the database.
-In addition, the bioclimatic region where each network is located is indicated with points of different shapes and colours.
-Graph (c) shows the the association between network connectances and the geometric mean of plant and pollinator species per network.
-Graph (d) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
+Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
+Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
 The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
-Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).", fig.width=10, fig.height=7, fig.align='center'}
+Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).
+Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
+In addition, the bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
 
 # Load libraries
-
-library(psych)
-# Load data
-data = readRDS("../Data/3_Final_data/Interaction_data.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-null_networks = readRDS("../Data/Working_files/null_networks.rds")
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks.rds")
-
-# Prepare data
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-              Month = lubridate::month(Date), 
-              Day = lubridate::day(Date))
-
-# Subset data for studies with only bumblebees
-v = c("13_Karise", "22_Kallnik")
-subsetting = unique(data$Study_id)
-
-long_format = data %>% 
-mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
-mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
-filter(!Study_id %in% v) %>% 
-select(Network_id, Plant_accepted_name, Pollinator_accepted_name)
-
-# Filter networks larger than 4 plants and 4 pollinator species
-v2 = long_format %>% 
-group_by(Network_id) %>% 
-summarise(Poll_spp = n_distinct(Plant_accepted_name),
-          Plant_spp = n_distinct(Pollinator_accepted_name)) %>% 
-ungroup() %>% 
-filter(Poll_spp >= 4 & Plant_spp >= 4) %>% 
-pull(Network_id)
-
-long_format = long_format %>% 
-filter(Network_id %in% v2)
-
-# Filter networks with more links than species
-links = long_format %>% 
-group_by(Network_id, Plant_accepted_name, Pollinator_accepted_name) %>%
-distinct() %>% 
-group_by(Network_id) %>% 
-summarise(links = length(Network_id))
-
-spp = long_format %>% 
-group_by(Network_id) %>% 
-summarise(n_plants = n_distinct(Plant_accepted_name),
-          n_polls = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(n_total = n_plants + n_polls)
-
-vector_to_include = left_join(links, spp)
-
-v3 = vector_to_include %>% 
-filter(links > n_total) %>% 
-pull(Network_id)
-
-long_format = long_format %>% 
-filter(Network_id %in% v3)
-
-# Explore how NODFc changes across the Y-X AXES (latitude-Longitude)
-long_format_coords = data %>% 
-mutate(Network_id = paste0(Study_id, "_", Network_id, "_", Year)) %>% 
-mutate(Network_id = str_replace_all(Network_id, " ","_")) %>% 
-select(Network_id, Latitude, Longitude, Bioregion) %>% 
-distinct()
-
-# Merge data
-metrics_by_network_coords = left_join(metrics_by_network, long_format_coords)
-
-# Connectance plots
-# 1st Latitude
-p1 = metrics_by_network_coords %>% 
-ggplot(aes(Latitude, Connectance)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.15, size=2.5, alpha=1) +
+library(dplyr)
+library(ggplot2)
+library(stringr)
+library(patchwork)
+
+# Load plots generated in scripts 7_2_Nestedness and 7_2_Connectance
+
+connectance_species = readRDS("../Data/Manuscript_info/Figure6_connectance_species.rds")
+connectance_latitude = readRDS("../Data/Manuscript_info/Figure6_connectance_latitude.rds")
+nestedness_z_scores = readRDS("../Data/Manuscript_info/nestedness_z_scores.rds")
+nestedness_latitude = readRDS("../Data/Manuscript_info/nestedness_latitude.rds")
+
+#Prepare first connectance plots
+library(betareg)
+model = betareg(Connectance ~ log_geometric_mean_spp, data = connectance_species)
+p1 = ggplot(connectance_species, aes(x = log_geometric_mean_spp, y = Connectance))+
+geom_point(pch = 21, fill = "azure3", size=2.5) +
+geom_line(aes(y = predict(model, connectance_species)), colour= "black")+
+xlab("log(Species)") +
 theme_bw() +
-ylab("Connectance") +
+ggtitle("(a)")
+
+p3 = connectance_latitude %>% 
+ggplot(aes(Latitude, residual_conectance)) +
+geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
 scale_fill_viridis_d() +
@@
 
-Although Europe...
-
-Geographic and taxonomic biases!
-Supporting reference:
-https://doi.org/10.1111/ecog.05926
-
-Networks, ref:
-
-https://www.biorxiv.org/content/10.1101/2020.04.02.021691v1.abstract
-
-Major changes in NW Europe occurred pre-1990
-https://doi.org/10.1111/ele.12121
-
-EuPPollNet represents the largest database of plant-pollinator interactions compiled to date and is a work in progress database as there are more studies, unsampled habitats, species, etc...
-
-Contextualise European diversity of plant and pollinator species with global biodiveristy
+EuPPollNet offers the largest set of plant-pollinator networks and plant-pollinator studies compiled to date at European level.
+The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million of interaction records.
+While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
+Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
+However, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
+This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
+In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require of numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
+
+Bees conducted the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
+For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were especially fly species rich or fly-dominated.
+These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
+In addition, beetles were only species rich in the Mediterranean region.
+Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their role as pollinators in this region [@herrera2019; @leon2022].
+The number of butterfly species and interactions were relatively low compared to the other taxa.
+While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
+This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
+Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted in average a third of the total interactions per network.
+Indeed, this proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights their key role as potential pollinator of many flowering plants.
+However, high honey bee densities can have negative impacts on wild pollinators and the plants they visit [@steffan2000; @geldmann2018; @herrera2020], emphasising the need to evaluate these impacts and implement actions if necessary.
+
+Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of plants in the database.
+This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
+While animal-centred sampling is likely to increase the plant:pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
+As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
+Overall, latitude did not explain major changes in residual connectance or normalised nestedness across networks.
+However, consistent with @olesen2002, residual connectance -the deviation from the expected connectance for a given network size- was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
+Thus, giving their association and that both metrics capture structural properties of species interactions, the potential ecological explanations that can help understanding their latitudinal trends can be similar.
+For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
+Therefore, we hypothesise that these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
+Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
+Therefore, while this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks.
+
+Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
+For instance, most plant-pollinator networks belongs to central Europe, while Eastern Europe and the Mediterranean region are under-represented.
+This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
+This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
+Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
+Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
+This will allow local and large scale analyses of the status and trends of plant-pollinator communities, effectively informing management and conservation actions.
+
+In conclusion, the EuPPollNet database enables researchers to explore spatial, taxonomic and structural properties of plant-pollinator networks within Europe.
+Importantly, the networks of the database differ in sampling effort and methodology, which can largely impact the resulting structure of interactions [@rivera2012; @chacoff2012].
+In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control and select the most suitable networks for their research questions.
+Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
+Although further evaluations of these findings are needed, this illustrates the numerous questions that can be answered with this set of networks.
+The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and spatio-temporal trends of plant-pollinator communities.
+Finally, we hope this database becomes an iterative resource that keeps growing and improving over time to better understand and conserve European biodiversity.
 
@@
 
-# Acknowledgments
-
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data. This research was funded by the H2020 European project Safeguard (101003476).
-
-# Author contributions
-
-# Supplementary material
+# ACKNOWLEDGEMENTS
+
+We thank all the taxonomists and ecologists that have made this database possible by contributing with their fieldwork data.
+
+# FUNDING INFORMATION
+
+This research was funded by the H2020 European project Safeguard (101003476).
+
+# CONFLICT OF INTEREST
+
+None.
+
+# DATA AVAILABILITY
+
+All data and code to produce this database and manuscript are available at Zenodo (LINK) and Github (LINK).
+
+\newpage
+
+# SUPPORTING INFORMATION
+
+**Title**: EuPPollNet: A European database of plant-pollinator networks
+
+**Authors**: `r authors`
+
+**Contains**:
+
+- Supplementary text 1
+
+- Figure S1
+
+- Figure S2
+
+- Figure S3
+
+\newpage
 
@@
       axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 1))
+      axis.text = element_text(face = "bold", size = 12))
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="80%", fig.cap= "\\textbf{Figure S3}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S3}.
 Association between nestedness (NODFc) and the geometric mean of plan and pollinator species per network", fig.width=3, fig.height=3}
+
+metrics_by_network_spp_number = readRDS("../Data/Working_files/metrics_by_network_spp_number.rds")
 

### 2024-06-03 Manuscript/EuPPollNet_Ms.qmd
edits from co-authors

@@
 mutate(Network_id = paste0(Study_id, Network_id))
+
 #Bee-syrphid Masterlist
@@
     Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
+    Variable == "Plant_original_name" ~ "Plant species name given by the authors",
     Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
@@
     Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
+    Variable == "Pollinator_original_name" ~ "Pollinator species name given by the authors",
     Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
@@
 **Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first three and last are ordered alphabetically*
+*All authors excluding the first four and last are ordered alphabetically*
 
@@
 **Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the intricate network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising its long-term persistence.
+Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
+However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks.
 One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences, while also providing other ecological variables of interest.
-Furthermore, this work provides an in-depth assessment of the taxonomic and sampling coverage of the database at the European level, complemented by analyses of key structural properties in plant-pollinator networks.
-We hope this database can help researchers to: 1) identify taxonomic, ecological, and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change on plant-pollinator networks to guide future conservation planning for both plant and pollinator species.
+Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
+This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
+Furthermore, this work assesses the taxonomic and sampling coverage of the database at the European level, and the key structural properties in plant-pollinator networks.
+We hope this database can facilitate future research aimed to fill in taxonomic, ecological, and geographical data gaps on plant-pollinator interactions, and explore the impacts of global change on plant-pollinator networks to guide future conservation planning for both plant and pollinator species.
 
@@
 
-**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats.
+**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
 The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
@@
 \small
-plant-pollinator networks, plant-pollinator interactions, flowering plants, Angiosperms, pollinators, nestedness, connectance
+plant-pollinator networks, species interactions, flowering plants, Angiosperms, pollinators, nestedness, connectance
 \endgroup
@@
 
-The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
+Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity.
 The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
+Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species status, trends, and conservation [@heberling2021; @zattara2021].
 However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
@@
 Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators.
+While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023] along with occurrence data [], is setting a foundation for the conservation of its flora and their pollinators.
 However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
@@
 Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
-
-Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
+Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need to keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
+
+Here, we present the EuPPollNet database, an acronym derived from European Plant-Pollinator Networks, which contains harmonized information on interaction data of plants and pollinators at European level.
+The pollinator taxonomic groups are composed by the main orders of entomofauna that visit and pollinate flowering plants in Europe.
 These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
@@
 
-This database is the result of one of the working packages of the European project Safeguard.
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
 Data acquisition followed a non-systematic approach.
-First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
+As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
 These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions within the database.
@@
 
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**).
+The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**).
 The different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
@@
 (\\textbf{b}) Number of studies by year within the database.
-(\\textbf{c}) Number of species and interactions across the 4 main pollinator orders at European level."}
+(\\textbf{c}) Number of species and interactions across the four main pollinator orders at European level."}
 
@@
 To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study.
+To build a plant-pollinator network matrix at a site level, users only need to group interactions by plant and pollinator species, site and study.
 Information about habitat type and bioclimatic region are also provided in this file.
@@
 The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
-In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [@kleijn2015].
-To that end, we also calculated the accumulation curve of pollinator species with increasing number plant species.
+In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [e.g., @kleijn2015 for crops].
+To that end, we also calculated the accumulation curve of pollinator species with an increasing number of plant species.
 The rarefied and extrapolated sampling curves were obtained with the help of the **iNEXT** package [@hsieh2016].
@@
 The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
+For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
 
@@
 Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
+Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
 The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
@@
 
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the vaznull function from the package **bipartite** [@dormann2008] that implements the null model from @vazquez2005.
+Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analysis in plant-pollinator networks and only compares each unique network against their randomized versions.
+We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008] that implements the null model from @vazquez2005.
 These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
@@
 The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 2d**).
-In addition, when we explore how common or rare are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 2e-2f**).
+In addition, when we explore how common are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 2e-2f**).
 For instance, most plants (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinators (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) are exclusively found in less than 1% of sampling sites; and the most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are only found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
@@
 
-The database contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+The database contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
 The coverage of the main pollinators groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
@@
 c1$Habitat = factor(c1$Habitat, levels = l2)
+
 #Create barplot on the side
@@
 Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
-Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes.
+Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located in lower latitudes.
 Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
@@
 EuPPollNet offers the largest set of plant-pollinator networks and plant-pollinator studies compiled to date at European level.
-The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million of interaction records.
+The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records.
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-However, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
+This likely reflects the fact that most plant and pollinator species are rare, and geographically restricted.
+For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
+Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
 This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
-In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require of numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
-
-Bees conducted the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
-For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were especially fly species rich or fly-dominated.
+In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
+
+Bees are responsible for the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
+For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
 These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only species rich in the Mediterranean region.
-Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their role as pollinators in this region [@herrera2019; @leon2022].
+In addition, beetles were only common in the Mediterranean region.
+Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
 The number of butterfly species and interactions were relatively low compared to the other taxa.
@@
 This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted in average a third of the total interactions per network.
-Indeed, this proportion is higher that the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights their key role as potential pollinator of many flowering plants.
-However, high honey bee densities can have negative impacts on wild pollinators and the plants they visit [@steffan2000; @geldmann2018; @herrera2020], emphasising the need to evaluate these impacts and implement actions if necessary.
+On the contrary, honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
+Indeed, this proportion is higher than the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights their key role as potential pollinators of many flowering plants.
+However, the rise of beekeeping in Europe has led to high honey bee densities, which can negatively impact wild pollinators and the plants they visit [@steffan2000; @magrach2017; @herrera2020], emphasising the need to evaluate these impacts and implement actions if necessary.
 
@@
 Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, most plant-pollinator networks belongs to central Europe, while Eastern Europe and the Mediterranean region are under-represented.
+For instance, most plant-pollinator networks belong to central Europe, while Eastern Europe and the Mediterranean region are under-represented.
 This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
@@
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
+Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still suffer from major threads such as climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
 Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
@@
 
-2) **Agricultural margings**: Sides of crops that can include any type of vegetation from low growing plants to trees.
+2) **Agricultural margins**: Sides of crops that can include any type of vegetation from low growing plants to trees.
 
@@
 The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included here agro-forestry areas and open to dense forest.
+We have included in this category agro-forestry areas and open to dense forest.
 Note that we have excluded from this category forest that contains sclerophyllous vegetation.
@@
 
-7) **Pastures**: Any type of low growing plant community that is highly influenced by human disturbance.
+7) **Intensive grassland**: Any type of low growing plant community that is highly influenced by human disturbance.
 For instance, agriculture, mowing, moderate to high grazing or urban environments.
@@
 Typical of the Mediterranean region.
-Note, that we have include in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
+Note, that we have included in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
 
@@
 
-10) **Natural grassslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
+10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
 

### 2024-06-03 Manuscript/EuPPollNet_word.qmd
edits from co-authors

@@
 The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy.
+Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity.
 The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021].
+Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species status, trends, and conservation [@heberling2021; @zattara2021].
 However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
@@
 Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
+Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need to keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
 
@@
 
-Bees conducted the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
-For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were especially fly species rich or fly-dominated.
+Bees are responsible for the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
+For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
 These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only species rich in the Mediterranean region.
-Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their role as pollinators in this region [@herrera2019; @leon2022].
+In addition, beetles were only common in the Mediterranean region.
+Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
 The number of butterfly species and interactions were relatively low compared to the other taxa.

### 2024-06-03 Scripts/4_Habitat_information.R

@@ -914,5 +914,9 @@ select(Study_id, Network_id, Authors_habitat, SafeNet_habitat, Latitude,
 check = habitat %>% 
 filter(is.na(SafeNet_habitat))
 
+#Recode to intensive grassland
+habitat = habitat %>%  
+mutate(SafeNet_habitat = recode_factor(SafeNet_habitat, "Pastures" = "Intensive grasslands")) %>% 
+mutate(SafeNet_habitat = recode_factor(SafeNet_habitat, "Semi-natural grassland" = "Semi-natural grasslands"))
 #Save data
 saveRDS(habitat, "Data/Working_files/Habitat.rds")

### 2024-06-04 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023] along with occurrence data [], is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
+While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023], along with occurrence data [@zattara2021], is setting a foundation for the conservation of its flora and their pollinators.
+However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
 Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
 Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
+Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
+For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
+Assembling and curating the existing information on EU plant-pollinator networks will guide research efforts, conservation planning and will set a foundation for future global change research.
 
@@
 Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
+For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
 Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
@@
 
-Here, we present the EuPPollNet database, an acronym derived from European Plant-Pollinator Networks, which contains harmonized information on interaction data of plants and pollinators at European level.
+Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on interaction data of plants and pollinators at European level.
 The pollinator taxonomic groups are composed by the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], which comprise almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
 To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions.
+In addition, we investigate how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across the latitudes range of studies and bioclimatic regions.
 Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations.
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time.
+EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
+In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
 This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
-We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
+We expect that EuPPollNet can be used to evaluate macro-ecological processes in plant-pollinator networks, guide conservation planning, and set a baseline for global change research.
 
@@
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-Data acquisition followed a non-systematic approach.
-As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions within the database.
+As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked to members of the Safeguard project.
+Second, the request was extended to data owners outside of the project.
+These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching studies on Google Scholar of under-represented regions within the database.
 To keep high quality standards that will allow robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
@@
 The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**).
-The different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant.
-Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species.
+The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
+Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a pollinator individual to the reproductive structure of a particular plant.
+Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species in the database.
 However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the EuPPollNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
+(\\textbf{a}) Locations of the studies in EuPPollNet showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
 The sizes of these shapes are proportional to the respective species counts.
 For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Number of studies by year within the database.
-(\\textbf{c}) Number of species and interactions across the four main pollinator orders at European level."}
+(\\textbf{b}) Number of studies by year in EuPPollNet.
+(\\textbf{c}) Number of species and interactions across the four main pollinator orders in EuPPollNet."}
 
@@
 In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study.
-The flower count data is provided in a separate file given that one third of studies lack floral measurements and that the units differ greatly across studies.
+The flower count data is provided in a separate file.
+Two thirds of studies contain information on floral abundance, but the methods and units differ greatly across studies.
 Finally, metadata at the study level is provided, including information about the authors, digital object identifier if available, sampling time and taxonomic coverage of the main pollinator groups for each study.
@@
 
-## Sampling coverage
-
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
-In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [e.g., @kleijn2015 for crops].
-To that end, we also calculated the accumulation curve of pollinator species with an increasing number of plant species.
-The rarefied and extrapolated sampling curves were obtained with the help of the **iNEXT** package [@hsieh2016].
-This was complemented with 100 random accumulation curves obtained programmatically to guide the visualization of the different rarefied curves.
-
 ## Taxonomic harmonization
 
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
+All plant and pollinator species names were checked and harmonized in R using **rgbif** [@chamberlain2022].
 The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
 For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) selected unmatched cases and fix orthographic errors; (iii) retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023] with the one of butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+In addition, taxonomic information at family, order and phylum level was downloaded for each species.
+
+For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
+
+For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023] with the one of butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
 
@@
 
+## Sampling coverage
+
+The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
+In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [e.g., @kleijn2015 for crops].
+To that end, we also calculated the accumulation curve of pollinator species with an increasing number of plant species.
+The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
+This was complemented with 100 random accumulation curves obtained programmatically to guide the visualization of the different rarefied curves.
+
 ## Habitat type and bioclimatic region
@@
 These categories allow a quick comparison and understanding of the habitat types from the database.
-However, we advise authors to revise this classification if they intend to rely on this field for their analyses as this is a non-fully objective process.
 Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
 Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
+The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites using a spatial joint from the **sf** package [@pebesma2018].
 
@@
 
-To provide a general overview of the main aspects of this set of plant-pollinator networks, we evaluated network patterns of connectance and nestedness across the latitudinal range of studies and bioclimatic regions.
+To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness and examined how these network metrics change across the latitudinal range of studies and bioclimatic regions.
 We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
@@
 # 3 | RESULTS
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Sampling coverage outputs
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Coverage
-poll_coverage = poll_output$DataInfo$SC *100
-plant_coverage = plant_output$DataInfo$SC *100
-int_coverage = int_output$DataInfo$SC *100
-plant_poll_coverage = plant_poll_output$DataInfo$SC *100
-
-#Increase doubling effort
-##Pollinators
-poll_increase = poll_output$iNextEst$Network_id
-poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
-poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
-poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
-##Plants
-plant_increase = plant_output$iNextEst$Network_id
-plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
-plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
-##Interactions
-int_increase = int_output$iNextEst$Network_id
-int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
-int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
-int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
-##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$Plant_id
-plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
-plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
-
-#Prepare results from common and rare poll across sampling sites
-#Pollinators
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-poll_total = nrow(poll_spread)
-polls_under_one_percent = poll_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-polls_under_one_percent = polls_under_one_percent/poll_total * 100
-#Find most common pollinator
-most_common_poll = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Pollinator_accepted_name)
-#Find most common poll percent
-most_common_poll_percent = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-#Plants
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-plant_total = nrow(plant_spread)
-plants_under_one_percent = plant_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-plants_under_one_percent = plants_under_one_percent/plant_total * 100  
-#Find most common plant
-most_common_plant = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Plant_accepted_name)
-#Find most common plant percent
-most_common_plant_percent = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-```
-
-## Sampling coverage
-
-The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 2a-2b**).
-The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 2c**).
-Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 2d**).
-In addition, when we explore how common are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 2e-2f**).
-For instance, most plants (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinators (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) are exclusively found in less than 1% of sampling sites; and the most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are only found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites.
-The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
-Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
-Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
-
-#Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
-#Create dataset
-d_poll = poll_output$iNextEst$Network_id
-#Check cols
-d_poll$col = d_poll$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           method = as.character(pollinators_curves$iteration), 
-           qD = pollinators_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(method == "observed")
-d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
-
-#Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col),size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_poll1_point, col = "black", size=2) + 
-ylab("Pollinator species") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000) +
-ggtitle("(a)") +
-theme(plot.margin = margin(r = 12))
-
-#Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
-#Create dataset
-d_plant = plant_output$iNextEst$Network_id
-#Check cols
-d_plant$col = d_plant$method
-
-#Now create tibble with same structure
-a_plant = tibble(t = plant_curves$Study_sampled, 
-           method = as.character(plant_curves$iteration), 
-           qD = plant_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(method == "observed")
-d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
-
-#Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_plant1_point, col = "black", size=2) + 
-ylab("Plant species") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000)+
-ggtitle("(b)")+
-theme(plot.margin = margin(r = 12))
-
-#Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
-#Create dataset
-d_int = int_output$iNextEst$Network_id
-#Check cols
-d_int$col = d_int$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = interactions_curves$Study_sampled, 
-           method = as.character(interactions_curves$iteration), 
-           qD = interactions_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(method == "observed")
-d_int1_lines = d_int1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Interactions") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000) +
-ylim(0, 30000)+
-ggtitle("(c)")
-
-#----------------#
-#Pollinators
-#----------------#
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-o1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.03, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
-xlab("Pollinator species") +
-ylab("Occurrence across sampling sites") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.35) + 
-annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
-annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=200, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3)  +
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
-                ylim=c(0,0.7), clip="off") +
-ggtitle("(e)")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-o2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.03, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
-xlab("Plant species") +
-ylab("Occurrence across sampling sites") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.35) + 
-annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
-annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=150, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3) + 
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
-                ylim=c(0, 0.7), clip="off") +
-ggtitle("(f)")
-
-#Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Create dataset
-d_pp = plant_poll_output$iNextEst$Plant_id
-#Check cols
-d_pp$col = d_pp$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           method = as.character(plant_poll_curves$iteration), 
-           qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(method == "observed")
-d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col)) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_pp1_point, col = "black") + 
-ylab("Pollinator species") +
-xlab("Plant species") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(d)") +
-scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
-theme(legend.position = "none")
-
-#Plot all
-p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
-      axis.title = element_text(size=12, face = "bold"),
-      axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-#a/b
-
-```
-
-\newpage
 
@@
 Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions of the database, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 3**).
+Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
 In addition, the presence or absence of interactions records for bees seems not to follow any sort of phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
-The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
+The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
 Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 3}.
-Phylogenetic representation of bee genera in the EuPPollNet database within the context of all European bee genera.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
+Phylogenetic and taxonomic coverage of the bee genera at European level.
 The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Phylogenetic representation of the plant families in the EuPPollNet database within the context of all flowering plant families occurring in Europe.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
+Phylogenetic and taxonomic coverage of the plant families at European level.
 The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Sampling coverage outputs
+poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
+plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
+int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
+plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
+
+#Coverage
+poll_coverage = poll_output$DataInfo$SC *100
+plant_coverage = plant_output$DataInfo$SC *100
+int_coverage = int_output$DataInfo$SC *100
+plant_poll_coverage = plant_poll_output$DataInfo$SC *100
+
+#Increase doubling effort
+##Pollinators
+poll_increase = poll_output$iNextEst$Network_id
+poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
+poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
+poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
+##Plants
+plant_increase = plant_output$iNextEst$Network_id
+plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
+plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
+plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
+##Interactions
+int_increase = int_output$iNextEst$Network_id
+int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
+int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
+int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
+##Pollinator by plant
+plant_poll_increase = plant_poll_output$iNextEst$Plant_id
+plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
+plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
+plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
+
+#Prepare results from common and rare poll across sampling sites
+#Pollinators
+poll_spread = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
+group_by(Network_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+ungroup() %>% 
+count(Pollinator_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
+#Check number of studies under 1%
+poll_total = nrow(poll_spread)
+polls_under_one_percent = poll_spread %>% 
+filter(Percent_total<1) %>% 
+nrow()
+polls_under_one_percent = polls_under_one_percent/poll_total * 100
+#Find most common pollinator
+most_common_poll = poll_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Pollinator_accepted_name)
+#Find most common poll percent
+most_common_poll_percent = poll_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Percent_total)
+#Plants
+plant_spread = data %>% 
+select(Plant_rank, Network_id, Plant_accepted_name) %>% 
+group_by(Network_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+ungroup() %>% 
+count(Plant_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
+#Check number of studies under 1%
+plant_total = nrow(plant_spread)
+plants_under_one_percent = plant_spread %>% 
+filter(Percent_total<1) %>% 
+nrow()
+plants_under_one_percent = plants_under_one_percent/plant_total * 100  
+#Find most common plant
+most_common_plant = plant_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Plant_accepted_name)
+#Find most common plant percent
+most_common_plant_percent = plant_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Percent_total)
+```
+
+## Sampling coverage
+
+The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
+This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 4a-4b**).
+The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
+However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
+Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
+The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
+In addition, when we explore how common are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 4e-4f**).
+For instance, most plants (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinators (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) are exclusively found in less than 1% of sampling sites; and the most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are only found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
+Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites.
+The solid black points indicate the number of species and interactions contained in the database.
+Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
+This last graph uses the same color and shape structure as the ones in the top panel.
+Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
+Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
+Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
+
+#Pollinators-----
+poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
+pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
+#Create dataset
+d_poll = poll_output$iNextEst$Network_id
+#Check cols
+d_poll$col = d_poll$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = pollinators_curves$Study_sampled, 
+           method = as.character(pollinators_curves$iteration), 
+           qD = pollinators_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_poll1 = bind_rows(d_poll, a_poll)
+d_poll1_point = d_poll1 %>%  filter(method == "observed")
+d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
+
+#Order levels
+d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col),size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_poll1_point, col = "black", size=2) + 
+ylab("Pollinator species") +
+xlab("Sampling sites") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000)+
+ylim(0, 3000) +
+ggtitle("(a)") +
+theme(plot.margin = margin(r = 12))
+
+#Plants-----
+plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
+plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
+#Create dataset
+d_plant = plant_output$iNextEst$Network_id
+#Check cols
+d_plant$col = d_plant$method
+
+#Now create tibble with same structure
+a_plant = tibble(t = plant_curves$Study_sampled, 
+           method = as.character(plant_curves$iteration), 
+           qD = plant_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_plant1 = bind_rows(d_plant, a_plant)
+d_plant1_point = d_plant1 %>%  filter(method == "observed")
+d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
+
+#Order levels
+d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col), size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_plant1_point, col = "black", size=2) + 
+ylab("Plant species") +
+xlab("Sampling sites") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000)+
+ylim(0, 3000)+
+ggtitle("(b)")+
+theme(plot.margin = margin(r = 12))
+
+#Interactions-----
+int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
+interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
+#Create dataset
+d_int = int_output$iNextEst$Network_id
+#Check cols
+d_int$col = d_int$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = interactions_curves$Study_sampled, 
+           method = as.character(interactions_curves$iteration), 
+           qD = interactions_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_int1 = bind_rows(d_int, a_poll)
+d_int1_point = d_int1 %>%  filter(method == "observed")
+d_int1_lines = d_int1 %>%  filter(!method == "observed")
+
+#Plot
+#Order levels
+d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col), size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_int1_point, col = "black", size=2) + 
+ylab("Interactions") +
+xlab("Sampling sites") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000) +
+ylim(0, 30000)+
+ggtitle("(c)")
+
+#----------------#
+#Pollinators
+#----------------#
+#Read data
+data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
+mutate(Network_id = paste0(Study_id, Network_id))  
+#First select cols of rank, accepted name and study ID
+#Then select unique levels by study ID
+#Now sum levels (maximum can be as maximum number of studies)
+poll_spread = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
+group_by(Network_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+ungroup() %>% 
+count(Pollinator_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = poll_spread %>%  
+select(Pollinator_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common pollinator for plotting
+most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
+select(Pollinator_accepted_name) %>% pull()
+
+#Plot
+o1 = ggplot(poll_spread, 
+aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
+geom_bar(stat = "identity", 
+           color = "#D55E00",
+           lwd = 0.03, fill = "#D55E00") +
+ylab("Shared pollinators (%)") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+  axis.ticks.x=element_blank(),
+  axis.ticks.y = element_line(size = 0.2),
+  panel.grid.major.x = element_blank(),
+  panel.grid.minor.y = element_blank(),
+  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
+xlab("Pollinator species") +
+ylab("Occurrence across sampling sites") + 
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+color = "black", size=0.35) + 
+annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
+annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
+annotate(geom="text", x=200, y=-0.05, label="Common spp."
+      ,color="black",  size = 3) +
+annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
+      ,color="black",  size = 3)  +
+scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
+coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
+                ylim=c(0,0.7), clip="off") +
+ggtitle("(e)")
+
+#----------------#
+#Plants
+#----------------#
+plant_spread = data %>% 
+select(Plant_rank, Network_id, Plant_accepted_name) %>% 
+group_by(Network_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+ungroup() %>% 
+count(Plant_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = plant_spread %>%  
+select(Plant_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common plant for plotting
+most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
+select(Plant_accepted_name) %>% pull()
+
+o2 = ggplot(plant_spread, 
+aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
+geom_bar(stat = "identity", 
+           color = "#009E73",
+           lwd = 0.03, fill = "#009E73") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+         axis.ticks.x=element_blank(),
+         axis.ticks.y = element_line(size = 0.2),
+         panel.grid.major.x = element_blank(),
+         panel.grid.minor.y = element_blank(),
+         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
+xlab("Plant species") +
+ylab("Occurrence across sampling sites") + 
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+          color = "black", size=0.35) + 
+annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
+annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
+      ,color="black",  fontface = 'italic', size = 3)+
+annotate(geom="text", x=150, y=-0.05, label="Common spp."
+      ,color="black",  size = 3) +
+annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
+      ,color="black",  size = 3) + 
+scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
+coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
+                ylim=c(0, 0.7), clip="off") +
+ggtitle("(f)")
+
+#Pollinator by plant species
+plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
+plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
+
+#Create dataset
+d_pp = plant_poll_output$iNextEst$Plant_id
+#Check cols
+d_pp$col = d_pp$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
+           method = as.character(plant_poll_curves$iteration), 
+           qD = plant_poll_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_pp1 = bind_rows(d_pp, a_poll)
+d_pp1_point = d_pp1 %>%  filter(method == "observed")
+d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
+
+#Plot
+#Order levels
+d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col)) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_pp1_point, col = "black") + 
+ylab("Pollinator species") +
+xlab("Plant species") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off") +
+ggtitle("(d)") +
+scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
+scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
+theme(legend.position = "none")
+
+#Plot all
+p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
+      axis.title = element_text(size=12, face = "bold"),
+      axis.text = element_text(size=10),
+      plot.title = element_text(size=18))
+
+#a/b
+
+```
+
+\newpage
+
 ## Habitat type and bioclimatic region
 
-The proportion of the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
+The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
 As expected, Hymenoptera was the main taxonomic order on the majority of habitats, exceeded only by Diptera on the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
-Overall, the proportion of Lepidoptera and Coleoptera was low across all habitats but Coleoptera showed a notably increase in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
+Overall, the proportion of Lepidoptera and Coleoptera species was low across all habitats but Coleoptera showed a notably increase in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
 Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
 In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
-Again, Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera seems to be only relevant in Mediterranean regions at European level.
+Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera seems to be only relevant in Mediterranean regions at European level.
 Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
+Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
 The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
@@
 These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only common in the Mediterranean region.
+In addition, beetles were only common floral visitors in the Mediterranean region.
 Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
@@
 
-This research was funded by the H2020 European project Safeguard (101003476).
+This research was funded by the H2020 European project Safeguard (101003476) and by the Federal State of Saxony-Anhalt (MLU-BioDivFund).
 

### 2024-06-05 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 pull()
+
+#Check number of distinct networks
+n_networks = data %>% 
+mutate(Network_id = paste0(Network_id,"_", Year)) %>% 
+distinct(Network_id) %>% 
+nrow(.)
+
 #Check number of distinct sampling sites
@@
 Furthermore, this work assesses the taxonomic and sampling coverage of the database at the European level, and the key structural properties in plant-pollinator networks.
-We hope this database can facilitate future research aimed to fill in taxonomic, ecological, and geographical data gaps on plant-pollinator interactions, and explore the impacts of global change on plant-pollinator networks to guide future conservation planning for both plant and pollinator species.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_sites, big.mark=",")` distinct locations (i.e., plant-pollinator networks), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided.
+We hope this database will facilitate future research that fills taxonomic, ecological, and geographical data gaps on plant-pollinator interactions.
+Also, we believe it will help evaluate patterns and drivers of plant-pollinator network change and guide future conservation planning for both plant and pollinator species.
+
+**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
+In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., family, order and phylum).
 
@@
 
-**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
+**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.
+All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
 
 **Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
+The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphid species at the European level.
 
@@
 However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
+
+The interactions between different plant and pollinator species within a community form complex networks.
+Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
+In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023].
+Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
+For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
+Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness.
+Hence, this emphasizes the importance of evaluating the spatio-temporal coverage of interactions to guide future sampling efforts and accurately understand the processes that shape plant-pollinator interactions.
 
@@
 
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
-In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023].
-Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need to keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
-
 Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on interaction data of plants and pollinators at European level.
@@
 In addition, we investigate how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across the latitudes range of studies and bioclimatic regions.
-Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations.
+Given that this database contains one of the largest sets of plant-pollinator networks collated to date, and considering that the prevalence of nestedness has been evaluated only with a relatively small number of networks [@bascompte2003], we also examine whether plant-pollinator networks are more nested than expected by chance.
 EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
@@
 To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix at a site level, users only need to group interactions by plant and pollinator species, site and study.
+To build a plant-pollinator network matrix within a single flowering season at a site level, users only need to group interactions by plant and pollinator species, site, study and year.
 Information about habitat type and bioclimatic region are also provided in this file.
@@
 For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023] with the one of butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+Coleoptera species names were only checked against the GBIF checklist.
 
@@
 The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it with the help of the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**; and for plants the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of **rtree** package [@li2023].
+We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it using the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**.
+For plants, the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of the **rtree** package [@li2023].
 
@@
 
-The different sites per study were described with a habitat type by the authors.
-As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
-For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
+The different sites were described with a habitat type by the authors of each study.
+We standardized  the habitat type using the author's description and land cover information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022].
 Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the EuPPollNet database (see habitat type definition in supplementary material).
@@
 
-The database contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
 The coverage of the main pollinators groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions of the database, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
-In addition, the presence or absence of interactions records for bees seems not to follow any sort of phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
+Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
+Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
+In addition, the presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
 The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
@@
 
-The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
+The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
 This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 4a-4b**).
@@
 However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
-Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
+When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
 The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
-In addition, when we explore how common are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 4e-4f**).
-For instance, most plants (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinators (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) are exclusively found in less than 1% of sampling sites; and the most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are only found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
+We find that a small portion of plant species and pollinator species are shared across a broad range of sampling sites and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of sampling sites (**Figures 4e-4f**).
+The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
 
@@
 In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
-Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera seems to be only relevant in Mediterranean regions at European level.
+Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera is only relevant in Mediterranean regions at European level.
 Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
@@
 
-EuPPollNet offers the largest set of plant-pollinator networks and plant-pollinator studies compiled to date at European level.
+EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level.
 The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records.
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-This likely reflects the fact that most plant and pollinator species are rare, and geographically restricted.
+This likely reflects the fact that most plant and pollinator species are rare and geographically restricted.
 For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
@@
 While animal-centred sampling is likely to increase the plant:pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
-Overall, latitude did not explain major changes in residual connectance or normalised nestedness across networks.
-However, consistent with @olesen2002, residual connectance -the deviation from the expected connectance for a given network size- was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
-Thus, giving their association and that both metrics capture structural properties of species interactions, the potential ecological explanations that can help understanding their latitudinal trends can be similar.
-For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
-Therefore, we hypothesise that these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
+EXPAND HERE!
+
+Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
+Networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
+These factors should increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
 Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
@@
 This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+This is especially relevant as Eastern Europe currently has vast landscapes of high quality semi-natural grasslands but is experiencing rapid land use change [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still suffer from major threads such as climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
+Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
 Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
@@
 In conclusion, the EuPPollNet database enables researchers to explore spatial, taxonomic and structural properties of plant-pollinator networks within Europe.
-Importantly, the networks of the database differ in sampling effort and methodology, which can largely impact the resulting structure of interactions [@rivera2012; @chacoff2012].
-In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control and select the most suitable networks for their research questions.
+In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
 Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
-Although further evaluations of these findings are needed, this illustrates the numerous questions that can be answered with this set of networks.
-The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and spatio-temporal trends of plant-pollinator communities.
+These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
+The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and trends of plant-pollinator communities.
 Finally, we hope this database becomes an iterative resource that keeps growing and improving over time to better understand and conserve European biodiversity.

### 2024-06-06 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 
-Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on interaction data of plants and pollinators at European level.
+Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on interaction data of plants and pollinators at the European level.
 The pollinator taxonomic groups are composed by the main orders of entomofauna that visit and pollinate flowering plants in Europe.
@@
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked to members of the Safeguard project.
+As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked from members of the Safeguard project.
 Second, the request was extended to data owners outside of the project.
-These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching studies on Google Scholar of under-represented regions within the database.
+These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
 To keep high quality standards that will allow robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
@@
 While animal-centred sampling is likely to increase the plant:pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-EXPAND HERE!
+In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
+Rare pollinators have been proven to be functionally important for plant species at large scales [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to observe these rare species and to effectively understand and protect plant-pollinator biodiversity.
 
@@
 Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
-Therefore, while this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks.
+Therefore, while this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks [@payrato2019].
 

### 2024-06-07 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, we investigate how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across the latitudes range of studies and bioclimatic regions.
+The overall number of potential pollinators and flowering plants in Europe was estimated with the help of the different checklists and sampling coverage.
+Finally, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
+
+JUSTIFY WITH A SENTENCE HERE:
+Check that the poll and flower number info is in results! and the phylo signal too.
+
+In addition, we investigate how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across the latitude range of studies and bioclimatic regions.
 Given that this database contains one of the largest sets of plant-pollinator networks collated to date, and considering that the prevalence of nestedness has been evaluated only with a relatively small number of networks [@bascompte2003], we also examine whether plant-pollinator networks are more nested than expected by chance.
@@
 The file contains a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
-To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
+Each row depicts a single interaction between a plant and a pollinator species.
 To build a plant-pollinator network matrix within a single flowering season at a site level, users only need to group interactions by plant and pollinator species, site, study and year.
@@
 In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study.
-The flower count data is provided in a separate file.
+The flower count data is provided in a separate file (.csv or .rds) that can be merged with the interaction data through the "Flower_data_merger" column.
 Two thirds of studies contain information on floral abundance, but the methods and units differ greatly across studies.
@@
 
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023] with the one of butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023], and butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
 Coleoptera species names were only checked against the GBIF checklist.
@@
 
-To assess the completeness of plant and pollinator species in the EuPPollNet database at European level, we used the aforementioned checklists for plants and pollinators.
+To assess the completeness of plant and pollinator species in the EuPPollNet database at the European level, we used the aforementioned checklists for plants and pollinators.
 Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
@@
 Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-The potential number of pollinator species at European level was estimated by adding the total number of species of bees, syrphids, and flies from the checklists, along with the extrapolated number of species from other insect pollinators.
+For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at the European level.
+The potential number of pollinator species at the European level was estimated by adding the total number of species of bees, syrphids, and flies from the checklists, along with the extrapolated number of species from other insect pollinators.
 This extrapolated number was estimated by assuming that the sampling coverage of these other insect pollinators is equal to the average coverage across bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
@@
 
-The different sites were described with a habitat type by the authors of each study.
-We standardized  the habitat type using the author's description and land cover information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022].
-Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the EuPPollNet database (see habitat type definition in supplementary material).
-These categories allow a quick comparison and understanding of the habitat types from the database.
+We describe the habitat type for each site using information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022], visual inspection of Google Earth imagery and the habitat classification from the authors.
+These different habitat categories (see definitions in supplementary text) allow a quick comparison and understanding of the habitat types from the database.
 Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
+Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned a biogeographical region to each site.
 The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites using a spatial joint from the **sf** package [@pebesma2018].
@@
 We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
-However, these network metrics are not independent of sampling effort (e.g., number of species) and to allow comparisons across them, we implemented “standardised” versions of connectance and nestedness.
+We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics.
 As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-This was done with the help of a Beta regression implemented with the package **betareg** [@cribari2010].
-Second, we implemented a metric of nedstedness (i.e., NODFc) that allows the comparison across networks as it corrects by connectance and the number of species [@song2017].
-This metric was calculated with the help of the **maxnodf** package [@hoeppke2021], which employs a computational eﬀicient approach to calculate it.
+The relationship between residual connectance and species richness was investigated using a beta-regression implemented with the package **betareg** [@cribari2010].
+Second, we used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
+This metric was calculated using the **maxnodf** package [@hoeppke2021].
 Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, we explored the association between connectance and nestedness with the number of species per network by conducting Kendall rank correlation coefficient in order to be able to compare the strength of this association across both network metrics.
-
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analysis in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008] that implements the null model from @vazquez2005.
+In addition, we explored the association between connectance and nestedness with the number of species per network using a Kendall rank correlation coefficient to compare the strength of associations between network structures and species richness across both network metrics.
+
+Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
+The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
+We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008].
 These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
@@
 As expected, Hymenoptera was the main taxonomic order on the majority of habitats, exceeded only by Diptera on the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
-Overall, the proportion of Lepidoptera and Coleoptera species was low across all habitats but Coleoptera showed a notably increase in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
+Overall, the proportion of Lepidoptera and Coleoptera species was low across all habitats but Coleoptera were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
 Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
@@
 
-Bees are responsible for the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
+Bees are responsible for the majority of the interactions at the metaweb level, but their relevance for network topology changed across habitats and bioclimatic regions.
 For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
@@
 On the contrary, honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-Indeed, this proportion is higher than the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights their key role as potential pollinators of many flowering plants.
-However, the rise of beekeeping in Europe has led to high honey bee densities, which can negatively impact wild pollinators and the plants they visit [@steffan2000; @magrach2017; @herrera2020], emphasising the need to evaluate these impacts and implement actions if necessary.
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of plants in the database.
+The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or the rise of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
+
+Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species in the database was almost double that of the plants.
 This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centred sampling is likely to increase the plant:pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
+While animal-centerred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
 In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
-Rare pollinators have been proven to be functionally important for plant species at large scales [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to observe these rare species and to effectively understand and protect plant-pollinator biodiversity.
-
-Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
+Rare pollinators are functionally important for plant species at large scales [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to observe these rare species and to effectively understand and protect plant-pollinator biodiversity.
+
+Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
 Networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
+These factors should increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
 Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
@@
 Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, most plant-pollinator networks belong to central Europe, while Eastern Europe and the Mediterranean region are under-represented.
+For instance, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
 This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-This is especially relevant as Eastern Europe currently has vast landscapes of high quality semi-natural grasslands but is experiencing rapid land use change [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+This lack of data is especially relevant for Eastern Europe which has vast landscapes of high quality semi-natural grasslands but is experiencing rapid land use change [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.

### 2024-06-10 Manuscript/EuPPollNet_Ms.qmd
Edits

@@
         \captionsetup[table]{labelformat=empty}
-
+        \usepackage{lineno}
+        \linenumbers
 bibliography: references.bib
@@
 This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
-Furthermore, this work assesses the taxonomic and sampling coverage of the database at the European level, and the key structural properties in plant-pollinator networks.
-We hope this database will facilitate future research that fills taxonomic, ecological, and geographical data gaps on plant-pollinator interactions.
-Also, we believe it will help evaluate patterns and drivers of plant-pollinator network change and guide future conservation planning for both plant and pollinator species.
+We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
+We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified.
+Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.
 
@@
 **Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
-The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
+The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
 
@@
 The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
-In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023].
-Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
+Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
+Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
+Although macro-ecological approaches make significant contributions to knowledge, such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @trojelsgaard2013].
 For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
 Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness.
-Hence, this emphasizes the importance of evaluating the spatio-temporal coverage of interactions to guide future sampling efforts and accurately understand the processes that shape plant-pollinator interactions.
+This emphasizes the importance of providing data in its rawest possible form in datasets that will be utilized for synthesis and macro-ecological studies.
 
@@
 
-Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on interaction data of plants and pollinators at the European level.
-The pollinator taxonomic groups are composed by the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], which comprise almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses.
+Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on plant-pollinator interactions at the European level.
+The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
+These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
 The overall number of potential pollinators and flowering plants in Europe was estimated with the help of the different checklists and sampling coverage.
@@
 
-JUSTIFY WITH A SENTENCE HERE:
-Check that the poll and flower number info is in results! and the phylo signal too.
-
-In addition, we investigate how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across the latitude range of studies and bioclimatic regions.
-Given that this database contains one of the largest sets of plant-pollinator networks collated to date, and considering that the prevalence of nestedness has been evaluated only with a relatively small number of networks [@bascompte2003], we also examine whether plant-pollinator networks are more nested than expected by chance.
+We investigated how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across latitudes and bioclimatic regions.
+Networks that are more nested than expected by chance exhibit greater redundancy in their ecological functions and more stability in the face of species loss.
+EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of nestedness, which has only been evaluated with a relatively small number of networks [@bascompte2003; @payrato2019].
 EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
@@
 These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
-To keep high quality standards that will allow robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
+To maintain high quality standards that will support robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
 
@@
 The EuPPollNet database is available in both .csv and .rds formats.
-The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
-The file contains a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
+The files contain a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
 Each row depicts a single interaction between a plant and a pollinator species.
@@
 For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at the European level.
-The potential number of pollinator species at the European level was estimated by adding the total number of species of bees, syrphids, and flies from the checklists, along with the extrapolated number of species from other insect pollinators.
-This extrapolated number was estimated by assuming that the sampling coverage of these other insect pollinators is equal to the average coverage across bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+The number of pollinator species at the European level was estimated by first estimating coverage for the three taxa that have checklists (bees, syrphids and butterflies).
+The mean coverage across these three groups (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`) was then used for the other taxa for which there is no checklist to extrapolate the number of pollinating species in Europe.
 
@@
 
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
+The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks.
 In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [e.g., @kleijn2015 for crops].
@@
 The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
-This was complemented with 100 random accumulation curves obtained programmatically to guide the visualization of the different rarefied curves.
+The different rarefied curves were complemented with 100 boostrapped accumulation curves.
 
@@
 
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness and examined how these network metrics change across the latitudinal range of studies and bioclimatic regions.
+To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network, and examined how these network metrics change across latitude of studies and bioclimatic regions in Europe.
 We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
@@
 The relationship between residual connectance and species richness was investigated using a beta-regression implemented with the package **betareg** [@cribari2010].
-Second, we used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
+We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
 This metric was calculated using the **maxnodf** package [@hoeppke2021].
 Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, we explored the association between connectance and nestedness with the number of species per network using a Kendall rank correlation coefficient to compare the strength of associations between network structures and species richness across both network metrics.
+In addition, we quantified the association between connectance and nestedness with the number of species per network using a Kendall rank correlation coefficient to compare the strength of associations between network structures and species richness across both network metrics.
 
@@
 
+Europe contains approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination.
 EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
@@
 Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
-In addition, the presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
+The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
 The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
+Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
 
@@
 
-The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 4a-4b**).
+The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_coverage)`% for both taxonomic groups.
+This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**).
 The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
@@
 The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
-We find that a small portion of plant species and pollinator species are shared across a broad range of sampling sites and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of sampling sites (**Figures 4e-4f**).
-The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
+We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**).
+The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.
 
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across networks.
+Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
 The solid black points indicate the number of species and interactions contained in the database.
@@
 This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
-Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
+Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks.
+Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network.
 Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
@@
 ylab("Pollinator species") +
-xlab("Sampling sites") +
+xlab("Networks") +
 theme_bw() +
@@
 ylab("Plant species") +
-xlab("Sampling sites") +
+xlab("Networks") +
 theme_bw() +
@@
 ylab("Interactions") +
-xlab("Sampling sites") +
+xlab("Networks") +
 theme_bw() +
@@
 xlab("Pollinator species") +
-ylab("Occurrence across sampling sites") + 
+ylab("Occurrence across networks") + 
 geom_vline(xintercept = quantile_50, linetype="dashed", 
@@
 xlab("Plant species") +
-ylab("Occurrence across sampling sites") + 
+ylab("Occurrence across networks") + 
 geom_vline(xintercept = quantile_50, linetype="dashed", 
@@
 However, the habitats that contain a higher number of sampling sites are `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies are `r studies_by_bioregion$Bioregion[[1]]` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r studies_by_bioregion$Bioregion[[2]]` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r studies_by_bioregion$Bioregion[[3]]` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r networks_by_bioregion$Bioregion[[1]]` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r networks_by_bioregion$Bioregion[[2]]` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r networks_by_bioregion$Bioregion[[3]]` (`r networks_by_bioregion$Bioregion_count[[3]]`).
+The bioclimatic regions with a higher number of studies are `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
 
@@
 This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-On the contrary, honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or the rise of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species in the database was almost double that of the plants.
+Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
+The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practice in Europe [@steffan2000; @magrach2017; @herrera2020].
+
+Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio), the observed number of pollinator species in the database was almost double that of the plants.
 This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
@@
 Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
-Networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
-Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
-Therefore, while this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks [@payrato2019].
+Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
+These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes.
+Finally, one third of networks were more nested than expected by chance.
+While this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result also suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks [@payrato2019].
 
@@
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
+Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], and the introduction of alien species [@vila2009; @vanbergen2018].
 Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
@@
 
-All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (LINK).
+All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
 
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S2}.
-Barplot indicating the number of sampling sites by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
+Barplot indicating the number of networks by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
 
@@
         space = "free", scales = "free_y") +
-ylab("Sampling sites") +
+ylab("Networks") +
 xlab(NULL) +

### 2024-06-14 Manuscript/EuPPollNet_Ms.qmd
updating authors

@@
 Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
-Although macro-ecological approaches make significant contributions to knowledge, such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @trojelsgaard2013].
+Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @trojelsgaard2013].
 For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
@@
 Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
+Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level.
 For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
@@
 To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
-The overall number of potential pollinators and flowering plants in Europe was estimated with the help of the different checklists and sampling coverage.
-Finally, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
-
-We investigated how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across latitudes and bioclimatic regions.
-Networks that are more nested than expected by chance exhibit greater redundancy in their ecological functions and more stability in the face of species loss.
-EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of nestedness, which has only been evaluated with a relatively small number of networks [@bascompte2003; @payrato2019].
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
+In addition, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
+
+EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks.
+For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where generalist species interact with all other species but specialists interact preferentially only with generalists, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
+
+Overall, EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
 In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
@@
 
-Europe contains approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination.
+Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination.
 EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
@@
 
+#Filter out 2 bioregions that only have a small study with data of a single day
+data = data %>% 
+filter(!Bioregion == "Steppic") %>% 
+filter(!Bioregion == "Pannonian")
+
 #Prepare orders for filtering
@@
 Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practice in Europe [@steffan2000; @magrach2017; @herrera2020].
+The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of intensive grasslands habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practiced in Europe [@steffan2000; @magrach2017; @herrera2020].
 

### 2024-06-14 Manuscript/EuPPollNet_word.qmd
updating authors

@@
 **Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the intricate network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising its long-term persistence.
+Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
+However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks.
 One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences, while also providing other ecological variables of interest.
-Furthermore, this work provides an in-depth assessment of the taxonomic and sampling coverage of the database at the European level, complemented by analyses of key structural properties in plant-pollinator networks.
-We hope this database can help researchers to: 1) identify taxonomic, ecological, and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change on plant-pollinator networks to guide future conservation planning for both plant and pollinator species.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_sites, big.mark=",")` distinct locations (i.e., plant-pollinator networks), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided.
+Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
+This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
+We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
+We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified.
+Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.
+
+**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
+In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., family, order and phylum).
 
@@
 
-**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
+**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.
+All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
 
 **Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.
-
-**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats.
-The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.
+The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphid species at the European level.
+
+**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
+The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
 
@@
 
-The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth.
 Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity.
@@
 
+The interactions between different plant and pollinator species within a community form complex networks.
+Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
+Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
+Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @trojelsgaard2013].
+For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
+Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness.
+This emphasizes the importance of providing data in its rawest possible form in datasets that will be utilized for synthesis and macro-ecological studies.
+
 Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016].
+While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023], along with occurrence data [@zattara2021], is setting a foundation for the conservation of its flora and their pollinators.
+However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
 Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
 Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
-In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023].
-Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016].
-Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012].
-Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need to keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.
-
-Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level.
-The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions.
-Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations.
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time.
+Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
+For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
+Assembling and curating the existing information on EU plant-pollinator networks will guide research efforts, conservation planning and will set a foundation for future global change research.
+
+Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on plant-pollinator interactions at the European level.
+The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
+These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
+The overall number of potential pollinators and flowering plants in Europe was estimated with the help of the different checklists and sampling coverage.
+Finally, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
+
+We investigated how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across latitudes and bioclimatic regions.
+Networks that are more nested than expected by chance exhibit greater redundancy in their ecological functions and more stability in the face of species loss [@thebault2010].
+EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of nestedness, which has only been evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019].
+EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
+In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
 This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
-We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.
+We expect that EuPPollNet can be used to evaluate macro-ecological processes in plant-pollinator networks, guide conservation planning, and set a baseline for global change research.
 
@@
 
-This database is the result of one of the working packages of the European project Safeguard.
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-Data acquisition followed a non-systematic approach.
-First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project.
-These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions within the database.
-To keep high quality standards that will allow robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
+As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked from members of the Safeguard project.
+Second, the request was extended to data owners outside of the project.
+These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
+To maintain high quality standards that will support robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
 
@@
 
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**).
-The different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant.
-Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species.
+The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**).
+The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
+Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a pollinator individual to the reproductive structure of a particular plant.
+Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species in the database.
 However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Approximate locations of the studies of the EuPPollNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
+(\\textbf{a}) Locations of the studies in EuPPollNet showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
 The sizes of these shapes are proportional to the respective species counts.
 For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
-(\\textbf{b}) Number of studies by year within the database.
-(\\textbf{c}) Number of species and interactions across the 4 main pollinator orders at European level."}
+(\\textbf{b}) Number of studies by year in EuPPollNet.
+(\\textbf{c}) Number of species and interactions across the four main pollinator orders in EuPPollNet."}
 
@@
 The EuPPollNet database is available in both .csv and .rds formats.
-The .csv file contains data in long format, while the .rds file stores the data in a list structure, with networks organized within each study.
-The file contains a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
-To provide a standard unit of plant-pollinator interaction across studies, each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix at a site level, authors only need to group interactions by plant and pollinator species, site and study.
+The files contain a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
+Each row depicts a single interaction between a plant and a pollinator species.
+To build a plant-pollinator network matrix within a single flowering season at a site level, users only need to group interactions by plant and pollinator species, site, study and year.
 Information about habitat type and bioclimatic region are also provided in this file.
 In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study.
-The flower count data is provided in a separate file given that one third of studies lack floral measurements and that the units differ greatly across studies.
+The flower count data is provided in a separate file (.csv or .rds) that can be merged with the interaction data through the "Flower_data_merger" column.
+Two thirds of studies contain information on floral abundance, but the methods and units differ greatly across studies.
 Finally, metadata at the study level is provided, including information about the authors, digital object identifier if available, sampling time and taxonomic coverage of the main pollinator groups for each study.
@@
 
-## Sampling coverage
-
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different sampling sites.
-In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [@kleijn2015].
-To that end, we also calculated the accumulation curve of pollinator species with increasing number plant species.
-The rarefied and extrapolated sampling curves were obtained with the help of the **iNEXT** package [@hsieh2016].
-This was complemented with 100 random accumulation curves obtained programmatically to guide the visualization of the different rarefied curves.
-
 ## Taxonomic harmonization
 
-All plant and pollinator species names were checked and standardized according to large scale taxonomic databases.
-To ensure reproducibility of the workflow, we have conducted this harmonization in R with **rgbif** [@chamberlain2022] as pivotal package to check for species names and retrieve further taxonomic information (i.e., phylum, order, family and genus) from the Global Biodiversity Information Facility (GBIF).
+All plant and pollinator species names were checked and harmonized in R using **rgbif** [@chamberlain2022].
 The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the old species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) selected unmatched cases and fix orthographic errors; (iii) retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023] with the one of butterflies [@wiemers2018]; (ii) then, we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we programmatically fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
+In addition, taxonomic information at family, order and phylum level was downloaded for each species.
+
+For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
+
+For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023], and butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
+Coleoptera species names were only checked against the GBIF checklist.
 
@@
 Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at European level.
-The potential number of pollinator species at European level was estimated by adding the total number of species of bees, syrphids, and flies from the checklists, along with the extrapolated number of species from other insect pollinators.
-This extrapolated number was estimated by assuming that the sampling coverage of these other insect pollinators is equal to the average coverage across bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at the European level.
+The number of pollinator species at the European level was estimated by first estimating coverage for the three taxa that have checklists (bees, syrphids and butterflies).
+The mean coverage across these three groups (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`) was then used for the other taxa for which there is no checklist to extrapolate the number of pollinating species in Europe.
 
@@
 The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it with the help of the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**; and for plants the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of **rtree** package [@li2023].
+We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it using the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**.
+For plants, the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of the **rtree** package [@li2023].
+
+## Sampling coverage
+
+The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks.
+In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [e.g., @kleijn2015 for crops].
+To that end, we also calculated the accumulation curve of pollinator species with an increasing number of plant species.
+The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
+The different rarefied curves were complemented with 100 boostrapped accumulation curves.
 
@@
 
-The different sites per study were described with a habitat type by the authors.
-As these habitats are not standardized across studies, they were standardized with the additional help of land cover information and visual checks on current satellite imagery.
-For each georeferenced site, the land cover information was extracted from Corine Land Cover (CLC, version 2018) with the help of the Terra package [@hijmans2022].
-Based on the habitat classification from the authors and the CLC classification, we created habitat categories that intend to summarize the diversity of habitats in the EuPPollNet database (see habitat type definition in supplementary material).
-These categories allow a quick comparison and understanding of the habitat types from the database.
-However, we advise authors to revise this classification if they intend to rely on this field for their analyses as this is a non-fully objective process.
+We describe the habitat type for each site using information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022], visual inspection of Google Earth imagery and the habitat classification from the authors.
+These different habitat categories (see definitions in supplementary text) allow a quick comparison and understanding of the habitat types from the database.
 Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore set of studies that share similar environmental conditions and species, we assigned to each site a biogeographical region.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites with the help of a spatial joint from the **sf** package [@pebesma2018].
+Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned a biogeographical region to each site.
+The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites using a spatial joint from the **sf** package [@pebesma2018].
 
@@
 
-To provide a general overview of the main aspects of this set of plant-pollinator networks, we evaluated network patterns of connectance and nestedness across the latitudinal range of studies and bioclimatic regions.
+To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network, and examined how these network metrics change across latitude of studies and bioclimatic regions in Europe.
 We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
-However, these network metrics are not independent of sampling effort (e.g., number of species) and to allow comparisons across them, we implemented “standardised” versions of connectance and nestedness.
+We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics.
 As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-This was done with the help of a Beta regression implemented with the package **betareg** [@cribari2010].
-Second, we implemented a metric of nedstedness (i.e., NODFc) that allows the comparison across networks as it corrects by connectance and the number of species [@song2017].
-This metric was calculated with the help of the **maxnodf** package [@hoeppke2021], which employs a computational eﬀicient approach to calculate it.
+The relationship between residual connectance and species richness was investigated using a beta-regression implemented with the package **betareg** [@cribari2010].
+We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
+This metric was calculated using the **maxnodf** package [@hoeppke2021].
 Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, we explored the association between connectance and nestedness with the number of species per network by conducting Kendall rank correlation coefficient in order to be able to compare the strength of this association across both network metrics.
-
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008, as this approach allow us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the vaznull function from the package **bipartite** [@dormann2008] that implements the null model from @vazquez2005.
+In addition, we quantified the association between connectance and nestedness with the number of species per network using a Kendall rank correlation coefficient to compare the strength of associations between network structures and species richness across both network metrics.
+
+Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
+The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
+We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008].
 These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
@@
 # 3 | RESULTS
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Sampling coverage outputs
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Coverage
-poll_coverage = poll_output$DataInfo$SC *100
-plant_coverage = plant_output$DataInfo$SC *100
-int_coverage = int_output$DataInfo$SC *100
-plant_poll_coverage = plant_poll_output$DataInfo$SC *100
-
-#Increase doubling effort
-##Pollinators
-poll_increase = poll_output$iNextEst$Network_id
-poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
-poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
-poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
-##Plants
-plant_increase = plant_output$iNextEst$Network_id
-plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
-plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
-##Interactions
-int_increase = int_output$iNextEst$Network_id
-int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
-int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
-int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
-##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$Plant_id
-plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
-plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
-
-#Prepare results from common and rare poll across sampling sites
-#Pollinators
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-poll_total = nrow(poll_spread)
-polls_under_one_percent = poll_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-polls_under_one_percent = polls_under_one_percent/poll_total * 100
-#Find most common pollinator
-most_common_poll = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Pollinator_accepted_name)
-#Find most common poll percent
-most_common_poll_percent = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-#Plants
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-plant_total = nrow(plant_spread)
-plants_under_one_percent = plant_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-plants_under_one_percent = plants_under_one_percent/plant_total * 100  
-#Find most common plant
-most_common_plant = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Plant_accepted_name)
-#Find most common plant percent
-most_common_plant_percent = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-```
-
-## Sampling coverage
-
-The estimated sampling coverage of plant and pollinator species within the database across the different sampling sites is approximately `r round(poll_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of sampling sites or networks (**Figure 2a-2b**).
-The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 2c**).
-Similarly, to the plant and pollinator species accumulation curves across sampling sites, when we consider the accumulated pollinator richness across sampled plant species, this curve also shows a "quasi-asymptotic" growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 2d**).
-In addition, when we explore how common or rare are plant and pollinator species across all sampling sites, we find that the majority of plant and pollinator species tend to be regionally specific, and only a minor portion of them is being shared across a broad range of sampling sites (**Figures 2e-2f**).
-For instance, most plants (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinators (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) are exclusively found in less than 1% of sampling sites; and the most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`*) species are only found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of sampling sites, respectively.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across sampling sites.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 sampling sites.
-The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across sampling sites.
-Species on the left (i.e., common) are found in many sampling sites, while species on the right (i.e., rare) are found in few or only a single sampling site.
-Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
-
-#Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
-#Create dataset
-d_poll = poll_output$iNextEst$Network_id
-#Check cols
-d_poll$col = d_poll$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           method = as.character(pollinators_curves$iteration), 
-           qD = pollinators_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(method == "observed")
-d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
-
-#Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col),size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_poll1_point, col = "black", size=2) + 
-ylab("Pollinator species") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000) +
-ggtitle("(a)") +
-theme(plot.margin = margin(r = 12))
-
-#Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
-#Create dataset
-d_plant = plant_output$iNextEst$Network_id
-#Check cols
-d_plant$col = d_plant$method
-
-#Now create tibble with same structure
-a_plant = tibble(t = plant_curves$Study_sampled, 
-           method = as.character(plant_curves$iteration), 
-           qD = plant_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(method == "observed")
-d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
-
-#Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_plant1_point, col = "black", size=2) + 
-ylab("Plant species") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000)+
-ggtitle("(b)")+
-theme(plot.margin = margin(r = 12))
-
-#Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
-#Create dataset
-d_int = int_output$iNextEst$Network_id
-#Check cols
-d_int$col = d_int$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = interactions_curves$Study_sampled, 
-           method = as.character(interactions_curves$iteration), 
-           qD = interactions_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(method == "observed")
-d_int1_lines = d_int1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Interactions") +
-xlab("Sampling sites") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000) +
-ylim(0, 30000)+
-ggtitle("(c)")
-
-#----------------#
-#Pollinators
-#----------------#
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-o1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.03, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
-xlab("Pollinator species") +
-ylab("Occurrence across sampling sites") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.35) + 
-annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
-annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=200, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3)  +
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
-                ylim=c(0,0.7), clip="off") +
-ggtitle("(e)")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-o2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.03, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
-xlab("Plant species") +
-ylab("Occurrence across sampling sites") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.35) + 
-annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
-annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=150, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3) + 
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
-                ylim=c(0, 0.7), clip="off") +
-ggtitle("(f)")
-
-#Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Create dataset
-d_pp = plant_poll_output$iNextEst$Plant_id
-#Check cols
-d_pp$col = d_pp$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           method = as.character(plant_poll_curves$iteration), 
-           qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(method == "observed")
-d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col)) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_pp1_point, col = "black") + 
-ylab("Pollinator species") +
-xlab("Plant species") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(d)") +
-scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
-theme(legend.position = "none")
-
-#Plot all
-p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
-      axis.title = element_text(size=12, face = "bold"),
-      axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-#a/b
-
-```
-
-\newpage
 
@@
 
-The database contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
+Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination.
+EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
 The coverage of the main pollinators groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions of the database, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 3**).
-In addition, the presence or absence of interactions records for bees seems not to follow any sort of phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
-The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 4**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also seems not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 3}.
-Phylogenetic representation of bee genera in the EuPPollNet database within the context of all European bee genera.
+Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
+Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
+The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
+The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
+Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
+Phylogenetic and taxonomic coverage of the bee genera at European level.
 The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
@@
 
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+
+#Sampling coverage outputs
+poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
+plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
+int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
+plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
+
+#Coverage
+poll_coverage = poll_output$DataInfo$SC *100
+plant_coverage = plant_output$DataInfo$SC *100
+int_coverage = int_output$DataInfo$SC *100
+plant_poll_coverage = plant_poll_output$DataInfo$SC *100
+
+#Increase doubling effort
+##Pollinators
+poll_increase = poll_output$iNextEst$Network_id
+poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
+poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
+poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
+##Plants
+plant_increase = plant_output$iNextEst$Network_id
+plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
+plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
+plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
+##Interactions
+int_increase = int_output$iNextEst$Network_id
+int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
+int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
+int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
+##Pollinator by plant
+plant_poll_increase = plant_poll_output$iNextEst$Plant_id
+plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
+plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
+plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
+
+#Prepare results from common and rare poll across sampling sites
+#Pollinators
+poll_spread = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
+group_by(Network_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+ungroup() %>% 
+count(Pollinator_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
+#Check number of studies under 1%
+poll_total = nrow(poll_spread)
+polls_under_one_percent = poll_spread %>% 
+filter(Percent_total<1) %>% 
+nrow()
+polls_under_one_percent = polls_under_one_percent/poll_total * 100
+#Find most common pollinator
+most_common_poll = poll_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Pollinator_accepted_name)
+#Find most common poll percent
+most_common_poll_percent = poll_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Percent_total)
+#Plants
+plant_spread = data %>% 
+select(Plant_rank, Network_id, Plant_accepted_name) %>% 
+group_by(Network_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+ungroup() %>% 
+count(Plant_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
+#Check number of studies under 1%
+plant_total = nrow(plant_spread)
+plants_under_one_percent = plant_spread %>% 
+filter(Percent_total<1) %>% 
+nrow()
+plants_under_one_percent = plants_under_one_percent/plant_total * 100  
+#Find most common plant
+most_common_plant = plant_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Plant_accepted_name)
+#Find most common plant percent
+most_common_plant_percent = plant_spread %>% 
+filter(Percent_total == max(Percent_total)) %>% 
+pull(Percent_total)
+```
+
+## Sampling coverage
+
+The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_coverage)`% for both taxonomic groups.
+This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**).
+The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
+However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
+When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
+The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
+We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**).
+The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across networks.
+Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
+The solid black points indicate the number of species and interactions contained in the database.
+Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
+This last graph uses the same color and shape structure as the ones in the top panel.
+Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks.
+Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network.
+Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
+
+#Pollinators-----
+poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
+pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
+#Create dataset
+d_poll = poll_output$iNextEst$Network_id
+#Check cols
+d_poll$col = d_poll$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = pollinators_curves$Study_sampled, 
+           method = as.character(pollinators_curves$iteration), 
+           qD = pollinators_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_poll1 = bind_rows(d_poll, a_poll)
+d_poll1_point = d_poll1 %>%  filter(method == "observed")
+d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
+
+#Order levels
+d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col),size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_poll1_point, col = "black", size=2) + 
+ylab("Pollinator species") +
+xlab("Networks") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000)+
+ylim(0, 3000) +
+ggtitle("(a)") +
+theme(plot.margin = margin(r = 12))
+
+#Plants-----
+plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
+plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
+#Create dataset
+d_plant = plant_output$iNextEst$Network_id
+#Check cols
+d_plant$col = d_plant$method
+
+#Now create tibble with same structure
+a_plant = tibble(t = plant_curves$Study_sampled, 
+           method = as.character(plant_curves$iteration), 
+           qD = plant_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_plant1 = bind_rows(d_plant, a_plant)
+d_plant1_point = d_plant1 %>%  filter(method == "observed")
+d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
+
+#Order levels
+d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col), size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_plant1_point, col = "black", size=2) + 
+ylab("Plant species") +
+xlab("Networks") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000)+
+ylim(0, 3000)+
+ggtitle("(b)")+
+theme(plot.margin = margin(r = 12))
+
+#Interactions-----
+int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
+interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
+#Create dataset
+d_int = int_output$iNextEst$Network_id
+#Check cols
+d_int$col = d_int$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = interactions_curves$Study_sampled, 
+           method = as.character(interactions_curves$iteration), 
+           qD = interactions_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_int1 = bind_rows(d_int, a_poll)
+d_int1_point = d_int1 %>%  filter(method == "observed")
+d_int1_lines = d_int1 %>%  filter(!method == "observed")
+
+#Plot
+#Order levels
+d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col), size = 0.65) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_int1_point, col = "black", size=2) + 
+ylab("Interactions") +
+xlab("Networks") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off")  +
+xlim(0,3000) +
+ylim(0, 30000)+
+ggtitle("(c)")
+
+#----------------#
+#Pollinators
+#----------------#
+#Read data
+data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
+mutate(Network_id = paste0(Study_id, Network_id))  
+#First select cols of rank, accepted name and study ID
+#Then select unique levels by study ID
+#Now sum levels (maximum can be as maximum number of studies)
+poll_spread = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
+group_by(Network_id) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+ungroup() %>% 
+count(Pollinator_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = poll_spread %>%  
+select(Pollinator_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common pollinator for plotting
+most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
+select(Pollinator_accepted_name) %>% pull()
+
+#Plot
+o1 = ggplot(poll_spread, 
+aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
+geom_bar(stat = "identity", 
+           color = "#D55E00",
+           lwd = 0.03, fill = "#D55E00") +
+ylab("Shared pollinators (%)") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+  axis.ticks.x=element_blank(),
+  axis.ticks.y = element_line(size = 0.2),
+  panel.grid.major.x = element_blank(),
+  panel.grid.minor.y = element_blank(),
+  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
+xlab("Pollinator species") +
+ylab("Occurrence across networks") + 
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+color = "black", size=0.35) + 
+annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
+annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
+annotate(geom="text", x=200, y=-0.05, label="Common spp."
+      ,color="black",  size = 3) +
+annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
+      ,color="black",  size = 3)  +
+scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
+coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
+                ylim=c(0,0.7), clip="off") +
+ggtitle("(e)")
+
+#----------------#
+#Plants
+#----------------#
+plant_spread = data %>% 
+select(Plant_rank, Network_id, Plant_accepted_name) %>% 
+group_by(Network_id) %>% 
+filter(Plant_rank == "SPECIES") %>% 
+distinct(Plant_accepted_name) %>% 
+ungroup() %>% 
+count(Plant_accepted_name) %>% 
+rename(n_networks_shared = n) %>% 
+arrange(-n_networks_shared) %>% 
+mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
+
+#Seelect quantile 50% from the total number of species
+spp_number = plant_spread %>%  
+select(Plant_accepted_name) %>% 
+n_distinct()
+df = seq(1, spp_number)
+quantile_50 = quantile(df, 0.5)
+
+#Obtain most common plant for plotting
+most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
+select(Plant_accepted_name) %>% pull()
+
+o2 = ggplot(plant_spread, 
+aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
+geom_bar(stat = "identity", 
+           color = "#009E73",
+           lwd = 0.03, fill = "#009E73") +
+theme_minimal()+
+theme(axis.text.x = element_blank(),
+         axis.ticks.x=element_blank(),
+         axis.ticks.y = element_line(size = 0.2),
+         panel.grid.major.x = element_blank(),
+         panel.grid.minor.y = element_blank(),
+         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
+xlab("Plant species") +
+ylab("Occurrence across networks") + 
+geom_vline(xintercept = quantile_50, linetype="dashed", 
+          color = "black", size=0.35) + 
+annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
+annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
+      ,color="black",  fontface = 'italic', size = 3)+
+annotate(geom="text", x=150, y=-0.05, label="Common spp."
+      ,color="black",  size = 3) +
+annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
+      ,color="black",  size = 3) + 
+scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
+coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
+                ylim=c(0, 0.7), clip="off") +
+ggtitle("(f)")
+
+#Pollinator by plant species
+plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
+plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
+
+#Create dataset
+d_pp = plant_poll_output$iNextEst$Plant_id
+#Check cols
+d_pp$col = d_pp$method
+
+#Now create tibble with same structure
+a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
+           method = as.character(plant_poll_curves$iteration), 
+           qD = plant_poll_curves$Unique_spp_cumulative,
+           col = "collectors")
+
+#Bind datasets 
+d_pp1 = bind_rows(d_pp, a_poll)
+d_pp1_point = d_pp1 %>%  filter(method == "observed")
+d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
+
+#Plot
+#Order levels
+d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+#Plot
+pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
+geom_line(aes(linetype = col)) +
+scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
+scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
+geom_ribbon(aes(ymin = qD.LCL, 
+    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
+scale_fill_manual(values = c("gray", "black", "red")) +
+geom_point(data = d_pp1_point, col = "black") + 
+ylab("Pollinator species") +
+xlab("Plant species") +
+theme_bw() +
+coord_cartesian(expand = FALSE, clip = "off") +
+ggtitle("(d)") +
+scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
+scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
+theme(legend.position = "none")
+
+#Plot all
+p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
+      axis.title = element_text(size=12, face = "bold"),
+      axis.text = element_text(size=10),
+      plot.title = element_text(size=18))
+
+#a/b
+
+```
+
+\newpage
+
 ## Habitat type and bioclimatic region
 
-The proportion of the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
+The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
 As expected, Hymenoptera was the main taxonomic order on the majority of habitats, exceeded only by Diptera on the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
-Overall, the proportion of Lepidoptera and Coleoptera was low across all habitats but Coleoptera showed a notably increase in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
+Overall, the proportion of Lepidoptera and Coleoptera species was low across all habitats but Coleoptera were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
 Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
 In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
-Again, Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera seems to be only relevant in Mediterranean regions at European level.
+Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera is only relevant in Mediterranean regions at European level.
 Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
@@
 However, the habitats that contain a higher number of sampling sites are `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies are `r studies_by_bioregion$Bioregion[[1]]` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r studies_by_bioregion$Bioregion[[2]]` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r studies_by_bioregion$Bioregion[[3]]` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r networks_by_bioregion$Bioregion[[1]]` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r networks_by_bioregion$Bioregion[[2]]` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r networks_by_bioregion$Bioregion[[3]]` (`r networks_by_bioregion$Bioregion_count[[3]]`).
+The bioclimatic regions with a higher number of studies are `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
 
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
+Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
 The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
@@
 
-EuPPollNet offers the largest set of plant-pollinator networks and plant-pollinator studies compiled to date at European level.
-The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million of interaction records.
+EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level.
+The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records.
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-Given that most of the plant-pollinator networks from the database are sampled on pastures, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-However, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
+This likely reflects the fact that most plant and pollinator species are rare and geographically restricted.
+For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
+Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
 This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
-In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require of numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
-
-Bees are responsible for the majority of the interactions at the metaweb level, but their relative relevance changed across habitats and bioclimatic regions.
+In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
+
+Bees are responsible for the majority of the interactions at the metaweb level, but their relevance for network topology changed across habitats and bioclimatic regions.
 For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
 These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only common in the Mediterranean region.
+In addition, beetles were only common floral visitors in the Mediterranean region.
 Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
@@
 This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-On the contrary, honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-Indeed, this proportion is higher than the one found when considering only natural communities at a global scale [~13%\; @hung2018], which highlights their key role as potential pollinators of many flowering plants.
-However, the rise of beekeeping in Europe has led to high honey bee densities, which can negatively impact wild pollinators and the plants they visit [@steffan2000; @magrach2017; @herrera2020], emphasising the need to evaluate these impacts and implement actions if necessary.
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of plants in the database.
+Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
+The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practice in Europe [@steffan2000; @magrach2017; @herrera2020].
+
+Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio), the observed number of pollinator species in the database was almost double that of the plants.
 This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centred sampling is likely to increase the plant:pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species.
-Overall, latitude did not explain major changes in residual connectance or normalised nestedness across networks.
-However, consistent with @olesen2002, residual connectance -the deviation from the expected connectance for a given network size- was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes.
-Thus, giving their association and that both metrics capture structural properties of species interactions, the potential ecological explanations that can help understanding their latitudinal trends can be similar.
-For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012].
-Therefore, we hypothesise that these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones, as these networks are likely to deviate more from a perfect nested or hierarchical structure compared to networks at higher latitudes.
-Finally, while one third of networks were more nested than expected by chance, two thirds did not show statistical difference with null expectations.
-Therefore, while this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks.
+While animal-centerred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
+In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
+Rare pollinators are functionally important for plant species at large scales [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to observe these rare species and to effectively understand and protect plant-pollinator biodiversity.
+
+Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
+Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
+These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes.
+Finally, one third of networks were more nested than expected by chance.
+While this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result also suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks [@payrato2019].
 
 Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, most plant-pollinator networks belongs to central Europe, while Eastern Europe and the Mediterranean region are under-represented.
+For instance, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
 This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-This is especially relevant as Eastern Europe is expected to continue developing and use their vast natural resources [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+This lack of data is especially relevant for Eastern Europe which has vast landscapes of high quality semi-natural grasslands but is experiencing rapid land use change [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium or The Netherlands) seem to have suffered most of the plant and pollinator biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-polllinator communities in Europe and across the globe still suffer from major threads such as climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], or the introduction of alien species [@vila2009; @vanbergen2018].
+Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], and the introduction of alien species [@vila2009; @vanbergen2018].
 Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
@@
 
-We thank all the taxonomists and ecologists that have made this database possible by contributing with their fieldwork data.
+We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data.
 
@@
 
-All data and code to produce this database and manuscript are available at Zenodo (LINK) and Github (LINK).
+All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
 

### 2024-06-14 Scripts/1_Processing_networks/20_Hoiss.R

@@ -122,9 +122,9 @@ rename(Metadata_fields = rowname, Metadata_info= V1) %>% as_tibble()
 
 #Prepare authorship data ----
 Authorship <- tibble(
-Coauthor_name = "Bernhard Hoiss",
-Orcid = NA,
-E_mail = "bernhard.hoiss@anl.bayern.de")
+Coauthor_name = c("Bernhard Hoiss", "Ingolf Steffan-Dewenter"),
+Orcid = c(NA, "0000-0003-1359-3944"),
+E_mail = c("bernhard.hoiss@anl.bayern.de", "ingolf.steffan@uni-wuerzburg.de"))
 
 #Save data ----
 #Create metadata list

### 2024-06-14 Scripts/1_Processing_networks/21_Hopfenmuller.R

@@ -128,9 +128,9 @@ rename(Metadata_fields = rowname, Metadata_info= V1) %>% as_tibble()
 
 #Prepare authorship data ----
 Authorship <- tibble(
-Coauthor_name = "Sebastian Hopfenmüller",
-Orcid = "0009-0004-5138-6342",
-E_mail = "")
+Coauthor_name = c("Sebastian Hopfenmüller", "Ingolf Steffan-Dewenter"),
+Orcid = c("0009-0004-5138-6342", "0000-0003-1359-3944"),
+E_mail = c("sebastian_hopfenmueller@web.de", "ingolf.steffan@uni-wuerzburg.de"))
 
 #Save data ----
 #Create metadata list

### 2024-06-14 Scripts/1_Processing_networks/3_Michez.R

@@ -102,7 +102,7 @@ Authorship <- data.frame(
   Coauthor_name = c("Denis Michez", "William Fiordaliso", "Sara Reverte"),
   Orcid = c("0000-0001-8880-1838", "0000-0002-9667-6800", "0000-0002-2924-3394"),
   E_mail = c("denis.michez@umons.ac.be", "william.fiordaliso@umons.ac.be", 
-             "sara.revertes@umons.ac.be"))
+             "sara.revertesaiz@umons.ac.be"))
 
 
 #Save data ----

### 2024-06-14 Scripts/4_Bioregion_information.R

@@ -7,7 +7,7 @@ library(sf)
 library(dplyr)
 
 #Load data 
-data = readRDS("Data/3_Final_data/Interactions_uncounted.rds")
+data = readRDS("Data/Working_files/Interactions_uncounted.rds")
 
 ##Read shapefile:
 bioregions <- sf::read_sf('Data/Working_files/Bio_Regions/BiogeoRegions2016.shp')

### 2024-06-14 Scripts/Metaweb_info/Authorship.R

@@ -53,10 +53,26 @@ filter(Coauthor_name == "Ignasi Bartomeus") %>%
   distinct(Coauthor_name, .keep_all = TRUE)
 data = data %>% 
 filter(!Coauthor_name == "Tiffany M. Knight" & !Coauthor_name =="Ignasi Bartomeus")
+#Fix one name
+data = data %>% 
+mutate(Coauthor_name = case_when(
+  Coauthor_name == "Elena Motivans Švara" ~ "Elena Motivans",
+  Coauthor_name == "Paula Dominguez Lapido" ~ "Paula Dominguez-Lapido",
+  T ~ Coauthor_name))
+
+
 #Prepare author data for article
 authors = data
+#Create alternative column to order authors
+authors$authors_ordered = word(authors$Coauthor_name,-1)
+#Fix manually some authors 
+#as the last name is not the correct one to assign order
+authors = authors %>% 
+mutate(authors_ordered = case_when(
+  authors_ordered == "Manincor" ~ "de",
+  T~authors_ordered))
 #Order dataset by last word of author's name
-ordered_data = authors[order(word(authors$Coauthor_name,-1)),]
+ordered_data = authors[order(authors$authors_ordered),]
 #Filter for duplicated author names
 ordered_authors = ordered_data %>% 
   distinct(Coauthor_name, .keep_all = TRUE)
@@ -164,11 +180,14 @@ TRUE ~ Affiliation3))
 ##Joanne M. Bennett----
 affiliations = affiliations %>% 
 mutate(Affiliation1 = case_when(Coauthor_name == "Joanne M. Bennett"  ~ 
-"German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig",
-TRUE ~ Affiliation1)) %>% 
+"Fenner School of Environment & Society, The Australian National University, Canberra, Australia",
+TRUE ~ Affiliation1))%>% 
 mutate(Affiliation2 = case_when(Coauthor_name == "Joanne M. Bennett"  ~ 
+"German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig",
+TRUE ~ Affiliation2)) %>% 
+mutate(Affiliation3 = case_when(Coauthor_name == "Joanne M. Bennett"  ~ 
[…]

### 2024-08-12 Manuscript/EuPPollNet_Ms.qmd
Editing text based co-authors comments

@@
 bibliography: references.bib
+csl: apa_6th.csl
 latex-tinytex: false
@@
 
+#Check number of species from Hymenop, Dipte, Lepidop and Coleop
+
+pollinator_spp = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+nrow(.)
+
+pollinator_spp_main_orders = data %>% 
+select(Pollinator_order, Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Pollinator_order == "Hymenoptera"|
+       Pollinator_order == "Diptera"|
+       Pollinator_order == "Lepidoptera"|
+       Pollinator_order == "HymenoColeopteraptera") %>% 
+distinct(Pollinator_accepted_name) %>% 
+nrow(.)
+
+spp_percent_main_order = pollinator_spp_main_orders/pollinator_spp *100
+
 ```
@@
 One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
+Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
 This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
@@
 \small
-plant-pollinator networks, species interactions, flowering plants, Angiosperms, pollinators, nestedness, connectance
+Angiosperms, connectance, flowering plants, nestedness, plant-pollinator networks, pollinators, species interactions 
 \endgroup
@@
 The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003].
+Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano2003] or nestedness [@bascompte2003].
 Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
-Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @trojelsgaard2013].
+Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @garcia2024].
 For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
@@
 However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
+Numerous works have studied plant-pollinator interactions in the last decades, generating thousands of plant-pollinator interaction networks worldwide.
 Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
@@
 The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
 To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
@@
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked from members of the Safeguard project.
+As this database is the result of one of the work packages of the European project Safeguard, first, data was directly requested from members of the Safeguard project.
 Second, the request was extended to data owners outside of the project.
@@
 The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a pollinator individual to the reproductive structure of a particular plant.
-Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species in the database.
+Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+Species that belong to other orders `r format(round(100-spp_percent_main_order, 2), nsmall = 2)`% are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
+Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database.
 However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).

### 2024-08-12 Manuscript/EuPPollNet_word.qmd
Editing text based co-authors comments

@@
 Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practice in Europe [@steffan2000; @magrach2017; @herrera2020].
+The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honey bees, the important role that honey bees are playing as pollinators in Europe, and/or because beekeeping is widely practice in Europe [@steffan2000; @magrach2017; @herrera2020].
 
@@
 This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centerred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
+While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
 In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
@@
 In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
-Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
+Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not an ubiquitous feature of all plant-pollinator networks.
 These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
@@
 
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data.
+We thank all the taxonomists and ecologists who have made this database possible by contributing their fieldwork data.
 

### 2024-08-14 Manuscript/EuPPollNet_Ms.qmd
Minor updates following co-authors comments

@@
 
-d = d %>% 
-mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
-filter(!Variable == "Longitude")
-
 #Add descriptor
@@
     Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
+    Variable == "Latitude" ~ "North-south position of the observed interaction in decimal degrees",
+        Variable == "Longitude" ~ "East-west position of the observed interaction in decimal degrees",
     Variable == "Date" ~ "Year, month and day when the observation took place",
@@
 **Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., family, order and phylum).
+In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).
 
@@
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-As this database is the result of one of the work packages of the European project Safeguard, first, data was directly requested from members of the Safeguard project.
+As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project.
 Second, the request was extended to data owners outside of the project.
 These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
-To maintain high quality standards that will support robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
+To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).
 
@@
 The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a pollinator individual to the reproductive structure of a particular plant.
+Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.
 Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
@@
 However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
-Notably,  _Apis mellifera_ represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
+Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
 
@@
 
-The EuPPollNet database is available in both .csv and .rds formats.
-The files contain a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
-Each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix within a single flowering season at a site level, users only need to group interactions by plant and pollinator species, site, study and year.
-Information about habitat type and bioclimatic region are also provided in this file.
-In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study.
-The flower count data is provided in a separate file (.csv or .rds) that can be merged with the interaction data through the "Flower_data_merger" column.
-Two thirds of studies contain information on floral abundance, but the methods and units differ greatly across studies.
-Finally, metadata at the study level is provided, including information about the authors, digital object identifier if available, sampling time and taxonomic coverage of the main pollinator groups for each study.
+The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble) + 1` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
+These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29).
+The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30).
+Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies.
+To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year.
+Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.
+
+\newpage
 
@@
 For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-In addition, taxonomic information at family, order and phylum level was downloaded for each species.
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzing matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
+In addition, taxonomic information at genus, family and order level was downloaded for each species.
+
+For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
 
@@
 EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinators groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
+The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
 Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
@@
 Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
-Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera is only relevant in Mediterranean regions at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
-The habitats sampled by a higher number of studies in the database are `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
-However, the habitats that contain a higher number of sampling sites are `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies are `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
+In this case, Hymenoptera was the predominant order in all bioclimatic regions, with Diptera being particularly abundant in the Alpine and Atlantic regions.
+Lepidoptera showed low proportions across all bioclimatic regions and Coleoptera were only relevant in the Mediterranean region at European level.
+Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differed across habitats and bioclimatic regions.
+The habitats sampled by a higher number of studies in the database were `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
+However, the habitats that contain a higher number of sampling sites were `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
+The bioclimatic regions with a higher number of studies were `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and those that contain a higher number of sampling sites were `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
 
@@
 
-Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio), the observed number of pollinator species in the database was almost double that of the plants.
+Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants.
 This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centerred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
+While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
 In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
@@
 This research was funded by the H2020 European project Safeguard (101003476) and by the Federal State of Saxony-Anhalt (MLU-BioDivFund).
+RT was supported by the Czech Science Foundation (Project No.
+21-24186M).
+NDM, NH, YP and FM were financially supported by the ANR ARSENIC project (grant no.
+14-CE02-0012), the ANR NGB project (grant no.
+17-CE32- 011), the Region Nord-Pas-de-Calais, the CNRS, the French Ministère de l'Enseignement Supérieur et de la Recherche, the Hauts-de-France Region and the European Regional Funds.
 
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S3}.
-Association between nestedness (NODFc) and the geometric mean of plan and pollinator species per network", fig.width=3, fig.height=3}
+Association between nestedness (NODFc) and the geometric mean of plant and pollinator species per network", fig.width=3, fig.height=3}
 

### 2024-08-14 Manuscript/Figures_and_Tables/Table_data_structure.Rmd
edits
Minor updates following co-authors comments

--- 2024-03-20 Manuscript/Table_data_structure.Rmd
@@
 #Prepare table with data structure
-data = readRDS("../Data/3_Final_data/Interaction_data.rds")
+data = readRDS("../../Data/3_Final_data/Interaction_data.rds")
 
@@
 
-d = d %>% 
-mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
-filter(!Variable == "Longitude")
+#d = d %>% 
+#mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
+#filter(!Variable == "Longitude")
 
@@
     Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
+    Variable == "Latitude" ~ "North-south position of the observed interaction in decimal degrees",
+        Variable == "Longitude" ~ "East-west position of the observed interaction in decimal degrees",
+
     Variable == "Date" ~ "Year, month and day when the observation took place",

### 2024-08-14 Manuscript/EuPPollNet_Ms.qmd
percent fo orders per study

@@
 
+#Calculate percentage of studies that recorded the main 4 orders
+metadata = readRDS("../Data/3_Final_data/Metadata.rds")
+#Some studies are divided by year
+#Get an average of the pollinator percentages across years first
+#Fix NA's
+metadata_study = metadata %>%
+group_by(Study_id) %>% 
+mutate(Hymenoptera_percent = as.numeric(na_if(Hymenoptera_percent, "NA"))) %>%  
+mutate(Diptera_percent = as.numeric(na_if(Diptera_percent, "NA"))) %>%  
+mutate(Lepidoptera_percent = as.numeric(na_if(Lepidoptera_percent, "NA"))) %>%  
+mutate(Coleoptera_percent = as.numeric(na_if(Coleoptera_percent, "NA"))) 
+#Calculate average
+mean_study = metadata_study  %>% 
+summarise(Hymenoptera_percent = mean(Hymenoptera_percent, na.rm = T),
+          Diptera_percent = mean(Diptera_percent, na.rm = T),
+          Lepidoptera_percent = mean(Lepidoptera_percent, na.rm = T),
+          Coleoptera_percent = mean(Coleoptera_percent, na.rm = T)) %>% 
+ mutate(across(c(Hymenoptera_percent, Diptera_percent, Lepidoptera_percent, Coleoptera_percent),
+                ~ ifelse(is.nan(.), NA, .))) 
+#Now calculate percentage of studies that record any of these 4 orders
+mean_percent_order = mean_study %>%
+  summarise(
+    Hymenoptera_percent_metaweb = sum(!is.na(Hymenoptera_percent)) / nrow(.) * 100,
+    Diptera_percent_metaweb = sum(!is.na(Diptera_percent)) / nrow(.) * 100,
+    Lepidoptera_percent_metaweb = sum(!is.na(Lepidoptera_percent)) / nrow(.) * 100,
+    Coleoptera_percent_metaweb = sum(!is.na(Coleoptera_percent)) / nrow(.) * 100)
+
+Hymenoptera_percent_metaweb = mean_percent_order$Hymenoptera_percent_metaweb
+Diptera_percent_metaweb = mean_percent_order$Diptera_percent_metaweb
+Lepidoptera_percent_metaweb = mean_percent_order$Lepidoptera_percent_metaweb
+Coleoptera_percent_metaweb = mean_percent_order$Coleoptera_percent_metaweb
+
 ```
@@
 Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+
+`r format(round(Hymenoptera_percent_metaweb, 2), nsmall = 2)`% of the studies documented interactions from Hymenoptera, `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% from Diptera, `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% from Lepidoptera, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% from Coleoptera.
+
 The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.

### 2024-08-14 Scripts/Metaweb_info/Study_level_Metadata.R

@@ -23,7 +23,7 @@ library(tidyr) #Reshape data (wide format)
 library(stringr)
 library(readr)
 #Load data 
-data = readRDS("Data/3_Final_data/Interactions_uncounted.rds")
+data = readRDS("Data/3_Final_data/Interaction_data.rds")
 data = data %>%
 dplyr::mutate(Year = lubridate::year(Date), 
                 Month = lubridate::month(Date), 
@@ -124,6 +124,64 @@ mutate(DOI = str_replace(DOI, " and", ";"))
 metadata = left_join(metadata, doi) %>% 
 relocate(DOI, .before = Year)
 
+#Add qualitative info of what orders where sampled
+#Based on authors info
+metadata = metadata %>% 
+mutate(Taxa_recorded = NA_character_) %>%
+mutate(Comment_taxa_recorded = NA_character_)
+
+#Studies by recorded floral visitors
+all_pollinators = c("1_Bartomeus", "2_Petanidou", 
+        "8_Biella", "9_Heleno", 
+        "10_Vanbergen", "11_Clough",
+        "14_Dupont", "15_Magrach",
+        "18_Bartomeus", "20_Hoiss",
+        "26_Castro", "27_Castro",
+        "48_Lara-Romero", "49_Hervias-Parejo",
+        "50_Hervias-Parejo", "51_Petanidou",
+        "52_Hadrava")
+
+only_bees = c("3_Michez")
+
+only_bumblebees = c("13_Karise", "22_Kallnik")
+
+bees_and_syrphids = c("6_Marini", "7_Scheper",
+                      "12_Ockinger", "16_Manincor",
+                      "19_Jauker", "23_Holzschuh",
+                      "24_Holzschuh", "25_Holzschuh",
+                      "28_Sutter", "29_Magrach",
+                      "30_Smith", "31_Roberts",
+                      "38_Maurer", "39_Schweiger")
+
+bees_syrphids_and_butterflies = c("32_ORourke", "33_Power",
+                                  "34_Stanley", "35_Mullen",
+                                  "36_Larkin", "37_White",
+                                  "47_Benadi")
+
+hymenoptera_diptera_lepidoptera = c("40_Knight", "41_Knight",
+                                    "42_Knight", "43_Knight",
+                                    "44_Knight", "45_Knight",
[…]

### 2024-08-15 Manuscript/EuPPollNet_Ms.qmd
updates on MS based on coauthors comments

@@
 
+#Check percentages of hymenoptea groups
+hymenoptera_percent_by_group = metadata %>% 
+ungroup() %>% 
+distinct(Study_id, Taxa_recorded) %>% 
+summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
+          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
+          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))
+
+bee_percent = hymenoptera_percent_by_group$Bee_studies
+bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
+all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies
+
+#Check percentages of diptera groups
+diptera_percent_by_group = metadata %>% 
+ungroup() %>% 
+distinct(Study_id, Taxa_recorded) %>% 
+summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
+          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
+          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
+                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)
+
+syrphid_percent = diptera_percent_by_group$Syrphid_studies
+all_diptera_percent = diptera_percent_by_group$Diptera_studies
+other_diptera_percent = diptera_percent_by_group$other_diptera_studies
+
 ```
@@
 Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-
-`r format(round(Hymenoptera_percent_metaweb, 2), nsmall = 2)`% of the studies documented interactions from Hymenoptera, `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% from Diptera, `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% from Lepidoptera, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% from Coleoptera.
-
+All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
 The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.
@@
 The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not to follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
+Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
 
@@
 #Coverage
-poll_coverage = poll_output$DataInfo$SC *100
-plant_coverage = plant_output$DataInfo$SC *100
+poll_sampling_coverage = poll_output$DataInfo$SC *100
+plant_sampling_coverage = plant_output$DataInfo$SC *100
 int_coverage = int_output$DataInfo$SC *100
@@
 
-The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_coverage)`% for both taxonomic groups.
+The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_sampling_coverage)`% for both taxonomic groups.
 This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**).
@@
 
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
+Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
 Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
@@
 Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
-Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located in lower latitudes.
+Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes.
 Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
@@
 Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
-In addition, the bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
+The bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
 
@@
 
-10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
+10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located at high elevation areas within Europe.
 

### 2024-08-15 Manuscript/EuPPollNet_Ms.qmd
edits on figure 2

@@
 
-Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination.
+Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that benefit from animal pollination.
 EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
@@
 The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Grey bars indicate the total number of species per genus at the European level."}
+Dark grey bars indicate the total number of species per genus at the European level."}
 
@@
                     color="black",
-                    offset = 0.137,
+                    offset = 0.146,
             axis.params=list(
@@
                          text.size  = 2.5,
-                         hjust  = 1,
+                         hjust  = 1.2,
                          vjust = 0.5,
@@
                          text.size  = 2.5,
-                         hjust  = 1,
+                         hjust  = -0.2,
                          vjust = 0.5,
-                         nbreak = 1,
+                         nbreak = 3,
                          fontface = 2
@@
          grid.params=list()) +
-annotate("text", x = 14.2, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.1, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
+annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
+annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
+annotate("text", x = 12.3, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
 new_scale_fill() +
@@
 scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Bee family coverage")
+ggtitle("Bee families")
 #--------------------------------------------------------------------------------------------#
@@
 #Plot
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
+# Plot for syrphid subfamily coverage with adjusted margins
+p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily, -Spp_number), Spp_number, fill = Group)) +
+  geom_bar(stat="identity", width = .5, position = "dodge", color = "black") +
   coord_cartesian(expand = FALSE) +
@@
   theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Syrphid subfamily coverage")
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    panel.border = element_rect(size = 1),
+    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 10)),
+    axis.title = element_text(face = "bold", size = 14),
+    axis.text = element_text(face = "bold", size = 12),
+    plot.margin = margin(10, 10, 20, 10)  # Adjusting the margins
+  ) +
+  scale_fill_manual(values = c("tan2", "lightgrey")) +
+  ggtitle("Syrphid subfamilies")
 #--------------------------------------------------------------------------------------------#
@@
 scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Butterfly family coverage")
+ggtitle("Butterfly families")
 

### 2024-08-16 Manuscript/EuPPollNet_Ms.qmd
edits following author's comments

@@
 geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.063, size=2.9, fontface=1) +
+geom_tiplab(linetype='dashed', linesize=.05, offset = 0.065, size=2.9, fontface=1) +
 scale_size(range = c(0,4.5), guide = 'none')  +
@@
                     color="black",
-                    offset = 0.146,
+                    offset = 0.148,
             axis.params=list(
@@
 annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.3, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
+annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
 new_scale_fill() +
-geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
-annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
-annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
-annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
-annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
-annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
-annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
+geom_point2(aes(subset=node==131), color='orange', size=3) +
+geom_point2(aes(subset=node==84), color='orange', size=3) +
+geom_point2(aes(subset=node==73), color='orange', size=3) +
+geom_point2(aes(subset=node==74), color='orange', size=3) +
+geom_point2(aes(subset=node==111), color='orange', size=3)+
+geom_point2(aes(subset=node==91), color='orange', size=3) +
+annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
+annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
+annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
+annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
+annotate("text", x = 2.55, y = 28.5, label = "Apidae", size= 3)+
+annotate("text", x = 2.25, y = 51.2, label = "Megachilidae", size= 3) +
 coord_cartesian(clip = "off")+

### 2024-08-16 Manuscript/EuPPollNet_word.qmd
edits following author's comments

@@
 One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
+Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and computationally reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
 This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.

### 2024-08-20 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 
-The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
+The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
 As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project.
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Locations of the studies in EuPPollNet showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
+(\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
 The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
+For visualization purposes, we have selected only a single location per study.
 (\\textbf{b}) Number of studies by year in EuPPollNet.
-(\\textbf{c}) Number of species and interactions across the four main pollinator orders in EuPPollNet."}
+(\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}.
+The total number of species and interactions is indicated in parentheses at the bottom."}
 
@@
 
-#Plot
-#p1 = ggplot(poll_spread, 
-#aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-#geom_bar(stat = "identity", 
-#           color = "#D55E00",
-#           lwd = 0.015, fill = "#D55E00") +
-#ylab("Shared pollinators (%)") +
-#theme_minimal()+
-#theme(axis.text.x = element_blank(),
-#  axis.ticks.x=element_blank(),
-#  axis.ticks.y = element_line(size = 0.2),
-#  plot.title = element_text(size = 8, face = "bold"), 
-#  axis.text.y = element_text(size=5),
-#  panel.grid.major.x = element_blank(),
-#  panel.grid.minor.y = element_blank(),
-#  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
-#xlab(NULL) +
-#ylab(NULL) + 
-#ggtitle("Pollinators")+
-#geom_vline(xintercept = quantile_50, linetype="dashed", 
-#color = "black", size=0.25) + 
-#annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = #1.5) +
-#annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  #fontface = 'italic', size = 1.5)+
-#scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
-#annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
-#coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, #max(poll_spread$Percent_total)), clip="off")
-#
-##----------------#
-##Plants
-##----------------#
-#plant_spread = data %>% 
-#select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-#group_by(Study_id) %>% 
-#filter(Plant_rank == "SPECIES") %>% 
-#distinct(Plant_accepted_name) %>% 
-#ungroup() %>% 
-#count(Plant_accepted_name) %>% 
-#rename(n_studies_shared = n) %>% 
-#arrange(-n_studies_shared) %>% 
-#mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-#  
-##Seelect quantile 50% from the total number of species
-#spp_number = plant_spread %>%  
-#select(Plant_accepted_name) %>% 
-#n_distinct()
-#df = seq(1, spp_number)
-#quantile_50 = quantile(df, 0.5)
-#
-##Obtain most common plant for plotting
-#most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-#select(Plant_accepted_name) %>% pull() 
-#
-#Plot
-#p2 = ggplot(plant_spread, 
-#aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-#geom_bar(stat = "identity", 
-#           color = "#009E73",
-#           lwd = 0.015, fill = "#009E73") +
-#theme_minimal()+
-#theme(axis.text.x = element_blank(),
-#         axis.ticks.x=element_blank(),
-#         axis.ticks.y = element_line(size = 0.2),
-#         axis.title.x = element_text(face = "bold", size = 7),            plot.title = #element_text(size = 8, face = "bold"),
-#         axis.text.y = element_text(size=5),
-#         panel.grid.major.x = element_blank(),
-#         panel.grid.minor.y = element_blank(),
-#         panel.grid.major.y = element_line(size=0.2, linetype =  2),
-#      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
-#xlab("Species") +
-#ylab("Occurrence across studies") + 
-#ggtitle("Plants") + 
-#geom_vline(xintercept = quantile_50, linetype="dashed", 
-#          color = "black", size=0.25) + 
-#annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
-#annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
-#      ,color="black",  fontface = 'italic', size = 1.5)+
-#annotate(geom="text", x=20, y=-0.08, label="Common spp."
-#      ,color="black",  size = 1.5) +
-#annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
-#      ,color="black",  size = 1.5) + 
-#scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = #label_number(accuracy = 0.01)) +
-#coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, #max(plant_spread$Percent_total)), clip="off")
-#
-
 data2 = data %>%  select(Study_id, Year) %>% 
@@
 main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
+
+#Calculate spp number and interactions separately
+data1_spp = data %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+group_by(Pollinator_order) %>% 
+summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
+
+data1_int = data %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+#filter(Pollinator_rank == "SPECIES") %>% 
+group_by(Pollinator_order) %>% 
+summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))
+
+data1 = left_join(data1_spp, data1_int)
+
 #Select main orders
-data1 = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
+spp = data %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
 filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.),
-          Interactions = length(Pollinator_accepted_name)/ nrow(.))
-
-# Combine Interactions and Species into a single column
+filter(Pollinator_rank == "SPECIES") %>% 
+summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
+pull()
+#Select main orders
+int = data %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+#filter(Pollinator_rank == "SPECIES") %>% 
+summarise(Interactions = length(Pollinator_accepted_name)) %>% 
+pull()
+
+#Combine Interactions and Species into a single column
 data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
@@
 ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6))
+theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6)) +
+scale_x_discrete(labels = c("Species\n (2172)", "Interactions\n (358,049)"))
 
@@
 The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks.
-In addition, an outstanding question in ecology is how many pollinators are required to pollinate flowering plants [e.g., @kleijn2015 for crops].
-To that end, we also calculated the accumulation curve of pollinator species with an increasing number of plant species.
+In addition, we computed the  accumulation curve of pollinator species with an increasing number of plant species as an indicator for how many pollinator species are likely responsible for the pollination of flowering plants [e.g., @kleijn2015 for crops].
 The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
-The different rarefied curves were complemented with 100 boostrapped accumulation curves.
+The different rarefied curves were complemented with 100 bootstrapped accumulation curves.
 

### 2024-08-20 Manuscript/Figures_and_Tables/1_Figure1.qmd
edits

--- 2024-03-27 Manuscript/1_Figure1.qmd
@@
 #Load data
-data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
+data = readRDS("../../Data/3_Final_data/Interaction_data.rds")
 
@@
 main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
+
+#Calculate spp number and interactions separately
+data1_spp = data %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+group_by(Pollinator_order) %>% 
+summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
+
+data1_int = data %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+#filter(Pollinator_rank == "SPECIES") %>% 
+group_by(Pollinator_order) %>% 
+summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))
+
+data1 = left_join(data1_spp, data1_int)
+
 #Select main orders
-data1 = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
+spp = data %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
 filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.),
-          Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
+filter(Pollinator_rank == "SPECIES") %>% 
+summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
+pull()
+#Select main orders
+int = data %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+#filter(Pollinator_rank == "SPECIES") %>% 
+summarise(Interactions = length(Pollinator_accepted_name)) %>% 
+pull()
 
-# Combine Interactions and Species into a single column
+#Combine Interactions and Species into a single column
 data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")

### 2024-08-20 Scripts/1_Processing_networks/54_Lazaro.R

@@ -1,0 +1,111 @@
+#DATASET NUMBER 54; Lazaro
+#Sent By A. Lazaro 
+
+source("Scripts/Processing/Functions/Empty_templates.R") #Read empty templates to compare with
+
+#Load libraries
+library(dplyr)
+library(stringr)
+library(readr)
+library(tibble)
+library(tidyr)
+library(lubridate)
+library(anytime)
+
+#Read empty templates to compare with
+source("Scripts/Processing/Functions/Empty_templates.R") 
+#Load function to unify structure of data
+source("Scripts/Processing/Functions/Change_str.R")
+
+#Prepare interaction data ----
+data = read_csv("Data/1_Raw_Data/54_Lazaro/Interaction_data.csv")
+
+#Check against template
+compare_variables(check_interaction_data, data)
+
+#All columns are there!
+#Only add if there is flower info available
+data = data %>% 
+mutate(Flower_data = "Yes") %>% 
+mutate(Flower_data_merger = paste0(word(Plant_species, 1), "_" , word(Plant_species,2), "_", Site_id, "_",
+                                   Day, "_", Month, "_", Year))
+#Unify structure of data
+data = change_str(data)
+
+#Split interaction data into dataframes within a list
+InteractionData = split(data, data$Site_id)
+
+#Prepare flower count data ---- 
+FlowerCount = read_csv("Data/1_Raw_data/54_Lazaro/Flower_count.csv")
+
+#Group by and calculate acerage per species, date and site
+#as there is no finer way to bind with the interaction dataset
+FlowerCount = FlowerCount %>% 
+group_by(Day, Month, Year, Site_id, Plant_species, Comments, Units) %>% 
+summarise(Flower_count = mean(Flower_count))
+
+#Set merger col
+FlowerCount = FlowerCount %>% 
+mutate(Flower_data_merger = paste0(word(Plant_species, 1), "_" , word(Plant_species,2), "_", Site_id, "_", Month,"_", Year))
+
+#Drop not needed vars
+FlowerCount = drop_variables(check_flower_count_data, FlowerCount) 
+
+#Set common structure
+FlowerCount = change_str2(FlowerCount)
+
+#Split interaction data into dataframes within a list
+FlowerCount = split(FlowerCount, FlowerCount$Site_id)
+
+
+#Prepare metadata data ----
+#Store unique cases of plants and polls
[…]

### 2024-08-20 Scripts/3_1_Pollinator_taxonomy.R

@@ -504,7 +504,7 @@ select(!c(Fixed1))
 
 #Conduct last fixes
 #----------------------------#
-#2)Fix synonyms typos (with help of worldflora)-----
+#2)Fix synonyms typos-----
 #----------------------------#
 
 #Note, each case has been searched manually and assessed
@@ -602,6 +602,14 @@ mutate(Fixed = case_when(
   Mismatch == "Diplazontinae sp1" ~ "Ichneumonidae",
   Mismatch == "Procridinae sp1" ~ "Zygaenidae",
   Mismatch == "Scolia flava" ~ "Scolia",
+  Mismatch == "Lenipithema humilis" ~ "Linepithema humile", #typo
+  Mismatch == "Componotus ruber" ~ "Camponotus ruber", #typo
+  Mismatch == "Componotus lateralis" ~ "Camponotus lateralis", #typo
+  Mismatch == "Dasysyrphus balteatus" ~ "Episyrphus balteatus", #typo
+  Mismatch == "Temnothorax algiricus" ~ "Temnothorax", #Higher rank
+  Mismatch == "Gelechioidea sp1" ~ "Lepidoptera", #Higher rank
+  Mismatch == "Chalcidoidea sp1" ~ "Hymenoptera", #Higher rank
+  Mismatch == "Yponomeutoidea sp1" ~ "Lepidoptera", #Higher rank
   T ~ Fixed)) %>% 
   rename(Old_name = Mismatch, Name = Fixed) 
 
@@ -988,7 +996,7 @@ levels(factor(all$Study_id))
 #Do this fo every new dataset that we add
 #Last one being checked is written within the filter argument
 subset_check = all %>% 
-filter(Study_id == "52_Hadrava") %>% 
+filter(Study_id == "53_Libran-embid") %>% 
 select(Old_name, Fixed_name, Rank, Status, Matchtype, Accepted_name, Unsure_id) %>% 
 distinct()
 

### 2024-08-20 Manuscript/EuPPollNet_Ms.qmd
Minor edits

@@
 theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6)) +
-scale_x_discrete(labels = c("Species\n (2172)", "Interactions\n (358,049)"))
+scale_x_discrete(labels = c("Species\n (2,172)", "Interactions\n (358,049)"))
 

### 2024-08-21 Manuscript/EuPPollNet_Ms.qmd
Changes following co-authors comments

@@
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
-As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project.
-Second, the request was extended to data owners outside of the project.
+As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022.
+Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024.
 These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
+While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019].
+This approach maximized the potential number of studies that could be incorporated in this database.
+The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*".
 To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).
@@
 The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented only diurnal interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
 All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
@@
 To assess the completeness of plant and pollinator species in the EuPPollNet database at the European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
+Specifically for plants, we refined the checklist to include only plants occurring in Europe and excluded taxonomic groups not associated with biotic pollination.
 We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at the European level.
-The number of pollinator species at the European level was estimated by first estimating coverage for the three taxa that have checklists (bees, syrphids and butterflies).
-The mean coverage across these three groups (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`) was then used for the other taxa for which there is no checklist to extrapolate the number of pollinating species in Europe.
+Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
+For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies at the European level, as these are the only taxonomic groups with species checklists available in Europe.
+In addition, the potential number of pollinator species in Europe was estimated by combining the total number of species of bees, syrphids, butterflies and species from other taxonomic groups.
+The total number of species from other taxonomic groups was extrapolated by assuming that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
 

### 2024-08-22 Manuscript/EuPPollNet_Ms.qmd
adressing comments from colleagues

@@
 EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks.
-For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where generalist species interact with all other species but specialists interact preferentially only with generalists, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
+For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where specialists species interact only with subsets of generalists species, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
 
@@
 
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**).
+The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations).
 The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented only diurnal interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
 All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure 1}.
 (\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
@@
 ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -5))
+theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -5))
 
@@
 theme_classic() +
-scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator order") +
+scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator\n order") +
 xlab(NULL) +
@@
      legend.key.size = unit(3, "mm"),
-     legend.margin=margin(0,0,0,0),
-     legend.box.margin=margin(-8,-5,-5,-5)) +
+     legend.margin=margin(0,2,0,0),
+     legend.box.margin=margin(-9,-9,-6,-6)) +
 guides(fill=guide_legend(nrow=2,byrow=TRUE))+
 ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6)) +
+theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -6)) +
 scale_x_discrete(labels = c("Species\n (2,172)", "Interactions\n (358,049)"))
@@
 Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies at the European level, as these are the only taxonomic groups with species checklists available in Europe.
-In addition, the potential number of pollinator species in Europe was estimated by combining the total number of species of bees, syrphids, butterflies and species from other taxonomic groups.
-The total number of species from other taxonomic groups was extrapolated by assuming that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+For pollinators, we evaluated the taxonomic coverage of taxonomic groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
+While there is not a good understanding of pollinator diversity in other taxonomic groups, it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
+Consequently, we provide an estimate for the total number of pollinators across the European continent.
 
@@
 EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
+The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
+Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
 Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across networks.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their unique pairwise interactions across networks.
 Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
@@
 Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network.
-Note that indeed \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
+Note that \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
 
@@
 geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Interactions") +
+ylab("Unique interactions") +
 xlab("Networks") +
@@
 The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-As expected, Hymenoptera was the main taxonomic order on the majority of habitats, exceeded only by Diptera on the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
-Overall, the proportion of Lepidoptera and Coleoptera species was low across all habitats but Coleoptera were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
+Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
+Overall, the proportion of flower visitors from Lepidoptera and Coleoptera were low across all habitats but Coleopteran flower visitors were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
 Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-In this case, Hymenoptera was the predominant order in all bioclimatic regions, with Diptera being particularly abundant in the Alpine and Atlantic regions.
-Lepidoptera showed low proportions across all bioclimatic regions and Coleoptera were only relevant in the Mediterranean region at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differed across habitats and bioclimatic regions.
+Hymenopterans were predominant across all bioclimatic regions and Dipterans were particularly abundant in the Alpine and Atlantic regions.
+Lepidopterans had low proportions across all bioclimatic regions and Coleopterans were only relevant in the Mediterranean region at European level.
+Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S3**) also differed across habitats and bioclimatic regions.
 The habitats sampled by a higher number of studies in the database were `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
@@
 
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
+Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
+Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
 We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**).
@@
 
-Bees are responsible for the majority of the interactions at the metaweb level, but their relevance for network topology changed across habitats and bioclimatic regions.
+Bees are responsible for the majority of the sampled interactions at the metaweb level.
+As not all surveys included all pollinator groups, this result may partly be influenced by the taxonomic groups sampled across studies.
+However, the relevance of bees and other pollinator orders for network topology changed across habitats and bioclimatic regions in accordance to the literature.
 For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
 These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only common floral visitors in the Mediterranean region.
-Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
+In addition, beetles were only commonly documented as floral visitors in the Mediterranean region.
+This study cannot determine whether pollination ecologists traditionally document flower-beetle interactions only in the Mediterranean, or if there are fewer flower visitations by beetles outside this region.
+Nevertheless, the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
 The number of butterfly species and interactions were relatively low compared to the other taxa.
@@
 
+- Figure S4
+
 \newpage
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S1}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure S1}. Location of all networks in the EuPPollNet database coloured by bioregion."}
+
+#Read data
+data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
+mutate(Network_id = paste0(Study_id, Network_id))
+
+#Prepare coordinates
+for_plotting = data %>% 
+select(Study_id, Latitude, Longitude, Bioregion) %>% 
+distinct()
+#Get countries and transfrom coordinates
+countries <- gisco_get_countries(
+resolution = 20) %>%
+st_transform(3035) %>% 
+sf::st_as_sf(crs = 4326)
+
+#Try to plot both together (same plot)
+map = ggplot(countries) +
+geom_sf(fill="floralwhite",color = "black", 
+        alpha = 1,linewidth=0.1) + 
+geom_point(data = for_plotting, alpha=0.75,size=1.2,pch=21,stroke = 0.25,aes(x= Longitude, y=Latitude, fill= Bioregion))+  
+coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
+         xlim = c(-14, 39), ylim = c(37, 70))+
+theme(panel.grid.major = element_blank(), 
+     legend.key = element_blank(), 
+     legend.text = element_text(size = 6), 
+     legend.title = element_text(face ="bold",  size = 8),
+     legend.spacing.x=unit(-0.5, "lines"), 
+     legend.key.width=unit(0.5, "lines"), 
+     legend.box.margin=margin(-10,-10,-10,-10),
+     axis.text = element_text(size=6,face="bold"),
+     axis.title = element_text(face="bold", size = 8),
+) +
+theme(panel.background = element_rect(fill = "aliceblue"),
+        panel.border = element_rect(colour = "black",
+        fill=NA, size=1)) + 
+ scale_size_continuous(name= "Species counts",range = c(0, 3.75))  +
+ylab("Latitude") +
+xlab("Longitude") +
+scale_x_continuous(breaks = seq(0, 20, by = 10)) +
+scale_y_continuous(breaks = seq(40, 60, by = 10)) +
+guides(fill = guide_legend(override.aes = list(size = 1.5, byrow = TRUE))) +
+scale_fill_manual(values = c("purple3", 
+                               "dodgerblue3", 
+                               "cyan4", 
+                               "forestgreen",
+                               "orange2",
+                               "tomato3",
+                               "khaki2"))
+
+#Plot everything
+map 
+```
+
+\newpage
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S2}.
 Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S2}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S3}.
 Barplot indicating the number of networks by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
+
+#Filter out 2 bioregions that only have a small study with data of a single day
+data = data %>% 
+filter(!Bioregion == "Steppic") %>% 
+filter(!Bioregion == "Pannonian")
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S3}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S4}.
 Association between nestedness (NODFc) and the geometric mean of plant and pollinator species per network", fig.width=3, fig.height=3}

### 2024-08-22 Manuscript/Figures_and_Tables/Exact_location.qmd
adressing comments from colleagues

--- 2024-08-22 Manuscript/EuPPollNet_Ms.qmd
@@
 ---
-title: "EuPPollNet: A European database of plant-pollinator networks"
+latex-tinytex: false
 format:
@@
         \newcommand{\elandscape}{\end{landscape}}
-        \usepackage{caption}
-        \captionsetup[figure]{labelformat=empty}
-        \captionsetup[table]{labelformat=empty}
-        \usepackage{lineno}
-        \linenumbers
-bibliography: references.bib
-csl: apa_6th.csl
-latex-tinytex: false
-link-citations: TRUE
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
 ---
@@
 library(dplyr) #For handling fata 
+library(ggplot2) #For plotting
+library(giscoR) #For plotting
 library(sf) #For handling coordinates and plotting
+library(patchwork)#For plotting (binding plots)
+library(ggstar) #For plotting (cool shapes)
 library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
-library(readr)
-library(phytools)
-#Load libraries for plotting
-library(giscoR) #Maps
-library(ggstar) #cool shapes as points
-library(ggplot2) #general plotting
-library(ggtree) #phylo trees
-library(ggpattern) #cool shapes in barplots
-library(patchwork) #plot organisation
-library(cowplot) #plot organisation
-library(viridis) #colours
-library(ggnewscale)
-library(ggtreeExtra)
-library(kableExtra)
-
-#In case of conflicts select the main dplyr functions
-library(conflicted)
-conflicts_prefer(dplyr::filter)
-conflicts_prefer(dplyr::select)
-
+library(tidyr)
+library(viridis)
 ```
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}
 
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
-
-#Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
-#Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
-                    header = T)
-#Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")
-
-#Work around to get the three cols
-#Note that month 1 and day 1 indicates the presence of a missing value
-#At least in Europe :)
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-
-#Check number of distinct networks
-n_networks = data %>% 
-mutate(Network_id = paste0(Network_id,"_", Year)) %>% 
-distinct(Network_id) %>% 
-nrow(.)
-
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Read author list 
-authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
-#Print authors
-authors = authors %>% 
-summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
-pull()
-#Read affiliation list 
-aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
-aff_list = aff_list %>% 
-summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
-pull()
-
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Calculate mean coverage and deviation
-coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
-mean_coverage = mean(coverage)
-sd_coverage = sd(coverage)
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-#Prepare table with all info 
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude" ~ "North-south position of the observed interaction in decimal degrees",
-        Variable == "Longitude" ~ "East-west position of the observed interaction in decimal degrees",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_original_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_original_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
-        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
-
-    TRUE ~ "Other"
-  ))
-
-#Check number of species and percentage of interactions of poll. groups
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-order_info = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
-          Species = n_distinct(Pollinator_accepted_name))
-
-Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
-select(Interactions) %>% pull()
-
-#Check percentage of studies with flower counts
-floral_counts_percent = data  %>%
-select(Study_id, Flower_data) %>% 
-distinct() %>% 
-group_by(Flower_data) %>% 
-summarise(Percentage = n()/ nrow(.)*100) %>% 
-filter(Flower_data == "Yes")
-
-#Check percentage of apis mellifera records
-all_records = data %>% nrow(.) 
-apis_records = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-apis_percent = apis_records/all_records * 100
-#Check average pecentage of interactions of apis by net
-a = data %>% 
-group_by(Network_id) %>% 
-summarise(n_interactions = length(Pollinator_accepted_name))
-b = data %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-summarise(n_interactions_apis = length(Pollinator_accepted_name))
-c = left_join(a,b)
-apis_average_network = c %>% 
-mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
-ungroup() %>% 
-summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
-pull()
-
-#Check number of species from Hymenop, Dipte, Lepidop and Coleop
-
-pollinator_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-pollinator_spp_main_orders = data %>% 
-select(Pollinator_order, Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera"|
-       Pollinator_order == "Diptera"|
-       Pollinator_order == "Lepidoptera"|
-       Pollinator_order == "HymenoColeopteraptera") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-spp_percent_main_order = pollinator_spp_main_orders/pollinator_spp *100
-
-#Calculate percentage of studies that recorded the main 4 orders
-metadata = readRDS("../Data/3_Final_data/Metadata.rds")
-#Some studies are divided by year
-#Get an average of the pollinator percentages across years first
-#Fix NA's
-metadata_study = metadata %>%
-group_by(Study_id) %>% 
-mutate(Hymenoptera_percent = as.numeric(na_if(Hymenoptera_percent, "NA"))) %>%  
-mutate(Diptera_percent = as.numeric(na_if(Diptera_percent, "NA"))) %>%  
-mutate(Lepidoptera_percent = as.numeric(na_if(Lepidoptera_percent, "NA"))) %>%  
-mutate(Coleoptera_percent = as.numeric(na_if(Coleoptera_percent, "NA"))) 
-#Calculate average
-mean_study = metadata_study  %>% 
-summarise(Hymenoptera_percent = mean(Hymenoptera_percent, na.rm = T),
-          Diptera_percent = mean(Diptera_percent, na.rm = T),
-          Lepidoptera_percent = mean(Lepidoptera_percent, na.rm = T),
-          Coleoptera_percent = mean(Coleoptera_percent, na.rm = T)) %>% 
- mutate(across(c(Hymenoptera_percent, Diptera_percent, Lepidoptera_percent, Coleoptera_percent),
-                ~ ifelse(is.nan(.), NA, .))) 
-#Now calculate percentage of studies that record any of these 4 orders
-mean_percent_order = mean_study %>%
-  summarise(
-    Hymenoptera_percent_metaweb = sum(!is.na(Hymenoptera_percent)) / nrow(.) * 100,
-    Diptera_percent_metaweb = sum(!is.na(Diptera_percent)) / nrow(.) * 100,
-    Lepidoptera_percent_metaweb = sum(!is.na(Lepidoptera_percent)) / nrow(.) * 100,
-    Coleoptera_percent_metaweb = sum(!is.na(Coleoptera_percent)) / nrow(.) * 100)
-
-Hymenoptera_percent_metaweb = mean_percent_order$Hymenoptera_percent_metaweb
-Diptera_percent_metaweb = mean_percent_order$Diptera_percent_metaweb
-Lepidoptera_percent_metaweb = mean_percent_order$Lepidoptera_percent_metaweb
-Coleoptera_percent_metaweb = mean_percent_order$Coleoptera_percent_metaweb
-
-#Check percentages of hymenoptea groups
-hymenoptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
-          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
-          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))
-
-bee_percent = hymenoptera_percent_by_group$Bee_studies
-bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
-all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies
-
-#Check percentages of diptera groups
-diptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
-          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
-          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
-                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)
-
-syrphid_percent = diptera_percent_by_group$Syrphid_studies
-all_diptera_percent = diptera_percent_by_group$Diptera_studies
-other_diptera_percent = diptera_percent_by_group$other_diptera_studies
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first four and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks.
-One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
-We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
-We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified.
-Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphid species at the European level.
-
-**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
-The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
-
-**KEYWORDS**\
-\begingroup
-\small
-Angiosperms, connectance, flowering plants, nestedness, plant-pollinator networks, pollinators, species interactions 
-\endgroup
-\pagebreak
-
-# 1 | INTRODUCTION
-
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species status, trends, and conservation [@heberling2021; @zattara2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano2003] or nestedness [@bascompte2003].
-Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
-Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @garcia2024].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
-Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness.
-This emphasizes the importance of providing data in its rawest possible form in datasets that will be utilized for synthesis and macro-ecological studies.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023], along with occurrence data [@zattara2021], is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, generating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level.
-For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
-Assembling and curating the existing information on EU plant-pollinator networks will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on plant-pollinator interactions at the European level.
-The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
-
-EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks.
-For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where specialists species interact only with subsets of generalists species, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
-
-Overall, EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
-This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
-We expect that EuPPollNet can be used to evaluate macro-ecological processes in plant-pollinator networks, guide conservation planning, and set a baseline for global change research.
-
-# 2 | METHODS
-
-## Data acquisition
-
-The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
-As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022.
-Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024.
-These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
-While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019].
-This approach maximized the potential number of studies that could be incorporated in this database.
-The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*".
-To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).
-
-## Dataset description
-
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations).
-The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.
-Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-Species that belong to other orders `r format(round(100-spp_percent_main_order, 2), nsmall = 2)`% are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
-Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database.
-However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
-Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have selected only a single location per study.
-(\\textbf{b}) Number of studies by year in EuPPollNet.
-(\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}.
-The total number of species and interactions is indicated in parentheses at the bottom."}
+#Load data
+data = readRDS("../../Data/3_Final_data/Interaction_data.rds")
 
@@
                 Day = lubridate::day(Date))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p2  = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-   axis.ticks.x = element_line(color="black"),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -5))
-
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-#Calculate spp number and interactions separately
-data1_spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
-
-data1_int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))
-
-data1 = left_join(data1_spp, data1_int)
-
-#Select main orders
-spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
-pull()
-#Select main orders
-int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Interactions = length(Pollinator_accepted_name)) %>% 
-pull()
-
-#Combine Interactions and Species into a single column
-data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
-
-#Set order of levels for plotting
-data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))
-
-#Plot
-p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
-geom_bar(stat = "identity", width = 0.7) +
-theme_classic() +
-scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator\n order") +
-xlab(NULL) +
-ylab("Percentage") + 
-scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
-theme(axis.text = element_text(size=6), 
-     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
-     axis.title.x = element_text(face="bold", size = 7),
-     legend.position = "top",
-     legend.title = element_text(size = 5, face = "bold"),
-     legend.text = element_text(size = 5),
-     legend.key.size = unit(3, "mm"),
-     legend.margin=margin(0,2,0,0),
-     legend.box.margin=margin(-9,-9,-6,-6)) +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))+
-ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -6)) +
-scale_x_discrete(labels = c("Species\n (2,172)", "Interactions\n (358,049)"))
-
-#plot2 = p3 / p1 / p2 
-plot2 = p2 / p3
-
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.485, 0.8))
-
-```
-
-\newpage
-
-## Data structure
-
-The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble) + 1` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
-These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29).
-The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30).
-Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies.
-To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year.
-Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.
-
-\newpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.} Column names and their descriptions within the EuPPollNet database.") %>%
-  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
-kable_styling(position = "center")
-
-```
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and harmonized in R using **rgbif** [@chamberlain2022].
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-In addition, taxonomic information at genus, family and order level was downloaded for each species.
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023], and butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-Coleoptera species names were only checked against the GBIF checklist.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the EuPPollNet database at the European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only plants occurring in Europe and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we evaluated the taxonomic coverage of taxonomic groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
-While there is not a good understanding of pollinator diversity in other taxonomic groups, it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
-Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
-Consequently, we provide an estimate for the total number of pollinators across the European continent.
-
-Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
-The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it using the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**.
-For plants, the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of the **rtree** package [@li2023].
-
-## Sampling coverage
-
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks.
-In addition, we computed the  accumulation curve of pollinator species with an increasing number of plant species as an indicator for how many pollinator species are likely responsible for the pollination of flowering plants [e.g., @kleijn2015 for crops].
-The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
-The different rarefied curves were complemented with 100 bootstrapped accumulation curves.
-
-## Habitat type and bioclimatic region
-
-We describe the habitat type for each site using information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022], visual inspection of Google Earth imagery and the habitat classification from the authors.
-These different habitat categories (see definitions in supplementary text) allow a quick comparison and understanding of the habitat types from the database.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned a biogeographical region to each site.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites using a spatial joint from the **sf** package [@pebesma2018].
-
-## Network analyses
-
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network, and examined how these network metrics change across latitude of studies and bioclimatic regions in Europe.
-We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
-We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics.
-As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-The relationship between residual connectance and species richness was investigated using a beta-regression implemented with the package **betareg** [@cribari2010].
-We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
-This metric was calculated using the **maxnodf** package [@hoeppke2021].
-Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, we quantified the association between connectance and nestedness with the number of species per network using a Kendall rank correlation coefficient to compare the strength of associations between network structures and species richness across both network metrics.
-
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
-The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008].
-These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
-Both connectance and nestedness (NODF) were estimated for each network by using the function _networklevel_ from **bipartite**.
-
-\newpage
-
-# 3 | RESULTS
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Taxonomic coverage outputs
-#Find number of plant and pollinator species
-##Pollinators
-poll_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-##Plants
-plant_spp = data %>% 
-select(Plant_rank, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-nrow(.)
-
-#Overall plant coverage
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Average plant family coverage
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData") %>% 
-filter(fam_group == "B")
-#Average coverage per family
-mean_plant_fam_cov = d1 %>% 
-ungroup() %>% 
-summarise(mean_plant_fam_cov = mean(Percent)) %>% 
-pull()
-
-#Percent of families with interaction records
-all_fam = nrow(d1)
-fam_with_data = d1 %>% 
-filter(Percent > 0) %>% 
-nrow()
-percent_plant_fam_with_data = fam_with_data/all_fam * 100
-
-#Calculate phylogenetic signal of presence-absence information
-#Read data and plot it
-tree = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-#Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-#select(label, categorical) %>% 
-ungroup() 
-#Check phylogenetic signal
-d1 = d1 %>% arrange(factor(label, levels = tree@phylo$tip.label))
-
-v = phylo_trait$categorical
-names(v) = tree@phylo$tip.label
-
-phylo_signal_plant = phylosig(tree@phylo, v, method="lambda", test=TRUE)
-lambda_phylosig_plants = phylo_signal_plant$lambda
-p_phylosig_plants = phylo_signal_plant$P
-
-#Load pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-#Check percentage of interactions from bees with and without apis
-#Check percentage of bees in the database
-bee_fam = c("Apidae", "Megachilidae", "Halictidae",
-            "Andrenidae", "Colletidae", "Melittidae")
-all_rows = data %>% nrow(.)
-#
-bee_rows = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-nrow(.)
-#Bee interactions
-bee_interactions_percent = bee_rows/all_rows *100
-#Eclude now apis
-bee_rows_non_apis = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-#
-bee_interactions_non_apis_percent = bee_rows_non_apis/bee_rows *100
-
-#Coverage across bee genera
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-all_rows_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-nrow(.)
-#
-data_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-filter(Percent > 0) %>% 
-nrow(.)
-#
-bee_genera_percent_coverage = data_bees/all_rows_bees * 100
-#Check average coverage across bee genera
-average_coverage_bee_genera = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-ungroup() %>% 
-summarise(average_coverage_bee_genera = mean(Percent)) %>% 
-pull()
-
-#Calculate phylogenetic signal of presence-absence information
-##Read data and plot it
-tree = readRDS("../Data/Manuscript_info/bee.tree100.rds")
-##Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-ungroup() 
-##Conduct left join to tibble of tip labels to organise order
-v = tibble(Genus1 = tree$tip.label) %>% 
-mutate(Genus1 = str_to_title(word(Genus1,1,sep = "_")))
-#
-v1 = left_join(v, phylo_trait)
-#
-v2 = v1$categorical
-names(v2) = tree$tip.label
-#
-phylo_signal_bees = phylosig(tree, v2, method="lambda", test=TRUE)
-lambda_phylosig_bees = phylo_signal_bees$lambda
-p_phylosig_bees = phylo_signal_bees$P
-
-#Check number of studies by habitat
-studies_by_habitat = data %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of sampling sites by habitat
-networks_by_habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of studies by bioregion
-studies_by_bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check number of sampling sites by bioregion
-networks_by_bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check percentage of Apis mellifera across networks
-apis_percent_networks = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>%
-group_by(Study_id) %>% 
-mutate(Apis = "Yes") %>% 
-distinct(Study_id, Apis) %>% 
-nrow() / nlevels(factor(data$Study_id)) *100
-
-```
-
-## Taxonomic coverage
-
-Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that benefit from animal pollination.
-EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
-The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
-The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic and taxonomic coverage of the bee genera at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Dark grey bars indicate the total number of species per genus at the European level."}
-
-#Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
-
-#Circular layour
-ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.065, size=2.9, fontface=1) +
-scale_size(range = c(0,4.5), guide = 'none')  +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-guides(fill=guide_legend(), size = guide_legend()) +
-theme(legend.position = "bottom") +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
-                    pwidth=0.13, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0.148,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-guides(fill = "none")+
-scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
-geom_fruit(geom=geom_bar,
-                    mapping=aes(y=label, x=Spp_number_Europe),
-                    pwidth=0.12, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = -0.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
-new_scale_fill() +
-geom_point2(aes(subset=node==131), color='orange', size=3) +
-geom_point2(aes(subset=node==84), color='orange', size=3) +
-geom_point2(aes(subset=node==73), color='orange', size=3) +
-geom_point2(aes(subset=node==74), color='orange', size=3) +
-geom_point2(aes(subset=node==111), color='orange', size=3)+
-geom_point2(aes(subset=node==91), color='orange', size=3) +
-annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
-annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
-annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
-annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
-annotate("text", x = 2.55, y = 28.5, label = "Apidae", size= 3)+
-annotate("text", x = 2.25, y = 51.2, label = "Megachilidae", size= 3) +
-coord_cartesian(clip = "off")+
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.6,0),
-      legend.direction = "horizontal")
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-Phylogenetic and taxonomic coverage of the plant families at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Sampling coverage outputs
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Coverage
-poll_sampling_coverage = poll_output$DataInfo$SC *100
-plant_sampling_coverage = plant_output$DataInfo$SC *100
-int_coverage = int_output$DataInfo$SC *100
-plant_poll_coverage = plant_poll_output$DataInfo$SC *100
-
-#Increase doubling effort
-##Pollinators
-poll_increase = poll_output$iNextEst$Network_id
-poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
-poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
-poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
-##Plants
-plant_increase = plant_output$iNextEst$Network_id
-plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
-plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
-##Interactions
-int_increase = int_output$iNextEst$Network_id
-int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
-int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
-int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
-##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$Plant_id
-plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
-plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
-
-#Prepare results from common and rare poll across sampling sites
-#Pollinators
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-poll_total = nrow(poll_spread)
-polls_under_one_percent = poll_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-polls_under_one_percent = polls_under_one_percent/poll_total * 100
-#Find most common pollinator
-most_common_poll = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Pollinator_accepted_name)
-#Find most common poll percent
-most_common_poll_percent = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-#Plants
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-plant_total = nrow(plant_spread)
-plants_under_one_percent = plant_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-plants_under_one_percent = plants_under_one_percent/plant_total * 100  
-#Find most common plant
-most_common_plant = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Plant_accepted_name)
-#Find most common plant percent
-most_common_plant_percent = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-```
-
-## Sampling coverage
-
-The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_sampling_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**).
-The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
-When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
-We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**).
-The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their unique pairwise interactions across networks.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
-The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks.
-Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network.
-Note that \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
-
-#Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
-#Create dataset
-d_poll = poll_output$iNextEst$Network_id
-#Check cols
-d_poll$col = d_poll$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           method = as.character(pollinators_curves$iteration), 
-           qD = pollinators_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(method == "observed")
-d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
-
-#Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col),size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_poll1_point, col = "black", size=2) + 
-ylab("Pollinator species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000) +
-ggtitle("(a)") +
-theme(plot.margin = margin(r = 12))
-
-#Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
-#Create dataset
-d_plant = plant_output$iNextEst$Network_id
-#Check cols
-d_plant$col = d_plant$method
-
-#Now create tibble with same structure
-a_plant = tibble(t = plant_curves$Study_sampled, 
-           method = as.character(plant_curves$iteration), 
-           qD = plant_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(method == "observed")
-d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
-
-#Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_plant1_point, col = "black", size=2) + 
-ylab("Plant species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000)+
-ggtitle("(b)")+
-theme(plot.margin = margin(r = 12))
-
-#Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
-#Create dataset
-d_int = int_output$iNextEst$Network_id
-#Check cols
-d_int$col = d_int$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = interactions_curves$Study_sampled, 
-           method = as.character(interactions_curves$iteration), 
-           qD = interactions_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(method == "observed")
-d_int1_lines = d_int1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Unique interactions") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000) +
-ylim(0, 30000)+
-ggtitle("(c)")
-
-#----------------#
-#Pollinators
-#----------------#
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-o1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.03, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
-xlab("Pollinator species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.35) + 
-annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
-annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=200, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3)  +
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
-                ylim=c(0,0.7), clip="off") +
-ggtitle("(e)")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-o2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.03, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
-xlab("Plant species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.35) + 
-annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
-annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=150, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3) + 
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
-                ylim=c(0, 0.7), clip="off") +
-ggtitle("(f)")
-
-#Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Create dataset
-d_pp = plant_poll_output$iNextEst$Plant_id
-#Check cols
-d_pp$col = d_pp$method
-
-#Now create tibble with same structure
-a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           method = as.character(plant_poll_curves$iteration), 
-           qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "collectors")
-
-#Bind datasets 
-d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(method == "observed")
-d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
-
-#Plot
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
-geom_line(aes(linetype = col)) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_pp1_point, col = "black") + 
-ylab("Pollinator species") +
-xlab("Plant species") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(d)") +
-scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
-theme(legend.position = "none")
-
-#Plot all
-p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
-      axis.title = element_text(size=12, face = "bold"),
-      axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-#a/b
-
-```
-
-\newpage
-
-## Habitat type and bioclimatic region
-
-The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
-Overall, the proportion of flower visitors from Lepidoptera and Coleoptera were low across all habitats but Coleopteran flower visitors were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
-Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-Hymenopterans were predominant across all bioclimatic regions and Dipterans were particularly abundant in the Alpine and Atlantic regions.
-Lepidopterans had low proportions across all bioclimatic regions and Coleopterans were only relevant in the Mediterranean region at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S3**) also differed across habitats and bioclimatic regions.
-The habitats sampled by a higher number of studies in the database were `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
-However, the habitats that contain a higher number of sampling sites were `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies were `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and those that contain a higher number of sampling sites were `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can contribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Karise is a very large study that only sampled bumblebees and
-#is providing unrealistic results for the Boreal region
-#Filter out
-#Maybe add to caption?
-data1 = data %>% 
-filter(!Study_id == "13_Karise")
-
-#Check number of studies by habitat
-habitat = data1 %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data1 %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(EuPPollNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(EuPPollNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 7),
-      axis.text = element_text(size=6),
-      panel.border = element_rect(size=1)) +
-xlab(NULL) +
-ylab("Proportion of species") +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 10, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = viridis(4),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-      axis.ticks.x = element_line(color="black"),
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 7),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black'),
-plot.margin = margin(0,0,0,0)) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-#Plot panel
-p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Outputs calculated in script 7_3
-#Load lm output connectance ~ latitude
-connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
-connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
-#Load lm output nestedness ~ latitude
-normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
-normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
-#Load values of connectance
-min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
-max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
-mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
-#Load values of nestedness
-min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
-max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
-mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
-#Load kendall corr from connectance and spp
-connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
-connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
-#Load kendall corr from nestedness and spp
-normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
-normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
-#Load z values
-#Save data
-z_percent = readRDS("../Data/Manuscript_info/z_percent.rds")
-no_diff_z_percent = z_percent %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent = z_percent %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent = z_percent %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-#Load correlation outputs of residual connectance and nestednes
-resid_connectance_nestedness_corr_tau = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_tau.rds")
-resid_connectance_nestedness_corr_pval = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_pval.rds")
-
-```
-
-## Network properties
-
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
-We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**).
-Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
-Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes.
-Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
-Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
-Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
-The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
-Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).
-Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
-The bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
-
-# Load libraries
-library(dplyr)
-library(ggplot2)
-library(stringr)
-library(patchwork)
-
-# Load plots generated in scripts 7_2_Nestedness and 7_2_Connectance
-
-connectance_species = readRDS("../Data/Manuscript_info/Figure6_connectance_species.rds")
-connectance_latitude = readRDS("../Data/Manuscript_info/Figure6_connectance_latitude.rds")
-nestedness_z_scores = readRDS("../Data/Manuscript_info/nestedness_z_scores.rds")
-nestedness_latitude = readRDS("../Data/Manuscript_info/nestedness_latitude.rds")
-
-#Prepare first connectance plots
-library(betareg)
-model = betareg(Connectance ~ log_geometric_mean_spp, data = connectance_species)
-p1 = ggplot(connectance_species, aes(x = log_geometric_mean_spp, y = Connectance))+
-geom_point(pch = 21, fill = "azure3", size=2.5) +
-geom_line(aes(y = predict(model, connectance_species)), colour= "black")+
-xlab("log(Species)") +
-theme_bw() +
-ggtitle("(a)")
-
-p3 = connectance_latitude %>% 
-ggplot(aes(Latitude, residual_conectance)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = connectance_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
-geom_ribbon(data = connectance_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ylab("Residual connectance") +
-ggtitle("(c)")
-
-#Prepare now nestedness plots
-# Plot
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-nestedness_z_scores = nestedness_z_scores %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-p2 = nestedness_z_scores %>% 
-ggplot(aes(x = z_score)) +
-geom_histogram(aes(y = ..density.., fill = infra_over_represented), 
-               bins = 100, alpha = 0.5, color = "black", position = "stack") +
-stat_function(data = nestedness_z_scores, fun = function(x) dnorm(x, mean = mean(nestedness_z_scores$z_score), sd = sd(nestedness_z_scores$z_score)) * 1.7, 
-              n = 1000, inherit.aes = FALSE, color = "gray18", size = 1) +
-theme_bw() +
-coord_cartesian(expand = FALSE) +
-geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-ylab("Density") +
-xlab("Z scores (NODF)")  +
-xlim(-6, 9) +
-ylim(0, 0.8) +
-scale_fill_manual(name = "Observed against null",
-                  limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                  labels = c("Less nested", "No difference", "More nested"),
-                  values = c("Under-represented" = "coral2", 
-                             "No statistical difference" = "palegreen3", 
-                             "Over-represented" = "cyan3")) +
-ggtitle("(b)")
-
-# 2nd Plot
-p4 = nestedness_latitude %>% 
-ggplot(aes(Latitude, Normalised_nestedness)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-theme_bw() +
-ylab("NODFc") +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
-geom_ribbon(data = nestedness_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ggtitle("(d)")
-
-# Layout plots
-#a = (p1 + p2) 
-#b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.position = "right")
-#a / b
-
-# Layout plots
-a = (p1 + p2)  & theme(legend.position = "right", legend.justification = "left")
-b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.justification = "left")
-a / b  & theme(axis.title =  element_text(size=16, face="bold"),
-               plot.title = element_text(size=22),
-               panel.border = element_rect(colour = "black",
-               fill=NA, size=1),
-               axis.text=element_text(size=12))
-
-```
-
-\newpage
-
-# 3 | DISCUSSION
-
-EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level.
-The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records.
-While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-This likely reflects the fact that most plant and pollinator species are rare and geographically restricted.
-For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
-This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
-In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
-
-Bees are responsible for the majority of the sampled interactions at the metaweb level.
-As not all surveys included all pollinator groups, this result may partly be influenced by the taxonomic groups sampled across studies.
-However, the relevance of bees and other pollinator orders for network topology changed across habitats and bioclimatic regions in accordance to the literature.
-For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
-These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only commonly documented as floral visitors in the Mediterranean region.
-This study cannot determine whether pollination ecologists traditionally document flower-beetle interactions only in the Mediterranean, or if there are fewer flower visitations by beetles outside this region.
-Nevertheless, the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
-The number of butterfly species and interactions were relatively low compared to the other taxa.
-While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
-This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of intensive grasslands habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practiced in Europe [@steffan2000; @magrach2017; @herrera2020].
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants.
-This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
-Rare pollinators are functionally important for plant species at large scales [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to observe these rare species and to effectively understand and protect plant-pollinator biodiversity.
-
-Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
-Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes.
-Finally, one third of networks were more nested than expected by chance.
-While this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result also suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks [@payrato2019].
-
-Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
-This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-This lack of data is especially relevant for Eastern Europe which has vast landscapes of high quality semi-natural grasslands but is experiencing rapid land use change [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
-These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], and the introduction of alien species [@vila2009; @vanbergen2018].
-Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
-This will allow local and large scale analyses of the status and trends of plant-pollinator communities, effectively informing management and conservation actions.
-
-In conclusion, the EuPPollNet database enables researchers to explore spatial, taxonomic and structural properties of plant-pollinator networks within Europe.
-In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
-Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
-These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
-The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and trends of plant-pollinator communities.
-Finally, we hope this database becomes an iterative resource that keeps growing and improving over time to better understand and conserve European biodiversity.
-
-\newpage
-
-# REFERENCES
-
-::: {#refs}
-:::
-
-# ACKNOWLEDGEMENTS
-
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data.
-
-# FUNDING INFORMATION
-
-This research was funded by the H2020 European project Safeguard (101003476) and by the Federal State of Saxony-Anhalt (MLU-BioDivFund).
-RT was supported by the Czech Science Foundation (Project No.
-21-24186M).
-NDM, NH, YP and FM were financially supported by the ANR ARSENIC project (grant no.
-14-CE02-0012), the ANR NGB project (grant no.
-17-CE32- 011), the Region Nord-Pas-de-Calais, the CNRS, the French Ministère de l'Enseignement Supérieur et de la Recherche, the Hauts-de-France Region and the European Regional Funds.
-
-# CONFLICT OF INTEREST
-
-None.
-
-# DATA AVAILABILITY
-
-All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
-
-\newpage
-
-# SUPPORTING INFORMATION
-
-**Title**: EuPPollNet: A European database of plant-pollinator networks
-
-**Authors**: `r authors`
-
-**Contains**:
-
-- Supplementary text 1
-
-- Figure S1
-
-- Figure S2
-
-- Figure S3
-
-- Figure S4
-
-\newpage
-
-**Supplementary text 1**
-
-Habitat definitions:
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-
-2) **Agricultural margins**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an  urban setting. Botanical gardens are included in this category.
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-
-5) **Forest/woodland understory**: Any plant community sampled under a wooded group of plants.
-The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included in this category agro-forestry areas and open to dense forest.
-Note that we have excluded from this category forest that contains sclerophyllous vegetation.
-
-6) **Semi-natural grassland**: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
-
-7) **Intensive grassland**: Any type of low growing plant community that is highly influenced by human disturbance.
-For instance, agriculture, mowing, moderate to high grazing or urban environments.
-Note that this category also includes old pastures with regrowth of woody vegetation.
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have included in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-
-10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located at high elevation areas within Europe.
-
-12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure S1}. Location of all networks in the EuPPollNet database coloured by bioregion."}
-
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
 
@@
 ```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S2}.
-Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
-
-#Prepare figure for supplementary material
-#Coverage across high taxonomic ranks of bees, syrphids and butterflies
-#--------------------------------------------------------------------------------------------#
-#Start from bees----
-#Bee families
-bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
-            "Andrenidae", "Colletidae","Melittidae")
-
-#Species per family within SafeNet
-bee_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-filter(Pollinator_family %in% bee_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-bee_spp_europe =  master_list %>% 
-select(Family, Genus, Species, GenusAndSpecies) %>% 
-filter(Family %in% bee_fam) %>% 
-group_by(Family) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-rename(Pollinator_family = Family) %>% 
-mutate(Group = "Europe")
-
-#Bind both datasets for plotting
-bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
-
-p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-ylab("Number of species") +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Bee families")
-#--------------------------------------------------------------------------------------------#
-#Continue with syrphids----
-#As we don't have subfamilies, let's obtained them from the master list
-syrphids = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
-rename(Pollinator_accepted_name = GenusAndSpecies)
-#Syrphid species safenet
-syrphid_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_family== "Syrphidae") %>% 
-distinct()
-#Add subfamily info
-syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
-#Summarise spp per subfamily
-syrphid_spp_safenet = syrphid_spp_safenet %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Get spp per subfamily at european level
-syrphid_spp_europe = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
-#Plot
-# Plot for syrphid subfamily coverage with adjusted margins
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily, -Spp_number), Spp_number, fill = Group)) +
-  geom_bar(stat="identity", width = .5, position = "dodge", color = "black") +
-  coord_cartesian(expand = FALSE) +
-  xlab(NULL) +
-  ylab(NULL) +
-  theme_bw() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    panel.border = element_rect(size = 1),
-    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 10)),
-    axis.title = element_text(face = "bold", size = 14),
-    axis.text = element_text(face = "bold", size = 12),
-    plot.margin = margin(10, 10, 20, 10)  # Adjusting the margins
-  ) +
-  scale_fill_manual(values = c("tan2", "lightgrey")) +
-  ggtitle("Syrphid subfamilies")
-#--------------------------------------------------------------------------------------------#
-#Continue with butterflies----
-#Do it at family level
-#Create vector with butterfly families
-butt_fam = c(unique(master_list1$family))
-#Species per family within SafeNet
-butterfly_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Lepidoptera") %>% 
-filter(Pollinator_family %in% butt_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-butterfly_spp_europe =  master_list1 %>% 
-select(family, genus, scientificName) %>% 
-group_by(family) %>% 
-summarise(Spp_number = n_distinct(scientificName)) %>% 
-rename(Pollinator_family = family) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
-#Plot
-p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Butterfly families")
-
-#--------------------------------------------------------------------------------------------#
-p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 14, face = "bold"),
-      axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 12))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S3}.
-Barplot indicating the number of networks by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat_bioregion = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat_bioregion = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-c = bind_rows(habitat, bioregion)
-
-ggplot(c, aes(Habitat_bioregion, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_bw()+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")),
-        space = "free", scales = "free_y") +
-ylab("Networks") +
-xlab(NULL) +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.x = element_line(color="black"),
-      strip.text.x = element_text(size=14, face = "bold"),
-      strip.background = element_rect(size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 14),
-axis.text = element_text(size=10))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S4}.
-Association between nestedness (NODFc) and the geometric mean of plant and pollinator species per network", fig.width=3, fig.height=3}
-
-metrics_by_network_spp_number = readRDS("../Data/Working_files/metrics_by_network_spp_number.rds")
-
-metrics_by_network_spp_number %>% 
-ggplot(aes(geometric_mean_spp, Normalised_nestedness)) +
-geom_point(pch = 21, fill = "azure3", size=1.5, stroke=0.25, alpha=1) +
-theme_bw() +
-xlab("Species") +
-ylab("Normalised nestedness (NODFc)") +
-theme(
-    axis.title = element_text(face = "bold", size = 10),
-    axis.text = element_text(size = 8))
-
-```

### 2024-08-28 Manuscript/EuPPollNet_Ms.qmd
minor edit
edits

@@
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
+
 This likely reflects the fact that most plant and pollinator species are rare and geographically restricted.
@@
 This study cannot determine whether pollination ecologists traditionally document flower-beetle interactions only in the Mediterranean, or if there are fewer flower visitations by beetles outside this region.
-Nevertheless, the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
+Nevertheless, the high proportion of beetles as floral visitors provides further support for their potential role as pollinators in the Mediterranean [@herrera2019; @leon2022].
 The number of butterfly species and interactions were relatively low compared to the other taxa.
@@
 Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of intensive grasslands habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practiced in Europe [@steffan2000; @magrach2017; @herrera2020].
+The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018], potentially reflecting the dominance of intensive grasslands habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practiced in Europe [@steffan2000; @magrach2017; @herrera2020].
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure S1}. Location of all networks in the EuPPollNet database coloured by bioregion."}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure S1}. Geographical location of all networks in the EuPPollNet database coloured by bioregion."}
 

### 2024-09-04 Manuscript/EuPPollNet_Ms.qmd
affiliations

@@
 The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*".
-To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).
+To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a floral visitor, referred to also as pollinators despite not evaluating pollination efficiency, and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).
 
@@
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-
-This likely reflects the fact that most plant and pollinator species are rare and geographically restricted.
+However, although our study likely has higher pollinator coverage, as we did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
 For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.

### 2024-09-04 Scripts/1_Processing_networks/4_Marini.R

@@ -6,6 +6,7 @@
 library(tidyverse)
 #Load function to unify structure of data
 source("Scripts/Processing/Functions/Change_str.R")
+source("Scripts/Processing/Functions/Empty_templates.R") #Read empty templates to compare with
 
 #Prepare interaction data ----
 data = read.csv("Data/1_Raw_data/4_5_6_Marini/Interaction_data.csv")
@@ -139,7 +140,7 @@ Metadata1 = as.data.frame(t(Metadata1)) %>%
 Authorship1 <- data.frame(
   Coauthor_name = c("Andree Cappellari", "Lorenzo Marini"),
   Orcid = c("0000-0002-6726-1323", "0000-0001-7429-7685"),
-  E_mail = c("andree.cappellari@phd.unipd.it", "lorenzo.marini@unipd.it"))
+  E_mail = c("andree.cappellari@unipd.it", "lorenzo.marini@unipd.it"))
 
 #Save data ----
 #Create list with all dataframes of interest

### 2024-09-05 Manuscript/EuPPollNet_Ms.qmd
running everything with the most updated data

--- 2024-08-28 Manuscript/EuPPollNet_Ms.qmd
@@
 The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*".
-To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).
+To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though the pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
 
@@
 Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we evaluated the taxonomic coverage of taxonomic groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
-While there is not a good understanding of pollinator diversity in other taxonomic groups, it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+For pollinators, we only evaluated the taxonomic coverage of groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
+While there is not a good understanding of pollinator diversity in other taxonomic groups (e.g., beetles), it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
 Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
@@
 
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network, and examined how these network metrics change across latitude of studies and bioclimatic regions in Europe.
+To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network in binary form (i.e., presence-absence) and examined how these network metrics change across different latitudes and bioclimatic regions in Europe.
 We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
@@
 As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-The relationship between residual connectance and species richness was investigated using a beta-regression implemented with the package **betareg** [@cribari2010].
+The relationship between residual connectance and species richness was investigated using beta regression.
+This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1.
+The model was implemented using the betareg package [@cribari2010].
 We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
@@
 Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, we quantified the association between connectance and nestedness with the number of species per network using a Kendall rank correlation coefficient to compare the strength of associations between network structures and species richness across both network metrics.
+In addition, to quantify how connectance and nestedness change with network size, we determined their association with the number of species per network using the Kendall rank correlation coefficient.
 
@@
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-
-This likely reflects the fact that most plant and pollinator species are rare and geographically restricted.
+However, although our study likely has a higher pollinator coverage, as we did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
 For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.

### 2024-09-05 Manuscript/EuPPollNet_Ms.qmd
check why this curve doesn't match now

@@
 ##Pollinators
-poll_increase = poll_output$iNextEst$Network_id
-poll_increase1 = poll_increase %>% filter(method == "observed") %>% pull(qD)
+poll_increase = poll_output$iNextEst$coverage_based
+poll_increase1 = poll_increase %>% filter(Method == "Observed") %>% pull(qD)
 poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
@@
 ##Plants
-plant_increase = plant_output$iNextEst$Network_id
-plant_increase1 = plant_increase %>% filter(method == "observed") %>% pull(qD)
+plant_increase = plant_output$iNextEst$coverage_based
+plant_increase1 = plant_increase %>% filter(Method == "Observed") %>% pull(qD)
 plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
@@
 ##Interactions
-int_increase = int_output$iNextEst$Network_id
-int_increase1 = int_increase %>% filter(method == "observed") %>% pull(qD)
+int_increase = int_output$iNextEst$coverage_based
+int_increase1 = int_increase %>% filter(Method == "Observed") %>% pull(qD)
 int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
@@
 ##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$Plant_id
-plant_poll_increase1 = plant_poll_increase %>% filter(method == "observed") %>% pull(qD)
+plant_poll_increase = plant_poll_output$iNextEst$coverage_based
+plant_poll_increase1 = plant_poll_increase %>% filter(Method == "Observed") %>% pull(qD)
 plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
@@
 #Create dataset
-d_poll = poll_output$iNextEst$Network_id
+d_poll = poll_output$iNextEst$coverage_based
 #Check cols
-d_poll$col = d_poll$method
+d_poll$col = d_poll$Method
 
@@
 a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           method = as.character(pollinators_curves$iteration), 
+           Method = as.character(pollinators_curves$iteration), 
            qD = pollinators_curves$Unique_spp_cumulative,
-           col = "collectors")
+           col = "Randomizations")
 
@@
 d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(method == "observed")
-d_poll1_lines = d_poll1 %>%  filter(!method == "observed")
-
+d_poll1_point = d_poll1 %>%  filter(Method == "Observed")
+d_poll1_lines = d_poll1 %>%  filter(!Method == "Observed")
+
+#Fix levels
+d_poll1_lines = d_poll1_lines %>% 
+mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
+                       Method == "Extrapolation" ~ "Extrapolated",
+                       T ~ col))
 #Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
 #Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = method, color = col)) +
+p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = Method, color = col)) +
 geom_line(aes(linetype = col),size = 0.65) +
@@
 #Create dataset
-d_plant = plant_output$iNextEst$Network_id
+d_plant = plant_output$iNextEst$coverage_based
 #Check cols
-d_plant$col = d_plant$method
-
+d_plant$col = d_plant$Method
 #Now create tibble with same structure
 a_plant = tibble(t = plant_curves$Study_sampled, 
-           method = as.character(plant_curves$iteration), 
+           Method = as.character(plant_curves$iteration), 
            qD = plant_curves$Unique_spp_cumulative,
-           col = "collectors")
+           col = "Randomizations")
 
@@
 d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(method == "observed")
-d_plant1_lines = d_plant1 %>%  filter(!method == "observed")
-
+d_plant1_point = d_plant1 %>%  filter(Method == "Observed")
+d_plant1_lines = d_plant1 %>%  filter(!Method == "Observed")
+
+#Fix levels
+d_plant1_lines = d_plant1_lines %>% 
+mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
+                       Method == "Extrapolation" ~ "Extrapolated",
+                       T ~ col))
 #Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
+d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
+
 #Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = method, color = col)) +
+p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = Method, color = col)) +
 geom_line(aes(linetype = col), size = 0.65) +
@@
 #Create dataset
-d_int = int_output$iNextEst$Network_id
+d_int = int_output$iNextEst$coverage_based
 #Check cols
-d_int$col = d_int$method
+d_int$col = d_int$Method
 
@@
 a_poll = tibble(t = interactions_curves$Study_sampled, 
-           method = as.character(interactions_curves$iteration), 
+           Method = as.character(interactions_curves$iteration), 
            qD = interactions_curves$Unique_spp_cumulative,
-           col = "collectors")
+           col = "Randomizations")
 
@@
 d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(method == "observed")
-d_int1_lines = d_int1 %>%  filter(!method == "observed")
-
+d_int1_point = d_int1 %>%  filter(Method == "Observed")
+d_int1_lines = d_int1 %>%  filter(!Method == "Observed")
+
+#Fix levels
+d_int1_lines = d_int1_lines %>% 
+mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
+                       Method == "Extrapolation" ~ "Extrapolated",
+                       T ~ col))
+#Order levels
+d_int1_lines$col = factor(d_int1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
 #Plot
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = method, color = col)) +
+p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = Method, color = col)) +
 geom_line(aes(linetype = col), size = 0.65) +
@@
 #Create dataset
-d_pp = plant_poll_output$iNextEst$Plant_id
+d_pp = plant_poll_output$iNextEst$coverage_based
 #Check cols
-d_pp$col = d_pp$method
+d_pp$col = d_pp$Method
 
@@
 a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           method = as.character(plant_poll_curves$iteration), 
+           Method = as.character(plant_poll_curves$iteration), 
            qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "collectors")
+           col = "Randomizations")
 
@@
 d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(method == "observed")
-d_pp1_lines = d_pp1 %>%  filter(!method == "observed")
-
+d_pp1_point = d_pp1 %>%  filter(Method == "Observed")
+d_pp1_lines = d_pp1 %>%  filter(!Method == "Observed")
+
+#Fix levels
+d_pp1_lines = d_pp1_lines %>% 
+mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
+                       Method == "Extrapolation" ~ "Extrapolated",
+                       T ~ col))
+#Order levels
+d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
 #Plot
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("collectors", "interpolated", "extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = method, color = col)) +
+pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = Method, color = col)) +
 geom_line(aes(linetype = col)) +
@@
 scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
+scale_y_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2600)) +
 theme(legend.position = "none")

### 2024-09-06 Manuscript/EuPPollNet_Ms.qmd
Fixing bug on sampling coverage, was a minor issue at the end

@@
 #Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_ploll.rds")
+plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_poll.rds")
 plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure S1}. Geographical location of all networks in the EuPPollNet database coloured by bioregion."}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure S1}. Geographical location of all networks in the EuPPollNet database coloured by bioregion."}
 
@@
      legend.title = element_text(face ="bold",  size = 8),
-     legend.spacing.x=unit(-0.5, "lines"), 
+     legend.spacing.x=unit(0.5, "lines"), 
      legend.key.width=unit(0.5, "lines"), 

### 2024-09-06 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
-Rare pollinators are functionally important for plant species at large scales [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to observe these rare species and to effectively understand and protect plant-pollinator biodiversity.
+Rare pollinators can be functionally important for plant species at the landscape level [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to identify these rare species across different regions and to effectively understand and protect plant-pollinator biodiversity.
 

### 2024-09-16 Manuscript/EuPPollNet_Ms.qmd
edits

@@
 This approach maximized the potential number of studies that could be incorporated in this database.
-The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*".
+The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'.
 To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though the pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
@@
 Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-Species that belong to other orders `r format(round(100-spp_percent_main_order, 2), nsmall = 2)`% are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
+Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
 Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database.
@@
 
-The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble) + 1` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
+The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble)` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
 These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29).
@@
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
+
+#Small edit to column number
+my_tibble = my_tibble %>% 
+mutate(Variable = paste0(seq(1:length(Variable)),". ", Variable))
 
@@
 
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network in binary form (i.e., presence-absence) and examined how these network metrics change across different latitudes and bioclimatic regions in Europe.
+To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network and examined how these network metrics change across different latitudes and bioclimatic regions in Europe.
 We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
@@
 As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-The relationship between residual connectance and species richness was investigated using beta regression.
+The relationship between residual connectance and species richness was investigated using a beta regression.
 This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1.
 The model was implemented using the betareg package [@cribari2010].
-We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
+We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species in comparison to other nestedness metrics that change with network size [@song2017].
 This metric was calculated using the **maxnodf** package [@hoeppke2021].
@@
 Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
-The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008].
-These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
-Both connectance and nestedness (NODF) were estimated for each network by using the function _networklevel_ from **bipartite**.
+The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions, avoiding the influence of network size.
+As NODF is a metric computed from binary matrices, we calculated 100 null models for the binarized version of each network using the 'curveball' algorithm [@strona2014].
+This algorithm implements the configuration model (i.e.
+random rewiring of all links, without self-links or double links) and thus keeps the exact number of connections per species (i.e., realised degree).
+Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species.
+Each empirical network was randomized 100 times using each method.
+These null models were implemented with the **vegan** [@oksanen2013] package.
+Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite**.
 
@@
 The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
+Dark grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
 
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their unique pairwise interactions across networks.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and the number of unique pairwise interactions across networks.
 Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
@@
 xlim(0,3000) +
-ylim(0, 30000)+
+ylim(0, 35000)+
 ggtitle("(c)")
@@
 color = "black", size=0.35) + 
-annotate(geom="text", x=1450, y=0.60, label="50th percentile",color="black", size = 3) +
+annotate(geom="text", x=1590, y=0.60, label="50th percentile",color="black", size = 3) +
 annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
@@
           color = "black", size=0.35) + 
-annotate(geom="text", x=950, y=0.60, label="50th percentile",color="black", size = 3) + 
+annotate(geom="text", x=1010, y=0.60, label="50th percentile",color="black", size = 3) + 
 annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
@@
 scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2600)) +
+scale_y_continuous(breaks = seq(0, 3000, by = 1000), limits=c(0,3000)) +
 theme(legend.position = "none")
@@
 #Load z values
-#Save data
-z_percent = readRDS("../Data/Manuscript_info/z_percent.rds")
-no_diff_z_percent = z_percent %>% 
+#VAZNULL
+z_percent_vaznull = readRDS("../Data/Manuscript_info/z_percent_vaznull.rds")
+no_diff_z_percent_vaznull = z_percent_vaznull %>% 
 filter(infra_over_represented == "No statistical difference") %>% 
 pull(percent)
-over_z_percent = z_percent %>% 
+over_z_percent_vaznull = z_percent_vaznull %>% 
 filter(infra_over_represented == "Over-represented") %>% 
 pull(percent)
-under_z_percent = z_percent %>% 
+under_z_percent_vaznull = z_percent_vaznull %>% 
 filter(infra_over_represented == "Under-represented") %>% 
 pull(percent)
+#CURVEBALL
+z_percent_curveball = readRDS("../Data/Manuscript_info/z_percent_curveball.rds")
+no_diff_z_percent_curveball = z_percent_curveball %>% 
+filter(infra_over_represented == "No statistical difference") %>% 
+pull(percent)
+over_z_percent_curveball = z_percent_curveball %>% 
+filter(infra_over_represented == "Over-represented") %>% 
+pull(percent)
+under_z_percent_curveball = z_percent_curveball %>% 
+filter(infra_over_represented == "Under-represented") %>% 
+pull(percent)
+
+#QUASISWAP
+z_percent_quasiswap = readRDS("../Data/Manuscript_info/z_percent_quasiswap.rds")
+
+no_diff_z_percent_quasiswap = z_percent_quasiswap %>% 
+filter(infra_over_represented == "No statistical difference") %>% 
+pull(percent)
+over_z_percent_quasiswap = z_percent_quasiswap %>% 
+filter(infra_over_represented == "Over-represented") %>% 
+pull(percent)
+under_z_percent_quasiswap = z_percent_quasiswap %>% 
+filter(infra_over_represented == "Under-represented") %>% 
+pull(percent)
+#rd2
+z_percent_rd2 = readRDS("../Data/Manuscript_info/z_percent_rd2.rds")
+
+no_diff_z_percent_rd2 = z_percent_rd2 %>% 
+filter(infra_over_represented == "No statistical difference") %>% 
+pull(percent)
+over_z_percent_rd2 = z_percent_rd2 %>% 
+filter(infra_over_represented == "Over-represented") %>% 
+pull(percent)
+under_z_percent_rd2 = z_percent_rd2 %>% 
+filter(infra_over_represented == "Under-represented") %>% 
+pull(percent)
+
 #Load correlation outputs of residual connectance and nestednes
@@
 Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
-We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**).
-Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
+Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**).
+The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested.
+The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested.
+Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
+(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
 Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes.
@@
 Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
-Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network).
+Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network with the curveball algorithm).
 The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
@@
 
-#Prepare now nestedness plots
-# Plot
+null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
+metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
+
+#Finally, find % of networks that differ statistically
+# Merge and process data
+d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
+
+d = d %>% 
+mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
+d = d %>% filter(!is.na((z_score)))
+
 p = 0.05 #cutoff probability 95% confidence
@@
 
-nestedness_z_scores = nestedness_z_scores %>%
+d1 = d %>%
 mutate(infra_over_represented = case_when(
@@
 
-p2 = nestedness_z_scores %>% 
+#Calculate percentages
+z_percent = d1 %>% 
+group_by(infra_over_represented) %>% 
+summarise(percent = length(infra_over_represented)/nrow(.) * 100)
+
+no_dif_z = z_percent %>% 
+filter(infra_over_represented == "No statistical difference")
+
+p2 = d1 %>% 
   ggplot(aes(x = z_score)) +
-geom_histogram(aes(y = ..density.., fill = infra_over_represented), 
-               bins = 100, alpha = 0.5, color = "black", position = "stack") +
-stat_function(data = nestedness_z_scores, fun = function(x) dnorm(x, mean = mean(nestedness_z_scores$z_score), sd = sd(nestedness_z_scores$z_score)) * 1.7, 
-              n = 1000, inherit.aes = FALSE, color = "gray18", size = 1) +
+  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
+                 bins = 60, alpha = 0.5, color = "black", position = "stack") +
   theme_bw() +
@@
   geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-ylab("Density") +
+  ylab("Number of networks") +
   xlab("Z scores (NODF)")  +
-xlim(-6, 9) +
-ylim(0, 0.8) +
+ # xlim(-6, 9) +
+ # ylim(0, 0.8) +
   scale_fill_manual(name = "Observed against null",
@@
                                "Over-represented" = "cyan3")) +
-ggtitle("(b)")
+ggtitle("(b)")+
+xlim(-6,6)
 
@@
 However, although our study likely has a higher pollinator coverage, as we did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
-For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
+For example, given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
 Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
@@
 Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018], potentially reflecting the dominance of intensive grasslands habitats in EuPPollNet, the highly generalised nature of honeybees, the important role that honeybees are playing as pollinators in Europe, and/or because beekeeping is widely practiced in Europe [@steffan2000; @magrach2017; @herrera2020].
+The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018].
+This potentially reflects the dominance of intensive grassland habitats in EuPPollNet, the highly generalised nature of honey bees, their native status and the important role that honeybees are playing as pollinators in Europe, and/or the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
 
@@
 These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes.
-Finally, one third of networks were more nested than expected by chance.
-While this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result also suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks [@payrato2019].
-
-Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
+Finally, most empirical networks (~85%) showed a non-nested structure.
+This result opposes to the idea that plant-pollinator networks tend to be nested [@bascompte2003], and support recent findings showing that nestedness is not a prevalent feature in plant-pollinator networks [@payrato2019].
+
+Despite this database covers a wide range of habitats across `r country_number` countries, it contains temporal and geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
+For instance, none of the studies in this database sampled nocturnal pollinators, which can impact our view of network structure [@garcia2024], and most studies were conducted during a single flowering season, limiting our ability to evaluate temporal trends of plant-pollinator communities in the face of environmental changes [@alarcon2008; @chacoff2018].
+In addition, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
 This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-This lack of data is especially relevant for Eastern Europe which has vast landscapes of high quality semi-natural grasslands but is experiencing rapid land use change [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+The lack of data for Eastern Europe, which contains vast landscapes with semi-natural grasslands experiencing rapid land use change [@sutcliffe2015], and for the Mediterranean region, which is severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023], is particularly concerning.
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
@@
       strip.text.x = element_text(size=14, face = "bold"),
-      strip.background = element_rect(size=0.5, linetype="solid"),
   axis.title = element_text(face = "bold", size = 14),
-axis.text = element_text(size=10))
+axis.text = element_text(size=10),
+    strip.background = element_rect(
+color=NA, fill=NA, size=0.5, linetype="solid"),
+panel.border = element_rect(colour = "black",
+        fill=NA, size=1))
 
@@
 ```
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S5}.
+Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with the 'quasiswap_count' algorithm (100 null models for each network).
+The quasiswap_count algorithm maintains row and column sums with fixed connectance."}
+
+library(dplyr)
+library(ggplot2)
+
+null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
+metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
+
+#Finally, find % of networks that differ statistically
+# Merge and process data
+d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
+
+d = d %>% 
+mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
+d = d %>% filter(!is.na((z_score)))
+
+p = 0.05 #cutoff probability 95% confidence
+critical_value = qnorm(p/2) #double tail probability divide by 2
+
+d1 = d %>%
+mutate(infra_over_represented = case_when(
+z_score < -abs(critical_value) ~ "Under-represented",
+between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
+    z_score > abs(critical_value) ~ "Over-represented"
+  ))
+
+#Calculate percentages
+z_percent = d1 %>% 
+group_by(infra_over_represented) %>% 
+summarise(percent = length(infra_over_represented)/nrow(.) * 100)
+
+no_dif_z = z_percent %>% 
+filter(infra_over_represented == "No statistical difference")
+
+# Plot
+ d1 %>% 
+  ggplot(aes(x = z_score)) +
+  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
+                 bins = 50, alpha = 0.5, color = "black", position = "stack") +
+  theme_bw() +
+  coord_cartesian(expand = FALSE) +
+  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
+  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
+  ylab("Number of networks") +
+  xlab("Z scores (NODFc)")  +
+ # xlim(-6, 9) +
+ # ylim(0, 0.8) +
+  scale_fill_manual(name = "Observed against null",
+                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
+                    labels = c("Less nested", "No difference", "More nested"),
+                    values = c("Under-represented" = "coral2", 
+                               "No statistical difference" = "palegreen3", 
+                               "Over-represented" = "cyan3")) +
+xlim(-6,6) +
+theme(plot.title = element_text(size = 24),
+      text = element_text(size = 20),
+      axis.title  = element_text(face="bold"))
+
+```

### 2024-09-19 Manuscript/EuPPollNet_Ms.qmd
Updating refs and creating script for EUNIS habitats

@@
 nrow(.)
+#Check number of distinct habitats
+n_habitats = data %>% 
+select(EuPPollNet_habitat) %>% 
+distinct() %>% 
+nrow(.)
 #Read author list 
@@
 
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
+**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from `r n_habitats` different natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
 All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
@@
 The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'.
-To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though the pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
+To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
 
@@
 These null models were implemented with the **vegan** [@oksanen2013] package.
-Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite**.
+Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite** [@dormann2008].
 
@@
 The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
+Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation and moors and heathland.
 Overall, the proportion of flower visitors from Lepidoptera and Coleoptera were low across all habitats but Coleopteran flower visitors were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
 Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-Hymenopterans were predominant across all bioclimatic regions and Dipterans were particularly abundant in the Alpine and Atlantic regions.
+Hymenopterans were abundant across all bioclimatic regions and Dipterans were particularly abundant in the Boreal, Alpine and Atlantic regions.
 Lepidopterans had low proportions across all bioclimatic regions and Coleopterans were only relevant in the Mediterranean region at European level.
@@
 Note that a single study can contribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
+Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.
+The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day.", fig.width=4.5, fig.height=8, fig.align='center'}
 
@@
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-However, although our study likely has a higher pollinator coverage, as we did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
+However, although EuPPollNet likely has a higher pollinator coverage, as it did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
 For example, given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
@@
 The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018].
-This potentially reflects the dominance of intensive grassland habitats in EuPPollNet, the highly generalised nature of honey bees, their native status and the important role that honeybees are playing as pollinators in Europe, and/or the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
+This potentially reflects the dominance of intensive grassland habitats in EuPPollNet and their widespread distribution across European landscapes [@isselstein2005], the highly generalised nature of honey bees, their native status and the important role that honeybees are playing as pollinators in Europe, and/or the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
 
@@
 Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes.
+These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes in Europe.
 Finally, most empirical networks (~85%) showed a non-nested structure.
@@
 
-Habitat definitions:
+EuPPollNet habitat definitions adapted from authors and CORINE Land Cover (CLC).
+To guide these definitions, they are connected to habitat types defined in the European Nature Information System (EUNIS 2021; <https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
+Specific examples from EUNIS habitats are provided when they can be precisely linked to EuPPollNet habitat types.
 
 1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
+This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘dry perennial anthropogenic herbaceous vegetation’, code **V38** ).
 
 2) **Agricultural margins**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an  urban setting. Botanical gardens are included in this category.
+This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘hedgerows’, code **V4** ).
+
+3) **Green urban areas**: Parks, private gardens or small pastures within an urban setting.
+Botanical gardens are included in this category.
+This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘cultivated areas of gardens and parks’, code **V2** ).
 
 4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-
-5) **Forest/woodland understory**: Any plant community sampled under a wooded group of plants.
-The forest could be embedded in an agricultural setting or in a fully natural scenario.
-We have included in this category agro-forestry areas and open to dense forest.
-Note that we have excluded from this category forest that contains sclerophyllous vegetation.
-
-6) **Semi-natural grassland**: Low growing plant community with relatively low disturbances but under low pressure such as seasonal mowing or extensive grazing.
-
-7) **Intensive grassland**: Any type of low growing plant community that is highly influenced by human disturbance.
-For instance, agriculture, mowing, moderate to high grazing or urban environments.
-Note that this category also includes old pastures with regrowth of woody vegetation.
+This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘intensive unmixed crops’, code **V11** ).
+
+5) **Forest/woodland understory**: Any plant community sampled within a wooded group of plants.
+The forest may be situated in an agricultural setting or in a fully natural scenario.
+We have included agroforestry areas as well as open to dense forests in this category but excluded forest that contains sclerophyllous vegetation.
+This category is similar to the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
+
+6) **Semi-natural grassland**: Acidic, neutral, or calcareous unimproved grasslands located at elevations lower than 1000 m that are managed extensively through mowing or grazing but have not been improved, i.e., they have never received significant fertilizer or herbicide inputs.
+This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R** ).
+
+7) **Intensive grassland**: Communities dominated by non-woody, low-growing plants that are heavily influenced by human disturbance, such as agriculture, mowing, moderate to high grazing, or urban environments.
+The plant composition is generally dominated by graminoid species.
+This category partially matches the EUNIS categories of ‘vegetated man made habitats’ (code **V**; e.g., category of ‘artificial grasslands and herb dominated habitats’, code **V4** ) and ‘grasslands and lands dominated by forbs, mosses or lichens’ (code **R** ).
 
@@
 Note, that we have included in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
+This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
+
+9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.  This category partially matches the **EUNIS habitat type**  ‘coastal habitats’ (code **N**).
 
 10) **Riparian vegetation**: Plant communities growing on river margins.
-
-10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located at high elevation areas within Europe.
-
-12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
+This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**) but also the category of ‘inland waters’ which is currently under review.
+
+10) **Montane to alpine grasslands**: Communities dominated by non-woody, low-growing plants that experience little or no human disturbance and are found in high-elevation areas (>1,000 m).
+This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R**; e.g., alpine and subalpine grasslands with code  **R4**).
+
+12) **Moors and heathland**: Low-growing woody vegetation (i.e., typically < 5 m) that is characteristic of poorly fertile soils where environmental conditions prevent the natural formation of forests.
+This category is similar to the **EUNIS habitat type** ‘heathland, scrub and tundra’ (code **S**).
 

### 2024-09-19 Scripts/Metaweb_checks/Eunis_checks.R

@@ -1,0 +1,27 @@
+library(sf)
+library(dplyr)
+
+
+#Read data
+data = readRDS("Data/Working_files/Interactions_uncounted.rds")
+
+levels(factor(data$EuPPollNet_habitat))
+
+data = data %>% 
+filter(EuPPollNet_habitat =="Moors and heathland")
+
+#Generate a tibble with coordinates
+coordinates = data %>% 
+select(Longitude, Latitude) %>% 
+distinct()
+
+#Read data
+natura_shapefile <- st_read("natura_2000/SHP/Natura2000_end2022_epsg3035.shp")
+st_crs(natura_shapefile)  # Check the CRS
+
+# Transform to match shapefile CRS
+coordinates_sf <- st_as_sf(coordinates, coords = c("Longitude", "Latitude"), crs = 4326)
+coordinates_sf_transformed <- st_transform(coordinates_sf, crs = st_crs(natura_shapefile))
+# Spatial join to find matching habitats
+habitats_at_coordinates <- st_join(coordinates_sf_transformed, natura_shapefile)
+

### 2024-09-21 Manuscript/EuPPollNet_Ms.qmd
Affiliations update

@@
 Finally, most empirical networks (~85%) showed a non-nested structure.
-This result opposes to the idea that plant-pollinator networks tend to be nested [@bascompte2003], and support recent findings showing that nestedness is not a prevalent feature in plant-pollinator networks [@payrato2019].
+This result contrasts with earlier work suggesting that plant-pollinator networks are typically nested [@bascompte2003] and it aligns with recent evidence showing that nestedness is not a common feature in plant-pollinator networks [@payrato2019].
 
@@
 
-We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data.
+We thank all the taxonomists and ecologists that has made this database possible by contributing with their fieldwork data.
 
@@
 - Figure S4
+
+- Figure S5
 

### 2024-09-21 Manuscript/EuPPollNet_word.qmd
Affiliations update

--- 2024-08-12 Manuscript/EuPPollNet_word.qmd
@@
 
+#Check number of species from Hymenop, Dipte, Lepidop and Coleop
+
+pollinator_spp = data %>% 
+select(Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+distinct(Pollinator_accepted_name) %>% 
+nrow(.)
+
+pollinator_spp_main_orders = data %>% 
+select(Pollinator_order, Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+filter(Pollinator_order == "Hymenoptera"|
+       Pollinator_order == "Diptera"|
+       Pollinator_order == "Lepidoptera"|
+       Pollinator_order == "HymenoColeopteraptera") %>% 
+distinct(Pollinator_accepted_name) %>% 
+nrow(.)
+
+spp_percent_main_order = pollinator_spp_main_orders/pollinator_spp *100
+
+#Calculate percentage of studies that recorded the main 4 orders
+metadata = readRDS("../Data/3_Final_data/Metadata.rds")
+#Some studies are divided by year
+#Get an average of the pollinator percentages across years first
+#Fix NA's
+metadata_study = metadata %>%
+group_by(Study_id) %>% 
+mutate(Hymenoptera_percent = as.numeric(na_if(Hymenoptera_percent, "NA"))) %>%  
+mutate(Diptera_percent = as.numeric(na_if(Diptera_percent, "NA"))) %>%  
+mutate(Lepidoptera_percent = as.numeric(na_if(Lepidoptera_percent, "NA"))) %>%  
+mutate(Coleoptera_percent = as.numeric(na_if(Coleoptera_percent, "NA"))) 
+#Calculate average
+mean_study = metadata_study  %>% 
+summarise(Hymenoptera_percent = mean(Hymenoptera_percent, na.rm = T),
+          Diptera_percent = mean(Diptera_percent, na.rm = T),
+          Lepidoptera_percent = mean(Lepidoptera_percent, na.rm = T),
+          Coleoptera_percent = mean(Coleoptera_percent, na.rm = T)) %>% 
+ mutate(across(c(Hymenoptera_percent, Diptera_percent, Lepidoptera_percent, Coleoptera_percent),
+                ~ ifelse(is.nan(.), NA, .))) 
+#Now calculate percentage of studies that record any of these 4 orders
+mean_percent_order = mean_study %>%
+  summarise(
+    Hymenoptera_percent_metaweb = sum(!is.na(Hymenoptera_percent)) / nrow(.) * 100,
+    Diptera_percent_metaweb = sum(!is.na(Diptera_percent)) / nrow(.) * 100,
+    Lepidoptera_percent_metaweb = sum(!is.na(Lepidoptera_percent)) / nrow(.) * 100,
+    Coleoptera_percent_metaweb = sum(!is.na(Coleoptera_percent)) / nrow(.) * 100)
+
+Hymenoptera_percent_metaweb = mean_percent_order$Hymenoptera_percent_metaweb
+Diptera_percent_metaweb = mean_percent_order$Diptera_percent_metaweb
+Lepidoptera_percent_metaweb = mean_percent_order$Lepidoptera_percent_metaweb
+Coleoptera_percent_metaweb = mean_percent_order$Coleoptera_percent_metaweb
+
+#Check percentages of hymenoptea groups
+hymenoptera_percent_by_group = metadata %>% 
+ungroup() %>% 
+distinct(Study_id, Taxa_recorded) %>% 
+summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
+          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
+          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))
+
+bee_percent = hymenoptera_percent_by_group$Bee_studies
+bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
+all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies
+
+#Check percentages of diptera groups
+diptera_percent_by_group = metadata %>% 
+ungroup() %>% 
+distinct(Study_id, Taxa_recorded) %>% 
+summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
+          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
+          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
+                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)
+
+syrphid_percent = diptera_percent_by_group$Syrphid_studies
+all_diptera_percent = diptera_percent_by_group$Diptera_studies
+other_diptera_percent = diptera_percent_by_group$other_diptera_studies
+
 ```
@@
 **Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., family, order and phylum).
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
+In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).
+
+**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from `r n_habitats` different natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
 All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
@@
 However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide.
+Numerous works have studied plant-pollinator interactions in the last decades, generating thousands of plant-pollinator interaction networks worldwide.
 Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.
+Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level.
 For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
@@
 The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@willemstein1987; @potts2015], accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
 To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
-The overall number of potential pollinators and flowering plants in Europe was estimated with the help of the different checklists and sampling coverage.
-Finally, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
-
-We investigated how two key structural metrics of plant-pollinator networks, connectance and nestedness, change across latitudes and bioclimatic regions.
-Networks that are more nested than expected by chance exhibit greater redundancy in their ecological functions and more stability in the face of species loss [@thebault2010].
-EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of nestedness, which has only been evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019].
-EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
+In addition, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
+
+EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks.
+For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where specialists species interact only with subsets of generalists species, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
+
+Overall, EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
 In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
@@
 
-The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent.
-As this database is the result of one of the working packages of the European project Safeguard, first, data was directly asked from members of the Safeguard project.
-Second, the request was extended to data owners outside of the project.
+The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
+As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022.
+Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024.
 These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
-To maintain high quality standards that will support robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.
+While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019].
+This approach maximized the potential number of studies that could be incorporated in this database.
+The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'.
+To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
 
@@
 
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**).
+The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations).
 The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as the sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a pollinator individual to the reproductive structure of a particular plant.
-Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species in the database.
+Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
+All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
+The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.
+Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
+Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
+Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database.
 However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
-Notably,  _Apis mellifera_ represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Locations of the studies in EuPPollNet showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
+Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure 1}.
+(\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
 The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have focused only on the European region with studies and selected a single location per study.
+For visualization purposes, we have selected only a single location per study.
 (\\textbf{b}) Number of studies by year in EuPPollNet.
-(\\textbf{c}) Number of species and interactions across the four main pollinator orders in EuPPollNet."}
+(\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}.
+The total number of species and interactions is indicated in parentheses at the bottom."}
 
@@
 ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -5))
+theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -5))
 
 main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
+
+#Calculate spp number and interactions separately
+data1_spp = data %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+filter(Pollinator_rank == "SPECIES") %>% 
+group_by(Pollinator_order) %>% 
+summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
+
+data1_int = data %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+#filter(Pollinator_rank == "SPECIES") %>% 
+group_by(Pollinator_order) %>% 
+summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))
+
+data1 = left_join(data1_spp, data1_int)
+
 #Select main orders
-data1 = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
+spp = data %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
 filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.),
-          Interactions = length(Pollinator_accepted_name)/ nrow(.))
-
-# Combine Interactions and Species into a single column
+filter(Pollinator_rank == "SPECIES") %>% 
+summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
+pull()
+#Select main orders
+int = data %>% 
+filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
+select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
+filter(Pollinator_order %in% main_orders) %>% 
+#filter(Pollinator_rank == "SPECIES") %>% 
+summarise(Interactions = length(Pollinator_accepted_name)) %>% 
+pull()
+
+#Combine Interactions and Species into a single column
 data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
@@
      legend.key.size = unit(3, "mm"),
-     legend.margin=margin(0,0,0,0),
-     legend.box.margin=margin(-8,-5,-5,-5)) +
+     legend.margin=margin(0,2,0,0),
+     legend.box.margin=margin(-9,-9,-6,-6)) +
 guides(fill=guide_legend(nrow=2,byrow=TRUE))+
 ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -6))
+theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -6)) +
+scale_x_discrete(labels = c("Species\n (2,172)", "Interactions\n (358,049)"))
 
@@
 
-The EuPPollNet database is available in both .csv and .rds formats.
-The files contain a total of `r nrow(my_tibble) + 1` columns that depict where and when the plant-pollinator interaction was observed (**Table 1**).
-Each row depicts a single interaction between a plant and a pollinator species.
-To build a plant-pollinator network matrix within a single flowering season at a site level, users only need to group interactions by plant and pollinator species, site, study and year.
-Information about habitat type and bioclimatic region are also provided in this file.
-In addition, the interaction dataset includes a column that describes the presence or absence of floral counts for each study.
-The flower count data is provided in a separate file (.csv or .rds) that can be merged with the interaction data through the "Flower_data_merger" column.
-Two thirds of studies contain information on floral abundance, but the methods and units differ greatly across studies.
-Finally, metadata at the study level is provided, including information about the authors, digital object identifier if available, sampling time and taxonomic coverage of the main pollinator groups for each study.
+The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble)` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
+These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29).
+The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30).
+Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies.
+To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year.
+Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.
+
+\newpage
 
 ```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
+
+#Small edit to column number
+my_tibble = my_tibble %>% 
+mutate(Variable = paste0(seq(1:length(Variable)),". ", Variable))
 
@@
 To assess the completeness of plant and pollinator species in the EuPPollNet database at the European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only European flowering plants and excluded taxonomic groups not associated with biotic pollination.
+Specifically for plants, we refined the checklist to include only plants occurring in Europe and excluded taxonomic groups not associated with biotic pollination.
 We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included some exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we compared only the taxonomic coverage of bees, syrphids and butterflies by using their species checklists at the European level.
-The number of pollinator species at the European level was estimated by first estimating coverage for the three taxa that have checklists (bees, syrphids and butterflies).
-The mean coverage across these three groups (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`) was then used for the other taxa for which there is no checklist to extrapolate the number of pollinating species in Europe.
+Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
+For pollinators, we only evaluated the taxonomic coverage of groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
+While there is not a good understanding of pollinator diversity in other taxonomic groups (e.g., beetles), it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
+Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
+Consequently, we provide an estimate for the total number of pollinators across the European continent.
 
@@
 
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network, and examined how these network metrics change across latitude of studies and bioclimatic regions in Europe.
+To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network and examined how these network metrics change across different latitudes and bioclimatic regions in Europe.
 We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
@@
 As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-The relationship between residual connectance and species richness was investigated using a beta-regression implemented with the package **betareg** [@cribari2010].
-We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species [@song2017].
+The relationship between residual connectance and species richness was investigated using a beta regression.
+This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1.
+The model was implemented using the betareg package [@cribari2010].
+We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species in comparison to other nestedness metrics that change with network size [@song2017].
 This metric was calculated using the **maxnodf** package [@hoeppke2021].
 Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, we quantified the association between connectance and nestedness with the number of species per network using a Kendall rank correlation coefficient to compare the strength of associations between network structures and species richness across both network metrics.
+In addition, to quantify how connectance and nestedness change with network size, we determined their association with the number of species per network using the Kendall rank correlation coefficient.
 
 Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
-The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions.
-We calculated 100 null models for each network with the help of the _vaznull_ function from the package **bipartite** [@dormann2008].
-These null networks have the same connectance and number of plant and pollinator species as the empirical ones, but different marginal totals.
-Both connectance and nestedness (NODF) were estimated for each network by using the function _networklevel_ from **bipartite**.
+The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions, avoiding the influence of network size.
+As NODF is a metric computed from binary matrices, we calculated 100 null models for the binarized version of each network using the 'curveball' algorithm [@strona2014].
+This algorithm implements the configuration model (i.e.
+random rewiring of all links, without self-links or double links) and thus keeps the exact number of connections per species (i.e., realised degree).
+Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species.
+Each empirical network was randomized 100 times using each method.
+These null models were implemented with the **vegan** [@oksanen2013] package.
+Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite** [@dormann2008].
 
@@
 
-Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that rely on animal pollination.
+Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that benefit from animal pollination.
 EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinators groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S1** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
+The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
+Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
 Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
@@
                     color="black",
-                    offset = 0.137,
+                    offset = 0.148,
             axis.params=list(
@@
          grid.params=list()) +
-annotate("text", x = 14.2, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.1, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
+annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
+annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
+annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
 new_scale_fill() +
-geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
-annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
-annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
-annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
-annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
-annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
-annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
+geom_point2(aes(subset=node==131), color='orange', size=3) +
+geom_point2(aes(subset=node==84), color='orange', size=3) +
+geom_point2(aes(subset=node==73), color='orange', size=3) +
+geom_point2(aes(subset=node==74), color='orange', size=3) +
+geom_point2(aes(subset=node==111), color='orange', size=3)+
+geom_point2(aes(subset=node==91), color='orange', size=3) +
+annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
+annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
+annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
+annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
+annotate("text", x = 2.55, y = 28.5, label = "Apidae", size= 3)+
+annotate("text", x = 2.25, y = 51.2, label = "Megachilidae", size= 3) +
 coord_cartesian(clip = "off")+
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and their interactions across networks.
+Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and the number of unique pairwise interactions across networks.
 Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
@@
 scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 2600, by = 1000), limits=c(0,2600)) +
+scale_y_continuous(breaks = seq(0, 3000, by = 1000), limits=c(0,3000)) +
 theme(legend.position = "none")
@@
 The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-As expected, Hymenoptera was the main taxonomic order on the majority of habitats, exceeded only by Diptera on the habitat categories of riparian vegetation, moors and heathland, and alpine grasslands.
-Overall, the proportion of Lepidoptera and Coleoptera species was low across all habitats but Coleoptera were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
+Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation and moors and heathland.
+Overall, the proportion of flower visitors from Lepidoptera and Coleoptera were low across all habitats but Coleopteran flower visitors were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
 Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-In this particular case, Hymenoptera was the predominant order in all bioclimatic regions with Diptera taking more importance in Alpine and Atlantic regions.
-Lepidoptera shows low proportions across all bioclimatic regions and Coleoptera is only relevant in Mediterranean regions at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S2**) also differs across habitats and bioclimatic regions.
-The habitats sampled by a higher number of studies in the database are `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
-However, the habitats that contain a higher number of sampling sites are `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies are `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and the ones that contain a higher number of sampling sites are `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
+Hymenopterans were abundant across all bioclimatic regions and Dipterans were particularly abundant in the Boreal, Alpine and Atlantic regions.
+Lepidopterans had low proportions across all bioclimatic regions and Coleopterans were only relevant in the Mediterranean region at European level.
+Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S3**) also differed across habitats and bioclimatic regions.
+The habitats sampled by a higher number of studies in the database were `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
+However, the habitats that contain a higher number of sampling sites were `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
+The bioclimatic regions with a higher number of studies were `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and those that contain a higher number of sampling sites were `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
 
@@
 Note that a single study can contribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}
+Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.
+The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day.", fig.width=4.5, fig.height=8, fig.align='center'}
+
+#Filter out 2 bioregions that only have a small study with data of a single day
+data = data %>% 
+filter(!Bioregion == "Steppic") %>% 
+filter(!Bioregion == "Pannonian")
 
@@
 
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**).
-We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**).
-Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**).
-Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located in lower latitudes.
+Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
+Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
+Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**).
+The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested.
+The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested.
+Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
+(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
+Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes.
 Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
@@
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-This likely reflects the fact that most plant and pollinator species are rare and geographically restricted.
-For example given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
+However, although EuPPollNet likely has a higher pollinator coverage, as it did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
+For example, given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
 Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
@@
 
-Bees are responsible for the majority of the interactions at the metaweb level, but their relevance for network topology changed across habitats and bioclimatic regions.
+Bees are responsible for the majority of the sampled interactions at the metaweb level.
+As not all surveys included all pollinator groups, this result may partly be influenced by the taxonomic groups sampled across studies.
+However, the relevance of bees and other pollinator orders for network topology changed across habitats and bioclimatic regions in accordance to the literature.
 For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
 These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only common floral visitors in the Mediterranean region.
-Although the networks from the database are visitation networks and do not capture pollinator efficiency [@ballantyne2015], the high proportion of beetles as floral visitors in the Mediterranean provides further support for their potential role as pollinators in this region [@herrera2019; @leon2022].
+In addition, beetles were only commonly documented as floral visitors in the Mediterranean region.
+This study cannot determine whether pollination ecologists traditionally document flower-beetle interactions only in the Mediterranean, or if there are fewer flower visitations by beetles outside this region.
+Nevertheless, the high proportion of beetles as floral visitors provides further support for their potential role as pollinators in the Mediterranean [@herrera2019; @leon2022].
 The number of butterfly species and interactions were relatively low compared to the other taxa.
@@
 Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities across the world [~13%\; @hung2018], potentially reflecting the dominance of pasture habitats in EuPPollNet, the highly generalised nature of honey bees, the important role that honey bees are playing as pollinators in Europe, and/or because beekeeping is widely practice in Europe [@steffan2000; @magrach2017; @herrera2020].
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio), the observed number of pollinator species in the database was almost double that of the plants.
+The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018].
+This potentially reflects the dominance of intensive grassland habitats in EuPPollNet and their widespread distribution across European landscapes [@isselstein2005], the highly generalised nature of honey bees, their native status and the important role that honeybees are playing as pollinators in Europe, and/or the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
+
+Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants.
 This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
@@
 Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes.
-Finally, one third of networks were more nested than expected by chance.
-While this supports the idea that plant-pollinator networks tend to be nested [@bascompte2003], this result also suggests that nestedness could be a less prevalent feature than previously thought for plant-pollinator networks [@payrato2019].
-
-Despite this database covering a wide range of habitats across `r country_number` countries, it contains geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
+These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes in Europe.
+Finally, most empirical networks (~85%) showed a non-nested structure.
+This result opposes to the idea that plant-pollinator networks tend to be nested [@bascompte2003], and support recent findings showing that nestedness is not a prevalent feature in plant-pollinator networks [@payrato2019].
+
+Despite this database covers a wide range of habitats across `r country_number` countries, it contains temporal and geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
+For instance, none of the studies in this database sampled nocturnal pollinators, which can impact our view of network structure [@garcia2024], and most studies were conducted during a single flowering season, limiting our ability to evaluate temporal trends of plant-pollinator communities in the face of environmental changes [@alarcon2008; @chacoff2018].
+In addition, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
 This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-This lack of data is especially relevant for Eastern Europe which has vast landscapes of high quality semi-natural grasslands but is experiencing rapid land use change [@sutcliffe2015], and the Mediterranean region is likely to be severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023].
+The lack of data for Eastern Europe, which contains vast landscapes with semi-natural grasslands experiencing rapid land use change [@sutcliffe2015], and for the Mediterranean region, which is severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023], is particularly concerning.
 These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
@@
 In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
-Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not an ubiquitous feature of all plant-pollinator networks.
+Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
 These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
@@
 
-We thank all the taxonomists and ecologists who have made this database possible by contributing their fieldwork data.
+We thank all the taxonomist and ecologist that has made this database possible by contributing with their fieldwork data.
 
@@
 Note, that we have included in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.
+This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
+
+9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.  This category partially matches the **EUNIS habitat type**  ‘coastal habitats’ (code **N**).
 
 10) **Riparian vegetation**: Plant communities growing on river margins.
-
-10) **Alpine grasslands**: Low growing plant communities with little or none human disturbance. Often located in high elevation areas within Europe.
-
-12) **Moors and heathland**: Low growing woody vegetation characteristic from low fertile soils near the coast or in alpine areas.
+This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**) but also the category of ‘inland waters’ which is currently under review.
+
+10) **Montane to alpine grasslands**: Communities dominated by non-woody, low-growing plants that experience little or no human disturbance and are found in high-elevation areas (>1,000 m).
+This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R**; e.g., alpine and subalpine grasslands with code  **R4**).
+
+12) **Moors and heathland**: Low-growing woody vegetation (i.e., typically < 5 m) that is characteristic of poorly fertile soils where environmental conditions prevent the natural formation of forests.
+This category is similar to the **EUNIS habitat type** ‘heathland, scrub and tundra’ (code **S**).
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S1}.
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure S1}. Geographical location of all networks in the EuPPollNet database coloured by bioregion."}
+
+#Read data
+data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
+mutate(Network_id = paste0(Study_id, Network_id))
+
+#Prepare coordinates
+for_plotting = data %>% 
+select(Study_id, Latitude, Longitude, Bioregion) %>% 
+distinct()
+#Get countries and transfrom coordinates
+countries <- gisco_get_countries(
+resolution = 20) %>%
+st_transform(3035) %>% 
+sf::st_as_sf(crs = 4326)
+
+#Try to plot both together (same plot)
+map = ggplot(countries) +
+geom_sf(fill="floralwhite",color = "black", 
+        alpha = 1,linewidth=0.1) + 
+geom_point(data = for_plotting, alpha=0.75,size=1.2,pch=21,stroke = 0.25,aes(x= Longitude, y=Latitude, fill= Bioregion))+  
+coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
+         xlim = c(-14, 39), ylim = c(37, 70))+
+theme(panel.grid.major = element_blank(), 
+     legend.key = element_blank(), 
+     legend.text = element_text(size = 6), 
+     legend.title = element_text(face ="bold",  size = 8),
+     legend.spacing.x=unit(0.5, "lines"), 
+     legend.key.width=unit(0.5, "lines"), 
+     legend.box.margin=margin(-10,-10,-10,-10),
+     axis.text = element_text(size=6,face="bold"),
+     axis.title = element_text(face="bold", size = 8),
+) +
+theme(panel.background = element_rect(fill = "aliceblue"),
+        panel.border = element_rect(colour = "black",
+        fill=NA, size=1)) + 
+ scale_size_continuous(name= "Species counts",range = c(0, 3.75))  +
+ylab("Latitude") +
+xlab("Longitude") +
+scale_x_continuous(breaks = seq(0, 20, by = 10)) +
+scale_y_continuous(breaks = seq(40, 60, by = 10)) +
+guides(fill = guide_legend(override.aes = list(size = 1.5, byrow = TRUE))) +
+scale_fill_manual(values = c("purple3", 
+                               "dodgerblue3", 
+                               "cyan4", 
+                               "forestgreen",
+                               "orange2",
+                               "tomato3",
+                               "khaki2"))
+
+#Plot everything
+map 
+```
+
+\newpage
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S2}.
 Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S3}.
-Association between nestedness (NODFc) and the geometric mean of plan and pollinator species per network", fig.width=3, fig.height=3}
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S4}.
+Association between nestedness (NODFc) and the geometric mean of plant and pollinator species per network", fig.width=3, fig.height=3}
 

### 2024-10-01 Manuscript/EuPPollNet_Ms.qmd
authors

@@
 bibliography: references.bib
-csl: apa_6th.csl
+csl: apa_7th.csl
 latex-tinytex: false
@@
 The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
-As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022.
+As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; <https://doi.org/10.3030/101003476>), first, data was directly requested from members of the Safeguard project in May 2022.
 Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024.
@@
 Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
-Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database.
+Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they account for a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
+Hymenoptera and Diptera contain the highest number of species, with each comprising approximately 1,000 species in the database.
 However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
@@
 
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
+For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and manually fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
 
@@
 
+#Load number of species from checklists and extrapolated one
+bee_potential_spp = readRDS("../Data/Manuscript_info/bee_potential_spp.RData")
+syrphid_potential_spp = readRDS("../Data/Manuscript_info/syrphid_potential_spp.RData")
+lepidoptera_potential_spp = readRDS("../Data/Manuscript_info/lepidoptera_potential_spp.RData")
+extrapolated_other_spp = readRDS("../Data/Manuscript_info/other_taxa_potential_spp.RData")
+extrapolated_other_spp = round(extrapolated_other_spp)
 ```
@@
 
-Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that benefit from animal pollination.
+Europe hosts approximately over 5,000 species of pollinators, including `r prettyNum(bee_potential_spp, big.mark=",")` bee species `r prettyNum(syrphid_potential_spp, big.mark=",")` syrphid species, `r prettyNum(lepidoptera_potential_spp, big.mark=",")` butterflies species, and about 1,700 species from other taxonomic groups, as estimated based on EuPPollNet average taxonomic coverage.
+Additionally, there are around 25,000 species of plants that benefit from animal pollination when excluding non-biotic pollinated species.
 EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic and taxonomic coverage of the bee genera at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Dark grey bars indicate the total number of species per genus at the European level."}
+Phylogenetic and taxonomic coverage of bee genera at European level.
+The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple.
+Additionally, the coverage of species recorded in EuPPollNet for each genus is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database, out of the total number of bee species in Europe.
+Dark grey bars represent the total number of species per genus at European level."}
 
@@
 annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
-annotate("text", x = 2.55, y = 28.5, label = "Apidae", size= 3)+
-annotate("text", x = 2.25, y = 51.2, label = "Megachilidae", size= 3) +
+annotate("text", x = 2.25, y = 26.8, label = "Megachilidae", size= 3)+
+annotate("text", x = 2.5, y = 51, label = "Apidae", size= 3) +
 coord_cartesian(clip = "off")+
@@
 Phylogenetic and taxonomic coverage of the plant families at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Dark grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
+The number of interactions recorded per family in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple.
+Additionally, the coverage of species recorded in EuPPollNet for each family is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database out of the total number of bee species per family in Europe.
+Dark grey bars represent the total number of species per family at European level.", fig.width=8, fig.height=8}
 
@@
 Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.
-The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day.", fig.width=4.5, fig.height=8, fig.align='center'}
+The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day and site.", fig.width=4.5, fig.height=8, fig.align='center'}
 
@@
 connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
+
+connectance_lat_pval = ifelse(connectance_lat_pval <0.01, 0.01, connectance_lat_pval)
+
 #Load lm output nestedness ~ latitude
@@
 normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
+
 #Load values of connectance
@@
 
+#To avoid having values under 0.01
+resid_connectance_nestedness_corr_pval = ifelse(resid_connectance_nestedness_corr_pval < 0.01, 0.01, resid_connectance_nestedness_corr_pval)
+
 ```
@@
 Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
+Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were not independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
 Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**).
@@
 Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
-(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
+(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ < `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ ~ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
 Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes.
-Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
+Note that residual connectance and normalised nestedness showed a moderate significant negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ < `r round(resid_connectance_nestedness_corr_pval, 2)`).
 
@@
 The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018].
-This potentially reflects the dominance of intensive grassland habitats in EuPPollNet and their widespread distribution across European landscapes [@isselstein2005], the highly generalised nature of honey bees, their native status and the important role that honeybees are playing as pollinators in Europe, and/or the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
+This potentially reflects the dominance of intensive grassland habitats in EuPPollNet and their widespread distribution across European landscapes [@isselstein2005], the highly generalised nature of honey bees, their native status and above all, the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
 
@@
 
-Despite this database covers a wide range of habitats across `r country_number` countries, it contains temporal and geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
+Although this database covers a wide range of habitats across `r country_number` countries, it contains temporal and geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
 For instance, none of the studies in this database sampled nocturnal pollinators, which can impact our view of network structure [@garcia2024], and most studies were conducted during a single flowering season, limiting our ability to evaluate temporal trends of plant-pollinator communities in the face of environmental changes [@alarcon2008; @chacoff2018].

### 2024-10-01 Scripts/1_Processing_networks/49_Hervias-Parejo.R

@@ -127,7 +127,7 @@ rename(Metadata_fields = rowname, Metadata_info= V1) %>% as_tibble()
 Authorship <- data.frame(
   Coauthor_name = c("Sandra Hervias-Parejo", "Anna Traveset"),
   Orcid = c("0000-0002-5377-3619", "0000-0002-1816-1334"),
-  E_mail = c("shervias@imedea.uib-csic.es", "atraveset@imedea.uib-csic.es"))
+  E_mail = c("shervias@imedea.uib-csic.es", "atraveset@csic.es"))
 
 #Save data ----
 #Create metadata list

### 2024-10-02 Manuscript/EuPPollNet_Ms.qmd
final edits

@@
 Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
-The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions, avoiding the influence of network size.
+The z-score approach only compares each unique network against their randomized versions, avoiding the influence of network size.
 As NODF is a metric computed from binary matrices, we calculated 100 null models for the binarized version of each network using the 'curveball' algorithm [@strona2014].
@@
 random rewiring of all links, without self-links or double links) and thus keeps the exact number of connections per species (i.e., realised degree).
-Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species.
+Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented a null model that reorganizes the quantitative networks before binarising them.
+To that end, we used the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species.
 Each empirical network was randomized 100 times using each method.
@@
 Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were not independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
+Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and, as expected, were not independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
+Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
+(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ < `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ ~ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
+Overall, networks towards higher latitudes showed lower residual connectance but similar nestedness compared to networks located at lower latitudes.
+Note that residual connectance and normalised nestedness showed a moderate significant negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ < `r round(resid_connectance_nestedness_corr_pval, 2)`).
 Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**).
@@
 The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested.
-Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
-(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ < `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ ~ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
-Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes.
-Note that residual connectance and normalised nestedness showed a moderate significant negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ < `r round(resid_connectance_nestedness_corr_pval, 2)`).
+Note that in both cases, NODFc is calculated on binarised matrices.
 
@@
 Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
+The solid fitted line indicates a significant association, while the dashed lines indicate a non-significant association.
 The bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
@@
   )) +
-geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
+geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE, linetype = "dashed")+
 geom_ribbon(data = nestedness_latitude, 
@@
 While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-However, although EuPPollNet likely has a higher pollinator coverage, as it did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
+This likely reflects that most plant and pollinator species are rare and geographically restricted.
 For example, given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
@@
 
-Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
+Consistent with @olesen2002, we found that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
 Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes in Europe.
-Finally, most empirical networks (~85%) showed a non-nested structure.
-This result contrasts with earlier work suggesting that plant-pollinator networks are typically nested [@bascompte2003] and it aligns with recent evidence showing that nestedness is not a common feature in plant-pollinator networks [@payrato2019].
+These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected at lower latitudes in Europe.
+However, our results cannot be extrapolated to lower latitudes outside Europe, as tropical systems might behave differently.
+Finally, a result that can be surprising is that most empirical networks (~85%) showed a non-nested structure.
+This result contrasts with earlier work suggesting that plant-pollinator networks are typically nested [@bascompte2003] and it aligns with recent evidence showing that nestedness is not a common feature in plant-pollinator networks when evaluated against restrictive null models that conserve the observed species degree [@payrato2019\; Figure S6].
+Note that while species degree distributions are sufficient to explain the emergence of nestedness, this does not preclude nestedness from being a useful metric for comparison across networks.
+For example, plant-pollinator networks are more nested than plant-herbivore networks precisely because these two network types differ in their composition of generalist and specialist species [@thebault2010].
 
@@
 In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
-Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
+Here, we have shown how connectance and nestedness change across their latitudinal range, and that plant-pollinator networks are as nested as expected given plant and pollinator generalist levels.
 These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
@@
 ```
+
+\newpage
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S6}.
+Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with a probabilistic null model from Bascompte et al., 2003 (100 null models for each network).
+This is the less restrictive approach, as it does not constrain the realised species degree.
+It first calculates the probability for each cell, and then if a random number between 0 and 1 exceeds this probability, an observed interaction is assigned to the cell.
+This shows that when the realised degree is not constrained, the number of networks that are statistically more nested is notably higher, highlighting that the nested structure is likely a byproduct of the species degree distribution."}
+
+library(dplyr)
+library(ggplot2)
+
+#Read data
+null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_nullmodel2.rds")
+metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
+
+#Finally, find % of networks that differ statistically
+# Merge and process data
+d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
+d = d %>% 
+mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
+
+p = 0.05 #cutoff probability 95% confidence
+critical_value = qnorm(p/2) #double tail probability divide by 2
+
+d1 = d %>%
+mutate(infra_over_represented = case_when(
+z_score < -abs(critical_value) ~ "Under-represented",
+between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
+    z_score > abs(critical_value) ~ "Over-represented"
+  ))
+
+#Calculate percentages
+z_percent = d1 %>% 
+group_by(infra_over_represented) %>% 
+summarise(percent = length(infra_over_represented)/nrow(.) * 100)
+
+# Plot
+ d1 %>% 
+  ggplot(aes(x = z_score)) +
+  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
+                 bins = 50, alpha = 0.5, color = "black", position = "stack") +
+  theme_bw() +
+  coord_cartesian(expand = FALSE) +
+  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
+  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
+  ylab("Number of networks") +
+  xlab("Z scores (NODFc)")  +
+ # xlim(-6, 9) +
+ # ylim(0, 0.8) +
+  scale_fill_manual(name = "Observed against null",
+                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
+                    labels = c("Less nested", "No difference", "More nested"),
+                    values = c("Under-represented" = "coral2", 
+                               "No statistical difference" = "palegreen3", 
+                               "Over-represented" = "cyan3")) +
+xlim(-6,6) +
+theme(plot.title = element_text(size = 24),
+      text = element_text(size = 20),
+      axis.title  = element_text(face="bold"))
+
+```

### 2024-10-02 Manuscript/Figures_and_Tables/Figure_S6_NULL.Rmd
final edits

--- 2024-09-21 Manuscript/EuPPollNet_Ms.qmd
@@
 ---
-title: "EuPPollNet: A European database of plant-pollinator networks"
-format:
-  pdf:
-    include-in-header: 
-      text: |
-        \usepackage{lscape}
-        \newcommand{\blandscape}{\begin{landscape}}
-        \newcommand{\elandscape}{\end{landscape}}
-        \usepackage{caption}
-        \captionsetup[figure]{labelformat=empty}
-        \captionsetup[table]{labelformat=empty}
-        \usepackage{lineno}
-        \linenumbers
-bibliography: references.bib
-csl: apa_6th.csl
-latex-tinytex: false
-link-citations: TRUE
-linkcolor: RoyalBlue
-urlcolor: RoyalBlue
+output:
+  pdf_document:
+header-includes:
+- \usepackage{caption}
+- \captionsetup[figure]{labelformat=empty}
+- \captionsetup[table]{labelformat=empty}
 ---
 
-```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr) #For handling fata 
-library(sf) #For handling coordinates and plotting
-library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
-library(readr)
-library(phytools)
-#Load libraries for plotting
-library(giscoR) #Maps
-library(ggstar) #cool shapes as points
-library(ggplot2) #general plotting
-library(ggtree) #phylo trees
-library(ggpattern) #cool shapes in barplots
-library(patchwork) #plot organisation
-library(cowplot) #plot organisation
-library(viridis) #colours
-library(ggnewscale)
-library(ggtreeExtra)
-library(kableExtra)
+```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S6}."}
 
-#In case of conflicts select the main dplyr functions
-library(conflicted)
-conflicts_prefer(dplyr::filter)
-conflicts_prefer(dplyr::select)
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
+library(dplyr)
+library(ggplot2)
 
 #Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
-
-#Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
-#Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
-                    header = T)
-#Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")
-
-#Work around to get the three cols
-#Note that month 1 and day 1 indicates the presence of a missing value
-#At least in Europe :)
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-
-#Check number of distinct networks
-n_networks = data %>% 
-mutate(Network_id = paste0(Network_id,"_", Year)) %>% 
-distinct(Network_id) %>% 
-nrow(.)
-
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Check number of distinct habitats
-n_habitats = data %>% 
-select(EuPPollNet_habitat) %>% 
-distinct() %>% 
-nrow(.)
-#Read author list 
-authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
-#Print authors
-authors = authors %>% 
-summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
-pull()
-#Read affiliation list 
-aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
-aff_list = aff_list %>% 
-summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
-pull()
-
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Calculate mean coverage and deviation
-coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
-mean_coverage = mean(coverage)
-sd_coverage = sd(coverage)
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-#Prepare table with all info 
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude" ~ "North-south position of the observed interaction in decimal degrees",
-        Variable == "Longitude" ~ "East-west position of the observed interaction in decimal degrees",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_original_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_original_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
-        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
-
-    TRUE ~ "Other"
-  ))
-
-#Check number of species and percentage of interactions of poll. groups
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-order_info = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
-          Species = n_distinct(Pollinator_accepted_name))
-
-Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
-select(Interactions) %>% pull()
-
-#Check percentage of studies with flower counts
-floral_counts_percent = data  %>%
-select(Study_id, Flower_data) %>% 
-distinct() %>% 
-group_by(Flower_data) %>% 
-summarise(Percentage = n()/ nrow(.)*100) %>% 
-filter(Flower_data == "Yes")
-
-#Check percentage of apis mellifera records
-all_records = data %>% nrow(.) 
-apis_records = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-apis_percent = apis_records/all_records * 100
-#Check average pecentage of interactions of apis by net
-a = data %>% 
-group_by(Network_id) %>% 
-summarise(n_interactions = length(Pollinator_accepted_name))
-b = data %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-summarise(n_interactions_apis = length(Pollinator_accepted_name))
-c = left_join(a,b)
-apis_average_network = c %>% 
-mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
-ungroup() %>% 
-summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
-pull()
-
-#Check number of species from Hymenop, Dipte, Lepidop and Coleop
-
-pollinator_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-pollinator_spp_main_orders = data %>% 
-select(Pollinator_order, Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera"|
-       Pollinator_order == "Diptera"|
-       Pollinator_order == "Lepidoptera"|
-       Pollinator_order == "HymenoColeopteraptera") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-spp_percent_main_order = pollinator_spp_main_orders/pollinator_spp *100
-
-#Calculate percentage of studies that recorded the main 4 orders
-metadata = readRDS("../Data/3_Final_data/Metadata.rds")
-#Some studies are divided by year
-#Get an average of the pollinator percentages across years first
-#Fix NA's
-metadata_study = metadata %>%
-group_by(Study_id) %>% 
-mutate(Hymenoptera_percent = as.numeric(na_if(Hymenoptera_percent, "NA"))) %>%  
-mutate(Diptera_percent = as.numeric(na_if(Diptera_percent, "NA"))) %>%  
-mutate(Lepidoptera_percent = as.numeric(na_if(Lepidoptera_percent, "NA"))) %>%  
-mutate(Coleoptera_percent = as.numeric(na_if(Coleoptera_percent, "NA"))) 
-#Calculate average
-mean_study = metadata_study  %>% 
-summarise(Hymenoptera_percent = mean(Hymenoptera_percent, na.rm = T),
-          Diptera_percent = mean(Diptera_percent, na.rm = T),
-          Lepidoptera_percent = mean(Lepidoptera_percent, na.rm = T),
-          Coleoptera_percent = mean(Coleoptera_percent, na.rm = T)) %>% 
- mutate(across(c(Hymenoptera_percent, Diptera_percent, Lepidoptera_percent, Coleoptera_percent),
-                ~ ifelse(is.nan(.), NA, .))) 
-#Now calculate percentage of studies that record any of these 4 orders
-mean_percent_order = mean_study %>%
-  summarise(
-    Hymenoptera_percent_metaweb = sum(!is.na(Hymenoptera_percent)) / nrow(.) * 100,
-    Diptera_percent_metaweb = sum(!is.na(Diptera_percent)) / nrow(.) * 100,
-    Lepidoptera_percent_metaweb = sum(!is.na(Lepidoptera_percent)) / nrow(.) * 100,
-    Coleoptera_percent_metaweb = sum(!is.na(Coleoptera_percent)) / nrow(.) * 100)
-
-Hymenoptera_percent_metaweb = mean_percent_order$Hymenoptera_percent_metaweb
-Diptera_percent_metaweb = mean_percent_order$Diptera_percent_metaweb
-Lepidoptera_percent_metaweb = mean_percent_order$Lepidoptera_percent_metaweb
-Coleoptera_percent_metaweb = mean_percent_order$Coleoptera_percent_metaweb
-
-#Check percentages of hymenoptea groups
-hymenoptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
-          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
-          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))
-
-bee_percent = hymenoptera_percent_by_group$Bee_studies
-bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
-all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies
-
-#Check percentages of diptera groups
-diptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
-          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
-          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
-                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)
-
-syrphid_percent = diptera_percent_by_group$Syrphid_studies
-all_diptera_percent = diptera_percent_by_group$Diptera_studies
-other_diptera_percent = diptera_percent_by_group$other_diptera_studies
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first four and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks.
-One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
-We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
-We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified.
-Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from `r n_habitats` different natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphid species at the European level.
-
-**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
-The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
-
-**KEYWORDS**\
-\begingroup
-\small
-Angiosperms, connectance, flowering plants, nestedness, plant-pollinator networks, pollinators, species interactions 
-\endgroup
-\pagebreak
-
-# 1 | INTRODUCTION
-
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species status, trends, and conservation [@heberling2021; @zattara2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano2003] or nestedness [@bascompte2003].
-Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
-Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @garcia2024].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
-Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness.
-This emphasizes the importance of providing data in its rawest possible form in datasets that will be utilized for synthesis and macro-ecological studies.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023], along with occurrence data [@zattara2021], is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, generating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level.
-For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
-Assembling and curating the existing information on EU plant-pollinator networks will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on plant-pollinator interactions at the European level.
-The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
-
-EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks.
-For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where specialists species interact only with subsets of generalists species, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
-
-Overall, EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
-This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
-We expect that EuPPollNet can be used to evaluate macro-ecological processes in plant-pollinator networks, guide conservation planning, and set a baseline for global change research.
-
-# 2 | METHODS
-
-## Data acquisition
-
-The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
-As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022.
-Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024.
-These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
-While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019].
-This approach maximized the potential number of studies that could be incorporated in this database.
-The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'.
-To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
-
-## Dataset description
-
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations).
-The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.
-Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
-Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database.
-However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
-Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have selected only a single location per study.
-(\\textbf{b}) Number of studies by year in EuPPollNet.
-(\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}.
-The total number of species and interactions is indicated in parentheses at the bottom."}
-
-#Fix dates
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p2  = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-   axis.ticks.x = element_line(color="black"),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -5))
-
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-#Calculate spp number and interactions separately
-data1_spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
-
-data1_int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))
-
-data1 = left_join(data1_spp, data1_int)
-
-#Select main orders
-spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
-pull()
-#Select main orders
-int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Interactions = length(Pollinator_accepted_name)) %>% 
-pull()
-
-#Combine Interactions and Species into a single column
-data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
-
-#Set order of levels for plotting
-data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))
-
-#Plot
-p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
-geom_bar(stat = "identity", width = 0.7) +
-theme_classic() +
-scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator\n order") +
-xlab(NULL) +
-ylab("Percentage") + 
-scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
-theme(axis.text = element_text(size=6), 
-     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
-     axis.title.x = element_text(face="bold", size = 7),
-     legend.position = "top",
-     legend.title = element_text(size = 5, face = "bold"),
-     legend.text = element_text(size = 5),
-     legend.key.size = unit(3, "mm"),
-     legend.margin=margin(0,2,0,0),
-     legend.box.margin=margin(-9,-9,-6,-6)) +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))+
-ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -6)) +
-scale_x_discrete(labels = c("Species\n (2,172)", "Interactions\n (358,049)"))
-
-#plot2 = p3 / p1 / p2 
-plot2 = p2 / p3
-
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.485, 0.8))
-
-```
-
-\newpage
-
-## Data structure
-
-The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble)` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
-These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29).
-The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30).
-Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies.
-To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year.
-Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.
-
-\newpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-#Small edit to column number
-my_tibble = my_tibble %>% 
-mutate(Variable = paste0(seq(1:length(Variable)),". ", Variable))
-
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.} Column names and their descriptions within the EuPPollNet database.") %>%
-  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
-kable_styling(position = "center")
-
-```
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and harmonized in R using **rgbif** [@chamberlain2022].
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-In addition, taxonomic information at genus, family and order level was downloaded for each species.
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023], and butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-Coleoptera species names were only checked against the GBIF checklist.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the EuPPollNet database at the European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only plants occurring in Europe and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we only evaluated the taxonomic coverage of groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
-While there is not a good understanding of pollinator diversity in other taxonomic groups (e.g., beetles), it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
-Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
-Consequently, we provide an estimate for the total number of pollinators across the European continent.
-
-Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
-The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it using the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**.
-For plants, the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of the **rtree** package [@li2023].
-
-## Sampling coverage
-
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks.
-In addition, we computed the  accumulation curve of pollinator species with an increasing number of plant species as an indicator for how many pollinator species are likely responsible for the pollination of flowering plants [e.g., @kleijn2015 for crops].
-The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
-The different rarefied curves were complemented with 100 bootstrapped accumulation curves.
-
-## Habitat type and bioclimatic region
-
-We describe the habitat type for each site using information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022], visual inspection of Google Earth imagery and the habitat classification from the authors.
-These different habitat categories (see definitions in supplementary text) allow a quick comparison and understanding of the habitat types from the database.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned a biogeographical region to each site.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites using a spatial joint from the **sf** package [@pebesma2018].
-
-## Network analyses
-
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network and examined how these network metrics change across different latitudes and bioclimatic regions in Europe.
-We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
-We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics.
-As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-The relationship between residual connectance and species richness was investigated using a beta regression.
-This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1.
-The model was implemented using the betareg package [@cribari2010].
-We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species in comparison to other nestedness metrics that change with network size [@song2017].
-This metric was calculated using the **maxnodf** package [@hoeppke2021].
-Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, to quantify how connectance and nestedness change with network size, we determined their association with the number of species per network using the Kendall rank correlation coefficient.
-
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
-The z-score approach allows us to compare our results with previous published nestedness analyses in plant-pollinator networks and only compares each unique network against their randomized versions, avoiding the influence of network size.
-As NODF is a metric computed from binary matrices, we calculated 100 null models for the binarized version of each network using the 'curveball' algorithm [@strona2014].
-This algorithm implements the configuration model (i.e.
-random rewiring of all links, without self-links or double links) and thus keeps the exact number of connections per species (i.e., realised degree).
-Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species.
-Each empirical network was randomized 100 times using each method.
-These null models were implemented with the **vegan** [@oksanen2013] package.
-Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite** [@dormann2008].
-
-\newpage
-
-# 3 | RESULTS
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Taxonomic coverage outputs
-#Find number of plant and pollinator species
-##Pollinators
-poll_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-##Plants
-plant_spp = data %>% 
-select(Plant_rank, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-nrow(.)
-
-#Overall plant coverage
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Average plant family coverage
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData") %>% 
-filter(fam_group == "B")
-#Average coverage per family
-mean_plant_fam_cov = d1 %>% 
-ungroup() %>% 
-summarise(mean_plant_fam_cov = mean(Percent)) %>% 
-pull()
-
-#Percent of families with interaction records
-all_fam = nrow(d1)
-fam_with_data = d1 %>% 
-filter(Percent > 0) %>% 
-nrow()
-percent_plant_fam_with_data = fam_with_data/all_fam * 100
-
-#Calculate phylogenetic signal of presence-absence information
-#Read data and plot it
-tree = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-#Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-#select(label, categorical) %>% 
-ungroup() 
-#Check phylogenetic signal
-d1 = d1 %>% arrange(factor(label, levels = tree@phylo$tip.label))
-
-v = phylo_trait$categorical
-names(v) = tree@phylo$tip.label
-
-phylo_signal_plant = phylosig(tree@phylo, v, method="lambda", test=TRUE)
-lambda_phylosig_plants = phylo_signal_plant$lambda
-p_phylosig_plants = phylo_signal_plant$P
-
-#Load pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-#Check percentage of interactions from bees with and without apis
-#Check percentage of bees in the database
-bee_fam = c("Apidae", "Megachilidae", "Halictidae",
-            "Andrenidae", "Colletidae", "Melittidae")
-all_rows = data %>% nrow(.)
-#
-bee_rows = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-nrow(.)
-#Bee interactions
-bee_interactions_percent = bee_rows/all_rows *100
-#Eclude now apis
-bee_rows_non_apis = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-#
-bee_interactions_non_apis_percent = bee_rows_non_apis/bee_rows *100
-
-#Coverage across bee genera
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-all_rows_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-nrow(.)
-#
-data_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-filter(Percent > 0) %>% 
-nrow(.)
-#
-bee_genera_percent_coverage = data_bees/all_rows_bees * 100
-#Check average coverage across bee genera
-average_coverage_bee_genera = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-ungroup() %>% 
-summarise(average_coverage_bee_genera = mean(Percent)) %>% 
-pull()
-
-#Calculate phylogenetic signal of presence-absence information
-##Read data and plot it
-tree = readRDS("../Data/Manuscript_info/bee.tree100.rds")
-##Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-ungroup() 
-##Conduct left join to tibble of tip labels to organise order
-v = tibble(Genus1 = tree$tip.label) %>% 
-mutate(Genus1 = str_to_title(word(Genus1,1,sep = "_")))
-#
-v1 = left_join(v, phylo_trait)
-#
-v2 = v1$categorical
-names(v2) = tree$tip.label
-#
-phylo_signal_bees = phylosig(tree, v2, method="lambda", test=TRUE)
-lambda_phylosig_bees = phylo_signal_bees$lambda
-p_phylosig_bees = phylo_signal_bees$P
-
-#Check number of studies by habitat
-studies_by_habitat = data %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of sampling sites by habitat
-networks_by_habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of studies by bioregion
-studies_by_bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check number of sampling sites by bioregion
-networks_by_bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check percentage of Apis mellifera across networks
-apis_percent_networks = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>%
-group_by(Study_id) %>% 
-mutate(Apis = "Yes") %>% 
-distinct(Study_id, Apis) %>% 
-nrow() / nlevels(factor(data$Study_id)) *100
-
-```
-
-## Taxonomic coverage
-
-Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that benefit from animal pollination.
-EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
-The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
-The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic and taxonomic coverage of the bee genera at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level.
-Dark grey bars indicate the total number of species per genus at the European level."}
-
-#Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
-
-#Circular layour
-ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.065, size=2.9, fontface=1) +
-scale_size(range = c(0,4.5), guide = 'none')  +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-guides(fill=guide_legend(), size = guide_legend()) +
-theme(legend.position = "bottom") +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
-                    pwidth=0.13, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0.148,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-guides(fill = "none")+
-scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
-geom_fruit(geom=geom_bar,
-                    mapping=aes(y=label, x=Spp_number_Europe),
-                    pwidth=0.12, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = -0.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
-new_scale_fill() +
-geom_point2(aes(subset=node==131), color='orange', size=3) +
-geom_point2(aes(subset=node==84), color='orange', size=3) +
-geom_point2(aes(subset=node==73), color='orange', size=3) +
-geom_point2(aes(subset=node==74), color='orange', size=3) +
-geom_point2(aes(subset=node==111), color='orange', size=3)+
-geom_point2(aes(subset=node==91), color='orange', size=3) +
-annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
-annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
-annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
-annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
-annotate("text", x = 2.55, y = 28.5, label = "Apidae", size= 3)+
-annotate("text", x = 2.25, y = 51.2, label = "Megachilidae", size= 3) +
-coord_cartesian(clip = "off")+
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.6,0),
-      legend.direction = "horizontal")
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-Phylogenetic and taxonomic coverage of the plant families at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale.
-Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization.
-The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level.
-Dark grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Sampling coverage outputs
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Coverage
-poll_sampling_coverage = poll_output$DataInfo$SC *100
-plant_sampling_coverage = plant_output$DataInfo$SC *100
-int_coverage = int_output$DataInfo$SC *100
-plant_poll_coverage = plant_poll_output$DataInfo$SC *100
-
-#Increase doubling effort
-##Pollinators
-poll_increase = poll_output$iNextEst$coverage_based
-poll_increase1 = poll_increase %>% filter(Method == "Observed") %>% pull(qD)
-poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
-poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
-##Plants
-plant_increase = plant_output$iNextEst$coverage_based
-plant_increase1 = plant_increase %>% filter(Method == "Observed") %>% pull(qD)
-plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
-##Interactions
-int_increase = int_output$iNextEst$coverage_based
-int_increase1 = int_increase %>% filter(Method == "Observed") %>% pull(qD)
-int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
-int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
-##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$coverage_based
-plant_poll_increase1 = plant_poll_increase %>% filter(Method == "Observed") %>% pull(qD)
-plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
-
-#Prepare results from common and rare poll across sampling sites
-#Pollinators
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-poll_total = nrow(poll_spread)
-polls_under_one_percent = poll_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-polls_under_one_percent = polls_under_one_percent/poll_total * 100
-#Find most common pollinator
-most_common_poll = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Pollinator_accepted_name)
-#Find most common poll percent
-most_common_poll_percent = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-#Plants
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-plant_total = nrow(plant_spread)
-plants_under_one_percent = plant_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-plants_under_one_percent = plants_under_one_percent/plant_total * 100  
-#Find most common plant
-most_common_plant = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Plant_accepted_name)
-#Find most common plant percent
-most_common_plant_percent = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-```
-
-## Sampling coverage
-
-The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_sampling_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**).
-The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
-When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
-We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**).
-The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and the number of unique pairwise interactions across networks.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
-The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks.
-Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network.
-Note that \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
-
-#Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
-#Create dataset
-d_poll = poll_output$iNextEst$coverage_based
-#Check cols
-d_poll$col = d_poll$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           Method = as.character(pollinators_curves$iteration), 
-           qD = pollinators_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(Method == "Observed")
-d_poll1_lines = d_poll1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_poll1_lines = d_poll1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col),size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_poll1_point, col = "black", size=2) + 
-ylab("Pollinator species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000) +
-ggtitle("(a)") +
-theme(plot.margin = margin(r = 12))
-
-#Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
-#Create dataset
-d_plant = plant_output$iNextEst$coverage_based
-#Check cols
-d_plant$col = d_plant$Method
-#Now create tibble with same structure
-a_plant = tibble(t = plant_curves$Study_sampled, 
-           Method = as.character(plant_curves$iteration), 
-           qD = plant_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(Method == "Observed")
-d_plant1_lines = d_plant1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_plant1_lines = d_plant1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-
-#Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_plant1_point, col = "black", size=2) + 
-ylab("Plant species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000)+
-ggtitle("(b)")+
-theme(plot.margin = margin(r = 12))
-
-#Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
-#Create dataset
-d_int = int_output$iNextEst$coverage_based
-#Check cols
-d_int$col = d_int$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = interactions_curves$Study_sampled, 
-           Method = as.character(interactions_curves$iteration), 
-           qD = interactions_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(Method == "Observed")
-d_int1_lines = d_int1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_int1_lines = d_int1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Unique interactions") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000) +
-ylim(0, 35000)+
-ggtitle("(c)")
-
-#----------------#
-#Pollinators
-#----------------#
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-o1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.03, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
-xlab("Pollinator species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.35) + 
-annotate(geom="text", x=1590, y=0.60, label="50th percentile",color="black", size = 3) +
-annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=200, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3)  +
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
-                ylim=c(0,0.7), clip="off") +
-ggtitle("(e)")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-o2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.03, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
-xlab("Plant species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.35) + 
-annotate(geom="text", x=1010, y=0.60, label="50th percentile",color="black", size = 3) + 
-annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=150, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3) + 
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
-                ylim=c(0, 0.7), clip="off") +
-ggtitle("(f)")
-
-#Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_poll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Create dataset
-d_pp = plant_poll_output$iNextEst$coverage_based
-#Check cols
-d_pp$col = d_pp$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           Method = as.character(plant_poll_curves$iteration), 
-           qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(Method == "Observed")
-d_pp1_lines = d_pp1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_pp1_lines = d_pp1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col)) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_pp1_point, col = "black") + 
-ylab("Pollinator species") +
-xlab("Plant species") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(d)") +
-scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 3000, by = 1000), limits=c(0,3000)) +
-theme(legend.position = "none")
-
-#Plot all
-p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
-      axis.title = element_text(size=12, face = "bold"),
-      axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-#a/b
-
-```
-
-\newpage
-
-## Habitat type and bioclimatic region
-
-The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation and moors and heathland.
-Overall, the proportion of flower visitors from Lepidoptera and Coleoptera were low across all habitats but Coleopteran flower visitors were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
-Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-Hymenopterans were abundant across all bioclimatic regions and Dipterans were particularly abundant in the Boreal, Alpine and Atlantic regions.
-Lepidopterans had low proportions across all bioclimatic regions and Coleopterans were only relevant in the Mediterranean region at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S3**) also differed across habitats and bioclimatic regions.
-The habitats sampled by a higher number of studies in the database were `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
-However, the habitats that contain a higher number of sampling sites were `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies were `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and those that contain a higher number of sampling sites were `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can contribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.
-The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Karise is a very large study that only sampled bumblebees and
-#is providing unrealistic results for the Boreal region
-#Filter out
-#Maybe add to caption?
-data1 = data %>% 
-filter(!Study_id == "13_Karise")
-
-#Check number of studies by habitat
-habitat = data1 %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data1 %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(EuPPollNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(EuPPollNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 7),
-      axis.text = element_text(size=6),
-      panel.border = element_rect(size=1)) +
-xlab(NULL) +
-ylab("Proportion of species") +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 10, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = viridis(4),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-      axis.ticks.x = element_line(color="black"),
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 7),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black'),
-plot.margin = margin(0,0,0,0)) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-#Plot panel
-p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Outputs calculated in script 7_3
-#Load lm output connectance ~ latitude
-connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
-connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
-#Load lm output nestedness ~ latitude
-normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
-normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
-#Load values of connectance
-min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
-max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
-mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
-#Load values of nestedness
-min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
-max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
-mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
-#Load kendall corr from connectance and spp
-connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
-connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
-#Load kendall corr from nestedness and spp
-normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
-normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
-#Load z values
-#VAZNULL
-z_percent_vaznull = readRDS("../Data/Manuscript_info/z_percent_vaznull.rds")
-no_diff_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-#CURVEBALL
-z_percent_curveball = readRDS("../Data/Manuscript_info/z_percent_curveball.rds")
-no_diff_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-
-#QUASISWAP
-z_percent_quasiswap = readRDS("../Data/Manuscript_info/z_percent_quasiswap.rds")
-
-no_diff_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-#rd2
-z_percent_rd2 = readRDS("../Data/Manuscript_info/z_percent_rd2.rds")
-
-no_diff_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-
-#Load correlation outputs of residual connectance and nestednes
-resid_connectance_nestedness_corr_tau = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_tau.rds")
-resid_connectance_nestedness_corr_pval = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_pval.rds")
-
-```
-
-## Network properties
-
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
-Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**).
-The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested.
-The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested.
-Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
-(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
-Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes.
-Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
-Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
-Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network with the curveball algorithm).
-The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
-Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).
-Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
-The bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
-
-# Load libraries
-library(dplyr)
-library(ggplot2)
-library(stringr)
-library(patchwork)
-
-# Load plots generated in scripts 7_2_Nestedness and 7_2_Connectance
-
-connectance_species = readRDS("../Data/Manuscript_info/Figure6_connectance_species.rds")
-connectance_latitude = readRDS("../Data/Manuscript_info/Figure6_connectance_latitude.rds")
-nestedness_z_scores = readRDS("../Data/Manuscript_info/nestedness_z_scores.rds")
-nestedness_latitude = readRDS("../Data/Manuscript_info/nestedness_latitude.rds")
-
-#Prepare first connectance plots
-library(betareg)
-model = betareg(Connectance ~ log_geometric_mean_spp, data = connectance_species)
-p1 = ggplot(connectance_species, aes(x = log_geometric_mean_spp, y = Connectance))+
-geom_point(pch = 21, fill = "azure3", size=2.5) +
-geom_line(aes(y = predict(model, connectance_species)), colour= "black")+
-xlab("log(Species)") +
-theme_bw() +
-ggtitle("(a)")
-
-p3 = connectance_latitude %>% 
-ggplot(aes(Latitude, residual_conectance)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = connectance_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
-geom_ribbon(data = connectance_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ylab("Residual connectance") +
-ggtitle("(c)")
-
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
+null_metrics_networks = readRDS("../../Data/Working_files/null_metrics_networks_nullmodel2.rds")
+metrics_by_network = readRDS("../../Data/Working_files/metrics_by_network.rds")
 
@@
 d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-
 d = d %>% 
 mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-d = d %>% filter(!is.na((z_score)))
 
@@
 summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-no_dif_z = z_percent %>% 
-filter(infra_over_represented == "No statistical difference")
-
-p2 = d1 %>% 
-  ggplot(aes(x = z_score)) +
-  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
-                 bins = 60, alpha = 0.5, color = "black", position = "stack") +
-  theme_bw() +
-  coord_cartesian(expand = FALSE) +
-  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-  ylab("Number of networks") +
-  xlab("Z scores (NODF)")  +
- # xlim(-6, 9) +
- # ylim(0, 0.8) +
-  scale_fill_manual(name = "Observed against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
-ggtitle("(b)")+
-xlim(-6,6)
-
-# 2nd Plot
-p4 = nestedness_latitude %>% 
-ggplot(aes(Latitude, Normalised_nestedness)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-theme_bw() +
-ylab("NODFc") +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
-geom_ribbon(data = nestedness_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ggtitle("(d)")
-
-# Layout plots
-#a = (p1 + p2) 
-#b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.position = "right")
-#a / b
-
-# Layout plots
-a = (p1 + p2)  & theme(legend.position = "right", legend.justification = "left")
-b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.justification = "left")
-a / b  & theme(axis.title =  element_text(size=16, face="bold"),
-               plot.title = element_text(size=22),
-               panel.border = element_rect(colour = "black",
-               fill=NA, size=1),
-               axis.text=element_text(size=12))
-
-```
-
-\newpage
-
-# 3 | DISCUSSION
-
-EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level.
-The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records.
-While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-However, although EuPPollNet likely has a higher pollinator coverage, as it did not exclude non-pollinating species (e.g., predators, herbivores, parasites) from the checklists, this likely reflects that most plant and pollinator species are rare and geographically restricted.
-For example, given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
-This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
-In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
-
-Bees are responsible for the majority of the sampled interactions at the metaweb level.
-As not all surveys included all pollinator groups, this result may partly be influenced by the taxonomic groups sampled across studies.
-However, the relevance of bees and other pollinator orders for network topology changed across habitats and bioclimatic regions in accordance to the literature.
-For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
-These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only commonly documented as floral visitors in the Mediterranean region.
-This study cannot determine whether pollination ecologists traditionally document flower-beetle interactions only in the Mediterranean, or if there are fewer flower visitations by beetles outside this region.
-Nevertheless, the high proportion of beetles as floral visitors provides further support for their potential role as pollinators in the Mediterranean [@herrera2019; @leon2022].
-The number of butterfly species and interactions were relatively low compared to the other taxa.
-While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
-This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018].
-This potentially reflects the dominance of intensive grassland habitats in EuPPollNet and their widespread distribution across European landscapes [@isselstein2005], the highly generalised nature of honey bees, their native status and the important role that honeybees are playing as pollinators in Europe, and/or the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants.
-This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
-Rare pollinators can be functionally important for plant species at the landscape level [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to identify these rare species across different regions and to effectively understand and protect plant-pollinator biodiversity.
-
-Consistent with @olesen2002, we find that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
-Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected and less nested networks at lower latitudes in Europe.
-Finally, most empirical networks (~85%) showed a non-nested structure.
-This result contrasts with earlier work suggesting that plant-pollinator networks are typically nested [@bascompte2003] and it aligns with recent evidence showing that nestedness is not a common feature in plant-pollinator networks [@payrato2019].
-
-Despite this database covers a wide range of habitats across `r country_number` countries, it contains temporal and geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, none of the studies in this database sampled nocturnal pollinators, which can impact our view of network structure [@garcia2024], and most studies were conducted during a single flowering season, limiting our ability to evaluate temporal trends of plant-pollinator communities in the face of environmental changes [@alarcon2008; @chacoff2018].
-In addition, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
-This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-The lack of data for Eastern Europe, which contains vast landscapes with semi-natural grasslands experiencing rapid land use change [@sutcliffe2015], and for the Mediterranean region, which is severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023], is particularly concerning.
-These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], and the introduction of alien species [@vila2009; @vanbergen2018].
-Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
-This will allow local and large scale analyses of the status and trends of plant-pollinator communities, effectively informing management and conservation actions.
-
-In conclusion, the EuPPollNet database enables researchers to explore spatial, taxonomic and structural properties of plant-pollinator networks within Europe.
-In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
-Here, we have shown how connectance and nestedness change across their latitudinal range and that nestedness is not a ubiquitous feature of all plant-pollinator networks.
-These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
-The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and trends of plant-pollinator communities.
-Finally, we hope this database becomes an iterative resource that keeps growing and improving over time to better understand and conserve European biodiversity.
-
-\newpage
-
-# REFERENCES
-
-::: {#refs}
-:::
-
-# ACKNOWLEDGEMENTS
-
-We thank all the taxonomists and ecologists that has made this database possible by contributing with their fieldwork data.
-
-# FUNDING INFORMATION
-
-This research was funded by the H2020 European project Safeguard (101003476) and by the Federal State of Saxony-Anhalt (MLU-BioDivFund).
-RT was supported by the Czech Science Foundation (Project No.
-21-24186M).
-NDM, NH, YP and FM were financially supported by the ANR ARSENIC project (grant no.
-14-CE02-0012), the ANR NGB project (grant no.
-17-CE32- 011), the Region Nord-Pas-de-Calais, the CNRS, the French Ministère de l'Enseignement Supérieur et de la Recherche, the Hauts-de-France Region and the European Regional Funds.
-
-# CONFLICT OF INTEREST
-
-None.
-
-# DATA AVAILABILITY
-
-All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
-
-\newpage
-
-# SUPPORTING INFORMATION
-
-**Title**: EuPPollNet: A European database of plant-pollinator networks
-
-**Authors**: `r authors`
-
-**Contains**:
-
-- Supplementary text 1
-
-- Figure S1
-
-- Figure S2
-
-- Figure S3
-
-- Figure S4
-
-- Figure S5
-
-\newpage
-
-**Supplementary text 1**
-
-EuPPollNet habitat definitions adapted from authors and CORINE Land Cover (CLC).
-To guide these definitions, they are connected to habitat types defined in the European Nature Information System (EUNIS 2021; <https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
-Specific examples from EUNIS habitats are provided when they can be precisely linked to EuPPollNet habitat types.
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘dry perennial anthropogenic herbaceous vegetation’, code **V38** ).
-
-2) **Agricultural margins**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘hedgerows’, code **V4** ).
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an urban setting.
-Botanical gardens are included in this category.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘cultivated areas of gardens and parks’, code **V2** ).
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘intensive unmixed crops’, code **V11** ).
-
-5) **Forest/woodland understory**: Any plant community sampled within a wooded group of plants.
-The forest may be situated in an agricultural setting or in a fully natural scenario.
-We have included agroforestry areas as well as open to dense forests in this category but excluded forest that contains sclerophyllous vegetation.
-This category is similar to the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
-
-6) **Semi-natural grassland**: Acidic, neutral, or calcareous unimproved grasslands located at elevations lower than 1000 m that are managed extensively through mowing or grazing but have not been improved, i.e., they have never received significant fertilizer or herbicide inputs.
-This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R** ).
-
-7) **Intensive grassland**: Communities dominated by non-woody, low-growing plants that are heavily influenced by human disturbance, such as agriculture, mowing, moderate to high grazing, or urban environments.
-The plant composition is generally dominated by graminoid species.
-This category partially matches the EUNIS categories of ‘vegetated man made habitats’ (code **V**; e.g., category of ‘artificial grasslands and herb dominated habitats’, code **V4** ) and ‘grasslands and lands dominated by forbs, mosses or lichens’ (code **R** ).
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have included in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.  This category partially matches the **EUNIS habitat type**  ‘coastal habitats’ (code **N**).
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**) but also the category of ‘inland waters’ which is currently under review.
-
-10) **Montane to alpine grasslands**: Communities dominated by non-woody, low-growing plants that experience little or no human disturbance and are found in high-elevation areas (>1,000 m).
-This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R**; e.g., alpine and subalpine grasslands with code  **R4**).
-
-12) **Moors and heathland**: Low-growing woody vegetation (i.e., typically < 5 m) that is characteristic of poorly fertile soils where environmental conditions prevent the natural formation of forests.
-This category is similar to the **EUNIS habitat type** ‘heathland, scrub and tundra’ (code **S**).
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure S1}. Geographical location of all networks in the EuPPollNet database coloured by bioregion."}
-
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude, Bioregion) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_point(data = for_plotting, alpha=0.75,size=1.2,pch=21,stroke = 0.25,aes(x= Longitude, y=Latitude, fill= Bioregion))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-     legend.key = element_blank(), 
-     legend.text = element_text(size = 6), 
-     legend.title = element_text(face ="bold",  size = 8),
-     legend.spacing.x=unit(0.5, "lines"), 
-     legend.key.width=unit(0.5, "lines"), 
-     legend.box.margin=margin(-10,-10,-10,-10),
-     axis.text = element_text(size=6,face="bold"),
-     axis.title = element_text(face="bold", size = 8),
-) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))  +
-ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(fill = guide_legend(override.aes = list(size = 1.5, byrow = TRUE))) +
-scale_fill_manual(values = c("purple3", 
-                               "dodgerblue3", 
-                               "cyan4", 
-                               "forestgreen",
-                               "orange2",
-                               "tomato3",
-                               "khaki2"))
-
-#Plot everything
-map 
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S2}.
-Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
-
-#Prepare figure for supplementary material
-#Coverage across high taxonomic ranks of bees, syrphids and butterflies
-#--------------------------------------------------------------------------------------------#
-#Start from bees----
-#Bee families
-bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
-            "Andrenidae", "Colletidae","Melittidae")
-
-#Species per family within SafeNet
-bee_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-filter(Pollinator_family %in% bee_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-bee_spp_europe =  master_list %>% 
-select(Family, Genus, Species, GenusAndSpecies) %>% 
-filter(Family %in% bee_fam) %>% 
-group_by(Family) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-rename(Pollinator_family = Family) %>% 
-mutate(Group = "Europe")
-
-#Bind both datasets for plotting
-bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
-
-p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-ylab("Number of species") +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Bee families")
-#--------------------------------------------------------------------------------------------#
-#Continue with syrphids----
-#As we don't have subfamilies, let's obtained them from the master list
-syrphids = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
-rename(Pollinator_accepted_name = GenusAndSpecies)
-#Syrphid species safenet
-syrphid_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_family== "Syrphidae") %>% 
-distinct()
-#Add subfamily info
-syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
-#Summarise spp per subfamily
-syrphid_spp_safenet = syrphid_spp_safenet %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Get spp per subfamily at european level
-syrphid_spp_europe = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
-#Plot
-# Plot for syrphid subfamily coverage with adjusted margins
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily, -Spp_number), Spp_number, fill = Group)) +
-  geom_bar(stat="identity", width = .5, position = "dodge", color = "black") +
-  coord_cartesian(expand = FALSE) +
-  xlab(NULL) +
-  ylab(NULL) +
-  theme_bw() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    panel.border = element_rect(size = 1),
-    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 10)),
-    axis.title = element_text(face = "bold", size = 14),
-    axis.text = element_text(face = "bold", size = 12),
-    plot.margin = margin(10, 10, 20, 10)  # Adjusting the margins
-  ) +
-  scale_fill_manual(values = c("tan2", "lightgrey")) +
-  ggtitle("Syrphid subfamilies")
-#--------------------------------------------------------------------------------------------#
-#Continue with butterflies----
-#Do it at family level
-#Create vector with butterfly families
-butt_fam = c(unique(master_list1$family))
-#Species per family within SafeNet
-butterfly_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Lepidoptera") %>% 
-filter(Pollinator_family %in% butt_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-butterfly_spp_europe =  master_list1 %>% 
-select(family, genus, scientificName) %>% 
-group_by(family) %>% 
-summarise(Spp_number = n_distinct(scientificName)) %>% 
-rename(Pollinator_family = family) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
-#Plot
-p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Butterfly families")
-
-#--------------------------------------------------------------------------------------------#
-p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 14, face = "bold"),
-      axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 12))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S3}.
-Barplot indicating the number of networks by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat_bioregion = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat_bioregion = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-c = bind_rows(habitat, bioregion)
-
-ggplot(c, aes(Habitat_bioregion, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_bw()+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")),
-        space = "free", scales = "free_y") +
-ylab("Networks") +
-xlab(NULL) +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.x = element_line(color="black"),
-      strip.text.x = element_text(size=14, face = "bold"),
-  axis.title = element_text(face = "bold", size = 14),
-axis.text = element_text(size=10),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-panel.border = element_rect(colour = "black",
-        fill=NA, size=1))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S4}.
-Association between nestedness (NODFc) and the geometric mean of plant and pollinator species per network", fig.width=3, fig.height=3}
-
-metrics_by_network_spp_number = readRDS("../Data/Working_files/metrics_by_network_spp_number.rds")
-
-metrics_by_network_spp_number %>% 
-ggplot(aes(geometric_mean_spp, Normalised_nestedness)) +
-geom_point(pch = 21, fill = "azure3", size=1.5, stroke=0.25, alpha=1) +
-theme_bw() +
-xlab("Species") +
-ylab("Normalised nestedness (NODFc)") +
-theme(
-    axis.title = element_text(face = "bold", size = 10),
-    axis.text = element_text(size = 8))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S5}.
-Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with the 'quasiswap_count' algorithm (100 null models for each network).
-The quasiswap_count algorithm maintains row and column sums with fixed connectance."}
-
-library(dplyr)
-library(ggplot2)
-
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-
-#Finally, find % of networks that differ statistically
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-d = d %>% filter(!is.na((z_score)))
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-#Calculate percentages
-z_percent = d1 %>% 
-group_by(infra_over_represented) %>% 
-summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-no_dif_z = z_percent %>% 
-filter(infra_over_represented == "No statistical difference")
 

### 2024-10-02 Manuscript/EuPPollNet_Ms.qmd
Minor edits on refs

@@
 Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
-Consequently, we provide an estimate for the total number of pollinators across the European continent.
+Consequently, we provide an estimate for the total number of potential pollinators across the European continent.
 
@@
 The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
+However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort, the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
 When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
@@
 Note that residual connectance and normalised nestedness showed a moderate significant negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ < `r round(resid_connectance_nestedness_corr_pval, 2)`).
-Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**).
+Empirical networks did not show statistically different nestedness (NODF) to the simulated ones (**Figure 6b** and **Figure S5**).
 The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested.
 The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested.
-Note that in both cases, NODFc is calculated on binarised matrices.
+Note that in both cases, NODF is calculated on binarised matrices.
 
@@
 Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected at lower latitudes in Europe.
+These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected networks at lower latitudes in Europe.
 However, our results cannot be extrapolated to lower latitudes outside Europe, as tropical systems might behave differently.
@@
 - Figure S5
+
+- Figure S6
 

### 2024-10-02 Manuscript/EuPPollNet_Ms.qmd
minor edits

@@
 ```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S6}.
-Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with a probabilistic null model from Bascompte et al., 2003 (100 null models for each network).
-This is the less restrictive approach, as it does not constrain the realised species degree.
-It first calculates the probability for each cell, and then if a random number between 0 and 1 exceeds this probability, an observed interaction is assigned to the cell.
-This shows that when the realised degree is not constrained, the number of networks that are statistically more nested is notably higher, highlighting that the nested structure is likely a byproduct of the species degree distribution."}
+Distribution of z-scores comparing the nestedness (NODF) of empirical networks with their randomized counterparts with the probabilistic null model from Bascompte et al., 2003 (100 null models for each network).
+This is a less restrictive approach, as it does not completely constrain the realized species degree distributions.
+In a nutshell, the null model first calculates the probability of interaction for each cell, and then if a random number between 0 and 1 exceeds this probability, an observed interaction is assigned to the cell.
+This shows that when the realised degree is not fully constrained, the number of networks that are statistically more nested is notably higher, highlighting that the nested structure is likely a byproduct of the species degree distribution."}
 

### 2024-10-03 Manuscript/Additional_information.qmd
journal format style

--- 2024-10-02 Manuscript/EuPPollNet_Ms.qmd
@@
 
-```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr) #For handling fata 
-library(sf) #For handling coordinates and plotting
-library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
-library(readr)
-library(phytools)
-#Load libraries for plotting
-library(giscoR) #Maps
-library(ggstar) #cool shapes as points
-library(ggplot2) #general plotting
-library(ggtree) #phylo trees
-library(ggpattern) #cool shapes in barplots
-library(patchwork) #plot organisation
-library(cowplot) #plot organisation
-library(viridis) #colours
-library(ggnewscale)
-library(ggtreeExtra)
-library(kableExtra)
-
-#In case of conflicts select the main dplyr functions
-library(conflicted)
-conflicts_prefer(dplyr::filter)
-conflicts_prefer(dplyr::select)
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
-
-#Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
-#Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
-                    header = T)
-#Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")
-
-#Work around to get the three cols
-#Note that month 1 and day 1 indicates the presence of a missing value
-#At least in Europe :)
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-
-#Check number of distinct networks
-n_networks = data %>% 
-mutate(Network_id = paste0(Network_id,"_", Year)) %>% 
-distinct(Network_id) %>% 
-nrow(.)
-
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Check number of distinct habitats
-n_habitats = data %>% 
-select(EuPPollNet_habitat) %>% 
-distinct() %>% 
-nrow(.)
-#Read author list 
-authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
-#Print authors
-authors = authors %>% 
-summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
-pull()
-#Read affiliation list 
-aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
-aff_list = aff_list %>% 
-summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
-pull()
-
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Calculate mean coverage and deviation
-coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
-mean_coverage = mean(coverage)
-sd_coverage = sd(coverage)
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-#Prepare table with all info 
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude" ~ "North-south position of the observed interaction in decimal degrees",
-        Variable == "Longitude" ~ "East-west position of the observed interaction in decimal degrees",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_original_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_original_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
-        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
-
-    TRUE ~ "Other"
-  ))
-
-#Check number of species and percentage of interactions of poll. groups
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-order_info = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
-          Species = n_distinct(Pollinator_accepted_name))
-
-Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
-select(Interactions) %>% pull()
-
-#Check percentage of studies with flower counts
-floral_counts_percent = data  %>%
-select(Study_id, Flower_data) %>% 
-distinct() %>% 
-group_by(Flower_data) %>% 
-summarise(Percentage = n()/ nrow(.)*100) %>% 
-filter(Flower_data == "Yes")
-
-#Check percentage of apis mellifera records
-all_records = data %>% nrow(.) 
-apis_records = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-apis_percent = apis_records/all_records * 100
-#Check average pecentage of interactions of apis by net
-a = data %>% 
-group_by(Network_id) %>% 
-summarise(n_interactions = length(Pollinator_accepted_name))
-b = data %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-summarise(n_interactions_apis = length(Pollinator_accepted_name))
-c = left_join(a,b)
-apis_average_network = c %>% 
-mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
-ungroup() %>% 
-summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
-pull()
-
-#Check number of species from Hymenop, Dipte, Lepidop and Coleop
-
-pollinator_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-pollinator_spp_main_orders = data %>% 
-select(Pollinator_order, Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera"|
-       Pollinator_order == "Diptera"|
-       Pollinator_order == "Lepidoptera"|
-       Pollinator_order == "HymenoColeopteraptera") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-spp_percent_main_order = pollinator_spp_main_orders/pollinator_spp *100
-
-#Calculate percentage of studies that recorded the main 4 orders
-metadata = readRDS("../Data/3_Final_data/Metadata.rds")
-#Some studies are divided by year
-#Get an average of the pollinator percentages across years first
-#Fix NA's
-metadata_study = metadata %>%
-group_by(Study_id) %>% 
-mutate(Hymenoptera_percent = as.numeric(na_if(Hymenoptera_percent, "NA"))) %>%  
-mutate(Diptera_percent = as.numeric(na_if(Diptera_percent, "NA"))) %>%  
-mutate(Lepidoptera_percent = as.numeric(na_if(Lepidoptera_percent, "NA"))) %>%  
-mutate(Coleoptera_percent = as.numeric(na_if(Coleoptera_percent, "NA"))) 
-#Calculate average
-mean_study = metadata_study  %>% 
-summarise(Hymenoptera_percent = mean(Hymenoptera_percent, na.rm = T),
-          Diptera_percent = mean(Diptera_percent, na.rm = T),
-          Lepidoptera_percent = mean(Lepidoptera_percent, na.rm = T),
-          Coleoptera_percent = mean(Coleoptera_percent, na.rm = T)) %>% 
- mutate(across(c(Hymenoptera_percent, Diptera_percent, Lepidoptera_percent, Coleoptera_percent),
-                ~ ifelse(is.nan(.), NA, .))) 
-#Now calculate percentage of studies that record any of these 4 orders
-mean_percent_order = mean_study %>%
-  summarise(
-    Hymenoptera_percent_metaweb = sum(!is.na(Hymenoptera_percent)) / nrow(.) * 100,
-    Diptera_percent_metaweb = sum(!is.na(Diptera_percent)) / nrow(.) * 100,
-    Lepidoptera_percent_metaweb = sum(!is.na(Lepidoptera_percent)) / nrow(.) * 100,
-    Coleoptera_percent_metaweb = sum(!is.na(Coleoptera_percent)) / nrow(.) * 100)
-
-Hymenoptera_percent_metaweb = mean_percent_order$Hymenoptera_percent_metaweb
-Diptera_percent_metaweb = mean_percent_order$Diptera_percent_metaweb
-Lepidoptera_percent_metaweb = mean_percent_order$Lepidoptera_percent_metaweb
-Coleoptera_percent_metaweb = mean_percent_order$Coleoptera_percent_metaweb
-
-#Check percentages of hymenoptea groups
-hymenoptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
-          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
-          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))
-
-bee_percent = hymenoptera_percent_by_group$Bee_studies
-bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
-all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies
-
-#Check percentages of diptera groups
-diptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
-          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
-          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
-                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)
-
-syrphid_percent = diptera_percent_by_group$Syrphid_studies
-all_diptera_percent = diptera_percent_by_group$Diptera_studies
-other_diptera_percent = diptera_percent_by_group$other_diptera_studies
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first four and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks.
-One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
-We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
-We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified.
-Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from `r n_habitats` different natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphid species at the European level.
-
-**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
-The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
-
-**KEYWORDS**\
-\begingroup
-\small
-Angiosperms, connectance, flowering plants, nestedness, plant-pollinator networks, pollinators, species interactions 
-\endgroup
-\pagebreak
-
-# 1 | INTRODUCTION
-
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species status, trends, and conservation [@heberling2021; @zattara2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano2003] or nestedness [@bascompte2003].
-Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
-Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @garcia2024].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
-Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness.
-This emphasizes the importance of providing data in its rawest possible form in datasets that will be utilized for synthesis and macro-ecological studies.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023], along with occurrence data [@zattara2021], is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, generating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level.
-For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
-Assembling and curating the existing information on EU plant-pollinator networks will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on plant-pollinator interactions at the European level.
-The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
-
-EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks.
-For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where specialists species interact only with subsets of generalists species, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
-
-Overall, EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
-This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
-We expect that EuPPollNet can be used to evaluate macro-ecological processes in plant-pollinator networks, guide conservation planning, and set a baseline for global change research.
-
-# 2 | METHODS
-
-## Data acquisition
-
-The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
-As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; <https://doi.org/10.3030/101003476>), first, data was directly requested from members of the Safeguard project in May 2022.
-Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024.
-These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
-While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019].
-This approach maximized the potential number of studies that could be incorporated in this database.
-The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'.
-To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
-
-## Dataset description
-
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations).
-The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.
-Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they account for a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
-Hymenoptera and Diptera contain the highest number of species, with each comprising approximately 1,000 species in the database.
-However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
-Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have selected only a single location per study.
-(\\textbf{b}) Number of studies by year in EuPPollNet.
-(\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}.
-The total number of species and interactions is indicated in parentheses at the bottom."}
-
-#Fix dates
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p2  = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-   axis.ticks.x = element_line(color="black"),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -5))
-
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-#Calculate spp number and interactions separately
-data1_spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
-
-data1_int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))
-
-data1 = left_join(data1_spp, data1_int)
-
-#Select main orders
-spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
-pull()
-#Select main orders
-int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Interactions = length(Pollinator_accepted_name)) %>% 
-pull()
-
-#Combine Interactions and Species into a single column
-data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
-
-#Set order of levels for plotting
-data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))
-
-#Plot
-p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
-geom_bar(stat = "identity", width = 0.7) +
-theme_classic() +
-scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator\n order") +
-xlab(NULL) +
-ylab("Percentage") + 
-scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
-theme(axis.text = element_text(size=6), 
-     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
-     axis.title.x = element_text(face="bold", size = 7),
-     legend.position = "top",
-     legend.title = element_text(size = 5, face = "bold"),
-     legend.text = element_text(size = 5),
-     legend.key.size = unit(3, "mm"),
-     legend.margin=margin(0,2,0,0),
-     legend.box.margin=margin(-9,-9,-6,-6)) +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))+
-ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -6)) +
-scale_x_discrete(labels = c("Species\n (2,172)", "Interactions\n (358,049)"))
-
-#plot2 = p3 / p1 / p2 
-plot2 = p2 / p3
-
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.485, 0.8))
-
-```
-
-\newpage
-
-## Data structure
-
-The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble)` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
-These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29).
-The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30).
-Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies.
-To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year.
-Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.
-
-\newpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-#Small edit to column number
-my_tibble = my_tibble %>% 
-mutate(Variable = paste0(seq(1:length(Variable)),". ", Variable))
-
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.} Column names and their descriptions within the EuPPollNet database.") %>%
-  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
-kable_styling(position = "center")
-
-```
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and harmonized in R using **rgbif** [@chamberlain2022].
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-In addition, taxonomic information at genus, family and order level was downloaded for each species.
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and manually fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023], and butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-Coleoptera species names were only checked against the GBIF checklist.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the EuPPollNet database at the European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only plants occurring in Europe and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we only evaluated the taxonomic coverage of groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
-While there is not a good understanding of pollinator diversity in other taxonomic groups (e.g., beetles), it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
-Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
-Consequently, we provide an estimate for the total number of pollinators across the European continent.
-
-Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
-The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it using the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**.
-For plants, the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of the **rtree** package [@li2023].
-
-## Sampling coverage
-
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks.
-In addition, we computed the  accumulation curve of pollinator species with an increasing number of plant species as an indicator for how many pollinator species are likely responsible for the pollination of flowering plants [e.g., @kleijn2015 for crops].
-The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
-The different rarefied curves were complemented with 100 bootstrapped accumulation curves.
-
-## Habitat type and bioclimatic region
-
-We describe the habitat type for each site using information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022], visual inspection of Google Earth imagery and the habitat classification from the authors.
-These different habitat categories (see definitions in supplementary text) allow a quick comparison and understanding of the habitat types from the database.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned a biogeographical region to each site.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites using a spatial joint from the **sf** package [@pebesma2018].
-
-## Network analyses
-
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network and examined how these network metrics change across different latitudes and bioclimatic regions in Europe.
-We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
-We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics.
-As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-The relationship between residual connectance and species richness was investigated using a beta regression.
-This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1.
-The model was implemented using the betareg package [@cribari2010].
-We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species in comparison to other nestedness metrics that change with network size [@song2017].
-This metric was calculated using the **maxnodf** package [@hoeppke2021].
-Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, to quantify how connectance and nestedness change with network size, we determined their association with the number of species per network using the Kendall rank correlation coefficient.
-
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
-The z-score approach only compares each unique network against their randomized versions, avoiding the influence of network size.
-As NODF is a metric computed from binary matrices, we calculated 100 null models for the binarized version of each network using the 'curveball' algorithm [@strona2014].
-This algorithm implements the configuration model (i.e.
-random rewiring of all links, without self-links or double links) and thus keeps the exact number of connections per species (i.e., realised degree).
-Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented a null model that reorganizes the quantitative networks before binarising them.
-To that end, we used the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species.
-Each empirical network was randomized 100 times using each method.
-These null models were implemented with the **vegan** [@oksanen2013] package.
-Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite** [@dormann2008].
-
-\newpage
-
-# 3 | RESULTS
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Taxonomic coverage outputs
-#Find number of plant and pollinator species
-##Pollinators
-poll_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-##Plants
-plant_spp = data %>% 
-select(Plant_rank, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-nrow(.)
-
-#Overall plant coverage
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Average plant family coverage
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData") %>% 
-filter(fam_group == "B")
-#Average coverage per family
-mean_plant_fam_cov = d1 %>% 
-ungroup() %>% 
-summarise(mean_plant_fam_cov = mean(Percent)) %>% 
-pull()
-
-#Percent of families with interaction records
-all_fam = nrow(d1)
-fam_with_data = d1 %>% 
-filter(Percent > 0) %>% 
-nrow()
-percent_plant_fam_with_data = fam_with_data/all_fam * 100
-
-#Calculate phylogenetic signal of presence-absence information
-#Read data and plot it
-tree = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-#Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-#select(label, categorical) %>% 
-ungroup() 
-#Check phylogenetic signal
-d1 = d1 %>% arrange(factor(label, levels = tree@phylo$tip.label))
-
-v = phylo_trait$categorical
-names(v) = tree@phylo$tip.label
-
-phylo_signal_plant = phylosig(tree@phylo, v, method="lambda", test=TRUE)
-lambda_phylosig_plants = phylo_signal_plant$lambda
-p_phylosig_plants = phylo_signal_plant$P
-
-#Load pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-#Check percentage of interactions from bees with and without apis
-#Check percentage of bees in the database
-bee_fam = c("Apidae", "Megachilidae", "Halictidae",
-            "Andrenidae", "Colletidae", "Melittidae")
-all_rows = data %>% nrow(.)
-#
-bee_rows = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-nrow(.)
-#Bee interactions
-bee_interactions_percent = bee_rows/all_rows *100
-#Eclude now apis
-bee_rows_non_apis = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-#
-bee_interactions_non_apis_percent = bee_rows_non_apis/bee_rows *100
-
-#Coverage across bee genera
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-all_rows_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-nrow(.)
-#
-data_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-filter(Percent > 0) %>% 
-nrow(.)
-#
-bee_genera_percent_coverage = data_bees/all_rows_bees * 100
-#Check average coverage across bee genera
-average_coverage_bee_genera = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-ungroup() %>% 
-summarise(average_coverage_bee_genera = mean(Percent)) %>% 
-pull()
-
-#Calculate phylogenetic signal of presence-absence information
-##Read data and plot it
-tree = readRDS("../Data/Manuscript_info/bee.tree100.rds")
-##Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-ungroup() 
-##Conduct left join to tibble of tip labels to organise order
-v = tibble(Genus1 = tree$tip.label) %>% 
-mutate(Genus1 = str_to_title(word(Genus1,1,sep = "_")))
-#
-v1 = left_join(v, phylo_trait)
-#
-v2 = v1$categorical
-names(v2) = tree$tip.label
-#
-phylo_signal_bees = phylosig(tree, v2, method="lambda", test=TRUE)
-lambda_phylosig_bees = phylo_signal_bees$lambda
-p_phylosig_bees = phylo_signal_bees$P
-
-#Check number of studies by habitat
-studies_by_habitat = data %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of sampling sites by habitat
-networks_by_habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of studies by bioregion
-studies_by_bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check number of sampling sites by bioregion
-networks_by_bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check percentage of Apis mellifera across networks
-apis_percent_networks = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>%
-group_by(Study_id) %>% 
-mutate(Apis = "Yes") %>% 
-distinct(Study_id, Apis) %>% 
-nrow() / nlevels(factor(data$Study_id)) *100
-
-#Load number of species from checklists and extrapolated one
-bee_potential_spp = readRDS("../Data/Manuscript_info/bee_potential_spp.RData")
-syrphid_potential_spp = readRDS("../Data/Manuscript_info/syrphid_potential_spp.RData")
-lepidoptera_potential_spp = readRDS("../Data/Manuscript_info/lepidoptera_potential_spp.RData")
-extrapolated_other_spp = readRDS("../Data/Manuscript_info/other_taxa_potential_spp.RData")
-extrapolated_other_spp = round(extrapolated_other_spp)
-```
-
-## Taxonomic coverage
-
-Europe hosts approximately over 5,000 species of pollinators, including `r prettyNum(bee_potential_spp, big.mark=",")` bee species `r prettyNum(syrphid_potential_spp, big.mark=",")` syrphid species, `r prettyNum(lepidoptera_potential_spp, big.mark=",")` butterflies species, and about 1,700 species from other taxonomic groups, as estimated based on EuPPollNet average taxonomic coverage.
-Additionally, there are around 25,000 species of plants that benefit from animal pollination when excluding non-biotic pollinated species.
-EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
-The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
-The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic and taxonomic coverage of bee genera at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple.
-Additionally, the coverage of species recorded in EuPPollNet for each genus is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database, out of the total number of bee species in Europe.
-Dark grey bars represent the total number of species per genus at European level."}
-
-#Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
-
-#Circular layour
-ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.065, size=2.9, fontface=1) +
-scale_size(range = c(0,4.5), guide = 'none')  +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-guides(fill=guide_legend(), size = guide_legend()) +
-theme(legend.position = "bottom") +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
-                    pwidth=0.13, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0.148,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-guides(fill = "none")+
-scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
-geom_fruit(geom=geom_bar,
-                    mapping=aes(y=label, x=Spp_number_Europe),
-                    pwidth=0.12, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = -0.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
-new_scale_fill() +
-geom_point2(aes(subset=node==131), color='orange', size=3) +
-geom_point2(aes(subset=node==84), color='orange', size=3) +
-geom_point2(aes(subset=node==73), color='orange', size=3) +
-geom_point2(aes(subset=node==74), color='orange', size=3) +
-geom_point2(aes(subset=node==111), color='orange', size=3)+
-geom_point2(aes(subset=node==91), color='orange', size=3) +
-annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
-annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
-annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
-annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
-annotate("text", x = 2.25, y = 26.8, label = "Megachilidae", size= 3)+
-annotate("text", x = 2.5, y = 51, label = "Apidae", size= 3) +
-coord_cartesian(clip = "off")+
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.6,0),
-      legend.direction = "horizontal")
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-Phylogenetic and taxonomic coverage of the plant families at European level.
-The number of interactions recorded per family in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple.
-Additionally, the coverage of species recorded in EuPPollNet for each family is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database out of the total number of bee species per family in Europe.
-Dark grey bars represent the total number of species per family at European level.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Sampling coverage outputs
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Coverage
-poll_sampling_coverage = poll_output$DataInfo$SC *100
-plant_sampling_coverage = plant_output$DataInfo$SC *100
-int_coverage = int_output$DataInfo$SC *100
-plant_poll_coverage = plant_poll_output$DataInfo$SC *100
-
-#Increase doubling effort
-##Pollinators
-poll_increase = poll_output$iNextEst$coverage_based
-poll_increase1 = poll_increase %>% filter(Method == "Observed") %>% pull(qD)
-poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
-poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
-##Plants
-plant_increase = plant_output$iNextEst$coverage_based
-plant_increase1 = plant_increase %>% filter(Method == "Observed") %>% pull(qD)
-plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
-##Interactions
-int_increase = int_output$iNextEst$coverage_based
-int_increase1 = int_increase %>% filter(Method == "Observed") %>% pull(qD)
-int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
-int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
-##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$coverage_based
-plant_poll_increase1 = plant_poll_increase %>% filter(Method == "Observed") %>% pull(qD)
-plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
-
-#Prepare results from common and rare poll across sampling sites
-#Pollinators
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-poll_total = nrow(poll_spread)
-polls_under_one_percent = poll_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-polls_under_one_percent = polls_under_one_percent/poll_total * 100
-#Find most common pollinator
-most_common_poll = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Pollinator_accepted_name)
-#Find most common poll percent
-most_common_poll_percent = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-#Plants
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-plant_total = nrow(plant_spread)
-plants_under_one_percent = plant_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-plants_under_one_percent = plants_under_one_percent/plant_total * 100  
-#Find most common plant
-most_common_plant = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Plant_accepted_name)
-#Find most common plant percent
-most_common_plant_percent = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-```
-
-## Sampling coverage
-
-The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_sampling_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**).
-The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
-When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
-We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**).
-The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and the number of unique pairwise interactions across networks.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
-The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks.
-Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network.
-Note that \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
-
-#Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
-#Create dataset
-d_poll = poll_output$iNextEst$coverage_based
-#Check cols
-d_poll$col = d_poll$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           Method = as.character(pollinators_curves$iteration), 
-           qD = pollinators_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(Method == "Observed")
-d_poll1_lines = d_poll1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_poll1_lines = d_poll1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col),size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_poll1_point, col = "black", size=2) + 
-ylab("Pollinator species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000) +
-ggtitle("(a)") +
-theme(plot.margin = margin(r = 12))
-
-#Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
-#Create dataset
-d_plant = plant_output$iNextEst$coverage_based
-#Check cols
-d_plant$col = d_plant$Method
-#Now create tibble with same structure
-a_plant = tibble(t = plant_curves$Study_sampled, 
-           Method = as.character(plant_curves$iteration), 
-           qD = plant_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(Method == "Observed")
-d_plant1_lines = d_plant1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_plant1_lines = d_plant1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-
-#Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_plant1_point, col = "black", size=2) + 
-ylab("Plant species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000)+
-ggtitle("(b)")+
-theme(plot.margin = margin(r = 12))
-
-#Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
-#Create dataset
-d_int = int_output$iNextEst$coverage_based
-#Check cols
-d_int$col = d_int$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = interactions_curves$Study_sampled, 
-           Method = as.character(interactions_curves$iteration), 
-           qD = interactions_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(Method == "Observed")
-d_int1_lines = d_int1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_int1_lines = d_int1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Unique interactions") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000) +
-ylim(0, 35000)+
-ggtitle("(c)")
-
-#----------------#
-#Pollinators
-#----------------#
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-o1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.03, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
-xlab("Pollinator species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.35) + 
-annotate(geom="text", x=1590, y=0.60, label="50th percentile",color="black", size = 3) +
-annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=200, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3)  +
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
-                ylim=c(0,0.7), clip="off") +
-ggtitle("(e)")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-o2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.03, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
-xlab("Plant species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.35) + 
-annotate(geom="text", x=1010, y=0.60, label="50th percentile",color="black", size = 3) + 
-annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=150, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3) + 
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
-                ylim=c(0, 0.7), clip="off") +
-ggtitle("(f)")
-
-#Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_poll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Create dataset
-d_pp = plant_poll_output$iNextEst$coverage_based
-#Check cols
-d_pp$col = d_pp$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           Method = as.character(plant_poll_curves$iteration), 
-           qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(Method == "Observed")
-d_pp1_lines = d_pp1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_pp1_lines = d_pp1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col)) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_pp1_point, col = "black") + 
-ylab("Pollinator species") +
-xlab("Plant species") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(d)") +
-scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 3000, by = 1000), limits=c(0,3000)) +
-theme(legend.position = "none")
-
-#Plot all
-p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
-      axis.title = element_text(size=12, face = "bold"),
-      axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-#a/b
-
-```
-
-\newpage
-
-## Habitat type and bioclimatic region
-
-The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation and moors and heathland.
-Overall, the proportion of flower visitors from Lepidoptera and Coleoptera were low across all habitats but Coleopteran flower visitors were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
-Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-Hymenopterans were abundant across all bioclimatic regions and Dipterans were particularly abundant in the Boreal, Alpine and Atlantic regions.
-Lepidopterans had low proportions across all bioclimatic regions and Coleopterans were only relevant in the Mediterranean region at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S3**) also differed across habitats and bioclimatic regions.
-The habitats sampled by a higher number of studies in the database were `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
-However, the habitats that contain a higher number of sampling sites were `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies were `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and those that contain a higher number of sampling sites were `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can contribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.
-The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day and site.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Karise is a very large study that only sampled bumblebees and
-#is providing unrealistic results for the Boreal region
-#Filter out
-#Maybe add to caption?
-data1 = data %>% 
-filter(!Study_id == "13_Karise")
-
-#Check number of studies by habitat
-habitat = data1 %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data1 %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(EuPPollNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(EuPPollNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 7),
-      axis.text = element_text(size=6),
-      panel.border = element_rect(size=1)) +
-xlab(NULL) +
-ylab("Proportion of species") +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 10, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = viridis(4),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-      axis.ticks.x = element_line(color="black"),
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 7),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black'),
-plot.margin = margin(0,0,0,0)) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-#Plot panel
-p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Outputs calculated in script 7_3
-#Load lm output connectance ~ latitude
-connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
-connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
-
-connectance_lat_pval = ifelse(connectance_lat_pval <0.01, 0.01, connectance_lat_pval)
-
-#Load lm output nestedness ~ latitude
-normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
-normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
-
-#Load values of connectance
-min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
-max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
-mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
-#Load values of nestedness
-min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
-max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
-mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
-#Load kendall corr from connectance and spp
-connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
-connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
-#Load kendall corr from nestedness and spp
-normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
-normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
-#Load z values
-#VAZNULL
-z_percent_vaznull = readRDS("../Data/Manuscript_info/z_percent_vaznull.rds")
-no_diff_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-#CURVEBALL
-z_percent_curveball = readRDS("../Data/Manuscript_info/z_percent_curveball.rds")
-no_diff_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-
-#QUASISWAP
-z_percent_quasiswap = readRDS("../Data/Manuscript_info/z_percent_quasiswap.rds")
-
-no_diff_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-#rd2
-z_percent_rd2 = readRDS("../Data/Manuscript_info/z_percent_rd2.rds")
-
-no_diff_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-
-#Load correlation outputs of residual connectance and nestednes
-resid_connectance_nestedness_corr_tau = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_tau.rds")
-resid_connectance_nestedness_corr_pval = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_pval.rds")
-
-#To avoid having values under 0.01
-resid_connectance_nestedness_corr_pval = ifelse(resid_connectance_nestedness_corr_pval < 0.01, 0.01, resid_connectance_nestedness_corr_pval)
-
-```
-
-## Network properties
-
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and, as expected, were not independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
-Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
-(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ < `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ ~ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
-Overall, networks towards higher latitudes showed lower residual connectance but similar nestedness compared to networks located at lower latitudes.
-Note that residual connectance and normalised nestedness showed a moderate significant negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ < `r round(resid_connectance_nestedness_corr_pval, 2)`).
-Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**).
-The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested.
-The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested.
-Note that in both cases, NODFc is calculated on binarised matrices.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
-Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
-Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network with the curveball algorithm).
-The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
-Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).
-Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
-The solid fitted line indicates a significant association, while the dashed lines indicate a non-significant association.
-The bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
-
-# Load libraries
-library(dplyr)
-library(ggplot2)
-library(stringr)
-library(patchwork)
-
-# Load plots generated in scripts 7_2_Nestedness and 7_2_Connectance
-
-connectance_species = readRDS("../Data/Manuscript_info/Figure6_connectance_species.rds")
-connectance_latitude = readRDS("../Data/Manuscript_info/Figure6_connectance_latitude.rds")
-nestedness_z_scores = readRDS("../Data/Manuscript_info/nestedness_z_scores.rds")
-nestedness_latitude = readRDS("../Data/Manuscript_info/nestedness_latitude.rds")
-
-#Prepare first connectance plots
-library(betareg)
-model = betareg(Connectance ~ log_geometric_mean_spp, data = connectance_species)
-p1 = ggplot(connectance_species, aes(x = log_geometric_mean_spp, y = Connectance))+
-geom_point(pch = 21, fill = "azure3", size=2.5) +
-geom_line(aes(y = predict(model, connectance_species)), colour= "black")+
-xlab("log(Species)") +
-theme_bw() +
-ggtitle("(a)")
-
-p3 = connectance_latitude %>% 
-ggplot(aes(Latitude, residual_conectance)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = connectance_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
-geom_ribbon(data = connectance_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ylab("Residual connectance") +
-ggtitle("(c)")
-
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-
-#Finally, find % of networks that differ statistically
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-d = d %>% filter(!is.na((z_score)))
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-#Calculate percentages
-z_percent = d1 %>% 
-group_by(infra_over_represented) %>% 
-summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-no_dif_z = z_percent %>% 
-filter(infra_over_represented == "No statistical difference")
-
-p2 = d1 %>% 
-  ggplot(aes(x = z_score)) +
-  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
-                 bins = 60, alpha = 0.5, color = "black", position = "stack") +
-  theme_bw() +
-  coord_cartesian(expand = FALSE) +
-  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-  ylab("Number of networks") +
-  xlab("Z scores (NODF)")  +
- # xlim(-6, 9) +
- # ylim(0, 0.8) +
-  scale_fill_manual(name = "Observed against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
-ggtitle("(b)")+
-xlim(-6,6)
-
-# 2nd Plot
-p4 = nestedness_latitude %>% 
-ggplot(aes(Latitude, Normalised_nestedness)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-theme_bw() +
-ylab("NODFc") +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE, linetype = "dashed")+
-geom_ribbon(data = nestedness_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ggtitle("(d)")
-
-# Layout plots
-#a = (p1 + p2) 
-#b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.position = "right")
-#a / b
-
-# Layout plots
-a = (p1 + p2)  & theme(legend.position = "right", legend.justification = "left")
-b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.justification = "left")
-a / b  & theme(axis.title =  element_text(size=16, face="bold"),
-               plot.title = element_text(size=22),
-               panel.border = element_rect(colour = "black",
-               fill=NA, size=1),
-               axis.text=element_text(size=12))
-
-```
-
-\newpage
-
-# 3 | DISCUSSION
-
-EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level.
-The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records.
-While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-This likely reflects that most plant and pollinator species are rare and geographically restricted.
-For example, given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
-This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
-In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
-
-Bees are responsible for the majority of the sampled interactions at the metaweb level.
-As not all surveys included all pollinator groups, this result may partly be influenced by the taxonomic groups sampled across studies.
-However, the relevance of bees and other pollinator orders for network topology changed across habitats and bioclimatic regions in accordance to the literature.
-For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
-These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only commonly documented as floral visitors in the Mediterranean region.
-This study cannot determine whether pollination ecologists traditionally document flower-beetle interactions only in the Mediterranean, or if there are fewer flower visitations by beetles outside this region.
-Nevertheless, the high proportion of beetles as floral visitors provides further support for their potential role as pollinators in the Mediterranean [@herrera2019; @leon2022].
-The number of butterfly species and interactions were relatively low compared to the other taxa.
-While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
-This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018].
-This potentially reflects the dominance of intensive grassland habitats in EuPPollNet and their widespread distribution across European landscapes [@isselstein2005], the highly generalised nature of honey bees, their native status and above all, the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants.
-This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
-Rare pollinators can be functionally important for plant species at the landscape level [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to identify these rare species across different regions and to effectively understand and protect plant-pollinator biodiversity.
-
-Consistent with @olesen2002, we found that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
-Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected at lower latitudes in Europe.
-However, our results cannot be extrapolated to lower latitudes outside Europe, as tropical systems might behave differently.
-Finally, a result that can be surprising is that most empirical networks (~85%) showed a non-nested structure.
-This result contrasts with earlier work suggesting that plant-pollinator networks are typically nested [@bascompte2003] and it aligns with recent evidence showing that nestedness is not a common feature in plant-pollinator networks when evaluated against restrictive null models that conserve the observed species degree [@payrato2019\; Figure S6].
-Note that while species degree distributions are sufficient to explain the emergence of nestedness, this does not preclude nestedness from being a useful metric for comparison across networks.
-For example, plant-pollinator networks are more nested than plant-herbivore networks precisely because these two network types differ in their composition of generalist and specialist species [@thebault2010].
-
-Although this database covers a wide range of habitats across `r country_number` countries, it contains temporal and geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, none of the studies in this database sampled nocturnal pollinators, which can impact our view of network structure [@garcia2024], and most studies were conducted during a single flowering season, limiting our ability to evaluate temporal trends of plant-pollinator communities in the face of environmental changes [@alarcon2008; @chacoff2018].
-In addition, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
-This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-The lack of data for Eastern Europe, which contains vast landscapes with semi-natural grasslands experiencing rapid land use change [@sutcliffe2015], and for the Mediterranean region, which is severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023], is particularly concerning.
-These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], and the introduction of alien species [@vila2009; @vanbergen2018].
-Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
-This will allow local and large scale analyses of the status and trends of plant-pollinator communities, effectively informing management and conservation actions.
-
-In conclusion, the EuPPollNet database enables researchers to explore spatial, taxonomic and structural properties of plant-pollinator networks within Europe.
-In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
-Here, we have shown how connectance and nestedness change across their latitudinal range, and that plant-pollinator networks are as nested as expected given plant and pollinator generalist levels.
-These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
-The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and trends of plant-pollinator communities.
-Finally, we hope this database becomes an iterative resource that keeps growing and improving over time to better understand and conserve European biodiversity.
-
-\newpage
-
-# REFERENCES
-
-::: {#refs}
-:::
-
 # ACKNOWLEDGEMENTS
@@
 None.
-
-# DATA AVAILABILITY
-
-All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
-
-\newpage
-
-# SUPPORTING INFORMATION
-
-**Title**: EuPPollNet: A European database of plant-pollinator networks
-
-**Authors**: `r authors`
-
-**Contains**:
-
-- Supplementary text 1
-
-- Figure S1
-
-- Figure S2
-
-- Figure S3
-
-- Figure S4
-
-- Figure S5
-
-\newpage
-
-**Supplementary text 1**
-
-EuPPollNet habitat definitions adapted from authors and CORINE Land Cover (CLC).
-To guide these definitions, they are connected to habitat types defined in the European Nature Information System (EUNIS 2021; <https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
-Specific examples from EUNIS habitats are provided when they can be precisely linked to EuPPollNet habitat types.
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘dry perennial anthropogenic herbaceous vegetation’, code **V38** ).
-
-2) **Agricultural margins**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘hedgerows’, code **V4** ).
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an urban setting.
-Botanical gardens are included in this category.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘cultivated areas of gardens and parks’, code **V2** ).
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘intensive unmixed crops’, code **V11** ).
-
-5) **Forest/woodland understory**: Any plant community sampled within a wooded group of plants.
-The forest may be situated in an agricultural setting or in a fully natural scenario.
-We have included agroforestry areas as well as open to dense forests in this category but excluded forest that contains sclerophyllous vegetation.
-This category is similar to the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
-
-6) **Semi-natural grassland**: Acidic, neutral, or calcareous unimproved grasslands located at elevations lower than 1000 m that are managed extensively through mowing or grazing but have not been improved, i.e., they have never received significant fertilizer or herbicide inputs.
-This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R** ).
-
-7) **Intensive grassland**: Communities dominated by non-woody, low-growing plants that are heavily influenced by human disturbance, such as agriculture, mowing, moderate to high grazing, or urban environments.
-The plant composition is generally dominated by graminoid species.
-This category partially matches the EUNIS categories of ‘vegetated man made habitats’ (code **V**; e.g., category of ‘artificial grasslands and herb dominated habitats’, code **V4** ) and ‘grasslands and lands dominated by forbs, mosses or lichens’ (code **R** ).
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have included in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.  This category partially matches the **EUNIS habitat type**  ‘coastal habitats’ (code **N**).
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**) but also the category of ‘inland waters’ which is currently under review.
-
-10) **Montane to alpine grasslands**: Communities dominated by non-woody, low-growing plants that experience little or no human disturbance and are found in high-elevation areas (>1,000 m).
-This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R**; e.g., alpine and subalpine grasslands with code  **R4**).
-
-12) **Moors and heathland**: Low-growing woody vegetation (i.e., typically < 5 m) that is characteristic of poorly fertile soils where environmental conditions prevent the natural formation of forests.
-This category is similar to the **EUNIS habitat type** ‘heathland, scrub and tundra’ (code **S**).
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure S1}. Geographical location of all networks in the EuPPollNet database coloured by bioregion."}
-
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude, Bioregion) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_point(data = for_plotting, alpha=0.75,size=1.2,pch=21,stroke = 0.25,aes(x= Longitude, y=Latitude, fill= Bioregion))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-     legend.key = element_blank(), 
-     legend.text = element_text(size = 6), 
-     legend.title = element_text(face ="bold",  size = 8),
-     legend.spacing.x=unit(0.5, "lines"), 
-     legend.key.width=unit(0.5, "lines"), 
-     legend.box.margin=margin(-10,-10,-10,-10),
-     axis.text = element_text(size=6,face="bold"),
-     axis.title = element_text(face="bold", size = 8),
-) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))  +
-ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(fill = guide_legend(override.aes = list(size = 1.5, byrow = TRUE))) +
-scale_fill_manual(values = c("purple3", 
-                               "dodgerblue3", 
-                               "cyan4", 
-                               "forestgreen",
-                               "orange2",
-                               "tomato3",
-                               "khaki2"))
-
-#Plot everything
-map 
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S2}.
-Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
-
-#Prepare figure for supplementary material
-#Coverage across high taxonomic ranks of bees, syrphids and butterflies
-#--------------------------------------------------------------------------------------------#
-#Start from bees----
-#Bee families
-bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
-            "Andrenidae", "Colletidae","Melittidae")
-
-#Species per family within SafeNet
-bee_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-filter(Pollinator_family %in% bee_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-bee_spp_europe =  master_list %>% 
-select(Family, Genus, Species, GenusAndSpecies) %>% 
-filter(Family %in% bee_fam) %>% 
-group_by(Family) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-rename(Pollinator_family = Family) %>% 
-mutate(Group = "Europe")
-
-#Bind both datasets for plotting
-bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
-
-p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-ylab("Number of species") +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Bee families")
-#--------------------------------------------------------------------------------------------#
-#Continue with syrphids----
-#As we don't have subfamilies, let's obtained them from the master list
-syrphids = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
-rename(Pollinator_accepted_name = GenusAndSpecies)
-#Syrphid species safenet
-syrphid_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_family== "Syrphidae") %>% 
-distinct()
-#Add subfamily info
-syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
-#Summarise spp per subfamily
-syrphid_spp_safenet = syrphid_spp_safenet %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Get spp per subfamily at european level
-syrphid_spp_europe = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
-#Plot
-# Plot for syrphid subfamily coverage with adjusted margins
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily, -Spp_number), Spp_number, fill = Group)) +
-  geom_bar(stat="identity", width = .5, position = "dodge", color = "black") +
-  coord_cartesian(expand = FALSE) +
-  xlab(NULL) +
-  ylab(NULL) +
-  theme_bw() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    panel.border = element_rect(size = 1),
-    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 10)),
-    axis.title = element_text(face = "bold", size = 14),
-    axis.text = element_text(face = "bold", size = 12),
-    plot.margin = margin(10, 10, 20, 10)  # Adjusting the margins
-  ) +
-  scale_fill_manual(values = c("tan2", "lightgrey")) +
-  ggtitle("Syrphid subfamilies")
-#--------------------------------------------------------------------------------------------#
-#Continue with butterflies----
-#Do it at family level
-#Create vector with butterfly families
-butt_fam = c(unique(master_list1$family))
-#Species per family within SafeNet
-butterfly_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Lepidoptera") %>% 
-filter(Pollinator_family %in% butt_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-butterfly_spp_europe =  master_list1 %>% 
-select(family, genus, scientificName) %>% 
-group_by(family) %>% 
-summarise(Spp_number = n_distinct(scientificName)) %>% 
-rename(Pollinator_family = family) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
-#Plot
-p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Butterfly families")
-
-#--------------------------------------------------------------------------------------------#
-p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 14, face = "bold"),
-      axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 12))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S3}.
-Barplot indicating the number of networks by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat_bioregion = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat_bioregion = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-c = bind_rows(habitat, bioregion)
-
-ggplot(c, aes(Habitat_bioregion, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_bw()+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")),
-        space = "free", scales = "free_y") +
-ylab("Networks") +
-xlab(NULL) +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.x = element_line(color="black"),
-      strip.text.x = element_text(size=14, face = "bold"),
-  axis.title = element_text(face = "bold", size = 14),
-axis.text = element_text(size=10),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-panel.border = element_rect(colour = "black",
-        fill=NA, size=1))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S4}.
-Association between nestedness (NODFc) and the geometric mean of plant and pollinator species per network", fig.width=3, fig.height=3}
-
-metrics_by_network_spp_number = readRDS("../Data/Working_files/metrics_by_network_spp_number.rds")
-
-metrics_by_network_spp_number %>% 
-ggplot(aes(geometric_mean_spp, Normalised_nestedness)) +
-geom_point(pch = 21, fill = "azure3", size=1.5, stroke=0.25, alpha=1) +
-theme_bw() +
-xlab("Species") +
-ylab("Normalised nestedness (NODFc)") +
-theme(
-    axis.title = element_text(face = "bold", size = 10),
-    axis.text = element_text(size = 8))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S5}.
-Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with the 'quasiswap_count' algorithm (100 null models for each network).
-The quasiswap_count algorithm maintains row and column sums with fixed connectance."}
-
-library(dplyr)
-library(ggplot2)
-
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-
-#Finally, find % of networks that differ statistically
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-d = d %>% filter(!is.na((z_score)))
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-#Calculate percentages
-z_percent = d1 %>% 
-group_by(infra_over_represented) %>% 
-summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-no_dif_z = z_percent %>% 
-filter(infra_over_represented == "No statistical difference")
-
-# Plot
- d1 %>% 
-  ggplot(aes(x = z_score)) +
-  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
-                 bins = 50, alpha = 0.5, color = "black", position = "stack") +
-  theme_bw() +
-  coord_cartesian(expand = FALSE) +
-  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-  ylab("Number of networks") +
-  xlab("Z scores (NODFc)")  +
- # xlim(-6, 9) +
- # ylim(0, 0.8) +
-  scale_fill_manual(name = "Observed against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
-xlim(-6,6) +
-theme(plot.title = element_text(size = 24),
-      text = element_text(size = 20),
-      axis.title  = element_text(face="bold"))
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S6}.
-Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with a probabilistic null model from Bascompte et al., 2003 (100 null models for each network).
-This is the less restrictive approach, as it does not constrain the realised species degree.
-It first calculates the probability for each cell, and then if a random number between 0 and 1 exceeds this probability, an observed interaction is assigned to the cell.
-This shows that when the realised degree is not constrained, the number of networks that are statistically more nested is notably higher, highlighting that the nested structure is likely a byproduct of the species degree distribution."}
-
-library(dplyr)
-library(ggplot2)
-
-#Read data
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_nullmodel2.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-
-#Finally, find % of networks that differ statistically
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-#Calculate percentages
-z_percent = d1 %>% 
-group_by(infra_over_represented) %>% 
-summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-# Plot
- d1 %>% 
-  ggplot(aes(x = z_score)) +
-  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
-                 bins = 50, alpha = 0.5, color = "black", position = "stack") +
-  theme_bw() +
-  coord_cartesian(expand = FALSE) +
-  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-  ylab("Number of networks") +
-  xlab("Z scores (NODFc)")  +
- # xlim(-6, 9) +
- # ylim(0, 0.8) +
-  scale_fill_manual(name = "Observed against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
-xlim(-6,6) +
-theme(plot.title = element_text(size = 24),
-      text = element_text(size = 20),
-      axis.title  = element_text(face="bold"))
-
-```

### 2024-10-03 Manuscript/EuPPollNet_main_text.qmd
journal format style

--- 2024-10-02 Manuscript/EuPPollNet_Ms.qmd
@@
 
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first four and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
 # Abstract
+
 \vspace*{-5mm}
@@
 
-# ACKNOWLEDGEMENTS
-
-We thank all the taxonomists and ecologists that has made this database possible by contributing with their fieldwork data.
-
-# FUNDING INFORMATION
-
-This research was funded by the H2020 European project Safeguard (101003476) and by the Federal State of Saxony-Anhalt (MLU-BioDivFund).
-RT was supported by the Czech Science Foundation (Project No.
-21-24186M).
-NDM, NH, YP and FM were financially supported by the ANR ARSENIC project (grant no.
-14-CE02-0012), the ANR NGB project (grant no.
-17-CE32- 011), the Region Nord-Pas-de-Calais, the CNRS, the French Ministère de l'Enseignement Supérieur et de la Recherche, the Hauts-de-France Region and the European Regional Funds.
-
-# CONFLICT OF INTEREST
-
-None.
-
 # DATA AVAILABILITY
@@
 All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
-
-\newpage
-
-# SUPPORTING INFORMATION
-
-**Title**: EuPPollNet: A European database of plant-pollinator networks
-
-**Authors**: `r authors`
-
-**Contains**:
-
-- Supplementary text 1
-
-- Figure S1
-
-- Figure S2
-
-- Figure S3
-
-- Figure S4
-
-- Figure S5
-
-- Figure S6
-
-\newpage
-
-**Supplementary text 1**
-
-EuPPollNet habitat definitions adapted from authors and CORINE Land Cover (CLC).
-To guide these definitions, they are connected to habitat types defined in the European Nature Information System (EUNIS 2021; <https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
-Specific examples from EUNIS habitats are provided when they can be precisely linked to EuPPollNet habitat types.
-
-1) **Ruderal vegetation**: Plants growing on highly disturbed sites such as road sides or mineral extraction sites.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘dry perennial anthropogenic herbaceous vegetation’, code **V38** ).
-
-2) **Agricultural margins**: Sides of crops that can include any type of vegetation from low growing plants to trees.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘hedgerows’, code **V4** ).
-
-3) **Green urban areas**: Parks, private gardens or small pastures within an urban setting.
-Botanical gardens are included in this category.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘cultivated areas of gardens and parks’, code **V2** ).
-
-4) **Agricultural land**: Includes any type of crop and any type of vegetation growing within them.
-This category partially matches the category of ‘vegetated man made habitats’ from EUNIS (code **V**; e.g., category of ‘intensive unmixed crops’, code **V11** ).
-
-5) **Forest/woodland understory**: Any plant community sampled within a wooded group of plants.
-The forest may be situated in an agricultural setting or in a fully natural scenario.
-We have included agroforestry areas as well as open to dense forests in this category but excluded forest that contains sclerophyllous vegetation.
-This category is similar to the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
-
-6) **Semi-natural grassland**: Acidic, neutral, or calcareous unimproved grasslands located at elevations lower than 1000 m that are managed extensively through mowing or grazing but have not been improved, i.e., they have never received significant fertilizer or herbicide inputs.
-This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R** ).
-
-7) **Intensive grassland**: Communities dominated by non-woody, low-growing plants that are heavily influenced by human disturbance, such as agriculture, mowing, moderate to high grazing, or urban environments.
-The plant composition is generally dominated by graminoid species.
-This category partially matches the EUNIS categories of ‘vegetated man made habitats’ (code **V**; e.g., category of ‘artificial grasslands and herb dominated habitats’, code **V4** ) and ‘grasslands and lands dominated by forbs, mosses or lichens’ (code **R** ).
-
-8) **Sclerophyllous vegetation**: Any type of system with a dominant shrub community adapted to drought.
-Typical of the Mediterranean region.
-Note, that we have included in this category also woodlands (open coniferous forest) where the shrub community was the main focus of the study.
-This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**).
-
-9) **Beaches, dunes, sands**: Plant communities growing on sandy soil.  This category partially matches the **EUNIS habitat type**  ‘coastal habitats’ (code **N**).
-
-10) **Riparian vegetation**: Plant communities growing on river margins.
-This category partially matches the **EUNIS habitat type**  ‘forest and other wooded land’ (code **T**) but also the category of ‘inland waters’ which is currently under review.
-
-10) **Montane to alpine grasslands**: Communities dominated by non-woody, low-growing plants that experience little or no human disturbance and are found in high-elevation areas (>1,000 m).
-This category partially matches the category of ‘grasslands and lands dominated by forbs, mosses or lichens’ from EUNIS (code **R**; e.g., alpine and subalpine grasslands with code  **R4**).
-
-12) **Moors and heathland**: Low-growing woody vegetation (i.e., typically < 5 m) that is characteristic of poorly fertile soils where environmental conditions prevent the natural formation of forests.
-This category is similar to the **EUNIS habitat type** ‘heathland, scrub and tundra’ (code **S**).
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure S1}. Geographical location of all networks in the EuPPollNet database coloured by bioregion."}
-
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude, Bioregion) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_point(data = for_plotting, alpha=0.75,size=1.2,pch=21,stroke = 0.25,aes(x= Longitude, y=Latitude, fill= Bioregion))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-     legend.key = element_blank(), 
-     legend.text = element_text(size = 6), 
-     legend.title = element_text(face ="bold",  size = 8),
-     legend.spacing.x=unit(0.5, "lines"), 
-     legend.key.width=unit(0.5, "lines"), 
-     legend.box.margin=margin(-10,-10,-10,-10),
-     axis.text = element_text(size=6,face="bold"),
-     axis.title = element_text(face="bold", size = 8),
-) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))  +
-ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(fill = guide_legend(override.aes = list(size = 1.5, byrow = TRUE))) +
-scale_fill_manual(values = c("purple3", 
-                               "dodgerblue3", 
-                               "cyan4", 
-                               "forestgreen",
-                               "orange2",
-                               "tomato3",
-                               "khaki2"))
-
-#Plot everything
-map 
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure S2}.
-Coverage of the EuPPollNet species for bees (family level), syrphids (subfamily level) and butterflies (family level) in relation to the total number of European species within these taxonomic groups.", fig.width=10, fig.height=4}
-
-#Prepare figure for supplementary material
-#Coverage across high taxonomic ranks of bees, syrphids and butterflies
-#--------------------------------------------------------------------------------------------#
-#Start from bees----
-#Bee families
-bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
-            "Andrenidae", "Colletidae","Melittidae")
-
-#Species per family within SafeNet
-bee_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-filter(Pollinator_family %in% bee_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-bee_spp_europe =  master_list %>% 
-select(Family, Genus, Species, GenusAndSpecies) %>% 
-filter(Family %in% bee_fam) %>% 
-group_by(Family) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-rename(Pollinator_family = Family) %>% 
-mutate(Group = "Europe")
-
-#Bind both datasets for plotting
-bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)
-
-p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-ylab("Number of species") +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Bee families")
-#--------------------------------------------------------------------------------------------#
-#Continue with syrphids----
-#As we don't have subfamilies, let's obtained them from the master list
-syrphids = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
-rename(Pollinator_accepted_name = GenusAndSpecies)
-#Syrphid species safenet
-syrphid_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_family== "Syrphidae") %>% 
-distinct()
-#Add subfamily info
-syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
-#Summarise spp per subfamily
-syrphid_spp_safenet = syrphid_spp_safenet %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Get spp per subfamily at european level
-syrphid_spp_europe = master_list %>% 
-filter(Family == "Syrphidae") %>% 
-select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
-group_by(Subfamily) %>% 
-summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
-#Plot
-# Plot for syrphid subfamily coverage with adjusted margins
-p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily, -Spp_number), Spp_number, fill = Group)) +
-  geom_bar(stat="identity", width = .5, position = "dodge", color = "black") +
-  coord_cartesian(expand = FALSE) +
-  xlab(NULL) +
-  ylab(NULL) +
-  theme_bw() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    panel.border = element_rect(size = 1),
-    plot.title = element_text(size = 14, face = "bold", margin = margin(b = 10)),
-    axis.title = element_text(face = "bold", size = 14),
-    axis.text = element_text(face = "bold", size = 12),
-    plot.margin = margin(10, 10, 20, 10)  # Adjusting the margins
-  ) +
-  scale_fill_manual(values = c("tan2", "lightgrey")) +
-  ggtitle("Syrphid subfamilies")
-#--------------------------------------------------------------------------------------------#
-#Continue with butterflies----
-#Do it at family level
-#Create vector with butterfly families
-butt_fam = c(unique(master_list1$family))
-#Species per family within SafeNet
-butterfly_spp_safenet = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Lepidoptera") %>% 
-filter(Pollinator_family %in% butt_fam) %>% 
-group_by(Pollinator_family) %>% 
-summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
-mutate(Group = "EuPPollNet")
-#Species per family within Europe
-butterfly_spp_europe =  master_list1 %>% 
-select(family, genus, scientificName) %>% 
-group_by(family) %>% 
-summarise(Spp_number = n_distinct(scientificName)) %>% 
-rename(Pollinator_family = family) %>% 
-mutate(Group = "Europe")
-#Bind both datasets for plotting
-butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
-#Plot
-p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
-geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
-coord_cartesian(expand = FALSE) +
-xlab(NULL) +
-ylab(NULL) +
-theme_bw() +
-theme(axis.text.x = element_text(angle = 45, hjust =1),
-      panel.border = element_rect(size=1)) +
-scale_fill_manual(values = c("tan2", "lightgrey"))+
-ggtitle("Butterfly families")
-
-#--------------------------------------------------------------------------------------------#
-p1+p2+p3 + plot_layout(guides = "collect") & 
-theme(plot.title = element_text(size = 14, face = "bold"),
-      axis.title = element_text(face = "bold", size = 14),
-      axis.text = element_text(face = "bold", size = 12))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.cap= "\\textbf{Figure S3}.
-Barplot indicating the number of networks by habitat and bioclimatic region within the database.", fig.width=8, fig.height=8}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Check number of studies by habitat
-habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat_bioregion = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-#Check number of studies by bioregion
-bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat_bioregion = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-c = bind_rows(habitat, bioregion)
-
-ggplot(c, aes(Habitat_bioregion, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_bw()+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")),
-        space = "free", scales = "free_y") +
-ylab("Networks") +
-xlab(NULL) +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.x = element_line(color="black"),
-      strip.text.x = element_text(size=14, face = "bold"),
-  axis.title = element_text(face = "bold", size = 14),
-axis.text = element_text(size=10),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-panel.border = element_rect(colour = "black",
-        fill=NA, size=1))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="60%", fig.cap= "\\textbf{Figure S4}.
-Association between nestedness (NODFc) and the geometric mean of plant and pollinator species per network", fig.width=3, fig.height=3}
-
-metrics_by_network_spp_number = readRDS("../Data/Working_files/metrics_by_network_spp_number.rds")
-
-metrics_by_network_spp_number %>% 
-ggplot(aes(geometric_mean_spp, Normalised_nestedness)) +
-geom_point(pch = 21, fill = "azure3", size=1.5, stroke=0.25, alpha=1) +
-theme_bw() +
-xlab("Species") +
-ylab("Normalised nestedness (NODFc)") +
-theme(
-    axis.title = element_text(face = "bold", size = 10),
-    axis.text = element_text(size = 8))
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S5}.
-Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with the 'quasiswap_count' algorithm (100 null models for each network).
-The quasiswap_count algorithm maintains row and column sums with fixed connectance."}
-
-library(dplyr)
-library(ggplot2)
-
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-
-#Finally, find % of networks that differ statistically
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-d = d %>% filter(!is.na((z_score)))
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-#Calculate percentages
-z_percent = d1 %>% 
-group_by(infra_over_represented) %>% 
-summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-no_dif_z = z_percent %>% 
-filter(infra_over_represented == "No statistical difference")
-
-# Plot
- d1 %>% 
-  ggplot(aes(x = z_score)) +
-  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
-                 bins = 50, alpha = 0.5, color = "black", position = "stack") +
-  theme_bw() +
-  coord_cartesian(expand = FALSE) +
-  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-  ylab("Number of networks") +
-  xlab("Z scores (NODFc)")  +
- # xlim(-6, 9) +
- # ylim(0, 0.8) +
-  scale_fill_manual(name = "Observed against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
-xlim(-6,6) +
-theme(plot.title = element_text(size = 24),
-      text = element_text(size = 20),
-      axis.title  = element_text(face="bold"))
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S6}.
-Distribution of z-scores comparing the nestedness of empirical networks with their randomized counterparts with a probabilistic null model from Bascompte et al., 2003 (100 null models for each network).
-This is the less restrictive approach, as it does not constrain the realised species degree.
-It first calculates the probability for each cell, and then if a random number between 0 and 1 exceeds this probability, an observed interaction is assigned to the cell.
-This shows that when the realised degree is not constrained, the number of networks that are statistically more nested is notably higher, highlighting that the nested structure is likely a byproduct of the species degree distribution."}
-
-library(dplyr)
-library(ggplot2)
-
-#Read data
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_nullmodel2.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-
-#Finally, find % of networks that differ statistically
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-#Calculate percentages
-z_percent = d1 %>% 
-group_by(infra_over_represented) %>% 
-summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-# Plot
- d1 %>% 
-  ggplot(aes(x = z_score)) +
-  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
-                 bins = 50, alpha = 0.5, color = "black", position = "stack") +
-  theme_bw() +
-  coord_cartesian(expand = FALSE) +
-  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-  ylab("Number of networks") +
-  xlab("Z scores (NODFc)")  +
- # xlim(-6, 9) +
- # ylim(0, 0.8) +
-  scale_fill_manual(name = "Observed against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
-xlim(-6,6) +
-theme(plot.title = element_text(size = 24),
-      text = element_text(size = 20),
-      axis.title  = element_text(face="bold"))
-
-```

### 2024-10-03 Manuscript/Supponting_information.qmd
journal format style

--- 2024-10-02 Manuscript/EuPPollNet_Ms.qmd
@@
 ```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-library(dplyr) #For handling fata 
-library(sf) #For handling coordinates and plotting
-library(scales) #For plotting (decimals on axes)
-library(lubridate) #To operate with dates
-library(tidyr) #Reshape data (wide format)
-library(stringr) #Fix/modify strings
-library(forcats)
+
+library(dplyr)
+library(sf)
+library(ggplot2)
+library(giscoR)
 library(readr)
-library(phytools)
-#Load libraries for plotting
-library(giscoR) #Maps
-library(ggstar) #cool shapes as points
-library(ggplot2) #general plotting
-library(ggtree) #phylo trees
-library(ggpattern) #cool shapes in barplots
-library(patchwork) #plot organisation
-library(cowplot) #plot organisation
-library(viridis) #colours
-library(ggnewscale)
-library(ggtreeExtra)
-library(kableExtra)
-
-#In case of conflicts select the main dplyr functions
-library(conflicted)
-conflicts_prefer(dplyr::filter)
-conflicts_prefer(dplyr::select)
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))
-
-#Bee-syrphid Masterlist
-master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
-#Load butterfly masterlist
-master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
-                    header = T)
-#Prepare table with data structure (as it would be made available to others)
-safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")
-
-#Work around to get the three cols
-#Note that month 1 and day 1 indicates the presence of a missing value
-#At least in Europe :)
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Country number
-country_number = data %>% 
-select(Country) %>% 
-n_distinct()
-#Study number
-study_number = data %>% 
-select(Study_id) %>% 
-n_distinct()
-#Average study number
-studies_by_country = data %>% 
-group_by(Country) %>% 
-summarise(study_number = n_distinct(Study_id)) %>% 
-summarise(mean(study_number)) %>% 
-pull()
-#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
-#Sampling method
-Sampling_method = data %>% 
-group_by(Study_id) %>% 
-distinct(Sampling_method) %>%  
-ungroup() %>% 
-count(Sampling_method) %>% 
-mutate(percent_method = n/sum(n) *100) %>% 
-mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
-#1st most common sampling methodology
-most_common_method = Sampling_method %>% 
-filter(percent_method == max(percent_method)) 
-#Minimum year
-min_year = data %>% 
-select(Year) %>% 
-slice(which.min(Year)) %>% 
-pull()
-#Maximum year
-max_year = data %>% 
-select(Year) %>% 
-slice(which.max(Year)) %>% 
-pull()
-#Number of interactions
-interaction_number = nrow(data)
-#Check number of accepted pollinators
-pollinator_spp_number = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name)%>%
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-#Check number of plant species
-plant_spp_number = data %>% 
-select(Plant_rank, Plant_status, 
-Plant_matchtype, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name) %>% 
-n_distinct()
-#Extract percentage of studies conducted during a single year
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
-group_by(Study_id) %>% 
-summarise(Count = n_distinct(Year))
-#Filter studies with only one year
-one_year = dates %>% 
-filter(Count == 1) %>% 
-pull() %>% 
-length() / nrow(dates) * 100
-#Obtain date information
-dates = data %>% 
-select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
-mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))
-
-dates1 = dates %>% 
-group_by(Study_id, Network_id, Sampling_method) %>% 
-summarise(Min_date = min(Date), 
-          Max_date= max(Date), 
-          Sampling_days = n_distinct(Date)) %>% 
-mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
-mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
-       .before=Study_id) %>% 
-arrange(Study_number) %>% 
-mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
-TRUE ~ NA)) %>% 
-mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
-TRUE ~ NA))
-#Extract min and max number of sampling days
-min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
-max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
-#Extract average of sampling days across networks and studies
-average_sampling_days = dates1 %>% 
-ungroup() %>% 
-group_by(Study_id) %>% 
-summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
-ungroup() %>% 
-summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
-pull()
-
-#Check number of distinct networks
-n_networks = data %>% 
-mutate(Network_id = paste0(Network_id,"_", Year)) %>% 
-distinct(Network_id) %>% 
-nrow(.)
-
-#Check number of distinct sampling sites
-n_sites = data %>% 
-select(Latitude, Longitude) %>% 
-distinct() %>% 
-nrow(.)
-#Check number of distinct bioregions
-n_bioregion = data %>% 
-select(Bioregion) %>% 
-distinct()  %>% 
-nrow(.)
-#Check number of distinct habitats
-n_habitats = data %>% 
-select(EuPPollNet_habitat) %>% 
-distinct() %>% 
-nrow(.)
+library(patchwork)
 #Read author list 
@@
 
-#Interactions at species level
-species_level = data %>% 
-filter(Plant_rank == "SPECIES") %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Plant_status == "ACCEPTED") %>% 
-filter(Pollinator_status == "ACCEPTED") %>% 
-nrow(.)
-species_level_int = species_level / nrow(data) * 100
-#Read here plant and pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Calculate mean coverage and deviation
-coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
-mean_coverage = mean(coverage)
-sd_coverage = sd(coverage)
-
-#Main orders percentage
-o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
-main_orders = data %>% 
-filter(Pollinator_order %in% o)
-main_orders_percent = nrow(main_orders) / nrow(data) * 100
-
-#Prepare table with all info 
-#Create variables
-vars = colnames(safenet_database)
-#Create table
-d = tibble(Variable = vars)
-
-#Add descriptor
-my_tibble = d %>%
-  mutate(Description = case_when(
-    Variable == "Study_id" ~ "Identifier of the study",
-    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
-    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
-    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
-    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
-    Variable == "Bioregion" ~ "European biogiographical regions",
-    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
-    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
-    Variable == "Latitude" ~ "North-south position of the observed interaction in decimal degrees",
-        Variable == "Longitude" ~ "East-west position of the observed interaction in decimal degrees",
-    Variable == "Date" ~ "Year, month and day when the observation took place",
-    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
-    Variable == "Plant_original_name" ~ "Plant species name given by the authors",
-    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
-    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
-    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
-    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
-    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
-    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-    Variable == "Pollinator_original_name" ~ "Pollinator species name given by the authors",
-    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
-    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
-    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
-    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
-    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
-        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
-        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",
-
-    TRUE ~ "Other"
-  ))
-
-#Check number of species and percentage of interactions of poll. groups
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-order_info = data %>% 
-select(Pollinator_order, Pollinator_accepted_name) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
-          Species = n_distinct(Pollinator_accepted_name))
-
-Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
-select(Interactions) %>% pull()
-
-#Check percentage of studies with flower counts
-floral_counts_percent = data  %>%
-select(Study_id, Flower_data) %>% 
-distinct() %>% 
-group_by(Flower_data) %>% 
-summarise(Percentage = n()/ nrow(.)*100) %>% 
-filter(Flower_data == "Yes")
-
-#Check percentage of apis mellifera records
-all_records = data %>% nrow(.) 
-apis_records = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-apis_percent = apis_records/all_records * 100
-#Check average pecentage of interactions of apis by net
-a = data %>% 
-group_by(Network_id) %>% 
-summarise(n_interactions = length(Pollinator_accepted_name))
-b = data %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-summarise(n_interactions_apis = length(Pollinator_accepted_name))
-c = left_join(a,b)
-apis_average_network = c %>% 
-mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
-ungroup() %>% 
-summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
-pull()
-
-#Check number of species from Hymenop, Dipte, Lepidop and Coleop
-
-pollinator_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-pollinator_spp_main_orders = data %>% 
-select(Pollinator_order, Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-filter(Pollinator_order == "Hymenoptera"|
-       Pollinator_order == "Diptera"|
-       Pollinator_order == "Lepidoptera"|
-       Pollinator_order == "HymenoColeopteraptera") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-
-spp_percent_main_order = pollinator_spp_main_orders/pollinator_spp *100
-
-#Calculate percentage of studies that recorded the main 4 orders
-metadata = readRDS("../Data/3_Final_data/Metadata.rds")
-#Some studies are divided by year
-#Get an average of the pollinator percentages across years first
-#Fix NA's
-metadata_study = metadata %>%
-group_by(Study_id) %>% 
-mutate(Hymenoptera_percent = as.numeric(na_if(Hymenoptera_percent, "NA"))) %>%  
-mutate(Diptera_percent = as.numeric(na_if(Diptera_percent, "NA"))) %>%  
-mutate(Lepidoptera_percent = as.numeric(na_if(Lepidoptera_percent, "NA"))) %>%  
-mutate(Coleoptera_percent = as.numeric(na_if(Coleoptera_percent, "NA"))) 
-#Calculate average
-mean_study = metadata_study  %>% 
-summarise(Hymenoptera_percent = mean(Hymenoptera_percent, na.rm = T),
-          Diptera_percent = mean(Diptera_percent, na.rm = T),
-          Lepidoptera_percent = mean(Lepidoptera_percent, na.rm = T),
-          Coleoptera_percent = mean(Coleoptera_percent, na.rm = T)) %>% 
- mutate(across(c(Hymenoptera_percent, Diptera_percent, Lepidoptera_percent, Coleoptera_percent),
-                ~ ifelse(is.nan(.), NA, .))) 
-#Now calculate percentage of studies that record any of these 4 orders
-mean_percent_order = mean_study %>%
-  summarise(
-    Hymenoptera_percent_metaweb = sum(!is.na(Hymenoptera_percent)) / nrow(.) * 100,
-    Diptera_percent_metaweb = sum(!is.na(Diptera_percent)) / nrow(.) * 100,
-    Lepidoptera_percent_metaweb = sum(!is.na(Lepidoptera_percent)) / nrow(.) * 100,
-    Coleoptera_percent_metaweb = sum(!is.na(Coleoptera_percent)) / nrow(.) * 100)
-
-Hymenoptera_percent_metaweb = mean_percent_order$Hymenoptera_percent_metaweb
-Diptera_percent_metaweb = mean_percent_order$Diptera_percent_metaweb
-Lepidoptera_percent_metaweb = mean_percent_order$Lepidoptera_percent_metaweb
-Coleoptera_percent_metaweb = mean_percent_order$Coleoptera_percent_metaweb
-
-#Check percentages of hymenoptea groups
-hymenoptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
-          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
-          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))
-
-bee_percent = hymenoptera_percent_by_group$Bee_studies
-bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
-all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies
-
-#Check percentages of diptera groups
-diptera_percent_by_group = metadata %>% 
-ungroup() %>% 
-distinct(Study_id, Taxa_recorded) %>% 
-summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
-          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
-          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
-                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)
-
-syrphid_percent = diptera_percent_by_group$Syrphid_studies
-all_diptera_percent = diptera_percent_by_group$Diptera_studies
-other_diptera_percent = diptera_percent_by_group$other_diptera_studies
-
-```
-
-\vspace*{-14mm}
-
-`r authors`
-
-**Corresponding author**= barragansljose@gmail.com\
-*All authors excluding the first four and last are ordered alphabetically*
-
-\begingroup
-\small
-
-`r aff_list`
-
-\small
-\endgroup
-
-\pagebreak
-
-# Abstract
-\vspace*{-5mm}
-
-**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks.
-One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
-Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
-We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
-We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified.
-Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).
-
-**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from `r n_habitats` different natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions.
-All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).
-
-**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
-
-**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species.
-The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphid species at the European level.
-
-**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
-The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
-
-**KEYWORDS**\
-\begingroup
-\small
-Angiosperms, connectance, flowering plants, nestedness, plant-pollinator networks, pollinators, species interactions 
-\endgroup
-\pagebreak
-
-# 1 | INTRODUCTION
-
-Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity.
-The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021].
-Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species status, trends, and conservation [@heberling2021; @zattara2021].
-However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.
-
-The interactions between different plant and pollinator species within a community form complex networks.
-Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano2003] or nestedness [@bascompte2003].
-Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study.
-Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @garcia2024].
-For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020].
-Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness.
-This emphasizes the importance of providing data in its rawest possible form in datasets that will be utilized for synthesis and macro-ecological studies.
-
-Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018].
-While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023], along with occurrence data [@zattara2021], is setting a foundation for the conservation of its flora and their pollinators.
-However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016].
-Numerous works have studied plant-pollinator interactions in the last decades, generating thousands of plant-pollinator interaction networks worldwide.
-Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024].
-Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level.
-For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context.
-Assembling and curating the existing information on EU plant-pollinator networks will guide research efforts, conservation planning and will set a foundation for future global change research.
-
-Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on plant-pollinator interactions at the European level.
-The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe.
-These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses.
-In addition, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.
-
-EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks.
-For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where specialists species interact only with subsets of generalists species, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].
-
-Overall, EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study.
-In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time.
-This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions.
-We expect that EuPPollNet can be used to evaluate macro-ecological processes in plant-pollinator networks, guide conservation planning, and set a baseline for global change research.
-
-# 2 | METHODS
-
-## Data acquisition
-
-The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995].
-As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; <https://doi.org/10.3030/101003476>), first, data was directly requested from members of the Safeguard project in May 2022.
-Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024.
-These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database.
-While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019].
-This approach maximized the potential number of studies that could be incorporated in this database.
-The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'.
-To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).
-
-## Dataset description
-
-The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations).
-The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet.
-Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%).
-All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans.
-The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions.
-Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%).
-Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they account for a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%).
-Hymenoptera and Diptera contain the highest number of species, with each comprising approximately 1,000 species in the database.
-However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**).
-Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure 1}.
-(\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study.
-The sizes of these shapes are proportional to the respective species counts.
-For visualization purposes, we have selected only a single location per study.
-(\\textbf{b}) Number of studies by year in EuPPollNet.
-(\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}.
-The total number of species and interactions is indicated in parentheses at the bottom."}
-
-#Fix dates
-data = data %>%
-dplyr::mutate(Year = lubridate::year(Date), 
-                Month = lubridate::month(Date), 
-                Day = lubridate::day(Date))
-
-#Prepare coordinates
-for_plotting = data %>% 
-select(Study_id, Latitude, Longitude) %>% 
-distinct()
-#Get countries and transfrom coordinates
-countries <- gisco_get_countries(
-resolution = 20) %>%
-st_transform(3035) %>% 
-sf::st_as_sf(crs = 4326)
-
-#Number of plant and poll species per network
-#-----------------------#
-#Pollinators------
-#-----------------------#
-#Check unique number of pollinator species
-PollNumber_by_ID = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-select(Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
-#Merge now with spp counts by network dataset
-poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-poll_data$coords <- data.frame(
-Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
-#Set coordinates
-poll_data$coords_sf = st_as_sf(poll_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-poll_data1 =  poll_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#-----------------------#
-#Plants-----
-#-----------------------#
-#Check unique number of plant species
-PlantNumber_by_ID = data %>% 
-select(Plant_rank, Plant_accepted_name, Study_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-select(Plant_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-summarise(Plant_counts = n_distinct(Plant_accepted_name))
-#Merge now with spp counts by network dataset
-plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
-#Create duplicate cord columns to operate on them
-plant_data$coords <- data.frame(
-Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
-#Set coordinates
-plant_data$coords_sf = st_as_sf(plant_data$coords, 
-  coords = c("Longitude", "Latitude"), crs = 4326)
-#Plot map of Europe
-#Very noisy with many coordinates per point,
-#Let's select a unique coordinate by study
-plant_data1 =  plant_data %>% 
-distinct(Study_id, .keep_all = TRUE)
-
-#Add vars
-plant_data1$Group = "Plant"
-poll_data1$Group = "Pollinator"
-#Rename cols
-plant_data2 = plant_data1 %>% 
-rename(Counts = Plant_counts)
-poll_data2 = poll_data1 %>% 
-rename(Counts = Pollinator_counts)
-#Bind rows
-all_data = bind_rows(plant_data2, poll_data2)
-
-#Try to plot both together (same plot)
-map = ggplot(countries) +
-geom_sf(fill="floralwhite",color = "black", 
-        alpha = 1,linewidth=0.1) + 
-geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
-           size=Counts, starshape = Group, color= Group))+  
-coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
-         xlim = c(-14, 39), ylim = c(37, 70))+
-theme(panel.grid.major = element_blank(), 
-      legend.position = ("top"),
-      legend.box = "vertical",
-      legend.background = element_blank(),
-      legend.box.background = element_blank(),
-      legend.key = element_blank(), 
-      legend.title = element_text(face ="bold",  size = 6),
-      legend.box.spacing = unit(0.2, "cm"),
-      legend.margin=margin(0,0,0,0),
-        legend.box.margin=margin(-10,-10,-10,-10),
-      axis.text = element_text(size=6,                  face="bold"),
-      axis.title = element_text(face="bold", size = 8),
-         legend.text = element_text(size = 6),                           legend.direction="horizontal",
-         legend.spacing.y = unit(-0.3, "cm"),
-      legend.title.align=0.5) +
-theme(panel.background = element_rect(fill = "aliceblue"),
-        panel.border = element_rect(colour = "black",
-        fill=NA, size=1)) + 
- scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
-scale_color_manual(values=c("#009E73", "#D55E00"),
-         guide="none") +
-scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
-             guide=guide_legend(reverse=TRUE,
-             override.aes=list(
-                 size=2,
-                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
-  ylab("Latitude") +
-xlab("Longitude") +
-scale_x_continuous(breaks = seq(0, 20, by = 10)) +
-scale_y_continuous(breaks = seq(40, 60, by = 10)) +
-guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
-ggtitle("(a)") +
-theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))
-
-#----------------#
-#Pollinators
-#----------------#
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
-group_by(Study_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_studies_shared = n) %>% 
-arrange(-n_studies_shared) %>% 
-mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-data2 = data %>%  select(Study_id, Year) %>% 
-distinct() %>% 
-group_by(Year) %>% 
-summarise(Counts = n())
-
-p2  = ggplot(data2, aes(Year, Counts)) +
-geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
-  axis.title.x = element_text(face = "bold", size = 8),
-  axis.text.y = element_text(size = 6),
-  axis.text.x = element_text(size = 6,  angle = 90), 
-  strip.text.x = element_text(size=12),
-   axis.ticks.x = element_line(color="black"),
-  strip.background = element_rect(
-  color="black", fill=NA, size=0.5, linetype="solid")) + 
-ylab("Number of studies") +
-coord_cartesian(expand = F, clip = "off") + 
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  panel.background = element_blank(),
-  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black')) +
-scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
-scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
-ggtitle("(b)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -5))
-
-main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
-
-#Calculate spp number and interactions separately
-data1_spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.))
-
-data1_int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-group_by(Pollinator_order) %>% 
-summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.))
-
-data1 = left_join(data1_spp, data1_int)
-
-#Select main orders
-spp = data %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Species = n_distinct(Pollinator_accepted_name)) %>% 
-pull()
-#Select main orders
-int = data %>% 
-filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
-select(Pollinator_order, Pollinator_accepted_name, Pollinator_rank) %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-#filter(Pollinator_rank == "SPECIES") %>% 
-summarise(Interactions = length(Pollinator_accepted_name)) %>% 
-pull()
-
-#Combine Interactions and Species into a single column
-data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")
-
-#Set order of levels for plotting
-data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))
-
-#Plot
-p3 = ggplot(data_long, aes(x = Variable, y = Value, fill = Pollinator_order)) + 
-geom_bar(stat = "identity", width = 0.7) +
-theme_classic() +
-scale_fill_viridis(discrete=TRUE, direction = -1, name="Pollinator\n order") +
-xlab(NULL) +
-ylab("Percentage") + 
-scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
-theme(axis.text = element_text(size=6), 
-     axis.title.y = element_text(face = "bold", size = 7, vjust = -2),  
-     axis.title.x = element_text(face="bold", size = 7),
-     legend.position = "top",
-     legend.title = element_text(size = 5, face = "bold"),
-     legend.text = element_text(size = 5),
-     legend.key.size = unit(3, "mm"),
-     legend.margin=margin(0,2,0,0),
-     legend.box.margin=margin(-9,-9,-6,-6)) +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))+
-ggtitle("(c)") +
-theme(plot.title = element_text(size = 10,hjust = -0.4, vjust = -6)) +
-scale_x_discrete(labels = c("Species\n (2,172)", "Interactions\n (358,049)"))
-
-#plot2 = p3 / p1 / p2 
-plot2 = p2 / p3
-
-#Plot everything
-map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.485, 0.8))
-
-```
-
-\newpage
-
-## Data structure
-
-The EuPPollNet database is available in both .csv and .rds formats and contains a total of `r nrow(my_tibble)` columns (**Table 1**), where each row represents a single interaction between a plant and a pollinator species.
-These columns include information about the study and network identifiers (columns 1 and 2), sampling method (3), habitat type as described by the author, and a unified habitat classification across studies (4 and 5), bioregion where the network is located (6), country, locality, and latitude-longitude coordinates (7 to 10), date of the interaction (11), number of interactions (12), taxonomic information about plants (13 to 20), taxonomic information about pollinators (21 to 28), and information about the availability of floral count data (29).
-The flower count data is provided in a separate file (.csv or .rds) and can be merged with the interaction data through the "Flower_data_merger" column (30).
-Note that although two-thirds of studies include information on floral abundance, the methods and units vary greatly across studies.
-To construct a plant-pollinator network matrix within a single flowering season at the site level, users should group interactions by plant and pollinator species, site, study, and year.
-Finally, metadata at the study level is provided in a separate file, including information about the authors, digital object identifier (if available), sampling time, and taxonomic coverage of the main pollinator groups for each study.
-
-\newpage
-
-```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.height=1, fig.width=10}
-
-#Small edit to column number
-my_tibble = my_tibble %>% 
-mutate(Variable = paste0(seq(1:length(Variable)),". ", Variable))
-
-kable(my_tibble, booktabs = T, linesep = "\\addlinespace", caption = "\\textbf{Table 1.} Column names and their descriptions within the EuPPollNet database.") %>%
-  kable_styling(full_width = F,font_size = 9) %>% row_spec(0, bold=T) %>% 
-kable_styling(position = "center")
-
-```
-
-## Taxonomic harmonization
-
-All plant and pollinator species names were checked and harmonized in R using **rgbif** [@chamberlain2022].
-The protocol for plants and pollinators is similar but slightly different given the availability of the different taxonomic resources.
-For transparency, we have included in the database the original species name, the new assigned name, and, if the name of the species is uncertain (e.g., species complex or species alike).
-In addition, taxonomic information at genus, family and order level was downloaded for each species.
-
-For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and manually fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.
-
-For pollinators: (i) we first created a checklist of species names for the most representative pollinator groups at the European level by combining the most up to date published checklists of bees and syrphids [@reverte2023], and butterflies [@wiemers2018]; (ii) we compared pollinator species names against the checklist and recovered some unmatched cases with restrictive fuzzy matching by using **stringdist** package [@van2014]; (iii) we fixed unmatched records when necessary and retrieved the taxonomic information for all species from GBIF; (iv) we fixed the non-found cases in the GBIF checklist and made sure that all species names from bees, syrphids and butterflies were named according to their respective species checklists.
-Coleoptera species names were only checked against the GBIF checklist.
-
-## Taxonomic coverage
-
-To assess the completeness of plant and pollinator species in the EuPPollNet database at the European level, we used the aforementioned checklists for plants and pollinators.
-Specifically for plants, we refined the checklist to include only plants occurring in Europe and excluded taxonomic groups not associated with biotic pollination.
-We did this by first excluding the families considered to have exclusively a wind pollination mode [see @culley2002], and then by filtering out the genera with wind or non-biotic pollination from families that exhibit both biotic and non-biotic pollination modes.
-Additionally, we manually included exotic species and added unresolved species names that were not present in the accepted names of the checklist at the current version of usage.
-For pollinators, we only evaluated the taxonomic coverage of groups with species checklists available in Europe (i.e., bees, syrphids and butterflies).
-While there is not a good understanding of pollinator diversity in other taxonomic groups (e.g., beetles), it was assumed that their coverage within the database is equal to the average coverage of bees, syrphids, and butterflies (mean coverage = `r round(mean_coverage, 2)`; sd = `r round(sd_coverage,2)`).
-Therefore, the total number of of flower visiting species from other taxonomic groups (i.e., non-bee, non syrphid and non-butterfly flower visitors) at European level was extrapolated by assuming that their coverage is equal to the mean coverage of bees, syrphids, and butterflies.
-Consequently, we provide an estimate for the total number of potential pollinators across the European continent.
-
-Finally, to evaluate if the presence-absence of interaction records for bees and flowering plants follows a phylogenetic pattern within the database, we calculated its phylogenetic signal  at genus and family level, respectively.
-The phylogenetic signal was calculated by using the _phylosig_ function from the **phytools** package [@revell2012].
-We extracted the phylogenetic information for bees from a genus level phylogeny [@hedtke2013] and processed it using the packages **ape** [@paradis2019], **MCMCglmm** [@hadfield2010] and **phytools**.
-For plants, the phylogenetic tree was obtained from a species level plant phylogeny [@smith2018] with the help of the **rtree** package [@li2023].
-
-## Sampling coverage
-
-The completeness of the EuPPollNet database was evaluated by exploring the rarefied accumulation curves of plant and pollinator species and their interactions across the different networks.
-In addition, we computed the  accumulation curve of pollinator species with an increasing number of plant species as an indicator for how many pollinator species are likely responsible for the pollination of flowering plants [e.g., @kleijn2015 for crops].
-The rarefied and extrapolated sampling curves were obtained using the **iNEXT** package [@hsieh2016].
-The different rarefied curves were complemented with 100 bootstrapped accumulation curves.
-
-## Habitat type and bioclimatic region
-
-We describe the habitat type for each site using information from Corine Land Cover (CLC, version 2018) extracted using the Terra package [@hijmans2022], visual inspection of Google Earth imagery and the habitat classification from the authors.
-These different habitat categories (see definitions in supplementary text) allow a quick comparison and understanding of the habitat types from the database.
-Moreover, Europe is characterized by a great variety of environmental conditions that harbor different biota.
-Thus, to allow authors to explore the set of studies that share similar environmental conditions and species, we assigned a biogeographical region to each site.
-The biogeographical regions were downloaded from the European Environment Agency (version 2016) and were matched to the different sites using a spatial joint from the **sf** package [@pebesma2018].
-
-## Network analyses
-
-To provide a general overview of the structure of plant-pollinator networks in EuPPollNet, we quantified connectance and nestedness for each network and examined how these network metrics change across different latitudes and bioclimatic regions in Europe.
-We selected these two network metrics as they are commonly evaluated in plant-pollinator network studies and capture structural properties with a straightforward interpretation.
-We implemented “standardised” versions of connectance and nestedness to account for the effect of sampling effort on network metrics.
-As connectance is negatively associated with network size [@jordano1987], we evaluated how network connectance was associated with the number of species (i.e., log of geometric mean of plants and pollinators) and extracted the residuals from this association (i.e., residual connectance) as a measurement of corrected connectance.
-The relationship between residual connectance and species richness was investigated using a beta regression.
-This approach was chosen because connectance displays a non-normal distribution with continuous values bounded between 0 and 1.
-The model was implemented using the betareg package [@cribari2010].
-We used NODFc to compare nestedness across networks, as it corrects by connectance and the number of species in comparison to other nestedness metrics that change with network size [@song2017].
-This metric was calculated using the **maxnodf** package [@hoeppke2021].
-Both residual connectance and NODFc were used as dependent variables to evaluate their association with latitude.
-In addition, to quantify how connectance and nestedness change with network size, we determined their association with the number of species per network using the Kendall rank correlation coefficient.
-
-Finally, to compare if networks are more or less nested than expected by chance, we employed the traditional z-score approach with the widely used nestedness metric (NODF) from @almeida2008.
-The z-score approach only compares each unique network against their randomized versions, avoiding the influence of network size.
-As NODF is a metric computed from binary matrices, we calculated 100 null models for the binarized version of each network using the 'curveball' algorithm [@strona2014].
-This algorithm implements the configuration model (i.e.
-random rewiring of all links, without self-links or double links) and thus keeps the exact number of connections per species (i.e., realised degree).
-Since null model selection can significantly influence statistical results [@kaiser2015], we also implemented a null model that reorganizes the quantitative networks before binarising them.
-To that end, we used the 'quasiswap_count' algorithm, which is a non-sequential algorithm for quantitative networks that maintains constant connectance and the number of connections per species.
-Each empirical network was randomized 100 times using each method.
-These null models were implemented with the **vegan** [@oksanen2013] package.
-Both connectance and nestedness (NODF) were estimated for each network using the function _networklevel_ from **bipartite** [@dormann2008].
-
-\newpage
-
-# 3 | RESULTS
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-#Taxonomic coverage outputs
-#Find number of plant and pollinator species
-##Pollinators
-poll_spp = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-nrow(.)
-##Plants
-plant_spp = data %>% 
-select(Plant_rank, Plant_accepted_name) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-nrow(.)
-
-#Overall plant coverage
-plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")
-
-#Average plant family coverage
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData") %>% 
-filter(fam_group == "B")
-#Average coverage per family
-mean_plant_fam_cov = d1 %>% 
-ungroup() %>% 
-summarise(mean_plant_fam_cov = mean(Percent)) %>% 
-pull()
-
-#Percent of families with interaction records
-all_fam = nrow(d1)
-fam_with_data = d1 %>% 
-filter(Percent > 0) %>% 
-nrow()
-percent_plant_fam_with_data = fam_with_data/all_fam * 100
-
-#Calculate phylogenetic signal of presence-absence information
-#Read data and plot it
-tree = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-#Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-#select(label, categorical) %>% 
-ungroup() 
-#Check phylogenetic signal
-d1 = d1 %>% arrange(factor(label, levels = tree@phylo$tip.label))
-
-v = phylo_trait$categorical
-names(v) = tree@phylo$tip.label
-
-phylo_signal_plant = phylosig(tree@phylo, v, method="lambda", test=TRUE)
-lambda_phylosig_plants = phylo_signal_plant$lambda
-p_phylosig_plants = phylo_signal_plant$P
-
-#Load pollinator coverage
-bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
-syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
-lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
-#Check percentage of interactions from bees with and without apis
-#Check percentage of bees in the database
-bee_fam = c("Apidae", "Megachilidae", "Halictidae",
-            "Andrenidae", "Colletidae", "Melittidae")
-all_rows = data %>% nrow(.)
-#
-bee_rows = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-nrow(.)
-#Bee interactions
-bee_interactions_percent = bee_rows/all_rows *100
-#Eclude now apis
-bee_rows_non_apis = data %>%
-filter(Pollinator_family %in% bee_fam) %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>% 
-nrow(.)
-#
-bee_interactions_non_apis_percent = bee_rows_non_apis/bee_rows *100
-
-#Coverage across bee genera
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-all_rows_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-nrow(.)
-#
-data_bees = d1 %>%  
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-filter(Percent > 0) %>% 
-nrow(.)
-#
-bee_genera_percent_coverage = data_bees/all_rows_bees * 100
-#Check average coverage across bee genera
-average_coverage_bee_genera = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-ungroup() %>% 
-summarise(average_coverage_bee_genera = mean(Percent)) %>% 
-pull()
-
-#Calculate phylogenetic signal of presence-absence information
-##Read data and plot it
-tree = readRDS("../Data/Manuscript_info/bee.tree100.rds")
-##Create binary trait to calculate phylo
-phylo_trait = d1 %>% 
-filter(fam_group == "a") %>% 
-mutate(Percent = case_when(is.na(Percent) ~ 0, TRUE ~ Percent)) %>% 
-mutate(categorical = case_when(Percent > 0 ~ 1, TRUE ~ 0)) %>% 
-ungroup() 
-##Conduct left join to tibble of tip labels to organise order
-v = tibble(Genus1 = tree$tip.label) %>% 
-mutate(Genus1 = str_to_title(word(Genus1,1,sep = "_")))
-#
-v1 = left_join(v, phylo_trait)
-#
-v2 = v1$categorical
-names(v2) = tree$tip.label
-#
-phylo_signal_bees = phylosig(tree, v2, method="lambda", test=TRUE)
-lambda_phylosig_bees = phylo_signal_bees$lambda
-p_phylosig_bees = phylo_signal_bees$P
-
-#Check number of studies by habitat
-studies_by_habitat = data %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of sampling sites by habitat
-networks_by_habitat = data %>% 
-select(Network_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) %>% 
-slice_max(Habitat_count, n = 3) %>% 
-mutate(EuPPollNet_habitat = tolower(EuPPollNet_habitat))
-#Check number of studies by bioregion
-studies_by_bioregion = data %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check number of sampling sites by bioregion
-networks_by_bioregion = data %>% 
-select(Network_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) %>% 
-slice_max(Bioregion_count, n = 3) %>% 
-mutate(Bioregion = tolower(Bioregion))
-#Check percentage of Apis mellifera across networks
-apis_percent_networks = data %>% 
-filter(Pollinator_accepted_name == "Apis mellifera") %>%
-group_by(Study_id) %>% 
-mutate(Apis = "Yes") %>% 
-distinct(Study_id, Apis) %>% 
-nrow() / nlevels(factor(data$Study_id)) *100
-
-#Load number of species from checklists and extrapolated one
-bee_potential_spp = readRDS("../Data/Manuscript_info/bee_potential_spp.RData")
-syrphid_potential_spp = readRDS("../Data/Manuscript_info/syrphid_potential_spp.RData")
-lepidoptera_potential_spp = readRDS("../Data/Manuscript_info/lepidoptera_potential_spp.RData")
-extrapolated_other_spp = readRDS("../Data/Manuscript_info/other_taxa_potential_spp.RData")
-extrapolated_other_spp = round(extrapolated_other_spp)
-```
-
-## Taxonomic coverage
-
-Europe hosts approximately over 5,000 species of pollinators, including `r prettyNum(bee_potential_spp, big.mark=",")` bee species `r prettyNum(syrphid_potential_spp, big.mark=",")` syrphid species, `r prettyNum(lepidoptera_potential_spp, big.mark=",")` butterflies species, and about 1,700 species from other taxonomic groups, as estimated based on EuPPollNet average taxonomic coverage.
-Additionally, there are around 25,000 species of plants that benefit from animal pollination when excluding non-biotic pollinated species.
-EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species.
-The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids).
-Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees.
-Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**).
-The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`).
-The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level.
-Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}.
-Phylogenetic and taxonomic coverage of bee genera at European level.
-The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple.
-Additionally, the coverage of species recorded in EuPPollNet for each genus is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database, out of the total number of bee species in Europe.
-Dark grey bars represent the total number of species per genus at European level."}
-
-#Read data
-tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
-n = readRDS("../Data/Manuscript_info/node_data.RData")
-
-#Circular layour
-ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = 0.065, size=2.9, fontface=1) +
-scale_size(range = c(0,4.5), guide = 'none')  +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-guides(fill=guide_legend(), size = guide_legend()) +
-theme(legend.position = "bottom") +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
-                    pwidth=0.13, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0.148,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = 1.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-guides(fill = "none")+
-scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
-geom_fruit(geom=geom_bar,
-                    mapping=aes(y=label, x=Spp_number_Europe),
-                    pwidth=0.12, 
-                    orientation="y", 
-                    stat="identity",
-                    color="black",
-                    offset = 0,
-            axis.params=list(
-                         axis = "x",
-                         text.size  = 2.5,
-                         hjust  = -0.2,
-                         vjust = 0.5,
-                         nbreak = 3,
-                         fontface = 2
-                     ),
-         grid.params=list()) +
-annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
-annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
-annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
-new_scale_fill() +
-geom_point2(aes(subset=node==131), color='orange', size=3) +
-geom_point2(aes(subset=node==84), color='orange', size=3) +
-geom_point2(aes(subset=node==73), color='orange', size=3) +
-geom_point2(aes(subset=node==74), color='orange', size=3) +
-geom_point2(aes(subset=node==111), color='orange', size=3)+
-geom_point2(aes(subset=node==91), color='orange', size=3) +
-annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
-annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
-annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
-annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
-annotate("text", x = 2.25, y = 26.8, label = "Megachilidae", size= 3)+
-annotate("text", x = 2.5, y = 51, label = "Apidae", size= 3) +
-coord_cartesian(clip = "off")+
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.6,0),
-      legend.direction = "horizontal")
-
-```
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.
-Phylogenetic and taxonomic coverage of the plant families at European level.
-The number of interactions recorded per family in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple.
-Additionally, the coverage of species recorded in EuPPollNet for each family is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database out of the total number of bee species per family in Europe.
-Dark grey bars represent the total number of species per family at European level.", fig.width=8, fig.height=8}
-
-library(ggplot2)
-library(ggtree)
-library(ggnewscale)
-library(ggtreeExtra)
-library(cowplot)
-
-#Save data and plot it in an Rmarkdown
-tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
-d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")
-
-#Check if there are differences in labels and order
-ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
-geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
-geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
-scale_size(range = c(0,4), guide = 'none') +
-scale_fill_continuous(type = "viridis", direction=-1) + 
-theme(legend.position = "bottom") +
-guides(fill=guide_legend(), size = guide_legend()) +
-new_scale_fill() +
-geom_fruit(data = d1, geom=geom_bar,
-   mapping=aes(y=label, x=Percent, fill = fam_group),
-   pwidth=0.1, 
-   orientation="y", 
-   stat="identity",
-   color="black",
-   offset = 0.028,
-   linewidth=0.35,
-   axis.params=list(
-     axis = "x",
-     text.size  = 1.5,
-     hjust  = 1,
-     vjust = 0.5,
-     nbreak = 3,
-     fontface = 2),
-  grid.params=list())+
-guides(fill = "none")+
-scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
-geom_fruit(geom=geom_bar,
-    mapping=aes(y=label, x=log(Spp_Europe +1)),
-    pwidth=0.1, 
-    orientation="y", 
-    stat="identity",
-    color="black",
-    offset = 0,
-    linewidth=0.35) +
-theme(legend.margin=unit(1,"cm"),
-      legend.spacing = unit(1, "cm"),
-      plot.margin = margin(4,0,4,0), 
-      legend.position = c(0.5,0.09),
-      legend.direction = "horizontal")
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Sampling coverage outputs
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Coverage
-poll_sampling_coverage = poll_output$DataInfo$SC *100
-plant_sampling_coverage = plant_output$DataInfo$SC *100
-int_coverage = int_output$DataInfo$SC *100
-plant_poll_coverage = plant_poll_output$DataInfo$SC *100
-
-#Increase doubling effort
-##Pollinators
-poll_increase = poll_output$iNextEst$coverage_based
-poll_increase1 = poll_increase %>% filter(Method == "Observed") %>% pull(qD)
-poll_increase2 = poll_increase %>% filter(t == max(t)) %>% pull(qD)
-poll_increase3 = (poll_increase2-poll_increase1)/poll_increase1 * 100
-##Plants
-plant_increase = plant_output$iNextEst$coverage_based
-plant_increase1 = plant_increase %>% filter(Method == "Observed") %>% pull(qD)
-plant_increase2 = plant_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_increase3 = (plant_increase2-plant_increase1)/plant_increase1 * 100
-##Interactions
-int_increase = int_output$iNextEst$coverage_based
-int_increase1 = int_increase %>% filter(Method == "Observed") %>% pull(qD)
-int_increase2 = int_increase %>% filter(t == max(t)) %>% pull(qD)
-int_increase3 = (int_increase2-int_increase1)/int_increase1 * 100
-##Pollinator by plant
-plant_poll_increase = plant_poll_output$iNextEst$coverage_based
-plant_poll_increase1 = plant_poll_increase %>% filter(Method == "Observed") %>% pull(qD)
-plant_poll_increase2 = plant_poll_increase %>% filter(t == max(t)) %>% pull(qD)
-plant_poll_increase3 = (plant_poll_increase2-plant_poll_increase1)/plant_poll_increase1 * 100
-
-#Prepare results from common and rare poll across sampling sites
-#Pollinators
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-poll_total = nrow(poll_spread)
-polls_under_one_percent = poll_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-polls_under_one_percent = polls_under_one_percent/poll_total * 100
-#Find most common pollinator
-most_common_poll = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Pollinator_accepted_name)
-#Find most common poll percent
-most_common_poll_percent = poll_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-#Plants
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id)*100)
-#Check number of studies under 1%
-plant_total = nrow(plant_spread)
-plants_under_one_percent = plant_spread %>% 
-filter(Percent_total<1) %>% 
-nrow()
-plants_under_one_percent = plants_under_one_percent/plant_total * 100  
-#Find most common plant
-most_common_plant = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Plant_accepted_name)
-#Find most common plant percent
-most_common_plant_percent = plant_spread %>% 
-filter(Percent_total == max(Percent_total)) %>% 
-pull(Percent_total)
-```
-
-## Sampling coverage
-
-The estimated sampling coverage of plant and pollinator species within EuPPollNet across the different networks is approximately `r round(poll_sampling_coverage)`% for both taxonomic groups.
-This suggests that the rarefied accumulation curves of both plant and pollinator species exhibit already a "quasi-asymptotic" growth of species richness by considering the current number of networks (**Figure 4a-4b**).
-The predicted observed species richness by doubling the sampling effort on the already sampled habitat types within the database will only increase pollinator richness by `r format(round(poll_increase3, 2), nsmall=2)`% and plant richness by `r format(round(plant_increase3, 2), nsmall=2)`%.
-However, the sampling coverage of interactions is `r format(round(int_coverage, 2), nsmall=2)`%, and by doubling the sampling effort, the predicted number of unique interactions recorded will have approximately a twofold increase (`r format(round(int_increase3, 2), nsmall=2)`%; **Figure 4c**).
-When we consider the accumulated pollinator richness across sampled plant species, this curve also shows a “quasi-asymptotic” growth with a sampling coverage value of `r format(round(plant_poll_coverage, 2), nsmall=2)`%.
-The predicted recorded pollinator species by doubling the number of plants sampled is expected to increase by `r format(round(plant_poll_increase3, 2), nsmall=2)`% (**Figure 4d**).
-We find that a small portion of plant species and pollinator species are shared across a broad range of networks and that most plant (`r format(round(plants_under_one_percent, 2), nsmall=2)`%) and pollinator (`r format(round(polls_under_one_percent, 2), nsmall=2)`%) species are exclusively found in less than 1% of networks (**Figures 4e-4f**).
-The most common plant (*`r most_common_plant`*) and pollinator (*`r most_common_poll`* when excluding *Apis mellifera*) species are found in `r format(round(most_common_plant_percent, 2), nsmall=2)`% and `r format(round(most_common_poll_percent, 2), nsmall=2)`% of networks, respectively.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 4}.
-Graphs (a-b-c) indicate the accumulation curves for pollinators, plants, and the number of unique pairwise interactions across networks.
-Grey solid lines represent 100 randomized accumulation curves, the black solid lines represent the interpolated curve (i.e., the mean across curves), and the red dashed lines illustrate the extrapolated curve for approximately 3000 networks.
-The solid black points indicate the number of species and interactions contained in the database.
-Graph (d) shows the accumulation curve of pollinator species across an increasing number of plant species.
-This last graph uses the same color and shape structure as the ones in the top panel.
-Graphs (e-f) indicate the percentage of occurrence (i.e., incidence) of plant and pollinator species across networks.
-Species on the left (i.e., common) are found in many networks, while species on the right (i.e., rare) are found in few or only a single network.
-Note that \\textit{Apis mellifera} is the most common pollinator but was excluded from this visualization.", fig.width=10, fig.height=6, out.width="115%"}
-
-#Pollinators-----
-poll_output = readRDS("../Data/Working_files/pollinator_sampling_coverage.rds")
-pollinators_curves = readRDS("../Data/Working_files/collectors_curves_pollinators.rds")
-#Create dataset
-d_poll = poll_output$iNextEst$coverage_based
-#Check cols
-d_poll$col = d_poll$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = pollinators_curves$Study_sampled, 
-           Method = as.character(pollinators_curves$iteration), 
-           qD = pollinators_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_poll1 = bind_rows(d_poll, a_poll)
-d_poll1_point = d_poll1 %>%  filter(Method == "Observed")
-d_poll1_lines = d_poll1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_poll1_lines = d_poll1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_poll1_lines$col = factor(d_poll1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-p1 = ggplot(d_poll1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col),size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_poll1_point, col = "black", size=2) + 
-ylab("Pollinator species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000) +
-ggtitle("(a)") +
-theme(plot.margin = margin(r = 12))
-
-#Plants-----
-plant_output = readRDS("../Data/Working_files/plant_sampling_coverage.rds")
-plant_curves = readRDS("../Data/Working_files/collectors_curves_plants.rds")
-#Create dataset
-d_plant = plant_output$iNextEst$coverage_based
-#Check cols
-d_plant$col = d_plant$Method
-#Now create tibble with same structure
-a_plant = tibble(t = plant_curves$Study_sampled, 
-           Method = as.character(plant_curves$iteration), 
-           qD = plant_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_plant1 = bind_rows(d_plant, a_plant)
-d_plant1_point = d_plant1 %>%  filter(Method == "Observed")
-d_plant1_lines = d_plant1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_plant1_lines = d_plant1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_plant1_lines$col = factor(d_plant1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-
-#Plot
-p2 = ggplot(d_plant1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_plant1_point, col = "black", size=2) + 
-ylab("Plant species") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000)+
-ylim(0, 3000)+
-ggtitle("(b)")+
-theme(plot.margin = margin(r = 12))
-
-#Interactions-----
-int_output = readRDS("../Data/Working_files/interactions_sampling_coverage.rds")
-interactions_curves = readRDS("../Data/Working_files/collectors_curves_interactions.rds")
-#Create dataset
-d_int = int_output$iNextEst$coverage_based
-#Check cols
-d_int$col = d_int$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = interactions_curves$Study_sampled, 
-           Method = as.character(interactions_curves$iteration), 
-           qD = interactions_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_int1 = bind_rows(d_int, a_poll)
-d_int1_point = d_int1 %>%  filter(Method == "Observed")
-d_int1_lines = d_int1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_int1_lines = d_int1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_int1_lines$col = factor(d_int1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-p3 = ggplot(d_int1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col), size = 0.65) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_int1_point, col = "black", size=2) + 
-ylab("Unique interactions") +
-xlab("Networks") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off")  +
-xlim(0,3000) +
-ylim(0, 35000)+
-ggtitle("(c)")
-
-#----------------#
-#Pollinators
-#----------------#
-#Read data
-data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
-mutate(Network_id = paste0(Study_id, Network_id))  
-#First select cols of rank, accepted name and study ID
-#Then select unique levels by study ID
-#Now sum levels (maximum can be as maximum number of studies)
-poll_spread = data %>% 
-select(Pollinator_rank, Pollinator_accepted_name, Network_id) %>% 
-group_by(Network_id) %>% 
-filter(Pollinator_rank == "SPECIES") %>% 
-distinct(Pollinator_accepted_name) %>% 
-ungroup() %>% 
-count(Pollinator_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = poll_spread %>%  
-select(Pollinator_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common pollinator for plotting
-most_common_poll = poll_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Pollinator_accepted_name) %>% pull()
-
-#Plot
-o1 = ggplot(poll_spread, 
-aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
-geom_bar(stat = "identity", 
-           color = "#D55E00",
-           lwd = 0.03, fill = "#D55E00") +
-ylab("Shared pollinators (%)") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-  axis.ticks.x=element_blank(),
-  axis.ticks.y = element_line(size = 0.2),
-  panel.grid.major.x = element_blank(),
-  panel.grid.minor.y = element_blank(),
-  panel.grid.major.y = element_line(size=0.2, linetype = 2)) +
-xlab("Pollinator species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-color = "black", size=0.35) + 
-annotate(geom="text", x=1590, y=0.60, label="50th percentile",color="black", size = 3) +
-annotate(geom="text", x=140, y=0.67, label=most_common_poll[1],color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=200, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1800, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3)  +
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0)) + 
-coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),
-                ylim=c(0,0.7), clip="off") +
-ggtitle("(e)")
-
-#----------------#
-#Plants
-#----------------#
-plant_spread = data %>% 
-select(Plant_rank, Network_id, Plant_accepted_name) %>% 
-group_by(Network_id) %>% 
-filter(Plant_rank == "SPECIES") %>% 
-distinct(Plant_accepted_name) %>% 
-ungroup() %>% 
-count(Plant_accepted_name) %>% 
-rename(n_networks_shared = n) %>% 
-arrange(-n_networks_shared) %>% 
-mutate(Percent_total = n_networks_shared / n_distinct(data$Network_id))
-
-#Seelect quantile 50% from the total number of species
-spp_number = plant_spread %>%  
-select(Plant_accepted_name) %>% 
-n_distinct()
-df = seq(1, spp_number)
-quantile_50 = quantile(df, 0.5)
-
-#Obtain most common plant for plotting
-most_common_plant = plant_spread %>% filter(n_networks_shared == max(n_networks_shared)) %>%
-select(Plant_accepted_name) %>% pull()
-
-o2 = ggplot(plant_spread, 
-aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
-geom_bar(stat = "identity", 
-           color = "#009E73",
-           lwd = 0.03, fill = "#009E73") +
-theme_minimal()+
-theme(axis.text.x = element_blank(),
-         axis.ticks.x=element_blank(),
-         axis.ticks.y = element_line(size = 0.2),
-         panel.grid.major.x = element_blank(),
-         panel.grid.minor.y = element_blank(),
-         panel.grid.major.y = element_line(size=0.2, linetype =  2)) +
-xlab("Plant species") +
-ylab("Occurrence across networks") + 
-geom_vline(xintercept = quantile_50, linetype="dashed", 
-          color = "black", size=0.35) + 
-annotate(geom="text", x=1010, y=0.60, label="50th percentile",color="black", size = 3) + 
-annotate(geom="text", x=100, y=0.4, label=most_common_plant[1]
-      ,color="black",  fontface = 'italic', size = 3)+
-annotate(geom="text", x=150, y=-0.05, label="Common spp."
-      ,color="black",  size = 3) +
-annotate(geom="text", x=1200, y=-0.05, label="\'Rare\' spp."
-      ,color="black",  size = 3) + 
-scale_y_continuous(breaks = seq(0, 0.7, by= 0.25), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
-coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),
-                ylim=c(0, 0.7), clip="off") +
-ggtitle("(f)")
-
-#Pollinator by plant species
-plant_poll_curves = readRDS("../Data/Working_files/collectors_curves_plant_poll.rds")
-plant_poll_output = readRDS("../Data/Working_files/plant_poll_spp_sampling_coverage.rds")
-
-#Create dataset
-d_pp = plant_poll_output$iNextEst$coverage_based
-#Check cols
-d_pp$col = d_pp$Method
-
-#Now create tibble with same structure
-a_poll = tibble(t = plant_poll_curves$Plant_sampled, 
-           Method = as.character(plant_poll_curves$iteration), 
-           qD = plant_poll_curves$Unique_spp_cumulative,
-           col = "Randomizations")
-
-#Bind datasets 
-d_pp1 = bind_rows(d_pp, a_poll)
-d_pp1_point = d_pp1 %>%  filter(Method == "Observed")
-d_pp1_lines = d_pp1 %>%  filter(!Method == "Observed")
-
-#Fix levels
-d_pp1_lines = d_pp1_lines %>% 
-mutate(col = case_when(Method == "Rarefaction" ~ "Interpolated",
-                       Method == "Extrapolation" ~ "Extrapolated",
-                       T ~ col))
-#Order levels
-d_pp1_lines$col = factor(d_pp1_lines$col, levels = c("Randomizations", "Interpolated", "Extrapolated"))
-#Plot
-pp1 = ggplot(d_pp1_lines, aes(x = t, y = qD, group = Method, color = col)) +
-geom_line(aes(linetype = col)) +
-scale_colour_manual(values = c("gray", "black", "red"), labels = c("Randomizations", "Interpolated", "Extrapolated")) +
-scale_linetype_manual(values = c("solid", "solid", "dashed"),  labels = c("Randomizations", "Interpolated", "Extrapolated")) + 
-geom_ribbon(aes(ymin = qD.LCL, 
-    ymax=qD.UCL, fill = col),alpha=0.2, colour = NA, show.legend = FALSE)+
-scale_fill_manual(values = c("gray", "black", "red")) +
-geom_point(data = d_pp1_point, col = "black") + 
-ylab("Pollinator species") +
-xlab("Plant species") +
-theme_bw() +
-coord_cartesian(expand = FALSE, clip = "off") +
-ggtitle("(d)") +
-scale_x_continuous(breaks = seq(0, 2800, by = 1000), limits=c(0,2800)) +
-scale_y_continuous(breaks = seq(0, 3000, by = 1000), limits=c(0,3000)) +
-theme(legend.position = "none")
-
-#Plot all
-p1 + p2 + p3 + pp1 + o1 + o2 + plot_layout(guides = "collect", nrow = 2, ncol=3) & theme(legend.title=element_blank(),
-      axis.title = element_text(size=12, face = "bold"),
-      axis.text = element_text(size=10),
-      plot.title = element_text(size=18))
-
-#a/b
-
-```
-
-\newpage
-
-## Habitat type and bioclimatic region
-
-The proportion of species from the major pollinator orders within the database differed across habitats and bioclimatic regions (**Figure 5**).
-Hymenoptera was the main taxonomic order in the majority of habitats, exceeded only by Diptera for the habitat categories of riparian vegetation and moors and heathland.
-Overall, the proportion of flower visitors from Lepidoptera and Coleoptera were low across all habitats but Coleopteran flower visitors were notably more abundant in sclerophyllous vegetation and beaches, dunes and sands habitat categories.
-Similar patterns were observed when exploring the pollinator proportions by bioclimatic region.
-Hymenopterans were abundant across all bioclimatic regions and Dipterans were particularly abundant in the Boreal, Alpine and Atlantic regions.
-Lepidopterans had low proportions across all bioclimatic regions and Coleopterans were only relevant in the Mediterranean region at European level.
-Notably, the number of studies (**Figure 5**) and sampling sites (**Figure S3**) also differed across habitats and bioclimatic regions.
-The habitats sampled by a higher number of studies in the database were `r studies_by_habitat$EuPPollNet_habitat[[1]]` (`r studies_by_habitat$Habitat_count[[1]]`), `r studies_by_habitat$EuPPollNet_habitat[[2]]` (`r studies_by_habitat$Habitat_count[[2]]`) and `r studies_by_habitat$EuPPollNet_habitat[[3]]` (`r studies_by_habitat$Habitat_count[[3]]`).
-However, the habitats that contain a higher number of sampling sites were `r networks_by_habitat$EuPPollNet_habitat[[1]]` (`r networks_by_habitat$Habitat_count[[1]]`), `r networks_by_habitat$EuPPollNet_habitat[[2]]` (`r networks_by_habitat$Habitat_count[[2]]`) and `r networks_by_habitat$EuPPollNet_habitat[[3]]` (`r networks_by_habitat$Habitat_count[[3]]`).
-The bioclimatic regions with a higher number of studies were `r str_to_title(studies_by_bioregion$Bioregion[[1]])` (`r studies_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(studies_by_bioregion$Bioregion[[2]])` (`r studies_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(studies_by_bioregion$Bioregion[[3]])` (`r studies_by_bioregion$Bioregion_count[[3]]`); and those that contain a higher number of sampling sites were `r str_to_title(networks_by_bioregion$Bioregion[[1]])` (`r networks_by_bioregion$Bioregion_count[[1]]`), `r str_to_title(networks_by_bioregion$Bioregion[[2]])` (`r networks_by_bioregion$Bioregion_count[[2]]`) and `r str_to_title(networks_by_bioregion$Bioregion[[3]])` (`r networks_by_bioregion$Bioregion_count[[3]]`).
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}.
-Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database.
-The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera.
-The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region.
-Note that a single study can contribute to more than one habitat or bioclimatic region.
-Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems.
-The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day and site.", fig.width=4.5, fig.height=8, fig.align='center'}
-
-#Filter out 2 bioregions that only have a small study with data of a single day
-data = data %>% 
-filter(!Bioregion == "Steppic") %>% 
-filter(!Bioregion == "Pannonian")
-
-#Prepare orders for filtering
-main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")
-
-#Karise is a very large study that only sampled bumblebees and
-#is providing unrealistic results for the Boreal region
-#Filter out
-#Maybe add to caption?
-data1 = data %>% 
-filter(!Study_id == "13_Karise")
-
-#Check number of studies by habitat
-habitat = data1 %>% 
-select(Study_id, EuPPollNet_habitat) %>% 
-distinct() %>%  
-group_by(EuPPollNet_habitat) %>% 
-summarise(Habitat_count = n()) 
-#Check number of studies by bioregion
-bioregion = data1 %>% 
-select(Study_id, Bioregion) %>% 
-distinct() %>%  
-group_by(Bioregion) %>% 
-summarise(Bioregion_count = n()) 
-#Check pols by habitat
-pollinator_percent = data1 %>%
-group_by(EuPPollNet_habitat, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(EuPPollNet_habitat) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Check pols by bioregion
-pollinator_percent_bio = data1 %>%
-group_by(Bioregion, Pollinator_order) %>%
-filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
-filter(Pollinator_order %in% main_orders) %>% 
-summarise(count = n()) %>%
-group_by(Bioregion) %>%
-mutate(Proportion = count/sum(count)) %>%
-select(-count) 
-#Preparing datasets for horizontal plot
-a1 = habitat %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat) %>% 
-rename(Count = Habitat_count)
-
-b1 = bioregion %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion) %>% 
-rename(Count = Bioregion_count)
-
-#Bind datasets
-c1 = bind_rows(a1,b1)
-
-#Create datasets and rename cols
-a2 = pollinator_percent_bio %>% 
-mutate(Group = "Bioregion") %>% 
-rename(Habitat = Bioregion)
-
-b2 = pollinator_percent %>% 
-mutate(Group = "Habitat") %>% 
-rename(Habitat = EuPPollNet_habitat)
-#Bind rows
-c2 = bind_rows(a2, b2)
-
-#Prepare to set ordered labels
-l2 = c2 %>% 
-filter(Pollinator_order == "Hymenoptera") %>% 
-arrange((Proportion)) %>% 
-pull(Habitat)
-#Set order of labels
-c2$Habitat = factor(c2$Habitat, levels = l2)
-
-#Set the right order of labels
-c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))
-
-p1 = ggplot(c2, 
-aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
-geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
-theme_bw() +
-theme(legend.position = "bottom",
-      plot.margin=unit(c(0,0,0,0), "mm"),
-      legend.text = element_text(size=6),
-      legend.key.size = unit(2, "mm"),
-      legend.spacing.y = unit(0.1, "cm"),
-       legend.box.spacing = unit(0, "pt"),
-      axis.title = element_text(face = "bold", size = 7),
-      axis.text = element_text(size=6),
-      panel.border = element_rect(size=1)) +
-xlab(NULL) +
-ylab("Proportion of species") +
-scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
-theme(strip.text.x = element_text(size = 10, face = "bold"),
-    strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid")) +
-scale_fill_manual(values = viridis(4),
-                  name = NULL) +
-coord_flip(expand = FALSE) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
-guides(fill=guide_legend(nrow=2,byrow=TRUE))
-
-#Unify labels
-c1$Habitat = factor(c1$Habitat, levels = l2)
-
-#Create barplot on the side
-p2 =ggplot(c1, aes(Habitat, Count)) +
-geom_col(fill = "gray25",width = 0.9, color="black", linewidth = 0.4) +
-theme_light() +
-theme(plot.margin=unit(c(0,0,0,0), "mm"),
-      axis.ticks.y = element_blank(), 
-      axis.ticks.x = element_line(color="black"),
-  axis.text.y = element_blank(), 
-      strip.text.x = element_text(size=12),
-      strip.background = element_rect(
-color=NA, fill=NA, size=0.5, linetype="solid"),
-  axis.title = element_text(face = "bold", size = 7),
-axis.text = element_text(size=6)) +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background = element_blank(),
-panel.border = element_blank(),
-axis.line = element_line(color = 'black'),
-plot.margin = margin(0,0,0,0)) +
-scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
-xlab(NULL) +
-ylab("Number of studies")+
-theme(legend.position = "none")+
-coord_flip(expand = F) +
-ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")
-
-#Plot panel
-p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
-
-```
-
-\newpage
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
-
-#Outputs calculated in script 7_3
-#Load lm output connectance ~ latitude
-connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
-connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
-
-connectance_lat_pval = ifelse(connectance_lat_pval <0.01, 0.01, connectance_lat_pval)
-
-#Load lm output nestedness ~ latitude
-normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
-normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
-
-#Load values of connectance
-min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
-max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
-mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
-#Load values of nestedness
-min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
-max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
-mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
-#Load kendall corr from connectance and spp
-connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
-connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
-#Load kendall corr from nestedness and spp
-normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
-normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
-#Load z values
-#VAZNULL
-z_percent_vaznull = readRDS("../Data/Manuscript_info/z_percent_vaznull.rds")
-no_diff_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_vaznull = z_percent_vaznull %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-#CURVEBALL
-z_percent_curveball = readRDS("../Data/Manuscript_info/z_percent_curveball.rds")
-no_diff_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_curveball = z_percent_curveball %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-
-#QUASISWAP
-z_percent_quasiswap = readRDS("../Data/Manuscript_info/z_percent_quasiswap.rds")
-
-no_diff_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_quasiswap = z_percent_quasiswap %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-#rd2
-z_percent_rd2 = readRDS("../Data/Manuscript_info/z_percent_rd2.rds")
-
-no_diff_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "No statistical difference") %>% 
-pull(percent)
-over_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "Over-represented") %>% 
-pull(percent)
-under_z_percent_rd2 = z_percent_rd2 %>% 
-filter(infra_over_represented == "Under-represented") %>% 
-pull(percent)
-
-#Load correlation outputs of residual connectance and nestednes
-resid_connectance_nestedness_corr_tau = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_tau.rds")
-resid_connectance_nestedness_corr_pval = readRDS("../Data/Manuscript_info/resid_connectance_nestedness_pval.rds")
-
-#To avoid having values under 0.01
-resid_connectance_nestedness_corr_pval = ifelse(resid_connectance_nestedness_corr_pval < 0.01, 0.01, resid_connectance_nestedness_corr_pval)
-
-```
-
-## Network properties
-
-Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**).
-Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and, as expected, were not independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**).
-Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks.
-(connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ < `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ ~ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**).
-Overall, networks towards higher latitudes showed lower residual connectance but similar nestedness compared to networks located at lower latitudes.
-Note that residual connectance and normalised nestedness showed a moderate significant negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ < `r round(resid_connectance_nestedness_corr_pval, 2)`).
-Empirical networks did not show statistically different nestedness (NODF) to the simulated ones (**Figure 6b** and **Figure S5**).
-The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested.
-The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested.
-Note that in both cases, NODF is calculated on binarised matrices.
-
-```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}.
-Graph (a) shows the association between network connectance and the geometric mean of plant and pollinator species per network on a log-scale with the respective fitted line from a Beta regression.
-Graph (b) shows the distribution of z-scores when comparing the nestedness from the empirical networks with their randomised counterparts (100 null models for each network with the curveball algorithm).
-The vertical red dashed lines represent the z critical value for a two tailed test with alpha = 0.05.
-Z-scores to the left of the first vertical red dashed line indicate that networks are less nested than expected by chance (red), those between the two dashed lines indicate no statistical difference from random expectations (green), and those to the right indicate that networks are more nested than expected by chance (blue).
-Graphs (c-d) show the fitted regression of residual connectance and nestedness across the latitudinal range of the studies from the database.
-The solid fitted line indicates a significant association, while the dashed lines indicate a non-significant association.
-The bioclimatic region of each network is indicated with points of different shapes and colours.", fig.width=10, fig.height=7, fig.align='center'}
-
-# Load libraries
-library(dplyr)
-library(ggplot2)
-library(stringr)
-library(patchwork)
-
-# Load plots generated in scripts 7_2_Nestedness and 7_2_Connectance
-
-connectance_species = readRDS("../Data/Manuscript_info/Figure6_connectance_species.rds")
-connectance_latitude = readRDS("../Data/Manuscript_info/Figure6_connectance_latitude.rds")
-nestedness_z_scores = readRDS("../Data/Manuscript_info/nestedness_z_scores.rds")
-nestedness_latitude = readRDS("../Data/Manuscript_info/nestedness_latitude.rds")
-
-#Prepare first connectance plots
-library(betareg)
-model = betareg(Connectance ~ log_geometric_mean_spp, data = connectance_species)
-p1 = ggplot(connectance_species, aes(x = log_geometric_mean_spp, y = Connectance))+
-geom_point(pch = 21, fill = "azure3", size=2.5) +
-geom_line(aes(y = predict(model, connectance_species)), colour= "black")+
-xlab("log(Species)") +
-theme_bw() +
-ggtitle("(a)")
-
-p3 = connectance_latitude %>% 
-ggplot(aes(Latitude, residual_conectance)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = connectance_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
-geom_ribbon(data = connectance_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ylab("Residual connectance") +
-ggtitle("(c)")
-
-null_metrics_networks = readRDS("../Data/Working_files/null_metrics_networks_quasiswap.rds")
-metrics_by_network = readRDS("../Data/Working_files/metrics_by_network.rds")
-
-#Finally, find % of networks that differ statistically
-# Merge and process data
-d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
-
-d = d %>% 
-mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)
-d = d %>% filter(!is.na((z_score)))
-
-p = 0.05 #cutoff probability 95% confidence
-critical_value = qnorm(p/2) #double tail probability divide by 2
-
-d1 = d %>%
-mutate(infra_over_represented = case_when(
-z_score < -abs(critical_value) ~ "Under-represented",
-between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
-    z_score > abs(critical_value) ~ "Over-represented"
-  ))
-
-#Calculate percentages
-z_percent = d1 %>% 
-group_by(infra_over_represented) %>% 
-summarise(percent = length(infra_over_represented)/nrow(.) * 100)
-
-no_dif_z = z_percent %>% 
-filter(infra_over_represented == "No statistical difference")
-
-p2 = d1 %>% 
-  ggplot(aes(x = z_score)) +
-  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
-                 bins = 60, alpha = 0.5, color = "black", position = "stack") +
-  theme_bw() +
-  coord_cartesian(expand = FALSE) +
-  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
-  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
-  ylab("Number of networks") +
-  xlab("Z scores (NODF)")  +
- # xlim(-6, 9) +
- # ylim(0, 0.8) +
-  scale_fill_manual(name = "Observed against null",
-                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
-                    labels = c("Less nested", "No difference", "More nested"),
-                    values = c("Under-represented" = "coral2", 
-                               "No statistical difference" = "palegreen3", 
-                               "Over-represented" = "cyan3")) +
-ggtitle("(b)")+
-xlim(-6,6)
-
-# 2nd Plot
-p4 = nestedness_latitude %>% 
-ggplot(aes(Latitude, Normalised_nestedness)) +
-geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
-theme_bw() +
-ylab("NODFc") +
-scale_fill_viridis_d() +
-scale_colour_viridis_d() +
-scale_shape_manual(values = c(
-  "Alpine" = 21,  # Circle
-  "Atlantic" = 22, # Triangle
-  "Boreal" = 23,   # Square
-  "Continental" = 24,  # Diamond
-  "Mediterranean" = 25  # Cross
-  )) +
-geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE, linetype = "dashed")+
-geom_ribbon(data = nestedness_latitude, 
-    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
-    fill = "grey70", alpha= 0.3) +
-theme_bw()+
-ggtitle("(d)")
-
-# Layout plots
-#a = (p1 + p2) 
-#b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.position = "right")
-#a / b
-
-# Layout plots
-a = (p1 + p2)  & theme(legend.position = "right", legend.justification = "left")
-b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.justification = "left")
-a / b  & theme(axis.title =  element_text(size=16, face="bold"),
-               plot.title = element_text(size=22),
-               panel.border = element_rect(colour = "black",
-               fill=NA, size=1),
-               axis.text=element_text(size=12))
-
-```
-
-\newpage
-
-# 3 | DISCUSSION
-
-EuPPollNet offers the largest set of plant-pollinator studies and networks compiled to date at European level.
-The database contains `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species with over a million interaction records.
-While the overall sampling coverage of species and interactions is relatively high across the sampled sites, the taxonomic coverage of plants and the main pollinator groups at the European level is still relatively low (i.e., `r round(plant_coverage, 1)`% for flowering plants and `r round(bee_coverage, 1)`% for bee species).
-This likely reflects that most plant and pollinator species are rare and geographically restricted.
-For example, given that most of the plant-pollinator networks from the database are sampled on intensive grasslands, and habitat heterogeneity is a crucial factor in understanding pollinator diversity at European level [@kleijn2015; @hass2018; @martinez2022], adding studies on other habitat types is likely to result in a rapid increase of the coverage of plant and pollinator species and their interactions.
-Indeed, plant and pollinator species were rarely shared across multiple sites, indicating that there are few "common" species and many "rare" ones at the metaweb or continental level.
-This high number of "rare" species results in an upward slope of the species or interaction accumulation curves [@thompson2003].
-In other words, minimal sampling efforts are capturing a substantial number of species and interactions, but achieving a comprehensive inventory will require numerous sampling events within and across habitats, particularly for plant-pollinator interactions.
-
-Bees are responsible for the majority of the sampled interactions at the metaweb level.
-As not all surveys included all pollinator groups, this result may partly be influenced by the taxonomic groups sampled across studies.
-However, the relevance of bees and other pollinator orders for network topology changed across habitats and bioclimatic regions in accordance to the literature.
-For instance, plant-pollinator communities in the Mediterranean were dominated by bees, while communities in Alpine or Boreal regions were fly species rich or fly-dominated.
-These patterns are consistent with our current understanding of bee diversity, which peaks in dry or temperate areas [@orr2021; @leclercq2023]; and with the fact that colder environments (i.e., altitude and latitude wise) harbor a larger fraction of fly pollinators compared to other taxa [@elberling1999; @lefebvre2018].
-In addition, beetles were only commonly documented as floral visitors in the Mediterranean region.
-This study cannot determine whether pollination ecologists traditionally document flower-beetle interactions only in the Mediterranean, or if there are fewer flower visitations by beetles outside this region.
-Nevertheless, the high proportion of beetles as floral visitors provides further support for their potential role as pollinators in the Mediterranean [@herrera2019; @leon2022].
-The number of butterfly species and interactions were relatively low compared to the other taxa.
-While Europe contains fewer butterfly species than other regions of the world [@ollerton2017], their relevance as pollinators is likely underestimated within this database.
-This is because a large fraction of studies (~40%) did not sample butterflies, and conventional sampling methods for monitoring other insect pollinators (e.g., bees or flies) may be inadequate for sampling plant-butterfly interactions [@isaac2011].
-Honey bees were present in `r round(apis_percent_networks, 1)`% of networks and conducted on average a third of the total interactions per network.
-The proportion of honey bees in networks across Europe is higher than in natural communities (i.e., large unmanaged assemblages of plant species) across the world [~13%\; @hung2018].
-This potentially reflects the dominance of intensive grassland habitats in EuPPollNet and their widespread distribution across European landscapes [@isselstein2005], the highly generalised nature of honey bees, their native status and above all, the widespread practice of beekeeping in Europe [@steffan2000; @magrach2017; @herrera2020].
-
-Although Europe contains a much larger number of flowering plants than pollinator species (~5 to 1 ratio according to our extrapolation from checklists), the observed number of pollinator species in the database was almost double that of the plants.
-This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018].
-While animal-centered sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants.
-In addition, we found that the accumulation curve of pollinators per plant species does not saturate, which indicates low redundancy of pollinators and that many are regionally “rare”.
-Rare pollinators can be functionally important for plant species at the landscape level [@winfree2018; @simpson2022], highlighting the need to conduct further sampling events to identify these rare species across different regions and to effectively understand and protect plant-pollinator biodiversity.
-
-Consistent with @olesen2002, we found that residual connectance (i.e., the deviation from the expected connectance for a given network size) was lower at higher latitudes, while normalised nestedness increased towards higher latitudes.
-Networks at lower latitudes in Europe are exposed to higher temperatures, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019], and the overall level of pollinator generalization is known to be higher at lower latitudes [@schleuning2012].
-These factors should increase the number of possible connections that can be established between plants and pollinators for a given network size, resulting in more connected networks at lower latitudes in Europe.
-However, our results cannot be extrapolated to lower latitudes outside Europe, as tropical systems might behave differently.
-Finally, a result that can be surprising is that most empirical networks (~85%) showed a non-nested structure.
-This result contrasts with earlier work suggesting that plant-pollinator networks are typically nested [@bascompte2003] and it aligns with recent evidence showing that nestedness is not a common feature in plant-pollinator networks when evaluated against restrictive null models that conserve the observed species degree [@payrato2019\; Figure S6].
-Note that while species degree distributions are sufficient to explain the emergence of nestedness, this does not preclude nestedness from being a useful metric for comparison across networks.
-For example, plant-pollinator networks are more nested than plant-herbivore networks precisely because these two network types differ in their composition of generalist and specialist species [@thebault2010].
-
-Although this database covers a wide range of habitats across `r country_number` countries, it contains temporal and geographical biases that can impact our understanding of plant-pollinator communities [@hughes2021].
-For instance, none of the studies in this database sampled nocturnal pollinators, which can impact our view of network structure [@garcia2024], and most studies were conducted during a single flowering season, limiting our ability to evaluate temporal trends of plant-pollinator communities in the face of environmental changes [@alarcon2008; @chacoff2018].
-In addition, most plant-pollinator networks are sampled from central Europe, while Eastern Europe and the Mediterranean region are underrepresented.
-This is consistent with previous studies which also report lack of plant-pollinator data for those regions [@bennett2018; @marshall2024], highlighting that this database shows existing patterns in data availability despite the absence of a systematic search for studies.
-The lack of data for Eastern Europe, which contains vast landscapes with semi-natural grasslands experiencing rapid land use change [@sutcliffe2015], and for the Mediterranean region, which is severely impacted by climate change [@jaworski2022; @duchenne2020; @pareja2023], is particularly concerning.
-These areas are well known for their rich pollinator diversity [@milivcic2018; @reverte2023], and their under-representation is likely contributing to the low taxonomic coverage of this database at the European level.
-Although some of the most well studied countries in Europe (e.g., Belgium, The Netherlands) have already experienced land use change and biodiversity loss at the end of the 20th century [@carvalheiro2013], plant-pollinator communities in Europe and across the globe still face current and future threats from climate change [@bartomeus2011; @duchenne2020], land use change [@reidsma2006; @batary2015], and the introduction of alien species [@vila2009; @vanbergen2018].
-Therefore, continuous monitoring programs are needed in order to evaluate spatio-temporal changes of species and their interactions across different European habitats and regions.
-This will allow local and large scale analyses of the status and trends of plant-pollinator communities, effectively informing management and conservation actions.
-
-In conclusion, the EuPPollNet database enables researchers to explore spatial, taxonomic and structural properties of plant-pollinator networks within Europe.
-In contrast to previous databases, EuPPollNet provides interaction data along with sampling information that could help researchers to better control for sampling effort and completeness and to select the most suitable networks for their research questions.
-Here, we have shown how connectance and nestedness change across their latitudinal range, and that plant-pollinator networks are as nested as expected given plant and pollinator generalist levels.
-These analyses aim to highlight the variability present across Europe in the structure of plant-pollinator networks and illustrate the  opportunities available to develop and test questions about spatio-temporal network change using EuPPollNet.
-The reproducible workflow allows researchers to adapt and reuse this database, enabling the continuous addition of new networks to better evaluate the status and trends of plant-pollinator communities.
-Finally, we hope this database becomes an iterative resource that keeps growing and improving over time to better understand and conserve European biodiversity.
-
-\newpage
-
-# REFERENCES
-
-::: {#refs}
-:::
-
-# ACKNOWLEDGEMENTS
-
-We thank all the taxonomists and ecologists that has made this database possible by contributing with their fieldwork data.
-
-# FUNDING INFORMATION
-
-This research was funded by the H2020 European project Safeguard (101003476) and by the Federal State of Saxony-Anhalt (MLU-BioDivFund).
-RT was supported by the Czech Science Foundation (Project No.
-21-24186M).
-NDM, NH, YP and FM were financially supported by the ANR ARSENIC project (grant no.
-14-CE02-0012), the ANR NGB project (grant no.
-17-CE32- 011), the Region Nord-Pas-de-Calais, the CNRS, the French Ministère de l'Enseignement Supérieur et de la Recherche, the Hauts-de-France Region and the European Regional Funds.
-
-# CONFLICT OF INTEREST
-
-None.
-
-# DATA AVAILABILITY
-
-All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
-
-\newpage
+#Bee-syrphid Masterlist
+master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
+#Load butterfly masterlist
+master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
+                    header = T)
+
+```
 
@@
 **Title**: EuPPollNet: A European database of plant-pollinator networks
-
-**Authors**: `r authors`
 

### 2024-10-03 Manuscript/Title_page.qmd
journal format style

---
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
        \captionsetup[table]{labelformat=empty}
        \usepackage{lineno}
        \linenumbers
bibliography: references.bib
csl: apa_7th.csl
latex-tinytex: false
link-citations: TRUE
linkcolor: RoyalBlue
urlcolor: RoyalBlue
---

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}

library(readr)
library(dplyr)

#Read author list 
authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
#Print authors
authors = authors %>% 
summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
pull()
#Read affiliation list 
aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
aff_list = aff_list %>% 
summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
pull()

```

\vspace*{-14mm}

**Title**:

EuPPollNet: A European database of plant-pollinator networks

**Authors**:

`r authors`

**Corresponding author** = barragansljose@gmail.com\
*All authors excluding the first four and last are ordered alphabetically*

**Affiliations**:

\begingroup
\small

`r aff_list`

\small
\endgroup

**Acknowledgements**:

We thank all the taxonomists and ecologists that has made this database possible by contributing with their fieldwork data.

**Biosketch**:

Jose B. Lanuza is a former master's, PhD, and postdoctoral researcher under Dr. Ignasi Bartomeus. Their work focuses on understanding the ecological processes that drive plant-pollinator interactions at different ecological scales. This research is another piece of the puzzle that will contribute to our understanding of how interactions change with various environmental drivers. Our work has primarily involved flowering plants and pollinators, where we have addressed questions related to their coexistence, interaction probability, and ecology. We have created the first database of Iberian bees for Spain and are strong advocates for reproducible science; this work is a perfect example of it. Further examples of the ongoing work and past projects can be consulted at Ignasi Bartomeus lab web page: <https://bartomeuslab.com/projects/>.

### 2024-10-03 Manuscript/EuPPollNet_main_text.qmd
Zenodo link

@@
 **Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
-The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
+The construction of the database is fully reproducible and can be accessed at the following link: <https://doi.org/10.5281/zenodo.13885274>.
 
@@
 
-All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
+All data and code to produce of this database and manuscript are available at Zenodo (<https://doi.org/10.5281/zenodo.13885274>) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).

### 2024-10-07 Manuscript/EuPPollNet_main_text.qmd
minor submission updates

--- 2024-10-03 Manuscript/EuPPollNet_main_text.qmd
@@
 
-**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants.
-Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning.
-However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks.
-One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time.
+**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity.
+However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interactions.
+Unfortunately, we lack robust generalisable data capturing how plant-pollinator communities are structured across space and time.
 Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest.
-This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences.
-We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
-We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified.
-Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.
-
-**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries.
-In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).
+In addition, we evaluate the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks.
+We believe EuPPollNet will stimulate research to address data gaps in plant-pollinator interactions and guide future efforts in conservation planning.
+
+**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks, which belong to `r study_number` different studies distributed across `r country_number` European countries.
+Information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information are also provided.
 
@@
 **Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`.
-All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).
 
@@
 
-**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats.
-The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.
+**Software format:** The database was built with R and is stored in “.rds” and “.csv” formats. Its construction is fully reproducible and can be accessed at: <https://doi.org/10.5281/zenodo.13897566>.
 
@@
 
-All data and code to produce of this database and manuscript are available at Zenodo (LINK) and Github (<https://github.com/JoseBSL/EuPPollNet.git>).
+All data and code to produce of this database and manuscript are available at Zenodo (<https://doi.org/10.5281/zenodo.13897566>).

### 2024-10-07 Manuscript/Title_page.qmd
Zenodo link
minor submission updates

@@
 
+Dr.
 Jose B.
-Lanuza is a former master's, PhD, and postdoctoral researcher under Dr.
+Lanuza is a long term collaborator of Dr.
 Ignasi Bartomeus.
@@
 This research is another piece of the puzzle that will contribute to our understanding of how interactions change with various environmental drivers.
-Our work has primarily involved flowering plants and pollinators, where we have addressed questions related to their coexistence, interaction probability, and ecology.
-We have created the first database of Iberian bees for Spain and are strong advocates for reproducible science; this work is a perfect example of it.
+Their work has primarily involved flowering plants and pollinators, where they have addressed questions related to their coexistence, interaction probability, and ecology.
+They have created the first database of Iberian bees for Spain and are strong advocates for reproducible science; this work is a perfect example of it.
 Further examples of the ongoing work and past projects can be consulted at Ignasi Bartomeus lab web page: <https://bartomeuslab.com/projects/>.
